### 2016-11-02 chapter1.tex
Initial Draft: Intro up through Observational Techniques

% Chapter 1: Introduction

\section{The Sun-Earth Environment}
The sun is both complex and highly dynamic.  The dynamics include a highly variable surface with strong magnetic fields energized plasma.  The surface can have sunspots, dark, relatively cool regions chracterized by intense magnetic fields, as well as coronal holes, which are low density areas that are a continuous source of plasma into the solar wind.  The intense magnetic fields associated with sunspots can break down releasing a large amount of energy and plasma, known as a solar flare.

The sun is always producing a continuous outflow of plasma known as the solar wind.  The solar wind typically has a density around 10 cm\(^{-3}\) at Earth and is traveling away from the sun at approximately 400 km/s.  However, there can be large and sudden outbursts of plasma from the sun such as solar flares or coronal mass ejections (CMEs) which can travel significanly faster.  The frequency of high energy events changes with the solar cycle, when the sun's activity changes regularly over a 22-year long cycle.  It takes 11 years for the sun to go from solar minimum, when solar activity is relatively low and there are few large events to solar maximum, when solar activity is generally high.

In addition to the outflow of plasma, the solar wind also consisted of the sun's magnetic field extended into the outer limits of the solar system, called the interplanetary magnetic field (IMF).  Because the sun has a short rotation period (about 11 days), the IMF gets twisted into a spiral, commonly called the Parker Spiral \textbf{CITE PARKER?}.  When the IMF reaches the Earth, the magnetic field is typically oriented at a \(45^\circ\) inclination relative to the ecliptic, but this is highly variable depending on solar wind conditions.

The IMF interacts with the Earth's own magnetic field, the magnetosphere.  The magnetosphere is approximately a dipole field, but dynamic pressure from the solar wind causes the sun facing side to be compressed and the rear side to be stretched into an extended tail.  In addition, magnetic reconnection can occur between the IMF and the magnetopause such that magnetic field lines that were closed in the magnetopause reconnect to magnetic field lines from the sun and become open to the solar wind.  This tends to happen in the polar regions and allows highly energized particles from the sun to flow down the magnetic field into the Earth's ionosphere.

\section{The Earth's Ionosphere}
\label{sec:ionosphere}
The Earth's ionosphere is a region of the upper atmosphere that ranges approximately between 50-1000 km in altitude where neutral gases have been excited so that they are partially ionoized, resulting in free electrons and ions.  Plasma is either partially or fully ionized gas.  When gases are heated, electrons are energized and removed from neutral particles, resulting in free ions and electrons.  Gases can either be partially ionized, in which only some of the molecules are ionized so charged particles are mixed with neutrals, or fully ionized such that all particles are ionized.

Two conflicting processes occure in the ionosphere to create a peak in electron density, referred to as the Chapman Layer \textbf{CTIE CHAPMAN?}.  Solar illumination and particle precipitation ionize neutrals, and the density of ions increase as altitude decreases because there are more neturals available to ionize.  However, below a certain point, the concentration of ions becomes large enough that recombination into neutral particles becomes a substantal factor, reducing the ion density.  The exact altitude where this density peak occurs is variable with time of day, location, season, and solar cycle.

\subsection{Ionosphere Regions}
\label{sec:ionosphere_regions}
The ionosphere has historically been divided into three distinct region.  The D region is located between 80 and 90 km in altitude and typically only exists in the daytime when the ionosphere is sunlit.  The E region typically exists between 90 and 120 km with a density peak between 105 and 110 km, depending on factors such as time of day and season.  Above 120 km is considered the F region, which has a density peak between 300 and 400 km.  The F region can actually have two peaks under some daytime conditions, notated as the F1 peak and the F2 peak.

Because the concentrations of ions and neutral particles changes with altitude, each of these regions have distinct plasma physical properties that make them unique and change how plasma structuring forms.  In the E region, electrons are magnetized so their motion is mostly due to gyration around the magnetic field.  However, the motion on ions is dominated by collisions between ions and neutrals, so ions in the E region are considered collisional.  In the F region where the neutral density is much lower, both ions and electrons are magnetized.  These differences vary the conductivity in both regions and can strongly impact how plasma waves develop, which will be discussed further in section \ref{sec:lit_experiment}.

\subsection{Plasma Structuring in the Polar Ionosphere}
\label{sec:polar_structure}
The polar ionosphere is a particularly interesting and dynamic region due to the presence of the magnetic poles and open field lines to the solar wind.  Plasma structuring can occur as variations in density, velocity, or temperature, but the main focus of this work will be density perturbations, often called irregularities.  Density irregularities in the polar cap can occur on scales ranging from thousands of kilometers to less than a centimeter.  Large-scale plasma irregularities are typically considered to be any density structuring on the scale of 1 -- 1000 km.  Some common examples are polar patches, polar holes, and sun-aligned arcs.  Polar patches are large density enhancements (usually at least twice the background plasma density) that travel across the polar cap with the background convection.  Polar holes are plasma density depletions that typically occur in the F region slightly poleward of the auroral oval.  Sun-aligned arcs are large density forms that stretch across the polar cap along the sun-earth line.  They are very narrow and often characterized by large and complex velocity shears on either side of the arc.

Intermediate-scale structuring typically ranges from 100 m -- 1 km.  This is also referred to as scintillation-causing structuring and is responsible for many of the negative space weather effects that are observed on Earth.  Radio signal scintillation through the ionosphere changes the phase and amplitude of the original signal, which can effect the ability that satelites have to communicate with ground recievers.  This can cause communication blackouts or introduce large errors into GPS calculations, effecting navigation.  Because communication and navigation technology is implemented in both every day civilian life as well as government and military intrests, the effects of this can be far reaching.

Small-scale structuring is generally considered to be all irregularities less than 100 m.  Although the effects of these scale sizes on navigation and communication systems are minimal, they can be the easiest to study because of large data sets collected from HF radars globally for several decades.  The irregularity sizes presented here are rough categories only and a high degree of coupling between different scales is generally considered reasonable.  For instance, large-scale structures such as polar patches may very well have internal intermediate- or small-scale structuring.  Some kind of turbulent cascade is often assumed to connect these different scales, but because various instruments or techniques usually only observe structuring on a particular scale, it is difficult to find direct evidence of this or how exactly it occurs.

\section{Observational Techniques and Models}
\section{SuperDARN}

%Figures
%- Map of radars (Northern and southern hemisphere)
%- Example convection map
%- Example ISR profile
%- Example MSISE profile?
%- Example of IGRF?
%- How does radar work diagram
%- Pulse sequence diagram

\section{CSR}
%-general introduction

Coherent Scatter Radars (CSR) have been used to study the Ionosphere for the last half century.  Radars are ground based instruments used to detect density structures in the ionosphere's plasma.  A basic radar consists of transmission and recieveing antenna.  The radar transmits a radio wave, which can reflect off a suititbly sized structure in the ionosphere.  Reflected waves can be detected by the recieving antennas.  The time difference between when the signal was transmitted and when it was detected  an then be used to determine how far away from the radar the target is.

The first instance of CSR being used to study plasma density structures in the ionosphere was the Scandanavian Twin Auroral Radar Experiment (STARE) in the 1970s and 1980s.  STARE consisted of two very-high frequency (VHF) radars with overlapping FoVs that were designed to measure FAIs in the E region of the ionosphere \citep{Greenwald1997}.  The advantage of two radars with overlaping FoVs was the ability to observe the same structures from two different orientations.  Because each radar can only measure the LoS velocity of a structure, two simultanious observations from different directions allows two different velocity vecotor compontents to be found, and hense the total velocity vector can be calculated.  This technique is still commonly used with CSR system.

There are three measurements typically made by CSR systems: backscatter power, line of sight (LoS) velocity, and spectral width.  Backscatter power is the power of the retuned signal that the radar's recievers detect.  Typically, a threshold is selected for the minimum power acceptable for a return to be considered an actual signal instead of background noise.  LoS velocity is the component of the total velocity that is measured along one beam of a radar.  Because radars measure velocity through the doppler shift of backscatter, only the component of the velocity vector that is along the radar beam can be measured.  Spectral width represents the width of the doppler power spectrum.
** EQUATIONS FOR LOS, WIDTH, ECT. **

%- Greenwald, 1985

CSR systems typically operate by transmitting pulses instead of continuously.  This allows the distance between the radar and the target to be found (based on the time difference between when the pulse is transmitted and recieved), but also limits the temporal resolution that the radar is capable of achieving.

%% INSERT RADAR PULSE DIAGRAM %%

The radar starts transmitting the pulse at \(t=-\Delta t\) and finishes at \(t=0\).  At \(t=?\), the pulse reaches the backscatter volume and begins reflecting off it.  At \(t=T-\Delta t\), the radar begins to received the the echoed signal, which finishs at \(t=T\).  The speed of the radar pulse in the Earth's atmosphere is approximated by the vacuum speed of light, \(c\).  The distance between the radar and the scattering volume is known as the range and is given in equation \ref{eqn:range}.
\begin{equation}
	\label{eqn:range}
	R = \frac{cT}{2}
\end{equation}
The size of the range gate is determined by the length of a radar pulse.  This is given in equation \ref{eqn:gatesize}.
\begin{equation}
	\label{eqn:gatesize}
	\Delta r = \frac{c\Delta t}{2}
\end{equation}
Longer pulses result in larger range gates, resulting in poorer spatial resolution.

When the radar transmits continuously, backscatter will be received, but because it is impossible to know when that signal was transmitted, the difference in time between transmission and receiving cannot be used to calculate how far away the backscatter volume is.  Conversly, if the radar only transmits a pulse of a certain length of a certain length and then "listens" for its return, the time between transmission and return is known and the distance the pulse traveled can be calculated and therefore the location of the backscattering volume.  However, another pulse cannot be transmitted until the first pulse returns, limiting the temporal resolution that can be achieved.  There are methods that can be used to improve this by allowing more than one pulse at a time to be sent and received, but the pulse method does inherently require a limited temporal resolution.  Increasing the length of a pulse can improve the temporal resolution, but decreases the spatial resolution.

More importantly than improving temporal resolution of data availabile, a short time between pulse returns is nessisary to allow dopler velocities to be calculated, particularly at long ranges.  The maximum dopler shifted frequency that can be observed by a radar is the Nyquist frequency, which is equivilent to half the sampling frquency.  **Explain where the Nyquist frequency comes from?**  Radars operating in single-pulse mode have a very limited sampiling frequency, as described above, particularly at long ranges where the pulse must travel a long distance.  This severely limits the dopler shift that can be measured, which places an upper limit on the LoS dopler velocity that can be calculated with this technique.  Using the frequency range of HF radars and the maximum range in the F-region ionosphere where backscatter would ideally be measured, single pulse mode results in a very low maximum doppler velocity, much lower than typical ionospheric flows.  **Example: Maximum doppler velocity that can be measured using single pulse at 2000 km range.**

%BENIFITS OF MULTIPULSE: Ponomarenko and Waters, 2006
Instead of waiting for each individual signal to be recieved before transmitting the next, HF radars typically employ a multipulse mode. In a multipulse mode, the radar transmits a series of pules with different timing between them.  
%AUTO CORRELATION FUNCTION

%- How to measure power, spectral width, LoS velocity
%	- What are these things?
%- Hanuise, 1993b
%- Baker, 1995

\subsection{SuperDARN}
- Overview
	- geography
	- number of radars
	- mid, auroral, high laditudes
	- history
		- STARE (Greenwald, 1978) - first use of CSR to study ionosphere, VHF radar
		- Goose Bay, built in 1983  (Greenwald, 1985) - first HF radar
	- measures power, spectral width, LoS velocity
	- decameter scale irregularities
	- 2 or more radars covering the same area can find the total velocity comonent
	- Network used to create convection maps
- Operational Specifications
	- Hardware - antennas?
	- Frequency
	- Modes
	- 8 pulse ACF (post 2011); 7 pulse before
		- 7 pulse - Greenwald 1983,1985; Farley 1972
		- 8 pulse - 2003 SDarn?
- How are convection maps produced?
	- Ruohoniemi et al. 1989
	- merge LoS measurments form 2+ radars with overlapping FoV (old method)
	- current method involves finding the "best fit" 2D ionospheric electrostatic potential
		- algorithm for this - Ruohoniemi and Baker, 1998
		- Legendre functions
		- Only use F-region measurements
		- Velocity data is supplimented with a statistical model when data point coverage from SuperDARN alone is insufficient

\section{Incoherent Scatter Radar}
Similar to CSRs, incoherent scatter radars (ISRs) are ground based radars that are used to probe the ionosphere.  However, they use a fundamentally different technique to observe structuring, which allows them to measure different things than CSRs.  Where as CSRs receive backscatter from large, "coherent" density structures in the ionosphere, ISRs operate by radar beams scattering due to the random thermal motion of electrons in the ionosphere.  The radar receivers then detect a spectra of frequencies with different power.  Analysis of this spectra will reveal the plasma frequency, and hence, the electron density, electron and ion temperature, and LoS velocity.
**CITE Gordon, 1958**

\section{Models}
Models can be useful tools for establishing the context of experimental measurments, especially in a global sense.  In particular, it is impractical to achieve full global coverage of all instruments, so models can be a useful way to fill in the gaps.  There are two main categories of models.  Theoretical models are based on first principles and global pictures are often found by running large computer simulations based on these first principles.  They can help give insight as to what causes specific phenomena because the process can be followed beginning to end.  Empirical models are based purely on measured data.  These can potentially be more accurate because they should replicate real measurments, but they also do not give any insight into how the system got to its ending state.  In addition, empirial models are often heavily averaged.
-empirical vs theory based models

\subsection{IRI}
The international reference ionosphere (IRI) is an empirical model of the Earth's ionosphere.  It was originally created in 1969 as a joint effort between the Comittee on Space Research (COSPAR) and the International Union of Radio Science (URSI) and has been periodically updated since then \citep{Rawer1978}.  The current version is IRI-2012 \citep{Bilitza2011}, which is used throughout the work done here.  It can be accessed either as an online module or by downloading and running the FORTRAN source code, both of which are available from the International Reference Ionosphere page on the NASA website (http://iri.gsfc.nasa.gov/).

The IRI model can produce outputs of electron density, temperatures of electrons, ions, and neutrals, ion composition (O+, H+, He+, O2+, NO+, N+), and total electron content (TEC).  The model requires the input of a location (latitude, longitude, and altitude) and time (year, date, and time).  Internally, the model has the additional capability of producing profiles in any of the input parameters.  This is particularly significant for height profiles, and allows the peak heights and densities of the E and F region to be calculated.

%- Empirical model of the global ionosphere
%- Parameters that can be calculated
%- How is it available?
%	- Online
%	- FORTRAN source code
%- Biltza, 1990

\subsection{MSISE}
The MSISE (Mass Spectrometer-Incoherent Scatter, Extended) model is an empirical model of the Earth's upper atmosphere.  It outputs overal neutral temperature and densities of particular species in the atmosphere, particularly O, N2, O2, He, Ar, H, and N.  Similar to the IRI model, input parameters include a location (latitude, longitude, height) and time (year, date, time), with the opportunity to produce profiles in any of those parameters.  The model was originally developed in 1986 to describe the neutral temperatures and densities above 100 km \citep{Batten1987}.  Currently, the 1990 rendition of the model is used (MSISE-90), which extends the lower altitudinal extent of the model and improves the predictions at higher altitudes \cite{Hedin1991}.
%- Empirical model of neutral atmosphere
%- parameters that can be calculated

\subsection{IGRF}
The international geomagnetic reference field (IGRF) is a model of the earth's magnetic field.  The model consists of mathematical expressions and corresponing coefficients that are updated regularly that represent both the Earth's magnetic field and how it changes over time \citep{Thebault2015}.  Because the Earth's magnetic field is in fact dynamic (the location of the geomagnetic poles can change by \(\sim\)10 km per year), the model constently needs to be updated not only to improve it but to keep it accurate with the present magnetic field.  It is the internationally agreed upon standard for the magnetic field strength and direction.
%- Earth's magnetic field

### 2016-11-08 chapter1.tex
Outline of derrivation for Theory and Simulations

@@
 \section{Observational Techniques and Models}
-\section{SuperDARN}
 
@@
 
-\section{CSR}
+\subsection{Coherent Scatter Radars}
 %-general introduction
@@
 Spectral width represents the width of the doppler power spectrum.
-** EQUATIONS FOR LOS, WIDTH, ECT. **
 
@@
 The speed of the radar pulse in the Earth's atmosphere is approximated by the vacuum speed of light, \(c\).
-The distance between the radar and the scattering volume is known as the range and is given in equation \ref{eqn:range}.
+The distance between the radar and the scattering volume is known as the range and is given in equation \ref{eqn:radar_range}.
 \begin{equation}
-	\label{eqn:range}
+	\label{eqn:radar_range}
 	R = \frac{cT}{2}
 \end{equation}
-The size of the range gate is determined by the length of a radar pulse.  This is given in equation \ref{eqn:gatesize}.
+The size of the range gate is determined by the length of a radar pulse.  This is given in equation \ref{eqn:range_gate}.
 \begin{equation}
-	\label{eqn:gatesize}
+	\label{eqn:range_gate}
 	\Delta r = \frac{c\Delta t}{2}
@@
 
-\subsection{SuperDARN}
-- Overview
-	- geography
-	- number of radars
-	- mid, auroral, high laditudes
-	- history
-		- STARE (Greenwald, 1978) - first use of CSR to study ionosphere, VHF radar
-		- Goose Bay, built in 1983  (Greenwald, 1985) - first HF radar
-	- measures power, spectral width, LoS velocity
-	- decameter scale irregularities
-	- 2 or more radars covering the same area can find the total velocity comonent
-	- Network used to create convection maps
-- Operational Specifications
-	- Hardware - antennas?
-	- Frequency
-	- Modes
-	- 8 pulse ACF (post 2011); 7 pulse before
-		- 7 pulse - Greenwald 1983,1985; Farley 1972
-		- 8 pulse - 2003 SDarn?
-- How are convection maps produced?
-	- Ruohoniemi et al. 1989
-	- merge LoS measurments form 2+ radars with overlapping FoV (old method)
-	- current method involves finding the "best fit" 2D ionospheric electrostatic potential
-		- algorithm for this - Ruohoniemi and Baker, 1998
-		- Legendre functions
-		- Only use F-region measurements
-		- Velocity data is supplimented with a statistical model when data point coverage from SuperDARN alone is insufficient
+\subsubsection{SuperDARN}
+The Super Dual Auroral Radar Network (SuperDARN) is a global network of HF CSRs that was designed to measure small-scale plasma structuring in the ionosphere and map the global plasma convection.
+The network currently consists of 35 operational radars between the northern and southern hemispheres distributed at mid-, high-, and polar latitudes.
+Following the STARE experiment \citep{Greenwald1978}, the first HF radar was built in Goose Bay, Canada in 1983, which would become the first of the SuperDARN radars \citep{Greenwald1985}.
 
-\section{Incoherent Scatter Radar}
+It is nessisary to use HF radars to study ionsospheric structures in the polar cap because the magnetic field is close to vertical and in order for the radar beam to meet the perpendicularity condition to observe FAIs, the beam must be refracted through a dense ionosphere.
+VHF radar beams are not refracted enough in the polar cap and the beam will pass through the ionosphere without reflecting off any FAIs.
+SuperDARN radars operate at a nominal 8--20 MHz frequency, which, according to the Bragg scatter equation for backscatter, corresponds to observing decameter-scale plasma waves.
+Each radar consists of 16 independent transmit and recieving antennas.
+Beams are electronically steerable and most radars have between 16 and 24 beams in normal operation mode, each beam being \(3.25^\circ\) in azimuth.
+Each beam consists of 75--100 range gates, which are generally either 15 km or 45 km in length \citep{Chisham2007}.
+SuperDARN radars were originally designed to used a 7 pulse ACF \citep{Farley1972,Greenwald1983,Greenwald1985}, however since 2011, most radars have begun using an 8 pulse sequence.
+
+One of the original purposes of SuperDARN was to produce maps of the plasma convection patterns in the polar caps using 2D velocity vectors.
+Originally this was accomplished by considering two radars with overlapping FoVs.
+If both recieved backscatter from the same scattering volume, two LoS velocities could be found for that scattering volume, which could be combined to find a 2D velocity vector \citep{Ruohoniemi1989}.
+The modern method for creating convection maps involves taking all data recorded by all radars in a particular hemisphere and finding the best fit to a 2D ionospheric electrostatic potential using Legendre functions \citep{Ruohonomiemi1998}.
+For convection maps, only data from the F region is considered.
+If there is not enough velocity data to confine the fit sufficiently, the measured data points are supplimented with a statistical model based on solar wind IMF conditions \textbf{CITE MCWILLIAMS???}.
+
+\subsection{Incoherent Scatter Radar}
 Similar to CSRs, incoherent scatter radars (ISRs) are ground based radars that are used to probe the ionosphere.
@@
 
-\section{Models}
+The first ISR was built in Arecibo, Puerto Rico in 1958.
+Since then, at least 8 other ISRs have been deployed around the world at equitoral, mid-, and high-latitudes.
+The most recent advancement has been the development of Advanced Modular Incoherent Scatter Radars (AMISR), of which there are currently three in use, one at the Poker Flat Rocket Range, just north of Fairbanks, AK and two in Resolute Bay, Canada.
+The advantage of these new systems is that they are electronical steerable, unlike older ISRs which consisted of a large dish that had to be moved manually to change the beam direction.
+This allows beams to be transmitted in many different directions at a very high time cadence, giving the radar the capability to make measurements in multiple directions almost simultaniously.
+This is very important for creating 2D maps of ionospheric conditions and tracking how they change in time.
+
+\subsection{IRI Model}
 Models can be useful tools for establishing the context of experimental measurments, especially in a global sense.
@@
 In addition, empirial models are often heavily averaged.
--empirical vs theory based models
 
-\subsection{IRI}
 The international reference ionosphere (IRI) is an empirical model of the Earth's ionosphere.
@@
 
-\subsection{MSISE}
-The MSISE (Mass Spectrometer-Incoherent Scatter, Extended) model is an empirical model of the Earth's upper atmosphere.
-It outputs overal neutral temperature and densities of particular species in the atmosphere, particularly O, N2, O2, He, Ar, H, and N.
-Similar to the IRI model, input parameters include a location (latitude, longitude, height) and time (year, date, time), with the opportunity to produce profiles in any of those parameters.
-The model was originally developed in 1986 to describe the neutral temperatures and densities above 100 km \citep{Batten1987}.
-Currently, the 1990 rendition of the model is used (MSISE-90), which extends the lower altitudinal extent of the model and improves the predictions at higher altitudes \cite{Hedin1991}.
-%- Empirical model of neutral atmosphere
-%- parameters that can be calculated
+\section{Theory and Observations}
+Much work has been done over the past several decades to try to understand the complicated picture of plasma structuring in the polar cap, both from theoretical analysis of plasma dispersion relations, simulations of plasma dynamics, and observations of plasma structuring using a variety of instruments.
 
-\subsection{IGRF}
-The international geomagnetic reference field (IGRF) is a model of the earth's magnetic field.
-The model consists of mathematical expressions and corresponing coefficients that are updated regularly that represent both the Earth's magnetic field and how it changes over time \citep{Thebault2015}.
-Because the Earth's magnetic field is in fact dynamic (the location of the geomagnetic poles can change by \(\sim\)10 km per year), the model constently needs to be updated not only to improve it but to keep it accurate with the present magnetic field.
-It is the internationally agreed upon standard for the magnetic field strength and direction.
-%- Earth's magnetic field
+\subsection{Theory and Simulations}
+One of the primary structuring mechanisms thought to influence the F region ionosphere is the gradient drift instability (GDI) \citep{Tsunoda1988}.
+GDI is operational when high-density purturbations in the plasma move to regions of lower background density and low-density purturbations in the plasma move to regions of higher background density, such that the wave amplitude grows relative to background conditions.
+As plasma drifts and carries irregularities with it, the ions in high density perturbations lag behind the electrons and create a purturbed electric field.
+This purturbed electric field combined with the background magnetic field causes the density perturbations to \(\vec{E}\times\vec{B}\) drift.
+If the drift causes high and low density perturbations to move as desribed above, the instability is operational and the perturbations grow.
+If instead high density perturbations drift into regions of higher background density and low density perturbations drift into regions of lower background density, the wave is damped.
+
+Some of the first 1D investigations of the linear GDI growth rate found that it can be described by a simple expression involving the plasma drift velocity, \(V_E\), and the gradient scale length, \(L\) \citep{Simon1963,Hoh1963,Linson1970}.
+\begin{equation}
+	\label{eqn:gdi_old}
+	\gamma = \frac{V_E}{L}
+\end{equation}
+This expression is only valid if the density gradient is in the same direction as the plasma drift velocity and wavevector direction is not considered.
+The results of \citet{Keskinen1982,Keskinen1983} do consider the wavevector, \(\vec{k}\), to some extent, but still assume a general relationship between density gradients and plasma drift \citep{Tsunoda1988}.
+\begin{equation}
+	\label{eqn:gdi_80s}
+	\gamma = \frac{k_y}{k}\frac{\vec{k}\cdot\vec{V}_E}{kL}
+\end{equation}
+The directional dependance of GDI has been shown to be significant, such that the growth rate can actually change signs (determining whether wave growth or damping occurs) with different wavevector directions, so it is important to consider this factor \citep{Makarevich2014c}.
+In addition, these expressions are only applicable in a particular altitudinal regieme, usually the F region.
+
+Recently, a new expression has been introduced that considers an arbetrary altitude, wavevector, and density gradient within the ionosphere, which allows FBI, GDI, an CCI to be considered with a single dispersion relation \citep{Makarevich2016a}.
+\begin{equation}
+	\label{eqn:gdi_Mak16}
+	(H_i-H_e)\omega = (H_i\vec{V}_{e0}-H_e\vec{V}_{i0})\cdot\vec{k}+(C_i-C_e)H_eH_i
+\end{equation}
+As before, \(\vec{k}\) is the wavevector and \(\vec{V}_{\alpha 0}\) is the velocity of species \(\alpha\).
+The species \(\alpha\) is given by either \(i\) for ions or \(e\) for electrons.
+The other quantities in equation \ref{eqn:gdi_Mak16} are not standard and described below.
+\begin{equation}
+	H_\alpha = S_\alpha F_\alpha + D_\alpha^{-1} F_\parallel \quad\quad\quad
+	C_\alpha = \frac{T_\alpha}{m_\alpha \Omega_\alpha}
+\end{equation}
+\begin{equation}
+	F_\alpha = i k_\perp^2 D_\alpha + \vec{G}\cdot\vec{k}_\perp D_\alpha + \vec{G}\cdot\vec{k}\times\uvec{b} \quad\quad\quad
+	F_\parallel = i k_\parallel^2 + \vec{G}\cdot\vec{k}_\parallel
+\end{equation}
+\begin{equation}
+	S_\alpha = \frac{1}{1+D_\alpha^2} \quad\quad\quad
+	D_\alpha = -\frac{i}{\Omega_\alpha}(\omega-\vec{k}\cdot{V}_{\alpha 0})+r_\alpha
+\end{equation}
+By considering certain limiting cases, an expression that is more relevant to the polar F region specifically can be resolved.
+\textbf{After SUBSTANTIAL algebraic manipulation...}  This results in the following expression for the GDI growth rate, which agrees well with the results of \citet{Makarevitch2014c}.
+\begin{equation}
+	\label{eqn:gdi}
+	\gamma = \frac{1}{1+\psi}\left(\uvec{k}\cdot\uvec{b}\times\vec{G}\right)\left(\uvec{k}\cdot\frac{\vec{E}}{B}-R\uvec{k}\cdot\frac{\vec{E}}{B}\times\uvec{b}\right)
+\end{equation}
+
+-How polar patches form
+-GDI (how small scale structuring forms)
+-Other instability mechanisms and why they aren't considered
+
+\subsection{Observations}
+-Patch formation here?
+-Dahlgren 2012 a, b
+-Semeter 2009
+-Lyne 2015 (rockets)
+-Moen 2012
+
+\section{Motivation and Objectives}

### 2016-11-14 chapter1.tex
Polar patch formation included in intro

@@
 \subsection{Theory and Simulations}
-One of the primary structuring mechanisms thought to influence the F region ionosphere is the gradient drift instability (GDI) \citep{Tsunoda1988}.
+
+Plasma structuring occurs in the ionosphere when conditions are such that wave growth occurs.
+Because different background conditions change factors such as the density of both neutrals and ions, the magnetic field strength, and so forth, the collison frequency, gyroradii, and ion inertial effects change at different altitudes.
+This causes a variety of different instability mechanisms to be operational depending on the background conditions.
+The Farley-Buneman instability (FBI), or the modified two-stream instability tends to be a major factor in the E region where ion inertial effects are high \citep{Farley1963,Buneman1963}.
+Also operational in the E region but more dominant at higher altitudes is the gradient-drift instability (GDI) \citep{Simon1963,Hoh1963,Linson1970}.
+When GDI is stable, wave growth can still occur sometimes along the magnetic field, which is known as the current-conductive instability (CCI).
+Additionally, there are shear-driven processes, known as the Kelvin-Helmholtz instability (KHI) \textbf{CITE???}.
+
+In the polar F-region ionosphere, GDI is typically considered the dominant structuring process \citep{Weber1984,Cerisier1985,Basu1988,Tsunoda1988}.
 GDI is operational when high-density purturbations in the plasma move to regions of lower background density and low-density purturbations in the plasma move to regions of higher background density, such that the wave amplitude grows relative to background conditions.
@@
 
-Recently, a new expression has been introduced that considers an arbetrary altitude, wavevector, and density gradient within the ionosphere, which allows FBI, GDI, an CCI to be considered with a single dispersion relation \citep{Makarevich2016a}.
+Recently, a new expression has been introduced that considers an arbitrary altitude, wavevector, and density gradient within the ionosphere, which allows FBI, GDI, an CCI to be considered with a single dispersion relation \citep{Makarevich2016a}.
 \begin{equation}
@@
 
--How polar patches form
--GDI (how small scale structuring forms)
--Other instability mechanisms and why they aren't considered
+In addition to analytical analysis of dispersion relations, GDI has also been studied through numerical simulations.
+\citet{Keskinen1982} found plasma waves on the trailing edge of a large density enhancement are unstable.
+\citet{Keskinen1990} expanded these results by considering coupling between the magnetosphere and ionosphere, but found that this coupling only has a substantial effect for very large-scale structuring.
+In addition, this simulation predicted instability evolution that was far faster than what observations had shown.
+3D simulations of GDI surrounding a plasma density enhancement were introduced in \citet{Guzdar1998}.
+\citet{Gondarenko1999} improved this 3D simulation further by including plasma dynamics parallel to the magnetic field and inertial effects.
+These factors are nessesary for the simulation to model secondary KHI and tertiary sheer-driven instability processes.
+These simulations provided insight in how exactly GDI structuring develops.
+Both the density and velocity spectra were found to be anisotropic, however increasing the impact of ion-inertial factors in simulations caused both spectra to become more isotropic \citep{Gondarenko2001}.
+Fluctuations in density were found that were as much at 10--20\% of the background plasma density \citep{Gondarenko2004a}.
+Although asymmetry between structuring on the leading and trailing edges of a large-scale density structure was observed in all simulations, rapidly changing convection velocity could complicate this \citep{Gondarenko2004b}.
+For instance, if a large scale structure was drifting one direction and convection patters change suddenly so that it is now drifting in the opposite direction, the leading edge becomes the trailing edge and vice versa and both edges may exhibit some structuring.
+Additionally, both edges may become structured if large velocity shears are initially present surrounding the density structure \citep{Gondarenko2006}.
+Large shears additionally cause KHI to be operational as a primary structuring mechanism, but over time GDI will still be the dominant structuring process.
 
 \subsection{Observations}
+
+Polar patches are large regions of enhanced plasma density in the polar F region \citep[][e.g.]{Weber1984,Buchau1985}.
+They are know to emerge on the dayside and then propagate with the background plasma convection across the polar cap to the nightside, where they either disintegrate or recombine with enhanced-density plasma in the auroral oval \textbf{CITE SOMETHING}.
+
+There have been at least 6 mechanisms proposed for how polar patches form on the dayside.
+\begin{enumerate}
+	\item Discrete changes in solar wind parameters, usch as IMF, density, speed, and pressure
+	\item Average flow patterns varying in time \citep{Anderson1988}
+	\item Transient magnetopause reconnection \citep{Lockwook1992}
+	\item Plasma production by cusp particle precipitation \citep{Roger1994,Millward1999}
+	\item Plasma flow jet channels that cut continuous streaches of plasma into segments 	\citep{Valladares1998}
+	\item Alfven wave coupling \citep{Prikryl1999}
+\end{enumerate}
+Although particle precipitation has been shown to create enhanced plasma density in the cusp region \citep{Roger1994}, it is unreasonable for this mechanism to account for the largest density enhancements observed in polar patches.
+These high densities must originate from the resivor of solar illuminated plasma on the dayside.
+Additionally, a mechanism is still required to separate patches from the cusp region.
+High speed flow channels can "cut" segments off of a tongue of high-density plasma that has been pulled into the polar cap by convection \citep{Valladares1994,Valladares1998}.
+Models have also shown changing large scale convection patterns quickly in time does produce density structures similar to patches \citep{Anderson1988}.
+However, the mechanism that seems to be dominant for the majority of polar patches observed is transient magnetic reconnection \citep{Carlson2012}.
+
+In order to discuss patch formation by transient magnetic reconnection, it is important to first understand the steady-state ionospheric convection patterns in the polar cap and how they relate to the magnetosphere.
+Figure \ref{fig:magnetosphere} is a schematic of how the IMF, magnetosphere, and ionosphere are all interconnected.
+In general, magnetic field lines in the magnetosphere can either be closed, meaning they connect only the north and south pole of the earth, or open, when one of the ends of the field lines originates from the sun such that the field line is connected directly to the IMF.
+The polar cap boundary is generally defined as the border between open and closed field lines, shown by the dark line in figure \ref{fig:magnetosphere}.
+By Faraday's Law, the total EMF around the polar cap boundary is equal to the rate of change of magnetic flux, equation \ref{eqn:pc_imf}.
+In the polar cap, the magnetic field strength is relatively constant at around \(5\time 10^-5\) T, so any changes in magnetic flux must be due to changes in the polar cap area, \(A_{pc}\) due to the polar cap boundary expanding or contracting \citep{Lockwood1992a}.
+\begin{equation}
+	\label{eqn:pc_imf}
+	-E = \frac{d\Phi_B}{dt} = B\frac{dA_{pc}}{dt}
+\end{equation}
+Magnetic reconnection often occurs on both the dayside and nightside of the magnetosphere, changing the amount of open magnetic flux, so the polar cap boundary can be highly dynamic.
+
+Magnetic reconnection between the IMF and the Earth's magnetosphere occurs along the X-line on the dayside magnetopause (blue line in figure \ref{fig:magnetosphere}).
+The X-line in the magnetopuase can be mapped along the magnetic field lines to the ionosphere, where it is referred to as the merging gap (pink line in figure \ref{fig:magnetosphere}).
+During reconnection events, the merging gap moves equatorward as the open magnetic flux into the polar cap increases \citep{Lockwood1992a}.
+In addition, the potential along the X-line also gets mapped the the merging gap, assuming there is minimal drop in potential along the magnetic field lines \citep{Lockwood1992b}.
+The combination of these two effects results in flux being transferred across the boundary at a rate equal to the applied voltage  along the X-line and hence plasma flow across the merging gap \citep{Lockwood1992b}, as seen in figure \ref{fig:patch_formation}.
+As the reconnection event continues, the flow across the merging gap creates a convection pattern, which starts to pull high density plasma from the dayside into the polar cap.
+If this continues consistently, eventually it can create a "tongue" of high-density plasma stretching between the dayside merging gap and the center of the polar cap, often called the tongue of ionization.
+However, when the reconnection burst ends, the merging gap begins to relax back towards the polar cap and brings the dense plasma with it, figure \ref{fig:patch_formation}.
+Before the flow stops completely, the patch "pinches off" of the ionization tounge and becomes an isolated island of high density plasma in the polar cap.
+The merging gap returns to its original position and the entire process can repeat with the next reconnection event.
+A series of reconnection events will create a string of isolated polar patches propigating through the center of the polar cap \citep{Lockwook1992b}.
+
+-HERE need to get into convection nitty gritty
+-patch formation, ect
+-possibly move this entire portion into theoretical section?
+
 -Patch formation here?

### 2016-11-15 chapter1.tex
Observation section patches and patch formation

@@
 By Faraday's Law, the total EMF around the polar cap boundary is equal to the rate of change of magnetic flux, equation \ref{eqn:pc_imf}.
-In the polar cap, the magnetic field strength is relatively constant at around \(5\time 10^-5\) T, so any changes in magnetic flux must be due to changes in the polar cap area, \(A_{pc}\) due to the polar cap boundary expanding or contracting \citep{Lockwood1992a}.
+In the polar cap, the magnetic field strength is relatively constant at around \(5\times 10^-5\) T, so any changes in magnetic flux must be due to changes in the polar cap area, \(A_{pc}\) due to the polar cap boundary expanding or contracting \citep{Lockwood1992a}.
 \begin{equation}
@@
 
--HERE need to get into convection nitty gritty
--patch formation, ect
--possibly move this entire portion into theoretical section?
+Polar patches were first observed in the 1980s using optical and radio techniques \citep{Weber1981,Buchau1983,Weber1984}.
+\citet{Weber1981} identified sun-aligned arcs that extend  along the sun-earth line through large parts of the polar cap using all-sky imaging photometers (ASIP).
+These arcs are over 1000 km long and \(\sim100\) km wide and are produced by soft electron precipitation.
+\citet{Buchau1983} used a combinations of Digisonde and ASIP measurements to identify large luminous patches in the polar cap in addition to sun-aligned arcs.
+These patches could have densities up to \(10^6\) cm\(^{-3}\) and tended to drift antisunward.
+\citet{Weber1984} found patches of enhanced ionization between 800--1000 km during moderately disturbed geomagnetic conditions, again using a combination of ASIP images, ionosonde measurements, and data from Dynamic Explorer 2 (DE-2).
+These finings confirmed that density enhancements tended to drift antisunward at speeds of 500-1000 m/s and patches were not colocated with enhanced particle precipitation within the polar cap, indicating that this is not the source of the density enhancement.
+Over time, the trailing edge of a patch became steeper than the leading edge and irregularities developed on the trailing edge.
+
+The theoretical mechanism for patch formation via transient magnetic reconnection has been described previously, Section \ref{sec:patch_formation}.
+\citet{Carlson2004} has provided experimental evidence to support this mechanism.
+Five signatures of transient magnetic reconnection were identified and then a patch formation event was examined for these signatures using the European Incoherent Scatter Scientific Association (EISCAT) Svalbard Radar.
+All five signatures were observed as expected, strengthening the idea that transient magnetopause reconnection is responsible for patch formation.
+\citet{Carlson2006} furthered this using a observations from the EISCAT Svalbard Radar, the EISCAT VHF radar in Troms\o, Norway, an ASIP, and a meridian scanning photometer (MSP).
+They managed to identify and track a series of patches directly from the subauroral plasma reservoir and show that the boundary moves equatorward before relaxing poleward, one of the main predictions of the transient magnetopause reconnection mechanism \citep{Lockwood1992}.
 

### 2016-11-17 chapter1.tex
Summary of observations of asymmetry

@@
 \begin{enumerate}
-	\item Discrete changes in solar wind parameters, usch as IMF, density, speed, and pressure
+	\item Discrete changes in solar wind parameters, usch as IMF, density, speed, and pressure \citep{Sojka1994}
 	\item Average flow patterns varying in time \citep{Anderson1988}
@@
 
+\citet{Coley1998} examined the occurrence of polar patches in both the northern and southern hemispheres with DE-2 and found that the frequency of patch occurrence was greatest when  the cusp was slightly dayward of the terminator.
+This creates the situation of an unsunlit polar cap, so background plasma density is low and blobs of enhanced density are easily visible.
+In addition, the cusp then provides a gateway of highly ionized sunlit plasma into the dark polar cap.
+\citet{Rodger1996} also considered occurrence of patches but with a HF radar in Halley, Antarctica.
+The diurnal variation in patch occurrence was found to peak at magnetic noon and the seasonal variation in equinox months.
+The only correlation between hourly averaged solar wind parameters that was found was a bias towards patches occurring with a negative IMF \(B_z\) component.
+
 The theoretical mechanism for patch formation via transient magnetic reconnection has been described previously, Section \ref{sec:patch_formation}.
@@
 
--Patch formation here?
--Dahlgren 2012 a, b
--Semeter 2009
--Lyne 2015 (rockets)
--Moen 2012
+Recent advancements have improved the ability to image density structures in the ionosphere, particularly using multi-instrument approaches.
+\citet{Semeter2009} introduced a method by which the density pattern in a particular volume could be imaged in three dimensions using AMISR systems.
+This technique was later used by \citet{Dahlgren2012a,Dahlgren2012b} to image a polar patch.
+\citet{Dahlgren2012b} additionally used ASIPs and SuperDARN radars to investigate the polar patch and found that backscatter tended to be observed more on the trailing edge of the patch.
+This agrees well with many studies that have examined the asymmetry of small-scale plasma structuring surrounding plasma patches.
+This asymmetry was first observed by \citet{Weber1984}.
+Later, \citet{Milan2002b} found that backscatter power from HF radars tended to be greater on the trailing edges of moving structures, which was attributed to GDI being unstable on the trailing edge but stable on the leading edge.
+\citet{Koustov2012} found similar results examining sun-aligned arcs.
+HF radar backscatter power tended to be greater trailing behind a sun-aligned arc moving through the polar cap that was identified with ASIP data.
+\citet{Moen2012} made direct measurements of plasma density structuring using the ICI-2 sounding rocket.
+This study provided evidence that decameter scale plasma structuring (the same as observed by HF radars) had spawned from kilometer scale density gradients and structuring was greatest where GDI was operational.
+\citet{Hosokawa2016} also considered the asymmetry of patches using an all-sky airglow imager (ASI) and found that the gradient on the leading edge of patches was typically 2--3 times steeper than that on the trailing edge.
+In addition, they found large finger-like structures on the trailing edge of patches that were 10s--100s of kilometers in size.
+These types of structures agree with GDI predictions, in particular they match the simulations of \citet{Gondarenko2004}.
 

### 2016-11-17 chapter1.tex
Motivation and Objectives added

@@
 \subsection{Coherent Scatter Radars}
-%-general introduction
+\label{sec:csr}
 
@@
 \subsubsection{SuperDARN}
+\label{sec:superdarn}
 The Super Dual Auroral Radar Network (SuperDARN) is a global network of HF CSRs that was designed to measure small-scale plasma structuring in the ionosphere and map the global plasma convection.
@@
 \subsection{Incoherent Scatter Radar}
+\label{sec:isr}
 Similar to CSRs, incoherent scatter radars (ISRs) are ground based radars that are used to probe the ionosphere.
@@
 \subsection{IRI Model}
+\label{sec:iri}
 Models can be useful tools for establishing the context of experimental measurments, especially in a global sense.
@@
 \subsection{Theory and Simulations}
+
+Polar patches are large regions of enhanced plasma density in the polar F region \citep[][e.g.]{Weber1984,Buchau1985}.
+They are know to emerge on the dayside and then propagate with the background plasma convection across the polar cap to the nightside, where they either disintegrate or recombine with enhanced-density plasma in the auroral oval \textbf{CITE SOMETHING}.
+
+There have been at least 6 mechanisms proposed for how polar patches form on the dayside.
+\begin{enumerate}
+	\item Discrete changes in solar wind parameters, usch as IMF, density, speed, and pressure \citep{Sojka1994}
+	\item Average flow patterns varying in time \citep{Anderson1988}
+	\item Transient magnetopause reconnection \citep{Lockwook1992}
+	\item Plasma production by cusp particle precipitation \citep{Roger1994,Millward1999}
+	\item Plasma flow jet channels that cut continuous streaches of plasma into segments 	\citep{Valladares1998}
+	\item Alfven wave coupling \citep{Prikryl1999}
+\end{enumerate}
+Although particle precipitation has been shown to create enhanced plasma density in the cusp region \citep{Roger1994}, it is unreasonable for this mechanism to account for the largest density enhancements observed in polar patches.
+These high densities must originate from the resivor of solar illuminated plasma on the dayside.
+Additionally, a mechanism is still required to separate patches from the cusp region.
+High speed flow channels can "cut" segments off of a tongue of high-density plasma that has been pulled into the polar cap by convection \citep{Valladares1994,Valladares1998}.
+Models have also shown changing large scale convection patterns quickly in time does produce density structures similar to patches \citep{Anderson1988}.
+However, the mechanism that seems to be dominant for the majority of polar patches observed is transient magnetic reconnection \citep{Carlson2012}.
+
+In order to discuss patch formation by transient magnetic reconnection, it is important to first understand the steady-state ionospheric convection patterns in the polar cap and how they relate to the magnetosphere.
+Figure \ref{fig:magnetosphere} is a schematic of how the IMF, magnetosphere, and ionosphere are all interconnected.
+In general, magnetic field lines in the magnetosphere can either be closed, meaning they connect only the north and south pole of the earth, or open, when one of the ends of the field lines originates from the sun such that the field line is connected directly to the IMF.
+The polar cap boundary is generally defined as the border between open and closed field lines, shown by the dark line in figure \ref{fig:magnetosphere}.
+By Faraday's Law, the total EMF around the polar cap boundary is equal to the rate of change of magnetic flux, equation \ref{eqn:pc_imf}.
+In the polar cap, the magnetic field strength is relatively constant at around \(5\times 10^-5\) T, so any changes in magnetic flux must be due to changes in the polar cap area, \(A_{pc}\) due to the polar cap boundary expanding or contracting \citep{Lockwood1992a}.
+\begin{equation}
+	\label{eqn:pc_imf}
+	-E = \frac{d\Phi_B}{dt} = B\frac{dA_{pc}}{dt}
+\end{equation}
+Magnetic reconnection often occurs on both the dayside and nightside of the magnetosphere, changing the amount of open magnetic flux, so the polar cap boundary can be highly dynamic.
+
+Magnetic reconnection between the IMF and the Earth's magnetosphere occurs along the X-line on the dayside magnetopause (blue line in figure \ref{fig:magnetosphere}).
+The X-line in the magnetopuase can be mapped along the magnetic field lines to the ionosphere, where it is referred to as the merging gap (pink line in figure \ref{fig:magnetosphere}).
+During reconnection events, the merging gap moves equatorward as the open magnetic flux into the polar cap increases \citep{Lockwood1992a}.
+In addition, the potential along the X-line also gets mapped the the merging gap, assuming there is minimal drop in potential along the magnetic field lines \citep{Lockwood1992b}.
+The combination of these two effects results in flux being transferred across the boundary at a rate equal to the applied voltage  along the X-line and hence plasma flow across the merging gap \citep{Lockwood1992b}, as seen in figure \ref{fig:patch_formation}.
+As the reconnection event continues, the flow across the merging gap creates a convection pattern, which starts to pull high density plasma from the dayside into the polar cap.
+If this continues consistently, eventually it can create a "tongue" of high-density plasma stretching between the dayside merging gap and the center of the polar cap, often called the tongue of ionization.
+However, when the reconnection burst ends, the merging gap begins to relax back towards the polar cap and brings the dense plasma with it, figure \ref{fig:patch_formation}.
+Before the flow stops completely, the patch "pinches off" of the ionization tounge and becomes an isolated island of high density plasma in the polar cap.
+The merging gap returns to its original position and the entire process can repeat with the next reconnection event.
+A series of reconnection events will create a string of isolated polar patches propigating through the center of the polar cap \citep{Lockwook1992b}.
 
@@
 
-Polar patches are large regions of enhanced plasma density in the polar F region \citep[][e.g.]{Weber1984,Buchau1985}.
-They are know to emerge on the dayside and then propagate with the background plasma convection across the polar cap to the nightside, where they either disintegrate or recombine with enhanced-density plasma in the auroral oval \textbf{CITE SOMETHING}.
-
-There have been at least 6 mechanisms proposed for how polar patches form on the dayside.
-\begin{enumerate}
-	\item Discrete changes in solar wind parameters, usch as IMF, density, speed, and pressure \citep{Sojka1994}
-	\item Average flow patterns varying in time \citep{Anderson1988}
-	\item Transient magnetopause reconnection \citep{Lockwook1992}
-	\item Plasma production by cusp particle precipitation \citep{Roger1994,Millward1999}
-	\item Plasma flow jet channels that cut continuous streaches of plasma into segments 	\citep{Valladares1998}
-	\item Alfven wave coupling \citep{Prikryl1999}
-\end{enumerate}
-Although particle precipitation has been shown to create enhanced plasma density in the cusp region \citep{Roger1994}, it is unreasonable for this mechanism to account for the largest density enhancements observed in polar patches.
-These high densities must originate from the resivor of solar illuminated plasma on the dayside.
-Additionally, a mechanism is still required to separate patches from the cusp region.
-High speed flow channels can "cut" segments off of a tongue of high-density plasma that has been pulled into the polar cap by convection \citep{Valladares1994,Valladares1998}.
-Models have also shown changing large scale convection patterns quickly in time does produce density structures similar to patches \citep{Anderson1988}.
-However, the mechanism that seems to be dominant for the majority of polar patches observed is transient magnetic reconnection \citep{Carlson2012}.
-
-In order to discuss patch formation by transient magnetic reconnection, it is important to first understand the steady-state ionospheric convection patterns in the polar cap and how they relate to the magnetosphere.
-Figure \ref{fig:magnetosphere} is a schematic of how the IMF, magnetosphere, and ionosphere are all interconnected.
-In general, magnetic field lines in the magnetosphere can either be closed, meaning they connect only the north and south pole of the earth, or open, when one of the ends of the field lines originates from the sun such that the field line is connected directly to the IMF.
-The polar cap boundary is generally defined as the border between open and closed field lines, shown by the dark line in figure \ref{fig:magnetosphere}.
-By Faraday's Law, the total EMF around the polar cap boundary is equal to the rate of change of magnetic flux, equation \ref{eqn:pc_imf}.
-In the polar cap, the magnetic field strength is relatively constant at around \(5\times 10^-5\) T, so any changes in magnetic flux must be due to changes in the polar cap area, \(A_{pc}\) due to the polar cap boundary expanding or contracting \citep{Lockwood1992a}.
-\begin{equation}
-	\label{eqn:pc_imf}
-	-E = \frac{d\Phi_B}{dt} = B\frac{dA_{pc}}{dt}
-\end{equation}
-Magnetic reconnection often occurs on both the dayside and nightside of the magnetosphere, changing the amount of open magnetic flux, so the polar cap boundary can be highly dynamic.
-
-Magnetic reconnection between the IMF and the Earth's magnetosphere occurs along the X-line on the dayside magnetopause (blue line in figure \ref{fig:magnetosphere}).
-The X-line in the magnetopuase can be mapped along the magnetic field lines to the ionosphere, where it is referred to as the merging gap (pink line in figure \ref{fig:magnetosphere}).
-During reconnection events, the merging gap moves equatorward as the open magnetic flux into the polar cap increases \citep{Lockwood1992a}.
-In addition, the potential along the X-line also gets mapped the the merging gap, assuming there is minimal drop in potential along the magnetic field lines \citep{Lockwood1992b}.
-The combination of these two effects results in flux being transferred across the boundary at a rate equal to the applied voltage  along the X-line and hence plasma flow across the merging gap \citep{Lockwood1992b}, as seen in figure \ref{fig:patch_formation}.
-As the reconnection event continues, the flow across the merging gap creates a convection pattern, which starts to pull high density plasma from the dayside into the polar cap.
-If this continues consistently, eventually it can create a "tongue" of high-density plasma stretching between the dayside merging gap and the center of the polar cap, often called the tongue of ionization.
-However, when the reconnection burst ends, the merging gap begins to relax back towards the polar cap and brings the dense plasma with it, figure \ref{fig:patch_formation}.
-Before the flow stops completely, the patch "pinches off" of the ionization tounge and becomes an isolated island of high density plasma in the polar cap.
-The merging gap returns to its original position and the entire process can repeat with the next reconnection event.
-A series of reconnection events will create a string of isolated polar patches propigating through the center of the polar cap \citep{Lockwook1992b}.
-
 Polar patches were first observed in the 1980s using optical and radio techniques \citep{Weber1981,Buchau1983,Weber1984}.
@@
 \section{Motivation and Objectives}
+
+It is known that the polar cap ionosphere is highly structured in a non-trivial manner \citep{Carlson2012}.
+In addition to being of purely scientific interest, this structuring creates complications when sending radio waves through the ionosphere.
+Because the structuring is highly irregular and varies on many spatial and temporal scales, unpredictable wave refraction, phase shifts, and amplitude attenuation can occur, making signals difficult to detect.
+This introduces problems in any system that involves ground-to-satellite communication, such as navigation or communication.
+Because of the increasing interconnectedness of modern society, any interruptions or errors in these systems can have broad impacts on infrastructure, including commercial and defense interests, as well as everyday life.
+
+Scientifically, the ionosphere is an important part of the highly coupled Sun-Earth environment.
+The polar cap ionosphere essentially serves as the boundary conditions for waves propagating through the magnetosphere.
+In addition, understanding plasma structuring on a range of scales is important for correctly interpreting backscatter from ionospheric radars.
+In particular, one of the main purposes of SuperDARN is to create large scale convection maps of both polar caps.
+These convection maps have been used to give context of both magnetosphere and thermosphere behavior and are widely used in space physics research.
+To map convection accurately, it is extremely important to understand when backscatter is observed and when it is absent, and why this is so.
+Understanding plasma instability and wave growth mechanisms is very important for answering these kinds of questions.
+Models and simulations of structuring in the polar caps incorporate structuring mechanisms so it is helpful if these are better understood, both the operation of a particular mechanism as well as how different mechanism are interconnected and which ones may be dominant under particular conditions.
+
+The aim of this work is to examine many different factors behind plasma irregularity production in the F region polar cap using both observations and theory.
+The majority of observational data has been found from ground based ionospheric radars, both CSR systems (Section \ref{sec:csr}) and ISR systems (Section \ref{sec:isr}), with particular emphasis on radars within the SuperDARN network, Section \ref{sec:superdarn}.
+However, these datasets have also been supplemented with information from models as well as other instruments when necessary.
+In addition, most experimental results are compared with predictions made by a linear fluid theory view of ionospheric plasma behavior.
+Specific objectives are as follows:
+
+The first objective of is to evaluate the extent of solar control over irregularity production in the polar F region.
+This considers a global view of small-scale plasma structuring.
+The occurrence of radar backscatter will be compared with both solar illumination (considered direct solar control) and IMF factors (considered indirect solar control).
+This gives a macrophysical view of where and when plasma irregularities are most likely to occur.
+
+The second objective is to investigate asymmetry in the GDI growth rate surrounding large-scale density structures.
+This is a modeling approach to identifying where plasma irregularities occur surrounding a polar patch and how they can be observed by a ground-based HF radar.
+The model allows certain factors to be considered that can be challenging to investigate experimentally becsuse it is difficult to isolate particular effects.
+The model used considers the direction of a large, elongated polar patch as well as it drift direction to be arbitrary in relation to the radar and independent of each other, which gives some insight as to how structuring would be expected with polar patches in orientations that are not often observed.
+
+The third objective is to investigate the relationship between factors that control GDI and small-scale plasma structuring.
+This combines the theoretical consideration of where growth rate is greatest with where structuring is observed the most around a polar patch.
+This is particularly important because even though plasma structuring is often observed on the trailing edge of polar patches and this is where GDI predicts the greatest growth rates, GDI theory actually predicts growth of much larger irregularites than the small scale structuring observed by HF radars.
+A direct comparison of the two improves our understanding of how growth rates are related to small-scale plasma structuring.

### 2016-11-18 chapter1.tex
Derivation of gamma completed

@@
 \end{equation}
-By considering certain limiting cases, an expression that is more relevant to the polar F region specifically can be resolved.
-\textbf{After SUBSTANTIAL algebraic manipulation...}  This results in the following expression for the GDI growth rate, which agrees well with the results of \citet{Makarevitch2014c}.
+The quantity \(r_\alpha\) is the ratio of the collision frequency to the gyrofrequency of a particular species, given by \(r_\alpha = \nu_\alpha/\Omega_\alpha\).
+
+It is useful to express Equation \ref{eqn:gdi_Mak16} in the following equivalent form by defining additional terms.
+\begin{equation}
+	\label{eqn:gdi_Mak16_alt}
+	\left(D_i-D_e\right)\omega = \vec{V}_d\cdot\vec{k}\left(\frac{Z_2}{Z_1}\right) + \left(D_i\vec{V}_{i0}-D_e\vec{V}_{e0}\right)\cdot\vec{k}+(C_i-C_e)k_{\perp}^2\left(\frac{Z_3}{Z_1}\right)
+\end{equation}
+In the F region where background plasma motion is dominated by electric fields perpendicular to the magnetic field, the differential plasma velocity, \(\vec{V}_d\), can be described as 
+\begin{equation}
+	\label{eqn:gdi_Mak16_Vd}
+	\vec{V}_d = \vec{V}_{e0}-\vec{V}_{i0} = (r_i-r_e)\left[s_es_i(1+\psi)\left(R\vec{V}_E-\frac{\vec{E}}{B}\right)\right]
+\end{equation}
+where, \(s_\alpha = (1+r_\alpha^2)^{-1}\).  Additionally,
+\begin{equation}
+\label{eqn:gdi_Mak16_Z}
+\begin{split}
+	Z_1 = i(1+Y)+a+Rb+ycK \quad\quad\quad 
+	Z_2 = R(i+a)-b \quad\quad\quad\quad\quad\quad\\
+	Z_3 = -\frac{\Psi(i+a)^2}{1+\Psi}-Rb(i+a)+\frac{b^2}{1+\Psi}-iy(y-ic)\left[\left(K+1-\Psi^{-1}\right)(i+a)+Rb\left(1-\Psi^{-1}\right)\right]\\+Ky^2(y-ic)^2
+\end{split}
+\end{equation}
+and
+\begin{equation}
+\begin{split}
+	a& = \frac{\vec{G}\cdot\vec{k}_\perp}{k_\perp^2} \quad\quad\quad
+	b = -\frac{\vec{G}\cdot\vec{k}\times\uvec{b}}{k_\perp^2} \quad\quad\quad
+	c = \frac{\vec{G}\cdot\vec{k}_\parallel}{k_\perp k_\parallel} \\
+	\Psi = -D_iD_e& \quad\quad
+	R = \frac{D_i+D_e}{1+\Psi} \quad\quad
+	y = \frac{k_\parallel}{k_\perp} \quad\quad
+	K = \left(1+\frac{1}{\Psi}\right)\left(1+R^2\right) \quad\quad
+	Y = Ky^2
+\end{split}
+\end{equation}.
+
+By considering certain limiting cases, an expression that is more relevant to the polar F region specifically can be resolved from Equation \ref{eqn:gdi_Mak16_alt}.
+First, assume a cold plasma such that the temperature of both ions and electrons is negligible.
+In this case, \(T_\alpha = 0\), so \(C_\alpha = 0\).
+Secondly assume all irregularities are perfectly aligned with the magnetic field such that \(k_\parallel = 0\).
+This means that \(\vec{k}_\perp = \vec{k}\), \(y=0\), and \(Y=0\).
+Finally, consider only low frequencies (\(\omega \ll \Omega_\alpha\)) and long wavelengths (\(\vec{k}\cdot\vec{V}_{\alpha 0} \\ \nu_\alpha\)).
+If \(D_\alpha\) is expressed as shown below, it can bee seen that these limits result in the first two terms being negligible such that \(D_\alpha \approx r_\alpha\).
+\begin{equation}
+	D_\alpha = -i\frac{\omega}{\Omega_\alpha} + i\frac{\vec{k}\cdot\vec{V}_{\alpha 0}}{\Omega_\alpha} + \frac{\nu_\alpha}{\Omega_\alpha} \approx \frac{\nu_\alpha}{\Omega_\alpha} = r_\alpha
+\end{equation}
+Equation \ref{eqn:gdi_Mak16_alt} can then be expressed as follows.
+\begin{equation}
+	\label{eqn:gdi_Mak16_Freg}
+	(r_i-r_e)\omega = \vec{V}_d\cdot\vec{k}\left(\frac{Z_2}{Z_1}\right)+(r_i\vec{V}_{i0}-r_e\vec{V}_{e0})
+\end{equation}
+The growth rate, \(\gamma\), is contained within the dispersion relation as \(\omega = \omega_r+i\gamma\), so the imaginary part of Equation \ref{eqn:gdi_Mak16_Freg} must be considered separately.
+\begin{equation}
+	\label{eqn:gdi_Mak16_gam}
+	(r_i-r_e)\gamma = \vec{V}_d\cdot\vec{k}\;\text{Im}\left\{\frac{Z_2}{Z_1}\right\}
+\end{equation}
+The differential plasma drift has been defined previously in Equation \ref{eqn:gdi_Mak16_Vd} but the imaginary component of \(Z_2/Z_1\) must still be identified.
+Using the definitions of \(Z_1\) and \(Z_2\) provided in Equation \ref{eqn:gdi_Mak16_Z} and the simplifying assumptions discussed above for the F region,
+\begin{equation}
+	\frac{Z_2}{Z_1} = \frac{R(i+a)-b}{i+a+Rb}
+\end{equation}.
+Multiplying both the numerator and denominator by \(-i+a+Rb\) removes the imaginary component from the denominator.
+Additionally the local approximation \(G/k_\perp \ll 1\) is taken, which results in \(a\), \(b\), and \(c\) being small so that any second order terms in these quantities can be neglected.
+\begin{equation}
+	\frac{Z_2}{Z_1} = R+ib(1+R^2)
+\end{equation}
+Using the above result and Equation \ref{eqn:gdi_Mak16_Vd}, Equation \ref{eqn:gdi_Mak16_gam} becomes
+\begin{equation}
+	(r_i-r_e)\gamma = (r_i-r_e)s_es_i\left(1+\psi\right)\left(R\vec{V}_E-\frac{E}{B}\right)\cdot\vec{k}b\left(1+R^2\right)
+\end{equation}
+Recognizing that \(s_es_i\left(1+R^2\right) = (1+\psi)^{-2}\) and \(b = \vec{G}\cdot\vec{k}\times\uvec{b}/k^2 = -\vec{k}\cdot\uvec{b}\times\vec{G}/k^2\) by vector identities a final expression for the GDI growth rate in the F region can be found, which agrees well with the results of \citet{Makarevitch2014c}.
 \begin{equation}
 	\label{eqn:gdi}
-	\gamma = \frac{1}{1+\psi}\left(\uvec{k}\cdot\uvec{b}\times\vec{G}\right)\left(\uvec{k}\cdot\frac{\vec{E}}{B}-R\uvec{k}\cdot\frac{\vec{E}}{B}\times\uvec{b}\right)
+	\gamma = \frac{1}{1+\psi}\left(\uvec{k}\cdot\uvec{b}\times\vec{G}\right)\uvec{k}\cdot\left(\frac{\vec{E}}{B}-R\vec{V}_E\right)
 \end{equation}
@@
 This combines the theoretical consideration of where growth rate is greatest with where structuring is observed the most around a polar patch.
-This is particularly important because even though plasma structuring is often observed on the trailing edge of polar patches and this is where GDI predicts the greatest growth rates, GDI theory actually predicts growth of much larger irregularites than the small scale structuring observed by HF radars.
+This is particularly important because even though plasma structuring is often observed on the trailing edge of polar patches and this is where GDI predicts the greatest growth rates, GDI theory actually predicts growth of much larger irregularities than the small scale structuring observed by HF radars.
 A direct comparison of the two improves our understanding of how growth rates are related to small-scale plasma structuring.

### 2016-11-23 chapter1.tex
Set up script for including graphics

@@
 In addition to the outflow of plasma, the solar wind also consisted of the sun's magnetic field extended into the outer limits of the solar system, called the interplanetary magnetic field (IMF).
-Because the sun has a short rotation period (about 11 days), the IMF gets twisted into a spiral, commonly called the Parker Spiral \textbf{CITE PARKER?}.
-When the IMF reaches the Earth, the magnetic field is typically oriented at a \(45^\circ\) inclination relative to the ecliptic, but this is highly variable depending on solar wind conditions.
+Because the sun has a short rotation period (about 11 days), the IMF gets twisted into a spiral, commonly called the Parker Spiral \citep{Parker1958}
+.  When the IMF reaches the Earth, the magnetic field is typically oriented at a \(45^\circ\) inclination relative to the ecliptic, but this is highly variable depending on solar wind conditions.
 
@@
 The exact altitude where this density peak occurs is variable with time of day, location, season, and solar cycle.
+
+\includegraphics[width=\textwidth]{densprofile.pdf}
 
@@
 It was originally created in 1969 as a joint effort between the Comittee on Space Research (COSPAR) and the International Union of Radio Science (URSI) and has been periodically updated since then \citep{Rawer1978}.
-The current version is IRI-2012 \citep{Bilitza2011}, which is used throughout the work done here.
+The current version is IRI-2012 \citep{Bilitza2014}, which is used throughout the work done here.
 It can be accessed either as an online module or by downloading and running the FORTRAN source code, both of which are available from the International Reference Ionosphere page on the NASA website (http://iri.gsfc.nasa.gov/).
@@
 
-Polar patches are large regions of enhanced plasma density in the polar F region \citep[][e.g.]{Weber1984,Buchau1985}.
+Polar patches are large regions of enhanced plasma density in the polar F region \citep[][e.g.]{Weber1984,Buchau1983}.
 They are know to emerge on the dayside and then propagate with the background plasma convection across the polar cap to the nightside, where they either disintegrate or recombine with enhanced-density plasma in the auroral oval \textbf{CITE SOMETHING}.
@@
 	\item Average flow patterns varying in time \citep{Anderson1988}
-	\item Transient magnetopause reconnection \citep{Lockwook1992}
-	\item Plasma production by cusp particle precipitation \citep{Roger1994,Millward1999}
+	\item Transient magnetopause reconnection \citep{Lockwook1992b}
+	\item Plasma production by cusp particle precipitation \citep{Rodger1994,Millward1999}
 	\item Plasma flow jet channels that cut continuous streaches of plasma into segments 	\citep{Valladares1998}
@@
 \citet{Carlson2006} furthered this using a observations from the EISCAT Svalbard Radar, the EISCAT VHF radar in Troms\o, Norway, an ASIP, and a meridian scanning photometer (MSP).
-They managed to identify and track a series of patches directly from the subauroral plasma reservoir and show that the boundary moves equatorward before relaxing poleward, one of the main predictions of the transient magnetopause reconnection mechanism \citep{Lockwood1992}.
+They managed to identify and track a series of patches directly from the subauroral plasma reservoir and show that the boundary moves equatorward before relaxing poleward, one of the main predictions of the transient magnetopause reconnection mechanism \citep{Lockwood1992b}.
 
@@
 This is a modeling approach to identifying where plasma irregularities occur surrounding a polar patch and how they can be observed by a ground-based HF radar.
-The model allows certain factors to be considered that can be challenging to investigate experimentally becsuse it is difficult to isolate particular effects.
+The model allows certain factors to be considered that can be challenging to investigate experimentally because it is difficult to isolate particular effects.
 The model used considers the direction of a large, elongated polar patch as well as it drift direction to be arbitrary in relation to the radar and independent of each other, which gives some insight as to how structuring would be expected with polar patches in orientations that are not often observed.

### 2016-11-23 thesis.sh

@@ -1,0 +1,6 @@
+
+#!/bin/bash
+
+gs -q -dSAFER -dBATCH -dNOPAUSE -sDEVICE=pdfwrite -dDEVICEWIDTHPOINTS=470 -dDEVICEHEIGHTPOINTS=350 -sOutputFile=/Users/ljlamarche/Desktop/UAFThesis/Figures/densprofile.pdf /Users/ljlamarche/Desktop/UAFThesis/Figures/densprofile.ps
+
+pdflatex LLamarcheThesis.tex

### 2016-11-23 chapter1.tex
Corrections made to Solar-Terrestrial Environment Section

@@
 
-\section{The Sun-Earth Environment}
-The sun is both complex and highly dynamic.
-The dynamics include a highly variable surface with strong magnetic fields energized plasma.
-The surface can have sunspots, dark, relatively cool regions chracterized by intense magnetic fields, as well as coronal holes, which are low density areas that are a continuous source of plasma into the solar wind.
+\section{The Solar-Terrestrial Environment}
+The solar terrestrial environment begins with the Sun in the center of the solar system.
+The Sun, like most stars, is composed of highly energized plasma that is gravitationally bound together.
+This makes it an extremely complex and highly dynamic system that is constantly interacting with the rest of the solar system.
+The surface of the Sun is populated with sunspots, dark, relatively cool regions with intense magnetic fields, as well as coronal holes, which are low density areas characterized with a continuous outflow of plasma.
 The intense magnetic fields associated with sunspots can break down releasing a large amount of energy and plasma, known as a solar flare.
+Planet-sized globs of plasma know as coronal mass ejections (CMEs) can be ejected from the surface, which can be large enough to maintain an internal magnetic field as they travel through the solar system.
+The frequency of sunspots and large outbursts of plasma changes with the 22-year long solar cycle.
+A large number of Sun spots, high occurrence of fast, dense plasma outbursts, and a highly variably magnetic field indicate solar maximum.
+Conversely, solar minimum is characterized by a low Sun spot number and slow and relatively steady outflow.
 
-The sun is always producing a continuous outflow of plasma known as the solar wind.
-The solar wind typically has a density around 10 cm\(^{-3}\) at Earth and is traveling away from the sun at approximately 400 km/s.
-However, there can be large and sudden outbursts of plasma from the sun such as solar flares or coronal mass ejections (CMEs) which can travel significanly faster.
-The frequency of high energy events changes with the solar cycle, when the sun's activity changes regularly over a 22-year long cycle.
-It takes 11 years for the sun to go from solar minimum, when solar activity is relatively low and there are few large events to solar maximum, when solar activity is generally high.
-
-In addition to the outflow of plasma, the solar wind also consisted of the sun's magnetic field extended into the outer limits of the solar system, called the interplanetary magnetic field (IMF).
-Because the sun has a short rotation period (about 11 days), the IMF gets twisted into a spiral, commonly called the Parker Spiral \citep{Parker1958}
-.  When the IMF reaches the Earth, the magnetic field is typically oriented at a \(45^\circ\) inclination relative to the ecliptic, but this is highly variable depending on solar wind conditions.
+The Sun interacts with the rest of the solar system through the solar wind, a continuous outflow of plasma that caries the Sun's magnetic field with it.
+The solar wind typically has a density around 10 cm\(^{-3}\) at Earth and is traveling away from the Sun at approximately 400 km/s.
+However, like the Sun, the solar wind is highly dynamic and these parameters can change drastically over a period of minutes.
+CMEs, for instance, create a substantial increase in density and are often proceeded by a shock.
+In addition to the outflow of plasma, the solar wind also consisted of the Sun's magnetic field extended into the outer limits of the solar system, called the interplanetary magnetic field (IMF).
+Because the Sun has a short rotation period (about 33 days), the IMF gets twisted into a spiral, commonly called the Parker Spiral \citep{Parker1958}.
+When the IMF reaches the Earth, the magnetic field is typically oriented at a \(45^\circ\) inclination relative to the ecliptic, but this is highly variable depending on solar wind conditions.
 
 The IMF interacts with the Earth's own magnetic field, the magnetosphere.
-The magnetosphere is approximately a dipole field, but dynamic pressure from the solar wind causes the sun facing side to be compressed and the rear side to be stretched into an extended tail.
-In addition, magnetic reconnection can occur between the IMF and the magnetopause such that magnetic field lines that were closed in the magnetopause reconnect to magnetic field lines from the sun and become open to the solar wind.
-This tends to happen in the polar regions and allows highly energized particles from the sun to flow down the magnetic field into the Earth's ionosphere.
+Without outside influences, the magnetosphere would be approximately a dipole field, but dynamic pressure from the solar wind causes the Sun facing side to be compressed to 6--10 Earth radii and the rear side to be stretched into an extended tail (hundreds of Earth radii).
+Because the solar wind is typically a supersonic flow, a bow shock forms on the dayside of the magnetosphere.
+The heated and compressed solar wind plasma created by the bow shock is known as the magnetosheath.
+The magnetopause is the actual boundary between solar wind plasma in the magnetosheath and magnetospheric plasma.
+However, the magnetosphere is not fully closed.
+At the magnetopause, magnetic reconnection can occur between the IMF and the Earth's magnetic field creating open field lines.
+Magnetic field lines originating from the Earth are then directly connected to the Sun through the solar wind, which consequently allows highly energized solar particles to enter the magnetosphere.
+Open field lines tend to occur around the Earth's polar caps, making these regions particularly interesting to study due to very complex behavior and coupled interactions between the Earth and the Sun.
+
+Closer to the Earth's surface, the magnetosphere also interacts with the Earth's ionosphere, a layer of partially ionized plasma in the upper atmosphere.
+The interaction between charged particles in the ionosphere with magnetic field lines from the magnetosphere couples these two regions and creates non-trivial dynamics.
+In addition, highly energized charged particles from the solar wind travel along open field lines and collide with both ions and neutral particles in the ionosphere, creating the aurora at high latitudes.
+Below the ionosphere, the thermosphere consists mostly of neutral particles, however neutral winds can influence the ionosphere through viscous interactions along the lower boundary.
+All of these regions contain incredibly complex dynamics and are additionally all interconnected, creating a complicated system that is extremely challenging to describe completely.
 
@@
 \label{sec:ionosphere}
-The Earth's ionosphere is a region of the upper atmosphere that ranges approximately between 50-1000 km in altitude where neutral gases have been excited so that they are partially ionoized, resulting in free electrons and ions.
-Plasma is either partially or fully ionized gas.
-When gases are heated, electrons are energized and removed from neutral particles, resulting in free ions and electrons.
-Gases can either be partially ionized, in which only some of the molecules are ionized so charged particles are mixed with neutrals, or fully ionized such that all particles are ionized.
-
-Two conflicting processes occure in the ionosphere to create a peak in electron density, referred to as the Chapman Layer \textbf{CTIE CHAPMAN?}.
+The Earth's ionosphere is a region of the upper atmosphere that ranges approximately between 50-1000 km in altitude where neutral gases have been excited so that they are partially ionoized, resulting in free electrons and ions mixed with neutral particles.
+Two conflicting processes occur in the ionosphere to create a peak in electron density, commonly referred to as the Chapman Layer \textbf{CTIE CHAPMAN?}.
 Solar illumination and particle precipitation ionize neutrals, and the density of ions increase as altitude decreases because there are more neturals available to ionize.
@@
 
+\begin{figure}
 	\includegraphics[width=\textwidth]{densprofile.pdf}
+	\caption{Altitude profile of electron density throughout the ionosphere.  Typical ranges of the D, E, and F regions are identified.}
+	\label{fig:densprofile}
+\end{figure}
 

### 2016-11-24 chapter1.tex
Corrections made to Ionosphere Section

@@
 \label{sec:ionosphere}
-The Earth's ionosphere is a region of the upper atmosphere that ranges approximately between 50-1000 km in altitude where neutral gases have been excited so that they are partially ionoized, resulting in free electrons and ions mixed with neutral particles.
+The Earth's ionosphere is a region of the upper atmosphere that ranges approximately between 50--1000 km in altitude where neutral gases have been excited so that they are partially ionoized, resulting in free electrons and ions mixed with neutral particles.
 Two conflicting processes occur in the ionosphere to create a peak in electron density, commonly referred to as the Chapman Layer \textbf{CTIE CHAPMAN?}.
@@
 The exact altitude where this density peak occurs is variable with time of day, location, season, and solar cycle.
+A profile of how electron density changes with altitude is shown in Figure \ref{fig:densprofile}.
 
@@
 
+Above 50 km, the primary neutral particles are atomic oxygen, O, molecular oxygen, \(O_2\), and molecular nitrogen, \(N_2\).
+Common ions at these altitudes are therefor mostly composed of oxygen and nitrogen, \(O^+, NO^+, O_2^+, N_2^+\), but hydrogen, \(H^+\), can contribute as well \citep{Kelley2009}.
+In the ionosphere, neutral densties are typically several orders of magnitude higher than ion densities, however even a small number of charged particles changes the behavior of the region dramatically.
+Because there are so many different ion species, plasma density is usually discussed simply in terms of electron density.
+
 \subsection{Ionosphere Regions}
 \label{sec:ionosphere_regions}
-The ionosphere has historically been divided into three distinct region.
+The ionosphere is typically divided into three region, Figure \ref{fig:densprofile}.
 The D region is located between 80 and 90 km in altitude and typically only exists in the daytime when the ionosphere is sunlit.
@@
 The F region can actually have two peaks under some daytime conditions, notated as the F1 peak and the F2 peak.
+The locations and peak densities of each of these regions is highly variable both diurnally and seasonally.
+In Figure \ref{fig:densprofile}, the daytime profile is shown with the solid line next to a nighttime profile with a dashed line.
+Note that night electron densities are lower at all altitudes due to the lack of photoionization durring nighttimes.
+In addition, the F-region peak moves to higher altitudes at night.
 
 Because the concentrations of ions and neutral particles changes with altitude, each of these regions have distinct plasma physical properties that make them unique and change how plasma structuring forms.
-In the E region, electrons are magnetized so their motion is mostly due to gyration around the magnetic field.
-However, the motion on ions is dominated by collisions between ions and neutrals, so ions in the E region are considered collisional.
-In the F region where the neutral density is much lower, both ions and electrons are magnetized.
-These differences vary the conductivity in both regions and can strongly impact how plasma waves develop, which will be discussed further in section \ref{sec:lit_experiment}.
+The main change that occurs is wether the the large-scale motion of a particle is controlled more by the magnetic field (the particle is magnetized) or by collisions with other particles (collisional).
+This can be determined by comparing the collision frequency, \(\nu_\alpha\), to the gyrofrequency, \(\Omega_\alpha\), of a particular species.
+If \(\Omega_\alpha \gg \nu_\alpha\), the particle completes many gyrations between each collision, so its motion is determined mostly by the magnetic field and it is magnetized.
+If \(\nu_\alpha \gg \Omega_\alpha\), the particle collides with other particles much more often than it complete a gyration, so it is considered collisional.
+The gyrofrquency is dependent on the mass and charge of the particle and the strength of the magnetic field, \(\Omega_\alpha = q_\alpha B/m_\alpha\), so is roughly constant through the ionosphere because the magnetic field strength does not change much through these altitudes.
+The collision frequency depends on the density of a species relative to the density of all other species in the plasma and can be calculated from a series of complicated expressions \citep{Schunk1980,Schunk2009}.
+In the E region, \(\Omega_e \gg \nu_e\), so electrons are magnetized.
+However, the motion on ions is dominated by collisions between ions and neutrals (\(\nu_i > \Omega_i\)), so ions in the E region are considered collisional.
+In the F region where the neutral density is much lower, both ions and electrons are magnetized, \(\Omega_\alpha \gg \nu_\alpha\).
+These differences in how charged particles move vary the conductivity and can strongly impact how plasma waves develop, which will be discussed further in Section \ref{sec:lit_theory}.
 
@@
 Plasma structuring can occur as variations in density, velocity, or temperature, but the main focus of this work will be density perturbations, often called irregularities.
-Density irregularities in the polar cap can occur on scales ranging from thousands of kilometers to less than a centimeter.
+Density irregularities in the polar cap can occur on scales ranging from thousands of kilometers to less than a centimeter \citep{Tsunoda1988}.
 Large-scale plasma irregularities are typically considered to be any density structuring on the scale of 1 -- 1000 km.
@@
 This can cause communication blackouts or introduce large errors into GPS calculations, effecting navigation.
-Because communication and navigation technology is implemented in both every day civilian life as well as government and military intrests, the effects of this can be far reaching.
+Because communication and navigation technology is implemented in both every day civilian life as well as government and military intrests, the effects of this kind of signal scintillation can be far reaching.
 
 Small-scale structuring is generally considered to be all irregularities less than 100 m.
-Although the effects of these scale sizes on navigation and communication systems are minimal, they can be the easiest to study because of large data sets collected from HF radars globally for several decades.
+Although the effects of these scale sizes on navigation and communication systems are minimal, they can be the easiest to study because of large data sets collected from high frequency ionospheric radars globally for several decades (discussed further in Section \ref{sec:superdarn}).
 The irregularity sizes presented here are rough categories only and a high degree of coupling between different scales is generally considered reasonable.
@@
 \label{sec:csr}
-
-Coherent Scatter Radars (CSR) have been used to study the Ionosphere for the last half century.
-Radars are ground based instruments used to detect density structures in the ionosphere's plasma.
-A basic radar consists of transmission and recieveing antenna.
-The radar transmits a radio wave, which can reflect off a suititbly sized structure in the ionosphere.
-Reflected waves can be detected by the recieving antennas.
-The time difference between when the signal was transmitted and when it was detected  an then be used to determine how far away from the radar the target is.
-
-The first instance of CSR being used to study plasma density structures in the ionosphere was the Scandanavian Twin Auroral Radar Experiment (STARE) in the 1970s and 1980s.
-STARE consisted of two very-high frequency (VHF) radars with overlapping FoVs that were designed to measure FAIs in the E region of the ionosphere \citep{Greenwald1997}.
-The advantage of two radars with overlaping FoVs was the ability to observe the same structures from two different orientations.
-Because each radar can only measure the LoS velocity of a structure, two simultanious observations from different directions allows two different velocity vecotor compontents to be found, and hense the total velocity vector can be calculated.
-This technique is still commonly used with CSR system.
+Coherent Scatter Radars (CSR) have been used to study the ionosphere for the last half century.
+The radar transmits a radio wave, which can reflect off plasma structuring in the ionosphere, and then be detected by the recieving antennas.
+The time difference between when the signal was transmitted and when it was detected and the power and phase of the returned signal can then be used to determine local characteristics of ionospheric plasma.
 
@@
 Because radars measure velocity through the doppler shift of backscatter, only the component of the velocity vector that is along the radar beam can be measured.
-Spectral width represents the width of the doppler power spectrum.
-
-%- Greenwald, 1985
+Spectral width represents the width of the doppler power spectrum \citep{Greenwald1985}.
 
@@
 %AUTO CORRELATION FUNCTION
+
+The first instance of CSR being used to study plasma density structures in the ionosphere was the Scandanavian Twin Auroral Radar Experiment (STARE) in the 1970s and 1980s.
+STARE consisted of two very-high frequency (VHF) radars with overlapping FoVs that were designed to measure FAIs in the E region of the ionosphere \citep{Greenwald1997}.
+The advantage of two radars with overlaping FoVs was the ability to observe the same structures from two different orientations.
+Because each radar can only measure the LoS velocity of a structure, two simultanious observations from different directions allows two different velocity vecotor compontents to be found, and hense the total velocity vector can be calculated.
+This technique is still commonly used with CSR system.
 

### 2016-11-28 chapter1.tex
Corrections made to CSR Section

@@
 The radar transmits a radio wave, which can reflect off plasma structuring in the ionosphere, and then be detected by the recieving antennas.
-The time difference between when the signal was transmitted and when it was detected and the power and phase of the returned signal can then be used to determine local characteristics of ionospheric plasma.
+The time difference between when the signal was transmitted and when it was detected and the power and phase of the returned signal can then be used to determine local characteristics of plasma irregularities.
 
@@
 Typically, a threshold is selected for the minimum power acceptable for a return to be considered an actual signal instead of background noise.
-LoS velocity is the component of the total velocity that is measured along one beam of a radar.
+LoS velocity is the component of the total plasma drift velocity that is measured along one beam of a radar.
 Because radars measure velocity through the doppler shift of backscatter, only the component of the velocity vector that is along the radar beam can be measured.
-Spectral width represents the width of the doppler power spectrum \citep{Greenwald1985}.
+Spectral width represents the width of the doppler power spectrum and can determine some of the characteristics of irregularities \citep{Greenwald1985}.
 
 CSR systems typically operate by transmitting pulses instead of continuously.
-This allows the distance between the radar and the target to be found (based on the time difference between when the pulse is transmitted and recieved), but also limits the temporal resolution that the radar is capable of achieving.
-
-%% INSERT RADAR PULSE DIAGRAM %%
-
-The radar starts transmitting the pulse at \(t=-\Delta t\) and finishes at \(t=0\).
-At \(t=?\), the pulse reaches the backscatter volume and begins reflecting off it.
-At \(t=T-\Delta t\), the radar begins to received the the echoed signal, which finishs at \(t=T\).
-The speed of the radar pulse in the Earth's atmosphere is approximated by the vacuum speed of light, \(c\).
-The distance between the radar and the scattering volume is known as the range and is given in equation \ref{eqn:radar_range}.
-\begin{equation}
-	\label{eqn:radar_range}
-	R = \frac{cT}{2}
-\end{equation}
-The size of the range gate is determined by the length of a radar pulse.  This is given in equation \ref{eqn:range_gate}.
+This allows the distance between the radar and the target to be found (based on the time difference between when the pulse is transmitted and received).
+The size of the range gate is determined by the length of a radar pulse, Equation \ref{eqn:range_gate}.
 \begin{equation}
@@
 Conversly, if the radar only transmits a pulse of a certain length of a certain length and then "listens" for its return, the time between transmission and return is known and the distance the pulse traveled can be calculated and therefore the location of the backscattering volume.
-However, another pulse cannot be transmitted until the first pulse returns, limiting the temporal resolution that can be achieved.
-There are methods that can be used to improve this by allowing more than one pulse at a time to be sent and received, but the pulse method does inherently require a limited temporal resolution.
-Increasing the length of a pulse can improve the temporal resolution, but decreases the spatial resolution.
+However, using this simple single-pulse method another pulse cannot be transmitted until the first pulse returns, limiting the temporal resolution that can be achieved.
+This can be improved by using a multipulse scheme, which will be discussed below.
 
-More importantly than improving temporal resolution of data availabile, a short time between pulse returns is nessisary to allow dopler velocities to be calculated, particularly at long ranges.
-The maximum dopler shifted frequency that can be observed by a radar is the Nyquist frequency, which is equivilent to half the sampling frquency.
-**Explain where the Nyquist frequency comes from?**  Radars operating in single-pulse mode have a very limited sampiling frequency, as described above, particularly at long ranges where the pulse must travel a long distance.
+A short time between pulse returns is nessisary to allow dopler velocities to be calculated, particularly at long ranges.
+The maximum dopler shifted frequency that can be observed by a radar is the Nyquist frequency, which is equivilent to half the sampling frequency, Equation \ref{eqn:nyquist}.
+\begin{equation}	
+	\label{eqn:nyquist}
+	f_n = \frac{f_s}{2}
+\end{equation}
+Radars operating in single-pulse mode have a very limited sampiling frequency, as described above, particularly at long ranges where the pulse must travel a long distance.
 This severely limits the dopler shift that can be measured, which places an upper limit on the LoS dopler velocity that can be calculated with this technique.
-Using the frequency range of HF radars and the maximum range in the F-region ionosphere where backscatter would ideally be measured, single pulse mode results in a very low maximum doppler velocity, much lower than typical ionospheric flows.
-**Example: Maximum doppler velocity that can be measured using single pulse at 2000 km range.**
+For an ionospheric radar observing a structure 2000 km away, it would take about 13 ms for a signal traveling at the speed of light to travel to the the structure and back to the radar, wich corresponds to a sampling frequency of \(f_s = 75\) Hz.
+The Nyquist frequency is then \(f_n = 37.5\) Hz, which corresponds to a maximum measurable Doppler velocity of \(\sim550\) m/s.
+Flows in the ionosphere have been known to exceed 1000--2000 m/s, so single pulse sampling is insufficient for these purposes.
 
 %BENIFITS OF MULTIPULSE: Ponomarenko and Waters, 2006
-Instead of waiting for each individual signal to be recieved before transmitting the next, HF radars typically employ a multipulse mode.
-In a multipulse mode, the radar transmits a series of pules with different timing between them.
-%AUTO CORRELATION FUNCTION
+Instead of waiting for each individual signal to be recieved before transmitting the next, ionospheric radars typically employ a multipulse mode.
+In a multipulse mode, the radar transmits a series of pules with different time intervals or lags between them.
+This can introduce complications when backscatter from different pulses at different ranges is received by the radar at the same time, referred to as cross-range interference, but these effects can be mitigated using correlation techniques.
+The smallest lag between two pulses is known as the multi-pulse increment, \(\tau\).
+All other lags are integer multiples of \(\tau\), which allows the calculation of the corresponding lags of the complex autocorrelation function (ACF).
+The ACF is used to find the spectral characteristics of the backscatter, from which the LoS velocity, spectral width, and backscatter power can be obtained.
+Overall, multi-pulse techniques are generally considered far more appropriate for ionospheric studies than single pulse \citep{Farley1972,Greenwald1983,Greenwald1985,Barthes1998,Ponomarenko2006}.
 
@@
 The Super Dual Auroral Radar Network (SuperDARN) is a global network of HF CSRs that was designed to measure small-scale plasma structuring in the ionosphere and map the global plasma convection.
-The network currently consists of 35 operational radars between the northern and southern hemispheres distributed at mid-, high-, and polar latitudes.
-Following the STARE experiment \citep{Greenwald1978}, the first HF radar was built in Goose Bay, Canada in 1983, which would become the first of the SuperDARN radars \citep{Greenwald1985}.
+The network currently consists of about 35 operational radars between the northern and southern hemispheres distributed at mid-, high-, and polar latitudes, Figure \ref{fig:superdarnmap}.
+Following the STARE experiment \citep{Greenwald1978}, the first HF ionospheric radar was built in Goose Bay, Canada in 1983, which would become the first of the SuperDARN radars \citep{Greenwald1985}.
+
+\begin{figure}
+	\includegraphics[width=\textwidth]{superdarnmap.pdf}
+	\caption{FoVs of all operational SuperDARN radars in both the northern (left) and southern (right) hemispheres.
+Polar latitude radars are shown with the red outlines.
+The radars at Rankin Inlet (northern hemisphere) and McMurdo Station (southern hemisphere) are shaded in red because these two instruments are particularly important in the following studies.
+Lines of constant magnetic latitude are shown in yellow at \(\Lambda = 40\deg, 60\deg, 80\deg\) in the northern hemisphere and at \(\Lambda = -40\deg,-60\deg,-80\deg\) in the southern hemisphere.}
+	\label{fig:superdarnmap}
+\end{figure}
 
@@
 Each radar consists of 16 independent transmit and recieving antennas.
-Beams are electronically steerable and most radars have between 16 and 24 beams in normal operation mode, each beam being \(3.25^\circ\) in azimuth.
+Beams are electronically steerable and most radars have between 16 and 24 beams in normal operation mode, each beam being \(3.25\deg\) in azimuth.
 Each beam consists of 75--100 range gates, which are generally either 15 km or 45 km in length \citep{Chisham2007}.
 SuperDARN radars were originally designed to used a 7 pulse ACF \citep{Farley1972,Greenwald1983,Greenwald1985}, however since 2011, most radars have begun using an 8 pulse sequence.
+The multi-pulse increment, \(\tau\), is 2400 \(\mu\)s, which corresponds to a Nyquist frequency of about 200 Hz, small enough to measure plasma drift velocities up to 3000 m/s.
+The spectral characteristics of backscatter of recovered from the ACF using the FITACF algorithm \citep{Ponomarenko2006}.
 
@@
 
+Because convection maps require data covering as wide a range of magnetic local time sectors as possible, SuperDARN radars are continuously operational, usually in a common mode.
+This creates a vast database of high-quality measurements of small-scale FAIs, which is very useful for other studies on irregularity occurrence and plasma structuring.
+
 \subsection{Incoherent Scatter Radar}
@@
 The radar receivers then detect a spectra of frequencies with different power.
-Analysis of this spectra will reveal the plasma frequency, and hence, the electron density, electron and ion temperature, and LoS velocity.
-**CITE Gordon, 1958**
+Analysis of this spectra will reveal the plasma frequency, and hence, the electron density, electron and ion temperature, and LoS velocity \citep{Gordon1958}.
 

### 2016-11-28 chapter1.tex
Corrections made to ISR Section

@@
 This creates a vast database of high-quality measurements of small-scale FAIs, which is very useful for other studies on irregularity occurrence and plasma structuring.
+In the studies presented here, data was primarily used from the SuperDARN radars located at Rankin Inlet, Canada (RKN) and McMurdo Station, Antarctica (MCM).
+The have been shaded in Figure \ref{fig:superdarnmap}.
 
@@
 Similar to CSRs, incoherent scatter radars (ISRs) are ground based radars that are used to probe the ionosphere.
-However, they use a fundamentally different technique to observe structuring, which allows them to measure different things than CSRs.
-Where as CSRs receive backscatter from large, "coherent" density structures in the ionosphere, ISRs operate by radar beams scattering due to the random thermal motion of electrons in the ionosphere.
-The radar receivers then detect a spectra of frequencies with different power.
-Analysis of this spectra will reveal the plasma frequency, and hence, the electron density, electron and ion temperature, and LoS velocity \citep{Gordon1958}.
+However, they use a fundamentally different technique to observe structuring, which allows them to measure parameters of the actual ionospheric plasma.
+Where as CSRs receive backscatter from relatively large, "coherent" density structures in the ionosphere, ISRs operate by radar beams scattering due to the random thermal motion of electrons in the plasma.
+The radar receivers then detect a spectra of frequencies with different power \citep{Gordon1958}.
 
-The first ISR was built in Arecibo, Puerto Rico in 1958.
+The scattering cross section for a volume of N electrons in the ionosphere is given by \(\sigma_n = N \sigma_e\) for radio waves with a wavelength small compared to the Debye length of the plasma, where \(\sigma_e\) is the scattering cross section of a single electron \citep{Gordon1958,Fejer1960}.
+The relationship to the plasma Debye length is important because at scale sizes smaller than the Debye length, the plasma is not capable of organized motion, so the motion of the elections is only due to their own thermal energy and not larger scale plasma waves.
+The signal returned from this kind of scatter was postulated to have a very low power and a very broad spectral width, however after the first ISR was built, the spectrum was found to be much narrower than previously expected \citep{Evans1969}.
+This can be attributed to the effect that the much slower ions have on free electrons in the plasma through ion acoustic waves \citep{Bowles1958}.
+The practical results of this is that ISRs can use radio waves of a much longer wavelength than originally anticipated, on the order of \(\sim\)1 m.
+
+Analysis of the returned power spectrum reveals the electron density, electron and ion temperatures, and LoS ion velocity \citep{Evans1969,Rishbeth1985,Nicolls2007}.
+Using a model of chemical composition at different altitudes, the average ion mass can also be found.
+In addition, 3D convection velocities and electric field vectors can be derived from LoS ion velocity measurements in some of the more modern ISRs by assuming both \(\vec{E}\) and \(\vec{V}_E\) are constant across the radar's FoV \citep{Heinselman2008}.
+Each instrument must be calibrated with known electron density measurements to account for noise and system constants \citep{Nicolls2007}.
+This is done using the plasma frequency measured from either the plasma line of the ISR spectra during summer daytime period or ionosonde measurements for all other times \citep{Bachivan2010,Themens2014}.
+
+The first ISR was built in Arecibo, Puerto Rico in the 1950s \citep{Gordon1958}.
 Since then, at least 8 other ISRs have been deployed around the world at equitoral, mid-, and high-latitudes.
@@
 The advantage of these new systems is that they are electronical steerable, unlike older ISRs which consisted of a large dish that had to be moved manually to change the beam direction.
-This allows beams to be transmitted in many different directions at a very high time cadence, giving the radar the capability to make measurements in multiple directions almost simultaniously.
-This is very important for creating 2D maps of ionospheric conditions and tracking how they change in time.
+This allows beams to be transmitted in many different directions at a very high time cadence, giving the radar the capability to make measurements in multiple directions almost simultaneously \citep{Nicolls2007a,Nicolls2007b,Bachivan2010}.
+This is very important for creating 2D maps of ionospheric conditions and tracking how they change in time \citep{Semeter2009,Dahlgren2012a,Dahlgren2012b}.
+AMISR systems typically have two types of pulses, a long pulse (LP) with 72 km range gates for F-region studies and an alternating code pulse (AC) with 4.5 km range gates gates for E region studies.
+The exact number and configuration of beams depends on the radar mode used and can be changed easily due to the system's electronic steering.
+In the common WorldDay mode, there are typically 11 beams and data is collected at \(\sim\)1 min intervals, although this can change slightly between different renditions of the WorldDay mode.
+
+One of the most important advantages of ISR systems over CSRs is the ability to directly measure electron density in the ionosphere.
+However, because ISRs are much scarcer and are only typically run on in WorldDay mode a few days out of every month, data is not as widely available.
+In addition, ISRs do not actually measure small-scale coherent plasma structures like the HF SuperDARN network does.
+However, they can directly measure large-scale structures and image density variations on scales larger that 100 km (limited by the number of beams and range gate size).
+For these region, it is often useful use a combination of CSR and ISR measurements in ionospheric plasma structuring studies, when possible.
+The primary ISR used in the following studies is the north face of the AMISR system at Resolute Bay, Canada, RISR-N.
+The FoV of RISR-N overlaps that of the RKN SuperDARN radar, making it ideal for these types of comparison studies.
 

### 2016-11-29 chapter1.tex
Corrections made to IRI Section, raytracing added

@@
 \label{sec:iri}
-Models can be useful tools for establishing the context of experimental measurments, especially in a global sense.
-In particular, it is impractical to achieve full global coverage of all instruments, so models can be a useful way to fill in the gaps.
-There are two main categories of models.
-Theoretical models are based on first principles and global pictures are often found by running large computer simulations based on these first principles.
-They can help give insight as to what causes specific phenomena because the process can be followed beginning to end.
-Empirical models are based purely on measured data.
-These can potentially be more accurate because they should replicate real measurments, but they also do not give any insight into how the system got to its ending state.
-In addition, empirial models are often heavily averaged.
-
-The international reference ionosphere (IRI) is an empirical model of the Earth's ionosphere.
-It was originally created in 1969 as a joint effort between the Comittee on Space Research (COSPAR) and the International Union of Radio Science (URSI) and has been periodically updated since then \citep{Rawer1978}.
+The International Reference Ionosphere (IRI) is an empirical model of the Earth's ionosphere.
+It was originally created in 1969 as a joint effort between the Committee on Space Research (COSPAR) and the International Union of Radio Science (URSI) and has been periodically updated since then \citep{Rawer1975,Rawer1978,Rawer1981,Bilitza1985,Bilitza1986,Bilitza1990,Bilitza1997,Bilitza2001,Bilitza2008}.
 The current version is IRI-2012 \citep{Bilitza2014}, which is used throughout the work done here.
-It can be accessed either as an online module or by downloading and running the FORTRAN source code, both of which are available from the International Reference Ionosphere page on the NASA website (http://iri.gsfc.nasa.gov/).
+It can be accessed either as an online module or by downloading and running the FORTRAN source code, both of which are available from the International Reference Ionosphere website (\url{http://iri.gsfc.nasa.gov/}).
 
@@
 The model requires the input of a location (latitude, longitude, and altitude) and time (year, date, and time).
-Internally, the model has the additional capability of producing profiles in any of the input parameters.
+Additionally, the model gives the option to choose between two F peak models and three bottomside thickness models.
+There are a variety of other optional input parameters such as sunspot number, ionospheric index, and F10.7 radio flux which can help make the output parameters more accurate for a particular situation.
+Internally, the model has the additional capability of producing profiles of any of the output parameters in either space or time.
 This is particularly significant for height profiles, and allows the peak heights and densities of the E and F region to be calculated.
 
-%- Empirical model of the global ionosphere
-%- Parameters that can be calculated
-%- How is it available?
-%	- Online
-%	- FORTRAN source code
-%- Biltza, 1990
+Although the IRI model is very useful for providing background plasma density conditions where no instruments are available to make real measurements, it is important to remember that it is a model and therefor only represents the average conditions.
+Although it is fairly accurate at mid-latitudes \citep{Coisson2006,Bilitza2012}, it does not necessarily represent variations within the polar cap well \citep{Themens2014,Makarevich2015}.
+In particular, the height of the F-region peak tends to be underestimated in daytime and the bottomside thickness does not show realistic seasonal and diurnal variation \citep{Themens2014}.
+Despite this, the IRI model is still a very important tool for establishing a density profile at any time in any location in the ionosphere.
+And example of an altitude profile of plasma density provided by the IRI model is shown in Figure \ref{fig:densprofile}.
+The displayed profiles shows how the electron density changes with altitude at local noon (day) and midnight (night) at \(75\deg\) N, \(0\deg\) E, geographic.
+
+One of the most important uses of the IRI model has been its integration into ray tracing simulations.
+It can be valuable to determine the degree to which HF radar beams refract through a dense ionosphere.
+This can be done through standard raytracing tools based on numerical solutions to the Hamiltonian raypath equations \citep{Haselgrove1963,Jones1975}, however these tools require a 2D density profile along the path that the radar beam travels.
+It is possible to assume a Gaussian distribution or Chapman layer as a simple model of the F-region peak, but it is much more instructive to use output from the IRI model, as it is sensitive to changes in latitude as well as seasonal and diurnal variations.
 

### 2016-11-29 thesis.sh
Incorperated superdarnmap figure

@@ -1,6 +1,8 @@
 
 #!/bin/bash
 
-gs -q -dSAFER -dBATCH -dNOPAUSE -sDEVICE=pdfwrite -dDEVICEWIDTHPOINTS=470 -dDEVICEHEIGHTPOINTS=350 -sOutputFile=/Users/ljlamarche/Desktop/UAFThesis/Figures/densprofile.pdf /Users/ljlamarche/Desktop/UAFThesis/Figures/densprofile.ps
+gs -q -dSAFER -dBATCH -dNOPAUSE -sDEVICE=pdfwrite -dDEVICEWIDTHPOINTS=530 -dDEVICEHEIGHTPOINTS=410 -sOutputFile=/Users/ljlamarche/Desktop/UAFThesis/Figures/densprofile.pdf /Users/ljlamarche/Desktop/UAFThesis/Figures/densprofile.ps
+
+gs -q -dSAFER -dBATCH -dNOPAUSE -sDEVICE=pdfwrite -dDEVICEWIDTHPOINTS=270 -dDEVICEHEIGHTPOINTS=520 -sOutputFile=/Users/ljlamarche/Desktop/UAFThesis/Figures/superdarnmap.pdf /Users/ljlamarche/Desktop/UAFThesis/Figures/SuperDARNmap.ps
 
 pdflatex LLamarcheThesis.tex

### 2016-12-01 chapter1.tex
Incorperated superdarnmap figure
Small corrections made throughout to references

@@
 
-%Figures
-%- Map of radars (Northern and southern hemisphere)
-%- Example convection map
-%- Example ISR profile
-%- Example MSISE profile?
-%- Example of IGRF?
-%- How does radar work diagram
-%- Pulse sequence diagram
-
 \subsection{Coherent Scatter Radars}
@@
 
-%BENIFITS OF MULTIPULSE: Ponomarenko and Waters, 2006
 Instead of waiting for each individual signal to be recieved before transmitting the next, ionospheric radars typically employ a multipulse mode.
@@
 
-%- How to measure power, spectral width, LoS velocity
-%	- What are these things?
-%- Hanuise, 1993b
-%- Baker, 1995
-
 \subsubsection{SuperDARN}
@@
 \begin{figure}
-	\includegraphics[width=\textwidth]{superdarnmap.pdf}
+	\includegraphics[width=\textwidth,angle=180]{superdarnmap.pdf}
 	\caption{FoVs of all operational SuperDARN radars in both the northern (left) and southern (right) hemispheres.
@@
 If both recieved backscatter from the same scattering volume, two LoS velocities could be found for that scattering volume, which could be combined to find a 2D velocity vector \citep{Ruohoniemi1989}.
-The modern method for creating convection maps involves taking all data recorded by all radars in a particular hemisphere and finding the best fit to a 2D ionospheric electrostatic potential using Legendre functions \citep{Ruohonomiemi1998}.
+The modern method for creating convection maps involves taking all data recorded by all radars in a particular hemisphere and finding the best fit to a 2D ionospheric electrostatic potential using Legendre functions \textbf{CITE???}.
 For convection maps, only data from the F region is considered.
@@
 
-Analysis of the returned power spectrum reveals the electron density, electron and ion temperatures, and LoS ion velocity \citep{Evans1969,Rishbeth1985,Nicolls2007}.
+Analysis of the returned power spectrum reveals the electron density, electron and ion temperatures, and LoS ion velocity \citep{Evans1969,Rishbeth1985,Nicolls2007a}.
 Using a model of chemical composition at different altitudes, the average ion mass can also be found.
 In addition, 3D convection velocities and electric field vectors can be derived from LoS ion velocity measurements in some of the more modern ISRs by assuming both \(\vec{E}\) and \(\vec{V}_E\) are constant across the radar's FoV \citep{Heinselman2008}.
-Each instrument must be calibrated with known electron density measurements to account for noise and system constants \citep{Nicolls2007}.
-This is done using the plasma frequency measured from either the plasma line of the ISR spectra during summer daytime period or ionosonde measurements for all other times \citep{Bachivan2010,Themens2014}.
+Each instrument must be calibrated with known electron density measurements to account for noise and system constants \citep{Nicolls2007a}.
+This is done using the plasma frequency measured from either the plasma line of the ISR spectra during summer daytime period or ionosonde measurements for all other times \citep{Bahcivan2010,Themens2014}.
 
@@
 Internally, the model has the additional capability of producing profiles of any of the output parameters in either space or time.
-This is particularly significant for height profiles, and allows the peak heights and densities of the E and F region to be calculated.
+This is particularly significant for height profiles, and allows the peak heights and densities o
+f the E and F region to be calculated.
 
@@
 And example of an altitude profile of plasma density provided by the IRI model is shown in Figure \ref{fig:densprofile}.
-The displayed profiles shows how the electron density changes with altitude at local noon (day) and midnight (night) at \(75\deg\) N, \(0\deg\) E, geographic.
+The displayed profiles shows how the electron density changes with altitude at local noon (day) and midnight (night) on January 1, 2007 at \(75\deg\) N, \(0\deg\) E, geographic.
 
@@
 \section{Theory and Observations}
-Much work has been done over the past several decades to try to understand the complicated picture of plasma structuring in the polar cap, both from theoretical analysis of plasma dispersion relations, simulations of plasma dynamics, and observations of plasma structuring using a variety of instruments.
+Much work has been done over the past several decades to try to understand the complicated picture of plasma structuring in the polar cap, including theoretical analysis of plasma dispersion relations, simulations of plasma dynamics, and observations of plasma structuring using a variety of instruments.
 
@@
 Additionally, a mechanism is still required to separate patches from the cusp region.
-High speed flow channels can "cut" segments off of a tongue of high-density plasma that has been pulled into the polar cap by convection \citep{Valladares1994,Valladares1998}.
+High speed flow channels can ``cut'' segments off of a tongue of high-density plasma that has been pulled into the polar cap by convection \citep{Valladares1994,Valladares1998}.
 Models have also shown changing large scale convection patterns quickly in time does produce density structures similar to patches \citep{Anderson1988}.
@@
 
+\begin{figure}
+	\includegraphics[width=\textwidth]{magnetosphere.pdf}
+	\caption{Diagram of the interactions between the IMF, magnetosphere, and polar cap ionosphere.
+Closed field lines in the Earth's magnetosphere are shown in grey while open field lines are light blue.
+The polar cap open-closed boundary is in green.
+Magnetic field lines that are part of the IMF are shown in dark blue.
+Magnetic reconnection occurs along the X-line (red) on the dayside of the magnetopause (dashed line).
+The X-line can be mapped back along magnetic field lines (orange) to the ionosphere, where their projection is the merging gap (pink dotted line).
+Durring reconnection events, the open-closed boundary expands equatorwards as more open field lines are formed and there is plasma flow across the merging gap.}
+	\label{fig:magnetosphere}
+\end{figure}
+
 Magnetic reconnection between the IMF and the Earth's magnetosphere occurs along the X-line on the dayside magnetopause (blue line in figure \ref{fig:magnetosphere}).
@@
 A series of reconnection events will create a string of isolated polar patches propigating through the center of the polar cap \citep{Lockwook1992b}.
+
+\begin{figure}
+	\includegraphics[width=\textwidth]{patch_formation.pdf}
+	\caption{Diagram of the polar patch formation through transient magnetopause reconnection.
+The open-closed boundary is the green line.
+Black contours represent plasma flow.
+Highly ionized (dense) daytime plasma is shown by the light orange regions.
+The burst of reconnection is occurring in Figures \ref{fig:patch_formation}a--\ref{fig:patch_formation}c but stops in Figures \ref{fig:patch_formation}d--\ref{fig:patch_formation}f.}
+	\label{fig:patch_formation}
+\end{figure}
 

### 2016-12-02 chapter1.tex
Corrections made to Experiment/Theory Sections

@@
 
-\section{Theory and Observations}
-Much work has been done over the past several decades to try to understand the complicated picture of plasma structuring in the polar cap, including theoretical analysis of plasma dispersion relations, simulations of plasma dynamics, and observations of plasma structuring using a variety of instruments.
+\section{A Brief Review of Plasma Structuring Theory and Observations}
+Much work has been done over the past several decades to study the complicated picture of plasma structuring in the polar cap, including theoretical analysis of plasma dispersion relations, simulations of plasma dynamics, and observations of plasma structuring using a variety of instruments.
+Here, a survey will be given of some of these advancements and the current state of understanding plasma structring on a range of scales.
 
 \subsection{Theory and Simulations}
+Large-scale structuring in the polar cap can take a variety of forms, Section \ref{sec:polar_structure}, but here we will focus on large regions of enhanced plasma density known as polar patches  \citep[][e.g.]{Weber1984,Weber1986,Buchau1983,Buchau1985}.
+The first observations of polar patches were made in the 1960s by \citet{Hill1963}.
+They can range in size up to 1500 km in diameter and can have peak densities as much as eight times that of the background ionospheric plasma \citep{Weber1986,Hosokawa2016}.
+Polar patches are know to emerge on the dayside and then propagate with the background plasma convection across the polar cap to the nightside, where they either disintegrate or recombine with enhanced-density plasma in the auroral oval \citep{Weber1985,Weber1986}.
 
-Polar patches are large regions of enhanced plasma density in the polar F region \citep[][e.g.]{Weber1984,Buchau1983}.
-They are know to emerge on the dayside and then propagate with the background plasma convection across the polar cap to the nightside, where they either disintegrate or recombine with enhanced-density plasma in the auroral oval \textbf{CITE SOMETHING}.
-
-There have been at least 6 mechanisms proposed for how polar patches form on the dayside.
+There have been at least 6 mechanisms proposed for how polar patches form on the dayside \citep{Crowley1996,Carlson2012}.
 \begin{enumerate}
@@
 
-In order to discuss patch formation by transient magnetic reconnection, it is important to first understand the steady-state ionospheric convection patterns in the polar cap and how they relate to the magnetosphere.
-Figure \ref{fig:magnetosphere} is a schematic of how the IMF, magnetosphere, and ionosphere are all interconnected.
-In general, magnetic field lines in the magnetosphere can either be closed, meaning they connect only the north and south pole of the earth, or open, when one of the ends of the field lines originates from the sun such that the field line is connected directly to the IMF.
-The polar cap boundary is generally defined as the border between open and closed field lines, shown by the dark line in figure \ref{fig:magnetosphere}.
-By Faraday's Law, the total EMF around the polar cap boundary is equal to the rate of change of magnetic flux, equation \ref{eqn:pc_imf}.
-In the polar cap, the magnetic field strength is relatively constant at around \(5\times 10^-5\) T, so any changes in magnetic flux must be due to changes in the polar cap area, \(A_{pc}\) due to the polar cap boundary expanding or contracting \citep{Lockwood1992a}.
+In order to discuss patch formation by transient magnetic reconnection, it is important to first understand the steady-state ionospheric convection patterns in the polar cap and how they relate to the magnetosphere \citep{Cowley1980}.
+Figure \ref{fig:magnetosphere} is a schematic of how the IMF, magnetosphere, and ionospheric polar cap are all interconnected.
+In general, magnetic field lines in the magnetosphere can either be closed (grey lines in Figure \ref{fig:magnetosphere}), meaning they connect only the north and south pole of the earth, or open (light blue lines in Figure \ref{fig:magnetosphere}), when one of the ends of the field lines originates from the sun such that the field line is connected directly to the IMF.
+IMF lines that are not connected to the earth are shown in dark blue in figure \ref{fig:magnetosphere}.
+The polar cap boundary is generally defined as the border between open and closed magnetic field lines, shown by the green line in Figure \ref{fig:magnetosphere}.
+
+By Faraday's Law, the total electromotive force, \(\xi\) around the polar cap boundary is equal to the rate of change of magnetic flux, \(\Phi_B\) within the polar cap, Equation \ref{eqn:pc_imf} \citep{Lockwood1992a}.
+In steady-state solutions, \(d\Phi_B/dt = 0\), so \(\xi=0\) by definition, however the polar cap is known to be a highly dynamic system, so changes in magnetic flux are expected.
 \begin{equation}
 	\label{eqn:pc_imf}
-	-E = \frac{d\Phi_B}{dt} = B\frac{dA_{pc}}{dt}
+	-\xi = \frac{d\Phi_B}{dt} = B\frac{dA_{pc}}{dt}
 \end{equation}
-Magnetic reconnection often occurs on both the dayside and nightside of the magnetosphere, changing the amount of open magnetic flux, so the polar cap boundary can be highly dynamic.
+In the polar cap at ionospheric altitudes, the magnetic field strength is relatively constant at around \(5\times 10^-5\) T, so any changes in magnetic flux must be due to changes in the polar cap area, \(A_{pc}\) due to the polar cap boundary expanding or contracting \citep{Lockwood1992a}.
+The polar cap boundary changes in response to magnetic reconnection between the magnetosphere and the IMF  on both the day and night sides changing the amount of open magnetic flux.
 
@@
 
-Magnetic reconnection between the IMF and the Earth's magnetosphere occurs along the X-line on the dayside magnetopause (blue line in figure \ref{fig:magnetosphere}).
-The X-line in the magnetopuase can be mapped along the magnetic field lines to the ionosphere, where it is referred to as the merging gap (pink line in figure \ref{fig:magnetosphere}).
-During reconnection events, the merging gap moves equatorward as the open magnetic flux into the polar cap increases \citep{Lockwood1992a}.
-In addition, the potential along the X-line also gets mapped the the merging gap, assuming there is minimal drop in potential along the magnetic field lines \citep{Lockwood1992b}.
-The combination of these two effects results in flux being transferred across the boundary at a rate equal to the applied voltage  along the X-line and hence plasma flow across the merging gap \citep{Lockwood1992b}, as seen in figure \ref{fig:patch_formation}.
-As the reconnection event continues, the flow across the merging gap creates a convection pattern, which starts to pull high density plasma from the dayside into the polar cap.
-If this continues consistently, eventually it can create a "tongue" of high-density plasma stretching between the dayside merging gap and the center of the polar cap, often called the tongue of ionization.
-However, when the reconnection burst ends, the merging gap begins to relax back towards the polar cap and brings the dense plasma with it, figure \ref{fig:patch_formation}.
-Before the flow stops completely, the patch "pinches off" of the ionization tounge and becomes an isolated island of high density plasma in the polar cap.
-The merging gap returns to its original position and the entire process can repeat with the next reconnection event.
-A series of reconnection events will create a string of isolated polar patches propigating through the center of the polar cap \citep{Lockwook1992b}.
+Magnetic reconnection between the IMF and the Earth's magnetosphere occurs along the X-line on the dayside magnetopause (red line in figure \ref{fig:magnetosphere}).
+The X-line in the magnetopuase can be mapped along the magnetic field lines (orange) to the ionosphere, where it is referred to as the merging gap (pink dotted line in figure \ref{fig:magnetosphere}) and lies along the steady state polar cap boundary.
+
+During reconnection events, the polar cap boundary moves equatorward as the open magnetic flux into the polar cap increases, Figure \ref{fig:magnetosphere} \citep{Cowley1991,Lockwood1992a}.
+In addition, the potential along the X-line also gets mapped to the merging gap, assuming there is minimal drop in potential along the magnetic field lines \citep{Lockwood1992b}.
+The combination of these two effects results in flux being transferred across the boundary at a rate equal to the applied voltage  along the X-line and hence plasma flow across the merging gap \citep{Lockwood1992b}, as seen in Figure \ref{fig:patch_formation}.
+
+Figure \ref{fig:patch_formation} is a schematic of how transient magnetopause reconnection forms polar patches.
+In all panels, the the dayside is the top of the figure and the green line shows the open-closed polar cap boundary, similar to Figure \ref{fig:magnetosphere}.
+Plasma flow contours are shown in black and the light orange region represents high density from the dayside reservoir.
+The reconnection burst starts at Figure \ref{fig:patch_formation}a and the increasing amount of open magnetic flux moves the dayside open-closed boundary equatorwards, as described above \citep{Cowley1991}.
+
+As reconnection continues in Figure \ref{fig:patch_formation}b, the open-closed boundary continues to expand equatorwards as more open flux is produced, which excites a convection pattern in the polar cap \citep{Cowley1991}.
+Simultaneously, this convection starts to transport dense dayside plasma polewards \citep{Lockwood1992b}.
+As more open flux is produced, Figure \ref{fig:patch_formation}c, the strength of the plasma flow increases and a large blob of dense, dayside plasma is pulled into the polar cap \citep{Lockwood1992b}.
+
+At Figure \ref{fig:patch_formation}d, the burst of reconnection stops and the open-closed boundary begins to relax polewards \citep{Cowley1991}.
+As it moves, it pulls the blob of dense plasma within the polar cap with it, separating it from the reservoir on the dayside \citep{Lockwood1992b}.
+Before the convection flow stops completely, Figure \ref{fig:patch_formation}e, the blob of dense plasma ``pinches off'' from the dayside plasma to become an independant patch \citep{Lockwood1992b}.
+The exact mechanism by which this happens is not well understood, but it is thought to be related to small variations in the IMF By component shifting the convection pattern slightly \citep{Cowley1980,Lockwood1992b}.
+By Figure \ref{fig:patch_formation}f, the open-closed boundary has returned completely to its original position and there is a newly formed patch within in the polar cap.
+The process can now repeat to form another patch within the polar cap.
+
+A series of reconnection bursts can create a line of patches all propagating across the polar cap.
+Reconnection bursts typically last about 2 minutes with anywhere between 7 and 25 minutes beween bursts \citep{Foster1984,Etemadi1988,Lockwood1992b}.
+However, observations have shown convection flow to be close to continuous for much longer periods.
+This can be explained by each reconnection burst exciting flow for a much longer interval than the time between bursts.
+In this way, a reconnection burst can excite plasma flow before the convection from the previous burst has stopped, and a series of overlapping reconnection events like this can create continuous plasma flow through the polar cap \citep{Cowley1991}.
 
@@
 
-Plasma structuring occurs in the ionosphere when conditions are such that wave growth occurs.
-Because different background conditions change factors such as the density of both neutrals and ions, the magnetic field strength, and so forth, the collison frequency, gyroradii, and ion inertial effects change at different altitudes.
-This causes a variety of different instability mechanisms to be operational depending on the background conditions.
+Plasma structuring, especially at smaller scales, is often discussed in terms of the growth or damping of plasma waves, or instabilities.
+As discussed previously in Section \ref{sec:ionosphere_regions}, the plasma characteristics change with altitude in the ionosphere, which causes a variety of different instability mechanisms to be operational depending on the region considered.
 The Farley-Buneman instability (FBI), or the modified two-stream instability tends to be a major factor in the E region where ion inertial effects are high \citep{Farley1963,Buneman1963}.
 Also operational in the E region but more dominant at higher altitudes is the gradient-drift instability (GDI) \citep{Simon1963,Hoh1963,Linson1970}.
-When GDI is stable, wave growth can still occur sometimes along the magnetic field, which is known as the current-conductive instability (CCI).
-Additionally, there are shear-driven processes, known as the Kelvin-Helmholtz instability (KHI) \textbf{CITE???}.
+If a wave has a component along the magnetic filed line, the current-conductive instability (CCI) can be operational even if GDI is stable.
+Addition instabilities mechanisms emerge if shears are considered, when \(\nabla E \neq 0\), such as the Kelvin-Helmholtz instability (KHI).
 
@@
 
-Some of the first 1D investigations of the linear GDI growth rate found that it can be described by a simple expression involving the plasma drift velocity, \(V_E\), and the gradient scale length, \(L\) \citep{Simon1963,Hoh1963,Linson1970}.
+Some of the first 1D investigations of the linear GDI growth rate found that it can be described by a simple expression involving the plasma drift velocity, \(V_E = \vec{E}\times\vec{B}\), and the gradient scale length, \(L = |n/(\partial n/\partial x)|\), Equation \ref{eqn:gdi_old} \citep{Simon1963,Hoh1963,Linson1970}.
 \begin{equation}
@@
 \end{equation}
-This expression is only valid if the density gradient is in the same direction as the plasma drift velocity and wavevector direction is not considered.
-The results of \citet{Keskinen1982,Keskinen1983} do consider the wavevector, \(\vec{k}\), to some extent, but still assume a general relationship between density gradients and plasma drift \citep{Tsunoda1988}.
+This expression is only valid if the density gradient is in the same direction as the plasma drift velocity and the wave is propagating perpendicular to both.
+The results of \citet{Keskinen1982,Keskinen1983}, Equation \ref{eqn:gdi_80s}, do consider an arbitrary wavevector, \(\vec{k}\), explicitly to some extent, but still assume a general relationship between density gradients and plasma drift \citep{Tsunoda1988}.
 \begin{equation}
@@
 The directional dependance of GDI has been shown to be significant, such that the growth rate can actually change signs (determining whether wave growth or damping occurs) with different wavevector directions, so it is important to consider this factor \citep{Makarevich2014c}.
-In addition, these expressions are only applicable in a particular altitudinal regieme, usually the F region.
+In addition, these expressions have historically only applicable in a particular altitudinal regieme, usually the F region.
 
-Recently, a new expression has been introduced that considers an arbitrary altitude, wavevector, and density gradient within the ionosphere, which allows FBI, GDI, an CCI to be considered with a single dispersion relation \citep{Makarevich2016a}.
+Recently, a new expression has been introduced that considers any arbitrary wavevector, density gradient, and altitude within the ionosphere, which allows FBI, GDI, an CCI to be considered with a single dispersion relation \citep{Makarevich2016a}.
 \begin{equation}
@@
 \end{equation}
-As before, \(\vec{k}\) is the wavevector and \(\vec{V}_{\alpha 0}\) is the velocity of species \(\alpha\).
+As before, \(\vec{k}\) is the wavevector, \(\omega\) is the wave frequency, and \(\vec{V}_{\alpha 0}\) is the velocity of species \(\alpha\).
 The species \(\alpha\) is given by either \(i\) for ions or \(e\) for electrons.
@@
 This means that \(\vec{k}_\perp = \vec{k}\), \(y=0\), and \(Y=0\).
-Finally, consider only low frequencies (\(\omega \ll \Omega_\alpha\)) and long wavelengths (\(\vec{k}\cdot\vec{V}_{\alpha 0} \\ \nu_\alpha\)).
+Finally, consider only low frequencies (\(\omega \ll \Omega_\alpha\)) and long wavelengths (\(\vec{k}\cdot\vec{V}_{\alpha 0} \gg \nu_\alpha\)).
 If \(D_\alpha\) is expressed as shown below, it can bee seen that these limits result in the first two terms being negligible such that \(D_\alpha \approx r_\alpha\).
@@
 Multiplying both the numerator and denominator by \(-i+a+Rb\) removes the imaginary component from the denominator.
-Additionally the local approximation \(G/k_\perp \ll 1\) is taken, which results in \(a\), \(b\), and \(c\) being small so that any second order terms in these quantities can be neglected.
+Additionally the local approximation \(G \ll k_\perp\) is taken, which results in \(a\), \(b\), and \(c\) being small so that any second order terms in these quantities can be neglected.
 \begin{equation}
@@
 \end{equation}
-Recognizing that \(s_es_i\left(1+R^2\right) = (1+\psi)^{-2}\) and \(b = \vec{G}\cdot\vec{k}\times\uvec{b}/k^2 = -\vec{k}\cdot\uvec{b}\times\vec{G}/k^2\) by vector identities a final expression for the GDI growth rate in the F region can be found, which agrees well with the results of \citet{Makarevitch2014c}.
+Recognizing that \(s_es_i\left(1+R^2\right) = (1+\psi)^{-2}\) and \(b = \vec{G}\cdot\vec{k}\times\uvec{b}/k^2 = -\vec{k}\cdot\uvec{b}\times\vec{G}/k^2\) by vector identities, a final expression for the GDI growth rate in the F region can be found, which agrees well with the results of \citet{Makarevitch2014c}.
 \begin{equation}
@@
 
-\subsection{Observations}
+\subsection{Observations of Polar Cap Structuring}
 
-Polar patches were first observed in the 1980s using optical and radio techniques \citep{Weber1981,Buchau1983,Weber1984}.
+Polar patches were first observed in the 1960s \citep{Hill1963}, but much of the work of characterizing them using optical and radio techniques was not accomplished until two decades later \citep{Weber1981,Weber1984,Weber1986,Buchau1983,Buchau1985}.
 \citet{Weber1981} identified sun-aligned arcs that extend  along the sun-earth line through large parts of the polar cap using all-sky imaging photometers (ASIP).
@@
 These patches could have densities up to \(10^6\) cm\(^{-3}\) and tended to drift antisunward.
+Additionally, \citet{Buchau1983} found that polar patches were more likely to be observed for southward IMF \(B_z\) while sun-aligned arcs were more common for northwards \(B_z\).
 \citet{Weber1984} found patches of enhanced ionization between 800--1000 km during moderately disturbed geomagnetic conditions, again using a combination of ASIP images, ionosonde measurements, and data from Dynamic Explorer 2 (DE-2).
-These finings confirmed that density enhancements tended to drift antisunward at speeds of 500-1000 m/s and patches were not colocated with enhanced particle precipitation within the polar cap, indicating that this is not the source of the density enhancement.
+These findings confirmed that density enhancements tended to drift antisunward at speeds of 500-1000 m/s and patches were not colocated with enhanced particle precipitation within the polar cap, indicating that this is not the source of the density enhancement.
 Over time, the trailing edge of a patch became steeper than the leading edge and irregularities developed on the trailing edge.
+Conversely, \citet{Buchau1985} found that plasma densities within patches were comparable with those measured at dayside sub-auroral latitudes, suggesting that dayside solar-illuminated plasma is the source of dense plasma that form polar patches.
 
@@
 In addition, the cusp then provides a gateway of highly ionized sunlit plasma into the dark polar cap.
-\citet{Rodger1996} also considered occurrence of patches but with a HF radar in Halley, Antarctica.
+Several studies had previously reported density enhanced patches convecting from the dayside through the cusp/throat region toward the polar cap \citep{Kelley1984,Foster1984,Foster1985,Foster1993}.
+This also agrees with observations made by \citet{Sojka1982,delaBeaujardiere1985}  \citet{Rodger1996} also considered occurrence of patches but with a HF radar in Halley, Antarctica.
 The diurnal variation in patch occurrence was found to peak at magnetic noon and the seasonal variation in equinox months.
 The only correlation between hourly averaged solar wind parameters that was found was a bias towards patches occurring with a negative IMF \(B_z\) component.
+A review of historical observations of polar patches has been provided by \citet{Crowley1996}
 
 The theoretical mechanism for patch formation via transient magnetic reconnection has been described previously, Section \ref{sec:patch_formation}.
-\citet{Carlson2004} has provided experimental evidence to support this mechanism.
-Five signatures of transient magnetic reconnection were identified and then a patch formation event was examined for these signatures using the European Incoherent Scatter Scientific Association (EISCAT) Svalbard Radar.
+A variety of experimental evidence has been presented to support this mechanism \citep{Cowley1998,Carlson2002}.
+Experiments run in the 1980s using the European Incoherent Scatter Scientific Association (EISCAT) Troms\o, Norway ISR showed that plasma flow occurred near noon shortly after southward IMF \(B_z\) arrives at the magnetopause \citep{Etemadi1988,Todd1988}.
+Later, \citet{Lockwood1993a,Lockwood1993b} showed that these plasma flows are pulsed in the region of the polar cap boundary, as predicted by \citet{Cowley1991}.
+Observations made by an ASIP in Svalbard in 1984 are fully consistent with patch formation through transient magnetopause reconnection \citep{Carlson1996,Carlson2002}.
+\citet{Carlson2004} identified five signatures of transient magnetic reconnection and then examined a patch formation event for these signatures using the EISCAT Svalbard Radar.
 All five signatures were observed as expected, strengthening the idea that transient magnetopause reconnection is responsible for patch formation.
-\citet{Carlson2006} furthered this using a observations from the EISCAT Svalbard Radar, the EISCAT VHF radar in Troms\o, Norway, an ASIP, and a meridian scanning photometer (MSP).
+\citet{Carlson2006} furthered this using a observations from the EISCAT Svalbard Radar, the EISCAT VHF radar in Troms\o an ASIP, and a meridian scanning photometer (MSP).
 They managed to identify and track a series of patches directly from the subauroral plasma reservoir and show that the boundary moves equatorward before relaxing poleward, one of the main predictions of the transient magnetopause reconnection mechanism \citep{Lockwood1992b}.
@@
 
-It is known that the polar cap ionosphere is highly structured in a non-trivial manner \citep{Carlson2012}.
+It is well accepted that the polar cap ionosphere is highly structured in a non-trivial manner \citep{Carlson2012}.
 In addition to being of purely scientific interest, this structuring creates complications when sending radio waves through the ionosphere.
@@
 In addition, most experimental results are compared with predictions made by a linear fluid theory view of ionospheric plasma behavior.
-Specific objectives are as follows:
+Specific objectives are as follows.
 
@@
 The model allows certain factors to be considered that can be challenging to investigate experimentally because it is difficult to isolate particular effects.
-The model used considers the direction of a large, elongated polar patch as well as it drift direction to be arbitrary in relation to the radar and independent of each other, which gives some insight as to how structuring would be expected with polar patches in orientations that are not often observed.
+The model used considers the direction of a large, elongated polar patch as well as its drift direction to be arbitrary in relation to the radar and independent of each other, which gives some insight as to how structuring would be expected with polar patches in orientations that are not often observed.
 
-The third objective is to investigate the relationship between factors that control GDI and small-scale plasma structuring.
-This combines the theoretical consideration of where growth rate is greatest with where structuring is observed the most around a polar patch.
+The third objective is to investigate the relationship between factors that control the linear GDI growth rate and small-scale plasma structuring.
+This combines the theoretical consideration of where growth rate is greatest with the experimental record of where structuring is observed around a polar patch.
 This is particularly important because even though plasma structuring is often observed on the trailing edge of polar patches and this is where GDI predicts the greatest growth rates, GDI theory actually predicts growth of much larger irregularities than the small scale structuring observed by HF radars.

### 2016-12-05 chapter1.tex
Small modifications recommended by Makarevich

@@
 Below the ionosphere, the thermosphere consists mostly of neutral particles, however neutral winds can influence the ionosphere through viscous interactions along the lower boundary.
-All of these regions contain incredibly complex dynamics and are additionally all interconnected, creating a complicated system that is extremely challenging to describe completely.
+Overall, the charged particle interactions in the ionosphere create a highly complex system, which is further complicated by coupling with both the thermosphere from below and the magnetosphere and solar wind from above.
 
@@
 The exact altitude where this density peak occurs is variable with time of day, location, season, and solar cycle.
-A profile of how electron density changes with altitude is shown in Figure \ref{fig:densprofile}.
+A profile of how electron density changes with altitude is shown in Figure \ref{fig:densprofile}, found from the International Reference Ionosphere (IRI) model run on January 1, 2007 at \(75\deg\) N, \(0\deg\) E, geographic (discussed further in Section \ref{sec:iri}).
 
@@
 	\includegraphics[width=\textwidth]{densprofile.pdf}
-	\caption{Altitude profile of electron density throughout the ionosphere.  Typical ranges of the D, E, and F regions are identified.}
+	\caption{Altitude profile of electron density throughout the ionosphere found from the IRI model.  Typical ranges of the D, E, and F regions are identified.}
 	\label{fig:densprofile}
@@
 
-\subsubsection{SuperDARN}
+\subsubsection{Super Dual Auroral Radar Network (SuperDARN)}
 \label{sec:superdarn}
@@
 \begin{figure}
-	\includegraphics[width=\textwidth,angle=180]{superdarnmap.pdf}
+	\includegraphics[width=1.3\textwidth,angle=90]{superdarnmap.pdf}
 	\caption{FoVs of all operational SuperDARN radars in both the northern (left) and southern (right) hemispheres.
@@
 If both recieved backscatter from the same scattering volume, two LoS velocities could be found for that scattering volume, which could be combined to find a 2D velocity vector \citep{Ruohoniemi1989}.
-The modern method for creating convection maps involves taking all data recorded by all radars in a particular hemisphere and finding the best fit to a 2D ionospheric electrostatic potential using Legendre functions \textbf{CITE???}.
+The modern method for creating convection maps involves taking all data recorded by all radars in a particular hemisphere and finding the best fit to a 2D ionospheric electrostatic potential using Legendre functions \citep{Ruohoniemi1998}.
 For convection maps, only data from the F region is considered.
-If there is not enough velocity data to confine the fit sufficiently, the measured data points are supplimented with a statistical model based on solar wind IMF conditions \textbf{CITE MCWILLIAMS???}.
+If there is not enough velocity data to confine the fit sufficiently, the measured data points are supplimented with a statistical model based on solar wind IMF conditions \citep{Ruohoniemi1998}.
 
@@
 
-\section{A Brief Review of Plasma Structuring Theory and Observations}
+\section{A Brief Review of Plasma Structuring Theory and Observations in the Polar Cap}
 Much work has been done over the past several decades to study the complicated picture of plasma structuring in the polar cap, including theoretical analysis of plasma dispersion relations, simulations of plasma dynamics, and observations of plasma structuring using a variety of instruments.
@@
 
-\subsection{Theory and Simulations}
+\subsection{Theory and Simulations of Polar Patches and Plasma Instabilities}
 Large-scale structuring in the polar cap can take a variety of forms, Section \ref{sec:polar_structure}, but here we will focus on large regions of enhanced plasma density known as polar patches  \citep[][e.g.]{Weber1984,Weber1986,Buchau1983,Buchau1985}.
 The first observations of polar patches were made in the 1960s by \citet{Hill1963}.
-They can range in size up to 1500 km in diameter and can have peak densities as much as eight times that of the background ionospheric plasma \citep{Weber1986,Hosokawa2016}.
+They can range in size up to 1500 km in diameter and can have peak densities as much as eight times that of the background ionospheric plasma \citep{Weber1986,Hosokawa2014}.
 Polar patches are know to emerge on the dayside and then propagate with the background plasma convection across the polar cap to the nightside, where they either disintegrate or recombine with enhanced-density plasma in the auroral oval \citep{Weber1985,Weber1986}.
@@
 	\item Average flow patterns varying in time \citep{Anderson1988}
-	\item Transient magnetopause reconnection \citep{Lockwook1992b}
+	\item Transient magnetopause reconnection \citep{Lockwood1992b}
 	\item Plasma production by cusp particle precipitation \citep{Rodger1994,Millward1999}
@@
 \begin{figure}
-	\includegraphics[width=\textwidth]{magnetosphere.pdf}
+	\includegraphics[width=0.75\textwidth,angle=270]{magnetosphere.pdf}
 	\caption{Diagram of the interactions between the IMF, magnetosphere, and polar cap ionosphere.
@@
 \begin{figure}
-	\includegraphics[width=\textwidth]{patch_formation.pdf}
+	\includegraphics[width=0.75\textwidth,angle=270]{patch_formation.pdf}
 	\caption{Diagram of the polar patch formation through transient magnetopause reconnection.
@@
 This agrees well with many studies that have examined the asymmetry of small-scale plasma structuring surrounding plasma patches.
+
 This asymmetry was first observed by \citet{Weber1984}.
-Later, \citet{Milan2002b} found that backscatter power from HF radars tended to be greater on the trailing edges of moving structures, which was attributed to GDI being unstable on the trailing edge but stable on the leading edge.
-\citet{Koustov2012} found similar results examining sun-aligned arcs.
-HF radar backscatter power tended to be greater trailing behind a sun-aligned arc moving through the polar cap that was identified with ASIP data.
+Backscatter power from HF radars tends to be greater on the trailing edges of both moving high-density polar structures \citep{Milan2002b} and sun-aligned arcs identified from ASIP data \citep{Koustov2012}, which is generally attributed to GDI being unstable on the trailing edge but stable on the leading edge of such large, moving density structures.
+Studies of patches using all-sky airglow imagers (ASI) found that the density gradient on the leading edge of polar patches tends to be 2--3 times steeper than that on the trailing edge \citep{Hosokawa2016}.
+Additionally, large finger-like structures on the trailing edge of patches have been identified that were 10s--100s of kilometers in size and agree with predictions made by GDI simulations \citep{Gondarenko2004}.
 \citet{Moen2012} made direct measurements of plasma density structuring using the ICI-2 sounding rocket.
-This study provided evidence that decameter scale plasma structuring (the same as observed by HF radars) had spawned from kilometer scale density gradients and structuring was greatest where GDI was operational.
-\citet{Hosokawa2016} also considered the asymmetry of patches using an all-sky airglow imager (ASI) and found that the gradient on the leading edge of patches was typically 2--3 times steeper than that on the trailing edge.
-In addition, they found large finger-like structures on the trailing edge of patches that were 10s--100s of kilometers in size.
-These types of structures agree with GDI predictions, in particular they match the simulations of \citet{Gondarenko2004}.
+This study provided evidence that decameter scale plasma structuring (the same as observed by HF radars) had spawned from kilometer scale density gradients (that observed by ASIP and ASI methods) and structuring was greatest where GDI was operational.
 
@@
 
-Scientifically, the ionosphere is an important part of the highly coupled Sun-Earth environment.
+The ionosphere is an important part of the highly coupled Sun-Earth environment.
 The polar cap ionosphere essentially serves as the boundary conditions for waves propagating through the magnetosphere.
@@
 
-The aim of this work is to examine many different factors behind plasma irregularity production in the F region polar cap using both observations and theory.
+The aim of this work is to examine some of the factors (such as global solar control and GDI directional dependance) behind plasma irregularity production in the F region polar cap using both observations and theory.
 The majority of observational data has been found from ground based ionospheric radars, both CSR systems (Section \ref{sec:csr}) and ISR systems (Section \ref{sec:isr}), with particular emphasis on radars within the SuperDARN network, Section \ref{sec:superdarn}.
@@
 A direct comparison of the two improves our understanding of how growth rates are related to small-scale plasma structuring.
+
+The outline of this thesis is as follows.
+Chapter \ref{sec:introduction} has provided an overview of the polar ionosphere and both macro- and microphysical plasma structuring mechanisms.
+Chapter \ref{sec:paper1} will discuss plasma irregular occurrence and the extent to which solar factors control this occurrence.
+Chapter \ref{sec:paper2} will consider the directional dependance of the GDI growth rate, and in particular give recommendations for the optimal orientation between a ground-based ionospheric radar and a large-scale density structure to observe small-scale structuring.
+Chapter \ref{sec:paper3} will examine coincident observations made an IRS of background plasma density conditions and a CSR of small-scale plasma structuring in the polar cap and provide evidence for small-scale structuring being directionally dependent in a similar fashion to GDI predictions for large-scale plasma structuring.
+Chapter \ref{sec:conclusion} will conclude this thesis and give a summary of the most significant results.
+The implications of these results will be discussed along with questions that are still open and recommendations for future research.

### 2016-12-06 introduction.tex
Restructured chapter1.tex as introduction.tex

--- 2016-12-05 chapter1.tex
@@
 % Chapter 1: Introduction
+\label{sec:introduction}
 
@@
 The advantage of these new systems is that they are electronical steerable, unlike older ISRs which consisted of a large dish that had to be moved manually to change the beam direction.
-This allows beams to be transmitted in many different directions at a very high time cadence, giving the radar the capability to make measurements in multiple directions almost simultaneously \citep{Nicolls2007a,Nicolls2007b,Bachivan2010}.
+This allows beams to be transmitted in many different directions at a very high time cadence, giving the radar the capability to make measurements in multiple directions almost simultaneously \citep{Nicolls2007a,Nicolls2007b,Bahcivan2010}.
 This is very important for creating 2D maps of ionospheric conditions and tracking how they change in time \citep{Semeter2009,Dahlgren2012a,Dahlgren2012b}.
@@
 \end{enumerate}
-Although particle precipitation has been shown to create enhanced plasma density in the cusp region \citep{Roger1994}, it is unreasonable for this mechanism to account for the largest density enhancements observed in polar patches.
+Although particle precipitation has been shown to create enhanced plasma density in the cusp region \citep{Rodger1994}, it is unreasonable for this mechanism to account for the largest density enhancements observed in polar patches.
 These high densities must originate from the resivor of solar illuminated plasma on the dayside.
@@
 \end{equation}
-Recognizing that \(s_es_i\left(1+R^2\right) = (1+\psi)^{-2}\) and \(b = \vec{G}\cdot\vec{k}\times\uvec{b}/k^2 = -\vec{k}\cdot\uvec{b}\times\vec{G}/k^2\) by vector identities, a final expression for the GDI growth rate in the F region can be found, which agrees well with the results of \citet{Makarevitch2014c}.
+Recognizing that \(s_es_i\left(1+R^2\right) = (1+\psi)^{-2}\) and \(b = \vec{G}\cdot\vec{k}\times\uvec{b}/k^2 = -\vec{k}\cdot\uvec{b}\times\vec{G}/k^2\) by vector identities, a final expression for the GDI growth rate in the F region can be found, which agrees well with the results of \citet{Makarevich2014c}.
 \begin{equation}
@@
 Studies of patches using all-sky airglow imagers (ASI) found that the density gradient on the leading edge of polar patches tends to be 2--3 times steeper than that on the trailing edge \citep{Hosokawa2016}.
-Additionally, large finger-like structures on the trailing edge of patches have been identified that were 10s--100s of kilometers in size and agree with predictions made by GDI simulations \citep{Gondarenko2004}.
+Additionally, large finger-like structures on the trailing edge of patches have been identified that were 10s--100s of kilometers in size and agree with predictions made by GDI simulations \citep{Gondarenko2004b}.
 \citet{Moen2012} made direct measurements of plasma density structuring using the ICI-2 sounding rocket.
@@
 The implications of these results will be discussed along with questions that are still open and recommendations for future research.
+
+\bibliographystyle{plainnat}
+\bibliography{references}

### 2016-12-06 thesis.sh

@@ -5,4 +5,14 @@ gs -q -dSAFER -dBATCH -dNOPAUSE -sDEVICE=pdfwrite -dDEVICEWIDTHPOINTS=530 -dDEVI
 
 gs -q -dSAFER -dBATCH -dNOPAUSE -sDEVICE=pdfwrite -dDEVICEWIDTHPOINTS=270 -dDEVICEHEIGHTPOINTS=520 -sOutputFile=/Users/ljlamarche/Desktop/UAFThesis/Figures/superdarnmap.pdf /Users/ljlamarche/Desktop/UAFThesis/Figures/SuperDARNmap.ps
 
+
+
+pdflatex LLamarcheThesis.tex
+
+bibtex LLamarcheThesis
+
+bibtex introduction
+
+pdflatex LLamarcheThesis.tex
+
 pdflatex LLamarcheThesis.tex

### 2016-12-06 introduction.tex
References incorportated into introduction

@@
 In addition, the cusp then provides a gateway of highly ionized sunlit plasma into the dark polar cap.
-Several studies had previously reported density enhanced patches convecting from the dayside through the cusp/throat region toward the polar cap \citep{Kelley1984,Foster1984,Foster1985,Foster1993}.
+Several studies had previously reported density enhanced patches convecting from the dayside through the cusp/throat region toward the polar cap \citep{Kelly1984,Foster1984,Foster1985,Foster1993}.
 This also agrees with observations made by \citet{Sojka1982,delaBeaujardiere1985}  \citet{Rodger1996} also considered occurrence of patches but with a HF radar in Halley, Antarctica.

### 2016-12-10 introduction.tex
Fixed INPROCEEDINGS issue, added custom .bst
Minor detail corrections made to introduction

@@
 The solar terrestrial environment begins with the Sun in the center of the solar system.
-The Sun, like most stars, is composed of highly energized plasma that is gravitationally bound together.
-This makes it an extremely complex and highly dynamic system that is constantly interacting with the rest of the solar system.
-The surface of the Sun is populated with sunspots, dark, relatively cool regions with intense magnetic fields, as well as coronal holes, which are low density areas characterized with a continuous outflow of plasma.
+The Sun is composed of highly energized plasma that is gravitationally bound together.
+The surface of the Sun has sunspots, which are dark, relatively cool regions with intense magnetic fields, as well as coronal holes, which are low density areas characterized with a continuous outflow of plasma.
 The intense magnetic fields associated with sunspots can break down releasing a large amount of energy and plasma, known as a solar flare.
-Planet-sized globs of plasma know as coronal mass ejections (CMEs) can be ejected from the surface, which can be large enough to maintain an internal magnetic field as they travel through the solar system.
-The frequency of sunspots and large outbursts of plasma changes with the 22-year long solar cycle.
-A large number of Sun spots, high occurrence of fast, dense plasma outbursts, and a highly variably magnetic field indicate solar maximum.
-Conversely, solar minimum is characterized by a low Sun spot number and slow and relatively steady outflow.
+Planet-sized masses of plasma known as coronal mass ejections (CMEs) can be ejected from the surface, which can be large enough to maintain an internal magnetic field as they travel through the solar system.
+The frequency of sunspots and large outbursts of plasma changes with the 11-year long solar cycle.
+A large number of sunspots, high occurrence of fast, dense plasma outbursts, and a highly variably magnetic field indicate solar maximum.
+Conversely, solar minimum is characterized by a low sunspot number and slow and relatively steady outflow.
+This thesis examines factors that control irregularity production in the Earth's ionosphere, which can be strongly influenced by solar activity.
 
 The Sun interacts with the rest of the solar system through the solar wind, a continuous outflow of plasma that caries the Sun's magnetic field with it.
-The solar wind typically has a density around 10 cm\(^{-3}\) at Earth and is traveling away from the Sun at approximately 400 km/s.
+The solar wind typically has a density around 10 cm\(^{-3}\) and is traveling away from the Sun at approximately 400 km/s at Earth's orbit.
 However, like the Sun, the solar wind is highly dynamic and these parameters can change drastically over a period of minutes.
-CMEs, for instance, create a substantial increase in density and are often proceeded by a shock.
-In addition to the outflow of plasma, the solar wind also consisted of the Sun's magnetic field extended into the outer limits of the solar system, called the interplanetary magnetic field (IMF).
+CMEs, for instance, create a substantial increase in density and are often preceded by a shock, a region of hot, compressed plasma caused by the CME moving at supersonic speeds.
+In addition to the outflow of plasma, the solar wind also consists of the Sun's magnetic field extended into the outer limits of the solar system, called the interplanetary magnetic field (IMF).
 Because the Sun has a short rotation period (about 33 days), the IMF gets twisted into a spiral, commonly called the Parker Spiral \citep{Parker1958}.
@@
 
-The IMF interacts with the Earth's own magnetic field, the magnetosphere.
-Without outside influences, the magnetosphere would be approximately a dipole field, but dynamic pressure from the solar wind causes the Sun facing side to be compressed to 6--10 Earth radii and the rear side to be stretched into an extended tail (hundreds of Earth radii).
+The IMF interacts with the Earth's own magnetic field, and this interaction creates the magnetosphere.
+Without the IMF, the magnetosphere would be approximately a dipole field, but dynamic pressure from the solar wind causes the Sun facing side to be compressed to 6--10 Earth radii and the rear side to be stretched into an extended tail (hundreds of Earth radii).
 Because the solar wind is typically a supersonic flow, a bow shock forms on the dayside of the magnetosphere.
@@
 Closer to the Earth's surface, the magnetosphere also interacts with the Earth's ionosphere, a layer of partially ionized plasma in the upper atmosphere.
-The interaction between charged particles in the ionosphere with magnetic field lines from the magnetosphere couples these two regions and creates non-trivial dynamics.
-In addition, highly energized charged particles from the solar wind travel along open field lines and collide with both ions and neutral particles in the ionosphere, creating the aurora at high latitudes.
-Below the ionosphere, the thermosphere consists mostly of neutral particles, however neutral winds can influence the ionosphere through viscous interactions along the lower boundary.
+The interaction between charged particles in the ionosphere with magnetic field lines couples the ionosphere, magnetosphere, and solar wind to create non-trivial dynamics.
+Additionally,  neutral winds in the thermosphere can influence the ionosphere through viscous interactions.
 Overall, the charged particle interactions in the ionosphere create a highly complex system, which is further complicated by coupling with both the thermosphere from below and the magnetosphere and solar wind from above.
@@
 \label{sec:ionosphere}
-The Earth's ionosphere is a region of the upper atmosphere that ranges approximately between 50--1000 km in altitude where neutral gases have been excited so that they are partially ionoized, resulting in free electrons and ions mixed with neutral particles.
-Two conflicting processes occur in the ionosphere to create a peak in electron density, commonly referred to as the Chapman Layer \textbf{CTIE CHAPMAN?}.
-Solar illumination and particle precipitation ionize neutrals, and the density of ions increase as altitude decreases because there are more neturals available to ionize.
+The Earth's ionosphere is a region of the upper atmosphere that ranges approximately between 50--1000 km in altitude where neutral gases have been partially ionized, resulting in free electrons and ions mixed with neutral particles.
+Two competing processes occur in the ionosphere to create a peak in electron density.
+Solar illumination and particle precipitation ionize neutrals, and the density of ions increase as altitude decreases because there are more neutral molecules available to ionize.
 However, below a certain point, the concentration of ions becomes large enough that recombination into neutral particles becomes a substantal factor, reducing the ion density.
 The exact altitude where this density peak occurs is variable with time of day, location, season, and solar cycle.
-A profile of how electron density changes with altitude is shown in Figure \ref{fig:densprofile}, found from the International Reference Ionosphere (IRI) model run on January 1, 2007 at \(75\deg\) N, \(0\deg\) E, geographic (discussed further in Section \ref{sec:iri}).
+A profile of how electron density changes with altitude is shown in Figure \ref{fig:densprofile}, found from the International Reference Ionosphere (IRI) model run on January 1, 2007 at \(75\deg\) N, \(0\deg\) E, geographic (the IRI model is discussed further in Section \ref{sec:iri}).
 
@@
 Above 50 km, the primary neutral particles are atomic oxygen, O, molecular oxygen, \(O_2\), and molecular nitrogen, \(N_2\).
-Common ions at these altitudes are therefor mostly composed of oxygen and nitrogen, \(O^+, NO^+, O_2^+, N_2^+\), but hydrogen, \(H^+\), can contribute as well \citep{Kelley2009}.
+Common ions at these altitudes are therefore mostly composed of oxygen and nitrogen, \(O^+, NO^+, O_2^+, N_2^+\), but hydrogen, \(H^+\), can contribute as well, particularly at higher altitudes \citep{Kelley2009}.
 In the ionosphere, neutral densties are typically several orders of magnitude higher than ion densities, however even a small number of charged particles changes the behavior of the region dramatically.
 Because there are so many different ion species, plasma density is usually discussed simply in terms of electron density.
+Plasmas are assumed to be quasineutral, meaning that electron and ion densities are approximately the same so the plasma does not have an overall charge.
+The electron density is equal to the sum of densities of all the ions in the plasma, or \(n = n_e = \sum n_i\).
 
-\subsection{Ionosphere Regions}
+\subsection{Ionospheric Regions}
 \label{sec:ionosphere_regions}
-The ionosphere is typically divided into three region, Figure \ref{fig:densprofile}.
+The ionosphere is typically divided into three regions, Figure \ref{fig:densprofile}.
 The D region is located between 80 and 90 km in altitude and typically only exists in the daytime when the ionosphere is sunlit.
-The E region typically exists between 90 and 120 km with a density peak between 105 and 110 km, depending on factors such as time of day and season.
-Above 120 km is considered the F region, which has a density peak between 300 and 400 km.
+The E region typically exists between 90 and 130 km with a density peak between 105 and 110 km, depending on factors such as time of day and season.
+Above 130 km is considered to be the F region, which has a density peak around 250 km, Figure \ref{fig:densprofile} \citet{Luhmann1995}.
 The F region can actually have two peaks under some daytime conditions, notated as the F1 peak and the F2 peak.
-The locations and peak densities of each of these regions is highly variable both diurnally and seasonally.
+The peak heights and densities of each of these regions is highly variable both diurnally and seasonally.
 In Figure \ref{fig:densprofile}, the daytime profile is shown with the solid line next to a nighttime profile with a dashed line.
-Note that night electron densities are lower at all altitudes due to the lack of photoionization durring nighttimes.
+Note that night electron densities are lower at all altitudes due to the lack of photoionization during nighttimes.
 In addition, the F-region peak moves to higher altitudes at night.
+The focus of this thesis will be the F region.
 
-Because the concentrations of ions and neutral particles changes with altitude, each of these regions have distinct plasma physical properties that make them unique and change how plasma structuring forms.
-The main change that occurs is wether the the large-scale motion of a particle is controlled more by the magnetic field (the particle is magnetized) or by collisions with other particles (collisional).
-This can be determined by comparing the collision frequency, \(\nu_\alpha\), to the gyrofrequency, \(\Omega_\alpha\), of a particular species.
+Because the concentrations of ions and neutral particles changes with altitude, each of these regions have distinct plasma physical properties that make them unique.
+The main issue is whether the the large-scale motion of a particle is controlled more by the magnetic field (the particle is magnetized) or by collisions with other particles (the particle is collisional).
+This can be determined by comparing the collision frequency, \(\nu_\alpha\), to the gyrofrequency, \(\Omega_\alpha\), of a particular species, where \(\alpha\) corresponds to either electrons, \(e\), or ions, \(i\).
 If \(\Omega_\alpha \gg \nu_\alpha\), the particle completes many gyrations between each collision, so its motion is determined mostly by the magnetic field and it is magnetized.
 If \(\nu_\alpha \gg \Omega_\alpha\), the particle collides with other particles much more often than it complete a gyration, so it is considered collisional.
-The gyrofrquency is dependent on the mass and charge of the particle and the strength of the magnetic field, \(\Omega_\alpha = q_\alpha B/m_\alpha\), so is roughly constant through the ionosphere because the magnetic field strength does not change much through these altitudes.
-The collision frequency depends on the density of a species relative to the density of all other species in the plasma and can be calculated from a series of complicated expressions \citep{Schunk1980,Schunk2009}.
+The gyrofrequency is dependent on the mass and charge of the particle and the strength of the magnetic field, \(\Omega_\alpha = q_\alpha B/m_\alpha\), so is roughly constant through the ionosphere because the magnetic field strength does not change much throughout these altitudes.
+The collision frequency depends on the density of a species relative to the density of all other species in the plasma and can be calculated from a series of standard expressions \citep{Schunk1980,Schunk2009}.
 In the E region, \(\Omega_e \gg \nu_e\), so electrons are magnetized.
@@
 In the F region where the neutral density is much lower, both ions and electrons are magnetized, \(\Omega_\alpha \gg \nu_\alpha\).
-These differences in how charged particles move vary the conductivity and can strongly impact how plasma waves develop, which will be discussed further in Section \ref{sec:lit_theory}.
+These differences in how charged particles move impact the conductivity and plasma waves development, which is further discussed in Section \ref{sec:lit_theory}.
 
@@
 \label{sec:polar_structure}
-The polar ionosphere is a particularly interesting and dynamic region due to the presence of the magnetic poles and open field lines to the solar wind.
-Plasma structuring can occur as variations in density, velocity, or temperature, but the main focus of this work will be density perturbations, often called irregularities.
+This thesis is concerned with the polar ionosphere, which is a particularly interesting and dynamic region due to the presence of open field lines to the solar wind.
+The main focus of this work will be plasma density perturbations, often called irregularities.
 Density irregularities in the polar cap can occur on scales ranging from thousands of kilometers to less than a centimeter \citep{Tsunoda1988}.
-Large-scale plasma irregularities are typically considered to be any density structuring on the scale of 1 -- 1000 km.
+Large-scale plasma irregularities are typically considered to be any density structure on the scale of greater than 30 km \citep{Kelley2009}.
 Some common examples are polar patches, polar holes, and sun-aligned arcs.
-Polar patches are large density enhancements (usually at least twice the background plasma density) that travel across the polar cap with the background convection.
-Polar holes are plasma density depletions that typically occur in the F region slightly poleward of the auroral oval.
+Polar patches are large density enhancements (usually at least twice the background plasma density) that travel across the polar cap with the background convection \citep{Weber1984,Valladares1994}.
+Polar holes are plasma density depletions that typically occur in the F region slightly poleward of the auroral oval \citep{Benson2001}.
 Sun-aligned arcs are large density forms that stretch across the polar cap along the sun-earth line.
-They are very narrow and often characterized by large and complex velocity shears on either side of the arc.
+They are very narrow and often characterized by large and complex velocity shears on either side of the arc \citep{Valladares1991}.
 
-Intermediate-scale structuring typically ranges from 100 m -- 1 km.
-This is also referred to as scintillation-causing structuring and is responsible for many of the negative space weather effects that are observed on Earth.
-Radio signal scintillation through the ionosphere changes the phase and amplitude of the original signal, which can effect the ability that satelites have to communicate with ground recievers.
-This can cause communication blackouts or introduce large errors into GPS calculations, effecting navigation.
-Because communication and navigation technology is implemented in both every day civilian life as well as government and military intrests, the effects of this kind of signal scintillation can be far reaching.
+Intermediate-scale plasma irregularities are those ranges from 100 m -- 30 km \citep{Kelley2009}.
+These are also referred to as scintillation-causing structures and are responsible for many of the negative space weather effects that are observed on Earth.
+One important example of a space weather effect is when the ionosphere changes the phase and amplitude of the original signal, called radio scintillation, which affects the ability of satellites to communicate with ground receivers.
+This can cause communication blackouts or introduce large errors into GPS/GNSS (Global Positioning System/Global Navigation Satellite System) calculations, affecting navigation.
+Because communication and navigation technology is implemented in both every day civilian life as well as government and military operations, the effects of this kind of signal scintillation can be far reaching.
 
-Small-scale structuring is generally considered to be all irregularities less than 100 m.
-Although the effects of these scale sizes on navigation and communication systems are minimal, they can be the easiest to study because of large data sets collected from high frequency ionospheric radars globally for several decades (discussed further in Section \ref{sec:superdarn}).
-The irregularity sizes presented here are rough categories only and a high degree of coupling between different scales is generally considered reasonable.
-For instance, large-scale structures such as polar patches may very well have internal intermediate- or small-scale structuring.
-Some kind of turbulent cascade is often assumed to connect these different scales, but because various instruments or techniques usually only observe structuring on a particular scale, it is difficult to find direct evidence of this or how exactly it occurs.
+Small-scale structuring is generally considered to be all irregularities less than 100 m \citep{Kelley2009}.
+Although small-scale irregularities are not responsible for scintillation that directly impact GPS/GNSS performance, they can be the easiest to study because of large data sets collected from high frequency ionospheric radars globally for several decades (discussed further in Section \ref{sec:superdarn}).
+The irregularity classifications presented here are rough categories only and a high degree of coupling between different scales occurs, which makes plasma structures of all scales (large, intermediate, and small) important for plasma structuring processes.
+For instance, large-scale structures such as polar patches may very well have internal intermediate- and/or small-scale structuring.
+A turbulent cascade is often assumed to connect these different scales, but because various instruments or techniques usually only observe structuring on a particular scale, it is difficult to find direct evidence of this or how exactly it occurs \citep{Kinter1985,Tsunoda1985}.
+A description of some of the particular instruments and models that are used in this thesis follows.
 
@@
 \label{sec:csr}
-Coherent Scatter Radars (CSR) have been used to study the ionosphere for the last half century.
-The radar transmits a radio wave, which can reflect off plasma structuring in the ionosphere, and then be detected by the recieving antennas.
-The time difference between when the signal was transmitted and when it was detected and the power and phase of the returned signal can then be used to determine local characteristics of plasma irregularities.
+In this thesis, we make extensive use of Coherent Scatter Radars (CSR), which have been used to study the ionosphere for the last half century.
+The radar transmits a radio wave, which can scatter off plasma structures in the ionosphere, and then be detected by the recieving antennas.
+The time difference between when the signal was transmitted and when it was detected and the power and phase of the returned signal can then be used to determine characteristics of plasma irregularities.
 
-There are three measurements typically made by CSR systems: backscatter power, line of sight (LoS) velocity, and spectral width.
-Backscatter power is the power of the retuned signal that the radar's recievers detect.
+There are three measurements typically made by CSR systems: backscatter power, line-of-sight (LoS) velocity, and spectral width.
+Backscatter power is the power of the returned signal that the radar's recievers detect.
 Typically, a threshold is selected for the minimum power acceptable for a return to be considered an actual signal instead of background noise.
-LoS velocity is the component of the total plasma drift velocity that is measured along one beam of a radar.
-Because radars measure velocity through the doppler shift of backscatter, only the component of the velocity vector that is along the radar beam can be measured.
-Spectral width represents the width of the doppler power spectrum and can determine some of the characteristics of irregularities \citep{Greenwald1985}.
+LoS velocity is the component of the total plasma drift velocity that is measured along the transmitted signal direction.
+Because radars measure velocity through the doppler shift of backscatter, only the component of the velocity vector can be measured.
+Spectral width represents the width of the doppler power spectrum and can determine characteristics of irregularities \citep{Greenwald1985}.
 
-CSR systems typically operate by transmitting pulses instead of continuously.
+CSR systems can operate by transmitting pulses instead of continuously.
 This allows the distance between the radar and the target to be found (based on the time difference between when the pulse is transmitted and received).
-The size of the range gate is determined by the length of a radar pulse, Equation \ref{eqn:range_gate}.
+The size of the range gate is determined by the length of a radar pulse, Equation \ref{eqn:range_gate}.\
 \begin{equation}
@@
 	\Delta r = \frac{c\Delta t}{2}
-\end{equation}
-Longer pulses result in larger range gates, resulting in poorer spatial resolution.
+\end{equation}.
+Here, \(\Delta r\) is the size of a range gate, \(\Delta t\) is the pulse length, and \(c\) is the speed of light.  Longer pulses result in larger range gates, resulting in poorer spatial resolution.
 
 When the radar transmits continuously, backscatter will be received, but because it is impossible to know when that signal was transmitted, the difference in time between transmission and receiving cannot be used to calculate how far away the backscatter volume is.
-Conversly, if the radar only transmits a pulse of a certain length of a certain length and then "listens" for its return, the time between transmission and return is known and the distance the pulse traveled can be calculated and therefore the location of the backscattering volume.
+Conversly, if the radar only transmits a pulse of a certain length of a certain length and then "listens" for its return, the time between transmission and return is known and the distance the pulse traveled can be calculated and therefore the location of the backscattering volume \citep{Farley1972,Greenwald1983}.
 However, using this simple single-pulse method another pulse cannot be transmitted until the first pulse returns, limiting the temporal resolution that can be achieved.
-This can be improved by using a multipulse scheme, which will be discussed below.
+This can be improved by using a multipulse scheme, which will be discussed below \citep{Farley1972,Greenwald1983,Greenwald1985}.
 
-A short time between pulse returns is nessisary to allow dopler velocities to be calculated, particularly at long ranges.
-The maximum dopler shifted frequency that can be observed by a radar is the Nyquist frequency, which is equivilent to half the sampling frequency, Equation \ref{eqn:nyquist}.
+A short time between pulse returns is necessary to allow Doppler velocities to be calculated, particularly at long ranges.
+The maximum Doppler shifted frequency that can be observed by a radar is the Nyquist frequency, \(f_n\), which is equivilent to half the sampling frequency, \(f_s\), Equation \ref{eqn:nyquist}
 \begin{equation}	
@@
 	f_n = \frac{f_s}{2}
-\end{equation}
-Radars operating in single-pulse mode have a very limited sampiling frequency, as described above, particularly at long ranges where the pulse must travel a long distance.
-This severely limits the dopler shift that can be measured, which places an upper limit on the LoS dopler velocity that can be calculated with this technique.
+\end{equation}.
+Radars operating in single-pulse mode must have a very low sampling frequency, as described above, particularly at long ranges where the pulse must travel a long distance.
+This severely limits the Doppler shift that can be measured, which places an upper limit on the LoS Doppler velocity that can be calculated with this technique.
 For an ionospheric radar observing a structure 2000 km away, it would take about 13 ms for a signal traveling at the speed of light to travel to the the structure and back to the radar, wich corresponds to a sampling frequency of \(f_s = 75\) Hz.
@@
 In a multipulse mode, the radar transmits a series of pules with different time intervals or lags between them.
-This can introduce complications when backscatter from different pulses at different ranges is received by the radar at the same time, referred to as cross-range interference, but these effects can be mitigated using correlation techniques.
+This can introduce complications when backscatter from different pulses at different ranges is received by the radar at the same time, referred to as cross-range interference, but these effects can be mitigated using correlation techniques \citep{Farley1972}.
 The smallest lag between two pulses is known as the multi-pulse increment, \(\tau\).
@@
 The ACF is used to find the spectral characteristics of the backscatter, from which the LoS velocity, spectral width, and backscatter power can be obtained.
-Overall, multi-pulse techniques are generally considered far more appropriate for ionospheric studies than single pulse \citep{Farley1972,Greenwald1983,Greenwald1985,Barthes1998,Ponomarenko2006}.
+Overall, multi-pulse techniques are generally considered far more appropriate for ionospheric studies than single pulse techniques \citep{Farley1972,Greenwald1983,Greenwald1985,Barthes1998,Ponomarenko2006}.
 
-The first instance of CSR being used to study plasma density structures in the ionosphere was the Scandanavian Twin Auroral Radar Experiment (STARE) in the 1970s and 1980s.
-STARE consisted of two very-high frequency (VHF) radars with overlapping FoVs that were designed to measure FAIs in the E region of the ionosphere \citep{Greenwald1997}.
-The advantage of two radars with overlaping FoVs was the ability to observe the same structures from two different orientations.
+The first observations of the ionosphere by CSRs were made in the 1930s \citep{Eckersley1937,Harang1938}.
+Over the next several decades, numerous used ground-based radars to advance our understanding of the ionosphere and radio aurora \citep{Hultqvist1964,Leadabrand1965,LangeHesse1967,Unwin1972,Sahr1996}.
+In the 1970s and 1980s, the Scandanavian Twin Auroral Radar Experiment (STARE) produced the first continuous large data-set, similar to how modern CSR systems are run.
+STARE consisted of two very high frequency (VHF) radars with overlapping FoVs that were designed to measure FAIs in the E region of the ionosphere \citep{Greenwald1997}.
+The advantage of two radars with overlaping FoVs was the ability to observe the same structures from two different directions.
 Because each radar can only measure the LoS velocity of a structure, two simultanious observations from different directions allows two different velocity vecotor compontents to be found, and hense the total velocity vector can be calculated.
@@
 \label{sec:superdarn}
-The Super Dual Auroral Radar Network (SuperDARN) is a global network of HF CSRs that was designed to measure small-scale plasma structuring in the ionosphere and map the global plasma convection.
+The primary instruments used within this body of work are radars within the Super Dual Auroral Radar Network (SuperDARN), a global network of high frequncy (HF) CSRs that was designed to measure small-scale plasma structures in the ionosphere and map the global plasma convection.
 The network currently consists of about 35 operational radars between the northern and southern hemispheres distributed at mid-, high-, and polar latitudes, Figure \ref{fig:superdarnmap}.
-Following the STARE experiment \citep{Greenwald1978}, the first HF ionospheric radar was built in Goose Bay, Canada in 1983, which would become the first of the SuperDARN radars \citep{Greenwald1985}.
+After the STARE experiment \citep{Greenwald1978}, the first HF ionospheric radar was built in Goose Bay, Canada in 1983, which would become the first of the SuperDARN radars \citep{Greenwald1985}.
 
@@
 
-It is nessisary to use HF radars to study ionsospheric structures in the polar cap because the magnetic field is close to vertical and in order for the radar beam to meet the perpendicularity condition to observe FAIs, the beam must be refracted through a dense ionosphere.
-VHF radar beams are not refracted enough in the polar cap and the beam will pass through the ionosphere without reflecting off any FAIs.
-SuperDARN radars operate at a nominal 8--20 MHz frequency, which, according to the Bragg scatter equation for backscatter, corresponds to observing decameter-scale plasma waves.
-Each radar consists of 16 independent transmit and recieving antennas.
+It is necessary to use HF radars to study ionsospheric structures in the polar cap because the magnetic field is close to vertical and in order for the radar beam to meet the perpendicularity condition to observe FAIs, the beam must be refracted through a dense ionosphere.
+VHF radar beams are not refracted enough in the polar cap and the beam will pass through the ionosphere without FAIs producing backscatter.
+SuperDARN radars operate nominally between 8--20 MHz, which, according to the Bragg scatter equation for backscatter, corresponds to observing decameter-scale plasma waves.
+Each radar consists of 16 independent antennas that transmit a radio wave and receive backscatter from the ionosphere.
 Beams are electronically steerable and most radars have between 16 and 24 beams in normal operation mode, each beam being \(3.25\deg\) in azimuth.
@@
 The multi-pulse increment, \(\tau\), is 2400 \(\mu\)s, which corresponds to a Nyquist frequency of about 200 Hz, small enough to measure plasma drift velocities up to 3000 m/s.
-The spectral characteristics of backscatter of recovered from the ACF using the FITACF algorithm \citep{Ponomarenko2006}.
+The spectral characteristics of backscatter are derived from the ACF using the FITACF algorithm \citep{Ponomarenko2006}.
 
@@
 For convection maps, only data from the F region is considered.
-If there is not enough velocity data to confine the fit sufficiently, the measured data points are supplimented with a statistical model based on solar wind IMF conditions \citep{Ruohoniemi1998}.
+If there is not enough velocity data to confine the fit sufficiently, the measured data points are supplemented with a statistical model based on solar wind IMF conditions \citep{Ruohoniemi1995,Ruohonieme2005}.
 
@@
 In the studies presented here, data was primarily used from the SuperDARN radars located at Rankin Inlet, Canada (RKN) and McMurdo Station, Antarctica (MCM).
-The have been shaded in Figure \ref{fig:superdarnmap}.
+These two radars have been shaded in Figure \ref{fig:superdarnmap}.
+MCM is the only polar radar in the southern hemisphere.
+Although the FoVs of other radars cover parts of the polar cap, they only to at either E region ranges or the furthest of the F region ranges, for which there is rarely much backscatter.
+There are currently three polar radars in the northern polar cap, but the FoV of RKN overlaps the Resolute Bay Incoherent Scatter Radar FoV (discussed in Section \ref{sec:isr}, which makes it particularly useful for multi-instrument studies of the ionosphere.
 
@@
 Similar to CSRs, incoherent scatter radars (ISRs) are ground based radars that are used to probe the ionosphere.
-However, they use a fundamentally different technique to observe structuring, which allows them to measure parameters of the actual ionospheric plasma.
-Where as CSRs receive backscatter from relatively large, "coherent" density structures in the ionosphere, ISRs operate by radar beams scattering due to the random thermal motion of electrons in the plasma.
-The radar receivers then detect a spectra of frequencies with different power \citep{Gordon1958}.
+However, they use a fundamentally different technique to observe structuring, which allows them to measure bulk plasma parameters in contrast to small-scale irregularity characteristics.
+While CSRs receive backscatter from relatively large, "coherent" density structures in the ionosphere, ISRs operate by radar scattering from waves due to the random thermal motion of electrons in the plasma.
+The radar receivers then measure a spectrum of frequencies with different power \citep{Gordon1958}.
 
@@
 In addition, 3D convection velocities and electric field vectors can be derived from LoS ion velocity measurements in some of the more modern ISRs by assuming both \(\vec{E}\) and \(\vec{V}_E\) are constant across the radar's FoV \citep{Heinselman2008}.
+More recently, a new algorithm has been developed that used the Lagrange method of undetermined multipliers to generate 3D convection velocity and electric field vectors without making this limiting assumption \citep{Nicolls2014}.
 Each instrument must be calibrated with known electron density measurements to account for noise and system constants \citep{Nicolls2007a}.
@@
 Since then, at least 8 other ISRs have been deployed around the world at equitoral, mid-, and high-latitudes.
-The most recent advancement has been the development of Advanced Modular Incoherent Scatter Radars (AMISR), of which there are currently three in use, one at the Poker Flat Rocket Range, just north of Fairbanks, AK and two in Resolute Bay, Canada.
-The advantage of these new systems is that they are electronical steerable, unlike older ISRs which consisted of a large dish that had to be moved manually to change the beam direction.
+The most recent advancement has been the development of Advanced Modular Incoherent Scatter Radars (AMISR), of which there are currently three in use, one at the Poker Flat Rocket Range, just north of Fairbanks, Alaska and two in Resolute Bay, Canada.
+The advantage of these new systems is that they are electronical steerable, unlike older ISRs which consisted of a large dish that had to be physically moved to change the beam direction.
 This allows beams to be transmitted in many different directions at a very high time cadence, giving the radar the capability to make measurements in multiple directions almost simultaneously \citep{Nicolls2007a,Nicolls2007b,Bahcivan2010}.
 This is very important for creating 2D maps of ionospheric conditions and tracking how they change in time \citep{Semeter2009,Dahlgren2012a,Dahlgren2012b}.
+
 AMISR systems typically have two types of pulses, a long pulse (LP) with 72 km range gates for F-region studies and an alternating code pulse (AC) with 4.5 km range gates gates for E region studies.
@@
 One of the most important advantages of ISR systems over CSRs is the ability to directly measure electron density in the ionosphere.
-However, because ISRs are much scarcer and are only typically run on in WorldDay mode a few days out of every month, data is not as widely available.
-In addition, ISRs do not actually measure small-scale coherent plasma structures like the HF SuperDARN network does.
+However, because ISRs are much scarcer and are only typically run on in WorldDay mode a few days out of every month, data are not as widely available.
+In addition, ISRs do not actually observe small-scale coherent plasma structures like the SuperDARN HF network does.
 However, they can directly measure large-scale structures and image density variations on scales larger that 100 km (limited by the number of beams and range gate size).
 For these region, it is often useful use a combination of CSR and ISR measurements in ionospheric plasma structuring studies, when possible.
-The primary ISR used in the following studies is the north face of the AMISR system at Resolute Bay, Canada, RISR-N.
-The FoV of RISR-N overlaps that of the RKN SuperDARN radar, making it ideal for these types of comparison studies.
+The primary ISR used in the present volume of work is the north face of the AMISR system at Resolute Bay, Canada, RISR-N.
+The FoV of RISR-N overlaps that of the RKN SuperDARN radar, making it ideal for these types of comparison studies, Chapters \ref{sec:paper1} and \ref{sec:paper3}.
 
@@
 \label{sec:iri}
+In addition to radars, models can be a useful tool for understanding the background conditions in the ionosphere, particular in regions where there is poor instrumental coverage.
 The International Reference Ionosphere (IRI) is an empirical model of the Earth's ionosphere.
@@
 The current version is IRI-2012 \citep{Bilitza2014}, which is used throughout the work done here.
-It can be accessed either as an online module or by downloading and running the FORTRAN source code, both of which are available from the International Reference Ionosphere website (\url{http://iri.gsfc.nasa.gov/}).
 
-The IRI model can produce outputs of electron density, temperatures of electrons, ions, and neutrals, ion composition (O+, H+, He+, O2+, NO+, N+), and total electron content (TEC).
+The IRI model produces outputs of electron density, temperatures of electrons, ions, and neutrals, ion composition (O+, H+, He+, O2+, NO+, N+), and total electron content (TEC).
 The model requires the input of a location (latitude, longitude, and altitude) and time (year, date, and time).
@@
 There are a variety of other optional input parameters such as sunspot number, ionospheric index, and F10.7 radio flux which can help make the output parameters more accurate for a particular situation.
-Internally, the model has the additional capability of producing profiles of any of the output parameters in either space or time.
-This is particularly significant for height profiles, and allows the peak heights and densities o
-f the E and F region to be calculated.
+Additionally, the model is capable of producing profiles of any of the output parameters in either space or time.
+This is particularly significant for height profiles, and allows the peak heights and densities of the E and F region to be calculated.
 
-Although the IRI model is very useful for providing background plasma density conditions where no instruments are available to make real measurements, it is important to remember that it is a model and therefor only represents the average conditions.
+Although the IRI model is very useful for providing background plasma density conditions where no instruments are available to make measurements, it is important to bear in mind that it is a model and therefor only represents the average conditions.
 Although it is fairly accurate at mid-latitudes \citep{Coisson2006,Bilitza2012}, it does not necessarily represent variations within the polar cap well \citep{Themens2014,Makarevich2015}.
 In particular, the height of the F-region peak tends to be underestimated in daytime and the bottomside thickness does not show realistic seasonal and diurnal variation \citep{Themens2014}.
+
 Despite this, the IRI model is still a very important tool for establishing a density profile at any time in any location in the ionosphere.
@@
 One of the most important uses of the IRI model has been its integration into ray tracing simulations.
-It can be valuable to determine the degree to which HF radar beams refract through a dense ionosphere.
-This can be done through standard raytracing tools based on numerical solutions to the Hamiltonian raypath equations \citep{Haselgrove1963,Jones1975}, however these tools require a 2D density profile along the path that the radar beam travels.
-It is possible to assume a Gaussian distribution or Chapman layer as a simple model of the F-region peak, but it is much more instructive to use output from the IRI model, as it is sensitive to changes in latitude as well as seasonal and diurnal variations.
+Throughout this work, it is valuable to determine the degree to which HF radar beams refract through a dense ionosphere.
+This can be done through standard raytracing tools based on numerical solutions to the Hamiltonian raypath equations, as shown in Chapter \ref{sec:paper1} \citep{Haselgrove1963,Jones1975}, however these tools require a 2D density profile along the path that the radar beam travels.
+It is possible to assume a Gaussian distribution or Chapman layer as a simple model of the F-region peak, but it is much more instructive to use output from the IRI model, as is done in this work, as it is sensitive to changes in latitude as well as seasonal and diurnal variations.
 
@@
 
-\subsection{Theory and Simulations of Polar Patches and Plasma Instabilities}
-Large-scale structuring in the polar cap can take a variety of forms, Section \ref{sec:polar_structure}, but here we will focus on large regions of enhanced plasma density known as polar patches  \citep[][e.g.]{Weber1984,Weber1986,Buchau1983,Buchau1985}.
+\subsection{Polar Patch Formation}
+\label{sec:lit_patches}
+Large-scale structuring in the polar cap can take a variety of forms, Section \ref{sec:polar_structure}, but here we will focus on large regions of enhanced plasma density known as polar patches  \citep[e.g.][]{Weber1984,Weber1986,Buchau1983,Buchau1985}.
 The first observations of polar patches were made in the 1960s by \citet{Hill1963}.
-They can range in size up to 1500 km in diameter and can have peak densities as much as eight times that of the background ionospheric plasma \citep{Weber1986,Hosokawa2014}.
+They can range in size from 100 km up to 1500 km in diameter and can have peak densities as much as eight times that of the background ionospheric plasma \citep{Weber1986,Hosokawa2014}.
 Polar patches are know to emerge on the dayside and then propagate with the background plasma convection across the polar cap to the nightside, where they either disintegrate or recombine with enhanced-density plasma in the auroral oval \citep{Weber1985,Weber1986}.
 
-There have been at least 6 mechanisms proposed for how polar patches form on the dayside \citep{Crowley1996,Carlson2012}.
+There have been at least 6 mechanisms proposed for how polar patches form on the dayside \citep{Crowley1996,Carlson2012}:
 \begin{enumerate}
@@
 	\item Alfven wave coupling \citep{Prikryl1999}
-\end{enumerate}
+\end{enumerate}.
 Although particle precipitation has been shown to create enhanced plasma density in the cusp region \citep{Rodger1994}, it is unreasonable for this mechanism to account for the largest density enhancements observed in polar patches.
-These high densities must originate from the resivor of solar illuminated plasma on the dayside.
+These high densities must originate from the reservoir of solar illuminated plasma on the dayside.
 Additionally, a mechanism is still required to separate patches from the cusp region.
 High speed flow channels can ``cut'' segments off of a tongue of high-density plasma that has been pulled into the polar cap by convection \citep{Valladares1994,Valladares1998}.
-Models have also shown changing large scale convection patterns quickly in time does produce density structures similar to patches \citep{Anderson1988}.
+Models have also shown that changing large scale convection patterns quickly in time does produce density structures similar to patches \citep{Anderson1988}.
 However, the mechanism that seems to be dominant for the majority of polar patches observed is transient magnetic reconnection \citep{Carlson2012}.
@@
 Figure \ref{fig:magnetosphere} is a schematic of how the IMF, magnetosphere, and ionospheric polar cap are all interconnected.
-In general, magnetic field lines in the magnetosphere can either be closed (grey lines in Figure \ref{fig:magnetosphere}), meaning they connect only the north and south pole of the earth, or open (light blue lines in Figure \ref{fig:magnetosphere}), when one of the ends of the field lines originates from the sun such that the field line is connected directly to the IMF.
-IMF lines that are not connected to the earth are shown in dark blue in figure \ref{fig:magnetosphere}.
+In general, magnetic field lines in the magnetosphere can either be closed (grey lines in Figure \ref{fig:magnetosphere}), meaning they connect only the north and south pole of the Earth, or open (light blue lines in Figure \ref{fig:magnetosphere}), when one of the ends of the field lines originates from the sun such that the field line is connected directly to the IMF.
+IMF lines that are not connected to the Earth are shown in dark blue in figure \ref{fig:magnetosphere}.
 The polar cap boundary is generally defined as the border between open and closed magnetic field lines, shown by the green line in Figure \ref{fig:magnetosphere}.
@@
 The X-line can be mapped back along magnetic field lines (orange) to the ionosphere, where their projection is the merging gap (pink dotted line).
-Durring reconnection events, the open-closed boundary expands equatorwards as more open field lines are formed and there is plasma flow across the merging gap.}
+During reconnection events, the open-closed boundary expands equatorwards as more open field lines are formed and there is plasma flow across the merging gap.}
 	\label{fig:magnetosphere}
@@
 
-Magnetic reconnection between the IMF and the Earth's magnetosphere occurs along the X-line on the dayside magnetopause (red line in figure \ref{fig:magnetosphere}).
-The X-line in the magnetopuase can be mapped along the magnetic field lines (orange) to the ionosphere, where it is referred to as the merging gap (pink dotted line in figure \ref{fig:magnetosphere}) and lies along the steady state polar cap boundary.
+Magnetic reconnection between the IMF and the Earth's magnetosphere occurs along the X-line on the dayside magnetopause (red line in Figure \ref{fig:magnetosphere}).
+The X-line in the magnetopause can be mapped along the magnetic field lines (orange) to the ionosphere, where it is referred to as the merging gap (pink dotted line in figure \ref{fig:magnetosphere}) and lies along the steady state polar cap boundary.
 
@@
 In addition, the potential along the X-line also gets mapped to the merging gap, assuming there is minimal drop in potential along the magnetic field lines \citep{Lockwood1992b}.
-The combination of these two effects results in flux being transferred across the boundary at a rate equal to the applied voltage  along the X-line and hence plasma flow across the merging gap \citep{Lockwood1992b}, as seen in Figure \ref{fig:patch_formation}.
+The combination of these two effects results in flux being transferred across the boundary at a rate equal to the applied voltage  along the X-line and hence plasma flow across the merging gap \citep{Lockwood1992b}, which will be examined further in Figure \ref{fig:patch_formation}.
 
-Figure \ref{fig:patch_formation} is a schematic of how transient magnetopause reconnection forms polar patches.
-In all panels, the the dayside is the top of the figure and the green line shows the open-closed polar cap boundary, similar to Figure \ref{fig:magnetosphere}.
+Figure \ref{fig:patch_formation} is a schematic of polar patches being formed through transient magnetopause reconnection.
+In all panels, the dayside is the top of the figure and the green line shows the open-closed polar cap boundary, similar to Figure \ref{fig:magnetosphere}.
 Plasma flow contours are shown in black and the light orange region represents high density from the dayside reservoir.
@@
 	\includegraphics[width=0.75\textwidth,angle=270]{patch_formation.pdf}
-	\caption{Diagram of the polar patch formation through transient magnetopause reconnection.
+	\caption{Polar patch formation through transient magnetopause reconnection.
 The open-closed boundary is the green line.
 Black contours represent plasma flow.
-Highly ionized (dense) daytime plasma is shown by the light orange regions.
+Dense daytime plasma is shown by the orange regions.
 The burst of reconnection is occurring in Figures \ref{fig:patch_formation}a--\ref{fig:patch_formation}c but stops in Figures \ref{fig:patch_formation}d--\ref{fig:patch_formation}f.}
@@
 
+\subsection{Theory of Plasma Instabilities}
+\label{sec:lit_instabilities}
 Plasma structuring, especially at smaller scales, is often discussed in terms of the growth or damping of plasma waves, or instabilities.
 As discussed previously in Section \ref{sec:ionosphere_regions}, the plasma characteristics change with altitude in the ionosphere, which causes a variety of different instability mechanisms to be operational depending on the region considered.
-The Farley-Buneman instability (FBI), or the modified two-stream instability tends to be a major factor in the E region where ion inertial effects are high \citep{Farley1963,Buneman1963}.
+The Farley-Buneman instability (FBI), or the modified two-stream instability tends to be a major factor in the E region where ion inertial effects are strong \citep{Farley1963,Buneman1963}.
 Also operational in the E region but more dominant at higher altitudes is the gradient-drift instability (GDI) \citep{Simon1963,Hoh1963,Linson1970}.
-If a wave has a component along the magnetic filed line, the current-conductive instability (CCI) can be operational even if GDI is stable.
-Addition instabilities mechanisms emerge if shears are considered, when \(\nabla E \neq 0\), such as the Kelvin-Helmholtz instability (KHI).
+If a wave has a component along the magnetic filed line, the current-conductive instability (CCI) can be operational even if GDI is not \citep{Hoh1960,Ossakow1979,Chaturvedi1981}.
+Additionally instability mechanisms emerge if shears are considered, when \(\nabla E \neq 0\), such as the Kelvin-Helmholtz instability (KHI) \citep{Kintner1977,DAngelo1965}.
 
@@
 GDI is operational when high-density purturbations in the plasma move to regions of lower background density and low-density purturbations in the plasma move to regions of higher background density, such that the wave amplitude grows relative to background conditions.
-As plasma drifts and carries irregularities with it, the ions in high density perturbations lag behind the electrons and create a purturbed electric field.
-This purturbed electric field combined with the background magnetic field causes the density perturbations to \(\vec{E}\times\vec{B}\) drift.
+As plasma drifts and carries irregularities with it, electrons and ions have slightly different drifts and create a perturbed electric field.
+This perturbed electric field combined with the background magnetic field causes the density perturbations to \(\vec{E}\times\vec{B}\) drift.
 If the drift causes high and low density perturbations to move as desribed above, the instability is operational and the perturbations grow.
@@
 
-Some of the first 1D investigations of the linear GDI growth rate found that it can be described by a simple expression involving the plasma drift velocity, \(V_E = \vec{E}\times\vec{B}\), and the gradient scale length, \(L = |n/(\partial n/\partial x)|\), Equation \ref{eqn:gdi_old} \citep{Simon1963,Hoh1963,Linson1970}.
+Some of the first 1D investigations of the linear GDI growth rate found that it can be described by a simple expression involving the plasma drift speed, \(V_E = |\vec{E}\times\vec{B}|/B^2\), and the gradient scale length, \(L = |n/(\partial n/\partial x)|\), Equation \ref{eqn:gdi_old} \citep{Simon1963,Hoh1963,Linson1970}
 \begin{equation}
@@
 	\gamma = \frac{V_E}{L}
-\end{equation}
-This expression is only valid if the density gradient is in the same direction as the plasma drift velocity and the wave is propagating perpendicular to both.
-The results of \citet{Keskinen1982,Keskinen1983}, Equation \ref{eqn:gdi_80s}, do consider an arbitrary wavevector, \(\vec{k}\), explicitly to some extent, but still assume a general relationship between density gradients and plasma drift \citep{Tsunoda1988}.
+\end{equation}.
+This expression is technically only valid if the density gradient, \(\nabla n\), is in the same direction as the plasma drift velocity, \(\vec{V}_E\), and the wave is propagating perpendicular to both.
+A more general expression that is used less often, Equation \ref{eqn:gdi_80s}, does consider an arbitrary wavevector, \(\vec{k}\), but still assumes density gradients parallel to the plasma drift \citep{Tsunoda1988}
 \begin{equation}
@@
 	\gamma = \frac{k_y}{k}\frac{\vec{k}\cdot\vec{V}_E}{kL}
-\end{equation}
-The directional dependance of GDI has been shown to be significant, such that the growth rate can actually change signs (determining whether wave growth or damping occurs) with different wavevector directions, so it is important to consider this factor \citep{Makarevich2014c}.
-In addition, these expressions have historically only applicable in a particular altitudinal regieme, usually the F region.
+\end{equation}.
+An even more general consideration involves arbitrary gradient and plasma drift directions \citep{Kesknien1982,Makarevich2014c}.
+In this more general case, the directional dependence of GDI has been shown to be significant, such that the growth rate can actually change signs (determining whether wave growth or damping occurs) with different wavevector directions, so it is important to consider this factor \citep{Makarevich2014c}.
+In addition, these expressions have historically only applicable in a particular altitudinal regime, usually the F region.
 
-Recently, a new expression has been introduced that considers any arbitrary wavevector, density gradient, and altitude within the ionosphere, which allows FBI, GDI, an CCI to be considered with a single dispersion relation \citep{Makarevich2016a}.
+Recently, a new expression has been introduced that considers any arbitrary wavevector, density gradient, and altitude within the ionosphere, which allows three electrostatic instabilities, FBI, GDI, and CCI, to be considered with a single dispersion relation \citep{Makarevich2016a}.
 \begin{equation}
@@
 \end{equation}
-As before, \(\vec{k}\) is the wavevector, \(\omega\) is the wave frequency, and \(\vec{V}_{\alpha 0}\) is the velocity of species \(\alpha\).
-The species \(\alpha\) is given by either \(i\) for ions or \(e\) for electrons.
+As before, \(\vec{k}\) is the wavevector, \(\omega\) is the wave frequency, and \(\vec{V}_{\alpha 0}\) is the background velocity of species \(\alpha\).
 The other quantities in equation \ref{eqn:gdi_Mak16} are not standard and described below.
@@
 \end{equation}
-Recognizing that \(s_es_i\left(1+R^2\right) = (1+\psi)^{-2}\) and \(b = \vec{G}\cdot\vec{k}\times\uvec{b}/k^2 = -\vec{k}\cdot\uvec{b}\times\vec{G}/k^2\) by vector identities, a final expression for the GDI growth rate in the F region can be found, which agrees well with the results of \citet{Makarevich2014c}.
+Recognizing that \(s_es_i\left(1+R^2\right) = (1+\psi)^{-2}\) and \(b = \vec{G}\cdot\vec{k}\times\uvec{b}/k^2 = -\vec{k}\cdot\uvec{b}\times\vec{G}/k^2\) by vector identities, a final expression for the GDI growth rate in the F region can be found, which agrees well with the results of \citet{Makarevich2014c}
 \begin{equation}
@@
 	\gamma = \frac{1}{1+\psi}\left(\uvec{k}\cdot\uvec{b}\times\vec{G}\right)\uvec{k}\cdot\left(\frac{\vec{E}}{B}-R\vec{V}_E\right)
-\end{equation}
+\end{equation}.
+Equation \ref{eqn:gdi} will be critical to the analysis done both in Chapter \ref{sec:paper2} and Chapter \ref{sec:paper3} in this thesis.
 
-In addition to analytical analysis of dispersion relations, GDI has also been studied through numerical simulations.
-\citet{Keskinen1982} found plasma waves on the trailing edge of a large density enhancement are unstable.
+%In addition to analytical analysis of dispersion relations, GDI has also been studied through numerical simulations.
+The advantage of simulations is that they can often be used to study phenomena at better time and spatial resolution than can be achieved by instruments and the senario of interest can be specified exactly.
+This is particularly useful for considering the degree of anisotropy that exists in plasma irregularities.
+
+Although linear GDI theory as described predicts clear dependence on wave vector direction \citep{Makarevich2014c}, this theory is only really valid in the long-wavelength limit, \(\vec{k}\cdot\vec{V}_E \ll \nu_\alpha\).
+In the F region, the ion-neutral collision frequency, \(\nu_i\), is about 10 Hz and if the drift speed, \(V_E\), is about 1000 m/s, fluid theory only applies for wavelengths greater than 100 m, larger than the decameter-scale waves that are detected by SuperDARN.
+This raises the important question of whether or not the same anisotropy that is predicted by linear GDI theory at larger scales can be expected in small-scale irregularities where nonlinear processes may be important.
+This issue has mostly been studied through numerical simulations due to the challenges associated with measuring a high-resolution 2D spectrum experimentally.
+
+The first simulations of GDI were done in the 1970s and successfully reproduced structures developing on the trailing edge of large density gradients \citep{Zabusky1973,Doles1976,Scannapieco1976,Ossakow1975,Ossakow1977}.
+Later, 2D simulations were expanded to include magnetosphere-ionosphere coupling, but it was found that this only had a substantial effect for very large-scale structures \citep{Keskinen1990}.
+3D simulations were introduced of GDI surrounding a large plasma density enhancement \citep{Guzdar1998} and were later improved by including plasma dynamics parallel to the magnetic field and inertial effects so that secondary KHI and tertiary shear-driven instability processes could be modeled \citep{Gondarenko1999}.
+These simulations consistently showed asymmetry between the leading and trailing edge of large-scale density structures and fluctuations in plasma density that were as much as 10--20\% of the background plasma \citep{Gondarenko2004a,Gondarenko2004b}.
+Plasma structuring can occur on the leading edge of large-scale density enhancements if the convection velocity changed rapidly such that the trailing edge becomes the leading edge and vice versa, or if there are large velocity shears initially present \citep{Gondarenko2004a,Gondarenko2004b}.
+
+For the most part, the results of nonlinear numerical simulations very closely match expectations from linear GDI theory.
+The spatial power spectra are anisotropic with most of the power concentrated in the direction where linear GDI growth is most favorable \citep{Keskinen1981,Keskinen1982,Gondarenko2001,Gondarenko2004b}.
+However, very strong shears or increasing the impact of ion-inertial factors can cause isotropy \citep{Gondarenko2001,Gondarenko2006}.
+
+%/  Additionally, 2D power spectrum could be calculated and initially considered plasma waves along a large density structure but were later expanded to include magnetosphere-ionosphere coupling \citep{Keskinen1981,Keskinen1982,Keskinen1990}.
+
+%\citet{Keskinen1982} found plasma waves on the trailing edge of a large density enhancement are unstable.
 \citet{Keskinen1990} expanded these results by considering coupling between the magnetosphere and ionosphere, but found that this coupling only has a substantial effect for very large-scale structuring.
@@
 \citet{Gondarenko1999} improved this 3D simulation further by including plasma dynamics parallel to the magnetic field and inertial effects.
-These factors are nessesary for the simulation to model secondary KHI and tertiary sheer-driven instability processes.
+These factors are necessary for the simulation to model secondary KHI and tertiary shear-driven instability processes.
 These simulations provided insight in how exactly GDI structuring develops.
-Both the density and velocity spectra were found to be anisotropic, however increasing the impact of ion-inertial factors in simulations caused both spectra to become more isotropic \citep{Gondarenko2001}.
-Fluctuations in density were found that were as much at 10--20\% of the background plasma density \citep{Gondarenko2004a}.
+
+%Both the density and velocity spectra were found to be anisotropic, however increasing the impact of ion-inertial factors in simulations caused both spectra to become more isotropic \citep{Gondarenko2001}.
+
+%Fluctuations in density were found that were as much at 10--20\% of the background plasma density \citep{Gondarenko2004a}.
 Although asymmetry between structuring on the leading and trailing edges of a large-scale density structure was observed in all simulations, rapidly changing convection velocity could complicate this \citep{Gondarenko2004b}.
@@
 \subsection{Observations of Polar Cap Structuring}
-
+\label{sec:lit_observations}
 Polar patches were first observed in the 1960s \citep{Hill1963}, but much of the work of characterizing them using optical and radio techniques was not accomplished until two decades later \citep{Weber1981,Weber1984,Weber1986,Buchau1983,Buchau1985}.
@@
 
-\bibliographystyle{plainnat}
+\bibliographystyle{uafthesis}
 \bibliography{references}

### 2017-01-03 introduction.tex
Rewrote observations of polar structuring section

@@
 \label{sec:introduction}
+
+The majority of mass in the universe exists as plasmas.
+Plasmas are fluid masses of charged particles, often formed when neutral atoms and molecules in a gas are ionized by external energy.
+Everyday examples of this is florecent lightbulbs, as well as flames.
+More exotic instances of plasma include nuclear fusion reactor experiments, most stars, nebula, and the interstellar medium.
+In fact, about 99\% of matter in the known universe exists in the plasma state.
+Of particular interest to this thesis is the plasma in Earth's ionosphere.
+
+Plasmas can be either fully ionized, where most particles in the plasma are ionized, or partially ionized, where only some fraction are ionized and the rest are neutral.
+Considering the dynamics of plasmas therefore requires not only collisions between particles to be considered but also charge-charge interactions and external electric and magnetic fields.
+These dynamics can generate waves in the plasma, which cause density perturbations.
+Additionally, variations in temperature and electric and magnetic field strength can occur on a variety of scales.
+This thesis focuses on mechanisms by which density variations in ionospheric plasma occur, often referred to as plasma structuring.
 
@@
 Large-scale structuring in the polar cap can take a variety of forms, Section \ref{sec:polar_structure}, but here we will focus on large regions of enhanced plasma density known as polar patches  \citep[e.g.][]{Weber1984,Weber1986,Buchau1983,Buchau1985}.
+One of the primary interests of this thesis is the relationships between irregularities of different scales and small-scale structuring has long been observed around polar patches \citep{Weber1984,Milan2002,Moen2012}.
+Although small-scale structures have also been observed around sun-aligned arcs \citep[e.g.][]{Koustov2012}, these structures also tend to be more complicated and often are surrounded by intense shears and electric fields \citep{Safargaleev2000,Aikio2002,Kozlovsky2007}.
+Additionally, polar patches occur relatively frequently, providing a good opportunity to study irregularity coupling at different scales \citep{Rodger1996}.
+In theory, many of the results presented in this work pertaining to the strong gradients on the edges of patches could also easily be applied to the edges of polar holes, but to date, far less attention has been paid to polar holes in the literature, making it challenging to accurately put observations in context \citep{Makarevich2015}.
+
 The first observations of polar patches were made in the 1960s by \citet{Hill1963}.
@@
 However, very strong shears or increasing the impact of ion-inertial factors can cause isotropy \citep{Gondarenko2001,Gondarenko2006}.
+The power spectra derived from these simulations are similar to those from experimental observations \citep{Baker1978,Kelley1979}, however ...
+\textbf{WHAT'S DIFFERENT AND IMPORTANT ABOUT PROJECT 3 HERE?}
 
@@
 
-%\citet{Keskinen1982} found plasma waves on the trailing edge of a large density enhancement are unstable.
+%\citet{Keskinen1982} found plasma waves on the trailing edge of at large density enhancement are unstable.
 \citet{Keskinen1990} expanded these results by considering coupling between the magnetosphere and ionosphere, but found that this coupling only has a substantial effect for very large-scale structuring.
@@
 Polar patches were first observed in the 1960s \citep{Hill1963}, but much of the work of characterizing them using optical and radio techniques was not accomplished until two decades later \citep{Weber1981,Weber1984,Weber1986,Buchau1983,Buchau1985}.
+Polar patches are density enhancements of as much as \(10^6\) cm\(^{-3}\) that occur primarially in the polar F region \citep{Buchau1983}.
+There were initially catagorized as extending 800--1000 km horizontally \citep{Weber1984}, but more recent studies have identified patches as large as 1500 km \citep{Hosokawa2014}.
+Patches tend to drift antisunwards at speeds of 500--1000 m/s, following background convection \citep{Buchau1983,Weber1984}.
+
+%\citet{Coley1998} examined the occurrence of polar patches in both the northern and southern hemispheres with DE-2 and found that
+
+Polar patch occurrence varies both diurnally and seasonally, with a diurnal peak at magnetic noon and a seasonal peak during equinox months \citep{Rodger1996}.
+Additionally, there is a bias towards patches occurring when there is a negative IMF \(B_z\) component \citep{Buchau1983,Rodger1996}.
+The frequency of patch occurrence is greatest when the cusp is slightly dayward of the terminator.
+This creates the situation of an unsunlit polar cap, so background plasma density is low and blobs of enhanced density are easily visible \citep{Coley1998}.
+In addition, the cusp then provides a gateway of highly ionized sunlit plasma into the dark polar cap.
+Patches convect from the dayside through the cusp/throat region towards the polar cap \citep{Kelly1984,Foster1984,Foster1985,Foster1993,Sojka1982,delaBeaujardiere1985}.
+
+%Several studies had previously reported density enhanced patches convecting from the dayside through the cusp/throat region toward the polar cap \citep{Kelly1984,Foster1984,Foster1985,Foster1993}.
+This also agrees with observations made by \citet{Sojka1982,delaBeaujardiere1985}
+
+%\citet{Rodger1996} also considered occurrence of patches but with a HF radar in Halley, Antarctica.
+The diurnal variation in patch occurrence was found to peak at magnetic noon and the seasonal variation in equinox months.
+The only correlation between hourly averaged solar wind parameters that was found was a bias towards patches occurring with a negative IMF \(B_z\) component.
+
+Over time, the trailing edge of a patch became steeper than the leading edge and irregularities developed on the trailing edge.
+
 \citet{Weber1981} identified sun-aligned arcs that extend  along the sun-earth line through large parts of the polar cap using all-sky imaging photometers (ASIP).
 These arcs are over 1000 km long and \(\sim100\) km wide and are produced by soft electron precipitation.
-\citet{Buchau1983} used a combinations of Digisonde and ASIP measurements to identify large luminous patches in the polar cap in addition to sun-aligned arcs.
+
+Additionally, \citet{Buchau1983} found that polar patches were more likely to be observed for southward IMF \(B_z\) while sun-aligned arcs were more common for northwards \(B_z\).
+
+%\citet{Buchau1983} used a combinations of Digisonde and ASIP measurements to identify large luminous patches in the polar cap in addition to sun-aligned arcs.
 These patches could have densities up to \(10^6\) cm\(^{-3}\) and tended to drift antisunward.
-Additionally, \citet{Buchau1983} found that polar patches were more likely to be observed for southward IMF \(B_z\) while sun-aligned arcs were more common for northwards \(B_z\).
-\citet{Weber1984} found patches of enhanced ionization between 800--1000 km during moderately disturbed geomagnetic conditions, again using a combination of ASIP images, ionosonde measurements, and data from Dynamic Explorer 2 (DE-2).
-These findings confirmed that density enhancements tended to drift antisunward at speeds of 500-1000 m/s and patches were not colocated with enhanced particle precipitation within the polar cap, indicating that this is not the source of the density enhancement.
+
+%\citet{Weber1984} found patches of enhanced ionization between 800--1000 km during moderately disturbed geomagnetic conditions, again using a combination of ASIP images, ionosonde measurements, and data from Dynamic Explorer 2 (DE-2).
+These findings confirmed that density enhancements tended to drift antisunward at speeds of 500-1000 m/s and patches were not colocated with enhanced particle within the polar cap, indicating that this is not the source of the density enhancement.
 Over time, the trailing edge of a patch became steeper than the leading edge and irregularities developed on the trailing edge.
+
+Patches are not colocated with enhanced particle precipitation within the polar cap, indicating that this is not the source of the density enhancement \citep{Weber1984}.
 Conversely, \citet{Buchau1985} found that plasma densities within patches were comparable with those measured at dayside sub-auroral latitudes, suggesting that dayside solar-illuminated plasma is the source of dense plasma that form polar patches.
-
-\citet{Coley1998} examined the occurrence of polar patches in both the northern and southern hemispheres with DE-2 and found that the frequency of patch occurrence was greatest when  the cusp was slightly dayward of the terminator.
-This creates the situation of an unsunlit polar cap, so background plasma density is low and blobs of enhanced density are easily visible.
-In addition, the cusp then provides a gateway of highly ionized sunlit plasma into the dark polar cap.
-Several studies had previously reported density enhanced patches convecting from the dayside through the cusp/throat region toward the polar cap \citep{Kelly1984,Foster1984,Foster1985,Foster1993}.
-This also agrees with observations made by \citet{Sojka1982,delaBeaujardiere1985}  \citet{Rodger1996} also considered occurrence of patches but with a HF radar in Halley, Antarctica.
-The diurnal variation in patch occurrence was found to peak at magnetic noon and the seasonal variation in equinox months.
-The only correlation between hourly averaged solar wind parameters that was found was a bias towards patches occurring with a negative IMF \(B_z\) component.
-A review of historical observations of polar patches has been provided by \citet{Crowley1996}
 
@@
 Experiments run in the 1980s using the European Incoherent Scatter Scientific Association (EISCAT) Troms\o, Norway ISR showed that plasma flow occurred near noon shortly after southward IMF \(B_z\) arrives at the magnetopause \citep{Etemadi1988,Todd1988}.
-Later, \citet{Lockwood1993a,Lockwood1993b} showed that these plasma flows are pulsed in the region of the polar cap boundary, as predicted by \citet{Cowley1991}.
+These plasma flows are pulsed in the region of the polar cap boundary \citep{Lockwood1993a,Lockwood1993b}, as predicted by \citet{Cowley1991}.
 Observations made by an ASIP in Svalbard in 1984 are fully consistent with patch formation through transient magnetopause reconnection \citep{Carlson1996,Carlson2002}.
@@
 All five signatures were observed as expected, strengthening the idea that transient magnetopause reconnection is responsible for patch formation.
-\citet{Carlson2006} furthered this using a observations from the EISCAT Svalbard Radar, the EISCAT VHF radar in Troms\o an ASIP, and a meridian scanning photometer (MSP).
-They managed to identify and track a series of patches directly from the subauroral plasma reservoir and show that the boundary moves equatorward before relaxing poleward, one of the main predictions of the transient magnetopause reconnection mechanism \citep{Lockwood1992b}.
+Furthermore, \citet{Carlson2006} identified and tracked a series of patches directly from the subauroral plasma reservoir and show that the boundary moves equatorward before relaxing poleward, one of the main predictions of the transient magnetopause reconnection mechanism \citep{Lockwood1992b}.
+A review of historical observations of polar patches has been provided by \citet{Crowley1996}.
 
@@
 
-This asymmetry was first observed by \citet{Weber1984}.
+Over time, the trailing edge of a patch became steeper than the leading edge and irregularities developed on the trailing edge \citep{Weber1984}.
 Backscatter power from HF radars tends to be greater on the trailing edges of both moving high-density polar structures \citep{Milan2002b} and sun-aligned arcs identified from ASIP data \citep{Koustov2012}, which is generally attributed to GDI being unstable on the trailing edge but stable on the leading edge of such large, moving density structures.
@@
 This study provided evidence that decameter scale plasma structuring (the same as observed by HF radars) had spawned from kilometer scale density gradients (that observed by ASIP and ASI methods) and structuring was greatest where GDI was operational.
+
+Small-scale plasma irregularities are typically studied with HF coherent scatter radars in the polar cap, Section \ref{sec:csr}.
+The factors that affect irregularity observation can roughly be divided into two groups, irregularity production and radar propagation.
+Irregularity production factors are related to the existance of plasma irregularities in the ionosphere while radar propagation factors are related to the abililty to observe irregularities with a ground based radar, Section \ref{sec:superdarn}.
+Within the polar caps, plasma irregularity production is much higher than at lower altitudes such that when background plasma conditions are favorable for radar propagation, FAIs are observed nearly continuously \citep{Bristow2011}.
+However, there are still definite diurnal, seasonal, and solar cycle trends \citep{Kane2012}.
+There is an increase in echo occurrence at solar maximum, particularly in the midnight sector \citep{Milan1997,Koustov2004}.
+Seasonally, high latitude radars observe a peak in echo occurrence durring equinox months \citep{Koustov2004}.
+Additionally, backscatter tends to occure at closer ranges in summer at midnight, but moves further away from the radar in other seasons at midday \citep{Milan1997}.
+In general, echo occurrence in high durring the day and low at night, but nighttime occurrence is directly proportional to the F peak density and daytime occurrence can be significantly reduced by a dense and conductive E layer.
+\citep{Koustov2004,Kane2012,Vickrey1982}.
+In the daytime, solar illumination causes the electron density of the E region to increase dramatically, making it highly conductive.
+This conductive layer can provide a conduit by which the potential differences associated with FAIs in the F region can be shorted out, effectively reducing F-region plasma irregularities \citep{Vickrey1982}.
+Both irreguarity production and radar propagation factors probably contribute to these regular variations of echo occurrence, because while a dense, daytime ionosphere creates favorable radar propigation conditions, it can also contribute to E-region shorting or smoothing of gradients which surpress irregularity production \citep{Koustov2004}.
+
+Irregularity production is impacted primarially by the presence of strong density gradients and electric fields in the ionosphere \citep{Koustov2004}.
+Properly oriented density gradients and electric fields are a nessesary condition for GDI to be operational, Section \ref{sec:lit_instbilities}, so the stronger these are, the more likely the irregularity production rate is high.
+Both backscatter power and occurrence increase with increasing electric field strength \citep{Ballatore2001,Danskin2002,Makarevich2014}.
+Similarly, backscatter occurrence increases when the IMF \(B_z\) component is strongly negative, implying a strong merging electric field.
+
+The primary radar propogation factor to consider is the background plasma density and how much the radar beam is refracted.
+If the background density is too low, the beam is under refracted and is never perpendicular to the magnetic field, however if the density is too high, the beam is over refracted, such that observation of small-scale FAIs is most favorable if the background density is between \(1\times 10^{11}\) m\(^{-3}\) and \(4\times 10^{11}\) m\(^{-3}\) \citep{Danskin2002,Makarevich2014}.
+If the D region is too dense, this can also in theory attenuate the signal and reduce the rate of returned backscatter, although observations have shown this is not a major factor in echo occurrence \citep{Danskin2002}.
+Finally, the radar equation predicts that the returned backscatter should be proportional to the perturbation density, \(\delta n\), squared, or, equivilently, the product of the background density and the fractional density, \(\delta n/n\), squared, Equation \ref{eqn:fractional_density}
+\begin{equation}
+	\label{eqn:fractional_density}
+	P \propto \delta n^2 \propto \left(\frac{\delta n}{n}\right)^2 n^2
+\end{equation}.
+There is no general agreement of what controls the fractional density, however, if it is approximately constant, then radar backscatter power could be directly proportional to \(n^2\).
+
+%Kane2012:
+%- diurnal, seasonal, solar cycle variation
+%- nighttime occurrence dirrectly proportional to peak electron density
+%- daytime occurrence significantly reduced by conductive E layer
+%Danskin2002:
+%- D region absorption not a major factor in echo occurrence
+%- under refraction occurs if n<1x10^11 m^-3
+%- over refraction occurs if n>4x10^11 m^-3
+%- echo power increases for stronger electric fields
+%Bristow2011:
+%- polar cap has nearly continuous production of decameter scale irregularities
+%- "near constant presnece of field-aligned density irregularities in the polar cap"
+%Milan1997:
+%- summer - backscatter observed at near ranges in midnight sector
+%- winter/equinox - backscatter observed at far ranges in midday sector
+%- frequency/boresite azimuth have a large impact on the ability to detect close backscatter
+%- increased backscatter at solar max
+%Makarevich2014:
+%- power increases with increasing density and electric field
+%- power drops significantly at large density values
+%- backscatter power proportional to n^2
+%Koustov2013:
+%- antisunward flows occur for negative Bz, sunward for positive
+%Koustov2004:
+%- increasced echo occurrence for solar max in midnight sector due to strong density gradients and electric fields
+%- noon sector does not see occurance increase with solar max
+%- high latitude radars have occurrence max in equinox months
+%- combination of propagation and production factors
+%Ballatore2001:
+%- backscatter occurrence increases for negative Bz
+%- backscatter occurrence increases with merging electric field strength
 

### 2017-01-03 thesis.sh
STE diagram started

@@ -16,3 +16,5 @@ bibtex introduction
 pdflatex LLamarcheThesis.tex
 
 pdflatex LLamarcheThesis.tex
+
+rm *.aux

### 2017-01-04 introduction.tex
Incorporated ISR map

@@
 Additionally, variations in temperature and electric and magnetic field strength can occur on a variety of scales.
-This thesis focuses on mechanisms by which density variations in ionospheric plasma occur, often referred to as plasma structuring.
+This thesis focuses on mechanisms by which density variations in ionospheric plasma occur, often referred to as plasma structures.
 
@@
 \begin{figure}
-	\includegraphics[width=1.3\textwidth,angle=90]{superdarnmap.pdf}
+	\includegraphics[width=\textwidth,angle=180]{SuperDARNmap.pdf}
 	\caption{FoVs of all operational SuperDARN radars in both the northern (left) and southern (right) hemispheres.
@@
 
+\begin{figure}
+	\includegraphics[width=\textwidth]{ISRmap.pdf}
+	\caption{FoVs of all ISRs in the northern polar region.
+The individual radars shown are identified by their commonly used acronyms: Poker Flat (PFISR), Resolute Bay North (RISR-N), Resolute Bay Canada (RISR-C), Sondrostrom (SFJ).
+RISR-N, which is primarially used throughout this thesis, is shown in orange.
+The FoV of the RKN SuperDARN radar is shown by the red outline.
+Lines of constant MLAT are shown at \(\Lambda=80\deg,60\deg\) in yellow.}
+	\label{fig:isrmap}
+\end{figure}
+
 The first ISR was built in Arecibo, Puerto Rico in the 1950s \citep{Gordon1958}.
 Since then, at least 8 other ISRs have been deployed around the world at equitoral, mid-, and high-latitudes.
-The most recent advancement has been the development of Advanced Modular Incoherent Scatter Radars (AMISR), of which there are currently three in use, one at the Poker Flat Rocket Range, just north of Fairbanks, Alaska and two in Resolute Bay, Canada.
+A map of the ISR systems currently operational in the northern polar cap is shown in Figure \ref{fig:isrmap}.
+The most recent advancement has been the development of Advanced Modular Incoherent Scatter Radars (AMISR), of which there are currently three in use, one at the Poker Flat Rocket Range (PFISR), just north of Fairbanks, Alaska and two in Resolute Bay, Canada (RISR-N and RISR-C).
 The advantage of these new systems is that they are electronical steerable, unlike older ISRs which consisted of a large dish that had to be physically moved to change the beam direction.
@@
 In the common WorldDay mode, there are typically 11 beams and data is collected at \(\sim\)1 min intervals, although this can change slightly between different renditions of the WorldDay mode.
+WorldDay mode creates the characteristic star shaped FoV seen in Figure \ref{fig:isrmap}
 
@@
 For these region, it is often useful use a combination of CSR and ISR measurements in ionospheric plasma structuring studies, when possible.
-The primary ISR used in the present volume of work is the north face of the AMISR system at Resolute Bay, Canada, RISR-N.
-The FoV of RISR-N overlaps that of the RKN SuperDARN radar, making it ideal for these types of comparison studies, Chapters \ref{sec:paper1} and \ref{sec:paper3}.
+The primary ISR used in the present volume of work is the north face of the AMISR system at Resolute Bay, Canada, RISR-N, shown in orange in Figure \ref{fig:isrmap}.
+The FoV of RISR-N overlaps that of the RKN SuperDARN radar, red outline in Figure \ref{fig:isrmap}, making it ideal for these types of comparison studies, Chapters \ref{sec:paper1} and \ref{sec:paper3}.
 

### 2017-01-04 thesis.sh

@@ -3,8 +3,9 @@
 
 gs -q -dSAFER -dBATCH -dNOPAUSE -sDEVICE=pdfwrite -dDEVICEWIDTHPOINTS=530 -dDEVICEHEIGHTPOINTS=410 -sOutputFile=/Users/ljlamarche/Desktop/UAFThesis/Figures/densprofile.pdf /Users/ljlamarche/Desktop/UAFThesis/Figures/densprofile.ps
 
-gs -q -dSAFER -dBATCH -dNOPAUSE -sDEVICE=pdfwrite -dDEVICEWIDTHPOINTS=270 -dDEVICEHEIGHTPOINTS=520 -sOutputFile=/Users/ljlamarche/Desktop/UAFThesis/Figures/superdarnmap.pdf /Users/ljlamarche/Desktop/UAFThesis/Figures/SuperDARNmap.ps
+gs -q -dSAFER -dBATCH -dNOPAUSE -sDEVICE=pdfwrite -dDEVICEWIDTHPOINTS=270 -dDEVICEHEIGHTPOINTS=520 -sOutputFile=/Users/ljlamarche/Desktop/UAFThesis/Figures/SuperDARNmap.pdf /Users/ljlamarche/Desktop/UAFThesis/Figures/SuperDARNmap.ps
 
+gs -q -dSAFER -dBATCH -dNOPAUSE -sDEVICE=pdfwrite -dDEVICEWIDTHPOINTS=510 -dDEVICEHEIGHTPOINTS=510 -sOutputFile=/Users/ljlamarche/Desktop/UAFThesis/Figures/ISRmap.pdf /Users/ljlamarche/Desktop/UAFThesis/Figures/ISRmap.ps
 
 
 pdflatex LLamarcheThesis.tex

### 2017-01-04 introduction.tex
Motivations and Objectives revised

@@
 
-It is well accepted that the polar cap ionosphere is highly structured in a non-trivial manner \citep{Carlson2012}.
-In addition to being of purely scientific interest, this structuring creates complications when sending radio waves through the ionosphere.
+The plasma in the polar cap ionosphere is highly structured in a non-trivial manner \citep{Tsunoda1988,Carlson2012}.
+In addition to being of purely scientific interest, this structuring creates complications when receiving radio signals through the ionosphere.
 Because the structuring is highly irregular and varies on many spatial and temporal scales, unpredictable wave refraction, phase shifts, and amplitude attenuation can occur, making signals difficult to detect.
-This introduces problems in any system that involves ground-to-satellite communication, such as navigation or communication.
+This introduces problems in any system that involves ground-to-satellite, satellite-to-ground, or over-the-horizon communication, such as navigation or communication.
 Because of the increasing interconnectedness of modern society, any interruptions or errors in these systems can have broad impacts on infrastructure, including commercial and defense interests, as well as everyday life.
@@
 
-The aim of this work is to examine some of the factors (such as global solar control and GDI directional dependance) behind plasma irregularity production in the F region polar cap using both observations and theory.
+The aim of this work is to investigate small-scale plasma irregularity production in the F-region polar cap.
+In particular, factors such as direct and indirect global solar control and GDI directional dependance will be considered.
 The majority of observational data has been found from ground based ionospheric radars, both CSR systems (Section \ref{sec:csr}) and ISR systems (Section \ref{sec:isr}), with particular emphasis on radars within the SuperDARN network, Section \ref{sec:superdarn}.
-However, these datasets have also been supplemented with information from models as well as other instruments when necessary.
-In addition, most experimental results are compared with predictions made by a linear fluid theory view of ionospheric plasma behavior.
-Specific objectives are as follows.
+However, these datasets have also been supplemented with information from models as well as satellite data when necessary.
+Results are investigated in the context of a linear fluid theory analysis of ionospheric plasma.
 
 The first objective of is to evaluate the extent of solar control over irregularity production in the polar F region.
-This considers a global view of small-scale plasma structuring.
 The occurrence of radar backscatter will be compared with both solar illumination (considered direct solar control) and IMF factors (considered indirect solar control).
@@
 The second objective is to investigate asymmetry in the GDI growth rate surrounding large-scale density structures.
-This is a modeling approach to identifying where plasma irregularities occur surrounding a polar patch and how they can be observed by a ground-based HF radar.
-The model allows certain factors to be considered that can be challenging to investigate experimentally because it is difficult to isolate particular effects.
-The model used considers the direction of a large, elongated polar patch as well as its drift direction to be arbitrary in relation to the radar and independent of each other, which gives some insight as to how structuring would be expected with polar patches in orientations that are not often observed.
+A model is used to identifying where plasma irregularities occur surrounding a polar patch and how they can be observed by a ground-based HF radar.
+The model considers a large, elongated polar patch drifting in an arbitrary direction relative to a radar and assumes irregularity production is directly dependent to the linear GDI growth rate.
+High linear growth rates predicted by the model are compared with classic observations of stronger backscatter power trailing drifting polar patches.
 
-The third objective is to investigate the relationship between factors that control the linear GDI growth rate and small-scale plasma structuring.
-This combines the theoretical consideration of where growth rate is greatest with the experimental record of where structuring is observed around a polar patch.
-This is particularly important because even though plasma structuring is often observed on the trailing edge of polar patches and this is where GDI predicts the greatest growth rates, GDI theory actually predicts growth of much larger irregularities than the small scale structuring observed by HF radars.
-A direct comparison of the two improves our understanding of how growth rates are related to small-scale plasma structuring.
+The third objective is to investigate directional dependancies of small-scale plasma irregularities.
+Two dimensional measurements of density gradients and electric fields are made such that the directionally dependent GDI growth rate can be calculated, which is then compared with plasma irregularity power and occurrence in a particular location.
+This gives insight into whether the production of small-scale irregularities is directionally dependent, similar to the linear GDI growth rate, or if the distribution is isotropic due to other nonlinear processes.
 
@@
 Chapter \ref{sec:introduction} has provided an overview of the polar ionosphere and both macro- and microphysical plasma structuring mechanisms.
-Chapter \ref{sec:paper1} will discuss plasma irregular occurrence and the extent to which solar factors control this occurrence.
-Chapter \ref{sec:paper2} will consider the directional dependance of the GDI growth rate, and in particular give recommendations for the optimal orientation between a ground-based ionospheric radar and a large-scale density structure to observe small-scale structuring.
-Chapter \ref{sec:paper3} will examine coincident observations made an IRS of background plasma density conditions and a CSR of small-scale plasma structuring in the polar cap and provide evidence for small-scale structuring being directionally dependent in a similar fashion to GDI predictions for large-scale plasma structuring.
-Chapter \ref{sec:conclusion} will conclude this thesis and give a summary of the most significant results.
-The implications of these results will be discussed along with questions that are still open and recommendations for future research.
+Chapter \ref{sec:paper1} presents an experimental investigation of decameter-scale plasma irregularities in the polar F-region and their solar control, with a particular focus on statistical analysis of data collected with a recently-deployed SuperDARN radar in the southern polar cap.
+The same results were also presented in a recent journal article \citep{Lamarche2015}.
+Chapter \ref{sec:paper2} presents a modeling study of the linear GDI growth rate surrounding a large, elongated polar patch drifting in an arbitrary direction within the FoV of a SuperDARN radar uses the results to provide insight as to optimal conditions from radars to observe irregularities.
+These results have also been recently published \citep{Lamarche2016}.
+Chapter \ref{sec:paper3} presents an experimental investigation of the directional dependence of small-scale plasma structures by considering simultaneous measurements of electric field and density gradients from an AMISR  system and FAIs observed by SuperDARN.
+Particular emphasis is placed on if small-scale structures are directionally dependent similar to the predictions of linear GDI theory.
+The same results have been submitted for publication \citep{Lamarche2017}.
+Chapter \ref{sec:conclusion} presents a summary of the most significant results of this body of work, their implications for interpretation of radar observations of plasma structures in the polar cap and suggestions for future research involving coherent and incoherent scatter radar facilities in the polar regiont.
 

### 2017-01-06 introduction.tex
STE figure incorporated into text

@@
 The Sun interacts with the rest of the solar system through the solar wind, a continuous outflow of plasma that caries the Sun's magnetic field with it.
-The solar wind typically has a density around 10 cm\(^{-3}\) and is traveling away from the Sun at approximately 400 km/s at Earth's orbit.
+This is shown in Figure \ref{fig:ste}.
+The solar wind, green, typically has a density around 10 cm\(^{-3}\) and is traveling away from the Sun at approximately 400 km/s at Earth's orbit.
 However, like the Sun, the solar wind is highly dynamic and these parameters can change drastically over a period of minutes.
 CMEs, for instance, create a substantial increase in density and are often preceded by a shock, a region of hot, compressed plasma caused by the CME moving at supersonic speeds.
-In addition to the outflow of plasma, the solar wind also consists of the Sun's magnetic field extended into the outer limits of the solar system, called the interplanetary magnetic field (IMF).
+In addition to the outflow of plasma, the solar wind also consists of the Sun's magnetic field extended into the outer limits of the solar system, called the interplanetary magnetic field (IMF), dark blue in Figure \ref{fig:ste}.
 Because the Sun has a short rotation period (about 33 days), the IMF gets twisted into a spiral, commonly called the Parker Spiral \citep{Parker1958}.
 When the IMF reaches the Earth, the magnetic field is typically oriented at a \(45^\circ\) inclination relative to the ecliptic, but this is highly variable depending on solar wind conditions.
+
+\begin{figure}
+	\includegraphics[width=0.7\textwidth,angle=270]{ste.pdf}
+	\caption{Diagram of the solar-terrestrial environment.
+The solar wind contains both a steady stream of charged particles and the IMF, dark blue.
+The red region represents the magnetosheath.
+The manetopause is shown by the black dashed line.
+Closed magnetic field lines from the earth are shown in grey while open field lines are shown in light blue.}
+	\label{fig:ste}
+\end{figure}
 
@@
 Because the solar wind is typically a supersonic flow, a bow shock forms on the dayside of the magnetosphere.
-The heated and compressed solar wind plasma created by the bow shock is known as the magnetosheath.
-The magnetopause is the actual boundary between solar wind plasma in the magnetosheath and magnetospheric plasma.
+The heated and compressed solar wind plasma created by the bow shock is known as the magnetosheath, the red region in Figure \ref{fig:ste}.
+The magnetopause, black dashed line, is the actual boundary between solar wind plasma in the magnetosheath and magnetospheric plasma.
 However, the magnetosphere is not fully closed.
 At the magnetopause, magnetic reconnection can occur between the IMF and the Earth's magnetic field creating open field lines.
+The last closed field line (closed field lines shown in grey) where reconnection tends to occur at the magnetopause is highlighted in yellow.
 Magnetic field lines originating from the Earth are then directly connected to the Sun through the solar wind, which consequently allows highly energized solar particles to enter the magnetosphere.
-Open field lines tend to occur around the Earth's polar caps, making these regions particularly interesting to study due to very complex behavior and coupled interactions between the Earth and the Sun.
+Open field lines, light blue, tend to occur around the Earth's polar caps, making these regions particularly interesting to study due to very complex behavior and coupled interactions between the Earth and the Sun.
 
@@
 
-%Kane2012:
-%- diurnal, seasonal, solar cycle variation
-%- nighttime occurrence dirrectly proportional to peak electron density
-%- daytime occurrence significantly reduced by conductive E layer
-%Danskin2002:
-%- D region absorption not a major factor in echo occurrence
-%- under refraction occurs if n<1x10^11 m^-3
-%- over refraction occurs if n>4x10^11 m^-3
-%- echo power increases for stronger electric fields
-%Bristow2011:
-%- polar cap has nearly continuous production of decameter scale irregularities
-%- "near constant presnece of field-aligned density irregularities in the polar cap"
-%Milan1997:
-%- summer - backscatter observed at near ranges in midnight sector
-%- winter/equinox - backscatter observed at far ranges in midday sector
-%- frequency/boresite azimuth have a large impact on the ability to detect close backscatter
-%- increased backscatter at solar max
-%Makarevich2014:
-%- power increases with increasing density and electric field
-%- power drops significantly at large density values
-%- backscatter power proportional to n^2
-%Koustov2013:
-%- antisunward flows occur for negative Bz, sunward for positive
-%Koustov2004:
-%- increasced echo occurrence for solar max in midnight sector due to strong density gradients and electric fields
-%- noon sector does not see occurance increase with solar max
-%- high latitude radars have occurrence max in equinox months
-%- combination of propagation and production factors
-%Ballatore2001:
-%- backscatter occurrence increases for negative Bz
-%- backscatter occurrence increases with merging electric field strength
-
 \section{Motivation and Objectives}

### 2017-01-10 conclusion.tex
Conlusion with Future Work

% Chapter 5: Conclusion

\chapter{Conclusions and Future Work}
\label{sec:conclusion}

Plasma irregularities in the F-region ionosphere have been observed for decades.  However, despite extensive data sets collected over these years, supprisingly little is known about these irregularities, especially at small scales.  This is in part due to the complex and turbulent nature of the ionosphere, particularly in the polar regions, but also due to some of the inherent limitations in traditional observation techniques which can make it very challenging to identify exactly what processes are operational.  In this Chapter, we summarize the results that have previously been discussed in Chapters \ref{sec:paper1}--\ref{sec:paper3} and consider what overarching conclusions can be derived from them which further our understanding of small-scale polar F-region plasma irregularities.

\section{Summary and Conclusions}
\label{sec:summary}

\section{Future Work}
\label{sec:futurework}

Although much progress has been made in understanding small-scale plasma structuring and HF radar observations, many of the results presented in this thesis give rise to further questions that must be addressed.

In Chapter \ref{sec:paper1}, the occurence of radar backscatter durring nighttime in the polar cap was found to be much higher than expected, particularly given that raytracing simulations predict that the orthagonality condition is rarely met and therefore there should be little to no backscatter.  The proposed hypothesis to explain this discreperency is that polar patches, which are not typically well represented by ionospheric density models, occur frequently in the nighttime F region and, on average, enhance the background plasma density enough that propagation conditions are sufficient for HF radar backscatter to regularly be observed.  Experimental surveys of polar patches have shown low occurrence at nighttimes, however the occurrence of patches and their intensity increase at equinox, which is where frequent nighttime backscatter is predominantly observed.  The expected signature in the radar data from this kind of senario would be a night with a series of isolate ``patches'' of backscatter.  An example of such an event was presented, Figure \ref{whatever the hell THIS'll be}, but it is unknown how often events like this occur.  Additionally, the intensity and size of a patch nessesary to cause the radar beam to refract sufficiently must be established, possibly through further use of raytracing simulations with customizable density profiles.

One of the major drawbacks to linear GDI theory is its limitation of applicability only to long wavelengths, which means if the F region, it cannot be directly applicable to decameter-scale irregularities measured by HF radars.  Generally, it is assumed that some form of turbulent cascade connects plasma structures of different scale sizes in the ionosphere such that predictions made at large scales can still be applied to decameter-scale observations.  This helps explain some of the observations of asymmetry around polar patches.  This asymmetry was also confirmed in the GDI growth rate from the results shown in Chapter \ref{sec:paper2}.  However, these results also predicted anisotropy in the growth rate, which is more challenging to confirm from small-scale irregularity obsersavtions, although some evidence of such has been provided in Chapter \ref{sec:paper3}.  It may be possible to further the theoretical understanding of GDI at small scales by developing a three dimentional picture of it using kinetic theory.  \citet{Basu1995} developed such a formulation for a specific geometry throughout all altitudes within the ionosphere, but expanding this to general vector directions would most likely be extremely mathematically intensive.

The results presented in Chapter \ref{sec:paper3} could also be improved with a more extensive data set.  The event chosen for analysis in that study was in 2011 and was selected on the basis of good data coverage from both RISR-N and RKN and a strong density gradient being apparent in the overlapping FoV.  Since then, RISR-C has been deployed with a FoV just south of RISR-N (Figure \ref{fig:isrmap}) such that observations between both radars can be coordinated and they can act as a single instrument with a FoV twice as large.  This allows a polar patch to be tracked over a longer distance and more individual data points to be captured.  Additionally, specific modes can be utilized that give a much finner grid of RISR beams, allowing measurments to be made closer together and smaller scale gradients to be be considered.  The downside to these modes is that it takes longer for each set of measurements to be made, so the temporal cadence decreases.  However, the required integration time for an ISR decreases if the background ionospheric density is higher, so the the temporal resolution can improve for measurments taken durring summer or daytime.  Another advantange to the RISR-N/RISR-C locations is that the SuperDARN radars at Clyde River (CLY) and Inuvik (INV) (Figure \ref{fig:superdarnmap}), both of which share at least part of the RISR-N/RISR-C FoV, but observe it from different directions.  This presents an unique oppertunity to further investigate the directional dependence of small-scale irregularities by observing the differences in backscatter observed from the same scattering volume at different directions.  Although straight forward in theory, this experiment would most likely prove nontrivial in practice.  Backscatter tends to be limited to the statistical high occurrence band discussed in Chapter \ref{sec:paper1}, and at least with INV, this band probably occurs at closer ranges than where the FoV overlaps other radars, so coincident backscatter is probably rare.  The probability of backscatter being observed is better when the background density is within the range where propogation conditions are most favorable, when the ionosphere is sunlit.  Additionally, it is adventagious to choose a time when the occurrence of backscatter is statistically high.  Based on results from Chapter \ref{sec:paper1}, the best time for such an experiment would therefore be daytime durring an equinox month.

### 2017-01-10 thesis.sh

@@ -14,8 +14,14 @@ bibtex LLamarcheThesis
 
 bibtex introduction
 
+bibtex conclusion
+
 pdflatex LLamarcheThesis.tex
 
 pdflatex LLamarcheThesis.tex
 
 rm *.aux
+rm *.bbl
+rm *.blg
+rm *.out
+rm *.log

### 2017-01-14 conclusion.tex
Conclusion draft

@@
 \label{sec:summary}
+
+Irregularities occur most often close to the terminator.
+This is likely due to the presence of a very large density gradient between high density solar illuminated plasma on the dayside and low density plasma on the nightside.
+In addition, if a SuperDARN radar itself is positioned close to the terminator, it is observing the density gradient along its boresight.
+If small-scale irregularity growth is directionally dependent, this is additionally the optimal orientation from which to observe the plasma structures, leading to increased echo occurrence.
+
+In the polar caps, backscatter is often observed at nighttime even when background plasma density is low and propagation conditions are too poor for FAIs to be observed with HF radar.
+This implies an additional source of nighttime plasma density that is not easily accounted for by models.
+Frequent polar patches that propogate from the dayside to the nightside could be the source of this additional density.
+The requirements for confirming this hypotheis will be detailed further in Section \ref{sec:futurework}.
+
+Nighttime observations of FAIs is also higher with a positive IMF By component.
+This could be due to the control that the By component over ionospheric convection patterns.
+In a convectional two-cell convection pattern, the plasma flows antisunward from dayside to nightside across the terminator.
+In this arrangement, the plasma flow, \(\vec{V}_E\), and the density gradient, \(\vec{G}\), are antiparallel, the least favorable orientation for GDI growth.
+However, with a significant IMF By component, the orientation of the convection cells rotates such that plasma is no longer convects antisunward, creating a situation where GDI is more favorable and plasma structures can develop more readily.
+Additionally, an IMF By component can contribute to the formation of polar patches by creating convection perturbations that separate high density patches from the daytime resevoir, as described in Section \ref{sec:lit_patches}.
+In this way, the IMF By component may be indirectly responsible for the increased observation of backscatter.
+If patch occurrence increases, it can contribute to the nighttime background density, as described above, which can improve propagation conditions so that FAIs can be observed by
+HF radars.
+
+In addition to simply providing additional background density so nighttime propagation conditions are sufficient for the radar to observe FAIs, polar patches can contribute to plasma irregularity formation directly by providing background density gradients for GDI to be operational on.
+This raises the important question of exactly how and where plasma structures form around polar patches.
+Observations have long suggested there is asymmetry in plasma structuring surrounding polar patches, but the results presented in Chapter \ref{sec:paper2} suggest that production may also be anisotropic.
+Asymmetry and anisotropy are independent such that the plasma structures can exhibit either of them independently or both or neither.
+
+The plasma drift and wavevector directions influence the asymmetry that is observed surrounding polar patches.
+The direction that the patch is drifting relative to its elongation direction impacts the orientations of \(\vec{G}\) and \(\vec{V}_E\) and controls the GDI growth rate.
+This also creates the asymmetry between the different edges of the patch because as the gradient direction changes, the irregularity growth rate changes.
+Additionally, linear theory predicts that irregularity growth is anisotropic, so the orientation of the patch relative to the radar can also impact the observed irregularities.
 

### 2017-01-15 conclusion.tex
Bibliography issues corrected

@@
 
-In Chapter \ref{sec:paper1}, the occurence of radar backscatter durring nighttime in the polar cap was found to be much higher than expected, particularly given that raytracing simulations predict that the orthagonality condition is rarely met and therefore there should be little to no backscatter.
+In Chapter \ref{sec:paper1}, the occurence of radar backscatter durring nighttime in the polar cap was found to be much higher than expected \citep{Bristow2011}, particularly given that raytracing simulations predict that the orthagonality condition is rarely met and therefore there should be little to no backscatter.
 The proposed hypothesis to explain this discreperency is that polar patches, which are not typically well represented by ionospheric density models, occur frequently in the nighttime F region and, on average, enhance the background plasma density enough that propagation conditions are sufficient for HF radar backscatter to regularly be observed.
-Experimental surveys of polar patches have shown low occurrence at nighttimes, however the occurrence of patches and their intensity increase at equinox, which is where frequent nighttime backscatter is predominantly observed.
+Experimental surveys of polar patches have shown low occurrence at nighttimes, however the occurrence of patches and their intensity increase at equinox, which is where frequent nighttime backscatter is predominantly observed \citep{Rodger1996}.
 The expected signature in the radar data from this kind of senario would be a night with a series of isolate ``patches'' of backscatter.
@@
 One of the major drawbacks to linear GDI theory is its limitation of applicability only to long wavelengths, which means if the F region, it cannot be directly applicable to decameter-scale irregularities measured by HF radars.
-Generally, it is assumed that some form of turbulent cascade connects plasma structures of different scale sizes in the ionosphere such that predictions made at large scales can still be applied to decameter-scale observations.
+Generally, it is assumed that some form of turbulent cascade connects plasma structures of different scale sizes in the ionosphere such that predictions made at large scales can still be applied to decameter-scale observations \citep{Tsunoda1988}.
 This helps explain some of the observations of asymmetry around polar patches.
@@
 Based on results from Chapter \ref{sec:paper1}, the best time for such an experiment would therefore be daytime durring an equinox month.
+
+\bibliographystyle{uafthesis}
+\bibliography{references}

### 2017-01-15 introduction.tex
Conlusion with Future Work
Paper 1 added

@@
 % Chapter 1: Introduction
+
+\chapter{Introduction}
 \label{sec:introduction}
@@
 \begin{figure}
-	\includegraphics[width=0.7\textwidth,angle=270]{ste.pdf}
+	\includegraphics[width=\textwidth,angle=0]{ste.pdf}
 	\caption{Diagram of the solar-terrestrial environment.
@@
 For convection maps, only data from the F region is considered.
-If there is not enough velocity data to confine the fit sufficiently, the measured data points are supplemented with a statistical model based on solar wind IMF conditions \citep{Ruohoniemi1995,Ruohonieme2005}.
+If there is not enough velocity data to confine the fit sufficiently, the measured data points are supplemented with a statistical model based on solar wind IMF conditions \citep{Ruohoniemi1995,Ruohoniemi2005}.
 

### 2017-01-15 paper1.tex
Paper 1 added

%\titlerunninghead{Solar Control of Polar Irregularities}

%\authoraddr{L. J. Lamarche and R. A. Makarevich, Geophysical Institute, University of Alaska Fairbanks, 903 Koyukuk Drive, PO Box 757320, Fairbanks, AK, 99775-7320, USA.
%(rmakarevich@alaska.edu)}

%\linenumbers*[1]
%\begin{document}

\chapter{Solar control of \(F\) region radar backscatter: Further insights from observations in the southern polar cap}

%\authors{Leslie J. Lamarche\altaffilmark{1} and Roman A. Makarevich \altaffilmark{1}}
%\vspace{-1.5cm}

%\altaffiltext{1}{Geophysical Institute and Department of Physics, University of Alaska Fairbanks, Fairbanks, AK, USA.}

%\begin{abstract}

The role of solar wind and illumination in production of small-scale \(F\) region plasma irregularities is investigated using a 4-year data set collected by the Super Dual Auroral Radar Network (SuperDARN) facility at the McMurdo station, Antarctica (MCM). Statistical analysis of ionospheric echoes detected by MCM shows that radar backscatter from the polar \(F\) region occurs in wide and persistent bands in range that exhibit systematic changes with local time, season, and solar cycle. It is demonstrated that all variations considered together form a distinct pattern. A comparison with \(F\) region model densities and raytracing simulations shows that this pattern is largely controlled by the \(F\) region solar-produced ionization during the day. During the night, however, MCM observations reveal a significant additional source of plasma density in the polar cap as compared with the model. An example of conjugate radar observations is presented that supports the idea of polar patches being this additional source of ionization on the nightside. Echo occurrence exhibits a clear peak near the solar terminator, which suggests that small-scale irregularities form in turbulent cascade from large scales. Further, echo occurrence is enhanced for particular IMF orientations during the night. Observations indicate that solar illumination control of irregularity production is strong and not restricted to the nightside. Indirect solar wind control is also exerted by the IMF-dependent convection pattern, since the gradient-drift instability favors certain orientations between the plasma density gradients and convection velocity.

%\end{abstract}

%\begin{article}
\section{Introduction}
\label{Sec:Intro}

Plasma density waves or irregularities are commonly observed in both the \(E\) and \(F\) regions of the ionosphere and with scales ranging from a few centimeters to several kilometers. Several methods have been used to study ionospheric structures, including in situ rocket and satellite measurements as well as remote sensing using radio and optical techniques \citep{Fejer1980}.  In this study, we focus on small-scale irregularities (\(\sim\)10 m) in the \(F\) region that are thought to be primarily due to the gradient-drift instability (GDI) \citep{Tsunoda1988}.

Production of these irregularities is reasonably well understood on a microphysical level, with GDI promoting irregularity growth for certain orientations of the density gradient vector relative to the convection electric field \citep{Kesknien1982a,Kesknien1982b,Makarevich2014c}. The role of global factors, on the other hand, such as geomagnetic activity or solar wind conditions is much less investigated. In particular, the extent of the solar control is largely unknown, both in terms of its direct effects due to varying ionization and  indirect effects due to varying interplanetary magnetic field (IMF) as explained below. This is particularly important to processes occurring in the polar cap where some of the strongest differences in solar illumination conditions occur and where geomagnetic field lines are directly connected to the IMF.

One fundamental question under investigation in whether solar illumination suppresses or enhances irregularity production. In the auroral zone, evidence has been presented that solar illumination reduces the rate at which irregularities are observed by coherent radars in the Super Dual Auroral Radar Network (SuperDARN) \citep{Ruohoniemi1997,Koustov2004}.  This is thought to be due to the background electron density increasing and smoothing the density gradients required to drive GDI \citep{Ruohoniemi1997,Koustov2004}. Preliminary studies in the polar ionosphere, however, indicated that echo occurrence is much reduced during nighttime \citep{Bristow2011}, implying that either daytime echo occurrence is enhanced or nighttime occurrence is suppressed. This needs to be further investigated to determine why and how irregularity production in the polar region may differ from that in the auroral region.

Observations of decameter-scale irregularities with SuperDARN have an advantage of significant datasets accumulated under a wide variety of observational conditions. SuperDARN utilizes radio refraction in the High Frequency (HF) band to observe \(F\) region echoes or backscatter from decameter-scale irregularities when radar beam achieves orthogonality with local magnetic field (orthogonality condition). Because of importance of propagation conditions for HF radar observations, the echo occurrence rates observed by SuperDARN are taken as a lower limit of the irregularity occurrence rates \citep{Chisham2007}.

Several recent SuperDARN studies have found clear seasonal and diurnal trends in the \(F\) region echo occurrence that formed a distinct pattern which appeared to be qualitatively similar to that of the solar zenith angle (SZA) variations \citep{Kane2012,Ghezelbash2014a,Ghezelbash2014b}. However, it has been challenging to determine how occurrence explicitly depends on SZA for all seasonal and time sectors. The nighttime occurrence taken separately was demonstrated to be closely related to SZA-controlled \(F\) region density  variations \citep{Kane2012}, but no similarly-clear relationship has been demonstrated for datasets including daytime points.

One difficulty is the need to consider propagation effects of plasma density. These arise because of \(F\)-region density controlling the amount of refraction and either under- or over-refraction reducing echo occurrence when orthogonality condition is not met \citep{Milan1997,Danskin2002}. This is also related to how echo occurrence is generally considered a lower limit on irregularities (described above). There is some ambiguity about whether changing irregularity production or detection by the radar is most responsible for any observed trends.

The production/detection ambiguity issue can be addressed by considering how both occurrence and distance from the radar (slant range) change in responses to various factors \citep{Kane2012,Ghezelbash2014b}. This is because range variations are expected to be almost entirely due to propagation effects and the differences in responses are therefore mostly likely due to changing production. In order to claim production as the dominant source of occurrence variability, occurrence analysis must explicitly considers the echo location. In the current study this approach is employed to develop a comprehensive model of occurrence location, quantify echo location trends, and explicitly consider these trends when developing occurrence estimates. This is in contrast with previous studies that considered echoes detected in the entire (or fixed part) of the radar's field-of-view (FoV) \citep[e.g.][]{Kumar2011,Kane2012}.

In addition to direct effects of solar illumination, there is also interest in the indirect ways in which the Sun may control irregularity production in the ionosphere. It has been shown that negative IMF Bz generally leads to higher echo occurrence in the auroral \(F\) region \citep{Ballatore2001}. An even stronger IMF control has been demonstrated for the auroral \(E\) region irregularities \citep{Makarevich2012}. Some dependence on substorm phase (and hence indirectly on IMF) is also expected and has been observed \citep{Wild2008}. One way in which the IMF and echo occurrence variations may be related is through the IMF-controlled convection pattern \citep{Ruohoniemi1996,Pettigrew2010}. Convection changes lead to different orientations between convection velocity and density gradient vectors, which may, in turn, change GDI contribution to structuring processes \citep{Makarevich2014c}. The feasibility of this scenario is largely unknown and the extent of IMF control of polar cap echo occurrence is yet to be evaluated.

The overall aim of this study is to evaluate the extent of solar control on the production of \(F\)-region density irregularities in the polar cap ionosphere. Specifically, the objectives are to (1) quantify variations in \(F\) region backscatter location for different time sectors, seasons, and solar cycle phases, (2) analyze direct effects of solar illumination in irregularity production utilizing a comprehensive model of echo location, and (3)  establish the relationship between echo occurrence and IMF in the polar cap \(F\) region.

\section{Experiment Configuration and Data Processing}
\label{Sec:Exp}

The primary instrument used in this study is the SuperDARN McMurdo radar located at McMurdo Station, Antarctica (77.88\(\deg\) S, 166.73\(\deg\) E, geographic) \citep{Bristow2011}. Using a standard SuperDARN 3-letter radar code, this radar is referred to as MCM. SuperDARN is an array of coherent HF radars used to study processes in the Earth's ionosphere and magnetosphere. Radars are located at mid and high latitudes in both the Northern and Southern Hemispheres. Radars measure a 17-lag auto-correlation function (ACF) from which estimates of the Doppler velocity, power, and spectral width of ionospheric echoes in 75--100 range bins for each of the 16--24 radar beams are obtained. Each radar beam spans \(\sim3.25\deg\) horizontally and \(\sim30\deg\) vertically, with the entire FoV of \(\sim52\deg\)--\(78\deg\) in azimuth. Range bins typically start at 180 km and have a length of 15 km or 45 km and a radar scans through all beams in 60 s or 120 s depending on the mode of operation. Other details on operations and technical characteristics of SuperDARN radars are given by \citet{Chisham2007}.

The magnetic latitude (MLAT) of MCM is 80.0\(\deg\) S, based on altitude-adjusted corrected geomagnetic (AACGM) coordinate system with 2010 coefficients \citep[see recent paper by][]{Shepherd2014}. The MCM FoV covers the southern polar cap and beam 14 of MCM is most conjugate with beam 1 of the SuperDARN Rankin Inlet (RKN) radar that overlooks the northern polar cap. The current study focuses on MCM observations, but in order to facilitate future comparisons with conjugate observations, the data from beam 14 of MCM are considered. MCM has been in routine operations since February 2010 and the current study has considered the data from the first 4 years of operations 2010--2013.

Standard SuperDARN criteria were used to eliminate ground scatter (\(\left| V-\Delta V \right| <30\) m/s and \(\left| W-\Delta W \right| <35\) m/s), where \(V\) is the Doppler velocity, \(W\) is the spectral width, and \(\Delta V\) and \(\Delta W\) represent uncertainties in these quantities. In addition, measurements with SNR \(<\) 3 dB or \(W>\) 500 m/s were eliminated to reduce contribution from interference and other spurious echoes. In addition to standard operation modes, each SuperDARN radar has a set time periods where the radar is run in special modes. In order to have as continuous of a dataset as possible for the statistical analysis, data were used both from the standard modes and from special modes when possible. Modes where the beam 14 was over-sampled (beam camping modes) or not sampled at all were not used.

In this study, we also employed the IMF measurements taken from the high-resolution OMNI database using spacecraft-specific dataset for the WIND satellite. This was the OMNI dataset that had the largest number of valid measurements throughout the period of interest 2010--2013. All data in the OMNI database are time shifted from the spacecraft location to Earth's bow shock nose (BSN). The data were further shifted by 9--12 min to account for propagation from the BSN to the high-latitude ionosphere using a method described by \citet{Khan1999} that was also employed by \citet{Makarevich2012}. The IMF By and Bz components were averaged over 10-min intervals. Any periods with IMF information missing were excluded from the analysis.

The model electron densities were also used in the current study. The electron densities were found from the International Reference Ionosphere (IRI) 2007 model \citep{Bilitza2008}. The IRI model is an empirical model of the ionosphere based on time of day, date, geographic location, altitude, and solar conditions, with electron density near the \(E\) and \(F\) region peaks being of particular interest for this study.  Using ionosonde terminology, these are also referred to below as model NmE and NmF2.

\section{Observations}
\label{Sec:Obs}

In order to examine diurnal, seasonal, and solar cycle trends in echo occurrence, the monthly average occurrence was found for each universal time (UT) 2-min interval and for each 45-km range bin. If measurements were conducted at a 2-min resolution, each radar cell had an occurrence of 1 or 0. For a higher resolution, the values were weighted accordingly so that if an echo was observed for every possible time within the 2-min interval, the overall echo occurrence was still 1. The data for each time/range cell were averaged over the entire month. Figures \ref{fig:month_avg_occ}a--d show monthly average occurrence expressed in percentage for four selected months in year 2010 versus UT and slant range. The four months presented were chosen to be representative of different seasons. Figures \ref{fig:month_avg_occ}e--h on the right similarly shows the monthly average peak \(F\)-region electron density NmF2 obtained from the IRI model. The cosine white contours in both left and right panels show the approximate range position of the \(F\)-region echo band for each month determined as described below.

\begin{figure}
\includegraphics[width=12.5 cm,angle=90]{month_density_occurance2010.pdf}
\caption{Monthly average plots of echo occurrence (left) and peak electron density (right) for 2010.  Four months are selected (one for each season) to show seasonal variations. In Figures \ref{fig:month_avg_occ}a--\ref{fig:month_avg_occ}d,  the pink line shows the 25\% occurrence contour that is used to define the echo band region. Black dots are the midpoints between the upper and lower portion of this contour.  A cosine function (white line) is fit to these points and used to model the echo band. The dashed white lines are the upper and lower bounds of the model band. The top right corner of each panel shows the parameters that characterize the band for that month: \(A\) is amplitude of cosine function (in km), \(B\) is time shift (in hrs), \(C\) is range shift from 0 (in km), and BW is the band width (constant distance between the upper and lower dashed white lines).  In Figures \ref{fig:month_avg_occ}e--\ref{fig:month_avg_occ}h the same echo location model is also given (white lines) along with expected echo locations based on raytracing simulations (dark red line).}
\label{fig:month_avg_occ}
\end{figure}

Figures \ref{fig:month_avg_occ}a--d show a distinct \(F\)-region band of high occurrence in every month at ranges 500--1500 km. Some \(E\)-region echoes are also observed at shorter ranges 180--300 km in June at 01--06 UT and in December at 17--24 UT. The \(F\)-region band range position and peak occurrence within the band was changing most clearly throughout the day in March and September, i.e. diurnal variation in both parameters is observed clearly in spring and fall months. To highlight the high echo occurrence band, a contour of 25\% occurrence is plotted in each left panel by the pink line and this shows a clear diurnal, cosine-like variation with UT in March and September.

The midpoints were also found between the top and bottom parts of the pink 25\% occurrence contour and the cosine function of the form
\begin{equation}\label{E:Model}
R\left(t\right) = A \cos\left[2\pi\left(t-B\right)/24\right] + C
\end{equation}
was fitted to these points. Here \(R\) is the range of the center of the band in kilometers, \(t\) is the time in hours, and  \(A\), \(B\), and \(C\) are fit coefficients with units of kilometers, hours, and kilometers, respectively. The white solid line in each panel of Figure \ref{fig:month_avg_occ} shows the resulting fit \(R\left(t\right)\). The white dashed lines represent the upper and lower limits of the band.  They were calculated by first finding the average distance \(\sigma\) between points in the 25\% occurrence contour and the fitted cosine at the same time of day. The original cosine fit was then shifted up or down by this amount as \(R\left(t\right) \pm \sigma\) to give the upper and lower limits of the band.  These limits match the pink 25\% occurrence contour well because all the points in the pink contour are approximately the same distance away from the central fit (white solid line). The band width (BW) could then be defined as \(2\sigma\). The fit parameters and band width are given in the top-right corner of Figures \ref{fig:month_avg_occ}a--d. For the rest of this study, ``echoes within the band'' refer to data points that fall between the upper and lower white dashed lines. From Figures \ref{fig:month_avg_occ}a--d, both the center position and upper/lower bounds for the band shown by the white lines capture the band and its diurnal variation in range very well for all months. Therefore, these estimates are used below to quantify the echo band behavior, particularly in context of expected electron densities.

MCM is oriented in such a way that it samples daytime (nighttime) ionosphere in roughly the first (second) half of the day in UT. The electron density on the right of Figure \ref{fig:month_avg_occ} clearly shows higher daytime densities that reach a peak near 05 UT at farther ranges and lower nighttime densities that reach a minimum near 17 UT. The \(F\)-region echo band thus was closer to the radar in the daytime (high densities) and farther away at night (low densities). This feature is most clear for the March and September datasets. It is much less clear in June, although one can still see a portion of the cosine-like variation at 02--10 UT in June in the pink contour. In this month, the white contours are mostly extrapolation outside of this interval, but they fit into the overall diurnal/seasonal pattern well (see also next analysis).  In December, the band is flat and diurnal variation is not evident.

In terms of diurnal variation of echo occurrence, there was a minimum occurrence at nighttime and a maximum in daytime hours. The minimum was much deeper and more extended in June so that essentially no echoes were observed at 12--24 UT. At this time, the expected location of echoes based on the cosine-fit model (white lines) is farther away from MCM than in any month and any UT interval, but no echoes are observed either at these ranges or any other ranges. This is in contrast with March and September, where occurrence is reduced at nighttime but not to zero. Overall, Figure \ref{fig:month_avg_occ} shows that range position of the band is largely controlled by the density conditions, while the echo occurrence variation is perhaps more complicated.

The final observation about Figure \ref{fig:month_avg_occ} concerns the results of raytracing simulations that were conducted for each hour of day 15 in each of the four months shown in Figure \ref{fig:month_avg_occ}. In these simulations, IRI densities in a 2D simulation domain aligned with beam 14 of MCM were used. The range locations where orthogonality condition was expected are plotted in Figures \ref{fig:month_avg_occ}e--h by the dark red line. The gaps during the nighttime in these lines in all months but December indicate that no orthogonality was expected at any range in MCM beam 14 during these periods. The red line is located between the solid and bottom dashed white lines, indicating a reasonable agreement between the expected and observed backscatter locations. The important exception is seen, however, for nighttime periods (or, more accurately, whenever densities were lower than \(\sim1.5\times10^5\) cm\(^{-3}\), dark blue colors). During those periods, no orthogonality was expected based on IRI densities, but significant (albeit reduced) amount of backscatter was observed in March and September. This implies a significant source of nighttime plasma density that is unaccounted for by the IRI model.

The same analysis was performed for the entire period of interest 2010--2013 and the band characteristics are presented in Figure \ref{fig:year_line}. It shows the band (a) position amplitude \(A\), (b) time shift \(B\), (c) range position \(C\), and (d) width (BW) plotted versus month.  Each year is shown as a line in a different color as indicated in the top-right corner of panel (a).

\begin{figure}
\includegraphics[width=11.0 cm,angle=90]{year_month_analysis_line.pdf}
\caption{The variations in the echo band (a) amplitude of the fitted cosine function \(A\), (b) time shift or time of day of the peak \(B\), (c) average position of the band or the vertical shift of the cosine function from zero \(C\), and (d) width (BW). Different lines are for different years with the legend given in Figure \ref{fig:year_line}a.}
\label{fig:year_line}
\end{figure}

All four parameters peak in austral winter and decrease in summer months. Solar cycle variation is evident in (a) the band position amplitude and (c) the average band position. In general, both parameters were larger in years closer to solar minimum (year 2010) and decreased moving towards solar maximum (2013).  In both cases, the variation existed primarily in the winter months (around June). In contrast, the seasonal variation was also evident in (b) the band time shift and (d) the band width, but no clear solar cycle trend was present. In fact, the band time shift was quite consistent between different years, except for winter months. This parameter was typically around 15 hrs. This is when a maximum of \(R\left(t\right)\) was observed in UT, i.e. when band position was located furthest away from the radar.

In the previous analysis, the focus was on band characteristics related to range and their variations. The next presentation examines the diurnal/seasonal/solar cycle pattern as a whole for both band range position and one other important characteristic, the echo occurrence within the band. For each month, the band position \(R\left(t\right)\) and the average occurrence between \(R\left(t\right)\pm \sigma\) were found similar to those presented in Figure \ref{fig:month_avg_occ}. A horizontal slice of each panel of Figure \ref{fig:year_color} represents the time series for one month. The band position \(R\left(t\right)\) (average echo occurrence) is shown in Figures \ref{fig:year_color}a--\ref{fig:year_color}d (Figures \ref{fig:year_color}e--\ref{fig:year_color}h) as a color contour plot representing the time series for all months of the year simultaneously; different rows refer to  different years. The left panels look much smoother than the right ones; this is simply because each line on the left is a fitted cosine \(R\left(t\right)\), while on the right it is the average ``raw'' occurrence between two fitted cosine lines \(R\left(t\right)\pm \sigma\). The white data gaps at the bottom of Figures \ref{fig:year_color}d and \ref{fig:year_color}h reflect the fact that no MCM data were collected in January 2010.

\begin{figure}
\includegraphics[width=12.0 cm,angle=90]{year_month_analysis_color.pdf}
\caption{Contour plots of the (a)--(d) echo band range and (e)--(h) occurrence versus UT and season. Each row shows the data for a different year as given in the top-right corner.}
\label{fig:year_color}
\end{figure}

In Figures \ref{fig:year_color}a--d, the band position \(R\left(t\right)\) peaks in nighttime (second half of the day) and in winter. This peak (yellow-to-red colors) is seen in the same plot area as the low occurrence plot cells in Figure \ref{fig:year_color}e--h (blue color). The amplitude of the range peak became consistently lower between 2010--2013.  This indicates that the high occurrence band moved closer to the radar for years further away from the solar minimum.

The minima in the band position \(R\left(t\right)\) (blue contours) were observed during the daytime and near equinoxes. This feature was most clear in 2010 when these minima were observed near 03 UT in March and September; these were also the two months presented in Figure \ref{fig:month_avg_occ}. The echo occurrence pattern, Figures \ref{fig:year_color}e--h, is less clear in terms of its maxima and their relationship with range minima. Typically, one equinoctial maximum was more pronounced as in 2011--2013, with no clear preference towards either March of September. In 2010, the two maxima are most similar (yellow contours), but the UT peak was shifted away from that of the range minimum (by about 5 hours). Overall, however, the range and occurrence patterns appear to be inverse of each other: larger range means smaller occurrence and vice versa.

The diurnal, seasonal, and solar-cycle variations in echo occurrence observed in Figures \ref{fig:month_avg_occ}--\ref{fig:year_color} can be explained by looking at the solar zenith angle (SZA) and ionospheric density conditions. SZA indicates how sunlit the ionosphere is and the associated degree of ionization, i.e. the \(E\)- and \(F\)-region electron densities, as explained in Section \ref{Sec:Intro}. The model NmE and NmF2 values have been employed for this analysis, Section \ref{Sec:Exp}. To keep model run times at a manageable level, here we consider 10-min cadence data including the average echo occurrence within the band (in Figures \ref{fig:month_avg_occ}--\ref{fig:year_color}, the 2-min data were considered).

Figures \ref{fig:scatter}a and \ref{fig:scatter}b present the results of this analysis for year 2010 in the form of scatter plots of band position versus (a) SZA and (b) model NmF2, while Figures \ref{fig:scatter}c and \ref{fig:scatter}d are scatter plots of echo occurrence versus (c) SZA and (d) model NmF2. In all four panels, the model NmE is shown by the color of the points, with the color bar given to the right. The points in the top row show some clustering in the form of closed loops (some of them appear just as horizontal lines). This is simply a consequence of the periodic nature of the cosine fit; each loop refers to the data for one month. The summer months exhibit little-to-no variation; these appear as horizontal loops-lines.

\begin{figure}
\includegraphics[width=17.0 cm]{scatter_E2010.pdf}
\caption{The echo band range versus the (a) solar zenith angle (SZA) and (b) model peak F2 electron density (NmF2). Each point is color-coded in model \(E\) region density. Figures \ref{fig:scatter}c and d show the same but for the  echo occurrence within the band.}
\label{fig:scatter}
\end{figure}

Figure \ref{fig:scatter}a shows a group of blue points referring to nighttime observations (SZA \(>\) 90\(\deg\)) that slope upwards from a band position range of \(\sim\)800 km to \(\sim\)2000 km with a SZA increasing. This indicates that the band moved into further ranges as radar sampled sectors farther away from sunset. The color of the points indicates that \(E\)-region electron density was typically lower at these times as well, as expected. In Figure \ref{fig:scatter}b, the same trend is seen as blue points starting at intermediate NmF2 values of \(\sim1.5 \times 10^5\) cm\(^{-3}\) and low ranges near 800 km, but increasing to ranges of \(\sim\)2000 km as NmF2 decreased to low values of \(\sim0.5 \times 10^5\) cm\(^{-3}\). This indicates that both representations (using SZA or model NmF2) are approximately equivalent: the band was further from the radar at low NmF2/high SZA values. At high NmF2/high NmE/low SZA values, i.e. during daytime conditions, the range did not change much. In summer months (horizontal loops mentioned above) in particular, this variation was virtually nonexistent. Overall, Figures \ref{fig:scatter}a and \ref{fig:scatter}b show a clearly-defined pattern of a decreasing band position \(R\left(t\right)\) with \(t\) moving away from sunset and SZA increasing from 90\(\deg\), NmF2 decreasing and NmE also decreasing.

The echo occurrence within the band shows a different pattern, Figures \ref{fig:scatter}c and \ref{fig:scatter}d. Occurrence increases towards SZA \(\approx90\deg\) during daytime, but then starts to decrease away from this value (Figure \ref{fig:scatter}c). The $E$-region density, on the other hand, appears to decrease monotonically with a SZA increasing. It has a peak at \(10 \times 10^4\) cm\(^{-3}\) at the smallest possible values of SZA \(\sim50\deg\), while occurrence peaks around 90\(\deg\). This means that echoes are most likely to be observed close to the terminator. There appears to be a similar peak when occurrence is plotted against \(F\)-region electron density (Figure \ref{fig:scatter}d). The NmF2 value corresponding to occurrence peak is between \(1.0\times 10^5\)--\(2.5\times 10^5\) cm\(^{-3}\) depending on the month.  At both lower and higher \(F\)-region densities, occurrence decreases. This indicates that there is a range \(F\)-region electron densities that favors the observation of echoes. In both Figures \ref{fig:scatter}c and \ref{fig:scatter}d, points where the \(E\)-region density is high (dark red points) have relatively low occurrence.  Based on Figure \ref{fig:scatter}c, these generally have the lowest occurrence of all daytime measurements.  Overall, Figures \ref{fig:scatter}c and \ref{fig:scatter}d show that echo occurrence peaks near the terminator where NmF2 is moderate and decreases for SZA increasing and decreasing from 90\(\deg\).

The IMF variations have been previously shown to affect HF radar observations of the auroral ionosphere in both the \(F\) region \citep{Ballatore2001} and \(E\) region \citep{Makarevich2012}. The IMF variations are a consequence of solar activity and are not related to the time of day or season, so they must be studied independently. IMF activity is also related to solar cycle, but the correlation is on a time scale of about a decade. In the following analysis, we attempt to determine whether there is a direct relationship between IMF behavior and echo occurrence on the scale of a few hours. We compare the MCM echo occurrence variations with the simultaneous IMF observations using a method analogous to that employed by \citet{Makarevich2012} in their study of auroral \(E\)-region irregularities (see also our section \ref{Sec:Exp} for a brief description of IMF data analysis).

The average occurrence within the band was found at 10-min intervals as in the previous analysis. The average IMF components over corresponding 10-min intervals were also found. Figure \ref{fig:ut_occ} presents the entire 4-year dataset of echo occurrence versus UT, with color representing average IMF component over each plot bin. The bin sizes were 1-h in time and 2\% in occurrence and Figure \ref{fig:ut_occ} shows contour plots made using this grid. The color bar for all three IMF components is shown to the right.

\begin{figure}
\includegraphics[width=15.0 cm]{IMF_vs_occ_MLT_avg.pdf}
\caption{IMF effects in echo occurrence. The echo occurrence versus UT was binned in 1-hr bins in UT and 2\% bins in occurrence and the average IMF (a) Bx, (b) By, and (c) Bz components over each plot cell are shown by the color.}
\label{fig:ut_occ}
\end{figure}

Figure \ref{fig:ut_occ}a and \ref{fig:ut_occ}c show minimal color structure in the first half of the day (0--12 UT; daytime observations). This indicates that in this time sector, neither positive nor negative IMF Bx or Bz was conducive to high echo occurrence. The dark blue region between 20\%--60\% occurrence and 0--10 UT of Figure \ref{fig:ut_occ}b (By component) indicates some association between low occurrence and negative By. The second half of the day (12--24; nighttime observations) exhibits a definite color structure for all three IMF components. Figure \ref{fig:ut_occ}a has an orange and yellow area between 18--21 UT at above 60\% occurrence. This means that in this time sector, high echo occurrence was observed during predominantly positive IMF Bx. Similarly, in Figure \ref{fig:ut_occ}b, the large blue area between 12--24 UT and above 50\% indicates an association between high occurrence and negative IMF By. Moreover, Figure \ref{fig:ut_occ}b shows that the color changes roughly monotonically from orange and yellow to green and blue with increasing occurrence. This means that higher occurrence was observed for more negative IMF By. A similar color structure is less clear in Figure \ref{fig:ut_occ}c (Bz  component), but still present. For example, Figure \ref{fig:ut_occ}c has a blue section above 70\% occurrence and between 16--24 UT, so high occurrence is also associated with negative IMF Bz in this sector. A similar change in color (from yellow to blue) with an increase in occurrence is also observed here, although it is not as monotonic (color pattern is somewhat patchy). Overall, this analysis indicates that higher nighttime occurrence is seen during periods of more positive Bx, negative By, and negative Bz.

Figure \ref{fig:by_bz} contains the same IMF data as Figure \ref{fig:ut_occ}, but here the IMF By and Bz effects are compared directly. IMF By and Bz are plotted on the horizontal and vertical axes, respectively, and color indicates relative occurrence, which was found as explained below. First, the average echo occurrence within the band over the whole month was found at 10-min intervals (these values were previously presented in Figure \ref{fig:year_color}, but at 2-min resolution). This time series was then subtracted from the average echo occurrence within the band for each day of the month (no averaging over the month). This resulted in a time series that approximately represents the variation in echo occurrence relative to a quiet-day season-specific diurnal average or the relative occurrence. This approach allows us to isolate the IMF effects from those due to diurnal and seasonal variations \citep{Makarevich12}. Points were divided into six groups based on UT and plotted in different panels of Figure \ref{fig:by_bz}.

\begin{figure}
\includegraphics[width=11.0 cm,angle=90]{occ_vs_IMF_avg.pdf}
\caption{Relative occurrence versus IMF By and Bz components for 6 different UT sectors. Each plot cell is color-coded based on the average relative occurrence over that cell.}
\label{fig:by_bz}
\end{figure}

Figures \ref{fig:by_bz}a--\ref{fig:by_bz}c (representing 0--12 UT) show very little color structure.  Each panel is uniformly patchy, indicating no preference towards relatively high or relatively low echo occurrence for any IMF conditions. This is consistent with the minimal color structure observed in the 0--12 UT interval in Figure \ref{fig:ut_occ}. In the second half of the day (Figure \ref{fig:by_bz}d--\ref{fig:by_bz}f), the left half of each panel is more red and orange, while the right half is more blue. In particular, Figure \ref{fig:by_bz}d has a peak in relative occurrence (orange) at (\(-3, -2\)) nT in the (By, Bz) coordinates and an occurrence trough (blue) at (\(+5, -3\)) nT. Figure \ref{fig:by_bz}e has a region of high relative occurrence between (\(-\)4--0, \(-\)8--0) nT and a region of low relative occurrence between 2--10 nT in IMF By.  Figure \ref{fig:by_bz}f generally has high relative occurrence in the bottom-left corner of the panel and low relative occurrence in the top-right corner of the panel. This means that high (low) relative occurrence is observed for periods of negative (positive) IMF By and Bz. Comparing Figures \ref{fig:by_bz}d--\ref{fig:by_bz}f, the dividing line between the red/orange areas and the blue area is close to vertical in Figure \ref{fig:by_bz}e (along By \(\sim\)0 nT), but rotates to be at almost 135\(\deg\) in Figure \ref{fig:by_bz}f (stretching between the top-left and bottom-right corners). Overall, these results indicate increasing importance of the Bz component in the last time sector, namely, positive Bz leads to negative relative occurrence. These results also agree with those shown in Figure \ref{fig:ut_occ}, where the patch of blue (negative Bz) was seen between 16--24 UT and above 70\% occurrence.

\section{Discussion}

In this study, a comprehensive data set collected by the SuperDARN McMurdo radar over 4 years has been employed to investigate formation of small-scale irregularities in the polar \(F\) region, with a particular focus on global factors. MCM routinely observes backscatter from ionospheric irregularities over extended time periods and large spatial areas within the southern polar cap (``persistent backscatter''), which allowed us to reliably quantify both backscatter location and occurrence characteristics and investigate their control by the solar wind and illumination. By considering both location and occurrence we achieved better understanding and differentiation between radio- and plasma-physical factors or, alternatively, between irregularity detection and production. Below we further discuss three groups of issues aligned with three objectives of this study.

\subsection{\(F\) Region Echo Band: A Comprehensive Model and Density Implications}

Previous investigations of small-scale irregularity formation in the polar cap \(F\) region have primarily used SuperDARN coherent radars in both northern and southern polar caps \citep{Bristow2011,Ghezelbash2014b,Koustov2014}. These observations revealed unusually high echo occurrence rates of \(F\) region backscatter that can exceeded 80\% \citep{Bristow2011,Ghezelbash2014b}. These polar cap occurrence rates significantly exceed similarly-derived occurrence rates at auroral latitudes where the highest reported values were near 60\% under much stronger solar activity conditions \citep[e.g.][]{Koustov2004}.

One possible reason for this difference is more favorable propagation conditions in the polar cap with greater chances to reach orthogonality with magnetic field, which is a necessary condition for Bragg scatter to be received by coherent scatter radars. Whether any observed occurrence trend is due to changing propagation or production is difficult to resolve based on HF radar echo occurrence observations alone, as explained in section \ref{Sec:Intro}. A reasonable assumption often utilized is that propagation will be mostly responsible for changing location of echo occurrence. Typically, the entire FoV (or its large portion) is considered in calculating occurrence rates and, assuming range variations stay within the FoV, occurrence variability will be mostly due to changing production \citep[e.g.][]{Kane2010}. A considerable improvement over this approach is to explicitly consider the trends in backscatter location as well \citep{Kane2012,Ghezelbash2014b}. This way, any variation in occurrence and location is most likely due to a production factor. The current study refines this approach by including the location information explicitly in the way occurrence estimates are obtained. Further, the information about variations in the typical ranges of backscatter is important by itself, particularly since it directly reflects variations in the \(F\) region density. This provides a useful (albeit somewhat indirect) way to monitor and analyze plasma production and transport processes in a region where few opportunities for direct continuous measurements of plasma density currently exist.

The two important outcomes of the current study related to this indirect density analysis were (1) development of an empirical model to describe location of \(F\) region backscatter and (2) analysis of characteristics derived from the model. The \(F\) region backscatter was predominantly observed in the form of a band that was well described by a simple model, equation (\ref{E:Model}). The \(F\) region echo band exhibited systematic diurnal and seasonal effects, Figure \ref{fig:month_avg_occ}, which was reflected in systematic changes of the fitted band characteristics \(A\), \(B\), \(C\), and band width (BW), Figure \ref{fig:year_line}. These changes were quite consistent with those expected for plasma density. For example, the echo band was moving to further ranges during the night while density expected from IRI model was decreasing, Figures \ref{fig:month_avg_occ}e--h. Similarly, the amplitude of range variations \(A\) and average position \(C\) of the echo band exhibited seasonal and solar cycle variations consistent with those expected for densities, Figures \ref{fig:year_line}a and c.

Another novel aspect of the current study was a comprehensive set of raytracing simulations that was conducted using IRI densities. In general, the diurnal and seasonal variations were very similar between simulations (red line in Figures \ref{fig:month_avg_occ}e--h) and observations (white lines). One important exception was observed during the night near equinoxes. No orthogonality was predicted by model simulations, whereas a significant (albeit reduced) amount of backscatter was observed, Figures \ref{fig:month_avg_occ}a and c. That is, an occurrence gap (i.e. no backscatter) similar to what was observed in austral winter (Figure \ref{fig:month_avg_occ}b) was expected but not observed.

This result is significant as it implies a considerable additional source of density which would provide additional refraction and bring backscatter to orthogonality at the ranges consistent with the model. This prompts an important question about the origin of this source. It is unlikely that it was local since this production is well described by IRI \citep[e.g.][]{Themens2014}. A reasonable assumption is then that additional density was transported from somewhere else. The two phenomena that can be responsible for this are polar patches \citep{Weber1984} and sun-aligned arcs \citep{Valladares1994}. The polar patches in particular are known to travel long distances across the polar cap and even well into the nightside auroral zone \citep{Moen2007,Oksavick2010}.

Direct density measurements are not available near MCM but more insights can be obtained from observations in the conjugate northern polar cap. Figure \ref{fig:patchy_example} shows an example of conjugate measurements during two selected events that fall within the equinoctial period of Figure \ref{fig:month_avg_occ}a. Shown are (a) IMF By and Bz components, (b) MCM line-of-sight velocity in beam 14 versus UT and range, (c) RKN velocity in nominally-conjugate beam 4, and (d) northern face of the Resolute Bay incoherent scatter radar (RISR-N) measurements of electron density versus UT and altitude on March 14, 2010. Figures \ref{fig:patchy_example}e--h present the same information but for a second event on April 6, 2010.

\begin{figure}
\includegraphics[width=18.0 cm]{Fig_RISR.pdf}
\caption{Conjugate measurements for two selected events on (left) March 14, 2010 and (right) April 6, 2010. Shown are the (a) and (e) IMF By (blue) and Bz (red) components, (b) and (f) MCM line-of-sight velocity versus UT and slant range in beam 14, (c) and (g) RKN velocity in beam 4, and (d) and (h) RISR-N electron density versus UT and altitude. The color bar is the same for the two middle rows and is given to the right of Figure \ref{fig:patchy_example}f.}
\label{fig:patchy_example}
\end{figure}

Figure \ref{fig:patchy_example}b shows an echo band that fits well with the average pattern of Figure \ref{fig:month_avg_occ}a (note more limited ranges for Figure \ref{fig:patchy_example}). This is a clear example of ``persistent backscatter'' at MCM where echoes are present throughout the entire day with only range changing in agreement with the model of equation (\ref{E:Model}). In contrast, Figure \ref{fig:patchy_example}f shows an example of ``patchy backscatter'' at MCM, particularly after 12 UT. The range variation is still roughly consistent with the model. The RKN backscatter pattern looks similar if one accounts for a \(\sim\)12-h  shift in UT between day- and nighttime observations. Thus nighttime echoes are observed by RKN near 08 UT during the first event (it is daytime at MCM) and they are more persistent. The daytime echoes are seen by RKN at closer ranges near 18 UT and they are also persistent, similar to MCM daytime echoes near 08 UT. During the second event, very few nighttime echoes are observed by RKN and those that are seen (near 02 UT) are quite patchy.

The key information about density is provided by RISR-N observations, Figures \ref{fig:patchy_example}d and h. A diurnal variation is evident on both days, with a broad nighttime minimum centered near 06--08 UT and a maximum centered at 18--20 UT. Significant perturbations on a background of this diurnal variation are also evident, e.g. relatively large densities near 06 UT at 300 km in Figure \ref{fig:patchy_example}d. These perturbations agree quite well with RKN observations in that when densities are enhanced so is backscatter. For example, lower densities are seen near 09 UT and this corresponds to a gap in echoes at 1600 km. Nighttime perturbations are weaker during the second event and this is a probable cause for absence of echoes during the night. Nighttime perturbations are also quite patchy in appearance for both events. The magnitude of perturbations is also comparable with background daytime densities observed at later UT. The RISR-N observations are thus consistent with polar patches being additional source of nighttime densities.

The IMF information presented in Figures \ref{fig:patchy_example}a and b also provides a clue into what conditions are associated with patchy backscatter. When IMF is relatively stable on a time scale of \(\sim\)20 min, as is the case in the first event, backscatter is persistent. When IMF exhibits strong perturbations in By with a period of \(\sim\)10 min, as is near 16 UT in the second event, nighttime backscatter is patchy. Daytime backscatter seen by RKN during the same interval does not seem to be affected, i.e. it is still persistent. Again, this is supportive of the polar patch scenario.

The presented statistical analysis of MCM observations thus imply that not only polar patches contribute to the nighttime densities but also that they do so on a regular basis. This causes the differences between night- and daytime densities implied by the observed backscatter ranges to be much smaller as compared with what is predicted by model raytracing predictions.

\subsection{\(F\) Region Irregularities: Suppression and Enhancement by Sunlight}

The two ideas about the role of solar illumination refer to (1) backscatter suppression through smoothing of density gradients that are needed for GDI \citep{Ruohoniemi1997,Koustov2004} and (2) backscatter enhancement by more favorable propagation conditions \citep{Bristow2011,Koustov2014}. In this section it is argued that these two ideas are not necessarily mutually exclusive and that both processes probably occur even in the same latitudinal domain.

Figure \ref{fig:month_ut} summarized observations and model results related to this issue for year 2010. It presents (a) MCM backscatter occurrence, (b) MCM backscatter range \(R(t)\), (c) model NmF2, and (d) SZA. Figures \ref{fig:month_ut}a and b are the same as Figures \ref{fig:year_color}d and h except that here 10-day (rather than monthly) occurrence averages are used and \(5\times5\)-window smoothing applied for occurrence. Figures \ref{fig:month_ut}c and d are remarkably similar in their contour patterns; one panel is an inversion of the other one, as expected for model NmF2 that is mostly SZA-controlled. The range panel (b) is also very similar to both, although here minima (darkest blue) are somewhat shifted in season from SZA minima/NmF2 maxima. This is not surprising since range variation represents changing propagation conditions and is mostly due to ionization locally-produced by the Sun.

The MCM occurrence shown in Figure \ref{fig:month_ut}a is also similar to other 3 panels for nighttime observations (deep occurrence minimum represented by blue colors in Figure \ref{fig:month_ut}a). However, there are both similarities and differences in daytime observations. Backscatter is enhanced in the dusk and dawn sectors (yellow and orange contours between 00 and 12 UT in Figure \ref{fig:month_ut}a). This is when density is also enhanced as compared to nighttime in Figure \ref{fig:month_ut}c due to some illumination at SZAs of 70$^\circ$--90$^\circ$ in Figure \ref{fig:month_ut}d. However, backscatter suppression by sunlight is also evident as a secondary occurrence minimum at 02 UT in austral summer coincides with the main NmF2 maximum/SZA minimum. Thus both effects/processes contribute to occurrence patterns.

\begin{figure}
\includegraphics[width=12.0 cm,angle=90]{month_ut_combined2010.pdf}
\caption{Contour plots of the (a) average occurrence within the band, (b) band range, (c) model peak F-region electron density (NmF2) and (d) solar zenith angle (SZA) versus UT and month in 2010. Color bars are given to the right. The white rectangle at the bottom of each panel is due to no data available in January 2010.}
\label{fig:month_ut}
\end{figure}

A different view of occurrence dependencies on NmF2, SZA, and NmE was offered in Figure \ref{fig:scatter} in a more explicit form. The range and occurrence dependencies were found to be quite different, which, as discussed above, indicates that propagation effects are not dominant in occurrence variations. This is in contrast with the previous study by \citet{Kane2012} that conducted separate analyses for night- and daytime observations. The current approach resulted in a single and well-defined pattern that has a clear occurrence peak near a SZA of 90\(\deg\), i.e. near the solar terminator. This is also in contrast with previous studies where a qualitative agreement was observed between contours of constant occurrence and SZA \citep{Kane12,Ghezelbesh2014b}. Occurrence decreases on both sides of SZA of 90\(\deg\) but most likely for different reasons.

A decrease at smaller SZA is consistent with the idea of sunlight wiping out gradients necessary for GDI to operate. Increase in NmE resulting in shorting of \(F\) region irregularities by the conducting \(E\) layer may have also contributed. A decrease at larger SZA, on the other hand, is associated with a decreasing NmF2, increasing range, and stronger under-refraction effects. Importantly, however, not all of this decrease is due to stronger under-refraction. This is because range was increasing, but still well within the radar's FoV, Figure \ref{fig:scatter}a. That is, at the ranges where echoes were expected based on range trends, echoes were present but reduced in occurrence, Figure \ref{fig:month_avg_occ}a and c. Thus an alternative explanation must be sought for an occurrence decrease at large SZA. This could be due the scattering cross-section being effectively proportional to the square of background density, a plasma-physical effect described by \citet{Makarevich2014b}. In addition, the role of gradients near the solar terminator will also diminish.

A clear occurrence peak near the solar terminator is a strong indication of importance of large-scale gradients in production of small-scale irregularities. A traditional way of thinking about this disparity between scales is in terms of turbulent cascade from large-scale structures to smaller irregularities \citep{Tsunoda1988}. In this process, large-scale structures can become unstable due to GDI and other structuring mechanisms and act as seeds for smaller-scale structures. The current observations thus indicate that large-scale density gradients associated with the solar terminator can be a significant source of seeding structures.

\subsection{Irregularity Production in the IMF-Controlled Convection Environment}

The current study has also presented the first investigation of indirect solar effects due to IMF variations on irregularity production in the polar cap. It is well established that IMF variations directly control many high-latitude phenomena including auroral electrojets \citep[e.g.][]{McPherron1988}, plasma convection patterns \citep{Heelis1984,Ruohoniemi1996,Ridley1998,Pettigrew2010} and occurrence of optical auroras \citep{Ostgaard2004}. In contrast, only two studies have investigated IMF control of small-scale irregularity production and these focused on observations in the auroral \(F\) region \citep{Ballatore2001} and auroral \(E\) region \citep{Makarevich2012}.

This study found no clear evidence on such a control in the dayside polar cap, Figures \ref{fig:ut_occ} and \ref{fig:by_bz}. On the nightside (12--24 UT), however, a clear pattern emerged where echo occurrence was enhanced for negative IMF By and Bz, Figure \ref{fig:ut_occ}. The pattern for IMF Bx was nearly an inversion of that for IMF By, implying a high degree of correlation between IMF By and \(-\)Bx. Interestingly, however, the IMF Bz pattern was somewhat different from that of IMF By, Figure \ref{fig:ut_occ}, and in fact less clearly defined. Nevertheless, occurrence was generally higher for more negative IMF Bz, Figure \ref{fig:ut_occ}c. This result is consistent with the study by \citet{Ballatore2001} that also found higher occurrence for negative IMF Bz in the auroral zone. The new and unexpected result of the current study was that IMF By effects were more clearly defined in the occurrence of backscatter from the polar cap. Another new result of the current study was that the occurrence dependence on IMF was changing in a systematic way with UT on the nightside, Figures \ref{fig:ut_occ} and \ref{fig:by_bz}. For example, near 12 UT occurrence was generally above 80\% when IMF By was on average below \(-\)1.5 nT (blue cells at the top of Figure \ref{fig:ut_occ}b). On the other hand, for the same IMF By, but at a later UT, occurrence was between 60\% and 100\%, i.e. blue area of the plot expanded lower in Figure \ref{fig:ut_occ}b.

The echo occurrence dependence on IMF in general can be interpreted in context of IMF-controlled convection environment. High-latitude convection patterns are directly controlled by solar wind variations. Statistical convection patterns are derived from satellites and radars and are routinely parameterized by IMF By and Bz \citep{Haaland2007, Pettigrew2010}. One mechanism through which IMF-controlled variations in convection may be responsible for changes in occurrence of plasma irregularities involves changing orientations between plasma density gradients \(\vec{G}\) and convection velocity \(\vec{V}_E\) vectors. GDI can more easily excite waves when a component \(\vec{G}\cdot\vec{V}_E\) is more positive, with \(\vec{G}\parallel\vec{V}_E\) being most favorable orientation \citep{Kesknein1982a,Makarevich2014c}. IMF-controlled changes in \(\vec{V}_E\) can thus potentially lead to changes in \(\vec{G}\cdot\vec{V}_E\), which, in turn, will change GDI growth rate and occurrence of GDI-produced irregularities. This is a feasible basic scenario, although some questions still remain as to the extent of its contribution to the trends observed and reported here. For example, gradient scales must be relatively small for GDI to contribute to decameter-scale irregularity generation, with one widely-accepted possibility of large-scale gradient structures serving as seeds for small-scale structuring processes \citep{Tsunoda1988}. In this context, the presented results suggest that time scales at which convection changes are shorter than those associated with breakup of large-scale structures and changes in prevalent gradient orientations at both large and small spatial scales. This is consistent with the idea of very fast propagation of convection disturbances \citep{Ridley1998,Ruohoniemi2002,Fiori2012,Taguchi2015}.

\section{Summary and Conclusions}

Statistical analysis of backscatter from small-scale plasma irregularities observed by the SuperDARN McMurdo radar in the southern polar cap indicate that solar control of small-scale irregularity production in the polar cap is strong, multifaceted, and not restricted to the dayside. It exists in the form of (1) direct effects of local solar ionization that produce density gradients near the solar terminator and wipe out gradients on the dayside, (2) dayside ionization transported to the nightside as polar patches that routinely improve propagation conditions, and (3) solar wind variations that are responsible for changes in plasma convection and density gradient environment. Other important findings are as follows.

1. Backscatter location can be described reasonably well by using a simple model involving a cosine fit to diurnal variation in echo band position. The derived fit characteristics of the echo band display a clear pattern of diurnal, seasonal, and solar cycle changes that is very similar to those of the solar zenith angle and model peak \(F\) region density. The band range position is, therefore, strongly controlled by the local solar illumination conditions in the polar cap. On the nightside, however, a comparison between observations and model raytracing simulations reveal a presence of significant additional source of ionization. Conjugate observations of backscatter and density indicate that polar patches transported from the dayside are a viable source for this ionization.

2. Backscatter occurrence statistic exhibits evidence of both suppression and enhancement by sunlight, with a clear occurrence peak observed near the solar terminator. Suppression occurs at the largest SZA values observed and is consistent with diminishing role of gradients in irregularity production and increasing role of conducting $E$ layer. Enhancement by sunlight is mostly associated with more favorable propagation conditions at intermediate SZA values and large-scale gradients near the solar terminator possibly serving as seeds for further structuring processes.

3. Indirect control by the solar wind variations is limited to the nightside, with enhanced backscatter occurrence observed during intervals of negative IMF By and Bz. The IMF control of small-scale irregularity production can occur through IMF-controlled variations in the high-latitude convection pattern that, in turn, lead to changes in mutual orientations between convection and prevalent gradient vectors, with plasma structuring favoring certain orientations.

%\acknowledgments This work was supported by NSF grants AGS-1248127 and AGS-1341902. SuperDARN data are freely available through the SuperDARN website at Virginia Polytechnic Institute and State University http://vt.superdarn.org/. Fortran source code for IRI 2007 is freely available for download from the Community Coordinated Modeling Center (CCMC) at http://ccmc.gsfc.nasa.gov/. OMNI data is available from NASA Goddard Space Flight Center Space Physics Data Facility at http://omniweb.gsfc.nasa.gov. The Resolute Bay incoherent scatter radar (North) is operated by SRI International on behalf of the US National Science Foundation under NSF Cooperative Agreement AGS-1133009. The data are accessible from the SRI International online database at http://amisr.sri.com/database/. The authors thank W. A. Bristow for useful suggestions and providing the original Fortran source code for raytracing simulations. The developed IDL raytracing code is available from the corresponding author.

%\newpage
\bibliographystyle{uafthesis}
\bibliography{references}

%\end{article}
%\end{document}

%3 points:

%Solar control of irregularity production is not restricted to dayside
%Production peak near solar terminator implies breakup of large-scale structures
%Importance of IMF-controlled convection orientations relative to gradients

### 2017-01-15 thesis.sh

@@ -2,17 +2,25 @@
 #!/bin/bash
 
 gs -q -dSAFER -dBATCH -dNOPAUSE -sDEVICE=pdfwrite -dDEVICEWIDTHPOINTS=530 -dDEVICEHEIGHTPOINTS=410 -sOutputFile=/Users/ljlamarche/Desktop/UAFThesis/Figures/densprofile.pdf /Users/ljlamarche/Desktop/UAFThesis/Figures/densprofile.ps
-
 gs -q -dSAFER -dBATCH -dNOPAUSE -sDEVICE=pdfwrite -dDEVICEWIDTHPOINTS=270 -dDEVICEHEIGHTPOINTS=520 -sOutputFile=/Users/ljlamarche/Desktop/UAFThesis/Figures/SuperDARNmap.pdf /Users/ljlamarche/Desktop/UAFThesis/Figures/SuperDARNmap.ps
-
 gs -q -dSAFER -dBATCH -dNOPAUSE -sDEVICE=pdfwrite -dDEVICEWIDTHPOINTS=510 -dDEVICEHEIGHTPOINTS=510 -sOutputFile=/Users/ljlamarche/Desktop/UAFThesis/Figures/ISRmap.pdf /Users/ljlamarche/Desktop/UAFThesis/Figures/ISRmap.ps
 
+gs -q -dSAFER -dBATCH -dNOPAUSE -sDEVICE=pdfwrite -dDEVICEWIDTHPOINTS=600 -dDEVICEHEIGHTPOINTS=850 -sOutputFile=/Users/ljlamarche/Desktop/UAFThesis/Figures/month_density_occurance2010.pdf /Users/ljlamarche/Desktop/UAFThesis/Figures/month_density_occurance2010.eps
+gs -q -dSAFER -dBATCH -dNOPAUSE -sDEVICE=pdfwrite -dDEVICEWIDTHPOINTS=600 -dDEVICEHEIGHTPOINTS=850 -sOutputFile=/Users/ljlamarche/Desktop/UAFThesis/Figures/year_month_analysis_line.pdf /Users/ljlamarche/Desktop/UAFThesis/Figures/year_month_analysis_line.eps
[…]

### 2017-01-17 introduction.tex
GDI derivation revised

@@
 	\includegraphics[width=\textwidth]{densprofile.pdf}
-	\caption{Altitude profile of electron density throughout the ionosphere found from the IRI model.  Typical ranges of the D, E, and F regions are identified.}
+	\caption{Altitude profile of electron density throughout the ionosphere found from the IRI model.  Typical ranges of the \(D\), \(E\), and \(F\) regions are identified.}
 	\label{fig:densprofile}
@@
 
-Above 50 km, the primary neutral particles are atomic oxygen, O, molecular oxygen, \(O_2\), and molecular nitrogen, \(N_2\).
-Common ions at these altitudes are therefore mostly composed of oxygen and nitrogen, \(O^+, NO^+, O_2^+, N_2^+\), but hydrogen, \(H^+\), can contribute as well, particularly at higher altitudes \citep{Kelley2009}.
+Above 50 km, the primary neutral particles are atomic oxygen, O, molecular oxygen, O\(_2\), and molecular nitrogen, N\(_2\).
+Common ions at these altitudes are therefore mostly composed of oxygen and nitrogen, O\(^+\), NO\(^+\), O\(_2^+\), N\(_2^+\), but hydrogen, H\(^+\), can contribute as well, particularly at higher altitudes \citep{Kelley2009}.
 In the ionosphere, neutral densties are typically several orders of magnitude higher than ion densities, however even a small number of charged particles changes the behavior of the region dramatically.
@@
 In the E region, \(\Omega_e \gg \nu_e\), so electrons are magnetized.
-However, the motion on ions is dominated by collisions between ions and neutrals (\(\nu_i > \Omega_i\)), so ions in the E region are considered collisional.
-In the F region where the neutral density is much lower, both ions and electrons are magnetized, \(\Omega_\alpha \gg \nu_\alpha\).
+However, the motion on ions is dominated by collisions between ions and neutrals (\(\nu_i > \Omega_i\)), so ions in the \(E\) region are considered collisional.
+In the \(F\) region where the neutral density is much lower, both ions and electrons are magnetized, \(\Omega_\alpha \gg \nu_\alpha\).
 These differences in how charged particles move impact the conductivity and plasma waves development, which is further discussed in Section \ref{sec:lit_theory}.
@@
 Polar patches are large density enhancements (usually at least twice the background plasma density) that travel across the polar cap with the background convection \citep{Weber1984,Valladares1994}.
-Polar holes are plasma density depletions that typically occur in the F region slightly poleward of the auroral oval \citep{Benson2001}.
+Polar holes are plasma density depletions that typically occur in the \(F\) region slightly poleward of the auroral oval \citep{Benson2001}.
 Sun-aligned arcs are large density forms that stretch across the polar cap along the sun-earth line.
@@
 When the radar transmits continuously, backscatter will be received, but because it is impossible to know when that signal was transmitted, the difference in time between transmission and receiving cannot be used to calculate how far away the backscatter volume is.
-Conversly, if the radar only transmits a pulse of a certain length of a certain length and then "listens" for its return, the time between transmission and return is known and the distance the pulse traveled can be calculated and therefore the location of the backscattering volume \citep{Farley1972,Greenwald1983}.
+Conversly, if the radar only transmits a pulse of a certain length of a certain length and then ``listens'' for its return, the time between transmission and return is known and the distance the pulse traveled can be calculated and therefore the location of the backscattering volume \citep{Farley1972,Greenwald1983}.
 However, using this simple single-pulse method another pulse cannot be transmitted until the first pulse returns, limiting the temporal resolution that can be achieved.
@@
 In the 1970s and 1980s, the Scandanavian Twin Auroral Radar Experiment (STARE) produced the first continuous large data-set, similar to how modern CSR systems are run.
-STARE consisted of two very high frequency (VHF) radars with overlapping FoVs that were designed to measure FAIs in the E region of the ionosphere \citep{Greenwald1997}.
+STARE consisted of two very high frequency (VHF) radars with overlapping FoVs that were designed to measure FAIs in the \(E\) region of the ionosphere \citep{Greenwald1997}.
 The advantage of two radars with overlaping FoVs was the ability to observe the same structures from two different directions.
@@
 The modern method for creating convection maps involves taking all data recorded by all radars in a particular hemisphere and finding the best fit to a 2D ionospheric electrostatic potential using Legendre functions \citep{Ruohoniemi1998}.
-For convection maps, only data from the F region is considered.
+For convection maps, only data from the \(F\) region is considered.
 If there is not enough velocity data to confine the fit sufficiently, the measured data points are supplemented with a statistical model based on solar wind IMF conditions \citep{Ruohoniemi1995,Ruohoniemi2005}.
@@
 However, they use a fundamentally different technique to observe structuring, which allows them to measure bulk plasma parameters in contrast to small-scale irregularity characteristics.
-While CSRs receive backscatter from relatively large, "coherent" density structures in the ionosphere, ISRs operate by radar scattering from waves due to the random thermal motion of electrons in the plasma.
+While CSRs receive backscatter from relatively large, ``coherent'' density structures in the ionosphere, ISRs operate by radar scattering from waves due to the random thermal motion of electrons in the plasma.
 The radar receivers then measure a spectrum of frequencies with different power \citep{Gordon1958}.
@@
 
-AMISR systems typically have two types of pulses, a long pulse (LP) with 72 km range gates for F-region studies and an alternating code pulse (AC) with 4.5 km range gates gates for E region studies.
+AMISR systems typically have two types of pulses, a long pulse (LP) with 72 km range gates for \(F\)-region studies and an alternating code pulse (AC) with 4.5 km range gates gates for \(E\)-region studies.
 The exact number and configuration of beams depends on the radar mode used and can be changed easily due to the system's electronic steering.
@@
 
-The IRI model produces outputs of electron density, temperatures of electrons, ions, and neutrals, ion composition (O+, H+, He+, O2+, NO+, N+), and total electron content (TEC).
+The IRI model produces outputs of electron density, temperatures of electrons, ions, and neutrals, ion composition (O\(^+\), H\(^+\), He\(^+\), O\(_2^+\), NO\(^+\), N\(^+\)), and total electron content (TEC).
 The model requires the input of a location (latitude, longitude, and altitude) and time (year, date, and time).
@@
 Additionally, the model is capable of producing profiles of any of the output parameters in either space or time.
-This is particularly significant for height profiles, and allows the peak heights and densities of the E and F region to be calculated.
+This is particularly significant for height profiles, and allows the peak heights and densities of the \(E\) and \(F\) region to be calculated.
 
@@
 As discussed previously in Section \ref{sec:ionosphere_regions}, the plasma characteristics change with altitude in the ionosphere, which causes a variety of different instability mechanisms to be operational depending on the region considered.
-The Farley-Buneman instability (FBI), or the modified two-stream instability tends to be a major factor in the E region where ion inertial effects are strong \citep{Farley1963,Buneman1963}.
-Also operational in the E region but more dominant at higher altitudes is the gradient-drift instability (GDI) \citep{Simon1963,Hoh1963,Linson1970}.
+The Farley-Buneman instability (FBI), or the modified two-stream instability tends to be a major factor in the \(E\) region where ion inertial effects are strong \citep{Farley1963,Buneman1963}.
+Also operational in the \(E\) region but more dominant at higher altitudes is the gradient-drift instability (GDI) \citep{Simon1963,Hoh1963,Linson1970}.
 If a wave has a component along the magnetic filed line, the current-conductive instability (CCI) can be operational even if GDI is not \citep{Hoh1960,Ossakow1979,Chaturvedi1981}.
@@
 
-Some of the first 1D investigations of the linear GDI growth rate found that it can be described by a simple expression involving the plasma drift speed, \(V_E = |\vec{E}\times\vec{B}|/B^2\), and the gradient scale length, \(L = |n/(\partial n/\partial x)|\), Equation \ref{eqn:gdi_old} \citep{Simon1963,Hoh1963,Linson1970}
+Some of the first 1D investigations of the linear GDI growth rate found that it can be described by a simple expression involving the plasma drift speed, \(V_E = |\vec{E}\times\vec{B}|/B^2\), and the gradient scale length, \(L^{-1} = |(\partial n/\partial x)/n|\), Equation \ref{eqn:gdi_old} \citep{Simon1963,Hoh1963,Linson1970}
 \begin{equation}
@@
 
-It is useful to express Equation \ref{eqn:gdi_Mak16} in the following equivalent form by defining additional terms.
+By taking certain limiting cases, expressions for the instability growth rate in different region can be derived from Equation \ref{eqn:gdi_Mak16}.
+These expressions are sumarized in Table \ref{tab:gdi_exps}.
+In addition to the quantities previously described in relation to Equation \ref{eqn:gdi_Mak16}, the following are useful for the expressions in Table \ref{tab:gdi_exps}.
 \begin{equation}
-	\label{eqn:gdi_Mak16_alt}
-	\left(D_i-D_e\right)\omega = \vec{V}_d\cdot\vec{k}\left(\frac{Z_2}{Z_1}\right) + \left(D_i\vec{V}_{i0}-D_e\vec{V}_{e0}\right)\cdot\vec{k}+(C_i-C_e)k_{\perp}^2\left(\frac{Z_3}{Z_1}\right)
-\end{equation}
-In the F region where background plasma motion is dominated by electric fields perpendicular to the magnetic field, the differential plasma velocity, \(\vec{V}_d\), can be described as 
-\begin{equation}
-	\label{eqn:gdi_Mak16_Vd}
+\begin{split}
+	b = -\frac{\vec{G}\cdot\vec{k}\times\uvec{b}}{k_\perp^2} \quad\quad\quad
+	y = \frac{k_\parallel}{k_\perp} \\
+	R = \frac{\sigma_H}{\sigma_P} = \frac{D_i+D_e}{1+\Psi} \quad\quad
+	\psi = -r_ir_e \quad\quad
+	\hat{\psi} = \psi\left(1+\frac{k_\parallel^2}{r_e^2 k_\perp^2}\right) \\
 	\vec{V}_d = \vec{V}_{e0}-\vec{V}_{i0} = (r_i-r_e)\left[s_es_i(1+\psi)\left(R\vec{V}_E-\frac{\vec{E}}{B}\right)\right]
-\end{equation}
-where, \(s_\alpha = (1+r_\alpha^2)^{-1}\).  Additionally,
-\begin{equation}
-\label{eqn:gdi_Mak16_Z}
-\begin{split}
-	Z_1 = i(1+Y)+a+Rb+ycK \quad\quad\quad 
-	Z_2 = R(i+a)-b \quad\quad\quad\quad\quad\quad\\
-	Z_3 = -\frac{\Psi(i+a)^2}{1+\Psi}-Rb(i+a)+\frac{b^2}{1+\Psi}-iy(y-ic)\left[\left(K+1-\Psi^{-1}\right)(i+a)+Rb\left(1-\Psi^{-1}\right)\right]\\+Ky^2(y-ic)^2
 \end{split}
 \end{equation}
-and
+
+For the purposes of this thesis, the cold plasma limit expression in Table \ref{tab:gdi_exps} is most recent and can be rewritten using the definition of \(b\), Equation \ref{eqn:gdi} \citep{Makarevich2014c,Makarevich2016a}.
 \begin{equation}
-\begin{split}
-	a& = \frac{\vec{G}\cdot\vec{k}_\perp}{k_\perp^2} \quad\quad\quad
-	b = -\frac{\vec{G}\cdot\vec{k}\times\uvec{b}}{k_\perp^2} \quad\quad\quad
-	c = \frac{\vec{G}\cdot\vec{k}_\parallel}{k_\perp k_\parallel} \\
-	\Psi = -D_iD_e& \quad\quad
-	R = \frac{D_i+D_e}{1+\Psi} \quad\quad
-	y = \frac{k_\parallel}{k_\perp} \quad\quad
-	K = \left(1+\frac{1}{\Psi}\right)\left(1+R^2\right) \quad\quad
-	Y = Ky^2
-\end{split}
-\end{equation}.
+	\label{eqn:gdi}
+	\gamma = \frac{1}{1+\psi}\left(\uvec{k}\cdot\uvec{b}\times\vec{G}\right)\uvec{k}\cdot\left(\frac{\vec{E}}{B}-R\vec{V}_E\right)
+\end{equation}
+Equation \ref{eqn:gdi} is applicable throughout the \(E\) and lower \(F\) regions.
 
-By considering certain limiting cases, an expression that is more relevant to the polar F region specifically can be resolved from Equation \ref{eqn:gdi_Mak16_alt}.
+%It is useful to express Equation \ref{eqn:gdi_Mak16} in the following equivalent form by defining additional terms.
+%\begin{equation}
+%	\label{eqn:gdi_Mak16_alt}
+%	\left(D_i-D_e\right)\omega = \vec{V}_d\cdot\vec{k}\left(\frac{Z_2}{Z_1}\right) + \left(D_i\vec{V}_{i0}-D_e\vec{V}_{e0}\right)\cdot\vec{k}+(C_i-C_e)k_{\perp}^2\left(\frac{Z_3}{Z_1}\right)
+%\end{equation}
+%In the \(F\) region where background plasma motion is dominated by electric fields perpendicular to the magnetic field, the differential plasma velocity, \(\vec{V}_d\), can be described as 
+%\begin{equation}
+%	\label{eqn:gdi_Mak16_Vd}
+%	\vec{V}_d = \vec{V}_{e0}-\vec{V}_{i0} = (r_i-r_e)\left[s_es_i(1+\psi)\left(R\vec{V}_E-\frac{\vec{E}}{B}\right)\right]
+%\end{equation}
+%where, \(s_\alpha = (1+r_\alpha^2)^{-1}\).  Additionally,
+%\begin{equation}
+%\label{eqn:gdi_Mak16_Z}
+%\begin{split}
+%	Z_1 = i(1+Y)+a+Rb+ycK \quad\quad\quad 
+%	Z_2 = R(i+a)-b \quad\quad\quad\quad\quad\quad\\
+%	Z_3 = -\frac{\Psi(i+a)^2}{1+\Psi}-Rb(i+a)+\frac{b^2}{1+\Psi}-iy(y-ic)\left[\left(K+1-\Psi^{-1}\right)(i+a)+Rb\left(1-\Psi^{-1}\right)\right]\\+Ky^2(y-ic)^2
+%\end{split}
+%\end{equation}
+%and
+%\begin{equation}
+%\begin{split}
+%	a& = \frac{\vec{G}\cdot\vec{k}_\perp}{k_\perp^2} \quad\quad\quad
+%	b = -\frac{\vec{G}\cdot\vec{k}\times\uvec{b}}{k_\perp^2} \quad\quad\quad
+%	c = \frac{\vec{G}\cdot\vec{k}_\parallel}{k_\perp k_\parallel} \\
+%	\Psi = -D_iD_e& \quad\quad
+%	R = \frac{D_i+D_e}{1+\Psi} \quad\quad
+%	y = \frac{k_\parallel}{k_\perp} \quad\quad
+%	K = \left(1+\frac{1}{\Psi}\right)\left(1+R^2\right) \quad\quad
+%	Y = Ky^2
+%\end{split}
+%\end{equation}.
+
+%By considering certain limiting cases, an expression that is more relevant to the polar F region specifically can be resolved from Equation \ref{eqn:gdi_Mak16_alt}.
 First, assume a cold plasma such that the temperature of both ions and electrons is negligible.
@@
 If \(D_\alpha\) is expressed as shown below, it can bee seen that these limits result in the first two terms being negligible such that \(D_\alpha \approx r_\alpha\).
-\begin{equation}
-	D_\alpha = -i\frac{\omega}{\Omega_\alpha} + i\frac{\vec{k}\cdot\vec{V}_{\alpha 0}}{\Omega_\alpha} + \frac{\nu_\alpha}{\Omega_\alpha} \approx \frac{\nu_\alpha}{\Omega_\alpha} = r_\alpha
-\end{equation}
-Equation \ref{eqn:gdi_Mak16_alt} can then be expressed as follows.
-\begin{equation}
-	\label{eqn:gdi_Mak16_Freg}
-	(r_i-r_e)\omega = \vec{V}_d\cdot\vec{k}\left(\frac{Z_2}{Z_1}\right)+(r_i\vec{V}_{i0}-r_e\vec{V}_{e0})
-\end{equation}
-The growth rate, \(\gamma\), is contained within the dispersion relation as \(\omega = \omega_r+i\gamma\), so the imaginary part of Equation \ref{eqn:gdi_Mak16_Freg} must be considered separately.
-\begin{equation}
-	\label{eqn:gdi_Mak16_gam}
-	(r_i-r_e)\gamma = \vec{V}_d\cdot\vec{k}\;\text{Im}\left\{\frac{Z_2}{Z_1}\right\}
-\end{equation}
-The differential plasma drift has been defined previously in Equation \ref{eqn:gdi_Mak16_Vd} but the imaginary component of \(Z_2/Z_1\) must still be identified.
+%\begin{equation}
+%	D_\alpha = -i\frac{\omega}{\Omega_\alpha} + i\frac{\vec{k}\cdot\vec{V}_{\alpha 0}}{\Omega_\alpha} + \frac{\nu_\alpha}{\Omega_\alpha} \approx \frac{\nu_\alpha}{\Omega_\alpha} = r_\alpha
+%\end{equation}
+%Equation \ref{eqn:gdi_Mak16_alt} can then be expressed as follows.
+%\begin{equation}
+%	\label{eqn:gdi_Mak16_Freg}
+%	(r_i-r_e)\omega = \vec{V}_d\cdot\vec{k}\left(\frac{Z_2}{Z_1}\right)+(r_i\vec{V}_{i0}-r_e\vec{V}_{e0})
+%\end{equation}
+%The growth rate, \(\gamma\), is contained within the dispersion relation as \(\omega = \omega_r+i\gamma\), so the imaginary part of Equation \ref{eqn:gdi_Mak16_Freg} must be considered separately.
+%\begin{equation}
+%	\label{eqn:gdi_Mak16_gam}
+%	(r_i-r_e)\gamma = \vec{V}_d\cdot\vec{k}\;\text{Im}\left\{\frac{Z_2}{Z_1}\right\}
+%\end{equation}
+%The differential plasma drift has been defined previously in Equation \ref{eqn:gdi_Mak16_Vd} but the imaginary component of \(Z_2/Z_1\) must still be identified.
 Using the definitions of \(Z_1\) and \(Z_2\) provided in Equation \ref{eqn:gdi_Mak16_Z} and the simplifying assumptions discussed above for the F region,
-\begin{equation}
-	\frac{Z_2}{Z_1} = \frac{R(i+a)-b}{i+a+Rb}
-\end{equation}.
-Multiplying both the numerator and denominator by \(-i+a+Rb\) removes the imaginary component from the denominator.
+%\begin{equation}
+%	\frac{Z_2}{Z_1} = \frac{R(i+a)-b}{i+a+Rb}
+%\end{equation}.
+%Multiplying both the numerator and denominator by \(-i+a+Rb\) removes the imaginary component from the denominator.
 Additionally the local approximation \(G \ll k_\perp\) is taken, which results in \(a\), \(b\), and \(c\) being small so that any second order terms in these quantities can be neglected.
-\begin{equation}
-	\frac{Z_2}{Z_1} = R+ib(1+R^2)
-\end{equation}
-Using the above result and Equation \ref{eqn:gdi_Mak16_Vd}, Equation \ref{eqn:gdi_Mak16_gam} becomes
-\begin{equation}
-	(r_i-r_e)\gamma = (r_i-r_e)s_es_i\left(1+\psi\right)\left(R\vec{V}_E-\frac{E}{B}\right)\cdot\vec{k}b\left(1+R^2\right)
-\end{equation}
-Recognizing that \(s_es_i\left(1+R^2\right) = (1+\psi)^{-2}\) and \(b = \vec{G}\cdot\vec{k}\times\uvec{b}/k^2 = -\vec{k}\cdot\uvec{b}\times\vec{G}/k^2\) by vector identities, a final expression for the GDI growth rate in the F region can be found, which agrees well with the results of \citet{Makarevich2014c}
-\begin{equation}
-	\label{eqn:gdi}
-	\gamma = \frac{1}{1+\psi}\left(\uvec{k}\cdot\uvec{b}\times\vec{G}\right)\uvec{k}\cdot\left(\frac{\vec{E}}{B}-R\vec{V}_E\right)
-\end{equation}.
-Equation \ref{eqn:gdi} will be critical to the analysis done both in Chapter \ref{sec:paper2} and Chapter \ref{sec:paper3} in this thesis.
+%\begin{equation}
+%	\frac{Z_2}{Z_1} = R+ib(1+R^2)
+%\end{equation}
+%Using the above result and Equation \ref{eqn:gdi_Mak16_Vd}, Equation \ref{eqn:gdi_Mak16_gam} becomes
+%\begin{equation}
+%	(r_i-r_e)\gamma = (r_i-r_e)s_es_i\left(1+\psi\right)\left(R\vec{V}_E-\frac{E}{B}\right)\cdot\vec{k}b\left(1+R^2\right)
+%\end{equation}
+%Recognizing that \(s_es_i\left(1+R^2\right) = (1+\psi)^{-2}\) and \(b = \vec{G}\cdot\vec{k}\times\uvec{b}/k^2 = -\vec{k}\cdot\uvec{b}\times\vec{G}/k^2\) by vector identities, a final expression for the GDI growth rate in the F region can be found, which agrees well with the results of \citet{Makarevich2014c}
+%\begin{equation}
+%	\label{eqn:gdi}
+%	\gamma = \frac{1}{1+\psi}\left(\uvec{k}\cdot\uvec{b}\times\vec{G}\right)\uvec{k}\cdot\left(\frac{\vec{E}}{B}-R\vec{V}_E\right)
+%\end{equation}.
+%Equation \ref{eqn:gdi} will be critical to the analysis done both in Chapter \ref{sec:paper2} and Chapter \ref{sec:paper3} in this thesis.
 
-%In addition to analytical analysis of dispersion relations, GDI has also been studied through numerical simulations.
-The advantage of simulations is that they can often be used to study phenomena at better time and spatial resolution than can be achieved by instruments and the senario of interest can be specified exactly.
-This is particularly useful for considering the degree of anisotropy that exists in plasma irregularities.
+\begin{table}
+\begin{tabular}{l l}
+Limiting Case & Growth Rate, \(\gamma\) \\
+\hline
+GDI/FBI in \(E\) Region & \(\gamma = \frac{1}{1+\hat{\psi}}\left[\frac{\hat{\psi}}{\nu_i}\left(\left(\frac{\vec{V}_d\cdot \vec{k}}{1+\hat{\psi}}\right)^2-C_s^2 k_\perp^2\right)+\frac{br_i\vec{V}_d\cdot \vec{k}}{1+\hat{\psi}}\right]\) \\
+GDI/CCI in \(F\) Region & \(\gamma = -\frac{b}{\psi+y^2}\left[\frac{\psi\vec{k}\cdot\vec{E}_{0\perp}+E_{0\parallel}k_\parallel}{B}\right]+C\left[r_ek_\perp^2-\frac{k_\parallel^2}{r_i}\left(1+\frac{r_i^2}{\psi+y^2}\right)\right]\) \\
+Cold Plasma in \(F\) Region & \(\gamma = \frac{b}{1+\psi}\left(R\vec{V}_E-\frac{\vec{E}_{0\perp}}{B}\right)\cdot\vec{k}\) \\
+\end{tabular}
+\caption{Irregularity growth rates in the ionosphere for three different limiting cases \citep{Makarevich2016a}.}
+\label{tab:gdi_exps}
+\end{table}
 
 Although linear GDI theory as described predicts clear dependence on wave vector direction \citep{Makarevich2014c}, this theory is only really valid in the long-wavelength limit, \(\vec{k}\cdot\vec{V}_E \ll \nu_\alpha\).
-In the F region, the ion-neutral collision frequency, \(\nu_i\), is about 10 Hz and if the drift speed, \(V_E\), is about 1000 m/s, fluid theory only applies for wavelengths greater than 100 m, larger than the decameter-scale waves that are detected by SuperDARN.
+In the \(F\) region, the ion-neutral collision frequency, \(\nu_i\), is about 10 Hz and if the drift speed, \(V_E\), is about 1000 m/s, fluid theory only applies for wavelengths greater than 100 m, larger than the decameter-scale waves that are detected by SuperDARN.
 This raises the important question of whether or not the same anisotropy that is predicted by linear GDI theory at larger scales can be expected in small-scale irregularities where nonlinear processes may be important.
@@
 However, very strong shears or increasing the impact of ion-inertial factors can cause isotropy \citep{Gondarenko2001,Gondarenko2006}.
-The power spectra derived from these simulations are similar to those from experimental observations \citep{Baker1978,Kelley1979}, however ...
-\textbf{WHAT'S DIFFERENT AND IMPORTANT ABOUT PROJECT 3 HERE?}
+The power spectra derived from these simulations are similar to those from experimental observations \citep{Baker1978,Kelley1979}, however it is computationally challenging to run simulations down to decameter scales and observations are generally limited to rocket and satellite techniques, which can only be used to calculate power spectra in one dimension \citep{Villain1986,Moen2012}.
 
@@
 Polar patches were first observed in the 1960s \citep{Hill1963}, but much of the work of characterizing them using optical and radio techniques was not accomplished until two decades later \citep{Weber1981,Weber1984,Weber1986,Buchau1983,Buchau1985}.
-Polar patches are density enhancements of as much as \(10^6\) cm\(^{-3}\) that occur primarially in the polar F region \citep{Buchau1983}.
+Polar patches are density enhancements of as much as \(10^6\) cm\(^{-3}\) that occur primarially in the polar \(F\) region \citep{Buchau1983}.
 There were initially catagorized as extending 800--1000 km horizontally \citep{Weber1984}, but more recent studies have identified patches as large as 1500 km \citep{Hosokawa2014}.
@@
 Additionally, backscatter tends to occure at closer ranges in summer at midnight, but moves further away from the radar in other seasons at midday \citep{Milan1997}.
-In general, echo occurrence in high durring the day and low at night, but nighttime occurrence is directly proportional to the F peak density and daytime occurrence can be significantly reduced by a dense and conductive E layer.
+In general, echo occurrence in high durring the day and low at night, but nighttime occurrence is directly proportional to the \(F\) peak density and daytime occurrence can be significantly reduced by a dense and conductive \(E\) layer.
 \citep{Koustov2004,Kane2012,Vickrey1982}.
-In the daytime, solar illumination causes the electron density of the E region to increase dramatically, making it highly conductive.
-This conductive layer can provide a conduit by which the potential differences associated with FAIs in the F region can be shorted out, effectively reducing F-region plasma irregularities \citep{Vickrey1982}.
-Both irreguarity production and radar propagation factors probably contribute to these regular variations of echo occurrence, because while a dense, daytime ionosphere creates favorable radar propigation conditions, it can also contribute to E-region shorting or smoothing of gradients which surpress irregularity production \citep{Koustov2004}.
+In the daytime, solar illumination causes the electron density of the \(E\) region to increase dramatically, making it highly conductive.
+This conductive layer can provide a conduit by which the potential differences associated with FAIs in the \(F\) region can be shorted out, effectively reducing \(F\)-region plasma irregularities \citep{Vickrey1982}.
+Both irreguarity production and radar propagation factors probably contribute to these regular variations of echo occurrence, because while a dense, daytime ionosphere creates favorable radar propigation conditions, it can also contribute to \(E\)-region shorting or smoothing of gradients which surpress irregularity production \citep{Koustov2004}.
 
@@
 Properly oriented density gradients and electric fields are a nessesary condition for GDI to be operational, Section \ref{sec:lit_instbilities}, so the stronger these are, the more likely the irregularity production rate is high.
-Both backscatter power and occurrence increase with increasing electric field strength \citep{Ballatore2001,Danskin2002,Makarevich2014}.
+Both backscatter power and occurrence increase with increasing electric field strength \citep{Ballatore2001,Danskin2002,Makarevich2014b}.
 Similarly, backscatter occurrence increases when the IMF \(B_z\) component is strongly negative, implying a strong merging electric field.
@@
 The primary radar propogation factor to consider is the background plasma density and how much the radar beam is refracted.
-If the background density is too low, the beam is under refracted and is never perpendicular to the magnetic field, however if the density is too high, the beam is over refracted, such that observation of small-scale FAIs is most favorable if the background density is between \(1\times 10^{11}\) m\(^{-3}\) and \(4\times 10^{11}\) m\(^{-3}\) \citep{Danskin2002,Makarevich2014}.
-If the D region is too dense, this can also in theory attenuate the signal and reduce the rate of returned backscatter, although observations have shown this is not a major factor in echo occurrence \citep{Danskin2002}.
+If the background density is too low, the beam is under refracted and is never perpendicular to the magnetic field, however if the density is too high, the beam is over refracted, such that observation of small-scale FAIs is most favorable if the background density is between \(1\times 10^{11}\) m\(^{-3}\) and \(4\times 10^{11}\) m\(^{-3}\) \citep{Danskin2002,Makarevich2014b}.
+If the \(D\) region is too dense, this can also in theory attenuate the signal and reduce the rate of returned backscatter, although observations have shown this is not a major factor in echo occurrence \citep{Danskin2002}.
 Finally, the radar equation predicts that the returned backscatter should be proportional to the perturbation density, \(\delta n\), squared, or, equivilently, the product of the background density and the fractional density, \(\delta n/n\), squared, Equation \ref{eqn:fractional_density}
@@
 
-The aim of this work is to investigate small-scale plasma irregularity production in the F-region polar cap.
+The aim of this work is to investigate small-scale plasma irregularity production in the \(F\)-region polar cap.
 In particular, factors such as direct and indirect global solar control and GDI directional dependance will be considered.
@@
 
-The first objective of is to evaluate the extent of solar control over irregularity production in the polar F region.
+The first objective of is to evaluate the extent of solar control over irregularity production in the polar \(F\) region.
 The occurrence of radar backscatter will be compared with both solar illumination (considered direct solar control) and IMF factors (considered indirect solar control).

### 2017-01-17 introduction.tex
Polar irregularity figure added

@@
 \label{sec:polar_structure}
+
 This thesis is concerned with the polar ionosphere, which is a particularly interesting and dynamic region due to the presence of open field lines to the solar wind.
 The main focus of this work will be plasma density perturbations, often called irregularities.
-Density irregularities in the polar cap can occur on scales ranging from thousands of kilometers to less than a centimeter \citep{Tsunoda1988}.
+Density irregularities in the polar cap can occur on scales ranging from thousands of kilometers to less than a centimeter \citep{Tsunoda1988}, and a summary of some of these irregularities is presented in Figure \ref{fig:polarirreg}.
 Large-scale plasma irregularities are typically considered to be any density structure on the scale of greater than 30 km \citep{Kelley2009}.
 Some common examples are polar patches, polar holes, and sun-aligned arcs.
-Polar patches are large density enhancements (usually at least twice the background plasma density) that travel across the polar cap with the background convection \citep{Weber1984,Valladares1994}.
+Polar patches are large density enhancements (usually at least twice the background plasma density) that travel across the polar cap with the background convection, Figure \ref{fig:polarirreg} \citep{Weber1984,Valladares1994}.
 Polar holes are plasma density depletions that typically occur in the \(F\) region slightly poleward of the auroral oval \citep{Benson2001}.
@@
 They are very narrow and often characterized by large and complex velocity shears on either side of the arc \citep{Valladares1991}.
+Large scale plasma structures and the density gradients that surround them often serve as a platform for smaller-scale structures to form as well.
+Polar patches, for instance, are known to form large finger-like structures along their trailing edges, Figure \ref{fig:polarirreg} \citep{Gondarenko2004,Hosokawa2016}.
+
+\begin{figure}
+	\includegraphics[width=0.8\textwidth]{irregularity.pdf}
+	\caption{Schematic of different scale size plasma density irregularities that occur in the polar cap.
+Polar patches, orange, move with the background ionospheric convection, indicated by green contours.
+Field-aligned irregularities, FAIs, are density perturbations that form along the magnetic field lines, pink lines.}
+	\label{fig:polarirreg}
+\end{figure}
 
@@
 Although small-scale irregularities are not responsible for scintillation that directly impact GPS/GNSS performance, they can be the easiest to study because of large data sets collected from high frequency ionospheric radars globally for several decades (discussed further in Section \ref{sec:superdarn}).
+Because electrons move far more easily along magnetic field lines than across them, density perturbations tend to be aligned with the magnetic field and are often known as field-aligned irregularities (FAI), Figure \ref{fig:polarirreg}.
 The irregularity classifications presented here are rough categories only and a high degree of coupling between different scales occurs, which makes plasma structures of all scales (large, intermediate, and small) important for plasma structuring processes.

### 2017-01-17 introduction.tex
References updated

@@
 Large scale plasma structures and the density gradients that surround them often serve as a platform for smaller-scale structures to form as well.
-Polar patches, for instance, are known to form large finger-like structures along their trailing edges, Figure \ref{fig:polarirreg} \citep{Gondarenko2004,Hosokawa2016}.
+Polar patches, for instance, are known to form large finger-like structures along their trailing edges, Figure \ref{fig:polarirreg} \citep{Gondarenko2004b,Hosokawa2016}.
 
@@
 Large-scale structuring in the polar cap can take a variety of forms, Section \ref{sec:polar_structure}, but here we will focus on large regions of enhanced plasma density known as polar patches  \citep[e.g.][]{Weber1984,Weber1986,Buchau1983,Buchau1985}.
-One of the primary interests of this thesis is the relationships between irregularities of different scales and small-scale structuring has long been observed around polar patches \citep{Weber1984,Milan2002,Moen2012}.
+One of the primary interests of this thesis is the relationships between irregularities of different scales and small-scale structuring has long been observed around polar patches \citep{Weber1984,Milan2002b,Moen2012}.
 Although small-scale structures have also been observed around sun-aligned arcs \citep[e.g.][]{Koustov2012}, these structures also tend to be more complicated and often are surrounded by intense shears and electric fields \citep{Safargaleev2000,Aikio2002,Kozlovsky2007}.
@@
 \end{equation}.
-An even more general consideration involves arbitrary gradient and plasma drift directions \citep{Kesknien1982,Makarevich2014c}.
+An even more general consideration involves arbitrary gradient and plasma drift directions \citep{Keskinen1982a,Makarevich2014c}.
 In this more general case, the directional dependence of GDI has been shown to be significant, such that the growth rate can actually change signs (determining whether wave growth or damping occurs) with different wavevector directions, so it is important to consider this factor \citep{Makarevich2014c}.

### 2017-01-17 paper1.tex
References updated

@@
 This resulted in a time series that approximately represents the variation in echo occurrence relative to a quiet-day season-specific diurnal average or the relative occurrence.
-This approach allows us to isolate the IMF effects from those due to diurnal and seasonal variations \citep{Makarevich12}.
+This approach allows us to isolate the IMF effects from those due to diurnal and seasonal variations \citep{Makarevich2012}.
 Points were divided into six groups based on UT and plotted in different panels of Figure \ref{fig:by_bz}.
@@
 The two phenomena that can be responsible for this are polar patches \citep{Weber1984} and sun-aligned arcs \citep{Valladares1994}.
-The polar patches in particular are known to travel long distances across the polar cap and even well into the nightside auroral zone \citep{Moen2007,Oksavick2010}.
+The polar patches in particular are known to travel long distances across the polar cap and even well into the nightside auroral zone \citep{Moen2007,Oksavik2010}.
 
@@
 near the solar terminator.
-This is also in contrast with previous studies where a qualitative agreement was observed between contours of constant occurrence and SZA \citep{Kane12,Ghezelbesh2014b}.
+This is also in contrast with previous studies where a qualitative agreement was observed between contours of constant occurrence and SZA \citep{Kane2012,Ghezelbash2014b}.
 Occurrence decreases on both sides of SZA of 90\(\deg\) but most likely for different reasons.
@@
 One mechanism through which IMF-controlled variations in convection may be responsible for changes in occurrence of plasma irregularities involves changing orientations between plasma density gradients \(\vec{G}\) and convection velocity \(\vec{V}_E\) vectors.
-GDI can more easily excite waves when a component \(\vec{G}\cdot\vec{V}_E\) is more positive, with \(\vec{G}\parallel\vec{V}_E\) being most favorable orientation \citep{Kesknein1982a,Makarevich2014c}.
+GDI can more easily excite waves when a component \(\vec{G}\cdot\vec{V}_E\) is more positive, with \(\vec{G}\parallel\vec{V}_E\) being most favorable orientation \citep{Keskinen1982a,Makarevich2014c}.
 IMF-controlled changes in \(\vec{V}_E\) can thus potentially lead to changes in \(\vec{G}\cdot\vec{V}_E\), which, in turn, will change GDI growth rate and occurrence of GDI-produced irregularities.

### 2017-01-17 thesis.sh

@@ -30,6 +30,6 @@ pdflatex LLamarcheThesis.tex
 
 rm *.aux
 rm *.bbl
-rm *.blg
+# rm *.blg
 rm *.out
 rm *.log

### 2017-01-17 conclusion.tex
Corrections made to Conclusion

@@
 
-Plasma irregularities in the F-region ionosphere have been observed for decades.
+Plasma irregularities in the \(F\)-region ionosphere have been observed for decades.
 However, despite extensive data sets collected over these years, supprisingly little is known about these irregularities, especially at small scales.
 This is in part due to the complex and turbulent nature of the ionosphere, particularly in the polar regions, but also due to some of the inherent limitations in traditional observation techniques which can make it very challenging to identify exactly what processes are operational.
-In this Chapter, we summarize the results that have previously been discussed in Chapters \ref{sec:paper1}--\ref{sec:paper3} and consider what overarching conclusions can be derived from them which further our understanding of small-scale polar F-region plasma irregularities.
+In this Chapter, we summarize the results that have previously been discussed in Chapters \ref{sec:paper1}--\ref{sec:paper3}, shown in Figure \ref{fig:conclusion}, and consider what overarching conclusions can be derived from them.
+Additionally, Section \ref{sec:futurework} will provide recommendations for future research which will help answer some of the still outstanding questions and further our understanding of small-scale polar \(F\)-region plasma irregularities.
 
-\section{Summary and Conclusions}
+\section{Summary of Results and Conclusions}
 \label{sec:summary}
 
-Irregularities occur most often close to the terminator.
-This is likely due to the presence of a very large density gradient between high density solar illuminated plasma on the dayside and low density plasma on the nightside.
-In addition, if a SuperDARN radar itself is positioned close to the terminator, it is observing the density gradient along its boresight.
-If small-scale irregularity growth is directionally dependent, this is additionally the optimal orientation from which to observe the plasma structures, leading to increased echo occurrence.
-
-In the polar caps, backscatter is often observed at nighttime even when background plasma density is low and propagation conditions are too poor for FAIs to be observed with HF radar.
+In the polar caps, backscatter is often observed at nighttime even when background plasma density is low and propagation conditions are traditionally thought to be too poor for FAIs to be observed with HF radar.
 This implies an additional source of nighttime plasma density that is not easily accounted for by models.
-Frequent polar patches that propogate from the dayside to the nightside could be the source of this additional density.
+Frequent polar patches that propagate from the dayside to the nightside could be the source of this additional density.
 The requirements for confirming this hypotheis will be detailed further in Section \ref{sec:futurework}.
@@
 Nighttime observations of FAIs is also higher with a positive IMF By component.
-This could be due to the control that the By component over ionospheric convection patterns.
-In a convectional two-cell convection pattern, the plasma flows antisunward from dayside to nightside across the terminator.
+This could be due to the control that the By component exerts over ionospheric convection patterns.
+In a convectional two-cell convection pattern with predominantly negative IMF Bz, the plasma flows antisunward from dayside to nightside across the terminator.
 In this arrangement, the plasma flow, \(\vec{V}_E\), and the density gradient, \(\vec{G}\), are antiparallel, the least favorable orientation for GDI growth.
-However, with a significant IMF By component, the orientation of the convection cells rotates such that plasma is no longer convects antisunward, creating a situation where GDI is more favorable and plasma structures can develop more readily.
-Additionally, an IMF By component can contribute to the formation of polar patches by creating convection perturbations that separate high density patches from the daytime resevoir, as described in Section \ref{sec:lit_patches}.
+However, with a significant IMF By component, the orientation of the convection cells rotates such that plasma no longer convects directly antisunward, creating a situation where GDI is more favorable and plasma structures can develop more readily.
+Additionally, an IMF By component can contribute to the formation of polar patches by creating convection perturbations that separate high density patches from the daytime reservoir, as described in Section \ref{sec:lit_patches}.
 In this way, the IMF By component may be indirectly responsible for the increased observation of backscatter.
@@
 This raises the important question of exactly how and where plasma structures form around polar patches.
-Observations have long suggested there is asymmetry in plasma structuring surrounding polar patches, but the results presented in Chapter \ref{sec:paper2} suggest that production may also be anisotropic.
+Observations have long suggested there is asymmetry in plasma structuring surrounding polar patches \citep{Weber1984,Milan2002b,Koustov2012}, but the results presented in Chapter \ref{sec:paper2} suggest that production may also be anisotropic, Figure \ref{fig:conclusion}.
 Asymmetry and anisotropy are independent such that the plasma structures can exhibit either of them independently or both or neither.
@@
 Additionally, linear theory predicts that irregularity growth is anisotropic, so the orientation of the patch relative to the radar can also impact the observed irregularities.
+In the \(F\) region, the GDI growth rate is highest for wavevectors perpendicular to the gradient, \(\vec{G}\).
+Although plasma waves can occur in all directions, a SuperDARN radar will only observe waves who's wavevectors are parallel to a radar beam.
+Because of this, irregularities as well as structure asymmetry are most likely to be maximized if the radar's boresight is parallel to the structure's elongation, Figure \ref{fig:conclusion}.
 
-\section{Future Work}
+Plasma irregularities are observed most often close to the terminator.
+This is likely due to the presence of a very large density gradient between high density solar illuminated plasma on the dayside and low density plasma on the nightside.
+The large background gradient promotes irregularity production through GDI.
+In addition, if a SuperDARN radar itself is positioned close to the terminator, it is observing the density gradient along its boresight, Figure \ref{fig:conclusion}.
+As discussed above, if small-scale irregularity growth is directionally dependent, this is additionally the optimal orientation from which to observe the plasma structures, leading to increased echo occurrence.
+
+\section{Recommendations for Future Research on \(F\) Region Plasma Irregularities}
 \label{sec:futurework}
@@
 
-In Chapter \ref{sec:paper1}, the occurence of radar backscatter durring nighttime in the polar cap was found to be much higher than expected \citep{Bristow2011}, particularly given that raytracing simulations predict that the orthagonality condition is rarely met and therefore there should be little to no backscatter.
-The proposed hypothesis to explain this discreperency is that polar patches, which are not typically well represented by ionospheric density models, occur frequently in the nighttime F region and, on average, enhance the background plasma density enough that propagation conditions are sufficient for HF radar backscatter to regularly be observed.
+In Chapter \ref{sec:paper1}, the occurence of radar backscatter durring nighttime in the polar cap was found to be much higher than expected \citep{Bristow2011}, particularly given that raytracing simulations predict that the orthagonality condition is rarely met and therefore there should be little to no backscatter, Figure \ref{fig:month_avg_occ}.
+The proposed hypothesis to explain this discreperency is that polar patches, which are not typically well represented by ionospheric density models, occur frequently in the nighttime \(F\) region and, on average, enhance the background plasma density enough that propagation conditions are sufficient for HF radar backscatter to regularly be observed.
 Experimental surveys of polar patches have shown low occurrence at nighttimes, however the occurrence of patches and their intensity increase at equinox, which is where frequent nighttime backscatter is predominantly observed \citep{Rodger1996}.
 The expected signature in the radar data from this kind of senario would be a night with a series of isolate ``patches'' of backscatter.
-An example of such an event was presented, Figure \ref{whatever the hell THIS'll be}, but it is unknown how often events like this occur.
+An example of such an event was presented, Figure \ref{fig:patchy_example}, but it is unknown how often events like this occur.
 Additionally, the intensity and size of a patch nessesary to cause the radar beam to refract sufficiently must be established, possibly through further use of raytracing simulations with customizable density profiles.
 
-One of the major drawbacks to linear GDI theory is its limitation of applicability only to long wavelengths, which means if the F region, it cannot be directly applicable to decameter-scale irregularities measured by HF radars.
+One of the major drawbacks to linear GDI theory is that it inherently assumes long wavelengths, which means if the \(F\) region, it technically should not be directly applicable to decameter-scale irregularities measured by HF radars.
 Generally, it is assumed that some form of turbulent cascade connects plasma structures of different scale sizes in the ionosphere such that predictions made at large scales can still be applied to decameter-scale observations \citep{Tsunoda1988}.
-This helps explain some of the observations of asymmetry around polar patches.
+This helps explain some of the observations of asymmetry around polar patches, which is predicted at large scales but often observed at decameter scales with radars \citep[e.g.][]{Weber1984,Milan2002b,Koustov2012}.
 This asymmetry was also confirmed in the GDI growth rate from the results shown in Chapter \ref{sec:paper2}.
@@
 It may be possible to further the theoretical understanding of GDI at small scales by developing a three dimentional picture of it using kinetic theory.
-\citet{Basu1995} developed such a formulation for a specific geometry throughout all altitudes within the ionosphere, but expanding this to general vector directions would most likely be extremely mathematically intensive.
+\citet{Basu1995} developed such a formulation for a specific geometry throughout all altitudes within the ionosphere, but expanding this to general vector directions would not be trivial, mathematically.
+However, modeling irregularity growth with such a formulation may be helpful to understand how plasma structures develop at such small scales.
 
@@
 This allows a polar patch to be tracked over a longer distance and more individual data points to be captured.
-Additionally, specific modes can be utilized that give a much finner grid of RISR beams, allowing measurments to be made closer together and smaller scale gradients to be be considered.
+Additionally, specific modes can be utilized that give a much finner grid of RISR beams, allowing measurements to be made closer together and smaller scale gradients to be be considered.
 The downside to these modes is that it takes longer for each set of measurements to be made, so the temporal cadence decreases.
-However, the required integration time for an ISR decreases if the background ionospheric density is higher, so the the temporal resolution can improve for measurments taken durring summer or daytime.
-Another advantange to the RISR-N/RISR-C locations is that the SuperDARN radars at Clyde River (CLY) and Inuvik (INV) (Figure \ref{fig:superdarnmap}), both of which share at least part of the RISR-N/RISR-C FoV, but observe it from different directions.
-This presents an unique oppertunity to further investigate the directional dependence of small-scale irregularities by observing the differences in backscatter observed from the same scattering volume at different directions.
+However, the required integration time for an ISR decreases if the background ionospheric density is higher, so the the temporal resolution can improve for measurements taken durring summer or daytime.
+Another advantage to the RISR-N/RISR-C locations is that the SuperDARN radars at Clyde River (CLY) and Inuvik (INV) (Figure \ref{fig:superdarnmap}), both of which share at least part of the RISR-N/RISR-C FoV, but observe it from different directions.
+This presents an unique opportunity to further investigate the directional dependence of small-scale irregularities by observing the differences in backscatter observed from the same scattering volume at different directions.
 Although straight forward in theory, this experiment would most likely prove nontrivial in practice.
-Backscatter tends to be limited to the statistical high occurrence band discussed in Chapter \ref{sec:paper1}, and at least with INV, this band probably occurs at closer ranges than where the FoV overlaps other radars, so coincident backscatter is probably rare.
-The probability of backscatter being observed is better when the background density is within the range where propogation conditions are most favorable, when the ionosphere is sunlit.
-Additionally, it is adventagious to choose a time when the occurrence of backscatter is statistically high.
+Backscatter tends to be limited to the statistical high occurrence band shown in Figure \ref{fig:month_avg_occ}, and at least with INV, this band probably occurs at closer ranges than where the FoV overlaps other radars, so coincident backscatter is probably rare.
+The probability of backscatter being observed is better when the background density is within the range where propagation conditions are most favorable, when the ionosphere is sunlit.
+Additionally, it is advantageous to choose a time when the occurrence of backscatter is statistically high.
 Based on results from Chapter \ref{sec:paper1}, the best time for such an experiment would therefore be daytime durring an equinox month.
+
+In Chapter \ref{sec:paper3}, both the occurrence and power of radar backscatter are used as proxies to indicate the ``amount'' of irregularities that exist in the ionosphere.
+However, the trends from these quantities are rarely in good agreement, giving rise to the question what, if anything, does each of them indicate about the plasma structuring in the ionosphere.
+The factors that control each need to be isolated, but this is nontrivial.
+Echo occurrence generally agrees with the linear GDI theory predictions but echo power only does for relatively low power.
+Irregularities may develop as predicted by linear GDI theory when their amplitudes are small but other nonlinear processes may impact larger density perturbations, such as shear-driven instabilities \citep{Gondarenko2006} or multistep processes \citep{Carlson2012}.
+By the radar equation, the power returned should be proportional to the product of the background density and fractional density, squared, Equation \ref{eqn:fractional_density}.
+However, there is no general consensus on what controls the fractional density in the \(F\) region, although it has been suggested that it is either constant or a function of convection speed, \(V_E\) \citep{Koustov1988,Haldoupis1990,Makarevich2014b}.
+In Chapter \ref{sec:paper3}, we additionally propose it may be a function of the instability growth rate, \(\gamma\).
+Because ground based radars are so often used to investigate ionospheric irregularities, this is an important issue to resolve to develop an accurate picture of how radar observations relate to plasma irregularities.
 

### 2017-01-18 abstract.tex
First draft of Abstract

% Abstract

Plasma irregularities in the ionosphere have been studied for decades using a variety of techniques.  A number of factors that contribute to the production and observation of plasma irregularities in the polar \(F\) region ionosphere are examined.  First, global factors related to both direct and indirect solar control are considered.  A band of ranges where echoes are statistically most likely to be observed is identified.  The location of this band varies diurnally and seasonally, but its motion relative to the radar can be explained by variations in the background plasma density effecting how far the beam must travel to be refracted to perpendicular to the magnetic field.  The high observed occurrence of nighttime backscatter implies an additional source of nighttime plasma density that is not accournted for by traditional models.  An example is presented that shows frequent polar patches may be the source of this density.  Backscatter occurrence peaks at the terminator.  Nighttime occurence is enhanced when a positive IMF By component is present, probably due to the convection pattern changing and creating a more favorable situation for gradient drift instability growth.

The gradient-drift instability is modeled within the field of view of a SuperDARN radar to identify when the growth rate is largest and examine asymmetry around a large-scale structure.  Dependencies on the relative directions of the gradient, plasma drift, and wavevector are all considered throughout the ionosphere.  In the \(F\) region, the strongest asymmetry is found when an elongated structure is oriented along the radar's boresight and moving perpendicular to its direction of elongation.  These results have implications for observations made with the SuperDARN network, but assume the directional dependencies predicted by linear GDI theory are applicable at decameter scales.

To test the accuracy of the predictions found from the model, irregularities surrounding a polar patch are analyzed in the context of linear GDI theory.  Backscatter power and occurrence from decameter scale irregularities are found using measurements from the SuperDARN radar at Rankin Inlet while background density gradients and electric fields are found from the north face of the Resolute Bay Incoherent Scatter Radar.  Particular emphasis is placed on directional dependence and wether small-scale irregularites show any signs of anisotropy.  While low powered echoes seem to follow the predictions of linear GDI theory, higher power echoes quickly exhibit much more complicated, nonlinear behavior.

### 2017-01-18 abstract.tex
Formatting updated

@@
 
-Plasma irregularities in the ionosphere have been studied for decades using a variety of techniques.
-A number of factors that contribute to the production and observation of plasma irregularities in the polar \(F\) region ionosphere are examined.
-First, global factors related to both direct and indirect solar control are considered.
+Plasma irregularities in the polar \(F\) region ionosphere are investigated primarily using ground based radars within the Super Dual Auroral Radar Network (SuperDARN) and the factors that contribute to the production and observation of these irregularities are considered.
+First, solar control on irregularity production is evaluated.
 A band of ranges where echoes are statistically most likely to be observed is identified.
-The location of this band varies diurnally and seasonally, but its motion relative to the radar can be explained by variations in the background plasma density effecting how far the beam must travel to be refracted to perpendicular to the magnetic field.
-The high observed occurrence of nighttime backscatter implies an additional source of nighttime plasma density that is not accournted for by traditional models.
+The location of this band varies diurnally and seasonally, but its motion relative to the radar can be explained by background density with solar illumination effecting radar beam propogation.
+The high observed occurrence of backscatter at night implies an additional source of nighttime plasma density that is not accounted for by traditional models.
 An example is presented that shows frequent polar patches may be the source of this density.
 Backscatter occurrence peaks at the terminator.
-Nighttime occurence is enhanced when a positive IMF By component is present, probably due to the convection pattern changing and creating a more favorable situation for gradient drift instability growth.
+Nighttime occurence is enhanced when a positive IMF By component is present, probably due to changes in the convection pattern creating a more favorable situation for gradient-drift instability growth.
 
-The gradient-drift instability is modeled within the field of view of a SuperDARN radar to identify when the growth rate is largest and examine asymmetry around a large-scale structure.
+The gradient-drift instability is modeled within the field of view of a SuperDARN radar to identify when the growth rate is largest and examine asymmetry around large-scale structures, such as polar patches.
 Dependencies on the relative directions of the gradient, plasma drift, and wavevector are all considered throughout the ionosphere.
@@
 
-To test the accuracy of the predictions found from the model, irregularities surrounding a polar patch are analyzed in the context of linear GDI theory.
+To test the accuracy of the predictions found from the model, irregularities surrounding a polar patch are analyzed in the context of linear gradient-drift instability theory.
 Backscatter power and occurrence from decameter scale irregularities are found using measurements from the SuperDARN radar at Rankin Inlet while background density gradients and electric fields are found from the north face of the Resolute Bay Incoherent Scatter Radar.
-Particular emphasis is placed on directional dependence and wether small-scale irregularites show any signs of anisotropy.
-While low powered echoes seem to follow the predictions of linear GDI theory, higher power echoes quickly exhibit much more complicated, nonlinear behavior.
+Particular emphasis is placed on directional dependence and wether small-scale irregularities show any signs of anisotropy.
+While low powered echoes seem to follow the predictions of linear theory, higher power echoes quickly exhibit much more complicated, nonlinear behavior.

### 2017-01-18 introduction.tex
Formatting updated

@@
 	\includegraphics[width=\textwidth,angle=0]{ste.pdf}
-	\caption{Diagram of the solar-terrestrial environment.
+	\caption[Solar-Terrestrial Environment]{Diagram of the solar-terrestrial environment.
 The solar wind contains both a steady stream of charged particles and the IMF, dark blue.
@@
 	\includegraphics[width=\textwidth]{densprofile.pdf}
-	\caption{Altitude profile of electron density throughout the ionosphere found from the IRI model.  Typical ranges of the \(D\), \(E\), and \(F\) regions are identified.}
+	\caption[Ionosphere density profile]{Altitude profile of electron density throughout the ionosphere found from the IRI model.  Typical ranges of the \(D\), \(E\), and \(F\) regions are identified.}
 	\label{fig:densprofile}
@@
 	\includegraphics[width=0.8\textwidth]{irregularity.pdf}
-	\caption{Schematic of different scale size plasma density irregularities that occur in the polar cap.
+	\caption[Plasma irregularities in the polar ionosphere]{Schematic of different scale size plasma density irregularities that occur in the polar cap.
 Polar patches, orange, move with the background ionospheric convection, indicated by green contours.
+Smaller-scale structures often move develop along the edges of polar patches where strong density gradients exist.
 Field-aligned irregularities, FAIs, are density perturbations that form along the magnetic field lines, pink lines.}
@@
 	\includegraphics[width=\textwidth,angle=180]{SuperDARNmap.pdf}
-	\caption{FoVs of all operational SuperDARN radars in both the northern (left) and southern (right) hemispheres.
+	\caption[SuperDARN map]{FoVs of all operational SuperDARN radars in both the northern (left) and southern (right) hemispheres.
 Polar latitude radars are shown with the red outlines.
@@
 	\includegraphics[width=\textwidth]{ISRmap.pdf}
-	\caption{FoVs of all ISRs in the northern polar region.
+	\caption[ISR map]{FoVs of all ISRs in the northern polar region.
 The individual radars shown are identified by their commonly used acronyms: Poker Flat (PFISR), Resolute Bay North (RISR-N), Resolute Bay Canada (RISR-C), Sondrostrom (SFJ).
@@
 	\includegraphics[width=0.75\textwidth,angle=270]{magnetosphere.pdf}
-	\caption{Diagram of the interactions between the IMF, magnetosphere, and polar cap ionosphere.
+	\caption[Magnetic reconnection footprint in the ionosphere]{Diagram of the interactions between the IMF, magnetosphere, and polar cap ionosphere.
 Closed field lines in the Earth's magnetosphere are shown in grey while open field lines are light blue.
@@
 The X-line can be mapped back along magnetic field lines (orange) to the ionosphere, where their projection is the merging gap (pink dotted line).
-During reconnection events, the open-closed boundary expands equatorwards as more open field lines are formed and there is plasma flow across the merging gap.}
+During reconnection events, the open-closed boundary expands equatorwards as more open field lines are formed and there is plasma flow across the merging gap.
+Figure adapted from \citet{Cowley1991}.}
 	\label{fig:magnetosphere}
@@
 	\includegraphics[width=0.75\textwidth,angle=270]{patch_formation.pdf}
-	\caption{Polar patch formation through transient magnetopause reconnection.
+	\caption[Polar patch formation]{Polar patch formation through transient magnetopause reconnection.
 The open-closed boundary is the green line.
@@
 Dense daytime plasma is shown by the orange regions.
-The burst of reconnection is occurring in Figures \ref{fig:patch_formation}a--\ref{fig:patch_formation}c but stops in Figures \ref{fig:patch_formation}d--\ref{fig:patch_formation}f.}
+The burst of reconnection is occurring in Figures \ref{fig:patch_formation}a--\ref{fig:patch_formation}c but stops in Figures \ref{fig:patch_formation}d--\ref{fig:patch_formation}f.
+Figure adapted from \citet{Cowley1991}.}
 	\label{fig:patch_formation}
@@
 \end{tabular}
-\caption{Irregularity growth rates in the ionosphere for three different limiting cases \citep{Makarevich2016a}.}
+\caption[Irregularity growth rates]{Irregularity growth rates in the ionosphere for three different limiting cases \citep{Makarevich2016a}.}
 \label{tab:gdi_exps}

### 2017-01-18 thesis.sh

@@ -5,14 +5,14 @@ gs -q -dSAFER -dBATCH -dNOPAUSE -sDEVICE=pdfwrite -dDEVICEWIDTHPOINTS=530 -dDEVI
 gs -q -dSAFER -dBATCH -dNOPAUSE -sDEVICE=pdfwrite -dDEVICEWIDTHPOINTS=270 -dDEVICEHEIGHTPOINTS=520 -sOutputFile=/Users/ljlamarche/Desktop/UAFThesis/Figures/SuperDARNmap.pdf /Users/ljlamarche/Desktop/UAFThesis/Figures/SuperDARNmap.ps
 gs -q -dSAFER -dBATCH -dNOPAUSE -sDEVICE=pdfwrite -dDEVICEWIDTHPOINTS=510 -dDEVICEHEIGHTPOINTS=510 -sOutputFile=/Users/ljlamarche/Desktop/UAFThesis/Figures/ISRmap.pdf /Users/ljlamarche/Desktop/UAFThesis/Figures/ISRmap.ps
 
-gs -q -dSAFER -dBATCH -dNOPAUSE -sDEVICE=pdfwrite -dDEVICEWIDTHPOINTS=600 -dDEVICEHEIGHTPOINTS=850 -sOutputFile=/Users/ljlamarche/Desktop/UAFThesis/Figures/month_density_occurance2010.pdf /Users/ljlamarche/Desktop/UAFThesis/Figures/month_density_occurance2010.eps
-gs -q -dSAFER -dBATCH -dNOPAUSE -sDEVICE=pdfwrite -dDEVICEWIDTHPOINTS=600 -dDEVICEHEIGHTPOINTS=850 -sOutputFile=/Users/ljlamarche/Desktop/UAFThesis/Figures/year_month_analysis_line.pdf /Users/ljlamarche/Desktop/UAFThesis/Figures/year_month_analysis_line.eps
-gs -q -dSAFER -dBATCH -dNOPAUSE -sDEVICE=pdfwrite -dDEVICEWIDTHPOINTS=600 -dDEVICEHEIGHTPOINTS=850 -sOutputFile=/Users/ljlamarche/Desktop/UAFThesis/Figures/year_month_analysis_color.pdf /Users/ljlamarche/Desktop/UAFThesis/Figures/year_month_analysis_color.eps
[…]

### 2017-01-20 paper1.tex
paper1 and paper2 added correctly

@@
 \begin{figure}
-\includegraphics[width=12.5 cm,angle=90]{month_density_occurance2010.pdf}
+\includegraphics[width=12.5 cm,angle=90]{p1f1.pdf}
 \caption{Monthly average plots of echo occurrence (left) and peak electron density (right) for 2010.
@@
 \begin{figure}
-\includegraphics[width=11.0 cm,angle=90]{year_month_analysis_line.pdf}
+\includegraphics[width=11.0 cm,angle=90]{p1f2.pdf}
 \caption{The variations in the echo band (a) amplitude of the fitted cosine function \(A\), (b) time shift or time of day of the peak \(B\), (c) average position of the band or the vertical shift of the cosine function from zero \(C\), and (d) width (BW).
@@
 \begin{figure}
-\includegraphics[width=12.0 cm,angle=90]{year_month_analysis_color.pdf}
+\includegraphics[width=12.0 cm,angle=90]{p1f3.pdf}
 \caption{Contour plots of the (a)--(d) echo band range and (e)--(h) occurrence versus UT and season. Each row shows the data for a different year as given in the top-right corner.}
@@
 \begin{figure}
-\includegraphics[width=17.0 cm]{scatter_E2010.pdf}
+\includegraphics[width=17.0 cm]{p1f4.pdf}
 \caption{The echo band range versus the (a) solar zenith angle (SZA) and (b) model peak F2 electron density (NmF2).
@@
 \begin{figure}
-\includegraphics[width=15.0 cm]{IMF_vs_occ_MLT_avg.pdf}
+\includegraphics[width=15.0 cm]{p1f5.pdf}
 \caption{IMF effects in echo occurrence.
@@
 
-\begin{figure}
-\includegraphics[width=11.0 cm,angle=90]{occ_vs_IMF_avg.pdf}
-\caption{Relative occurrence versus IMF By and Bz components for 6 different UT sectors. Each plot cell is color-coded based on the average relative occurrence over that cell.}
-\label{fig:by_bz}
-\end{figure}
+%\begin{figure}
+%\includegraphics[width=11.0 cm,angle=90]{occ_vs_IMF_avg.pdf}
+%\caption{Relative occurrence versus IMF By and Bz components for 6 different UT sectors. Each plot cell is color-coded based on the average relative occurrence over that cell.}
+%\label{fig:by_bz}
+%\end{figure}
 
@@
 \begin{figure}
-\includegraphics[width=18.0 cm]{Fig_RISR.pdf}
+\includegraphics[width=18.0 cm]{p1f6.pdf}
 \caption{Conjugate measurements for two selected events on (left) March 14, 2010 and (right) April 6, 2010.
@@
 \begin{figure}
-\includegraphics[width=12.0 cm,angle=90]{month_ut_combined2010.pdf}
+\includegraphics[width=12.0 cm,angle=90]{p1f7.pdf}
 \caption{Contour plots of the (a) average occurrence within the band, (b) band range, (c) model peak F-region electron density (NmF2) and (d) solar zenith angle (SZA) versus UT and month in 2010.

### 2017-01-20 paper2.tex
paper1 and paper2 added correctly

%\documentclass[angeo, manuscript]{copernicus}

%\begin{document}

%\linenumbers

\chapter{A modeling study of asymmetries in plasma irregularity characteristics near gradient reversals}

%\Author[1]{Leslie J.}{Lamarche}
%\Author[1]{Roman A.}{Makarevich}

%\affil[1]{Geophysical Institute and Department of Physics, University of Alaska Fairbanks, Fairbanks, Alaska, USA}

%\runningtitle{Asymmetries in Plasma Irregularities}

%\runningauthor{Lamarche}

%\correspondence{Leslie J. Lamarche (ljlamarche@alaska.edu)}

%\received{}
%\pubdiscuss{} %% only important for two-stage journals
%\revised{}
%\accepted{}
%\published{}

%\firstpage{1}

%\maketitle

%\begin{abstract}

Asymmetries in plasma density irregularity generation between the leading and trailing edges of the large-scale plasma density structures in the high-latitude ionosphere are investigated. A model is developed that evaluates the gradient-drift instability (GDI) growth rate differences across the gradient reversal that is applicable at all propagation directions and for the broad range of altitudes spanning the entire lower ionosphere. In particular, the model describes asymmetries that would be observed by an oblique scanning radar near density structures in the polar cap such as elongated polar patches. The dependencies on the relative orientations between the directions of the gradient reversal, plasma convection, and wave propagation are examined at different altitudinal regions. At all altitudes, the largest asymmetries are expected for observations along the gradient reversals, e.g. when an elongated structure is oriented along the radar boresite. The convection direction that results in the strongest asymmetries exhibits a strong dependence on the altitude, with the optimal convection being parallel to the gradient reversal in the E region, perpendicular to it in the F region, and at some angle between these extremes in the transitional region. Implications for observations of polar patches by oblique scanning radars within the Super Dual Auroral Radar Network are discussed. It is demonstrated that the wave propagation direction relative to the prevalent convection and gradient directions plays a critical role in controlling both the irregularity growth rate and its asymmetries near gradient reversals.

%\end{abstract}

\section{Introduction}
Plasma density structures in the high-latitude ionosphere range in size from a few centimeters to hundreds of kilometers. In the ionospheric F region, they typically move with the large-scale plasma convection, i.e. their phase velocity is the \(\vec{E}\times\vec{B}\) drift velocity.  Observations of large-scale density structures in the polar cap ionosphere such as polar patches and sun-aligned arcs have shown some asymmetry in the small-scale irregularity characteristics between their leading and trailing edges \citep[e.g.][]{Weber1984}. In general, trailing edges exhibit stronger structuring, which is often interpreted in terms of more favorable conditions for the gradient-drift instability (GDI) along the trailing edge \citep{Weber1984,Milan2002a, Koustov2012,Moen2012}.

The GDI growth rate is generally dependent on the relative directions of the background density gradient vector \(\nabla n_0\) and the convection electric field \(\vec{E}\) or, equivalently, the convection velocity \(\vec{V}_E=\vec{E}\times\vec{B}/B^2\). It is widely accepted that the maximum GDI growth rate in the F region occurs when the density gradient \(\nabla n_0\) is parallel to the convection drift velocity \(\vec{V}_E\) \citep{Keskinen1982a}, while in the E region the maximum occurs when \(\nabla n_0\parallel \vec{E}\). An analytic expression for the GDI growth rate in the F region has been derived in numerous studies, with most of these derivations considering a particular subset of mutual orientations between \(\nabla n_0\), \(\vec{V}_E\), and the wave vector \(\vec{k}\). This typically results in the upper limit of the GDI growth rate of \(\gamma=V_E/L\), where \(L\) is a characteristic gradient scale length defined through \(L^{-1}\equiv |\nabla n_0|/n_0\). More general vector configurations and growth rate expressions in the F region have also been considered \citep{Keskinen1982a,Makarevich2014c}, but their implications for studying the asymmetries between leading and trailing edges of large-scale structures have not been investigated. Further, it has been demonstrated recently that the GDI growth rate patterns change dramatically between the E and F regions \citep{Makarevich2014c}. This means that one expects different signatures of small-scale irregularities and their asymmetries near large-scale structures at different altitudinal regions.

Oblique-sounding coherent radars within the Super Dual Auroral Radar Network (SuperDARN) routinely detect small-scale irregularities in a broad range of altitudes and directions \citep{Chisham2007}. Importantly, both E and F region irregularities are observed by the same radar \citep[e.g. recent papers by][]{Lamarche2015,Forsythe2015} and measurements of the background convection velocity \(\vec{V}_E\) are routinely derived from F-region observations. In addition to their ability to observe large-scale structures in context of background plasma convection, SuperDARN radars provided important observational evidence in support of the GDI mechanism being responsible for the irregularity formation and asymmetries near gradient reversals \citep{Milan2002a,Koustov2012}. Because these radars sample a particular range of the wave vector directions \(\vec{k}\), the issue of the growth rate dependence on \(\vec{k}\) is of critical importance. In particular, it is largely unknown if there are specific \(\vec{k}\) directions or, equivalently, specific radar orientations that are more favorable for observing large-scale ionospheric structures and asymmetries between their leading and trailing edges.

The aim of this study is to investigate the asymmetries between the GDI growth rates near gradient reversals that often occur in the vicinity of large-scale density structures. The specific objectives are to (1) develop a quantitative approach for comparing the growth rates on either side of the structure,  (2) analyse how the growth rates and asymmetries change for different orientations of the background density gradient and convection velocity and determine if there is an optimal radar orientation for which the asymmetry is the strongest, and (3) investigate how the growth rate and asymmetries change between the ionospheric E and F layers.

\section{Modeling of Growth Rates within Radar's Field-of-View}\label{Sec:2}
In this study, we investigate behavior of the GDI growth rate within the field-of-view (FoV) of an oblique scanning radar with a particular focus on asymmetries between the leading and trailing edges of large-scale density structures in the ionosphere. Since growth rate exhibits a strong dependence on the altitude through its dependence on the ratio between the Hall to Pedersen conductivities \(R=\sigma_H/\sigma_P\) \citep{Makarevich2014c}, it is important to understand and model altitudinal dependencies of related parameters.

Figure \ref{fig:1}a presents altitudinal dependencies of several parameters of interest. Shown are the ratios of collision frequencies and gyrofrequencies for electrons \(-r_e=\nu_e/|\Omega_e|\) (blue) and ions \(r_i=\nu_i/\Omega_i\) (green), the anisotropy parameter \(\psi=-r_i r_e\) (pink), and the conductivity ratio \(R=\sigma_H/\sigma_P = \left(r_i+r_e\right)/\left(1+\psi\right)\) (dashed black line on top of the green one). The above notations are the same as in the recent study by \citet{Makarevich2014c}, where the electrons gyrofrequency \(\Omega_e=-|e|B/m_e\) and the ratio \(r_e=\nu_e/\Omega_e\) are negative, while other quantities of interest are positive. The \(x\) axis in Figure \ref{fig:1}a does not have any units since all presented parameters are dimensionless quantities.

The collision frequencies used to calculate these parameters were derived from atmospheric neutral densities found from the Mass Spectrometer - Incoherent Scatter Extended (MSISE) model.  Standard expressions for collision frequencies given by \citet{Schunk1978} were used. MSISE-90 is an empirical model of the densities and temperatures of neutral species in the Earth's atmosphere \citep{Hedin1991}.  For this study, the neutral densities were found for February 15, 2014 at 20 UT at McMurdo station, Antarctica (77.88\(\deg\)S, 166.73\(\deg\)E, geographic), however the parameters considered are expected to be relatively stable so this selection does not constrain the results.

\begin{figure}
\includegraphics[width=\textwidth,angle=180]{f01.pdf}
\caption{(a) Altitudinal variations of parameters of interest. Shown are the ratios of collision frequencies and gyrofrequencies for electrons \(-r_e=\nu_e/|\Omega_e|\) (blue) and ions \(r_i=\nu_i/\Omega_i\) (green), the anisotropy parameter \(\psi=-r_i r_e\) (pink), and the conductivity ratio \(R=\sigma_H/\sigma_P = \left(r_i+r_e\right)/\left(1+\psi\right)\) (dashed black line on top of the green one).  (b) Virtual height of radar scattering volume versus the slant range. The areas shaded with red, green, and blue color refer to the E region, transitional region, and F region, respectively. (c) Geometry of observations and the gradient-drift instability (GDI) growth rate pattern in the radar's field-of-view (FoV). The plasma density gradient reversal is shown by the the heavy white line. The gradient vectors \(\vec{G}=\nabla n_0/n_0\) point towards the reversal in all points within the FoV as shown by representative pink vectors. Also shown are the angle \(\theta\) between the gradient reversal and the horizontal axis (dashed black line),  the angle \(\phi\) between the convection velocity direction \(\vec{V}_E\) (red arrow) and the horizontal axis. The electric field \(\vec{E}\)  and the wave vector  \(\vec{k}\) are shown by the green and yellow arrows, respectively. The normalized GDI growth rate at every cell in the radar's FoV was calculated and the radar cells are colored accordingly, with the color bar shown in the bottom-right corner.  The E region, transitional region, and F region are outlined in red, green, and blue, respectively.}
\label{fig:1}
\end{figure}

We develop a model that is based on a typical observational geometry of individual SuperDARN radars. Each radar within SuperDARN scans through 16 look directions (radar beams) separated by 3.25\(\deg\) in azimuth and receives returns from 75--100 range gates separated by 45 km \citep{Chisham2007}. For an oblique-sounding radar like SuperDARN, different range gates refer to different altitudes, with a virtual height model normally used to characterize this altitudinal dependence of the radar scattering volume as a function of the slant range. Figure \ref{fig:1}b shows the standard SuperDARN virtual height model \citep{Chisham2008}. The ranges where the beam is within the E region (red), F region (blue), and transitional region (green) are indicated. The virtual height model assumes a particular height depending on the slant range. The model is based on where the signal is typically coming from for any given range. In the standard model the virtual height is constant throughout the E and F regions, but changes rapidly in the transitional region between 600 km and 800 km as shown in Figure \ref{fig:1}b. For individual observations in each beam, the height variations will be more gradual, but on average it is represented well by the model shown in Figure \ref{fig:1}b. A revised virtual height model in which virtual height changes with slant range has also been developed \citep{Chisham2008}, but since in the current study we were interested in the differences between the E and F region behavior (rather than small differences within each region), a standard virtual height model has been used.

Figure \ref{fig:1}c presents a typical observational geometry in the horizontal plane. Here we show a FoV of the radar at McMurdo Antarctic station with the geographic South Pole located to the left and down of panel (c) (in the center of concentric circles that show geographic parallels). The same model can be used for any SuperDARN radar in exactly the same way, with the only reason why McMurdo radar is presented here being that this radar often observes large-scale structures that are of interest to this study \citep{Bristow2011,Lamarche2015}. Other high-latitude radars in the network also observe these structures including the Hankasalmi \citep{Milan2002a} and Rankin Inlet \citep{Koustov2012} radars in the northern hemisphere and the present results can be also used to interpret these observations.

Figure \ref{fig:1}c presents a fan plot of the McMurdo radar's FoV with a large-scale structure across the FoV (thick white line). It it assumed that the field of gradient vectors \(\vec{G}=\nabla n_0/n_0\) is uniform in magnitude, with directions always pointing towards the structure throughout the FoV. In other words, the direction of the density gradient reverses at this structure, which will be referred to as the gradient reversal for the remainder of this paper. An example of a structure near which such a reversal may occur is a sun-aligned arc. Three representative pairs of the gradient vector \(\vec{G}\) are shown with pink arrows.

In order to quantify effects due to different structure orientations and motion directions, two angles were used. The first angle describes the structure orientation relative to the radar. This is the angle \(\theta\) between the gradient reversal direction and the direction perpendicular to the boresite (FoV center direction); this perpendicular direction is shown by the dashed horizontal line in Figure \ref{fig:1}c. The second angle \(\phi\) is the angle between the structure drift velocity (red vector in Figure \ref{fig:1}c) and the dashed horizontal line. Since these structures tend to move with the background convection velocity \citep{Makarevich2015b}, this drift velocity is simply \(\vec{V}_E= \vec{E}\times\vec{B}/B^2\). The leading and trailing edges of the structure are determined by the direction of \(\vec{V}_E\).  The direction of the electric field \(\vec{E}\) (green vector) is also defined by \(\phi\) because \(\vec{E}\) is perpendicular to \(\vec{V}_E\). The magnetic field, \(\vec{B}\), is assumed to be straight out of the page, which is approximately consistent with the orientation of the Earth's magnetic field in the southern polar cap and so \(\vec{E}\) is simply rotated counterclockwise by 90\(\deg\) from \(\vec{V}_E\). The direction of the wave vector \(\vec{k}\) (yellow vector) is determined by the location of the point within the radar FoV, as \(\vec{k}\) will be opposite to the direction of the radar beam.

Finally, the GDI growth rates were found for each radar beam-range cell as follows. The location and height of the center of each radar cell was found with a standard positioning method utilizing virtual heights shown in Figure \ref{fig:1}b. The height was used to find the value of the conductivity ratio \(R\) using information presented in Figure \ref{fig:1}a.  Using these quantities, the GDI growth rate at any point within the radar FoV can be calculated \citep[eqn. 18]{Makarevich2014c} using
\begin{equation}
\label{eqn:Mak14gr}
\gamma = \frac{GV_E}{1+\psi}\left(\hat{\vec{k}}\cdot\hat{\vec{b}}\times\hat{\vec{g}}\right)\left(\hat{\vec{k}}\cdot\hat{\vec{e}}-R\hat{\vec{k}}\cdot\hat{\vec{e}}\times\hat{\vec{b}}\right) = \frac{GV_E}{1+\psi}\gamma^\prime.
\end{equation}
Here, \(\hat{\vec{k}}\), \(\hat{\vec{b}}\), \(\hat{\vec{g}}\), and \(\hat{\vec{e}}\) represent unit vectors in the direction of the wave vector \(\vec{k}\), magnetic field \(\vec{B}\), density gradient \(\vec{G}\), and electric field \(\vec{E}\), respectively.  The quantity \(GV_E\) will be the same for all vector orientations and heights. The anisotropy parameter \(\psi\) will change at different heights, but it is always small relative to 1, Figure \ref{fig:1}a, so its effect is minimal.  Since this study focuses primarily on different relative orientations of the gradient and convection directions and altitude regimes, it is convenient to use a normalized growth rate \(\gamma^\prime\).  For the remainder of this paper, the GDI growth rate will refer to the normalized growth rate \(\gamma^\prime\) in arbitrary units, with the prime symbol omitted for brevity.

Each radar cell in Figure \ref{fig:1}c is color-coded in the GDI growth rate at its center, with the color bar given in the bottom-right corner of Figure \ref{fig:1}c.  On the left side of the radar FoV, the growth rate is positive on the leading edge of the gradient reversal and negative on the trailing edge, but both are very close to zero.  Conversely, the growth rate is negative on the leading edge and positive on the trailing edge on the right side of the FoV.  In addition, both the negative and positive values are much larger (\(\sim\mp0.5\)) than on the left side. For this presentation, a gradient field that is uniform in magnitude was assumed; this was done to illustrate the most salient features of the GDI growth rate pattern and in particular effects due to variations in direction and altitude. In practice, the gradient magnitude will also change depending on the distance from the structure, but the directional and altitudinal effects in asymmetries will be the same. This is further examined in Section \ref{Sec:Rev} where a more realistic approach is employed. The important common features are that the GDI growth rate and its asymmetries are strongly dependent on the altitude and direction of observations.  The direction of observation is directly related to the wave propagation direction since for every beam the radar samples a narrow range of propagation directions.  The present study has considered a dependence on the observational direction, but an alternative presentation of the same effect is to consider a fixed scattering volume with variable propagation directions.  This alternate presentation has been shown in \citet{Makarevich2014c}, their figure 1 and figure 2.  These figures show the GDI growth rate in all wavevector directions for set gradient and drift velocity directions for the E and F regions, respectively.

Figure \ref{fig:1}c presented the growth rate pattern for one particular orientation of the gradient reversal and plasma convection. Figure \ref{fig:2} presents the same analysis for four different orientations and locations of the gradient reversal. In all cases, the convection direction is assumed to be perpendicular to the gradient reversal direction (\(\phi-\theta=90\deg\)).  This is because large-scale structures, such as polar patches tend to drift perpendicular to their elongation direction \citep{Makarevich2004c,Makarevich2015b}. This requirement was convenient for presenting different growth rate patterns, but will be relaxed later in this study.

Figure \ref{fig:2}a shows a gradient reversal through the center of the F-region portion of the FoV that is perpendicular to the boresite (\(\theta=0\deg\)).  The growth rate on both sides of the gradient reversal is small (close to zero), but it is slightly negative on the leading edge and slightly positive on the trailing edge. This difference is more evident near the edge beams of the FoV.  In the transitional region, the growth rate is again close to zero near the boresite, but increases to \(\sim0.7\) on either edge of the beam.  The E region shows a completely different pattern, where the growth rate is negative on the left side and smoothly transitions to positive on the right side.  In addition, the most extreme values for growth rate in the E region are above \(\pm1.0\), i.e. much higher than what is observed in the F region.

\begin{figure}
\includegraphics[width=\textwidth]{f02.pdf}
\caption{The GDI growth rate within the radar's FoV for four cases of gradient reversal orientation and position (white line).  In all cases, the convection direction \(\vec{V}_E\) (red arrow) is perpendicular to the gradient reversal direction.  The orientation of the vectors \(\vec{G}\), \(\vec{V}_E\), \(\vec{E}\), and \(\vec{k}\) are shown in each panel, using the same colors as in Figure \ref{fig:1}c.  The GDI growth rate at each radar cell is given by the cell color, with the color bar given in the bottom-left corner of panel (a). Panels (a) and (b) show a gradient reversal centered in the F region, while panels (c) and (d) show the reversal in the transitional region.  The angles \(\theta\) and \(\phi\) (as defined in Figure \ref{fig:1}c) are also given in the bottom-right corner of each panel.}
\label{fig:2}
\end{figure}

Figure \ref{fig:2}b shows an inclined gradient reversal through the FoV in the F region (\(\theta=40\deg\)). Again, the growth rate is negative on the leading edge and positive on the trailing edge, but the difference is much more dramatic.  The growth rate depends on \(\vec{k}\), so the magnitude of difference across the reversal again changes for different beams. This results in a much greater contrast in the growth rate in the right side of the FoV. The normalized growth rate is \(-0.7\) on the leading edge and \(+0.7\) on the trailing edge. This is in contrast with the difference on the FoV left side where the values are only \(\sim\mp0.1\).  The transitional region shows a similar pattern with lower positive values on the left side and growth rate above 1.0 on the right side.  The growth rate is above 1.0 throughout the E region as well.

Figure \ref{fig:2}c presents the same configuration as Figure \ref{fig:2}a, except that a gradient reversal is centered in the transitional region. Similar to Figure \ref{fig:2}a, the growth rate is close to zero on both sizes of the reversal in the center of the radar FoV.  However, there are more extreme values (\(\sim\mp0.5\)) near the edge beams.  Figure \ref{fig:2}d shows an inclined gradient reversal in the transitional region.  The growth rate difference across the gradient is again more extreme further away from the boresite. Overall, this analysis demonstrated that both the gradient reversal orientation and its position in the FoV are important in controlling the asymmetries between the leading and trailing edges.

\section{Modeling of Growth Rate Differences Across the Gradient Reversal}\label{Sec:Rev}

In order to quantify the above-noted differences in the growth rate for different orientations of the gradient reversal, the average change in the growth rate across the reversal \(\Delta\gamma_{avg}\) was found for any general configuration of the gradient reversal and the convection direction. The approach presented in Figures \ref{fig:1}c and \ref{fig:2} was useful for presenting and emphasizing general spatial patterns in the growth rate. In this section, we employ two new methods that are more robust, independent of the way the the radar FoV is divided into beam-range cells, and more suitable for evaluating the growth rate differences in the vicinity of large-scale structures. These two new methods are described below.

\subsection{Numerical Method}
\label{sec:num}
In the first method, evenly spaced points were selected along the gradient reversal within the radar FoV. These points did not necessarily correspond to real range gates or beams used in the standard SuperDARN radar configuration.  A virtual radar beam was constructed that intersected each point.  From this, the virtual slant range and beam azimuth were found, which were used to find the wave vector \(\vec{k}\) and the height corresponding to that point.  The growth rates on opposite sides of the gradient reversal were found as follows.  For the leading edge, the gradient vector \(\vec{G}\) is set to be perpendicular to the gradient reversal and in a direction such that the component of \(\vec{V}_E\) along \(\vec{G}\) is negative, i.e. \(\vec{V}_E\cdot \vec{G}<0\). For the trailing edge, the opposite direction is selected such that \(\vec{V}_E\cdot \vec{G}>0\). These two cases can be thought of as the limiting cases when two points on either side of the reversal are both moved towards the reversal until they are both colocated on it.  The GDI growth rate for both cases can then be calculated using Eq. (\ref{eqn:Mak14gr}) as described above.  The difference in growth rate across the reversal, \(\Delta\gamma\), can then be found by subtracting the trailing edge case from the leading edge case.  The average difference in growth rate across the reversal for the entire FoV, \(\Delta\gamma_{avg}\), can be found by averaging \(\Delta\gamma\) of every point initially considered along the gradient reversal.

\subsection{Analytic Method}\label{sec:ana}

An alternative approach is to develop an analytic expression for \(\Delta\gamma_{avg}\) for any general direction of the gradient reversal \(\theta\) and convection \(\phi\) as described in Appendix \ref{app:1}. The result is
\begin{equation}
\label{eqn:gr}
\Delta\gamma_{avg}=H(\theta,\phi)\left[\sin(\phi-\theta)-C\sin(\phi+\theta)+R\left(\cos(\phi-\theta)-C\cos(\phi+\theta)\right)\right], 
\end{equation}
with 
\begin{equation}
\label{eqn:Heaviside}
H(\theta,\phi)=\Theta(\phi-\theta+\pi)-\Theta(\theta-\phi-\pi)+2\Theta(\phi-\theta-\pi)-2\Theta(\phi-\theta),
\end{equation}
function \(\Theta\) being the Heaviside step function, and \(C=0.87\) for SuperDARN observations, Appendix \ref{app:1}. This expression can be readily used to compute \(\Delta\gamma_{avg}\) for any given combination of \(\theta\) and \(\phi\). One limitation of this expression is, however, that it can only be applied when the conductivity ratio \(R\) is approximately constant, i.e. in the E region or in the F region, but not in the transitional region.

\subsection{General Vector Orientations}

The two methods for finding the average difference in the GDI growth rate across the gradient reversal can now be employed to characterize different orientations. In this section, we consider a more general case of different relative orientations of the gradient reversal and the convection direction, not necessarily restricted to the case of \(\theta- \phi= 90\deg\) that was considered in Section \ref{Sec:2}. In other words, we consider all possible differences \(\theta- \phi\).

Figure \ref{fig:3} examines the growth rate difference across the gradient reversal for 4 selected angle differences \(\theta- \phi\) as indicated in the top-right corner of each panel. In each case, the average growth rate difference, notated simply by \(\Delta\gamma\) for the remainder of this paper, is calculated for different \(\theta\) values from \(0\deg\) to \(360\deg\). In each plot, the convection direction \(\phi\) changes with \(\theta\) so that their difference remains constant. The average growth rate difference is shown on the \(y\) axis in arbitrary units. The numerical method results, Section \ref{sec:num}, are shown for the averages (solid lines) and standard deviations (shaded regions). The blue and red curves correspond to the E- and F-region cases. The analytic method results, Section \ref{sec:ana}, are also presented for the expected average differences (dashed lines). Because the growth rate difference is calculated by subtracting the trailing edge \(\gamma\) from the leading edge \(\gamma\), negative \(\Delta\gamma\) indicates larger instability growth on the trailing edge than the leading edge. In addition, large \(\Delta\gamma\) (either positive or negative) indicates a strong contrast between the two sides of the gradient reversal.

\begin{figure}
\includegraphics[width=\textwidth,angle=180]{f03.pdf}
\caption{The difference in the GDI growth rate across the gradient reversal for four selected angles between the reversal and convection direction.  The gradient reversal direction \(\theta\) is on the \(x\)-axis and the average growth rate difference across the reversal \(\Delta\gamma\) is on the \(y\)-axis for each panel.  The difference between \(\phi\) and \(\theta\) is given in the top-right corner of each plot. The difference in the average growth rate \(\Delta\gamma\) computed using a numerical approach (see text for details) is shown by the solid lines, with the shaded regions representing the standard deviation.  The dashed lines show the behavior of \(\Delta\gamma\) calculated using an analytic expression. The trends for both the E-region (red) and F-region (blue) cases are shown in each panel.}
\label{fig:3}
\end{figure}

Figure \ref{fig:3}a shows  a case of \(\phi-\theta=0\deg\), i.e. the case when the convection direction is parallel to the gradient reversal. This corresponds to \(\vec{V}_E \perp \vec{G}\) or \(\vec{E} \parallel \vec{G}\) on the trailing edge, which is the condition for most favorable gradient orientation in the E region \citep{Makarevich2014c}. The growth rate difference in the E region (red) is negative at all gradient reversal directions \(\theta\) and it changes with \(\theta\) in a sinusoidal fashion. It has the lowest \(\Delta\gamma\) minimum values observed out of all 4 cases (\(\sim-10\) at \(90\deg\) and \(270\deg\)). The average growth rate difference for the F region (blue) also shows a periodic behavior with \(\theta\); it is positive for \(0\deg\)--\(90\deg\) and \(180\deg\)--\(270\deg\) and negative for all other angles. The amplitude of oscillations is much smaller than in the E region. The segments where \(\Delta\gamma>0\) correspond to the growth rate on the leading edge exceeding that on the trailing edge. However, \(\Delta\gamma\approx 1\) and \(\Delta\gamma\approx -1\) are the maximum and minimum values, respectively, so the contrast across the gradient reversal is relatively small regardless of whether the leading or trailing edge has the greater growth rate. The important conclusion from this analysis is that the contrast or the asymmetry magnitude depends strongly on the structure orientation angle \(\theta\), which confirms the previous results based on two angles \(\theta\) only, Figures \ref{fig:2}a and b.

Figures \ref{fig:3}b and \ref{fig:3}c present the same analysis, but with a set angle of \(30\deg\) and \(60\deg\), respectively, between the gradient reversal and the convection direction.  These show the gradual transition between the extreme cases of the convection direction parallel to the gradient reversal, Figure \ref{fig:3}a, and the convection direction perpendicular to the reversal, Figure \ref{fig:3}d.  The general E-region pattern starts to move towards more positive values and the F-region pattern moves towards more negative values.  The location of the peaks in growth rate difference changes as well. For the E region, the location of the lowest \(\Delta\gamma\) moves gradually from \(\theta\approx 90\deg, 270\deg\) when \(\phi-\theta=0\deg\) to \(\theta\approx 45\deg, 225\deg\) when \(\phi-\theta=90\deg\). Similarly, for the F region, the location of the lowest \(\Delta\gamma\) moves from \(\theta\approx 135\deg, 315\deg\) to \(\theta\approx 90\deg, 270\deg\) between these two situations.

Figure \ref{fig:3}d shows the case when the convection direction is perpendicular to the gradient reversal, which corresponds to \(\vec{V}_E \parallel \vec{G}\).  From previous studies, this is the condition where the growth rate on the trailing edge is the greatest in the F region.  Similar to Figure \ref{fig:3}a, the growth rate difference in Figure \ref{fig:3}d for the F region (blue) is negative for all directions of the gradient reversal and has the lowest values observed for the F region (\(\sim-2\)) at \(90\deg\) and \(270\deg\). This corresponds to the largest contrast between the leading and trailing edge of the structure. Thus the largest difference is observed when the growth rate peaks. This only true however for the largest difference, i.e. at a particular $\theta$ value where this largest magnitude is found. For other \(\theta\) values, the difference may be larger or smaller than at other values of \(\phi-\theta\). The E-region (red) extreme values (maximum \(\Delta\gamma\approx 4\) and minimum \(\Delta\gamma\approx -6\)) have the lowest magnitude out of all 4 cases shown in panels (a)--(d).  This indicates that the contrast in the E region is the smallest for this particular orientation, although it is still much larger than that observed in the F region.

Throughout all 4 cases, there are several common features. The amplitude of variations of the growth rate difference in the F region is consistently much less than that in the E region.  Generally, there is very good agreement between the numerical method and the analytic method in both the E and F regions.  Typically, the results from the analytic method fall well within one standard deviation of the numerical method.  The exceptions to this occur when \(\theta\) is \(90\deg\) or \(270\deg\).  The discrepancy in the results in these locations is due to the fact that the two methods have some inherent differences. In these cases, the gradient reversal structure passes through the radar FoV along its boresite or central beam, so only points along that line (and that azimuth) are considered for the numerical approach, while the whole range of azimuths in the analytic approach is still integrated over. For most gradient reversal orientations, however, this difference is negligible.

The analysis presented in Figure \ref{fig:3} can be extended to include any arbitrary orientation between the gradient reversal and convection direction by allowing \(\phi\) and \(\theta\) to vary independently.  Both the numerical method and the analytical method can be used to find the average difference in growth rate for any direction of the gradient reversal and plasma convection in the E and F regions.  Figure \ref{fig:4} presents the results of this analysis in the form of contour plots of the growth rate difference \(\Delta\gamma\) versus angles \(\theta\) and \(\phi\). Figures \ref{fig:4}a and \ref{fig:4}b show the results of the analytic method in the E and F regions, respectively. The remaining 4 panels show the results from the numerical method in (c) the E region, (d) the F region, (e) transitional region, and (f) all regions combined. With exception of Figure \ref{fig:4}e, different values of \(\Delta\gamma\) along the gradient reversal were included in the averaging only for the points with specified virtual heights, e.g. 115 km for Figure \ref{fig:4}c. In Figure \ref{fig:4}f, all points along the gradient reversal within the FoV were considered.

The discontinuities in \(\Delta\gamma\) that occur on diagonal lines \(\phi=\theta\), \(\phi=\theta+180\deg\), \(\phi=\theta-180\deg\) are due to the fact that the convection velocity vector crosses the gradient reversal at these points and the leading and trailing edges of the gradient reversal switch.  This feature is common in all panels in Figure \ref{fig:4}. On these discontinuities, the \(\Delta\gamma\) values always switch their sign, but their magnitude is the same on both sides of the discontinuity. An alternative presentation would be to only show magnitude, but the current version also presents information about the sign and the sense of the change, which is also useful.

\begin{figure}
\includegraphics[width=\textwidth,angle=180]{f04.pdf}
\caption{Contour plots of the growth rate difference \(\Delta\gamma\)  versus the two angles of interest. The direction of the gradient reversal \(\theta\) is on the \(x\)-axis and the convection direction \(\phi\) is on the \(y\)-axis.  The color bar is given to the right. Panels (a) and (b) show the results using an analytic approach in the E and F regions, respectively.  Panels (c)--(f) show the results using a numerical approach in the E, F, transitional, and all regions, respectively. The dashed black and white line in panels (c) and (f) indicates the average value of the conductivity ratio \(R\) along the gradient reversal for different angles \(\theta\) in the transitional region and all regions, respectively. The scale for \(R\) is shown on the right of panels (c) and (f).}
\label{fig:4}
\end{figure}

Figure \ref{fig:4}a shows the average growth rate difference across the gradient reversal found by the analytic method in the E region.  The same information but for 4 specific cases of \(\phi-\theta\) were shown in Figure \ref{fig:3}.  Specifically, each dashed red line in Figure \ref{fig:3} represents one diagonal cross section in Figure \ref{fig:4}a. For example, the red dashed line in Figure \ref{fig:3}a refers to the main diagonal line \(\phi=\theta\) and the red dashed line in Figure \ref{fig:3}d refers to the line \(\phi=\theta+90\deg\). The peak magnitudes in \(\Delta\gamma\) occur at \(\theta=90\deg,270\deg\) and \(\phi=90\deg,270\deg\).  This agrees with what is shown in Figure \ref{fig:3}a. Physically, this means that the largest contrast in growth rate across the gradient reversal occurs when the reversal is passing through the radar FoV along the boresite and convection is parallel or antiparallel to the reversal. The lowest magnitudes of \(\Delta\gamma\) occur approximately at the lines \(\theta=0\deg\), \(\theta=180\deg\), \(\phi=0\deg\), and \(\phi=180\deg\). At these locations the difference is close to zero (green color). This means that contrast in the growth rate is the least if either the gradient reversal or the convection direction is perpendicular to the boresite.

Figure \ref{fig:4}b presents the same information but for the F region. This refers to the blue dashed lines in Figure \ref{fig:3}.  In this case, the peak magnitudes in \(\Delta\gamma\) occur at \(\theta=90\deg,270\deg\) and \(\phi=0\deg,180\deg\).  This means that the greatest contrast in growth rate occurs when the gradient reversal is parallel to the boresite and the convection is perpendicular to it.  The smallest magnitudes of \(\Delta\gamma\) occur at the lines \(\theta=0\deg\), \(\theta=180\deg\), \(\phi=90\deg\), and \(\phi=270\deg\). This means that the least contrast across the reversal will occur when the gradient reversal is perpendicular to the boresite or when the convection direction is either parallel or antiparallel to the radar boresite.

Figure \ref{fig:4}c and d present the growth rate difference in the E and F regions, respectively, found with the numerical method.  A comparison of Figures \ref{fig:4}a and b with Figures \ref{fig:4}c and d shows that the results of the analytic and numerical methods are very similar for both the E and F regions. The results for the numerical method (Figures \ref{fig:4}c and d) are not as smooth as the analytic method results, but this is simply due to the discrete steps that are used in the numerical method. Comparison between the E and F region patterns, however, reveals a few interesting features. First, the location of the minima moves from \(\phi=90\deg,270\deg\) in the E region to \(\phi=0\deg,180\deg\) in the F region.  These results agree with what was observed in Figure \ref{fig:3}.  In addition, there are large regions of strongly positive \(\Delta\gamma\) (dark red) in the E region (Figures \ref{fig:4}a and \ref{fig:4}c) that do not appear in the F region.  There are some areas of positive \(\Delta\gamma\) in the F region, but they are much more limited and have a maximum value of only \(\sim 1\).  This means that while there are significant areas where the growth rate on the leading edge is greater than that on the trailing edge in the E region, the growth rate on the trailing edge almost always exceeds the leading edge in the F region. This is an important generalization of the well-known result that growth rates are more positive on the trailing edge, where \(\vec{G}\cdot \vec{V}_E>0\).

Figures \ref{fig:4}e shows results of the numerical analysis of the transitional region. This panel does not have a comparison with the analytic method because the analytic method requires a constant conductivity ratio \(R\) which changes rapidly from \(\sim\)5.08 in the E region to 0.001 in the F region, Figure \ref{fig:1}a. To demonstrate this, the average \(R\) value along the gradient reversal is plotted using the black-white dashed line and the scale shown on the right. The ratio depends on the gradient reversal direction \(\theta\), but not on the convection direction \(\phi\).  As angle \(\theta\) changes, the gradient reversal passes through different portions of the FoV with different heights. Different altitudes correspond to hugely different values of \(R\), which causes the variation in the conductivity ratio shown in Figure \ref{fig:4}e. On average, however, the conductivity ratio is somewhere in between its values in the E and F regions, so the pattern shown is between the E region pattern and F region pattern. The minimum is between \(\phi=90\deg\) (the E region value) and \(\phi=180\deg\) (the F region value) and has a value of \(\sim -3\). This will be further examined in Figure \ref{fig:5}.

Finally, Figure \ref{fig:4}f presents the same analysis for all regions considered together, i.e. with averaging done over all points along the gradient reversal regardless of their virtual height. For most angles of the gradient reversal, only the F region will be sampled, Figure \ref{fig:1}c. The gradient reversal will only pass through the transitional and E region if it is close to parallel with the boresite, i.e. when \(\theta\) is close to either \(90\deg\) or \(270\deg\).  This is why the majority of the pattern in Figure \ref{fig:4}f resembles the F region patterns (Figure \ref{fig:4}d), but changes sharply around \(\theta=90\deg\) and \(\theta=270\deg\) due to contributions from the E region pattern (Figure \ref{fig:4}c).  This can also be seen in the conductivity ratio. For most angles, average \(R\) only has contribution from the F region, so it is at the F-region value (very close to zero on this scale).  However, close to \(\theta=90\deg\) and \(\theta=270\deg\) average \(R\) includes E-region values (\(\sim5\)), so it increases sharply.  This causes \(\Delta\gamma\) to be relatively low with values between \(-1.5\) and \(+1.5\) for most angles \(\theta\); it changes around \(\theta=90\deg\) and \(\theta=270\deg\) to have values less than \(-2\).  Overall, this analysis shows that both the peak asymmetry and where it is seen in \(\theta\) and \(\phi\) change significantly with altitude.

\subsection{Behavior in the Transitional Region}
Figure \ref{fig:5} shows a more detailed analysis of the transitional region.  The transitional region was divided into segments that were 2 km in slant range. In each of these segments, the analysis presented in Figure \ref{fig:4} was performed and the average height and average conductivity ratio were found.  Figure \ref{fig:5}a shows the average conductivity ratio (black line) and height (orange line) from 550 km to 850 km in slant range.  The ranges where the beam is in the E (red), F (blue), and transitional (green) regions are shaded. Both the height and the conductivity ratio are constant in the E and F regions since virtual height is constant at these ranges in the standard virtual height model, Figure \ref{fig:1}b. The height increases linearly in the transitional region, while the conductivity ratio decreases rapidly between 600 km and 616 km (shaded in yellow) at the lowest ranges within the transitional region. The \(R\) value continues to decrease throughout the transitional region, but above \(\sim620\) km in the slant range, it is so close to zero that further changes do not affect the results appreciably.  The region highlighted in yellow, where most of the change in \(R\) occurs, is examined more closely in Figures \ref{fig:5}b and \ref{fig:5}c.

\begin{figure}
\includegraphics[width=\textwidth,angle=180]{f05.pdf}
\caption{Analysis of the transitional region.  (a) The average conductivity ratio \(R\) (black line) and the radar virtual height (orange line) versus the slant range for the entire transitional region and parts of the E and F regions. The scales for the conductivity ratio and height are given on the left and right of panel (a), respectively.  The E (red), transitional (yellow and green) and F (blue) regions are highlighted.  The region highlighted with yellow is examined closer in panels (b) and (c).  Panel (b) is similar to panel (a), but focuses on slant ranges 600--616 km, where most of the change in the conductivity ratio occurs.  In addition to the \(R\) and height continuous variations, it also gives the average value of \(R\) for each 2-km virtual range gate. The same 2-km virtual gates are considered in panel (c) that shows a similar analysis to Figure \ref{fig:4}. The top row shows the results of the numerical approach (N) and the bottom row shows the analytic approach results (A).}
\label{fig:5}
\end{figure}

Figure \ref{fig:5}b has the same format as Figure \ref{fig:5}a, but it focuses on the 16-km-wide region in the slant range highlighted with yellow. In these 16 km, the conductivity ratio (black line) decreases by approximately one order of magnitude, from \(\sim\)5 to 0.5. The height (orange line) increases from \(\sim\)110 km to 130 km. These changes make a large difference in the patterns of difference of growth rate, as described below.

Each small panel in Figure \ref{fig:5}c has the same format as Figure \ref{fig:4}a. The top row shows the numerical analysis and the bottom show the analytic analysis of each 2 km segment.  The value of \(R\) for the analytic method was taken from the average \(R\) for each 2 km segment, as shown in Figure \ref{fig:5}b. From 600--602 km, \(R\) is close to the E region value (\(\sim5.1\)), so the patterns for \(\Delta\gamma\) are close to those for the E region (Figures \ref{fig:4}a and c).  However, by 614--616 km, the value of \(R\) has decreased substantially and the patterns resemble the F-region patterns (Figures \ref{fig:4}b and d) much more closely.  From 600 km to 616 km, the location of lowest value of \(\Delta\gamma\) changes from \(\phi=90\deg,270\deg\) to \(\phi=0\deg,180\deg\).  This represents the change from where the lowest \(\Delta\gamma\) values are in the E region to their location in the F region. In addition, the lowest value of \(\Delta\gamma\) also increases from less than \(-8\) (E region) to \(\sim-3\) (F region).  The areas of highly positive \(\Delta\gamma\) that appeared in the E region (Figures \ref{fig:4}a and c) are not evident above 606 km. This indicates that after this point, there is no longer a significant subset of orientations where the growth rate on the leading edge is larger than that on the trailing edge.

\section{Discussion}
In this study, we have applied the GDI linear fluid theory to study the asymmetries in the GDI growth rate on either side of a large-scale density structure, particularly as applied to observations with oblique scanning radars such as SuperDARN.  The important new aspect of this study is explicit consideration of varying wave vector direction and altitude and their effects on asymmetries in the plasma structuring processes. We focus on the first-order effects due to the differences in the GDI growth rate that are often involved in interpreting asymmetries in intensity and occurrence characteristics of  small-scale plasma irregularities in the high-latitude ionosphere.  Below, we discuss the applicability of these results and their implication for interpretation of radar observations near gradient reversals.

The original expression used to define the GDI growth rate, Equation \ref{eqn:Mak14gr}, has been derived under the assumption of negligible inertial effects \citep{Makarevich2014c}, i.e. in the low-frequency and long-wavelength limit: \(\omega\ll \nu_{in}\), \(\lambda\gg 2\pi V_E \nu_{in}^{-1}\). The fluid theory itself also applies directly to long wavelengths, i.e scales much larger than the ion mean free path. The observations with which we compare our results refer to small-scale irregularities and, more specifically, decameter-scale waves in the E and F regions as well as in the transitional region, in the case of SuperDARN observations. In the E region up to 130 km in altitude, these conditions may be satisfied for decameter scales \citep[e.g.][]{Fejer1984}. In the F region, the long-wavelength case refers to significantly larger scales (\(\sim\) 1000 m) below which the fluid theory is not necessarily directly applicable. However, observational studies have found the behavior of decameter-scale irregularities in the F region to be consistent with many fluid theory predictions, including their propagation with the \(\vec{E}\times\vec{B}\) drift velocity and stronger waves on the trailing edge of gradient reversals. This leads to a common interpretation that decameter-scale waves observed by SuperDARN and  kilometer-scale waves in the F region described by a standard fluid approach are closely related, possibly by a turbulent cascade \citep[e.g.][]{Tsunoda1988}.  Within the above discussed limits of applicability of the basic equation (\ref{eqn:Mak14gr}) employed in the present modeling, GDI is operational throughout the wide range of ionospheric altitudes spanning both the E and F regions. This is because there is always a range of propagation and gradient directions for which growth rates are positive in both the E and F regions.

In this study we used the SuperDARN radar at McMurdo Station, Antarctica to model spatial variations of the GDI growth rate within this radar's field-of-view and the resulting asymmetries near gradient reversals. Despite this, the results should be also applicable to any of the high-latitude SuperDARN radars in either the northern or southern hemisphere which observe large-scale density structures with gradient reversals. To be more accurate, the parameters of the neutral density model would have to be changed to match a particular location, but these corrections would be very minor. More generally, these results should be the same for any oblique scanning coherent scatter radar at high latitudes.  In principle, this technique could also be applied to auroral and mid-latitude radars, but at these latitudes, the Earth's magnetic field is further away from the vertical, which affects the assumption that \(\vec{B}\) is perpendicular to the radar's FoV and all other vector quantities are within the same plane. In addition, extended gradient reversals are much rarer at mid- and auroral latitudes than in the polar cap, so this model may be less applicable for interpreting observations in these regions.

Although this study is relevant to observations with obliquely scanning HF radars, care should be taken with how the results are used to interpret experimental data.  In this study, only the linear growth rate of GDI was considered. Even though radar backscatter power and occurrence trends are often interpreted using the GDI growth rates in the linear regime \citep[e.g.][]{Oksavik2012,Moen2012}, there are other factors that may affect echo power and occurrence.  In the polar cap, HF radars rely on beam refraction through a dense ionosphere to observe field-aligned irregularities \citep{Bristow2011,Koustov2014}.  In addition, high background electron density in both the E and F regions can smooth gradients required for GDI and limit instability growth \citep{Ruohoniemi1997,Koustov2004}.  Large conductivity in the E region can also short out F region irregularities \citep{Danskin2002,Kane2012,Lamarche2015}.  An enhanced D region may act to attenuate backscatter signals, although this effect is not always obvious \citep{Dansin2002}.  Nonlinear processes in the F region have been theoretically shown to alter the dependence on the propagation direction while still showing significant anisotropy similar to that predicted by the linear theory \citep{Keskinen81,Keskinen82a,Keskinen83}.  The situation in the E region appears to be somewhat different, with some echoes also exhibiting strong anisotropy \citep[e.g.][]{Ierkic1980,Makarevich2008} and other echoes showing isotropic power variation \citep{Ierkic1980,Koustov2001} presumably because of nonlinear processes.  The results of the present study are thus more likely to apply in the F region.

The present model used an idealized gradient reversal superimposed on the radar's FoV, meant to simulate an elongated structure with higher density than the background, such as elongated polar patches \citep{Hosokawa2014} or sun-aligned arcs \citep{Koustov2012}.  In reality, any structure like this will be spread spatially, and the pattern of GDI growth rate around it will be more complicated.  In addition, plasma flows around the gradient reversal may also be more complicated than the simple uniform flow considered here.  Sun-aligned arcs in particular are known to have complex shears in plasma flow locally across the density enhancement due to converging electric fields \citep[e.g.][]{Kozlovsky2007}, which would complicate growth rate calculations. However, the present model focused on the large-scale motion of density structures, which tends to be relatively consistent on time scales of at least half an hour, and may be still valid for an overall picture of where greater structuring is anticipated.

The results from Figure \ref{fig:4} show that \(\Delta\gamma\) is almost always negative in the F region, which implies that the growth rate on the trailing edge of a large-scale density structure is almost always larger than that on the leading edge.  This agrees with previous observations of polar patches and sun-aligned arcs in the F region ionosphere that consistently show a greater irregularity intensity on the trailing edge than the leading edge of these structures \citep{Weber1984,Milan2002a,Koustov2012}.  Greater power from radar backscatter is indicative of more decameter-scale structuring, which agrees with a greater growth rate on the trailing edge.  Some studies have shown radar backscatter (and presumably plasma structuring) occurs within the entire volume of polar patches \citep{Hosokawa2009}.  The present model cannot replicate this result as it does not consider structures of arbitrary shapes.  The observed patches are often irregularly shaped, which could result in a limited number of places along their edges where the relative gradient/plasma drift orientation is such that no growth would occur \citep{Hosokawa2009}.  In this way, these observations are not necessarily inconsistent with the results presented here, but they cannot be represented by the model in its current form.

In the E region, large values of \(\Delta\gamma\) that are both positive and negative indicate that the growth rate can be high on both the leading and trailing edge.  Several previous studies have investigated large-scale density structures in the E region \citep{Milan1999b,Milan2001a,Milan2002a}, but none of these have focused on asymmetries in irregularity characteristics associated with gradient reversals so it is difficult to compare these modeling results with observations.

\citet{Koustov2008} presented data in their figure 5 that showed smaller echo power on the leading edge of a sun-aligned arc in the F region, but strong backscatter power on both sides of the arc in the short range gates, corresponding to the E-region or transitional region.  The arc was approximately parallel to the radar boresite and moved towards the left of the radar FoV, corresponding to \(\Delta\gamma\approx 0\) in our model (\(\theta=90\deg,\phi\approx 180\deg\) in figure \ref{fig:4}c).  Slight changes in the convection direction (\(\phi\)) could create both the condition where \(\Delta\gamma>0\) (greater growth on leading edge) or \(\Delta\gamma<0\) (greater growth on trailing edge).  Although this proposed mechanism seems reasonable, it is important to note that the sun-aligned arc examined in \citet{Koustov2008} has been mapped to an altitude of 250 km and there is no information about its vertical extent, so it is possible that the density enhancement does not reach the E region at all and the structuring is due to a different process.

An important new result of this study was that the difference in growth rate depends strongly on the orientation of the gradient reversal and convection direction in both the E and F regions. In general, the maximum \(\Delta\gamma\) occurred at \(\theta=90\deg,270\deg\) in both the E and F regions, regardless of \(\phi\).  This corresponds to the gradient reversal being aligned along the radar boresite with the gradient direction being perpendicular to the boresite. On one hand, this results is perhaps not surprising since this is the direction where the peak of the growth rate \(\gamma\) itself occurs \citep{Makarevich2014c}. On the other hand, we found that this is also the direction of the peak in the difference \(\Delta\gamma\) and in the average difference \(\Delta\gamma_{avg}\) and this latter result in particular does not immediately follows from the growth rate behavior itself.

In the F region, the peak in the \(\Delta\gamma_{avg}\) magnitude occurs when the convection direction is perpendicular to the radar boresite \(\phi=90\deg\). In the E region, the peak is when the convection direction is either parallel or antiparallel to the boresite \(\phi=0\deg, 180\deg\). This corresponds to the convection direction being parallel to the wave vector. In the transitional region, the ``optimal'' convection direction changes gradually from \(\phi=0\deg\) to \(\phi=90\deg\). This is a prediction that has not been tested experimentally. To test it, one would need to follow the track of a large elongated polar patch or sun-aligned arc.  Although polar patches can retain their form in the F region for several hours \citep{Hosokawa2014}, it is likely that the patch would move a significant lateral distance from the boresite.  The structure may have to be observed by multiple SuperDARN radars from different directions unless a highly elongated patch (>1500 km) is observed, similar to that described in \citet{Hosokawa2014}.

Another challenge in observing altitudinal transitions is associated with a requirements for a large density structure to have a significant vertical extent, at least from the lower F region through a significant portion of the transitional region. The vertical extent of these structures has not been investigated systematically, although several examples of density enhancements associated with these structures have been presented. Thus \citet{Schlatter2013} examined density enhancements due to high-energy auroral precipitation in the E and F region between \(\sim100\) km and \(\sim150\) km, see their Figure 1a. \citet{Jayachandran2012} estimated the maximum vertical thickness of sun-aligned arcs as \(\sim84\) km.  Sun-aligned arcs have been modeled to have regions of enhanced density at altitudes as low as 100 km and as high as 300 km \citep{Crain1994}.  This is somewhat  supported by observations that show high-density areas at a similar range of altitudes \citep{Valladares1991}. These previous investigations thus provide some support for the idea that it may be possible to find a single density structure with large enough vertical extent to cover the necessary range.

\section{Conclusions}
A model of asymmetries in plasma wave generation due to sharp reversals in the plasma density gradient near elongated density structures in the high-latitude ionosphere revealed the following:

\begin{enumerate}
\item Growth rate of plasma irregularities generated in the linear regime of the gradient-drift instability exhibits asymmetry between leading and trailing edges of convecting elongated density structures such as polar patches.  Both the GDI growth rate and its asymmetry expressed as the average difference between growth rates on the leading and trailing edges are strongly dependent on the wave propagation direction. Asymmetry is maximized for wave vectors parallel or antiparallel to the elongation direction of the gradient reversal at any altitude in the lower ionosphere.
\item Asymmetry is also strongly dependent on the background convection direction relative to the gradient reversal, but this dependence changes with altitude. Maximum asymmetry occurs when the background convection is parallel to a gradient reversal in the E region, but perpendicular to a gradient reversal in the F region. There is a relatively smooth transition between configurations associated with maximum asymmetry in the E and F layers. In the E region, either the leading or trailing edge can have significant positive (or negative) growth rate, resulting in the asymmetry favoring either edge, depending on the structure orientation. This is in contrast with the F region, where growth rate on the trailing edge is almost always larger than the growth rate on the leading edge, but with a much lower absolute maximum than in the E region. This strong altitudinal dependence is due to the rapid decrease of the ratio of Hall and Pedersen conductivities at higher altitudes.
\item An oblique scanning radar is optimally oriented to observe large-scale and elongated density structures in the high-latitude ionosphere when the radar's boresite is closest in direction to the structure elongation for any given altitude and convection direction since this configuration is associated with the largest contrast between leading and trailing edges. The convection direction that leads to the largest contrast depends on the altitude and for HF radars that predominantly receive backscatter from the F region the optimal convection direction is perpendicular to the boresite.
\end{enumerate}

\appendix
\section{Analytic Expression for the GDI Growth Rate Difference}\label{app:1}

In this section, we derive an analytic expression for the average difference in the GDI growth rate across the gradient reversal passing through the radar's viewing area, Figure \ref{fig:1}c. The starting point is Equation (\ref{eqn:Mak14gr}). Consider a single point that lies on the gradient reversal within the radar's FoV. In the geometry illustrated in Figure \ref{fig:1}c, the relevant vectors at this point can be written as \(\vec{\hat k}=-\sin\alpha\vec{\hat{e}}_1-\cos\alpha\vec{\hat{e}}_2\), \(\vec{\hat e}=-\sin\phi\vec{\hat{e}}_1+\cos\phi\vec{\hat{e}}_2\), \(\hat{\vec{g}}=\pm[\sin\theta\vec{\hat{e}}_1-\cos\theta\vec{\hat{e}}_2]\), where angles \(\theta\) and \(\phi\) were previously defined in Section \ref{Sec:2}, angle \(\alpha\) is the angle between the wave vector and the boresite direction, and unit vectors \(\vec{\hat{e}}_1\)  and \(\vec{\hat{e}}_2\) are  perpendicular and parallel to the boresite, respectively.

Using these expressions, all the terms in Equation \ref{eqn:Mak14gr} can be expressed as functions of \(\alpha\), \(\theta\), and \(\phi\)
\begin{equation}
\gamma = \pm\sin(\alpha+\theta)[\cos(\alpha+\phi)-R\sin(\alpha+\phi)].
\end{equation}
This is the GDI growth rate at a given point for the two opposing gradient directions that will occur on either side of the gradient reversal.  If \(\phi<\theta-180\deg\) or \(\theta<\phi<\theta+180\deg\), the negative sign represents the leading edge and the positive sign represents the trailing edge, so the growth rate difference is negative.  If \(\phi<\theta\) or \(\theta+180\deg<\phi<\theta+360\deg\), the positive sign represents the leading edge and the negative sign the trailing edge, so the growth rate difference is positive.  The expression for growth rate difference is then given by 
\begin{equation}
\label{eqn:dg2}
\Delta\gamma =
\begin{cases}
2\sin(\alpha+\theta)[\cos(\alpha+\phi)-R\sin(\alpha+\phi)], &\quad \phi<\theta \text{ or } \theta+180\deg<\phi<\theta+360\deg,\\
-2\sin(\alpha+\theta)[\cos(\alpha+\phi)-R\sin(\alpha+\phi)], & \quad \phi<\theta-180\deg \text{ or } \theta<\phi<\theta+180\deg.
\end{cases}
\end{equation}
Equation \ref{eqn:dg2} can be expressed as one equation using the Heaviside step function $\Theta$.  First, let us define two functions of \(\theta\) and \(\phi\)
\begin{equation}
\label{eqn:H1}
H_1(\theta,\phi) \equiv \Theta((\theta-\pi)-\phi)+\Theta(\phi-\theta)-\Theta(\phi-(\theta+\pi)),
\end{equation}
\begin{equation}
\label{eqn:H2}
H_2(\theta,\phi) \equiv \Theta(\phi-(\theta-\pi))-\Theta(\phi-\theta)+\Theta(\phi-(\theta+\pi)).
\end{equation}
Equation \ref{eqn:H1} gives \(H_1(\theta,\phi)\), which returns 1 for \(\phi<\theta-180\deg\) and \(\theta<\phi<\theta+180\deg\), and 0 for all other values of \(\phi\).  Equation \ref{eqn:H2} gives \(H_2(\theta,\phi)\), which returns 1 for \(\phi<\theta\) and \(\theta+180\deg<\phi<\theta+360\deg\), and 0 for all other values of \(\phi\).  A single expression for $\Delta\gamma$ is constructed using these two function as
\begin{equation}
\Delta\gamma = 2 H_1(\theta,\phi)\sin(\alpha+\theta)[\cos(\alpha+\phi)-R\sin(\alpha+\phi)]-2 H_2(\theta,\phi)\sin(\alpha+\theta)[\cos(\alpha+\phi)-R\sin(\alpha+\phi)].
\end{equation}

This expression can be rewritten in a shorter form if we define \(\mathrm{H}(\theta,\phi)\equiv H_2(\theta,\phi) - H_1(\theta,\phi)\) 
\begin{equation}
\label{eqn:dg}
\Delta\gamma = -2H(\theta,\phi)\sin(\alpha+\theta)[\cos(\alpha+\phi)-R\sin(\alpha+\phi)].
\end{equation}
Equation \ref{eqn:dg} gives the difference in the GDI growth rate across the gradient reversal for any point along the reversal for any arbitrary orientation of the gradient reversal and convection direction. To find the average difference in growth rate, we integrate over \(\alpha\) across the radar FoV between \(-\alpha_0\) and \(+\alpha_0\) with respect to the boresite. In this integration, we assume that the conductivity ratio \(R\) is a constant for any combination of angles, which is a fair assumption for observations in the E region or in the F region. 
\begin{equation}
\label{eqn:dgavg}
\Delta\gamma_{\mathrm{avg}} = \frac{1}{2\alpha_0}\int_{-\alpha_0}^{\alpha_0}\Delta\gamma d\alpha = -\frac{H(\theta,\phi)}{\alpha_0}\int_{-\alpha_0}^{\alpha_0}\sin(\alpha+\theta)[\cos(\alpha+\phi)-R\sin(\alpha+\phi)]d\alpha
\end{equation}
The integral in equation \ref{eqn:dgavg} is easily evaluated and the final expression for the average difference in GDI growth rate across the gradient reversal becomes
\begin{equation}
\Delta\gamma_{\mathrm{avg}} = H(\theta,\phi)\left[\sin(\phi-\theta)-C\sin(\phi+\theta)+R\left(\cos(\phi-\theta)-C\cos(\phi+\theta)\right)\right], 
\end{equation}
where a new constant  has been defined: \(C\equiv\sin(2\alpha_0)/2\alpha_0\). For a typical SuperDARN radar, \(\alpha_0=26\deg\) and \(C\approx 0.87\).

%\begin{acknowledgements}
%This work was supported by NSF grants AGS-1248127 and PLR-1443504. The MSISE-90 model is available from the GSFC/SPDF OMNIWeb interface at http://omniweb.gsfc.nasa.gov/.
%\end{acknowledgements}

\bibliographystyle{uafthesis}
\bibliography{references}

%\newpage

%\addtocounter{figure}{-1}\renewcommand{\thefigure}{\arabic{figure}a}

%\end{document}

### 2017-01-20 thesis.sh

@@ -17,11 +17,9 @@ gs -q -dSAFER -dBATCH -dNOPAUSE -sDEVICE=pdfwrite -dDEVICEWIDTHPOINTS=510 -dDEVI
 pdflatex LLamarcheThesis.tex
 
 bibtex LLamarcheThesis
-
 bibtex introduction
-
 bibtex paper1
-
+bibtex paper2
 bibtex conclusion
 
 pdflatex LLamarcheThesis.tex

### 2017-01-20 conclusion.tex
Conlusion figure incorporated into text
Formatting updated
.svg figures incorporated into thesis correctly

@@
 
-\chapter{Conclusions and Future Work}
+\chapter{Conclusion}
 \label{sec:conclusion}
@@
 The requirements for confirming this hypotheis will be detailed further in Section \ref{sec:futurework}.
+
+\begin{figure}
+  \includegraphics[width=\textwidth]{conclusion.pdf}
+  \caption[Irregularity production factors]{Summary of the main conclusions in this thesis as to what factors most effect pasma irregularity production.}
+  \label{fig:conclusion}
+\end{figure}
 
@@
 By the radar equation, the power returned should be proportional to the product of the background density and fractional density, squared, Equation \ref{eqn:fractional_density}.
-However, there is no general consensus on what controls the fractional density in the \(F\) region, although it has been suggested that it is either constant or a function of convection speed, \(V_E\) \citep{Koustov1988,Haldoupis1990,Makarevich2014b}.
+However, there is no general consensus on what controls the fractional density in the \(F\) region, although it has been suggested that it is either constant or a function of convection speed, \(V_E\) \citep{Kustov1988,Haldoupis1990,Makarevich2014b}.
 In Chapter \ref{sec:paper3}, we additionally propose it may be a function of the instability growth rate, \(\gamma\).

### 2017-01-20 introduction.tex
.svg figures incorporated into thesis correctly

@@
 For instance, large-scale structures such as polar patches may very well have internal intermediate- and/or small-scale structuring.
-A turbulent cascade is often assumed to connect these different scales, but because various instruments or techniques usually only observe structuring on a particular scale, it is difficult to find direct evidence of this or how exactly it occurs \citep{Kinter1985,Tsunoda1985}.
+A turbulent cascade is often assumed to connect these different scales, but because various instruments or techniques usually only observe structuring on a particular scale, it is difficult to find direct evidence of this or how exactly it occurs \citep{Kintner1985,Tsunoda1985}.
 A description of some of the particular instruments and models that are used in this thesis follows.
@@
 Although the IRI model is very useful for providing background plasma density conditions where no instruments are available to make measurements, it is important to bear in mind that it is a model and therefor only represents the average conditions.
-Although it is fairly accurate at mid-latitudes \citep{Coisson2006,Bilitza2012}, it does not necessarily represent variations within the polar cap well \citep{Themens2014,Makarevich2015}.
+Although it is fairly accurate at mid-latitudes \citep{Coisson2006,Bilitza2012}, it does not necessarily represent variations within the polar cap well \citep{Themens2014,Makarevich2015b}.
 In particular, the height of the F-region peak tends to be underestimated in daytime and the bottomside thickness does not show realistic seasonal and diurnal variation \citep{Themens2014}.
@@
 Also operational in the \(E\) region but more dominant at higher altitudes is the gradient-drift instability (GDI) \citep{Simon1963,Hoh1963,Linson1970}.
-If a wave has a component along the magnetic filed line, the current-conductive instability (CCI) can be operational even if GDI is not \citep{Hoh1960,Ossakow1979,Chaturvedi1981}.
+If a wave has a component along the magnetic field line, the current-conductive instability (CCI) can be operational even if GDI is not \citep{Hoh1960,Ossakow1979,Chaturvedi1981}.
 Additionally instability mechanisms emerge if shears are considered, when \(\nabla E \neq 0\), such as the Kelvin-Helmholtz instability (KHI) \citep{Kintner1977,DAngelo1965}.
@@
 For the most part, the results of nonlinear numerical simulations very closely match expectations from linear GDI theory.
-The spatial power spectra are anisotropic with most of the power concentrated in the direction where linear GDI growth is most favorable \citep{Keskinen1981,Keskinen1982,Gondarenko2001,Gondarenko2004b}.
+The spatial power spectra are anisotropic with most of the power concentrated in the direction where linear GDI growth is most favorable \citep{Keskinen1981a,Keskinen1981b,Keskinen1982a,Gondarenko2001,Gondarenko2004b}.
 However, very strong shears or increasing the impact of ion-inertial factors can cause isotropy \citep{Gondarenko2001,Gondarenko2006}.

### 2017-01-20 paper1.tex
.svg figures incorporated into thesis correctly

@@
 
-Production of these irregularities is reasonably well understood on a microphysical level, with GDI promoting irregularity growth for certain orientations of the density gradient vector relative to the convection electric field \citep{Kesknien1982a,Kesknien1982b,Makarevich2014c}.
+Production of these irregularities is reasonably well understood on a microphysical level, with GDI promoting irregularity growth for certain orientations of the density gradient vector relative to the convection electric field \citep{Keskinen1982a,Keskinen1982b,Makarevich2014c}.
 The role of global factors, on the other hand, such as geomagnetic activity or solar wind conditions is much less investigated.

### 2017-01-20 paper2.tex
.svg figures incorporated into thesis correctly

@@
 In all cases, the convection direction is assumed to be perpendicular to the gradient reversal direction (\(\phi-\theta=90\deg\)).
-This is because large-scale structures, such as polar patches tend to drift perpendicular to their elongation direction \citep{Makarevich2004c,Makarevich2015b}.
+This is because large-scale structures, such as polar patches tend to drift perpendicular to their elongation direction \citep{Makarevitch2004c,Makarevich2015b}.
 This requirement was convenient for presenting different growth rate patterns, but will be relaxed later in this study.
@@
 Large conductivity in the E region can also short out F region irregularities \citep{Danskin2002,Kane2012,Lamarche2015}.
-An enhanced D region may act to attenuate backscatter signals, although this effect is not always obvious \citep{Dansin2002}.
-Nonlinear processes in the F region have been theoretically shown to alter the dependence on the propagation direction while still showing significant anisotropy similar to that predicted by the linear theory \citep{Keskinen81,Keskinen82a,Keskinen83}.
+An enhanced D region may act to attenuate backscatter signals, although this effect is not always obvious \citep{Danskin2002}.
+Nonlinear processes in the F region have been theoretically shown to alter the dependence on the propagation direction while still showing significant anisotropy similar to that predicted by the linear theory \citep{Keskinen1981b,Keskinen1982a,Keskinen1983}.
 The situation in the E region appears to be somewhat different, with some echoes also exhibiting strong anisotropy \citep[e.g.][]{Ierkic1980,Makarevich2008} and other echoes showing isotropic power variation \citep{Ierkic1980,Koustov2001} presumably because of nonlinear processes.

### 2017-01-20 thesis.sh

@@ -5,14 +5,12 @@ gs -q -dSAFER -dBATCH -dNOPAUSE -sDEVICE=pdfwrite -dDEVICEWIDTHPOINTS=530 -dDEVI
 gs -q -dSAFER -dBATCH -dNOPAUSE -sDEVICE=pdfwrite -dDEVICEWIDTHPOINTS=270 -dDEVICEHEIGHTPOINTS=520 -sOutputFile=/Users/ljlamarche/Desktop/UAFThesis/Figures/SuperDARNmap.pdf /Users/ljlamarche/Desktop/UAFThesis/Figures/SuperDARNmap.ps
 gs -q -dSAFER -dBATCH -dNOPAUSE -sDEVICE=pdfwrite -dDEVICEWIDTHPOINTS=510 -dDEVICEHEIGHTPOINTS=510 -sOutputFile=/Users/ljlamarche/Desktop/UAFThesis/Figures/ISRmap.pdf /Users/ljlamarche/Desktop/UAFThesis/Figures/ISRmap.ps
 
-# gs -q -dSAFER -dBATCH -dNOPAUSE -sDEVICE=pdfwrite -dDEVICEWIDTHPOINTS=600 -dDEVICEHEIGHTPOINTS=850 -sOutputFile=/Users/ljlamarche/Desktop/UAFThesis/Figures/month_density_occurance2010.pdf /Users/ljlamarche/Desktop/UAFThesis/Figures/month_density_occurance2010.eps
-# gs -q -dSAFER -dBATCH -dNOPAUSE -sDEVICE=pdfwrite -dDEVICEWIDTHPOINTS=600 -dDEVICEHEIGHTPOINTS=850 -sOutputFile=/Users/ljlamarche/Desktop/UAFThesis/Figures/year_month_analysis_line.pdf /Users/ljlamarche/Desktop/UAFThesis/Figures/year_month_analysis_line.eps
-# gs -q -dSAFER -dBATCH -dNOPAUSE -sDEVICE=pdfwrite -dDEVICEWIDTHPOINTS=600 -dDEVICEHEIGHTPOINTS=850 -sOutputFile=/Users/ljlamarche/Desktop/UAFThesis/Figures/year_month_analysis_color.pdf /Users/ljlamarche/Desktop/UAFThesis/Figures/year_month_analysis_color.eps
[…]

### 2017-01-24 abstract.tex
Corrections made to Introduction

@@
 
-Plasma irregularities in the polar \(F\) region ionosphere are investigated primarily using ground based radars within the Super Dual Auroral Radar Network (SuperDARN) and the factors that contribute to the production and observation of these irregularities are considered.
+Factors that contribute to the production and observation of plasma irregularities in the polar \(F\) region ionosphere are investigated.
+The primary focus of this thesis is decameter-scale irregularities observed with radars in the Super Dual Auroral Radar Network (SuperDARN).
+
 First, solar control on irregularity production is evaluated.
-A band of ranges where echoes are statistically most likely to be observed is identified.
-The location of this band varies diurnally and seasonally, but its motion relative to the radar can be explained by background density with solar illumination effecting radar beam propogation.
-The high observed occurrence of backscatter at night implies an additional source of nighttime plasma density that is not accounted for by traditional models.
+A band of ranges where echoes are statistically probable is identified.
+The location of this band within a radar's field of view varies diurnally and seasonally, but its motion can be explained through variations in background density with solar illumination effecting radar beam propagation.
+Backscatter occurrence peaks at the terminator.
+High backscatter occurrence at night implies an additional source of nighttime plasma density that is not accounted for by traditional models.
 An example is presented that shows frequent polar patches may be the source of this density.
-Backscatter occurrence peaks at the terminator.
-Nighttime occurence is enhanced when a positive IMF By component is present, probably due to changes in the convection pattern creating a more favorable situation for gradient-drift instability growth.
+Nighttime occurrence is enhanced when a positive IMF By component is present, probably due to changes in the convection pattern creating a more favorable situation for gradient-drift instability growth.
 
@@
 To test the accuracy of the predictions found from the model, irregularities surrounding a polar patch are analyzed in the context of linear gradient-drift instability theory.
-Backscatter power and occurrence from decameter scale irregularities are found using measurements from the SuperDARN radar at Rankin Inlet while background density gradients and electric fields are found from the north face of the Resolute Bay Incoherent Scatter Radar.
-Particular emphasis is placed on directional dependence and wether small-scale irregularities show any signs of anisotropy.
+Backscatter power and occurrence from decameter-scale irregularities are found using measurements from the SuperDARN radar at Rankin Inlet while background density gradients and electric fields are found from the north face of the Resolute Bay Incoherent Scatter Radar.
+Particular emphasis is placed on directional dependence and whether small-scale irregularities show any signs of anisotropy.
 While low powered echoes seem to follow the predictions of linear theory, higher power echoes quickly exhibit much more complicated, nonlinear behavior.

### 2017-01-24 introduction.tex
Corrections made to Introduction

@@
 
-The majority of mass in the universe exists as plasmas.
 Plasmas are fluid masses of charged particles, often formed when neutral atoms and molecules in a gas are ionized by external energy.
-Everyday examples of this is florecent lightbulbs, as well as flames.
-More exotic instances of plasma include nuclear fusion reactor experiments, most stars, nebula, and the interstellar medium.
+Instances of plasma include everything from florescent lightbulbs and flames to more exotic examples such as nuclear fusion reactor experiments, most stars, nebula, and the interstellar medium.
 In fact, about 99\% of matter in the known universe exists in the plasma state.
@@
 
-Plasmas can be either fully ionized, where most particles in the plasma are ionized, or partially ionized, where only some fraction are ionized and the rest are neutral.
-Considering the dynamics of plasmas therefore requires not only collisions between particles to be considered but also charge-charge interactions and external electric and magnetic fields.
+Ionospheric plasma is partially ionized, meaning only a fraction of the particles are ionized and the rest are neutral.
+Considering the dynamics of plasmas therefore requires not only collisions between particles to be considered but also charge-charge interactions and external and internal electric and magnetic fields.
 These dynamics can generate waves in the plasma, which cause density perturbations.
@@
 The Sun is composed of highly energized plasma that is gravitationally bound together.
-The surface of the Sun has sunspots, which are dark, relatively cool regions with intense magnetic fields, as well as coronal holes, which are low density areas characterized with a continuous outflow of plasma.
+The surface of the Sun is highly dynamic, populated with sunspots, which are dark, relatively cool regions with intense magnetic fields, as well as coronal holes, which are low density areas characterized with a continuous outflow of plasma.
 The intense magnetic fields associated with sunspots can break down releasing a large amount of energy and plasma, known as a solar flare.
@@
 This is shown in Figure \ref{fig:ste}.
-The solar wind, green, typically has a density around 10 cm\(^{-3}\) and is traveling away from the Sun at approximately 400 km/s at Earth's orbit.
+The solar wind, represented by the green arrows, typically has a density around 10 cm\(^{-3}\) and is traveling away from the Sun at approximately 400 km/s at Earth's orbit.
 However, like the Sun, the solar wind is highly dynamic and these parameters can change drastically over a period of minutes.
 CMEs, for instance, create a substantial increase in density and are often preceded by a shock, a region of hot, compressed plasma caused by the CME moving at supersonic speeds.
-In addition to the outflow of plasma, the solar wind also consists of the Sun's magnetic field extended into the outer limits of the solar system, called the interplanetary magnetic field (IMF), dark blue in Figure \ref{fig:ste}.
+In addition to the outflow of plasma, the solar wind also consists of the Sun's magnetic field extended into the outer limits of the solar system, called the interplanetary magnetic field (IMF), dark blue lines in Figure \ref{fig:ste}.
 Because the Sun has a short rotation period (about 33 days), the IMF gets twisted into a spiral, commonly called the Parker Spiral \citep{Parker1958}.
-When the IMF reaches the Earth, the magnetic field is typically oriented at a \(45^\circ\) inclination relative to the ecliptic, but this is highly variable depending on solar wind conditions.
+When the IMF reaches the Earth, the magnetic field is typically oriented at a \(45\deg\) inclination relative to the ecliptic, but this is highly variable depending on solar wind conditions.
 
@@
 	\caption[Solar-Terrestrial Environment]{Diagram of the solar-terrestrial environment.
-The solar wind contains both a steady stream of charged particles and the IMF, dark blue.
+The solar wind, green arrows, contains both a steady stream of charged particles and the IMF, dark blue.
 The red region represents the magnetosheath.
 The manetopause is shown by the black dashed line.
-Closed magnetic field lines from the earth are shown in grey while open field lines are shown in light blue.}
+Closed magnetic field lines from the earth are shown in grey while open field lines are shown in light blue.
+The ionosphere is shown in light green above the Earth's surface.}
 	\label{fig:ste}
@@
 The magnetopause, black dashed line, is the actual boundary between solar wind plasma in the magnetosheath and magnetospheric plasma.
+Magnetic field lines that are part of the closed magnetosphere are shown in grey in Figure \ref{fig:ste}.
 However, the magnetosphere is not fully closed.
-At the magnetopause, magnetic reconnection can occur between the IMF and the Earth's magnetic field creating open field lines.
-The last closed field line (closed field lines shown in grey) where reconnection tends to occur at the magnetopause is highlighted in yellow.
+At the magnetopause, magnetic reconnection can occur between the IMF and the Earth's magnetic field creating open field lines, light blue in Figure \ref{fig:ste}.
+This reconnection occurs on last closed field line intersecting the magnetopause (highlighted in yellow in Figure \ref{fig:ste}).
 Magnetic field lines originating from the Earth are then directly connected to the Sun through the solar wind, which consequently allows highly energized solar particles to enter the magnetosphere.
-Open field lines, light blue, tend to occur around the Earth's polar caps, making these regions particularly interesting to study due to very complex behavior and coupled interactions between the Earth and the Sun.
+Open field lines tend to occur around the Earth's polar caps, making these regions particularly interesting to study due to very complex behavior and coupled interactions between the Earth and the Sun.
 
-Closer to the Earth's surface, the magnetosphere also interacts with the Earth's ionosphere, a layer of partially ionized plasma in the upper atmosphere.
+Closer to the Earth's surface, the magnetosphere also interacts with the Earth's ionosphere, a layer of partially ionized plasma in the upper atmosphere, light green in Figure \ref{fig:ste}.
 The interaction between charged particles in the ionosphere with magnetic field lines couples the ionosphere, magnetosphere, and solar wind to create non-trivial dynamics.
@@
 Solar illumination and particle precipitation ionize neutrals, and the density of ions increase as altitude decreases because there are more neutral molecules available to ionize.
-However, below a certain point, the concentration of ions becomes large enough that recombination into neutral particles becomes a substantal factor, reducing the ion density.
+However, below a certain point, the concentration of ions becomes large enough that recombination into neutral particles becomes a substantial factor, reducing the ion density.
 The exact altitude where this density peak occurs is variable with time of day, location, season, and solar cycle.
@@
 Common ions at these altitudes are therefore mostly composed of oxygen and nitrogen, O\(^+\), NO\(^+\), O\(_2^+\), N\(_2^+\), but hydrogen, H\(^+\), can contribute as well, particularly at higher altitudes \citep{Kelley2009}.
-In the ionosphere, neutral densties are typically several orders of magnitude higher than ion densities, however even a small number of charged particles changes the behavior of the region dramatically.
+In the ionosphere, neutral densities are typically several orders of magnitude higher than ion densities, however even a small number of charged particles changes the behavior of the region dramatically.
 Because there are so many different ion species, plasma density is usually discussed simply in terms of electron density.
@@
 Density irregularities in the polar cap can occur on scales ranging from thousands of kilometers to less than a centimeter \citep{Tsunoda1988}, and a summary of some of these irregularities is presented in Figure \ref{fig:polarirreg}.
+
 Large-scale plasma irregularities are typically considered to be any density structure on the scale of greater than 30 km \citep{Kelley2009}.
 Some common examples are polar patches, polar holes, and sun-aligned arcs.
-Polar patches are large density enhancements (usually at least twice the background plasma density) that travel across the polar cap with the background convection, Figure \ref{fig:polarirreg} \citep{Weber1984,Valladares1994}.
+Polar patches are large density enhancements (usually at least twice the background plasma density) that travel across the polar cap with the background convection, Figure \ref{fig:polarirreg}a \citep{Weber1984,Valladares1994}.
 Polar holes are plasma density depletions that typically occur in the \(F\) region slightly poleward of the auroral oval \citep{Benson2001}.
@@
 Large scale plasma structures and the density gradients that surround them often serve as a platform for smaller-scale structures to form as well.
-Polar patches, for instance, are known to form large finger-like structures along their trailing edges, Figure \ref{fig:polarirreg} \citep{Gondarenko2004b,Hosokawa2016}.
+Polar patches, for instance, are known to form large finger-like structures along their trailing edges, Figure \ref{fig:polarirreg}b \citep{Gondarenko2004b,Hosokawa2016}.
 
 \begin{figure}
-	\includegraphics[width=0.8\textwidth]{irregularity.pdf}
+	\includegraphics[width=0.5\textwidth]{irregularity.pdf}
 	\caption[Plasma irregularities in the polar ionosphere]{Schematic of different scale size plasma density irregularities that occur in the polar cap.
@@
 This can cause communication blackouts or introduce large errors into GPS/GNSS (Global Positioning System/Global Navigation Satellite System) calculations, affecting navigation.
-Because communication and navigation technology is implemented in both every day civilian life as well as government and military operations, the effects of this kind of signal scintillation can be far reaching.
 
@@
 Although small-scale irregularities are not responsible for scintillation that directly impact GPS/GNSS performance, they can be the easiest to study because of large data sets collected from high frequency ionospheric radars globally for several decades (discussed further in Section \ref{sec:superdarn}).
-Because electrons move far more easily along magnetic field lines than across them, density perturbations tend to be aligned with the magnetic field and are often known as field-aligned irregularities (FAI), Figure \ref{fig:polarirreg}.
+Because electrons move far more easily along magnetic field lines than across them, density perturbations tend to be aligned with the magnetic field and are often known as field-aligned irregularities (FAI), Figure \ref{fig:polarirreg}d.
 The irregularity classifications presented here are rough categories only and a high degree of coupling between different scales occurs, which makes plasma structures of all scales (large, intermediate, and small) important for plasma structuring processes.
@@
 In this thesis, we make extensive use of Coherent Scatter Radars (CSR), which have been used to study the ionosphere for the last half century.
-The radar transmits a radio wave, which can scatter off plasma structures in the ionosphere, and then be detected by the recieving antennas.
+The radar transmits a radio wave, which can scatter off plasma structures in the ionosphere, and then be detected by the receiving antennas.
 The time difference between when the signal was transmitted and when it was detected and the power and phase of the returned signal can then be used to determine characteristics of plasma irregularities.
@@
 There are three measurements typically made by CSR systems: backscatter power, line-of-sight (LoS) velocity, and spectral width.
-Backscatter power is the power of the returned signal that the radar's recievers detect.
+Backscatter power is the power of the returned signal that the radar's receivers detect.
 Typically, a threshold is selected for the minimum power acceptable for a return to be considered an actual signal instead of background noise.
@@
 This allows the distance between the radar and the target to be found (based on the time difference between when the pulse is transmitted and received).
-The size of the range gate is determined by the length of a radar pulse, Equation \ref{eqn:range_gate}.\
+The size of the range gate is determined by the length of a radar pulse, Equation \ref{eqn:range_gate}.
 \begin{equation}
@@
 When the radar transmits continuously, backscatter will be received, but because it is impossible to know when that signal was transmitted, the difference in time between transmission and receiving cannot be used to calculate how far away the backscatter volume is.
-Conversly, if the radar only transmits a pulse of a certain length of a certain length and then ``listens'' for its return, the time between transmission and return is known and the distance the pulse traveled can be calculated and therefore the location of the backscattering volume \citep{Farley1972,Greenwald1983}.
+Conversely, if the radar only transmits a pulse of a certain length of a certain length and then ``listens'' for its return, the time between transmission and return is known and the distance the pulse traveled can be calculated and therefore the location of the backscattering volume \citep{Farley1972,Greenwald1983}.
 However, using this simple single-pulse method another pulse cannot be transmitted until the first pulse returns, limiting the temporal resolution that can be achieved.
@@
 A short time between pulse returns is necessary to allow Doppler velocities to be calculated, particularly at long ranges.
-The maximum Doppler shifted frequency that can be observed by a radar is the Nyquist frequency, \(f_n\), which is equivilent to half the sampling frequency, \(f_s\), Equation \ref{eqn:nyquist}
+The maximum Doppler shifted frequency that can be observed by a radar is the Nyquist frequency, \(f_n\), which is equivalent to half the sampling frequency, \(f_s\), Equation \ref{eqn:nyquist}
 \begin{equation}	
@@
 This severely limits the Doppler shift that can be measured, which places an upper limit on the LoS Doppler velocity that can be calculated with this technique.
-For an ionospheric radar observing a structure 2000 km away, it would take about 13 ms for a signal traveling at the speed of light to travel to the the structure and back to the radar, wich corresponds to a sampling frequency of \(f_s = 75\) Hz.
+For an ionospheric radar observing a structure 2000 km away, it would take about 13 ms for a signal traveling at the speed of light to travel to the the structure and back to the radar, which corresponds to a sampling frequency of \(f_s = 75\) Hz.
 The Nyquist frequency is then \(f_n = 37.5\) Hz, which corresponds to a maximum measurable Doppler velocity of \(\sim550\) m/s.
@@
 
-Instead of waiting for each individual signal to be recieved before transmitting the next, ionospheric radars typically employ a multipulse mode.
+Instead of waiting for each individual signal to be received before transmitting the next, ionospheric radars typically employ a multipulse mode.
 In a multipulse mode, the radar transmits a series of pules with different time intervals or lags between them.
@@
 The first observations of the ionosphere by CSRs were made in the 1930s \citep{Eckersley1937,Harang1938}.
-Over the next several decades, numerous used ground-based radars to advance our understanding of the ionosphere and radio aurora \citep{Hultqvist1964,Leadabrand1965,LangeHesse1967,Unwin1972,Sahr1996}.
+Over the next several decades, numerous used ground-based radars to advance our understanding of the ionosphere and radio aurora \citep{Hultqvist1964,Leadabrand1965,Unwin1972,Sahr1996}.
 In the 1970s and 1980s, the Scandanavian Twin Auroral Radar Experiment (STARE) produced the first continuous large data-set, similar to how modern CSR systems are run.
 STARE consisted of two very high frequency (VHF) radars with overlapping FoVs that were designed to measure FAIs in the \(E\) region of the ionosphere \citep{Greenwald1997}.
-The advantage of two radars with overlaping FoVs was the ability to observe the same structures from two different directions.
-Because each radar can only measure the LoS velocity of a structure, two simultanious observations from different directions allows two different velocity vecotor compontents to be found, and hense the total velocity vector can be calculated.
+The advantage of two radars with overlapping FoVs was the ability to observe the same structures from two different directions.
+Because each radar can only measure the LoS velocity of a structure, two simultaneous observations from different directions allows two different velocity vector components to be found, and hence the total velocity vector can be calculated.
 This technique is still commonly used with CSR system.
@@
 \label{sec:superdarn}
-The primary instruments used within this body of work are radars within the Super Dual Auroral Radar Network (SuperDARN), a global network of high frequncy (HF) CSRs that was designed to measure small-scale plasma structures in the ionosphere and map the global plasma convection.
+The primary instruments used within this body of work are radars within the Super Dual Auroral Radar Network (SuperDARN), a global network of high frequency (HF) CSRs that was designed to measure small-scale plasma structures in the ionosphere and map the global plasma convection.
 The network currently consists of about 35 operational radars between the northern and southern hemispheres distributed at mid-, high-, and polar latitudes, Figure \ref{fig:superdarnmap}.
@@
 
-It is necessary to use HF radars to study ionsospheric structures in the polar cap because the magnetic field is close to vertical and in order for the radar beam to meet the perpendicularity condition to observe FAIs, the beam must be refracted through a dense ionosphere.
+It is necessary to use HF radars to study ionospheric structures in the polar cap because the magnetic field is close to vertical and in order for the radar beam to meet the perpendicularity condition to observe FAIs, the beam must be refracted through a dense ionosphere.
 VHF radar beams are not refracted enough in the polar cap and the beam will pass through the ionosphere without FAIs producing backscatter.
@@
 Originally this was accomplished by considering two radars with overlapping FoVs.
-If both recieved backscatter from the same scattering volume, two LoS velocities could be found for that scattering volume, which could be combined to find a 2D velocity vector \citep{Ruohoniemi1989}.
+If both received backscatter from the same scattering volume, two LoS velocities could be found for that scattering volume, which could be combined to find a 2D velocity vector \citep{Ruohoniemi1989}.
 The modern method for creating convection maps involves taking all data recorded by all radars in a particular hemisphere and finding the best fit to a 2D ionospheric electrostatic potential using Legendre functions \citep{Ruohoniemi1998}.
@@
 The first ISR was built in Arecibo, Puerto Rico in the 1950s \citep{Gordon1958}.
-Since then, at least 8 other ISRs have been deployed around the world at equitoral, mid-, and high-latitudes.
+Since then, at least 8 other ISRs have been deployed around the world at equatorial, mid-, and high-latitudes.
 A map of the ISR systems currently operational in the northern polar cap is shown in Figure \ref{fig:isrmap}.
 The most recent advancement has been the development of Advanced Modular Incoherent Scatter Radars (AMISR), of which there are currently three in use, one at the Poker Flat Rocket Range (PFISR), just north of Fairbanks, Alaska and two in Resolute Bay, Canada (RISR-N and RISR-C).
-The advantage of these new systems is that they are electronical steerable, unlike older ISRs which consisted of a large dish that had to be physically moved to change the beam direction.
+The advantage of these new systems is that they are electronically steerable, unlike older ISRs which consisted of a large dish that had to be physically moved to change the beam direction.
 This allows beams to be transmitted in many different directions at a very high time cadence, giving the radar the capability to make measurements in multiple directions almost simultaneously \citep{Nicolls2007a,Nicolls2007b,Bahcivan2010}.
@@
 Much work has been done over the past several decades to study the complicated picture of plasma structuring in the polar cap, including theoretical analysis of plasma dispersion relations, simulations of plasma dynamics, and observations of plasma structuring using a variety of instruments.
-Here, a survey will be given of some of these advancements and the current state of understanding plasma structring on a range of scales.
+Here, a survey will be given of some of these advancements and the current state of understanding plasma structuring on a range of scales.
 
@@
 \begin{figure}
-	\includegraphics[width=0.75\textwidth,angle=270]{magnetosphere.pdf}
+	\includegraphics[width=0.75\textwidth,angle=0]{magnetosphere.pdf}
 	\caption[Magnetic reconnection footprint in the ionosphere]{Diagram of the interactions between the IMF, magnetosphere, and polar cap ionosphere.
@@
 
-During reconnection events, the polar cap boundary moves equatorward as the open magnetic flux into the polar cap increases, Figure \ref{fig:magnetosphere} \citep{Cowley1991,Lockwood1992a}.
+During reconnection events, the polar cap boundary moves equatorwards as the open magnetic flux into the polar cap increases, Figure \ref{fig:magnetosphere} \citep{Cowley1991,Lockwood1992a}.
 In addition, the potential along the X-line also gets mapped to the merging gap, assuming there is minimal drop in potential along the magnetic field lines \citep{Lockwood1992b}.
@@
 As it moves, it pulls the blob of dense plasma within the polar cap with it, separating it from the reservoir on the dayside \citep{Lockwood1992b}.
-Before the convection flow stops completely, Figure \ref{fig:patch_formation}e, the blob of dense plasma ``pinches off'' from the dayside plasma to become an independant patch \citep{Lockwood1992b}.
+Before the convection flow stops completely, Figure \ref{fig:patch_formation}e, the blob of dense plasma ``pinches off'' from the dayside plasma to become an independent patch \citep{Lockwood1992b}.
 The exact mechanism by which this happens is not well understood, but it is thought to be related to small variations in the IMF By component shifting the convection pattern slightly \citep{Cowley1980,Lockwood1992b}.
@@
 A series of reconnection bursts can create a line of patches all propagating across the polar cap.
-Reconnection bursts typically last about 2 minutes with anywhere between 7 and 25 minutes beween bursts \citep{Foster1984,Etemadi1988,Lockwood1992b}.
+Reconnection bursts typically last about 2 minutes with anywhere between 7 and 25 minutes between bursts \citep{Foster1984,Etemadi1988,Lockwood1992b}.
 However, observations have shown convection flow to be close to continuous for much longer periods.
@@
 \begin{figure}
-	\includegraphics[width=0.75\textwidth,angle=270]{patch_formation.pdf}
+	\includegraphics[width=0.75\textwidth,angle=0]{patch_formation.pdf}
 	\caption[Polar patch formation]{Polar patch formation through transient magnetopause reconnection.
@@
 In the polar F-region ionosphere, GDI is typically considered the dominant structuring process \citep{Weber1984,Cerisier1985,Basu1988,Tsunoda1988}.
-GDI is operational when high-density purturbations in the plasma move to regions of lower background density and low-density purturbations in the plasma move to regions of higher background density, such that the wave amplitude grows relative to background conditions.
+GDI is operational when high-density perturbations in the plasma move to regions of lower background density and low-density perturbations in the plasma move to regions of higher background density, such that the wave amplitude grows relative to background conditions.
 As plasma drifts and carries irregularities with it, electrons and ions have slightly different drifts and create a perturbed electric field.
 This perturbed electric field combined with the background magnetic field causes the density perturbations to \(\vec{E}\times\vec{B}\) drift.
-If the drift causes high and low density perturbations to move as desribed above, the instability is operational and the perturbations grow.
+If the drift causes high and low density perturbations to move as described above, the instability is operational and the perturbations grow.
 If instead high density perturbations drift into regions of higher background density and low density perturbations drift into regions of lower background density, the wave is damped.
@@
 Polar patches were first observed in the 1960s \citep{Hill1963}, but much of the work of characterizing them using optical and radio techniques was not accomplished until two decades later \citep{Weber1981,Weber1984,Weber1986,Buchau1983,Buchau1985}.
-Polar patches are density enhancements of as much as \(10^6\) cm\(^{-3}\) that occur primarially in the polar \(F\) region \citep{Buchau1983}.
-There were initially catagorized as extending 800--1000 km horizontally \citep{Weber1984}, but more recent studies have identified patches as large as 1500 km \citep{Hosokawa2014}.
+Polar patches are density enhancements of as much as \(10^6\) cm\(^{-3}\) that occur primarily in the polar \(F\) region \citep{Buchau1983}.
+There were initially categorized as extending 800--1000 km horizontally \citep{Weber1984}, but more recent studies have identified patches as large as 1500 km \citep{Hosokawa2014}.
 Patches tend to drift antisunwards at speeds of 500--1000 m/s, following background convection \citep{Buchau1983,Weber1984}.
@@
 Additionally, there is a bias towards patches occurring when there is a negative IMF \(B_z\) component \citep{Buchau1983,Rodger1996}.
-The frequency of patch occurrence is greatest when the cusp is slightly dayward of the terminator.
-This creates the situation of an unsunlit polar cap, so background plasma density is low and blobs of enhanced density are easily visible \citep{Coley1998}.
+The frequency of patch occurrence is greatest when the cusp is slightly daywards of the terminator.
+This creates the situation of a dark polar cap, so background plasma density is low and blobs of enhanced density are easily visible \citep{Coley1998}.
 In addition, the cusp then provides a gateway of highly ionized sunlit plasma into the dark polar cap.
@@
 All five signatures were observed as expected, strengthening the idea that transient magnetopause reconnection is responsible for patch formation.
-Furthermore, \citet{Carlson2006} identified and tracked a series of patches directly from the subauroral plasma reservoir and show that the boundary moves equatorward before relaxing poleward, one of the main predictions of the transient magnetopause reconnection mechanism \citep{Lockwood1992b}.
+Furthermore, \citet{Carlson2006} identified and tracked a series of patches directly from the subauroral plasma reservoir and show that the boundary moves equatorwards before relaxing poleward, one of the main predictions of the transient magnetopause reconnection mechanism \citep{Lockwood1992b}.
 A review of historical observations of polar patches has been provided by \citet{Crowley1996}.
@@
 The factors that affect irregularity observation can roughly be divided into two groups, irregularity production and radar propagation.
-Irregularity production factors are related to the existance of plasma irregularities in the ionosphere while radar propagation factors are related to the abililty to observe irregularities with a ground based radar, Section \ref{sec:superdarn}.
+Irregularity production factors are related to the existence of plasma irregularities in the ionosphere while radar propagation factors are related to the ability to observe irregularities with a ground based radar, Section \ref{sec:superdarn}.
 Within the polar caps, plasma irregularity production is much higher than at lower altitudes such that when background plasma conditions are favorable for radar propagation, FAIs are observed nearly continuously \citep{Bristow2011}.
@@
 There is an increase in echo occurrence at solar maximum, particularly in the midnight sector \citep{Milan1997,Koustov2004}.
-Seasonally, high latitude radars observe a peak in echo occurrence durring equinox months \citep{Koustov2004}.
-Additionally, backscatter tends to occure at closer ranges in summer at midnight, but moves further away from the radar in other seasons at midday \citep{Milan1997}.
-In general, echo occurrence in high durring the day and low at night, but nighttime occurrence is directly proportional to the \(F\) peak density and daytime occurrence can be significantly reduced by a dense and conductive \(E\) layer.
+Seasonally, high latitude radars observe a peak in echo occurrence during equinox months \citep{Koustov2004}.
+Additionally, backscatter tends to occur at closer ranges in summer at midnight, but moves further away from the radar in other seasons at midday \citep{Milan1997}.
+In general, echo occurrence in high during the day and low at night, but nighttime occurrence is directly proportional to the \(F\) peak density and daytime occurrence can be significantly reduced by a dense and conductive \(E\) layer.
 \citep{Koustov2004,Kane2012,Vickrey1982}.
@@
 This conductive layer can provide a conduit by which the potential differences associated with FAIs in the \(F\) region can be shorted out, effectively reducing \(F\)-region plasma irregularities \citep{Vickrey1982}.
-Both irreguarity production and radar propagation factors probably contribute to these regular variations of echo occurrence, because while a dense, daytime ionosphere creates favorable radar propigation conditions, it can also contribute to \(E\)-region shorting or smoothing of gradients which surpress irregularity production \citep{Koustov2004}.
+Both irregularity production and radar propagation factors probably contribute to these regular variations of echo occurrence, because while a dense, daytime ionosphere creates favorable radar propagation conditions, it can also contribute to \(E\)-region shorting or smoothing of gradients which surpress irregularity production \citep{Koustov2004}.
 
-Irregularity production is impacted primarially by the presence of strong density gradients and electric fields in the ionosphere \citep{Koustov2004}.
-Properly oriented density gradients and electric fields are a nessesary condition for GDI to be operational, Section \ref{sec:lit_instbilities}, so the stronger these are, the more likely the irregularity production rate is high.
+Irregularity production is impacted primarily by the presence of strong density gradients and electric fields in the ionosphere \citep{Koustov2004}.
+Properly oriented density gradients and electric fields are a necessary condition for GDI to be operational, Section \ref{sec:lit_instbilities}, so the stronger these are, the more likely the irregularity production rate is high.
 Both backscatter power and occurrence increase with increasing electric field strength \citep{Ballatore2001,Danskin2002,Makarevich2014b}.
@@
 
-The primary radar propogation factor to consider is the background plasma density and how much the radar beam is refracted.
+The primary radar propagation factor to consider is the background plasma density and how much the radar beam is refracted.
 If the background density is too low, the beam is under refracted and is never perpendicular to the magnetic field, however if the density is too high, the beam is over refracted, such that observation of small-scale FAIs is most favorable if the background density is between \(1\times 10^{11}\) m\(^{-3}\) and \(4\times 10^{11}\) m\(^{-3}\) \citep{Danskin2002,Makarevich2014b}.
 If the \(D\) region is too dense, this can also in theory attenuate the signal and reduce the rate of returned backscatter, although observations have shown this is not a major factor in echo occurrence \citep{Danskin2002}.
-Finally, the radar equation predicts that the returned backscatter should be proportional to the perturbation density, \(\delta n\), squared, or, equivilently, the product of the background density and the fractional density, \(\delta n/n\), squared, Equation \ref{eqn:fractional_density}
+Finally, the radar equation predicts that the returned backscatter should be proportional to the perturbation density, \(\delta n\), squared, or, equivalently, the product of the background density and the fractional density, \(\delta n/n\), squared, Equation \ref{eqn:fractional_density}
 \begin{equation}
@@
 The same results have been submitted for publication \citep{Lamarche2017}.
-Chapter \ref{sec:conclusion} presents a summary of the most significant results of this body of work, their implications for interpretation of radar observations of plasma structures in the polar cap and suggestions for future research involving coherent and incoherent scatter radar facilities in the polar regiont.
+Chapter \ref{sec:conclusion} presents a summary of the most significant results of this body of work, their implications for interpretation of radar observations of plasma structures in the polar cap and suggestions for future research involving coherent and incoherent scatter radar facilities in the polar region.
 

### 2017-01-25 conclusion.tex
Revisions made to Conclusion

@@
 Plasma irregularities in the \(F\)-region ionosphere have been observed for decades.
-However, despite extensive data sets collected over these years, supprisingly little is known about these irregularities, especially at small scales.
+However, despite extensive data sets collected over these years, surprisingly little is known about these irregularities, especially at small scales.
 This is in part due to the complex and turbulent nature of the ionosphere, particularly in the polar regions, but also due to some of the inherent limitations in traditional observation techniques which can make it very challenging to identify exactly what processes are operational.
@@
 
+Radar backscatter from ionospheric irregularities is most likely to be observed at particular slant ranges in the radar FoV, Figure \ref{fig:conclusion}.
+The location of this window of high echo occurrence slant ranges varies diurnally and seasonally, primarily due to changing background plasma density in the ionosphere.
+The background density controls how far a radar beam must travel to be perpendicular to the magnetic field through refraction and observe FAIs.
+This can create the illusion that small-scale plasma irregularities are more common in certain physical locations while in reality, they may exist over a much wider area but it is easier to observe them in particular locations.
+
 In the polar caps, backscatter is often observed at nighttime even when background plasma density is low and propagation conditions are traditionally thought to be too poor for FAIs to be observed with HF radar.
 This implies an additional source of nighttime plasma density that is not easily accounted for by models.
-Frequent polar patches that propagate from the dayside to the nightside could be the source of this additional density.
-The requirements for confirming this hypotheis will be detailed further in Section \ref{sec:futurework}.
+Frequent polar patches that propagate from the dayside to the nightside could be the source of this additional density, Figure \ref{fig:conclusion}.
+The requirements for confirming this hypothesis will be detailed further in Section \ref{sec:futurework}.
 
@@
   \includegraphics[width=\textwidth]{conclusion.pdf}
-  \caption[Irregularity production factors]{Summary of the main conclusions in this thesis as to what factors most effect pasma irregularity production.}
+  \caption[Irregularity production factors]{Summary of the main conclusions in this thesis as to what factors most effect small-scale plasma irregularity production.
+Polar patches (orange) drift antisunwards across the polar cap following the background plasma convection (red contours).
+Nighttime irregularity observation is enhanced with a positive IMF By component.
+Production is maximized near the terminator.
+Small-scale irregularity production surrounding a large-scale structure is found to exhibit both asymmetry and anisotropy.
+Radars are most likely to observe this asymmetry if the patch is elongated along the radar boresight.}
   \label{fig:conclusion}
@@
 
-Nighttime observations of FAIs is also higher with a positive IMF By component.
+Nighttime observations of FAIs is also higher with a positive IMF By component, Figure \ref{fig:conclusion}.
 This could be due to the control that the By component exerts over ionospheric convection patterns.
@@
 This also creates the asymmetry between the different edges of the patch because as the gradient direction changes, the irregularity growth rate changes.
-Additionally, linear theory predicts that irregularity growth is anisotropic, so the orientation of the patch relative to the radar can also impact the observed irregularities.
+Additionally, linear theory predicts that irregularity growth is anisotropic, so the orientation of the patch relative to the radar can also impact the observed irregularities because the rate at which irregularities develop changes in different directions.
 In the \(F\) region, the GDI growth rate is highest for wavevectors perpendicular to the gradient, \(\vec{G}\).
@@
 As discussed above, if small-scale irregularity growth is directionally dependent, this is additionally the optimal orientation from which to observe the plasma structures, leading to increased echo occurrence.
+
+The production and characteristics of small-scale plasma irregularities and the factors that control them are complex.
+Radar studies must differentiate between production and propagation factors, determining if in fact irregularity production is more favorable or if background conditions just make observations of irregularities more likely.
+This thesis helps advance our understanding by identifying a number of situations where small-scale irregularities are observed and establishing plausible reasons why either production is favorable according to plasma theory or radar observation is likely.
 

### 2017-01-25 introduction.tex
Revisions made to Conclusion

@@
 	\includegraphics[width=0.5\textwidth]{irregularity.pdf}
-	\caption[Plasma irregularities in the polar ionosphere]{Schematic of different scale size plasma density irregularities that occur in the polar cap.
-Polar patches, orange, move with the background ionospheric convection, indicated by green contours.
-Smaller-scale structures often move develop along the edges of polar patches where strong density gradients exist.
-Field-aligned irregularities, FAIs, are density perturbations that form along the magnetic field lines, pink lines.}
+	\caption[Plasma irregularities in the polar ionosphere]{Plasma density irregularities occur in the polar cap at different scale sizes.
+Polar patches, orange, move with the background ionospheric convection, indicated by red contours (a).
+Smaller-scale structures often move develop along the edges of polar patches where strong density gradients exist (b).
+Intermediate scale irregularities are primarily responsible for disrupting radio signals traveling through the ionosphere (c).
+Field-aligned irregularities, FAIs, are density perturbations that form along the magnetic field lines, light blue (d).}
 	\label{fig:polarirreg}

### 2017-01-27 paper1.tex
Corrections made to Introduction
Paper1 fixed to be published version

@@
-%\titlerunninghead{Solar Control of Polar Irregularities}
-
-%\authoraddr{L. J. Lamarche and R. A. Makarevich, Geophysical Institute, University of Alaska Fairbanks, 903 Koyukuk Drive, PO Box 757320, Fairbanks, AK, 99775-7320, USA.
-%(rmakarevich@alaska.edu)}
-
-%\linenumbers*[1]
-%\begin{document}
+% Chapter 2: paper1
 
 \chapter{Solar control of \(F\) region radar backscatter: Further insights from observations in the southern polar cap}
-
-%\authors{Leslie J. Lamarche\altaffilmark{1} and Roman A. Makarevich \altaffilmark{1}}
-%\vspace{-1.5cm}
-
-%\altaffiltext{1}{Geophysical Institute and Department of Physics, University of Alaska Fairbanks, Fairbanks, AK, USA.}
 
@@
 
-%\begin{article}
 \section{Introduction}
@@
 Observations of decameter-scale irregularities with SuperDARN have an advantage of significant datasets accumulated under a wide variety of observational conditions.
-SuperDARN utilizes radio refraction in the High Frequency (HF) band to observe \(F\) region echoes or backscatter from decameter-scale irregularities when radar beam achieves orthogonality with local magnetic field (orthogonality condition).
+SuperDARN utilizes radio refraction in the High-Frequency (HF) band to observe \(F\) region echoes or backscatter from decameter-scale irregularities when radar beam achieves orthogonality with local magnetic field (orthogonality condition).
 Because of importance of propagation conditions for HF radar observations, the echo occurrence rates observed by SuperDARN are taken as a lower limit of the irregularity occurrence rates \citep{Chisham2007}.
@@
 However, it has been challenging to determine how occurrence explicitly depends on SZA for all seasonal and time sectors.
-The nighttime occurrence taken separately was demonstrated to be closely related to SZA-controlled \(F\) region density  variations \citep{Kane2012}, but no similarly-clear relationship has been demonstrated for datasets including daytime points.
+The nighttime occurrence, taken separately, was demonstrated to be closely related to SZA-controlled \(F\) region density  variations \citep{Kane2012}, but no similarly clear relationship has been demonstrated for datasets including daytime points.
+One possible reason is the competing effects of the electric filed, particularly in the noon sector near the cusp region \citep{Ghezelbash2014a,Ghezelbash2014b}.
 
 One difficulty is the need to consider propagation effects of plasma density.
-These arise because of \(F\)-region density controlling the amount of refraction and either under- or over-refraction reducing echo occurrence when orthogonality condition is not met \citep{Milan1997,Danskin2002}.
-This is also related to how echo occurrence is generally considered a lower limit on irregularities (described above).
+These arise because of \(F\)-region density controlling the amount of refraction and either underrefraction or overrefraction reducing echo occurrence when orthogonality condition is not met \citep{Milan1997,Danskin2002}.
+This is also related to how echo occurrence is generally considered a lower limit on irregularities, as described above.
 There is some ambiguity about whether changing irregularity production or detection by the radar is most responsible for any observed trends.
@@
 This is because range variations are expected to be almost entirely due to propagation effects and the differences in responses are therefore mostly likely due to changing production.
-In order to claim production as the dominant source of occurrence variability, occurrence analysis must explicitly considers the echo location.
+In order to claim production as the dominant source of occurrence variability, occurrence analysis must explicitly consider echo location.
 In the current study this approach is employed to develop a comprehensive model of occurrence location, quantify echo location trends, and explicitly consider these trends when developing occurrence estimates.
@@
 
-The primary instrument used in this study is the SuperDARN McMurdo radar located at McMurdo Station, Antarctica (77.88\(\deg\) S, 166.73\(\deg\) E, geographic) \citep{Bristow2011}.
+The primary instrument used in this study is the SuperDARN McMurdo radar located at McMurdo Station, Antarctica (77.88\(\deg\) S, 166.73\(\deg\) E, geographic) \citep[Figure 1]{Bristow2011}.
 Using a standard SuperDARN 3-letter radar code, this radar is referred to as MCM.
 SuperDARN is an array of coherent HF radars used to study processes in the Earth's ionosphere and magnetosphere.
-Radars are located at mid and high latitudes in both the Northern and Southern Hemispheres.
+Radars are located at middle and high latitudes in both the Northern and Southern Hemispheres.
 Radars measure a 17-lag auto-correlation function (ACF) from which estimates of the Doppler velocity, power, and spectral width of ionospheric echoes in 75--100 range bins for each of the 16--24 radar beams are obtained.
@@
 
-The magnetic latitude (MLAT) of MCM is 80.0\(\deg\) S, based on altitude-adjusted corrected geomagnetic (AACGM) coordinate system with 2010 coefficients \citep[see recent paper by][]{Shepherd2014}.
+The magnetic latitude (MLAT) of MCM is 80.0\(\deg\) S, based on the altitude-adjusted corrected geomagnetic (AACGM) coordinate system with 2010 coefficients (see recent paper by \citet{Shepherd2014}).
 The MCM FoV covers the southern polar cap and beam 14 of MCM is most conjugate with beam 1 of the SuperDARN Rankin Inlet (RKN) radar that overlooks the northern polar cap.
@@
 Standard SuperDARN criteria were used to eliminate ground scatter (\(\left| V-\Delta V \right| <30\) m/s and \(\left| W-\Delta W \right| <35\) m/s), where \(V\) is the Doppler velocity, \(W\) is the spectral width, and \(\Delta V\) and \(\Delta W\) represent uncertainties in these quantities.
-In addition, measurements with SNR \(<\) 3 dB or \(W>\) 500 m/s were eliminated to reduce contribution from interference and other spurious echoes.
+In addition, measurements with signal-to-noise ratio \(<\) 3 dB or \(W>\) 500 m/s were eliminated to reduce contribution from interference and other spurious echoes.
 In addition to standard operation modes, each SuperDARN radar has a set time periods where the radar is run in special modes.
@@
 Any periods with IMF information missing were excluded from the analysis.
+In addition, the Polar Cap South (PCS) index data derived from magnetometer measurements at Vostock Antarctic station have been used \citep{Troshichev1988}.
+The PC index is a standard index used to quantify responses of various polar cap phenomena to geomagnetic disturbance \citep[e.g.][]{Fiori2009}.
 
 The model electron densities were also used in the current study.
-The electron densities were found from the International Reference Ionosphere (IRI) 2007 model \citep{Bilitza2008}.
-The IRI model is an empirical model of the ionosphere based on time of day, date, geographic location, altitude, and solar conditions, with electron density near the \(E\) and \(F\) region peaks being of particular interest for this study.
-Using ionosonde terminology, these are also referred to below as model NmE and NmF2.
+The electron densities were found from the International Reference Ionosphere (IRI) 2012 model \citep{Bilitza2014}.
+The IRI model is an empirical model of the ionosphere based on time of day, date, geographic location, altitude, and solar conditions, with the \(E\) region and \(F\) region peak electron densities (\(N_m E\) and \(N_m F_2\)) being of particular interest for this study.
+The IRI model uses the International Union of Radio Science (URSI) or Consultative Committee on International Radio (CCIR) coefficient maps of the critical frequencies ot produce \(N_m E\) and \(N_m F_2\), with both the URSI and CCIR options implemented in the current study.
+In addition, IRI 2012 has three options for the bottomside thickness parameter B0, with all three options used in the current study.
 
@@
 In order to examine diurnal, seasonal, and solar cycle trends in echo occurrence, the monthly average occurrence was found for each universal time (UT) 2-min interval and for each 45-km range bin.
-If measurements were conducted at a 2-min resolution, each radar cell had an occurrence of 1 or 0.
-For a higher resolution, the values were weighted accordingly so that if an echo was observed for every possible time within the 2-min interval, the overall echo occurrence was still 1.
-The data for each time/range cell were averaged over the entire month.
-Figures \ref{fig:month_avg_occ}a--d show monthly average occurrence expressed in percentage for four selected months in year 2010 versus UT and slant range.
+Figures \ref{fig:month_avg_occ}a--\ref{fig:month_avg_occ}d show monthly average occurrence expressed in percentage for four selected months in year 2010 versus UT and slant range.
 The four months presented were chosen to be representative of different seasons.
-Figures \ref{fig:month_avg_occ}e--h on the right similarly shows the monthly average peak \(F\)-region electron density NmF2 obtained from the IRI model.
-The cosine white contours in both left and right panels show the approximate range position of the \(F\)-region echo band for each month determined as described below.
+Figures \ref{fig:month_avg_occ}e--\ref{fig:month_avg_occ}h on the right similarly shows the monthly average peak \(F\)-region electron density \(N_mF_2\) obtained from the IRI model using its default option of the URSI coefficient maps and the bottomside thickness B0 from \citet{Altadill2009}.
+For the other five IRI options, the plasma density profiles were also calculated and used in ray tracing calculations, as discussed below.
+The \(N_mF_2\) values for these other five options are not presented here.
+As compared with the IRI 2012 default option, the average percent difference for the three CCIR options were 15\%--22\% in March, June, and September and \(\sim9\)\% in December.
+For the three URSI options, they were 3\%--9\% in March, June, and September and \(\sim2\)\% in December.
+Finally, the cosine white contours in both Figures \ref{fig:month_avg_occ}a--\ref{fig:month_avg_occ}d and \ref{fig:month_avg_occ}e--\ref{fig:month_avg_occ}h show the approximate range position of the \(F\)-region echo band for each month determined as described below.
 
@@
 \includegraphics[width=12.5 cm,angle=90]{p1f1.pdf}
-\caption{Monthly average plots of echo occurrence (left) and peak electron density (right) for 2010.
+\caption{Monthly average plots of (a--d) echo occurrence and (e--h) peak electron density for 2010.
 Four months are selected (one for each season) to show seasonal variations.
@@
 The dashed white lines are the upper and lower bounds of the model band.
-The top right corner of each panel shows the parameters that characterize the band for that month: \(A\) is amplitude of cosine function (in km), \(B\) is time shift (in hrs), \(C\) is range shift from 0 (in km), and BW is the band width (constant distance between the upper and lower dashed white lines).
-In Figures \ref{fig:month_avg_occ}e--\ref{fig:month_avg_occ}h the same echo location model is also given (white lines) along with expected echo locations based on raytracing simulations (dark red line).}
+The top right corner of each panel shows the parameters that characterize the band for that month: \(A\) is amplitude of cosine function, \(B\) is time shift, \(C\) is range shift from 0, and BW is the band width.
+In Figures \ref{fig:month_avg_occ}e--\ref{fig:month_avg_occ}h the same echo location model is also given (white lines) along with expected echo locations based on ray tracing simulations (solid lines).
+The red (grey) lines refer to three URSI (CCIR) options (see text for details).}
 \label{fig:month_avg_occ}
@@
 
-Figures \ref{fig:month_avg_occ}a--d show a distinct \(F\)-region band of high occurrence in every month at ranges 500--1500 km.
+Figures \ref{fig:month_avg_occ}a--\ref{fig:month_avg_occ}d show a distinct \(F\)-region band of high occurrence in every month at ranges 500--1500 km.
 Some \(E\)-region echoes are also observed at shorter ranges 180--300 km in June at 01--06 UT and in December at 17--24 UT.
+In addition, Figures \ref{fig:month_avg_occ}a, \ref{fig:month_avg_occ}c, and \ref{fig:month_avg_occ}d show some echoes with low-occurrence near 04 UT.
+These observations refer to the region just equatorward of the polar cap near the noon sector (looking through the polar cap) and thus may represent a radar signature of the dayside cusp \citep{Yeoman2002}.
+In the following analysis, we focus on the maid \(F\)-region band between 500 km and 1500 km which refers to observations in the polar cap.
+
 The \(F\)-region band range position and peak occurrence within the band was changing most clearly throughout the day in March and September, i.e.
 diurnal variation in both parameters is observed clearly in spring and fall months.
-To highlight the high echo occurrence band, a contour of 25\% occurrence is plotted in each left panel by the pink line and this shows a clear diurnal, cosine-like variation with UT in March and September.
+To highlight the high echo occurrence band, a contour of 25\% occurrence is plotted in each Figure \ref{fig:month_avg_occ}a--\ref{fig:month_avg_occ}d by the pink line.
+The midpoints were also found between the top and bottom parts of the pink 25\% occurrence contour.
+These are plotted by the small black dots and referred to below as the band position \(r(t)\), where range \(r\) is in kilometers and time \(t\) is in hours.
 
-The midpoints were also found between the top and bottom parts of the pink 25\% occurrence contour and the cosine function of the form
+Both the 25\% occurrence contours and the band position \(r(t)\) exhibited a clear cosine-like variation with a 24-h period in March and September 2010.
+Very similar variations were observed in other years and other near-equinoctial months.
+Further, similar variations were also observed for many individual days.
+IN summer months, the variation was largely flat, Figure \ref{fig:month_avg_occ}d, which can be also represented by a cosine variation shifted from zero with a small amplitude.
+To further characterize this diurnal variation, a cosine function of the form
 \begin{equation}\label{E:Model}
@@
 \end{equation}
-was fitted to these points.
-Here \(R\) is the range of the center of the band in kilometers, \(t\) is the time in hours, and  \(A\), \(B\), and \(C\) are fit coefficients with units of kilometers, hours, and kilometers, respectively.
-The white solid line in each panel of Figure \ref{fig:month_avg_occ} shows the resulting fit \(R\left(t\right)\).
+was fitted to the band position values \(r(t)\).
+Here \(R(t)\) is the fitted range of the center of the band in kilometers, and  \(A\), \(B\), and \(C\) are fit coefficients with units of kilometers, hours, and kilometers, respectively.
+The white solid line in Figures \ref{fig:month_avg_occ}a--\ref{fig:month_avg_occ}h shows the resulting fit \(R\left(t\right)\).
 The white dashed lines represent the upper and lower limits of the band.
@@
 The original cosine fit was then shifted up or down by this amount as \(R\left(t\right) \pm \sigma\) to give the upper and lower limits of the band.
-These limits match the pink 25\% occurrence contour well because all the points in the pink contour are approximately the same distance away from the central fit (white solid line).
-The band width (BW) could then be defined as \(2\sigma\).
-The fit parameters and band width are given in the top-right corner of Figures \ref{fig:month_avg_occ}a--d.
+The bandwidth (BW) could then be defined as \(2\sigma\).
+The fit parameters and band width are given in the top right corner of Figures \ref{fig:month_avg_occ}a--\ref{fig:month_avg_occ}d.
 For the rest of this study, ``echoes within the band'' refer to data points that fall between the upper and lower white dashed lines.
-From Figures \ref{fig:month_avg_occ}a--d, both the center position and upper/lower bounds for the band shown by the white lines capture the band and its diurnal variation in range very well for all months.
+From Figures \ref{fig:month_avg_occ}a--\ref{fig:month_avg_occ}d, both the center position and upper/lower bounds for the band shown by the white lines capture the band and its diurnal variation in range very well for all months.
 Therefore, these estimates are used below to quantify the echo band behavior, particularly in context of expected electron densities.
@@
 MCM is oriented in such a way that it samples daytime (nighttime) ionosphere in roughly the first (second) half of the day in UT.
-The electron density on the right of Figure \ref{fig:month_avg_occ} clearly shows higher daytime densities that reach a peak near 05 UT at farther ranges and lower nighttime densities that reach a minimum near 17 UT.
-The \(F\)-region echo band thus was closer to the radar in the daytime (high densities) and farther away at night (low densities).
-This feature is most clear for the March and September datasets.
+The electron density profiles change significantly between the nighttime and daytime periods, with one important consequence being higher peak densities \(N_mF_2\) during the day.
+This is illustrated in Figure \ref{fig:month_avg_occ}e--\ref{fig:month_avg_occ}h that clearly shows higher daytime \(N_mF_2\) values that reach a peak near 05 UT at farther ranges and lower nighttime \(N_mF_2\) values that reach a minimum near 17 UT.
+The other important effect is an increase in the peak height \(h_mF_2\) during the night (not presented here).
+
+The \(F\)-region echo band was also closer to the radar in the daytime and farther away at night.
+This feature is most clear for the March and September data sets.
 It is much less clear in June, although one can still see a portion of the cosine-like variation at 02--10 UT in June in the pink contour.
@@
 The minimum was much deeper and more extended in June so that essentially no echoes were observed at 12--24 UT.
-At this time, the expected location of echoes based on the cosine-fit model (white lines) is farther away from MCM than in any month and any UT interval, but no echoes are observed either at these ranges or any other ranges.
+At this time, the expected location of echoes based on the cosine-fit model (white lines) is farther away from MCM than in any month and any UT interval within uncertainty, but no echoes are observed either at these ranges or any other ranges.
 This is in contrast with March and September, where occurrence is reduced at nighttime but not to zero.
@@
 
-The final observation about Figure \ref{fig:month_avg_occ} concerns the results of raytracing simulations that were conducted for each hour of day 15 in each of the four months shown in Figure \ref{fig:month_avg_occ}.
-In these simulations, IRI densities in a 2D simulation domain aligned with beam 14 of MCM were used.
-The range locations where orthogonality condition was expected are plotted in Figures \ref{fig:month_avg_occ}e--h by the dark red line.
-The gaps during the nighttime in these lines in all months but December indicate that no orthogonality was expected at any range in MCM beam 14 during these periods.
-The red line is located between the solid and bottom dashed white lines, indicating a reasonable agreement between the expected and observed backscatter locations.
-The important exception is seen, however, for nighttime periods (or, more accurately, whenever densities were lower than \(\sim1.5\times10^5\) cm\(^{-3}\), dark blue colors).
-During those periods, no orthogonality was expected based on IRI densities, but significant (albeit reduced) amount of backscatter was observed in March and September.
+The final observation about Figure \ref{fig:month_avg_occ} concerns the results of raytracing simulations that were conducted for each hour of day 15 in each of the four months shown in Figure \ref{fig:month_avg_occ} using a version of standard ray tracing tools based on numerical solutions of the Hamiltonian raypath equations \citep{Haselgrove1963,Jones1975} that allow the consideration of arbitrary density distributions.
+In these simulations, IRI densities in a 2-D simulation domain aligned with beam 14 of MCM were used.
+That is, plasma densities employed were varying both horizontally and vertically.
+The grid bins in the horizontal and vertical directions were 5 km and 1 km, respectively, which is well below the radar cell size of 45 km.
+Rays separated by \(1\deg\) in elevation were found under the assumption of ordinary mode propagation.
+Locations were founds where the ray was nearly perpendicular (within \(1\deg\)) to the local magnetic field based on the International Geomagnetic Reference Field 12 model \citep{Thebault2015}.
+
+The range locations where orthogonality condition was expected are plotted in Figures \ref{fig:month_avg_occ}e--\ref{fig:month_avg_occ}h by six lines corresponding to six different options of the IRI model described above.
+The red (grey) kines refer to three URSI (CCIR) options.
+These lines are seen between the solid and bottom dashed white lines, indicating a reasonable agreement between the expected and observed backscatter locations.
+The important exception is seen, however, for nighttime periods (low-density periods).
+During these periods, no orthogonality was expected based on IRI densities, but significant (albeit reduced) amount of backscatter was observed in March and September.
 This implies a significant source of nighttime plasma density that is unaccounted for by the IRI model.
+IRI is known to underestimate \(N_mF_2\) during equinoctial nighttime periods in the northern polar cap \citep{Themens2014}, and our observations suggest that a similar underestimation also occurs in the southern polar cap.
+The UT extent of the period with no expected orthogonality is reduced in the CCIR option, which suggests a better ionospheric representation in this option.
+This result is also consistent with that reported by \citet{Themens2014}.
 
 The same analysis was performed for the entire period of interest 2010--2013 and the band characteristics are presented in Figure \ref{fig:year_line}.
-It shows the band (a) position amplitude \(A\), (b) time shift \(B\), (c) range position \(C\), and (d) width (BW) plotted versus month.
-Each year is shown as a line in a different color as indicated in the top-right corner of panel (a).
-
-\begin{figure}
-\includegraphics[width=11.0 cm,angle=90]{p1f2.pdf}
-\caption{The variations in the echo band (a) amplitude of the fitted cosine function \(A\), (b) time shift or time of day of the peak \(B\), (c) average position of the band or the vertical shift of the cosine function from zero \(C\), and (d) width (BW).
-Different lines are for different years with the legend given in Figure \ref{fig:year_line}a.}
-\label{fig:year_line}
-\end{figure}
-
+It shows the band position amplitude \(A\), Figure \ref{fig:year_line}a, time shift \(B\), Figure \ref{fig:year_line}b, range position \(C\), Figure \ref{fig:year_line}c, and width (BW), Figure \ref{fig:year_line}d, plotted versus month.
+Each year is shown as a line in a different color as indicated in the top-right corner of Figure \ref{fig:year_line}a.
 All four parameters peak in austral winter and decrease in summer months.
-Solar cycle variation is evident in (a) the band position amplitude and (c) the average band position.
+Solar cycle variation is evident in the band position amplitude, Figure \ref{fig:year_line}a, and the average band position, Figure \ref{fig:year_line}c.
 In general, both parameters were larger in years closer to solar minimum (year 2010) and decreased moving towards solar maximum (2013).
 In both cases, the variation existed primarily in the winter months (around June).
-In contrast, the seasonal variation was also evident in (b) the band time shift and (d) the band width, but no clear solar cycle trend was present.
+In contrast, the seasonal variation was also evident in the band time shift, Figure \ref{fig:year_line}b, and the band width, Figure \ref{fig:year_line}d, but no clear solar cycle trend was present.
 In fact, the band time shift was quite consistent between different years, except for winter months.
@@
 
+\begin{figure}
+\includegraphics[width=11.0 cm,angle=90]{p1f2.pdf}
+\caption{The variations in the echo band (a) amplitude of the fitted cosine function \(A\), (b) time shift or time of day of the peak \(B\), (c) average position of the band or the vertical shift of the cosine function from zero \(C\), and (d) width (BW).
+Different lines are for different years with the legend given in Figure \ref{fig:year_line}a.
+The fit errors are given by the vertical bars.}
+\label{fig:year_line}
+\end{figure}
+
 In the previous analysis, the focus was on band characteristics related to range and their variations.
@@
 A horizontal slice of each panel of Figure \ref{fig:year_color} represents the time series for one month.
-The band position \(R\left(t\right)\) (average echo occurrence) is shown in Figures \ref{fig:year_color}a--\ref{fig:year_color}d (Figures \ref{fig:year_color}e--\ref{fig:year_color}h) as a color contour plot representing the time series for all months of the year simultaneously; different rows refer to  different years.
-The left panels look much smoother than the right ones; this is simply because each line on the left is a fitted cosine \(R\left(t\right)\), while on the right it is the average ``raw'' occurrence between two fitted cosine lines \(R\left(t\right)\pm \sigma\).
-The white data gaps at the bottom of Figures \ref{fig:year_color}d and \ref{fig:year_color}h reflect the fact that no MCM data were collected in January 2010.
+The band position \(r\left(t\right)\) is shown in Figures \ref{fig:year_color}a--\ref{fig:year_color}d as a color contour plot representing the time series for all months of the year simultaneously; different rows refer to  different years.
+Similarly, the average echo occurrence is shown in Figures \ref{fig:year_color}e--\ref{fig:year_color}h.
+Figures \ref{fig:year_color}a--\ref{fig:year_color}d also present the fitted band position R(t) as line contours using the same colour scheme.
+
+The white gaps appear in Figures \ref{fig:year_color}a--\ref{fig:year_color}d during the periods with no estimates for \(r(t)\).
+For the fitted band position \(R(t)\), these gaps are filled except for January 2010 when no MCM data were available.
+Outside the gaps, the \(R(t)\) contours match well the underlying color pattern.
+INside the gaps, they represent extrapolation of the same pattern.
+This information is alos useful ot analyze, particularly considering that the fitted errors for the months that contibute the most to the gaps were less tan 20\%, Figures \ref{fig:month_avg_occ}b and \ref{fig:year_line}.
 
@@
 \includegraphics[width=12.0 cm,angle=90]{p1f3.pdf}
-\caption{Contour plots of the (a)--(d) echo band range and (e)--(h) occurrence versus UT and season. Each row shows the data for a different year as given in the top-right corner.}
+\caption{Contour plots of the (a--d) echo band position \(r(t)\) and (e)--(h) occurrence versus UT and season.
+Each row shows the data for a different year as given in the top right corner.
+The unfilled color contours in Figures \ref{fig:year_color}a--\ref{fig:year_color}d show the fitted band position \(R(t)\) using the same color scheme.}
 \label{fig:year_color}
@@
 
-In Figures \ref{fig:year_color}a--d, the band position \(R\left(t\right)\) peaks in nighttime (second half of the day) and in winter.
-This peak (yellow-to-red colors) is seen in the same plot area as the low occurrence plot cells in Figure \ref{fig:year_color}e--h (blue color).
-The amplitude of the range peak became consistently lower between 2010--2013.
-This indicates that the high occurrence band moved closer to the radar for years further away from the solar minimum.
-
-The minima in the band position \(R\left(t\right)\) (blue contours) were observed during the daytime and near equinoxes.
-This feature was most clear in 2010 when these minima were observed near 03 UT in March and September; these were also the two months presented in Figure \ref{fig:month_avg_occ}.
-The echo occurrence pattern, Figures \ref{fig:year_color}e--h, is less clear in terms of its maxima and their relationship with range minima.
+In Figures \ref{fig:year_color}a--\ref{fig:year_color}d, the band position \(r\left(t\right)\) is larger near the edges of the gap.
+The fitted band position \(R(t)\) also shows a peak near the center of the gap (winter nighttime periods).
+The peak values for both \(r(t)\) and \(R(t)\) became lower between 2010 and 2013.
+This indicates that the high-occurrence band moved closer to the radar for years further away from the solar minimum.
+The minima in \(r(t)\) and \(R\left(t\right)\) were observed during the daytime and near equinoxes.
+This feature was most clear in 2010 when these minima were observed near 03 UT in March and September; these were also the two months presented in Figures \ref{fig:month_avg_occ}a and \ref{fig:month_avg_occ}c.
+The echo occurrence pattern, Figures \ref{fig:year_color}e--\ref{fig:year_color}h, is less clear in terms of its maxima and their relationship with range minima.
 Typically, one equinoctial maximum was more pronounced as in 2011--2013, with no clear preference towards either March of September.
@@
 the \(E\)- and \(F\)-region electron densities, as explained in Section \ref{Sec:Intro}.
-The model NmE and NmF2 values have been employed for this analysis, Section \ref{Sec:Exp}.
+The model \(N_mE\) and \(N_mF_2\) values have been employed for this analysis, Section \ref{Sec:Exp}.
 To keep model run times at a manageable level, here we consider 10-min cadence data including the average echo occurrence within the band (in Figures \ref{fig:month_avg_occ}--\ref{fig:year_color}, the 2-min data were considered).
 
-Figures \ref{fig:scatter}a and \ref{fig:scatter}b present the results of this analysis for year 2010 in the form of scatter plots of band position versus (a) SZA and (b) model NmF2, while Figures \ref{fig:scatter}c and \ref{fig:scatter}d are scatter plots of echo occurrence versus (c) SZA and (d) model NmF2.
-In all four panels, the model NmE is shown by the color of the points, with the color bar given to the right.
-The points in the top row show some clustering in the form of closed loops (some of them appear just as horizontal lines).
+Figures \ref{fig:scatter}a and \ref{fig:scatter}b present the results of this analysis for year 2010 in the form of scatter plots of band position versus (Figure \ref{fig:scatter}a) SZA and (Figure \ref{fig:scatter}b) model \(N_mF_2\), while Figures \ref{fig:scatter}c and \ref{fig:scatter}d are scatter plots of echo occurrence versus (Figure \ref{fig:scatter}c) SZA and (Figure \ref{fig:scatter}d) model \(N_mF_2\).
+In all four panels, the model \(N_mE\) is shown by the color of the points, with the color bar given to the right.
+The points in Figures \ref{fig:scatter}a and \ref{fig:scatter}b show some clustering in the form of closed loops (some of them appear just as horizontal lines).
 This is simply a consequence of the periodic nature of the cosine fit; each loop refers to the data for one month.
@@
 \includegraphics[width=17.0 cm]{p1f4.pdf}
-\caption{The echo band range versus the (a) solar zenith angle (SZA) and (b) model peak F2 electron density (NmF2).
+\caption{The echo band range versus the (a) solar zenith angle (SZA) and (b) model peak F2 electron density (\(N_mF_2\)).
 Each point is color-coded in model \(E\) region density.
-Figures \ref{fig:scatter}c and d show the same but for the  echo occurrence within the band.}
+Figures \ref{fig:scatter}c and \ref{fig:scatter}d show the same as Figures \ref{fig:scatter}a and \ref{fig:scatter}b but for the  echo occurrence within the band.}
 \label{fig:scatter}
@@
 
-Figure \ref{fig:scatter}a shows a group of blue points referring to nighttime observations (SZA \(>\) 90\(\deg\)) that slope upwards from a band position range of \(\sim\)800 km to \(\sim\)2000 km with a SZA increasing.
+Figure \ref{fig:scatter}a shows a group of blue points referring to nighttime observations (\(\text{SZA}>90\deg\)) that slope upwards from a band position range of \(\sim\)800 km to \(\sim\)2000 km with a SZA increasing.
 This indicates that the band moved into further ranges as radar sampled sectors farther away from sunset.
 The color of the points indicates that \(E\)-region electron density was typically lower at these times as well, as expected.
-In Figure \ref{fig:scatter}b, the same trend is seen as blue points starting at intermediate NmF2 values of \(\sim1.5 \times 10^5\) cm\(^{-3}\) and low ranges near 800 km, but increasing to ranges of \(\sim\)2000 km as NmF2 decreased to low values of \(\sim0.5 \times 10^5\) cm\(^{-3}\).
-This indicates that both representations (using SZA or model NmF2) are approximately equivalent: the band was further from the radar at low NmF2/high SZA values.
-At high NmF2/high NmE/low SZA values, i.e.
+In Figure \ref{fig:scatter}b, the same trend is seen as blue points starting at intermediate \(N_mF_2\) values of \(\sim1.5 \times 10^5\) cm\(^{-3}\) and low ranges near 800 km, but increasing to ranges of \(\sim\)2000 km as \(N_mF_2\) decreased to low values of \(\sim0.5 \times 10^5\) cm\(^{-3}\).
+This indicates that both representations (using SZA or model \(N_mF_2\)) are approximately equivalent: the band was further from the radar at low \(N_mF_2\)/high SZA values.
+At high \(N_mF_2\)/high \(N_mE\)/low SZA values, i.e.
 during daytime conditions, the range did not change much.
 In summer months (horizontal loops mentioned above) in particular, this variation was virtually nonexistent.
-Overall, Figures \ref{fig:scatter}a and \ref{fig:scatter}b show a clearly-defined pattern of a decreasing band position \(R\left(t\right)\) with \(t\) moving away from sunset and SZA increasing from 90\(\deg\), NmF2 decreasing and NmE also decreasing.
+Overall, Figures \ref{fig:scatter}a and \ref{fig:scatter}b show a clearly-defined pattern of a decreasing band position with \(t\) moving away from sunset and SZA increasing from 90\(\deg\), \(N_mF_2\) decreasing, and \(N_mE\) also decreasing.
 
@@
 There appears to be a similar peak when occurrence is plotted against \(F\)-region electron density (Figure \ref{fig:scatter}d).
-The NmF2 value corresponding to occurrence peak is between \(1.0\times 10^5\)--\(2.5\times 10^5\) cm\(^{-3}\) depending on the month.
+The \(N_mF_2\) value corresponding to occurrence peak is between \(1.0\times 10^5\) and \(2.5\times 10^5\) cm\(^{-3}\) depending on the month.
 At both lower and higher \(F\)-region densities, occurrence decreases.
@@
 Based on Figure \ref{fig:scatter}c, these generally have the lowest occurrence of all daytime measurements.
-Overall, Figures \ref{fig:scatter}c and \ref{fig:scatter}d show that echo occurrence peaks near the terminator where NmF2 is moderate and decreases for SZA increasing and decreasing from 90\(\deg\).
+Overall, Figures \ref{fig:scatter}c and \ref{fig:scatter}d show that echo occurrence peaks near the terminator where \(N_mF_2\) is moderate and decreases for SZA increasing and decreasing from 90\(\deg\).
 
@@
 In the following analysis, we attempt to determine whether there is a direct relationship between IMF behavior and echo occurrence on the scale of a few hours.
-We compare the MCM echo occurrence variations with the simultaneous IMF observations using a method analogous to that employed by \citet{Makarevich2012} in their study of auroral \(E\)-region irregularities (see also our section \ref{Sec:Exp} for a brief description of IMF data analysis).
-
-The average occurrence within the band was found at 10-min intervals as in the previous analysis.
-The average IMF components over corresponding 10-min intervals were also found.
-Figure \ref{fig:ut_occ} presents the entire 4-year dataset of echo occurrence versus UT, with color representing average IMF component over each plot bin.
-The bin sizes were 1-h in time and 2\% in occurrence and Figure \ref{fig:ut_occ} shows contour plots made using this grid.
-The color bar for all three IMF components is shown to the right.
+We compare the MCM echo occurrence variations with the simultaneous IMF observations using a method analogous to that employed by \citet{Makarevich2012} in their study of auroral \(E\)-region irregularities (see also our Section \ref{Sec:Exp} for a brief description of IMF data analysis).
+In addition, we evaluate a dependence on the PCS index using the same approach.
 
@@
 \includegraphics[width=15.0 cm]{p1f5.pdf}
-\caption{IMF effects in echo occurrence.
-The echo occurrence versus UT was binned in 1-hr bins in UT and 2\% bins in occurrence and the average IMF (a) Bx, (b) By, and (c) Bz components over each plot cell are shown by the color.}
+\caption{The MCM echo occurrence versus UT color coded in the average (a) IMF By, (b) IMF Bz, and (c) PCS values.}
 \label{fig:ut_occ}
@@
 
-Figure \ref{fig:ut_occ}a and \ref{fig:ut_occ}c show minimal color structure in the first half of the day (0--12 UT; daytime observations).
-This indicates that in this time sector, neither positive nor negative IMF Bx or Bz was conducive to high echo occurrence.
-The dark blue region between 20\%--60\% occurrence and 0--10 UT of Figure \ref{fig:ut_occ}b (By component) indicates some association between low occurrence and negative By.
-The second half of the day (12--24; nighttime observations) exhibits a definite color structure for all three IMF components.
-Figure \ref{fig:ut_occ}a has an orange and yellow area between 18--21 UT at above 60\% occurrence.
-This means that in this time sector, high echo occurrence was observed during predominantly positive IMF Bx.
-Similarly, in Figure \ref{fig:ut_occ}b, the large blue area between 12--24 UT and above 50\% indicates an association between high occurrence and negative IMF By.
-Moreover, Figure \ref{fig:ut_occ}b shows that the color changes roughly monotonically from orange and yellow to green and blue with increasing occurrence.
-This means that higher occurrence was observed for more negative IMF By.
-A similar color structure is less clear in Figure \ref{fig:ut_occ}c (Bz  component), but still present.
-For example, Figure \ref{fig:ut_occ}c has a blue section above 70\% occurrence and between 16--24 UT, so high occurrence is also associated with negative IMF Bz in this sector.
-A similar change in color (from yellow to blue) with an increase in occurrence is also observed here, although it is not as monotonic (color pattern is somewhat patchy).
-Overall, this analysis indicates that higher nighttime occurrence is seen during periods of more positive Bx, negative By, and negative Bz.
-
-Figure \ref{fig:by_bz} contains the same IMF data as Figure \ref{fig:ut_occ}, but here the IMF By and Bz effects are compared directly.
-IMF By and Bz are plotted on the horizontal and vertical axes, respectively, and color indicates relative occurrence, which was found as explained below.
-First, the average echo occurrence within the band over the whole month was found at 10-min intervals (these values were previously presented in Figure \ref{fig:year_color}, but at 2-min resolution).
-This time series was then subtracted from the average echo occurrence within the band for each day of the month (no averaging over the month).
-This resulted in a time series that approximately represents the variation in echo occurrence relative to a quiet-day season-specific diurnal average or the relative occurrence.
-This approach allows us to isolate the IMF effects from those due to diurnal and seasonal variations \citep{Makarevich2012}.
-Points were divided into six groups based on UT and plotted in different panels of Figure \ref{fig:by_bz}.
-
-%\begin{figure}
-%\includegraphics[width=11.0 cm,angle=90]{occ_vs_IMF_avg.pdf}
-%\caption{Relative occurrence versus IMF By and Bz components for 6 different UT sectors. Each plot cell is color-coded based on the average relative occurrence over that cell.}
-%\label{fig:by_bz}
-%\end{figure}
-
-Figures \ref{fig:by_bz}a--\ref{fig:by_bz}c (representing 0--12 UT) show very little color structure.
-Each panel is uniformly patchy, indicating no preference towards relatively high or relatively low echo occurrence for any IMF conditions.
-This is consistent with the minimal color structure observed in the 0--12 UT interval in Figure \ref{fig:ut_occ}.
-In the second half of the day (Figure \ref{fig:by_bz}d--\ref{fig:by_bz}f), the left half of each panel is more red and orange, while the right half is more blue.
-In particular, Figure \ref{fig:by_bz}d has a peak in relative occurrence (orange) at (\(-3, -2\)) nT in the (By, Bz) coordinates and an occurrence trough (blue) at (\(+5, -3\)) nT.
-Figure \ref{fig:by_bz}e has a region of high relative occurrence between (\(-\)4--0, \(-\)8--0) nT and a region of low relative occurrence between 2--10 nT in IMF By.
-Figure \ref{fig:by_bz}f generally has high relative occurrence in the bottom-left corner of the panel and low relative occurrence in the top-right corner of the panel.
-This means that high (low) relative occurrence is observed for periods of negative (positive) IMF By and Bz.
-Comparing Figures \ref{fig:by_bz}d--\ref{fig:by_bz}f, the dividing line between the red/orange areas and the blue area is close to vertical in Figure \ref{fig:by_bz}e (along By \(\sim\)0 nT), but rotates to be at almost 135\(\deg\) in Figure \ref{fig:by_bz}f (stretching between the top-left and bottom-right corners).
-Overall, these results indicate increasing importance of the Bz component in the last time sector, namely, positive Bz leads to negative relative occurrence.
-These results also agree with those shown in Figure \ref{fig:ut_occ}, where the patch of blue (negative Bz) was seen between 16--24 UT and above 70\% occurrence.
+The average occurrence within the band was found at 1-h UT bins using the entire 4 year data set.
+The average IMF components and PCS values were also found.
+The data were binned using 1-h UT and 2\% occurrence bins and smoothed using a \(3\times3\) window.
+Figure \ref{fig:ut_occ} presents contour plots of echo occurrence  versus UT, with color representing average (a) IMF By, (b) IMF Bz, and (c) PCS index.
+Figures \ref{fig:ut_occ}a and \ref{fig:ut_occ}b show almost no features in the first half of the day (0--12 UT; daytime observations).
+The second half of the day (12--24 UT; nighttime observations) exhibits a definite color structure for both IMF components.
+The echo occurrence was generally higher for more negative By ad Bz.
+Moreover, in Figure \ref{fig:ut_occ}a, the bins with IMF By below \(-2\) nT (blue colors) ar elocated at progressively lower occurrence values.
+This feature is less evident for IMF Bz, which indicates that IMF By component has a stronger association with the F-region echo occurrence than IMF Bz.
+In Figure \ref{fig:ut_occ}c, the average PCS index is somewhat enhanced before 12 UT, a result consisten with previous statistical analysis of the PCS index \citep[e.g.][Figure 4]{Troshichev2007}.
+There is no clear relationship between the echo occurrence and the PCS index except for the nighttime occurrence to be enhanced at large PCS values and the daytime occurrence to be either high or low at large PCS values.
 
@@
 
-\subsection{\(F\) Region Echo Band: A Comprehensive Model and Density Implications}
+\subsection{\(\mathbf{F}\) Region Echo Band: A Comprehensive Model and Density Implications}
 
@@
 One possible reason for this difference is more favorable propagation conditions in the polar cap with greater chances to reach orthogonality with magnetic field, which is a necessary condition for Bragg scatter to be received by coherent scatter radars.
-Whether any observed occurrence trend is due to changing propagation or production is difficult to resolve based on HF radar echo occurrence observations alone, as explained in section \ref{Sec:Intro}.
+Whether any observed occurrence trend is due to changing propagation or production is difficult to resolve based on HF radar echo occurrence observations alone, as explained in Section \ref{Sec:Intro}.
 A reasonable assumption often utilized is that propagation will be mostly responsible for changing location of echo occurrence.
@@
 The current study refines this approach by including the location information explicitly in the way occurrence estimates are obtained.
-Further, the information about variations in the typical ranges of backscatter is important by itself, particularly since it directly reflects variations in the \(F\) region density.
+Further, the information about variations in the typical ranges of backscatter is important by itself, particularly since it directly reflects variations in the \(F\)-region density.
 This provides a useful (albeit somewhat indirect) way to monitor and analyze plasma production and transport processes in a region where few opportunities for direct continuous measurements of plasma density currently exist.
 
-The two important outcomes of the current study related to this indirect density analysis were (1) development of an empirical model to describe location of \(F\) region backscatter and (2) analysis of characteristics derived from the model.
-The \(F\) region backscatter was predominantly observed in the form of a band that was well described by a simple model, equation (\ref{E:Model}).
-The \(F\) region echo band exhibited systematic diurnal and seasonal effects, Figure \ref{fig:month_avg_occ}, which was reflected in systematic changes of the fitted band characteristics \(A\), \(B\), \(C\), and band width (BW), Figure \ref{fig:year_line}.
+The two important outcomes of the current study related to this indirect density analysis were (1) development of an empirical model to describe location of \(F\)-region backscatter and (2) analysis of characteristics derived from the model.
+The \(F\)-region backscatter was predominantly observed in the form of a band that was well described by a simple model, Equation (\ref{E:Model}).
+The \(F\) region echo band exhibited systematic diurnal and seasonal variations, Figure \ref{fig:month_avg_occ}, which was reflected in systematic changes of the fitted band characteristics \(A\), \(B\), \(C\), and band width (BW), Figure \ref{fig:year_line}.
 These changes were quite consistent with those expected for plasma density.
-For example, the echo band was moving to further ranges during the night while density expected from IRI model was decreasing, Figures \ref{fig:month_avg_occ}e--h.
-Similarly, the amplitude of range variations \(A\) and average position \(C\) of the echo band exhibited seasonal and solar cycle variations consistent with those expected for densities, Figures \ref{fig:year_line}a and c.
+For example, the echo band was moving to further ranges during the night while density expected from IRI model was decreasing, Figures \ref{fig:month_avg_occ}e--\ref{fig:month_avg_occ}h.
+Similarly, the amplitude of range variations \(A\) and average position \(C\) of the echo band exhibited seasonal and solar cycle variations consistent with those expected for densities, Figures \ref{fig:year_line}a and \ref{fig:year_line}c.
 
-Another novel aspect of the current study was a comprehensive set of raytracing simulations that was conducted using IRI densities.
-In general, the diurnal and seasonal variations were very similar between simulations (red line in Figures \ref{fig:month_avg_occ}e--h) and observations (white lines).
+Another novel aspect of the current study was a comprehensive set of ray tracing simulations that was conducted using IRI densities.
+In general, the diurnal and seasonal variations were very similar between simulations (red line in Figures \ref{fig:month_avg_occ}e--\ref{fig:month_avg_occ}h) and observations (white lines).
 One important exception was observed during the night near equinoxes.
-No orthogonality was predicted by model simulations, whereas a significant (albeit reduced) amount of backscatter was observed, Figures \ref{fig:month_avg_occ}a and c.
+No orthogonality was predicted by model simulations, whereas a significant (albeit reduced) amount of backscatter was observed, Figures \ref{fig:month_avg_occ}a and \ref{fig:month_avg_occ}c.
 That is, an occurrence gap (i.e.
@@
 
-This result is significant as it implies a considerable additional source of density which would provide additional refraction and bring backscatter to orthogonality at the ranges consistent with the model.
+This result is significant as it implies a considerable additional source of density as compared with IRI outputs.
+This density would provide additional refraction and bring backscatter to orthogonality at the ranges consistent with the occurrence model.
 This prompts an important question about the origin of this source.
-It is unlikely that it was local since this production is well described by IRI \citep[e.g.][]{Themens2014}.
-A reasonable assumption is then that additional density was transported from somewhere else.
-The two phenomena that can be responsible for this are polar patches \citep{Weber1984} and sun-aligned arcs \citep{Valladares1994}.
+An improved agreement between the observed and IRI modeled densities has been recently reported during periods of solar production-dominated dynamics for both URSI and CCIR options \citep{Themens2014}.
+The increased errors during periods of little solar-driven production were attributed ot transport processes \citep{Themens2014}.
+Following the same idea, a reasonable assumption is that additional density implied by our observations was transported from somewhere else.
+Several phenomena may be responsible for this including polar patches \citep{Wever1984}, Sun-aligned arcs \citep{Valladares1994}, and storm-enhanced densities \citep{Foster1993}.
 The polar patches in particular are known to travel long distances across the polar cap and even well into the nightside auroral zone \citep{Moen2007,Oksavik2010}.
+Some support for the polar patches as a feasible source for the plasma density also comes from the density measurements presented below.
 
-Direct density measurements are not available near MCM but more insights can be obtained from observations in the conjugate northern polar cap.
+Direct density measurements at a comparable temporal resolution were not available near MCM during the period under investigation, but more insights can be obtained from observations in the conjugate northern polar cap.
 Figure \ref{fig:patchy_example} shows an example of conjugate measurements during two selected events that fall within the equinoctial period of Figure \ref{fig:month_avg_occ}a.
-Shown are (a) IMF By and Bz components, (b) MCM line-of-sight velocity in beam 14 versus UT and range, (c) RKN velocity in nominally-conjugate beam 4, and (d) northern face of the Resolute Bay incoherent scatter radar (RISR-N) measurements of electron density versus UT and altitude on March 14, 2010.
-Figures \ref{fig:patchy_example}e--h present the same information but for a second event on April 6, 2010.
+Shown are IMF By and Bz components, Figure \ref{fig:patchy_example}a, MCM line-of-sight velocity in beam 14 versus UT and range, Figure \ref{fig:patchy_example}b, RKN velocity in nominally-conjugate beam 4, Figure \ref{fig:patchy_example}c, and northern face of the Resolute Bay incoherent scatter radar (RISR-N) measurements of electron density versus UT and altitude, Figure \ref{fig:patchy_example}d, on March 14, 2010.
+Figures \ref{fig:patchy_example}e--\ref{fig:patchy_example}h present the same information but for a second event on April 6, 2010.
 
@@
 \includegraphics[width=18.0 cm]{p1f6.pdf}
-\caption{Conjugate measurements for two selected events on (left) March 14, 2010 and (right) April 6, 2010.
-Shown are the (a) and (e) IMF By (blue) and Bz (red) components, (b) and (f) MCM line-of-sight velocity versus UT and slant range in beam 14, (c) and (g) RKN velocity in beam 4, and (d) and (h) RISR-N electron density versus UT and altitude.
-The color bar is the same for the two middle rows and is given to the right of Figure \ref{fig:patchy_example}f.}
+\caption{Conjugate measurements for two selected events on (a--d) March 14, 2010 and (e-h) April 6, 2010.
+Shown are the IMF By (blue) and Bz (red) components (Figures \ref{fig:patchy_example}a and \ref{fig:patchy_example}e), MCM line-of-sight velocity versus UT and slant range in beam 14 (Figures \ref{fig:patchy_example}b and \ref{fig:patchy_example}f), RKN velocity in beam 4 (Figure \ref{fig:patchy_example}c and \ref{fig:patchy_example}g), and RISR-N electron density versus UT and altitude (Figure \ref{fig:patchy_example}d and \ref{fig:patchy_example}h).
+The color bar is the same for Figures \ref{fig:patchy_example}b, \ref{fig:patchy_example}f, \ref{fig:patchy_example}c, and \ref{fig:patchy_example}g and is given to the right of Figure \ref{fig:patchy_example}f.}
 \label{fig:patchy_example}
@@
 Figure \ref{fig:patchy_example}b shows an echo band that fits well with the average pattern of Figure \ref{fig:month_avg_occ}a (note more limited ranges for Figure \ref{fig:patchy_example}).
-This is a clear example of ``persistent backscatter'' at MCM where echoes are present throughout the entire day with only range changing in agreement with the model of equation (\ref{E:Model}).
+This is a clear example of persistent backscatter at MCM where echoes are present throughout the entire day with only range changing in agreement with the model of Equation (\ref{E:Model}).
 In contrast, Figure \ref{fig:patchy_example}f shows an example of ``patchy backscatter'' at MCM, particularly after 12 UT.
 The range variation is still roughly consistent with the model.
-The RKN backscatter pattern looks similar if one accounts for a \(\sim\)12-h  shift in UT between day- and nighttime observations.
-Thus nighttime echoes are observed by RKN near 08 UT during the first event (it is daytime at MCM) and they are more persistent.
+
+The backscatter patches move away form MCM in Figure \ref{fig:patchy_example}b.
+For example, the patch near 17 UT moved by \(\sim700\) km in \(\sim\)20-min, which translates into \(\sim600\) m/s velocity.
+This is consistent with the Doppler velocity measured by MCM within the patch (dark blue colors).
+Similar consistency between the patch drift and plasma convection velocities was observed by \citet{Bristow2011} using MCM observations and, more recently, by \citet{Makarevich2015} using direct observations of 2-D convection velocity and density enhancements by RISR-N in the northern polar cap.
+This impolies that polar patches tend to move with the background convection, at least deep within the polar cap.
+
+In Figure \ref{fig:patchy_example}c, the RKN backscatter pattern looks similar if one accounts for a \(\sim\)12-h  shift in UT between daytime and nighttime observations.
+Thus, nighttime echoes are observed by RKN near 08 UT during the first event (it is daytime at MCM) and they are more persistent.
 The daytime echoes are seen by RKN at closer ranges near 18 UT and they are also persistent, similar to MCM daytime echoes near 08 UT.
@@
 
-The key information about density is provided by RISR-N observations, Figures \ref{fig:patchy_example}d and h.
+The key information about density is provided by RISR-N observations, Figures \ref{fig:patchy_example}d and \ref{fig:patchy_example}h.
 A diurnal variation is evident on both days, with a broad nighttime minimum centered near 06--08 UT and a maximum centered at 18--20 UT.
@@
 
-The IMF information presented in Figures \ref{fig:patchy_example}a and b also provides a clue into what conditions are associated with patchy backscatter.
+The IMF information presented in Figures \ref{fig:patchy_example}a and \ref{fig:patchy_example}b also provides a clue into what conditions are associated with patchy backscatter.
 When IMF is relatively stable on a time scale of \(\sim\)20 min, as is the case in the first event, backscatter is persistent.
@@
 Again, this is supportive of the polar patch scenario.
+In this scenario, polar patches contribute to the nighttime densities and the present MCM observations in context of ray tracing simulations suggest that they do on a regular basis.
 
-The presented statistical analysis of MCM observations thus imply that not only polar patches contribute to the nighttime densities but also that they do so on a regular basis.
-This causes the differences between night- and daytime densities implied by the observed backscatter ranges to be much smaller as compared with what is predicted by model raytracing predictions.
-
-\subsection{\(F\) Region Irregularities: Suppression and Enhancement by Sunlight}
+\subsection{\(\mathbf{F}\) Region Irregularities: Suppression and Enhancement by Sunlight}
 
@@
 Figure \ref{fig:month_ut} summarized observations and model results related to this issue for year 2010.
-It presents (a) MCM backscatter occurrence, (b) MCM backscatter range \(R(t)\), (c) model NmF2, and (d) SZA.
-Figures \ref{fig:month_ut}a and b are the same as Figures \ref{fig:year_color}d and h except that here 10-day (rather than monthly) occurrence averages are used and \(5\times5\)-window smoothing applied for occurrence.
-Figures \ref{fig:month_ut}c and d are remarkably similar in their contour patterns; one panel is an inversion of the other one, as expected for model NmF2 that is mostly SZA-controlled.
-The range panel (b) is also very similar to both, although here minima (darkest blue) are somewhat shifted in season from SZA minima/NmF2 maxima.
+It presents MCM backscatter occurrence, Figure \ref{fig:month_ut}a, MCM backscatter range \(R(t)\), Figure \ref{fig:month_ut} b, model \(N_mF_2\), Figure \ref{fig:month_ut}c, and SZA, Figure \ref{fig:month_ut}.
+Figures \ref{fig:month_ut}a and \ref{fig:month_ut}b are the same as Figures \ref{fig:year_color}d and \ref{fig:year_color}h except that here 10-day (rather than monthly) occurrence averages are used and \(5\times5\) window smoothing applied for occurrence.
+Figures \ref{fig:month_ut}c and \ref{fig:month_ut}d are remarkably similar in their contour patterns; one panel is an inversion of the other one, as expected for model \(N_mF_2\) that is mostly SZA-controlled.
+Figure \ref{fig:month_ut}b (range) is also very similar to both, although here minima (darkest blue) are somewhat shifted in season from SZA minima/\(N_mF_2\) maxima.
 This is not surprising since range variation represents changing propagation conditions and is mostly due to ionization locally-produced by the Sun.
 
-The MCM occurrence shown in Figure \ref{fig:month_ut}a is also similar to other 3 panels for nighttime observations (deep occurrence minimum represented by blue colors in Figure \ref{fig:month_ut}a).
+The MCM occurrence shown in Figure \ref{fig:month_ut}a is also similar to other three panels for nighttime observations (deep occurrence minimum represented by blue colors in Figure \ref{fig:month_ut}a).
 However, there are both similarities and differences in daytime observations.
 Backscatter is enhanced in the dusk and dawn sectors (yellow and orange contours between 00 and 12 UT in Figure \ref{fig:month_ut}a).
-This is when density is also enhanced as compared to nighttime in Figure \ref{fig:month_ut}c due to some illumination at SZAs of 70$^\circ$--90$^\circ$ in Figure \ref{fig:month_ut}d.
-However, backscatter suppression by sunlight is also evident as a secondary occurrence minimum at 02 UT in austral summer coincides with the main NmF2 maximum/SZA minimum.
+This is when density is also enhanced as compared to nighttime in Figure \ref{fig:month_ut}c due to some illumination at SZAs of 70\(\deg\)--90\(\deg\) in Figure \ref{fig:month_ut}d.
+However, backscatter suppression by sunlight is also evident as a secondary occurrence minimum at 02 UT in austral summer coincides with the main \(N_mF_2\) maximum/SZA minimum.
 Thus both effects/processes contribute to occurrence patterns.
@@
 \includegraphics[width=12.0 cm,angle=90]{p1f7.pdf}
-\caption{Contour plots of the (a) average occurrence within the band, (b) band range, (c) model peak F-region electron density (NmF2) and (d) solar zenith angle (SZA) versus UT and month in 2010.
+\caption{Contour plots of the (a) average occurrence within the band, (b) band range, (c) model peak F-region electron density (\(N_mF_2\)) and (d) solar zenith angle (SZA) versus UT and month in 2010.
 Color bars are given to the right.
@@
 
-A different view of occurrence dependencies on NmF2, SZA, and NmE was offered in Figure \ref{fig:scatter} in a more explicit form.
+A different view of occurrence dependencies on \(N_mF_2\), SZA, and \(N_mE\) was offered in Figure \ref{fig:scatter} in a more explicit form.
 The range and occurrence dependencies were found to be quite different, which, as discussed above, indicates that propagation effects are not dominant in occurrence variations.
@@
 
-A decrease at smaller SZA is consistent with the idea of sunlight wiping out gradients necessary for GDI to operate.
-Increase in NmE resulting in shorting of \(F\) region irregularities by the conducting \(E\) layer may have also contributed.
-A decrease at larger SZA, on the other hand, is associated with a decreasing NmF2, increasing range, and stronger under-refraction effects.
-Importantly, however, not all of this decrease is due to stronger under-refraction.
+A decrease at smaller SZA is consistent with the idea of sunlight smoothing out gradients necessary for GDI to operate.
+An increase in \(N_mE\) resulting in the perturbation electric fields being shorted out by the conducting \(E\) layer may have also contributed.
+A decrease at larger SZA, on the other hand, is associated with a decreasing \(N_mF_2\), increasing range, and stronger underrefraction effects.
+Importantly, however, not all of this decrease is due to stronger underrefraction.
 This is because range was increasing, but still well within the radar's FoV, Figure \ref{fig:scatter}a.
-That is, at the ranges where echoes were expected based on range trends, echoes were present but reduced in occurrence, Figure \ref{fig:month_avg_occ}a and c.
+That is, at the ranges where echoes were expected based on range trends, echoes were present but reduced in occurrence, Figure \ref{fig:month_avg_occ}a and \ref{fig:month_avg_occ}c.
 Thus an alternative explanation must be sought for an occurrence decrease at large SZA.
@@
 
-This study found no clear evidence on such a control in the dayside polar cap, Figures \ref{fig:ut_occ} and \ref{fig:by_bz}.
+This study found no clear evidence on such a control in the dayside polar cap, Figure \ref{fig:ut_occ}.
 On the nightside (12--24 UT), however, a clear pattern emerged where echo occurrence was enhanced for negative IMF By and Bz, Figure \ref{fig:ut_occ}.
-The pattern for IMF Bx was nearly an inversion of that for IMF By, implying a high degree of correlation between IMF By and \(-\)Bx.
 Interestingly, however, the IMF Bz pattern was somewhat different from that of IMF By, Figure \ref{fig:ut_occ}, and in fact less clearly defined.
@@
 The new and unexpected result of the current study was that IMF By effects were more clearly defined in the occurrence of backscatter from the polar cap.
-Another new result of the current study was that the occurrence dependence on IMF was changing in a systematic way with UT on the nightside, Figures \ref{fig:ut_occ} and \ref{fig:by_bz}.
-For example, near 12 UT occurrence was generally above 80\% when IMF By was on average below \(-\)1.5 nT (blue cells at the top of Figure \ref{fig:ut_occ}b).
-On the other hand, for the same IMF By, but at a later UT, occurrence was between 60\% and 100\%, i.e.
-blue area of the plot expanded lower in Figure \ref{fig:ut_occ}b.
+Another new result of the current study was that the occurrence dependence on IMF was changing in a systematic way with UT on the nightside, Figures \ref{fig:ut_occ}.
 
@@
 
+One other mechanism through which IMF can exert influence over echo occurrence rates is through its indirect control of geomagnetic substorms.
+It is well known that enhanced particle precipitation and convection electric fields during substorms can be a significant source for plasma structures in the auroral region \citep[e.g.][]{Tsunoda1988}.
+At decameter scales, however, propagation conditions appear ot be a major limiting factor, with broad reductions in the echo occurrence rates observed in the nightside auroral zone \citep{Wild2008}.
+On the other hand, this is not the case in the polar cap, where no noticeable change in the cho occurrence is observed \citep[Figures 1 and 2]{Wild2008}.
+The absence of noticeable occurrence dependence on geomagnetic activity represented by PCS, Figure \ref{fig:ut_occ}c, is possibly related with that.
+Thus, IMF effects are most likely to be observed in the echo occurrence through its control of the convection environment as described above.
+
 \section{Summary and Conclusions}
@@
 The band range position is, therefore, strongly controlled by the local solar illumination conditions in the polar cap.
-On the nightside, however, a comparison between observations and model raytracing simulations reveal a presence of significant additional source of ionization.
+On the nightside, however, a comparison between observations and model ray tracing simulations reveal a presence of significant additional source of ionization.
 Conjugate observations of backscatter and density indicate that polar patches transported from the dayside are a viable source for this ionization.
@@
 Backscatter occurrence statistic exhibits evidence of both suppression and enhancement by sunlight, with a clear occurrence peak observed near the solar terminator.
-Suppression occurs at the largest SZA values observed and is consistent with diminishing role of gradients in irregularity production and increasing role of conducting $E$ layer.
+Suppression occurs at the largest SZA values observed and is consistent with diminishing role of gradients in irregularity production and increasing role of conducting \(E\) layer.
 Enhancement by sunlight is mostly associated with more favorable propagation conditions at intermediate SZA values and large-scale gradients near the solar terminator possibly serving as seeds for further structuring processes.

### 2017-01-28 introduction.tex
Corrections made to Introduction

@@
 Plasmas are fluid masses of charged particles, often formed when neutral atoms and molecules in a gas are ionized by external energy.
-Instances of plasma include everything from florescent lightbulbs and flames to more exotic examples such as nuclear fusion reactor experiments, most stars, nebula, and the interstellar medium.
-In fact, about 99\% of matter in the known universe exists in the plasma state.
+Instances of plasma include everything from florescent lightbulbs and flames to more exotic examples such as nuclear fusion reactor experiments, most stars, nebulas, and the interstellar medium.
+In fact, about 99\% of visible matter in the known universe exists in the plasma state.
 Of particular interest to this thesis is the plasma in Earth's ionosphere.
@@
 These dynamics can generate waves in the plasma, which cause density perturbations.
-Additionally, variations in temperature and electric and magnetic field strength can occur on a variety of scales.
+Similarly, variations in plasma temperature and electric and magnetic field strength can occur on a variety of scales.
 This thesis focuses on mechanisms by which density variations in ionospheric plasma occur, often referred to as plasma structures.
@@
 \section{The Solar-Terrestrial Environment}
-The solar terrestrial environment begins with the Sun in the center of the solar system.
+The solar-terrestrial environment begins with the Sun in the center of the solar system.
 The Sun is composed of highly energized plasma that is gravitationally bound together.
-The surface of the Sun is highly dynamic, populated with sunspots, which are dark, relatively cool regions with intense magnetic fields, as well as coronal holes, which are low density areas characterized with a continuous outflow of plasma.
+The surface of the Sun is highly dynamic, populated with sunspots, which are dark and relatively cool regions with intense magnetic fields, as well as coronal holes, which are low-density areas characterized with a continuous outflow of plasma.
 The intense magnetic fields associated with sunspots can break down releasing a large amount of energy and plasma, known as a solar flare.
 Planet-sized masses of plasma known as coronal mass ejections (CMEs) can be ejected from the surface, which can be large enough to maintain an internal magnetic field as they travel through the solar system.
-The frequency of sunspots and large outbursts of plasma changes with the 11-year long solar cycle.
-A large number of sunspots, high occurrence of fast, dense plasma outbursts, and a highly variably magnetic field indicate solar maximum.
+The frequency of sunspots and large outbursts of plasma changes with the 11-year-long solar cycle.
+A large number of sunspots, high occurrence of fast and dense plasma outbursts, and a highly variable magnetic field indicate solar maximum.
 Conversely, solar minimum is characterized by a low sunspot number and slow and relatively steady outflow.
@@
 
-The Sun interacts with the rest of the solar system through the solar wind, a continuous outflow of plasma that caries the Sun's magnetic field with it.
+The Sun interacts with the rest of the solar system through the solar wind, a continuous outflow of plasma that carries the Sun's magnetic field with it.
 This is shown in Figure \ref{fig:ste}.
-The solar wind, represented by the green arrows, typically has a density around 10 cm\(^{-3}\) and is traveling away from the Sun at approximately 400 km/s at Earth's orbit.
+The solar wind, represented by the green arrows in Figure \ref{fig:ste}, typically has a density around 10 cm\(^{-3}\) and is traveling away from the Sun at approximately 400 km/s at the Earth's orbit.
 However, like the Sun, the solar wind is highly dynamic and these parameters can change drastically over a period of minutes.
 CMEs, for instance, create a substantial increase in density and are often preceded by a shock, a region of hot, compressed plasma caused by the CME moving at supersonic speeds.
-In addition to the outflow of plasma, the solar wind also consists of the Sun's magnetic field extended into the outer limits of the solar system, called the interplanetary magnetic field (IMF), dark blue lines in Figure \ref{fig:ste}.
+In addition to the outflow of plasma, the solar wind also contains the Sun's magnetic field extended into the outer limits of the solar system, called the interplanetary magnetic field (IMF), dark blue lines in Figure \ref{fig:ste}.
 Because the Sun has a short rotation period (about 33 days), the IMF gets twisted into a spiral, commonly called the Parker Spiral \citep{Parker1958}.
@@
 	\includegraphics[width=\textwidth,angle=0]{ste.pdf}
-	\caption[Solar-Terrestrial Environment]{Diagram of the solar-terrestrial environment.
-The solar wind, green arrows, contains both a steady stream of charged particles and the IMF, dark blue.
-The red region represents the magnetosheath.
-The manetopause is shown by the black dashed line.
-Closed magnetic field lines from the earth are shown in grey while open field lines are shown in light blue.
-The ionosphere is shown in light green above the Earth's surface.}
+	\caption[Solar-terrestrial environment]{Diagram of the solar-terrestrial environment.
+The solar wind, green arrows, contains both a steady stream of charged particles and the interplanetary magnetic field, dark blue.
+The red region represents the magnetosheath, the shock layer formed due to the supersonic plasma outflow from the Sun.
+The magnetopause encloses the magnetosphere dominated by the geomagnetic field and is shown by the black dashed line.
+Closed magnetic field lines from the Earth are shown in grey, while open field lines are shown in light blue.
+The ionosphere is shown in yellow above the Earth's surface.}
 	\label{fig:ste}
@@
 The IMF interacts with the Earth's own magnetic field, and this interaction creates the magnetosphere.
-Without the IMF, the magnetosphere would be approximately a dipole field, but dynamic pressure from the solar wind causes the Sun facing side to be compressed to 6--10 Earth radii and the rear side to be stretched into an extended tail (hundreds of Earth radii).
+Without the IMF, the magnetosphere would be approximately in a dipole shape, but dynamic pressure from the solar wind causes the Sun-facing side to be compressed to 6--10 Earth radii and the opposite side to be stretched into an extended tail (hundreds of Earth radii).
 Because the solar wind is typically a supersonic flow, a bow shock forms on the dayside of the magnetosphere.
@@
 
-Closer to the Earth's surface, the magnetosphere also interacts with the Earth's ionosphere, a layer of partially ionized plasma in the upper atmosphere, light green in Figure \ref{fig:ste}.
+Closer to the Earth's surface, the magnetosphere also interacts with the Earth's ionosphere, a layer of partially ionized plasma in the upper atmosphere, yellow in Figure \ref{fig:ste}.
 The interaction between charged particles in the ionosphere with magnetic field lines couples the ionosphere, magnetosphere, and solar wind to create non-trivial dynamics.
@@
 \label{sec:ionosphere}
-The Earth's ionosphere is a region of the upper atmosphere that ranges approximately between 50--1000 km in altitude where neutral gases have been partially ionized, resulting in free electrons and ions mixed with neutral particles.
-Two competing processes occur in the ionosphere to create a peak in electron density.
-Solar illumination and particle precipitation ionize neutrals, and the density of ions increase as altitude decreases because there are more neutral molecules available to ionize.
-However, below a certain point, the concentration of ions becomes large enough that recombination into neutral particles becomes a substantial factor, reducing the ion density.
-The exact altitude where this density peak occurs is variable with time of day, location, season, and solar cycle.
-A profile of how electron density changes with altitude is shown in Figure \ref{fig:densprofile}, found from the International Reference Ionosphere (IRI) model run on January 1, 2007 at \(75\deg\) N, \(0\deg\) E, geographic (the IRI model is discussed further in Section \ref{sec:iri}).
+The focus of this body of work is the Earth's ionosphere, a region of the upper atmosphere that extends approximately between 50--1000 km in altitude where neutral gases have been partially ionized, resulting in free electrons and ions mixed with neutral particles.
+Two competing processes occur in the ionosphere to create a peak in the electron density.
+Solar illumination and particle precipitation ionize neutrals, and the density of ions increases as altitude decreases because there are more neutral molecules available to ionize.
+However, below a certain point, the concentration of ions becomes large enough that recombination becomes a substantial factor, reducing the ion density.
+The exact altitude where this density peak occurs varries with time of day, location, season, and solar cycle.
+Representative model examples of electron density changes with altitude are shown in Figure \ref{fig:densprofile}, found from the International Reference Ionosphere (IRI) model run on January 1, 2007 at \(75\deg\) N, \(0\deg\) E, geographic (the IRI model is discussed further in Section \ref{sec:iri}).
 
@@
 	\includegraphics[width=\textwidth]{densprofile.pdf}
-	\caption[Ionosphere density profile]{Altitude profile of electron density throughout the ionosphere found from the IRI model.  Typical ranges of the \(D\), \(E\), and \(F\) regions are identified.}
+	\caption[Ionosphere density profile]{Altitude profiles of electron density throughout the ionosphere found from the IRI model.  Typical extents of the \(D\), \(E\), and \(F\) regions are identified.}
 	\label{fig:densprofile}
@@
 The ionosphere is typically divided into three regions, Figure \ref{fig:densprofile}.
-The D region is located between 80 and 90 km in altitude and typically only exists in the daytime when the ionosphere is sunlit.
-The E region typically exists between 90 and 130 km with a density peak between 105 and 110 km, depending on factors such as time of day and season.
-Above 130 km is considered to be the F region, which has a density peak around 250 km, Figure \ref{fig:densprofile} \citet{Luhmann1995}.
-The F region can actually have two peaks under some daytime conditions, notated as the F1 peak and the F2 peak.
+The \(D\) region is located between 80 and 90 km in altitude and typically only exists in the daytime when the ionosphere is sunlit.
+The \(E\) region typically exists between 90 and 130 km with a density peak between 105 and 110 km, depending on factors such as time of day and season.
+Above 130 km is considered to be the F region, which has a density peak around 250 km, Figure \ref{fig:densprofile} \citep{Kivelson1995}.
+The \(F\) region can actually have two peaks under some daytime conditions, notated as the \(F1\) peak and the \(F2\) peak.
 The peak heights and densities of each of these regions is highly variable both diurnally and seasonally.
-In Figure \ref{fig:densprofile}, the daytime profile is shown with the solid line next to a nighttime profile with a dashed line.
-Note that night electron densities are lower at all altitudes due to the lack of photoionization during nighttimes.
-In addition, the F-region peak moves to higher altitudes at night.
-The focus of this thesis will be the F region.
+In Figure \ref{fig:densprofile}, the daytime profile is shown with the solid line next to a nighttime profile shown with a dashed line.
+Note that night electron densities are lower at all altitudes due to the lack of photoionization during the night time.
+In addition, the \(F\)-region peak moves to higher altitudes at night.
+The focus of this thesis is on the \(F\) region.
 
 Because the concentrations of ions and neutral particles changes with altitude, each of these regions have distinct plasma physical properties that make them unique.
-The main issue is whether the the large-scale motion of a particle is controlled more by the magnetic field (the particle is magnetized) or by collisions with other particles (the particle is collisional).
-This can be determined by comparing the collision frequency, \(\nu_\alpha\), to the gyrofrequency, \(\Omega_\alpha\), of a particular species, where \(\alpha\) corresponds to either electrons, \(e\), or ions, \(i\).
+The main issue is whether the large-scale motion of a particle is controlled more by the magnetic field (the particle is magnetized) or by collisions with other particles (the particle is collisional).
+This can be determined by comparing the collision frequency, \(\nu_\alpha\), to the gyrofrequency, \(\Omega_\alpha\), of a particular species, where \(\alpha\) refers to either electrons, \(e\), or ions, \(i\).
 If \(\Omega_\alpha \gg \nu_\alpha\), the particle completes many gyrations between each collision, so its motion is determined mostly by the magnetic field and it is magnetized.
 If \(\nu_\alpha \gg \Omega_\alpha\), the particle collides with other particles much more often than it complete a gyration, so it is considered collisional.
-The gyrofrequency is dependent on the mass and charge of the particle and the strength of the magnetic field, \(\Omega_\alpha = q_\alpha B/m_\alpha\), so is roughly constant through the ionosphere because the magnetic field strength does not change much throughout these altitudes.
+The gyrofrequency is dependent on the mass and charge of the particle and the strength of the magnetic field, \(\Omega_\alpha = q_\alpha B/m_\alpha\), so is roughly constant throughout the ionosphere because the magnetic field strength does not change much throughout these altitudes.
 The collision frequency depends on the density of a species relative to the density of all other species in the plasma and can be calculated from a series of standard expressions \citep{Schunk1980,Schunk2009}.
@@
 In the \(F\) region where the neutral density is much lower, both ions and electrons are magnetized, \(\Omega_\alpha \gg \nu_\alpha\).
-These differences in how charged particles move impact the conductivity and plasma waves development, which is further discussed in Section \ref{sec:lit_theory}.
+These differences in how charged particles move impact the conductivity and plasma waves development, which is further discussed in Section \ref{sec:lit_instabilities}.
 
@@
 
-This thesis is concerned with the polar ionosphere, which is a particularly interesting and dynamic region due to the presence of open field lines to the solar wind.
-The main focus of this work will be plasma density perturbations, often called irregularities.
+This thesis is concerned with the polar ionosphere, which is a particularly interesting and dynamic region due to the presence of open field lines that are directly connected to the solar wind.
+The main focus of this work is plasma density waves or structures, also known as irregularities.
 Density irregularities in the polar cap can occur on scales ranging from thousands of kilometers to less than a centimeter \citep{Tsunoda1988}, and a summary of some of these irregularities is presented in Figure \ref{fig:polarirreg}.
@@
 Polar patches are large density enhancements (usually at least twice the background plasma density) that travel across the polar cap with the background convection, Figure \ref{fig:polarirreg}a \citep{Weber1984,Valladares1994}.
-Polar holes are plasma density depletions that typically occur in the \(F\) region slightly poleward of the auroral oval \citep{Benson2001}.
-Sun-aligned arcs are large density forms that stretch across the polar cap along the sun-earth line.
+Polar holes are plasma density depletions that typically occur in the \(F\) region slightly poleward of the auroral oval \citep{Brinton1978}.
+Sun-aligned arcs are large density forms that stretch across the polar cap along the Sun-Earth line.
 They are very narrow and often characterized by large and complex velocity shears on either side of the arc \citep{Valladares1991}.
-Large scale plasma structures and the density gradients that surround them often serve as a platform for smaller-scale structures to form as well.
+Large-scale plasma structures and the density gradients that surround them often serve as a platform for smaller-scale structures to form as well.
 Polar patches, for instance, are known to form large finger-like structures along their trailing edges, Figure \ref{fig:polarirreg}b \citep{Gondarenko2004b,Hosokawa2016}.
@@
 	\caption[Plasma irregularities in the polar ionosphere]{Plasma density irregularities occur in the polar cap at different scale sizes.
-Polar patches, orange, move with the background ionospheric convection, indicated by red contours (a).
-Smaller-scale structures often move develop along the edges of polar patches where strong density gradients exist (b).
-Intermediate scale irregularities are primarily responsible for disrupting radio signals traveling through the ionosphere (c).
-Field-aligned irregularities, FAIs, are density perturbations that form along the magnetic field lines, light blue (d).}
+Polar patches, orange in Figures \ref{fig:polarirreg}a and \ref{fig:polarirreg}b, move with the background ionospheric convection, indicated by red arrows.
+Smaller-scale structures often develop along the edges of polar patches where strong density gradients exist, Figure \ref{fig:polarirreg}b.
+Intermediate-scale irregularities are primarily responsible for disrupting radio signals traveling through the ionosphere, Figure \ref{fig:polarirreg}c.
+Field-aligned irregularities, FAIs, are density perturbations that form along the magnetic field lines, Figure \ref{fig:polarirreg}d.}
 	\label{fig:polarirreg}
@@
 
-Intermediate-scale plasma irregularities are those ranges from 100 m -- 30 km \citep{Kelley2009}.
-These are also referred to as scintillation-causing structures and are responsible for many of the negative space weather effects that are observed on Earth.
+Intermediate-scale plasma irregularities are those with sizes of 100 m -- 30 km \citep{Kelley2009}.
+These are also referred to as scintillation-causing structures and are responsible for many of the aderse space weather effects that are observed on Earth.
 One important example of a space weather effect is when the ionosphere changes the phase and amplitude of the original signal, called radio scintillation, which affects the ability of satellites to communicate with ground receivers.
@@
 Small-scale structuring is generally considered to be all irregularities less than 100 m \citep{Kelley2009}.
-Although small-scale irregularities are not responsible for scintillation that directly impact GPS/GNSS performance, they can be the easiest to study because of large data sets collected from high frequency ionospheric radars globally for several decades (discussed further in Section \ref{sec:superdarn}).
+Although small-scale irregularities are not responsible for scintillation that directly impact GPS/GNSS performance, they can be the easiest to study because of large datasets collected with a large number of ionospheric radars around the world that have been in operation for many decades (discussed further in Section \ref{sec:superdarn}).
 Because electrons move far more easily along magnetic field lines than across them, density perturbations tend to be aligned with the magnetic field and are often known as field-aligned irregularities (FAI), Figure \ref{fig:polarirreg}d.
@@
 Typically, a threshold is selected for the minimum power acceptable for a return to be considered an actual signal instead of background noise.
-LoS velocity is the component of the total plasma drift velocity that is measured along the transmitted signal direction.
-Because radars measure velocity through the doppler shift of backscatter, only the component of the velocity vector can be measured.
-Spectral width represents the width of the doppler power spectrum and can determine characteristics of irregularities \citep{Greenwald1985}.
+LoS velocity is the component of the total plasma irregularity velocity that is measured along the transmitted signal direction.
+Because radars measure velocity through the Doppler shift of backscatter, only a component of the velocity vector can be measured.
+Spectral width represents the width of the Doppler power spectrum and can be used to infer other characteristics of irregularities, such as their lifetime \citep{Greenwald1985}.
 
@@
 This allows the distance between the radar and the target to be found (based on the time difference between when the pulse is transmitted and received).
-The size of the range gate is determined by the length of a radar pulse, Equation \ref{eqn:range_gate}.
+The size of the range gate is determined by the length of a radar pulse, Equation \ref{eqn:range_gate}
 \begin{equation}
 	\label{eqn:range_gate}
-	\Delta r = \frac{c\Delta t}{2}
-\end{equation}.
+	\Delta r = \frac{c\Delta t}{2}.
+\end{equation}
 Here, \(\Delta r\) is the size of a range gate, \(\Delta t\) is the pulse length, and \(c\) is the speed of light.  Longer pulses result in larger range gates, resulting in poorer spatial resolution.
@@
 When the radar transmits continuously, backscatter will be received, but because it is impossible to know when that signal was transmitted, the difference in time between transmission and receiving cannot be used to calculate how far away the backscatter volume is.
-Conversely, if the radar only transmits a pulse of a certain length of a certain length and then ``listens'' for its return, the time between transmission and return is known and the distance the pulse traveled can be calculated and therefore the location of the backscattering volume \citep{Farley1972,Greenwald1983}.
+Conversely, if the radar only transmits a pulse of a certain length of a certain length and then ``listens'' for its return, the time between transmission and return is known and the distance the pulse traveled can be calculated and therefore the location of the backscattering volume can be found \citep{Farley1972,Greenwald1983}.
 However, using this simple single-pulse method another pulse cannot be transmitted until the first pulse returns, limiting the temporal resolution that can be achieved.
@@
 	\label{eqn:nyquist}
-	f_n = \frac{f_s}{2}
-\end{equation}.
+	f_n = \frac{f_s}{2}.
+\end{equation}
 Radars operating in single-pulse mode must have a very low sampling frequency, as described above, particularly at long ranges where the pulse must travel a long distance.
-This severely limits the Doppler shift that can be measured, which places an upper limit on the LoS Doppler velocity that can be calculated with this technique.
-For an ionospheric radar observing a structure 2000 km away, it would take about 13 ms for a signal traveling at the speed of light to travel to the the structure and back to the radar, which corresponds to a sampling frequency of \(f_s = 75\) Hz.
+This severely limits the Doppler shift that can be measured, which places an upper limit on the LoS velocity that can be calculated with this technique.
+For an ionospheric radar observing a structure 2000 km away, it would take about 13 ms for a signal traveling at the speed of light to travel to the structure and back to the radar, which corresponds to a sampling frequency of \(f_s = 75\) Hz.
 The Nyquist frequency is then \(f_n = 37.5\) Hz, which corresponds to a maximum measurable Doppler velocity of \(\sim550\) m/s.
@@
 Instead of waiting for each individual signal to be received before transmitting the next, ionospheric radars typically employ a multipulse mode.
-In a multipulse mode, the radar transmits a series of pules with different time intervals or lags between them.
+In a multipulse mode, the radar transmits a series of pulses with different time intervals or lags between them.
 This can introduce complications when backscatter from different pulses at different ranges is received by the radar at the same time, referred to as cross-range interference, but these effects can be mitigated using correlation techniques \citep{Farley1972}.
@@
 The first observations of the ionosphere by CSRs were made in the 1930s \citep{Eckersley1937,Harang1938}.
-Over the next several decades, numerous used ground-based radars to advance our understanding of the ionosphere and radio aurora \citep{Hultqvist1964,Leadabrand1965,Unwin1972,Sahr1996}.
+Over the next several decades, numerous ground-based radars have been used to advance our understanding of the ionosphere and radio aurora \citep{Hultqvist1964,Leadabrand1965,Unwin1972,Sahr1996}.
 In the 1970s and 1980s, the Scandanavian Twin Auroral Radar Experiment (STARE) produced the first continuous large data-set, similar to how modern CSR systems are run.
-STARE consisted of two very high frequency (VHF) radars with overlapping FoVs that were designed to measure FAIs in the \(E\) region of the ionosphere \citep{Greenwald1997}.
+STARE consisted of two very high frequency (VHF) radars with overlapping fields-of-view (FoV) that were designed to measure FAIs in the \(E\) region of the ionosphere \citep{Greenwald1997}.
 The advantage of two radars with overlapping FoVs was the ability to observe the same structures from two different directions.
@@
 \label{sec:superdarn}
-The primary instruments used within this body of work are radars within the Super Dual Auroral Radar Network (SuperDARN), a global network of high frequency (HF) CSRs that was designed to measure small-scale plasma structures in the ionosphere and map the global plasma convection.
+The primary instruments used within this body of work are radars within the Super Dual Auroral Radar Network (SuperDARN), a global network of high frequency (HF) CSRs that was designed to observe small-scale plasma structures in the ionosphere and map the global plasma convection.
 The network currently consists of about 35 operational radars between the northern and southern hemispheres distributed at mid-, high-, and polar latitudes, Figure \ref{fig:superdarnmap}.
@@
 	\includegraphics[width=\textwidth,angle=180]{SuperDARNmap.pdf}
-	\caption[SuperDARN map]{FoVs of all operational SuperDARN radars in both the northern (left) and southern (right) hemispheres.
-Polar latitude radars are shown with the red outlines.
-The radars at Rankin Inlet (northern hemisphere) and McMurdo Station (southern hemisphere) are shaded in red because these two instruments are particularly important in the following studies.
+	\caption[SuperDARN map]{Fields-of-view of all operational SuperDARN radars in both the northern (left) and southern (right) hemispheres.
+Polar-latitude radars are shown with the red outlines.
+The radars at Rankin Inlet (northern hemisphere) and McMurdo Station (southern hemisphere) are shaded in red because these two instruments are particularly important to this body of work.
 Lines of constant magnetic latitude are shown in yellow at \(\Lambda = 40\deg, 60\deg, 80\deg\) in the northern hemisphere and at \(\Lambda = -40\deg,-60\deg,-80\deg\) in the southern hemisphere.}
@@
 
-It is necessary to use HF radars to study ionospheric structures in the polar cap because the magnetic field is close to vertical and in order for the radar beam to meet the perpendicularity condition to observe FAIs, the beam must be refracted through a dense ionosphere.
+The most important condition for observation of FAIs with CSRs is the so-called perpendicularity condition, according to which the radar beam has to reach the scattering volume perpendicular to the local magnetic field direction.
+For this reason, it is necessary to use HF radars to study ionospheric structures in the polar cap because the magnetic field is close to vertical and the beam must be refracted through a dense ionosphere to reach perpendicularity with the magnetic field.
 VHF radar beams are not refracted enough in the polar cap and the beam will pass through the ionosphere without FAIs producing backscatter.
-SuperDARN radars operate nominally between 8--20 MHz, which, according to the Bragg scatter equation for backscatter, corresponds to observing decameter-scale plasma waves.
+SuperDARN radars operate nominally between 8--20 MHz, which, according to the Bragg scatter condition, refers to decameter-scale plasma waves.
 Each radar consists of 16 independent antennas that transmit a radio wave and receive backscatter from the ionosphere.
 Beams are electronically steerable and most radars have between 16 and 24 beams in normal operation mode, each beam being \(3.25\deg\) in azimuth.
-Each beam consists of 75--100 range gates, which are generally either 15 km or 45 km in length \citep{Chisham2007}.
-SuperDARN radars were originally designed to used a 7 pulse ACF \citep{Farley1972,Greenwald1983,Greenwald1985}, however since 2011, most radars have begun using an 8 pulse sequence.
+Each beam consists of 75--100 range gates, which are generally either 15 km, 30 km, or 45 km in length \citep{Chisham2007}.
+SuperDARN radars were originally designed to used a 7-pulse ACF \citep{Farley1972,Greenwald1983,Greenwald1985}, however since 2011, most radars have begun using an 8-pulse sequence.
 The multi-pulse increment, \(\tau\), is 2400 \(\mu\)s, which corresponds to a Nyquist frequency of about 200 Hz, small enough to measure plasma drift velocities up to 3000 m/s.
@@
 
-One of the original purposes of SuperDARN was to produce maps of the plasma convection patterns in the polar caps using 2D velocity vectors.
+One of the original purposes of SuperDARN was to produce maps of the plasma convection patterns at high latitudes using 2D velocity vectors.
 Originally this was accomplished by considering two radars with overlapping FoVs.
-If both received backscatter from the same scattering volume, two LoS velocities could be found for that scattering volume, which could be combined to find a 2D velocity vector \citep{Ruohoniemi1989}.
+If both received backscatter from the same scattering volume, two LoS velocities could be found and combined to find a 2D velocity vector \citep{Ruohoniemi1989}.
 The modern method for creating convection maps involves taking all data recorded by all radars in a particular hemisphere and finding the best fit to a 2D ionospheric electrostatic potential using Legendre functions \citep{Ruohoniemi1998}.
 For convection maps, only data from the \(F\) region is considered.
-If there is not enough velocity data to confine the fit sufficiently, the measured data points are supplemented with a statistical model based on solar wind IMF conditions \citep{Ruohoniemi1995,Ruohoniemi2005}.
+If there is not enough velocity data to confine the fit sufficiently, the measured data points are supplemented with a statistical model based on the IMF conditions \citep{Ruohoniemi1995,Ruohoniemi2005}.
 
@@
 This creates a vast database of high-quality measurements of small-scale FAIs, which is very useful for other studies on irregularity occurrence and plasma structuring.
-In the studies presented here, data was primarily used from the SuperDARN radars located at Rankin Inlet, Canada (RKN) and McMurdo Station, Antarctica (MCM).
+In the studies presented here, data were primarily used from the SuperDARN radars located at Rankin Inlet, Canada (RKN) and McMurdo Station, Antarctica (MCM).
 These two radars have been shaded in Figure \ref{fig:superdarnmap}.
-MCM is the only polar radar in the southern hemisphere.
-Although the FoVs of other radars cover parts of the polar cap, they only to at either E region ranges or the furthest of the F region ranges, for which there is rarely much backscatter.
-There are currently three polar radars in the northern polar cap, but the FoV of RKN overlaps the Resolute Bay Incoherent Scatter Radar FoV (discussed in Section \ref{sec:isr}, which makes it particularly useful for multi-instrument studies of the ionosphere.
+MCM is the only polar radar in the southern hemisphere that has been operational for more than 5 years.
+Although the FoVs of other radars cover parts of the polar cap, these parts refer to either \(E\) region ranges or the furthest of the \(F\) region ranges, for which there is rarely much backscatter.
+There are currently three polar radars in the northern polar cap, but the FoV of RKN overlaps the viewing area of an important complimentary instrument, the Resolute Bay Incoherent Scatter Radar FoV, discussed in Section \ref{sec:isr}, which makes it particularly useful for multi-instrument studies of the ionosphere.
 
-\subsection{Incoherent Scatter Radar}
+\subsection{Resolute Bay Incoherent Scatter Radar (RISR)}
 \label{sec:isr}
-Similar to CSRs, incoherent scatter radars (ISRs) are ground based radars that are used to probe the ionosphere.
+Similar to CSRs, incoherent scatter radars (ISRs) are ground-based radars that are used to probe the ionosphere.
 However, they use a fundamentally different technique to observe structuring, which allows them to measure bulk plasma parameters in contrast to small-scale irregularity characteristics.
@@
 
-The scattering cross section for a volume of N electrons in the ionosphere is given by \(\sigma_n = N \sigma_e\) for radio waves with a wavelength small compared to the Debye length of the plasma, where \(\sigma_e\) is the scattering cross section of a single electron \citep{Gordon1958,Fejer1960}.
-The relationship to the plasma Debye length is important because at scale sizes smaller than the Debye length, the plasma is not capable of organized motion, so the motion of the elections is only due to their own thermal energy and not larger scale plasma waves.
+The scattering cross section for a volume of \(N\) electrons in the ionosphere is given by \(\sigma_n = N \sigma_e\) for radio waves with a wavelength small compared to the Debye length of the plasma, where \(\sigma_e\) is the scattering cross section of a single electron \citep{Gordon1958,Fejer1960}.
+The relationship to the plasma Debye length is important because at scale sizes smaller than the Debye length, the plasma is not capable of organized motion, so the motion of the elections is only due to their own thermal energy and not larger-scale plasma waves.
 The signal returned from this kind of scatter was postulated to have a very low power and a very broad spectral width, however after the first ISR was built, the spectrum was found to be much narrower than previously expected \citep{Evans1969}.
@@
 In addition, 3D convection velocities and electric field vectors can be derived from LoS ion velocity measurements in some of the more modern ISRs by assuming both \(\vec{E}\) and \(\vec{V}_E\) are constant across the radar's FoV \citep{Heinselman2008}.
-More recently, a new algorithm has been developed that used the Lagrange method of undetermined multipliers to generate 3D convection velocity and electric field vectors without making this limiting assumption \citep{Nicolls2014}.
+More recently, a new algorithm has been developed that used the Lagrange method of undetermined multipliers to generate 2D convection velocity and electric field vectors without making this limiting assumption \citep{Nicolls2014}.
 Each instrument must be calibrated with known electron density measurements to account for noise and system constants \citep{Nicolls2007a}.
@@
 	\caption[ISR map]{FoVs of all ISRs in the northern polar region.
-The individual radars shown are identified by their commonly used acronyms: Poker Flat (PFISR), Resolute Bay North (RISR-N), Resolute Bay Canada (RISR-C), Sondrostrom (SFJ).
-RISR-N, which is primarially used throughout this thesis, is shown in orange.
-The FoV of the RKN SuperDARN radar is shown by the red outline.
+The individual radars shown are identified by their commonly used acronyms: Poker Flat (PFISR), Resolute Bay North (RISR-N), Resolute Bay Canada (RISR-C), Sondrestrom (SFJ).
+RISR-N, which is primarily used throughout this thesis, is shown in orange.
+The FoV of the SuperDARN RKN radar is shown by the red outline.
 Lines of constant MLAT are shown at \(\Lambda=80\deg,60\deg\) in yellow.}
@@
 Since then, at least 8 other ISRs have been deployed around the world at equatorial, mid-, and high-latitudes.
-A map of the ISR systems currently operational in the northern polar cap is shown in Figure \ref{fig:isrmap}.
-The most recent advancement has been the development of Advanced Modular Incoherent Scatter Radars (AMISR), of which there are currently three in use, one at the Poker Flat Rocket Range (PFISR), just north of Fairbanks, Alaska and two in Resolute Bay, Canada (RISR-N and RISR-C).
+A map of the ISR systems currently operational near the northern polar cap is shown in Figure \ref{fig:isrmap}.
+The most recent advancement has been the development of Advanced Modular Incoherent Scatter Radars (AMISR) \citep{Heinselman2008}, of which there are currently three in use, one at the Poker Flat Rocket Range (PFISR), just north of Fairbanks, Alaska and two in Resolute Bay, Canada (RISR-N and RISR-C).
 The advantage of these new systems is that they are electronically steerable, unlike older ISRs which consisted of a large dish that had to be physically moved to change the beam direction.
-This allows beams to be transmitted in many different directions at a very high time cadence, giving the radar the capability to make measurements in multiple directions almost simultaneously \citep{Nicolls2007a,Nicolls2007b,Bahcivan2010}.
+This allows beams to be transmitted in many different directions at a very high time cadence, giving the radar the capability to make measurements in multiple directions almost simultaneously \citep{Nicolls2007a,Nicolls2007b,Heinselman2008,Bahcivan2010}.
 This is very important for creating 2D maps of ionospheric conditions and tracking how they change in time \citep{Semeter2009,Dahlgren2012a,Dahlgren2012b}.
 
-AMISR systems typically have two types of pulses, a long pulse (LP) with 72 km range gates for \(F\)-region studies and an alternating code pulse (AC) with 4.5 km range gates gates for \(E\)-region studies.
+AMISR systems typically have two types of pulses, a long pulse (LP) with 72-km range gates for \(F\)-region studies and an alternating code pulse (AC) with 4.5-km range gates gates for \(E\)-region studies.
 The exact number and configuration of beams depends on the radar mode used and can be changed easily due to the system's electronic steering.
 In the common WorldDay mode, there are typically 11 beams and data is collected at \(\sim\)1 min intervals, although this can change slightly between different renditions of the WorldDay mode.
-WorldDay mode creates the characteristic star shaped FoV seen in Figure \ref{fig:isrmap}
 
 One of the most important advantages of ISR systems over CSRs is the ability to directly measure electron density in the ionosphere.
-However, because ISRs are much scarcer and are only typically run on in WorldDay mode a few days out of every month, data are not as widely available.
+However, because ISRs are much scarcer and are only typically run in WorldDay mode a few days out of every month, data are not as widely available.
 In addition, ISRs do not actually observe small-scale coherent plasma structures like the SuperDARN HF network does.
 However, they can directly measure large-scale structures and image density variations on scales larger that 100 km (limited by the number of beams and range gate size).
-For these region, it is often useful use a combination of CSR and ISR measurements in ionospheric plasma structuring studies, when possible.
+For these reasons, it is useful to use a combination of CSR and ISR measurements in ionospheric structuring studies, whenever possible.
 The primary ISR used in the present volume of work is the north face of the AMISR system at Resolute Bay, Canada, RISR-N, shown in orange in Figure \ref{fig:isrmap}.
-The FoV of RISR-N overlaps that of the RKN SuperDARN radar, red outline in Figure \ref{fig:isrmap}, making it ideal for these types of comparison studies, Chapters \ref{sec:paper1} and \ref{sec:paper3}.
+The FoV of RISR-N overlaps that of the RKN SuperDARN radar, red outline in Figure \ref{fig:isrmap}, making it ideal for these types of comparison studies in the polar cap.
+The two experimental studies that employ both the SuperDARN RKN and RISR-N systems are presented in Chapters \ref{sec:paper1} and \ref{sec:paper3}.
 
-\subsection{IRI Model}
+\subsection{International Reference Ionosphere (IRI) Model}
 \label{sec:iri}
-In addition to radars, models can be a useful tool for understanding the background conditions in the ionosphere, particular in regions where there is poor instrumental coverage.
+In addition to radars, models can be a useful tool for understanding the background conditions in the ionosphere, particular in regions with poor instrumental coverage.
 The International Reference Ionosphere (IRI) is an empirical model of the Earth's ionosphere.
@@
 
-Although the IRI model is very useful for providing background plasma density conditions where no instruments are available to make measurements, it is important to bear in mind that it is a model and therefor only represents the average conditions.
+Although the IRI model is very useful for providing background plasma density conditions where no instruments are available to make measurements, it is important to bear in mind that it is a model and therefore only represents the average conditions.
 Although it is fairly accurate at mid-latitudes \citep{Coisson2006,Bilitza2012}, it does not necessarily represent variations within the polar cap well \citep{Themens2014,Makarevich2015b}.
@@
 Despite this, the IRI model is still a very important tool for establishing a density profile at any time in any location in the ionosphere.
-And example of an altitude profile of plasma density provided by the IRI model is shown in Figure \ref{fig:densprofile}.
+An example of an altitude profile of plasma density provided by the IRI model was shown in Figure \ref{fig:densprofile}.
 The displayed profiles shows how the electron density changes with altitude at local noon (day) and midnight (night) on January 1, 2007 at \(75\deg\) N, \(0\deg\) E, geographic.
@@
 \section{A Brief Review of Plasma Structuring Theory and Observations in the Polar Cap}
-Much work has been done over the past several decades to study the complicated picture of plasma structuring in the polar cap, including theoretical analysis of plasma dispersion relations, simulations of plasma dynamics, and observations of plasma structuring using a variety of instruments.
-Here, a survey will be given of some of these advancements and the current state of understanding plasma structuring on a range of scales.
+Much work has been done over the past several decades to study the complex picture of plasma structuring in the polar cap, including theoretical analysis of plasma dispersion relations, simulations of plasma dynamics, and observations of plasma structuring using a variety of instruments.
+In this section, a survey is given of some of these advancements focusing on large- and small-scale structuring processes in the polar cap.
 
@@
 \label{sec:lit_patches}
-Large-scale structuring in the polar cap can take a variety of forms, Section \ref{sec:polar_structure}, but here we will focus on large regions of enhanced plasma density known as polar patches  \citep[e.g.][]{Weber1984,Weber1986,Buchau1983,Buchau1985}.
-One of the primary interests of this thesis is the relationships between irregularities of different scales and small-scale structuring has long been observed around polar patches \citep{Weber1984,Milan2002b,Moen2012}.
+Large-scale structuring in the polar cap can take a variety of forms, Section \ref{sec:polar_structure}, but here the focus is on large regions of enhanced plasma density known as polar patches  \citep[e.g.][]{Weber1984,Weber1986,Buchau1983,Buchau1985}.
+One of the primary interests of this thesis is the relationships between irregularities of different scales and small-scale structuring.
+This structuring has long been observed around polar patches \citep{Weber1984,Milan2002b,Moen2012}.
 Although small-scale structures have also been observed around sun-aligned arcs \citep[e.g.][]{Koustov2012}, these structures also tend to be more complicated and often are surrounded by intense shears and electric fields \citep{Safargaleev2000,Aikio2002,Kozlovsky2007}.
@@
 The first observations of polar patches were made in the 1960s by \citet{Hill1963}.
-They can range in size from 100 km up to 1500 km in diameter and can have peak densities as much as eight times that of the background ionospheric plasma \citep{Weber1986,Hosokawa2014}.
-Polar patches are know to emerge on the dayside and then propagate with the background plasma convection across the polar cap to the nightside, where they either disintegrate or recombine with enhanced-density plasma in the auroral oval \citep{Weber1985,Weber1986}.
+They can range in size from 100 km up to 1500 km and can have peak densities as much as eight times that of the background ionospheric plasma \citep{Weber1986,Hosokawa2014}.
+Polar patches are known to emerge on the dayside and then propagate with the background plasma convection across the polar cap to the nightside, where they either disintegrate or recombine with enhanced-density plasma in the auroral oval \citep{Weber1985,Weber1986}.
 
@@
 \begin{enumerate}
-	\item Discrete changes in solar wind parameters, usch as IMF, density, speed, and pressure \citep{Sojka1994}
+	\item Discrete changes in solar wind parameters, such as IMF, density, speed, and pressure \citep{Sojka1994}
 	\item Average flow patterns varying in time \citep{Anderson1988}
@@
 	\item Plasma flow jet channels that cut continuous streaches of plasma into segments 	\citep{Valladares1998}
-	\item Alfven wave coupling \citep{Prikryl1999}
-\end{enumerate}.
+	\item Alfven wave coupling \citep{Prikryl1999}.
+\end{enumerate}
 Although particle precipitation has been shown to create enhanced plasma density in the cusp region \citep{Rodger1994}, it is unreasonable for this mechanism to account for the largest density enhancements observed in polar patches.
@@
 High speed flow channels can ``cut'' segments off of a tongue of high-density plasma that has been pulled into the polar cap by convection \citep{Valladares1994,Valladares1998}.
-Models have also shown that changing large scale convection patterns quickly in time does produce density structures similar to patches \citep{Anderson1988}.
+Models have also shown that fast-changing large-scale convection patterns does produce density structures similar to patches \citep{Anderson1988}.
 However, the mechanism that seems to be dominant for the majority of polar patches observed is transient magnetic reconnection \citep{Carlson2012}.
@@
 Figure \ref{fig:magnetosphere} is a schematic of how the IMF, magnetosphere, and ionospheric polar cap are all interconnected.
-In general, magnetic field lines in the magnetosphere can either be closed (grey lines in Figure \ref{fig:magnetosphere}), meaning they connect only the north and south pole of the Earth, or open (light blue lines in Figure \ref{fig:magnetosphere}), when one of the ends of the field lines originates from the sun such that the field line is connected directly to the IMF.
-IMF lines that are not connected to the Earth are shown in dark blue in figure \ref{fig:magnetosphere}.
+In general, magnetic field lines in the magnetosphere can either be closed (grey lines in Figure \ref{fig:magnetosphere}), meaning they connect only the north and south pole of the Earth, or open (light blue lines in Figure \ref{fig:magnetosphere}), when one of the ends of the field lines originates from the Sun such that the field line is connected directly to the IMF.
+IMF lines that are not connected to the Earth are shown in dark blue in Figure \ref{fig:magnetosphere}.
 The polar cap boundary is generally defined as the border between open and closed magnetic field lines, shown by the green line in Figure \ref{fig:magnetosphere}.
@@
 By Faraday's Law, the total electromotive force, \(\xi\) around the polar cap boundary is equal to the rate of change of magnetic flux, \(\Phi_B\) within the polar cap, Equation \ref{eqn:pc_imf} \citep{Lockwood1992a}.
-In steady-state solutions, \(d\Phi_B/dt = 0\), so \(\xi=0\) by definition, however the polar cap is known to be a highly dynamic system, so changes in magnetic flux are expected.
+In steady-state solutions, \(d\Phi_B/dt = 0\), so \(\xi=0\) by definition, however the polar cap is known to be a highly dynamic system, so changes in magnetic flux are expected
 \begin{equation}
 	\label{eqn:pc_imf}
-	-\xi = \frac{d\Phi_B}{dt} = B\frac{dA_{pc}}{dt}
+	-\xi = \frac{d\Phi_B}{dt} = B\frac{dA_{pc}}{dt}.
 \end{equation}
-In the polar cap at ionospheric altitudes, the magnetic field strength is relatively constant at around \(5\times 10^-5\) T, so any changes in magnetic flux must be due to changes in the polar cap area, \(A_{pc}\) due to the polar cap boundary expanding or contracting \citep{Lockwood1992a}.
-The polar cap boundary changes in response to magnetic reconnection between the magnetosphere and the IMF  on both the day and night sides changing the amount of open magnetic flux.
+In the polar cap at ionospheric altitudes, the magnetic field strength is relatively constant at around \(5\times 10^{-5}\) T, so any changes in magnetic flux must be due to changes in the polar cap area, \(A_{pc}\) due to the polar cap boundary expanding or contracting \citep{Lockwood1992a}.
+The polar cap boundary changes in response to magnetic reconnection between the magnetosphere and the IMF  on both the day- and night-sides changing the amount of open magnetic flux.
 
@@
 	\caption[Magnetic reconnection footprint in the ionosphere]{Diagram of the interactions between the IMF, magnetosphere, and polar cap ionosphere.
-Closed field lines in the Earth's magnetosphere are shown in grey while open field lines are light blue.
+Closed field lines in the Earth's magnetosphere are shown in grey, while open field lines are light blue.
 The polar cap open-closed boundary is in green.
@@
 Magnetic reconnection occurs along the X-line (red) on the dayside of the magnetopause (dashed line).
-The X-line can be mapped back along magnetic field lines (orange) to the ionosphere, where their projection is the merging gap (pink dotted line).
+The X-line can be mapped back along magnetic field lines (orange) to the ionosphere, where its projection is the merging gap (pink dotted line).
 During reconnection events, the open-closed boundary expands equatorwards as more open field lines are formed and there is plasma flow across the merging gap.
@@
 Magnetic reconnection between the IMF and the Earth's magnetosphere occurs along the X-line on the dayside magnetopause (red line in Figure \ref{fig:magnetosphere}).
-The X-line in the magnetopause can be mapped along the magnetic field lines (orange) to the ionosphere, where it is referred to as the merging gap (pink dotted line in figure \ref{fig:magnetosphere}) and lies along the steady state polar cap boundary.
+The X-line in the magnetopause can be mapped along the magnetic field lines (orange) to the ionosphere, where it is referred to as the merging gap (pink dotted line in Figure \ref{fig:magnetosphere}) and lies along the steady state polar cap boundary.
 
@@
 The exact mechanism by which this happens is not well understood, but it is thought to be related to small variations in the IMF By component shifting the convection pattern slightly \citep{Cowley1980,Lockwood1992b}.
-By Figure \ref{fig:patch_formation}f, the open-closed boundary has returned completely to its original position and there is a newly formed patch within in the polar cap.
+By the time shown in Figure \ref{fig:patch_formation}f, the open-closed boundary has returned completely to its original position and there is a newly formed patch within the polar cap.
 The process can now repeat to form another patch within the polar cap.
@@
 A series of reconnection bursts can create a line of patches all propagating across the polar cap.
-Reconnection bursts typically last about 2 minutes with anywhere between 7 and 25 minutes between bursts \citep{Foster1984,Etemadi1988,Lockwood1992b}.
+Reconnection bursts typically last about 2 min with anywhere between 7 min and 25 min between bursts \citep{Foster1984,Etemadi1988,Lockwood1992b}.
 However, observations have shown convection flow to be close to continuous for much longer periods.
@@
 Dense daytime plasma is shown by the orange regions.
-The burst of reconnection is occurring in Figures \ref{fig:patch_formation}a--\ref{fig:patch_formation}c but stops in Figures \ref{fig:patch_formation}d--\ref{fig:patch_formation}f.
+The burst of reconnection is occurring in Figures \ref{fig:patch_formation}a--\ref{fig:patch_formation}c, but stops in Figures \ref{fig:patch_formation}d--\ref{fig:patch_formation}f.
 Figure adapted from \citet{Cowley1991}.}
@@
 If a wave has a component along the magnetic field line, the current-conductive instability (CCI) can be operational even if GDI is not \citep{Hoh1960,Ossakow1979,Chaturvedi1981}.
-Additionally instability mechanisms emerge if shears are considered, when \(\nabla E \neq 0\), such as the Kelvin-Helmholtz instability (KHI) \citep{Kintner1977,DAngelo1965}.
+Additional instability mechanisms emerge if shears are considered, when \(\nabla E \neq 0\), such as the Kelvin-Helmholtz instability (KHI) \citep{Kintner1977,DAngelo1965}.
 
-In the polar F-region ionosphere, GDI is typically considered the dominant structuring process \citep{Weber1984,Cerisier1985,Basu1988,Tsunoda1988}.
+In the polar F-region ionosphere, GDI is typically considered to be the dominant structuring process \citep{Weber1984,Cerisier1985,Basu1988,Tsunoda1988,Fukumoto2000}.
 GDI is operational when high-density perturbations in the plasma move to regions of lower background density and low-density perturbations in the plasma move to regions of higher background density, such that the wave amplitude grows relative to background conditions.
 As plasma drifts and carries irregularities with it, electrons and ions have slightly different drifts and create a perturbed electric field.
-This perturbed electric field combined with the background magnetic field causes the density perturbations to \(\vec{E}\times\vec{B}\) drift.
-If the drift causes high and low density perturbations to move as described above, the instability is operational and the perturbations grow.
+This perturbed electric field combined with the background magnetic field causes the density perturbations to drift in the \(\vec{E}\times\vec{B}\) direction.
+If the drift causes high- and low-density perturbations to move as described above, the instability is operational and the perturbations grow.
 If instead high density perturbations drift into regions of higher background density and low density perturbations drift into regions of lower background density, the wave is damped.
 
-Some of the first 1D investigations of the linear GDI growth rate found that it can be described by a simple expression involving the plasma drift speed, \(V_E = |\vec{E}\times\vec{B}|/B^2\), and the gradient scale length, \(L^{-1} = |(\partial n/\partial x)/n|\), Equation \ref{eqn:gdi_old} \citep{Simon1963,Hoh1963,Linson1970}
+Some of the first 1D investigations of the linear GDI growth rate found that it can be described by the following simple expression involving the plasma drift speed, \(V_E = |\vec{E}\times\vec{B}|/B^2\), and the gradient scale length, \(L^{-1} = |(\partial n/\partial x)/n|\) \citep{Simon1963,Hoh1963,Linson1970}
 \begin{equation}
 	\label{eqn:gdi_old}
-	\gamma = \frac{V_E}{L}
-\end{equation}.
+	\gamma = \frac{V_E}{L}.
+\end{equation}
 This expression is technically only valid if the density gradient, \(\nabla n\), is in the same direction as the plasma drift velocity, \(\vec{V}_E\), and the wave is propagating perpendicular to both.
-A more general expression that is used less often, Equation \ref{eqn:gdi_80s}, does consider an arbitrary wavevector, \(\vec{k}\), but still assumes density gradients parallel to the plasma drift \citep{Tsunoda1988}
+A more general expression that is used less often does consider an arbitrary wavevector, \(\vec{k}\), but still assumes density gradients parallel to the plasma drift \citep{Tsunoda1988}:
 \begin{equation}
 	\label{eqn:gdi_80s}
-	\gamma = \frac{k_y}{k}\frac{\vec{k}\cdot\vec{V}_E}{kL}
-\end{equation}.
+	\gamma = \frac{k_y}{k}\frac{\vec{k}\cdot\vec{V}_E}{kL}.
+\end{equation}
 An even more general consideration involves arbitrary gradient and plasma drift directions \citep{Keskinen1982a,Makarevich2014c}.
-In this more general case, the directional dependence of GDI has been shown to be significant, such that the growth rate can actually change signs (determining whether wave growth or damping occurs) with different wavevector directions, so it is important to consider this factor \citep{Makarevich2014c}.
-In addition, these expressions have historically only applicable in a particular altitudinal regime, usually the F region.
+In this more general case, the directional dependence of GDI has been shown to be significant, such that the growth rate can actually change signs (determining whether wave growth or damping occurs) with different wavevector and gradient directions.
+In addition, these expressions have historically been developed for a particular altitudinal regime, usually in the \(F\) region.
 
-Recently, a new expression has been introduced that considers any arbitrary wavevector, density gradient, and altitude within the ionosphere, which allows three electrostatic instabilities, FBI, GDI, and CCI, to be considered with a single dispersion relation \citep{Makarevich2016a}.
+Recently, a new expression has been introduced that considers any arbitrary wavevector, density gradient, and altitude within the ionosphere, which allows three electrostatic instabilities, FBI, GDI, and CCI, to be considered within a single dispersion relation \citep{Makarevich2016a}
 \begin{equation}
 	\label{eqn:gdi_Mak16}
-	(H_i-H_e)\omega = (H_i\vec{V}_{e0}-H_e\vec{V}_{i0})\cdot\vec{k}+(C_i-C_e)H_eH_i
+	(H_i-H_e)\omega = (H_i\vec{V}_{e0}-H_e\vec{V}_{i0})\cdot\vec{k}+(C_i-C_e)H_eH_i.
 \end{equation}
-As before, \(\vec{k}\) is the wavevector, \(\omega\) is the wave frequency, and \(\vec{V}_{\alpha 0}\) is the background velocity of species \(\alpha\).
-The other quantities in equation \ref{eqn:gdi_Mak16} are not standard and described below.
+Here, \(\vec{k}\) is the wavevector, \(\omega\) is the wave frequency, and \(\vec{V}_{\alpha 0}\) is the background velocity of species \(\alpha\).
+All other quantities in Equation \ref{eqn:gdi_Mak16} are not standard and described below.
 \begin{equation}
+\begin{split}
 	H_\alpha = S_\alpha F_\alpha + D_\alpha^{-1} F_\parallel \quad\quad\quad 
-	C_\alpha = \frac{T_\alpha}{m_\alpha \Omega_\alpha}
-\end{equation}
-\begin{equation}
+	C_\alpha = \frac{T_\alpha}{m_\alpha \Omega_\alpha} \\
 	F_\alpha = i k_\perp^2 D_\alpha + \vec{G}\cdot\vec{k}_\perp D_\alpha + \vec{G}\cdot\vec{k}\times\uvec{b} \quad\quad\quad
-	F_\parallel = i k_\parallel^2 + \vec{G}\cdot\vec{k}_\parallel
-\end{equation}
-\begin{equation}
+	F_\parallel = i k_\parallel^2 + \vec{G}\cdot\vec{k}_\parallel \\
 	S_\alpha = \frac{1}{1+D_\alpha^2} \quad\quad\quad
 	D_\alpha = -\frac{i}{\Omega_\alpha}(\omega-\vec{k}\cdot{V}_{\alpha 0})+r_\alpha
+\end{split}
 \end{equation}
+Throughout this thesis, \(T_\alpha\) and \(m_\alpha\) refer to the temperature and mass of species \(\alpha\), respectively.
+The unit vector \(\uvec{b}\) is in the direction of the magnetic field \(\vec{B}\) and \(\vec{k}_\perp\) is the component of the wavevector \(\vec{k}\) perpendicular to the magnetic field while \(\vec{k}_\parallel\) is the component parallel to it.
+The gradient strength, \(\vec{G}\), is defined by \(\vec{G} = \nabla n/n\) such that the magnitude of the gradient strength is equivalent to the inverse gradient scale length, \(G = L^{-1}\).
 The quantity \(r_\alpha\) is the ratio of the collision frequency to the gyrofrequency of a particular species, given by \(r_\alpha = \nu_\alpha/\Omega_\alpha\).
 
-By taking certain limiting cases, expressions for the instability growth rate in different region can be derived from Equation \ref{eqn:gdi_Mak16}.
-These expressions are sumarized in Table \ref{tab:gdi_exps}.
-In addition to the quantities previously described in relation to Equation \ref{eqn:gdi_Mak16}, the following are useful for the expressions in Table \ref{tab:gdi_exps}.
+By considering certain limiting cases, expressions for the instability growth rate in different regions can be derived from Equation \ref{eqn:gdi_Mak16}.
+These expressions are summarized in Table \ref{tab:gdi_exps}.
+In addition to the quantities previously described in relation to Equation \ref{eqn:gdi_Mak16}, the following definitions are useful for the expressions in Table \ref{tab:gdi_exps}.
 \begin{equation}
@@
 
-For the purposes of this thesis, the cold plasma limit expression in Table \ref{tab:gdi_exps} is most recent and can be rewritten using the definition of \(b\), Equation \ref{eqn:gdi} \citep{Makarevich2014c,Makarevich2016a}.
+For the purposes of this thesis, the cold plasma limit expression in Table \ref{tab:gdi_exps} is most relevant and can be rewritten using the definition of \(b\) \citep{Makarevich2014c,Makarevich2016a}
 \begin{equation}
 	\label{eqn:gdi}
-	\gamma = \frac{1}{1+\psi}\left(\uvec{k}\cdot\uvec{b}\times\vec{G}\right)\uvec{k}\cdot\left(\frac{\vec{E}}{B}-R\vec{V}_E\right)
+	\gamma = \frac{1}{1+\psi}\left(\uvec{k}\cdot\uvec{b}\times\vec{G}\right)\uvec{k}\cdot\left(\frac{\vec{E}}{B}-R\vec{V}_E\right).
 \end{equation}
 Equation \ref{eqn:gdi} is applicable throughout the \(E\) and lower \(F\) regions.
-
-%It is useful to express Equation \ref{eqn:gdi_Mak16} in the following equivalent form by defining additional terms.
-%\begin{equation}
-%	\label{eqn:gdi_Mak16_alt}
-%	\left(D_i-D_e\right)\omega = \vec{V}_d\cdot\vec{k}\left(\frac{Z_2}{Z_1}\right) + \left(D_i\vec{V}_{i0}-D_e\vec{V}_{e0}\right)\cdot\vec{k}+(C_i-C_e)k_{\perp}^2\left(\frac{Z_3}{Z_1}\right)
-%\end{equation}
-%In the \(F\) region where background plasma motion is dominated by electric fields perpendicular to the magnetic field, the differential plasma velocity, \(\vec{V}_d\), can be described as 
-%\begin{equation}
-%	\label{eqn:gdi_Mak16_Vd}
-%	\vec{V}_d = \vec{V}_{e0}-\vec{V}_{i0} = (r_i-r_e)\left[s_es_i(1+\psi)\left(R\vec{V}_E-\frac{\vec{E}}{B}\right)\right]
-%\end{equation}
-%where, \(s_\alpha = (1+r_\alpha^2)^{-1}\).  Additionally,
-%\begin{equation}
-%\label{eqn:gdi_Mak16_Z}
-%\begin{split}
-%	Z_1 = i(1+Y)+a+Rb+ycK \quad\quad\quad 
-%	Z_2 = R(i+a)-b \quad\quad\quad\quad\quad\quad\\
-%	Z_3 = -\frac{\Psi(i+a)^2}{1+\Psi}-Rb(i+a)+\frac{b^2}{1+\Psi}-iy(y-ic)\left[\left(K+1-\Psi^{-1}\right)(i+a)+Rb\left(1-\Psi^{-1}\right)\right]\\+Ky^2(y-ic)^2
-%\end{split}
-%\end{equation}
-%and
-%\begin{equation}
-%\begin{split}
-%	a& = \frac{\vec{G}\cdot\vec{k}_\perp}{k_\perp^2} \quad\quad\quad
-%	b = -\frac{\vec{G}\cdot\vec{k}\times\uvec{b}}{k_\perp^2} \quad\quad\quad
-%	c = \frac{\vec{G}\cdot\vec{k}_\parallel}{k_\perp k_\parallel} \\
-%	\Psi = -D_iD_e& \quad\quad
-%	R = \frac{D_i+D_e}{1+\Psi} \quad\quad
-%	y = \frac{k_\parallel}{k_\perp} \quad\quad
-%	K = \left(1+\frac{1}{\Psi}\right)\left(1+R^2\right) \quad\quad
-%	Y = Ky^2
-%\end{split}
-%\end{equation}.
-
-%By considering certain limiting cases, an expression that is more relevant to the polar F region specifically can be resolved from Equation \ref{eqn:gdi_Mak16_alt}.
-First, assume a cold plasma such that the temperature of both ions and electrons is negligible.
-In this case, \(T_\alpha = 0\), so \(C_\alpha = 0\).
-Secondly assume all irregularities are perfectly aligned with the magnetic field such that \(k_\parallel = 0\).
-This means that \(\vec{k}_\perp = \vec{k}\), \(y=0\), and \(Y=0\).
-Finally, consider only low frequencies (\(\omega \ll \Omega_\alpha\)) and long wavelengths (\(\vec{k}\cdot\vec{V}_{\alpha 0} \gg \nu_\alpha\)).
-If \(D_\alpha\) is expressed as shown below, it can bee seen that these limits result in the first two terms being negligible such that \(D_\alpha \approx r_\alpha\).
-%\begin{equation}
-%	D_\alpha = -i\frac{\omega}{\Omega_\alpha} + i\frac{\vec{k}\cdot\vec{V}_{\alpha 0}}{\Omega_\alpha} + \frac{\nu_\alpha}{\Omega_\alpha} \approx \frac{\nu_\alpha}{\Omega_\alpha} = r_\alpha
-%\end{equation}
-%Equation \ref{eqn:gdi_Mak16_alt} can then be expressed as follows.
-%\begin{equation}
-%	\label{eqn:gdi_Mak16_Freg}
-%	(r_i-r_e)\omega = \vec{V}_d\cdot\vec{k}\left(\frac{Z_2}{Z_1}\right)+(r_i\vec{V}_{i0}-r_e\vec{V}_{e0})
-%\end{equation}
-%The growth rate, \(\gamma\), is contained within the dispersion relation as \(\omega = \omega_r+i\gamma\), so the imaginary part of Equation \ref{eqn:gdi_Mak16_Freg} must be considered separately.
-%\begin{equation}
-%	\label{eqn:gdi_Mak16_gam}
-%	(r_i-r_e)\gamma = \vec{V}_d\cdot\vec{k}\;\text{Im}\left\{\frac{Z_2}{Z_1}\right\}
-%\end{equation}
-%The differential plasma drift has been defined previously in Equation \ref{eqn:gdi_Mak16_Vd} but the imaginary component of \(Z_2/Z_1\) must still be identified.
-Using the definitions of \(Z_1\) and \(Z_2\) provided in Equation \ref{eqn:gdi_Mak16_Z} and the simplifying assumptions discussed above for the F region,
-%\begin{equation}
-%	\frac{Z_2}{Z_1} = \frac{R(i+a)-b}{i+a+Rb}
-%\end{equation}.
-%Multiplying both the numerator and denominator by \(-i+a+Rb\) removes the imaginary component from the denominator.
-Additionally the local approximation \(G \ll k_\perp\) is taken, which results in \(a\), \(b\), and \(c\) being small so that any second order terms in these quantities can be neglected.
-%\begin{equation}
-%	\frac{Z_2}{Z_1} = R+ib(1+R^2)
-%\end{equation}
-%Using the above result and Equation \ref{eqn:gdi_Mak16_Vd}, Equation \ref{eqn:gdi_Mak16_gam} becomes
-%\begin{equation}
-%	(r_i-r_e)\gamma = (r_i-r_e)s_es_i\left(1+\psi\right)\left(R\vec{V}_E-\frac{E}{B}\right)\cdot\vec{k}b\left(1+R^2\right)
-%\end{equation}
-%Recognizing that \(s_es_i\left(1+R^2\right) = (1+\psi)^{-2}\) and \(b = \vec{G}\cdot\vec{k}\times\uvec{b}/k^2 = -\vec{k}\cdot\uvec{b}\times\vec{G}/k^2\) by vector identities, a final expression for the GDI growth rate in the F region can be found, which agrees well with the results of \citet{Makarevich2014c}
-%\begin{equation}
-%	\label{eqn:gdi}
-%	\gamma = \frac{1}{1+\psi}\left(\uvec{k}\cdot\uvec{b}\times\vec{G}\right)\uvec{k}\cdot\left(\frac{\vec{E}}{B}-R\vec{V}_E\right)
-%\end{equation}.
-%Equation \ref{eqn:gdi} will be critical to the analysis done both in Chapter \ref{sec:paper2} and Chapter \ref{sec:paper3} in this thesis.
 
@@
 
-Although linear GDI theory as described predicts clear dependence on wave vector direction \citep{Makarevich2014c}, this theory is only really valid in the long-wavelength limit, \(\vec{k}\cdot\vec{V}_E \ll \nu_\alpha\).
-In the \(F\) region, the ion-neutral collision frequency, \(\nu_i\), is about 10 Hz and if the drift speed, \(V_E\), is about 1000 m/s, fluid theory only applies for wavelengths greater than 100 m, larger than the decameter-scale waves that are detected by SuperDARN.
-This raises the important question of whether or not the same anisotropy that is predicted by linear GDI theory at larger scales can be expected in small-scale irregularities where nonlinear processes may be important.
+Although linear GDI theory as described above predicts clear dependence on wavevector direction \citep{Makarevich2014c}, this theory is only valid in the long-wavelength limit, \(\vec{k}\cdot\vec{V}_E \ll \nu_\alpha\).
+In the \(F\) region, the ion-neutral collision frequency, \(\nu_i\), is about 10 Hz and if the drift speed, \(V_E\), is about 1000 m/s, fluid theory only applies to wavelengths greater than 100 m, larger than the decameter-scale waves that are detected by SuperDARN.
+This raises the important question of whether or not the same anisotropy that is predicted by linear GDI theory at larger scales can be expected in small-scale irregularities where other processes (not described by the linear fluid theory) may be important.
 This issue has mostly been studied through numerical simulations due to the challenges associated with measuring a high-resolution 2D spectrum experimentally.
 
-The first simulations of GDI were done in the 1970s and successfully reproduced structures developing on the trailing edge of large density gradients \citep{Zabusky1973,Doles1976,Scannapieco1976,Ossakow1975,Ossakow1977}.
+The first simulations of GDI were done in the 1970s; they have successfully reproduced structures developing on the trailing edge of large density gradients \citep{Zabusky1973,Doles1976,Scannapieco1976,Ossakow1975,Ossakow1977}.
 Later, 2D simulations were expanded to include magnetosphere-ionosphere coupling, but it was found that this only had a substantial effect for very large-scale structures \citep{Keskinen1990}.
 3D simulations were introduced of GDI surrounding a large plasma density enhancement \citep{Guzdar1998} and were later improved by including plasma dynamics parallel to the magnetic field and inertial effects so that secondary KHI and tertiary shear-driven instability processes could be modeled \citep{Gondarenko1999}.
-These simulations consistently showed asymmetry between the leading and trailing edge of large-scale density structures and fluctuations in plasma density that were as much as 10--20\% of the background plasma \citep{Gondarenko2004a,Gondarenko2004b}.
+These simulations consistently showed asymmetry between the leading and trailing edge of large-scale density structures and fluctuations in plasma density that were as much as 10\%--20\% of the background plasma \citep{Gondarenko2004a,Gondarenko2004b}.
 Plasma structuring can occur on the leading edge of large-scale density enhancements if the convection velocity changed rapidly such that the trailing edge becomes the leading edge and vice versa, or if there are large velocity shears initially present \citep{Gondarenko2004a,Gondarenko2004b}.
@@
 For the most part, the results of nonlinear numerical simulations very closely match expectations from linear GDI theory.
-The spatial power spectra are anisotropic with most of the power concentrated in the direction where linear GDI growth is most favorable \citep{Keskinen1981a,Keskinen1981b,Keskinen1982a,Gondarenko2001,Gondarenko2004b}.
-However, very strong shears or increasing the impact of ion-inertial factors can cause isotropy \citep{Gondarenko2001,Gondarenko2006}.
+In particular, it was found that the spatial power spectra are anisotropic with most of the power concentrated in the direction where linear GDI growth is most favorable \citep{Keskinen1981a,Keskinen1981b,Keskinen1982a,Gondarenko2001,Gondarenko2004b}.
+However, very large shears or stronger ion-inertial factors can cause isotropy \citep{Gondarenko2001,Gondarenko2006}.
 The power spectra derived from these simulations are similar to those from experimental observations \citep{Baker1978,Kelley1979}, however it is computationally challenging to run simulations down to decameter scales and observations are generally limited to rocket and satellite techniques, which can only be used to calculate power spectra in one dimension \citep{Villain1986,Moen2012}.
@@
 
-Polar patch occurrence varies both diurnally and seasonally, with a diurnal peak at magnetic noon and a seasonal peak during equinox months \citep{Rodger1996}.
+Polar patch occurrence varies both diurnally and seasonally, with a diurnal peak at magnetic noon and a seasonal peak during equinoctial months \citep{Rodger1996}.
 Additionally, there is a bias towards patches occurring when there is a negative IMF \(B_z\) component \citep{Buchau1983,Rodger1996}.
 The frequency of patch occurrence is greatest when the cusp is slightly daywards of the terminator.
-This creates the situation of a dark polar cap, so background plasma density is low and blobs of enhanced density are easily visible \citep{Coley1998}.
+This creates a dark polar cap, so background plasma density is low and blobs of enhanced density are easily visible \citep{Coley1998}.
 In addition, the cusp then provides a gateway of highly ionized sunlit plasma into the dark polar cap.
 Patches convect from the dayside through the cusp/throat region towards the polar cap \citep{Kelly1984,Foster1984,Foster1985,Foster1993,Sojka1982,delaBeaujardiere1985}.
+A comprehensive review of polar patch observations has been provided by \citet{Crowley1996}.
 
@@
 
-Over time, the trailing edge of a patch became steeper than the leading edge and irregularities developed on the trailing edge.
+%Over time, the trailing edge of a patch became steeper than the leading edge and irregularities developed on the trailing edge.
 
-\citet{Weber1981} identified sun-aligned arcs that extend  along the sun-earth line through large parts of the polar cap using all-sky imaging photometers (ASIP).
+%\citet{Weber1981} identified sun-aligned arcs that extend  along the sun-earth line through large parts of the polar cap using all-sky imaging photometers (ASIP).
 These arcs are over 1000 km long and \(\sim100\) km wide and are produced by soft electron precipitation.
 
-Additionally, \citet{Buchau1983} found that polar patches were more likely to be observed for southward IMF \(B_z\) while sun-aligned arcs were more common for northwards \(B_z\).
+%Additionally, \citet{Buchau1983} found that polar patches were more likely to be observed for southward IMF \(B_z\) while sun-aligned arcs were more common for northwards \(B_z\).
 
@@
 
-Patches are not colocated with enhanced particle precipitation within the polar cap, indicating that this is not the source of the density enhancement \citep{Weber1984}.
+%Patches are not colocated with enhanced particle precipitation within the polar cap, indicating that this is not the source of the density enhancement \citep{Weber1984}.
 Conversely, \citet{Buchau1985} found that plasma densities within patches were comparable with those measured at dayside sub-auroral latitudes, suggesting that dayside solar-illuminated plasma is the source of dense plasma that form polar patches.
 
-The theoretical mechanism for patch formation via transient magnetic reconnection has been described previously, Section \ref{sec:patch_formation}.
+The theoretical mechanism for patch formation via transient magnetic reconnection has been described previously, Section \ref{sec:lit_patches}.
 A variety of experimental evidence has been presented to support this mechanism \citep{Cowley1998,Carlson2002}.
-Experiments run in the 1980s using the European Incoherent Scatter Scientific Association (EISCAT) Troms\o, Norway ISR showed that plasma flow occurred near noon shortly after southward IMF \(B_z\) arrives at the magnetopause \citep{Etemadi1988,Todd1988}.
+Experiments run in the 1980s using the European Incoherent Scatter (EISCAT) radar facility in Troms\o, Norway showed that plasma flow occurred near noon shortly after southward IMF \(B_z\) arrives at the magnetopause \citep{Etemadi1988,Todd1988}.
 These plasma flows are pulsed in the region of the polar cap boundary \citep{Lockwood1993a,Lockwood1993b}, as predicted by \citet{Cowley1991}.
-Observations made by an ASIP in Svalbard in 1984 are fully consistent with patch formation through transient magnetopause reconnection \citep{Carlson1996,Carlson2002}.
-\citet{Carlson2004} identified five signatures of transient magnetic reconnection and then examined a patch formation event for these signatures using the EISCAT Svalbard Radar.
+Observations made by an all sky imaging photometer (ASIP) in Svalbard in 1984 were fully consistent with patch formation through transient magnetopause reconnection \citep{Carlson1996,Carlson2002}.
+\citet{Carlson2004} identified five signatures of transient magnetic reconnection and then examined a patch formation event for these signatures using the EISCAT Svalbard radar.
 All five signatures were observed as expected, strengthening the idea that transient magnetopause reconnection is responsible for patch formation.
 Furthermore, \citet{Carlson2006} identified and tracked a series of patches directly from the subauroral plasma reservoir and show that the boundary moves equatorwards before relaxing poleward, one of the main predictions of the transient magnetopause reconnection mechanism \citep{Lockwood1992b}.
-A review of historical observations of polar patches has been provided by \citet{Crowley1996}.
 
-Recent advancements have improved the ability to image density structures in the ionosphere, particularly using multi-instrument approaches.
-\citet{Semeter2009} introduced a method by which the density pattern in a particular volume could be imaged in three dimensions using AMISR systems.
+Besides polar cap patch formation and evolution, much of the previous experimental effort focused on small-scale structuring processes near polar patches.
+One of the key observations in this regard was that, over time, the trailing edge of a patch became steeper than the leading edge and irregularities developed on the trailing edge \citep{Weber1984}.
+Backscatter power from HF radars also tended to be greater on the trailing edges of both moving polar patches \citep{Milan2002b} and sun-aligned arcs identified from ASIP data \citep{Koustov2012}, which is generally attributed to GDI being unstable on the trailing edge but stable on the leading edge of such large, moving density structures.
+Studies of patches using all-sky airglow imagers (ASI) found that the density gradient on the leading edge of polar patches tends to be 2--3 times steeper than that on the trailing edge \citep{Hosokawa2016}.
+Additionally, large finger-like structures on the trailing edge of patches have been identified that were tens to hundreds of kilometers in size and agreed with predictions made by GDI simulations \citep{Gondarenko2004b}.
+\citet{Moen2012} made direct measurements of plasma density structuring using the ICI-2 sounding rocket.
+This study provided evidence that decameter-scale plasma structuring (the same as observed by HF radars) had spawned from kilometer-scale density gradients (observed by ASIP and ASI methods) and structuring was greatest where GDI was operational according to estimates of the growth rate.
+
+Recent experimental advancements have improved the ability to image density structures in the ionosphere, particularly using multi-instrument approaches.
+\citet{Semeter2009} introduced a method by which the density pattern in a particular volume could be imaged in 3D using AMISR systems.
 This technique was later used by \citet{Dahlgren2012a,Dahlgren2012b} to image a polar patch.
-\citet{Dahlgren2012b} additionally used ASIPs and SuperDARN radars to investigate the polar patch and found that backscatter tended to be observed more on the trailing edge of the patch.
+\citet{Dahlgren2012b} additionally used ASIPs and SuperDARN radars to observe the polar patch and found that backscatter tended to be observed more on the trailing edge of the patch.
 This agrees well with many studies that have examined the asymmetry of small-scale plasma structuring surrounding plasma patches.
 
-Over time, the trailing edge of a patch became steeper than the leading edge and irregularities developed on the trailing edge \citep{Weber1984}.
-Backscatter power from HF radars tends to be greater on the trailing edges of both moving high-density polar structures \citep{Milan2002b} and sun-aligned arcs identified from ASIP data \citep{Koustov2012}, which is generally attributed to GDI being unstable on the trailing edge but stable on the leading edge of such large, moving density structures.
-Studies of patches using all-sky airglow imagers (ASI) found that the density gradient on the leading edge of polar patches tends to be 2--3 times steeper than that on the trailing edge \citep{Hosokawa2016}.
-Additionally, large finger-like structures on the trailing edge of patches have been identified that were 10s--100s of kilometers in size and agree with predictions made by GDI simulations \citep{Gondarenko2004b}.
-\citet{Moen2012} made direct measurements of plasma density structuring using the ICI-2 sounding rocket.
-This study provided evidence that decameter scale plasma structuring (the same as observed by HF radars) had spawned from kilometer scale density gradients (that observed by ASIP and ASI methods) and structuring was greatest where GDI was operational.
+Even though observations of plasma irregularities near polar patches present certain advantages, small-scale irregularities may be present in the polar cap regardless of polar patch presence.
+Plasma irregularities at decameter scales are typically studied with HF coherent scatter radars, Section \ref{sec:csr}.
+The factors that affect irregularity observation can roughly be divided into two groups: irregularity production and radar propagation.
+Irregularity production factors are related to the existence of plasma irregularities in the ionosphere, while radar propagation factors are related to the ability to observe irregularities with a ground-based radar, Section \ref{sec:superdarn}.
 
-Small-scale plasma irregularities are typically studied with HF coherent scatter radars in the polar cap, Section \ref{sec:csr}.
-The factors that affect irregularity observation can roughly be divided into two groups, irregularity production and radar propagation.
-Irregularity production factors are related to the existence of plasma irregularities in the ionosphere while radar propagation factors are related to the ability to observe irregularities with a ground based radar, Section \ref{sec:superdarn}.
-Within the polar caps, plasma irregularity production is much higher than at lower altitudes such that when background plasma conditions are favorable for radar propagation, FAIs are observed nearly continuously \citep{Bristow2011}.
-However, there are still definite diurnal, seasonal, and solar cycle trends \citep{Kane2012}.
+Recent radar observations suggested that in the polar cap, plasma irregularity production may be much higher than at lower latitudes such that when background plasma conditions are favorable for radar propagation, FAIs are observed nearly continuously \citep{Bristow2011}.
+However, there are still definite diurnal, seasonal, and solar cycle trends as far as radar observations at lower, auroral latitudes are concerned \citep{Kane2012}.
 There is an increase in echo occurrence at solar maximum, particularly in the midnight sector \citep{Milan1997,Koustov2004}.
-Seasonally, high latitude radars observe a peak in echo occurrence during equinox months \citep{Koustov2004}.
+Seasonally, high-latitude radars observe a peak in echo occurrence during equinoctial months \citep{Koustov2004}.
+
 Additionally, backscatter tends to occur at closer ranges in summer at midnight, but moves further away from the radar in other seasons at midday \citep{Milan1997}.
-In general, echo occurrence in high during the day and low at night, but nighttime occurrence is directly proportional to the \(F\) peak density and daytime occurrence can be significantly reduced by a dense and conductive \(E\) layer.
-\citep{Koustov2004,Kane2012,Vickrey1982}.
+In general, echo occurrence is high during the day and low at night, but nighttime occurrence is directly proportional to the \(F\) peak density and daytime occurrence can be significantly reduced by a dense and conductive \(E\) layer \citep{Koustov2004,Kane2012}.
 In the daytime, solar illumination causes the electron density of the \(E\) region to increase dramatically, making it highly conductive.
@@
 
-Irregularity production is impacted primarily by the presence of strong density gradients and electric fields in the ionosphere \citep{Koustov2004}.
-Properly oriented density gradients and electric fields are a necessary condition for GDI to be operational, Section \ref{sec:lit_instbilities}, so the stronger these are, the more likely the irregularity production rate is high.
-Both backscatter power and occurrence increase with increasing electric field strength \citep{Ballatore2001,Danskin2002,Makarevich2014b}.
-Similarly, backscatter occurrence increases when the IMF \(B_z\) component is strongly negative, implying a strong merging electric field.
+The two most important production factors for irregularities produced through GDI are density gradients and convection velocity or, equivalently, convection electric fields, Section \ref{sec:lit_instabilities}.
+Properly-oriented density gradients and electric fields are a necessary condition for GDI to be operational, Section \ref{sec:lit_instabilities}, and the stronger these are, the more likely the irregularity production rate will be high.
+Experimental studies focused on the latter factor, with both backscatter power and occurrence showing an increase with increasing electric field strength \citep{Danskin2002,Makarevich2014b}.
 
-The primary radar propagation factor to consider is the background plasma density and how much the radar beam is refracted.
-If the background density is too low, the beam is under refracted and is never perpendicular to the magnetic field, however if the density is too high, the beam is over refracted, such that observation of small-scale FAIs is most favorable if the background density is between \(1\times 10^{11}\) m\(^{-3}\) and \(4\times 10^{11}\) m\(^{-3}\) \citep{Danskin2002,Makarevich2014b}.
+Amongst the propagation factors, arguably the primary one is the background plasma density and the associated amount of the radar beam refraction.
+If the background density is too low, the beam is under refracted and is never perpendicular to the magnetic field, however if the density is too high, the beam is over refracted.
+Previous studies show that observation of small-scale FAIs is most favorable if the background density is between \(1\times 10^{11}\) m\(^{-3}\) and \(4\times 10^{11}\) m\(^{-3}\) \citep{Danskin2002,Makarevich2014b}.
 If the \(D\) region is too dense, this can also in theory attenuate the signal and reduce the rate of returned backscatter, although observations have shown this is not a major factor in echo occurrence \citep{Danskin2002}.
-Finally, the radar equation predicts that the returned backscatter should be proportional to the perturbation density, \(\delta n\), squared, or, equivalently, the product of the background density and the fractional density, \(\delta n/n\), squared, Equation \ref{eqn:fractional_density}
-\begin{equation}
-	\label{eqn:fractional_density}
-	P \propto \delta n^2 \propto \left(\frac{\delta n}{n}\right)^2 n^2
-\end{equation}.
-There is no general agreement of what controls the fractional density, however, if it is approximately constant, then radar backscatter power could be directly proportional to \(n^2\).
+
+Generally, both production and propagation factors are important in determining when irregularities are observed by ground-based HF radars, so it is important to distinguish between the two to interpret data correctly.
+Additionally, the relationship between plasma irregularities of different scales is important to understand how small-scale irregularities develop in the context of larger ionospheric dynamics.
 
@@
 In addition to being of purely scientific interest, this structuring creates complications when receiving radio signals through the ionosphere.
-Because the structuring is highly irregular and varies on many spatial and temporal scales, unpredictable wave refraction, phase shifts, and amplitude attenuation can occur, making signals difficult to detect.
+Because the structuring is highly irregular and varies on many spatial and temporal scales, unpredictable wave refraction, phase shifts, and amplitude attenuation can occur, making radio signals difficult to detect.
 This introduces problems in any system that involves ground-to-satellite, satellite-to-ground, or over-the-horizon communication, such as navigation or communication.
@@
 In addition, understanding plasma structuring on a range of scales is important for correctly interpreting backscatter from ionospheric radars.
-In particular, one of the main purposes of SuperDARN is to create large scale convection maps of both polar caps.
-These convection maps have been used to give context of both magnetosphere and thermosphere behavior and are widely used in space physics research.
+In particular, one of the main purposes of SuperDARN is to create large-scale convection maps of both hemispheres including polar cap regions.
+These convection maps have been used to give context of both the magnetosphere and thermosphere and are widely used in space physics research.
 To map convection accurately, it is extremely important to understand when backscatter is observed and when it is absent, and why this is so.
-Understanding plasma instability and wave growth mechanisms is very important for answering these kinds of questions.
-Models and simulations of structuring in the polar caps incorporate structuring mechanisms so it is helpful if these are better understood, both the operation of a particular mechanism as well as how different mechanism are interconnected and which ones may be dominant under particular conditions.
+Understanding plasma instability and wave growth mechanisms is very important for answering these questions.
+Furthermore, models and simulations of ionospheric dynamics in the polar caps often incorporate plasma structuring details, such as instability growth mechanism.
+Although GDI is often considered the dominant instability mechanism in the \(F\) region, it is helpful to understand which other mechanisms may be operational and how they are interconnected.
 
 The aim of this work is to investigate small-scale plasma irregularity production in the \(F\)-region polar cap.
-In particular, factors such as direct and indirect global solar control and GDI directional dependance will be considered.
-The majority of observational data has been found from ground based ionospheric radars, both CSR systems (Section \ref{sec:csr}) and ISR systems (Section \ref{sec:isr}), with particular emphasis on radars within the SuperDARN network, Section \ref{sec:superdarn}.
-However, these datasets have also been supplemented with information from models as well as satellite data when necessary.
-Results are investigated in the context of a linear fluid theory analysis of ionospheric plasma.
+Both global and local control of irregularity production is studied, i.e.
+we consider global solar control through solar illumination and solar wind ans well as much more local control by plasma density gradients and convection electric field.
+The majority of observational data have been measured by ground-based ionospheric radars, both CSR systems (Section \ref{sec:csr}) and ISR systems (Section \ref{sec:isr}), with particular emphasis on radars within the SuperDARN network, Section \ref{sec:superdarn}.
+These datasets have also been complimented with information from models as well as satellite data when necessary.
+Experimental results are considered in the context of a linear fluid theory of the gradient-drift instability.
 
-The first objective of is to evaluate the extent of solar control over irregularity production in the polar \(F\) region.
-The occurrence of radar backscatter will be compared with both solar illumination (considered direct solar control) and IMF factors (considered indirect solar control).
-This gives a macrophysical view of where and when plasma irregularities are most likely to occur.
+The first objective of this body of work is to evaluate the extent of solar control over irregularity production in the polar \(F\) region.
+The occurrence of radar backscatter is considered in the contex both solar illumination (considered direct solar control) and IMF factors (considered indirect solar control).
+This gives a global or macrophysical view of where and when plasma irregularities are most likely to occur.
 
@@
 The model considers a large, elongated polar patch drifting in an arbitrary direction relative to a radar and assumes irregularity production is directly dependent to the linear GDI growth rate.
-High linear growth rates predicted by the model are compared with classic observations of stronger backscatter power trailing drifting polar patches.
+The model results are expected to contribute to more accurate interpretation of observations of irregularities near polar patches.
 
-The third objective is to investigate directional dependancies of small-scale plasma irregularities.
-Two dimensional measurements of density gradients and electric fields are made such that the directionally dependent GDI growth rate can be calculated, which is then compared with plasma irregularity power and occurrence in a particular location.
-This gives insight into whether the production of small-scale irregularities is directionally dependent, similar to the linear GDI growth rate, or if the distribution is isotropic due to other nonlinear processes.
+The third objective is to investigate directional dependencies of small-scale plasma irregularities produced through the gradient-drift instability operating in the linear regiem.
+Two-dimensional measurements of density gradients and electric fields are made such that the directionally-dependent GDI growth rate can be calculated, which is then compared with plasma irregularity power and occurrence in a particular location.
+This gives insight into whether the production of small-scale irregularities is directionally dependent, similar to the linear GDI growth rate, or if the distribution with the propagation direction is isotropic due to other processes.
 
 The outline of this thesis is as follows.
-Chapter \ref{sec:introduction} has provided an overview of the polar ionosphere and both macro- and microphysical plasma structuring mechanisms.
+Chapter \ref{sec:introduction} provides an overview of the polar ionosphere and both plasma structuring mechanisms at both large and small scales.
 Chapter \ref{sec:paper1} presents an experimental investigation of decameter-scale plasma irregularities in the polar F-region and their solar control, with a particular focus on statistical analysis of data collected with a recently-deployed SuperDARN radar in the southern polar cap.
 The same results were also presented in a recent journal article \citep{Lamarche2015}.
-Chapter \ref{sec:paper2} presents a modeling study of the linear GDI growth rate surrounding a large, elongated polar patch drifting in an arbitrary direction within the FoV of a SuperDARN radar uses the results to provide insight as to optimal conditions from radars to observe irregularities.
+Chapter \ref{sec:paper2} presents a modeling study of the linear GDI growth rate of plasma irregularities near a large, elongated polar patch drifting in an arbitrary direction within the viewing area of an oblique-scanning radar such as SuperDARN.
 These results have also been recently published \citep{Lamarche2016}.
-Chapter \ref{sec:paper3} presents an experimental investigation of the directional dependence of small-scale plasma structures by considering simultaneous measurements of electric field and density gradients from an AMISR  system and FAIs observed by SuperDARN.
-Particular emphasis is placed on if small-scale structures are directionally dependent similar to the predictions of linear GDI theory.
+Chapter \ref{sec:paper3} presents an experimental investigation of the directional dependence of small-scale plasma structures by considering simultaneous measurements of electric field and density gradients from an AMISR  system and FAIs observed by SuperDARN in the context of predictions made by linear GDI theory and the model results presented in Chapter \ref{sec:paper3}.
 The same results have been submitted for publication \citep{Lamarche2017}.

### 2017-01-29 paper1.tex
Formatting updated

@@
 \section{Introduction}
-\label{Sec:Intro}
+\label{sec:p1intro}
 
@@
 \section{Experiment Configuration and Data Processing}
-\label{Sec:Exp}
+\label{sec:p1experiment}
 
@@
 \section{Observations}
-\label{Sec:Obs}
+\label{sec:p1observations}
 
@@
 \begin{figure}
-\includegraphics[width=12.5 cm,angle=90]{p1f1.pdf}
-\caption{Monthly average plots of (a--d) echo occurrence and (e--h) peak electron density for 2010.
+	\centering
+	\includegraphics[height=\textwidth,angle=90]{p1f1.pdf}
+	\caption[Monthly average echo occurrence and electron density]{Monthly average plots of (a--d) echo occurrence and (e--h) peak electron density for 2010.
 Four months are selected (one for each season) to show seasonal variations.
@@
 To further characterize this diurnal variation, a cosine function of the form
-\begin{equation}\label{E:Model}
+\begin{equation}\label{eqn:band}
 R\left(t\right) = A \cos\left[2\pi\left(t-B\right)/24\right] + C
@@
 \begin{figure}
-\includegraphics[width=11.0 cm,angle=90]{p1f2.pdf}
-\caption{The variations in the echo band (a) amplitude of the fitted cosine function \(A\), (b) time shift or time of day of the peak \(B\), (c) average position of the band or the vertical shift of the cosine function from zero \(C\), and (d) width (BW).
+	\centering
+	\includegraphics[height=\textwidth,angle=90]{p1f2.pdf}
+	\caption[Trends in high-occurrence band]{The variations in the echo band (a) amplitude of the fitted cosine function \(A\), (b) time shift or time of day of the peak \(B\), (c) average position of the band or the vertical shift of the cosine function from zero \(C\), and (d) width (BW).
 Different lines are for different years with the legend given in Figure \ref{fig:year_line}a.
@@
 \begin{figure}
-\includegraphics[width=12.0 cm,angle=90]{p1f3.pdf}
-\caption{Contour plots of the (a--d) echo band position \(r(t)\) and (e)--(h) occurrence versus UT and season.
+	\centering
+	\includegraphics[height=\textwidth,angle=90]{p1f3.pdf}
+	\caption[Solar cycle trends of high-occurrence band]{Contour plots of the (a--d) echo band position \(r(t)\) and (e)--(h) occurrence versus UT and season.
 Each row shows the data for a different year as given in the top right corner.
@@
 SZA indicates how sunlit the ionosphere is and the associated degree of ionization, i.e.
-the \(E\)- and \(F\)-region electron densities, as explained in Section \ref{Sec:Intro}.
-The model \(N_mE\) and \(N_mF_2\) values have been employed for this analysis, Section \ref{Sec:Exp}.
+the \(E\)- and \(F\)-region electron densities, as explained in Section \ref{sec:p1intro}.
+The model \(N_mE\) and \(N_mF_2\) values have been employed for this analysis, Section \ref{sec:p1experiment}.
 To keep model run times at a manageable level, here we consider 10-min cadence data including the average echo occurrence within the band (in Figures \ref{fig:month_avg_occ}--\ref{fig:year_color}, the 2-min data were considered).
@@
 \begin{figure}
-\includegraphics[width=17.0 cm]{p1f4.pdf}
-\caption{The echo band range versus the (a) solar zenith angle (SZA) and (b) model peak F2 electron density (\(N_mF_2\)).
+	\centering
+	\includegraphics[width=\textwidth]{p1f4.pdf}
+	\caption[Band position and occurrence versus SZA and \(N_mF_2\)]{The echo band range versus the (a) solar zenith angle (SZA) and (b) model peak F2 electron density (\(N_mF_2\)).
 Each point is color-coded in model \(E\) region density.
@@
 In the following analysis, we attempt to determine whether there is a direct relationship between IMF behavior and echo occurrence on the scale of a few hours.
-We compare the MCM echo occurrence variations with the simultaneous IMF observations using a method analogous to that employed by \citet{Makarevich2012} in their study of auroral \(E\)-region irregularities (see also our Section \ref{Sec:Exp} for a brief description of IMF data analysis).
+We compare the MCM echo occurrence variations with the simultaneous IMF observations using a method analogous to that employed by \citet{Makarevich2012} in their study of auroral \(E\)-region irregularities (see also our Section \ref{sec:p1experiment} for a brief description of IMF data analysis).
 In addition, we evaluate a dependence on the PCS index using the same approach.
@@
 \begin{figure}
-\includegraphics[width=15.0 cm]{p1f5.pdf}
-\caption{The MCM echo occurrence versus UT color coded in the average (a) IMF By, (b) IMF Bz, and (c) PCS values.}
+	\centering
+	\includegraphics[width=\textwidth]{p1f5.pdf}
+	\caption[Occurrence versus UT and IMF conditions]{The MCM echo occurrence versus UT color coded in the average (a) IMF By, (b) IMF Bz, and (c) PCS values.}
 	\label{fig:ut_occ}
@@
 \section{Discussion}
+\label{sec:p1discussion}
 
@@
 One possible reason for this difference is more favorable propagation conditions in the polar cap with greater chances to reach orthogonality with magnetic field, which is a necessary condition for Bragg scatter to be received by coherent scatter radars.
-Whether any observed occurrence trend is due to changing propagation or production is difficult to resolve based on HF radar echo occurrence observations alone, as explained in Section \ref{Sec:Intro}.
+Whether any observed occurrence trend is due to changing propagation or production is difficult to resolve based on HF radar echo occurrence observations alone, as explained in Section \ref{sec:p1intro}.
 A reasonable assumption often utilized is that propagation will be mostly responsible for changing location of echo occurrence.
@@
 The two important outcomes of the current study related to this indirect density analysis were (1) development of an empirical model to describe location of \(F\)-region backscatter and (2) analysis of characteristics derived from the model.
-The \(F\)-region backscatter was predominantly observed in the form of a band that was well described by a simple model, Equation (\ref{E:Model}).
+The \(F\)-region backscatter was predominantly observed in the form of a band that was well described by a simple model, Equation (\ref{eqn:band}).
 The \(F\) region echo band exhibited systematic diurnal and seasonal variations, Figure \ref{fig:month_avg_occ}, which was reflected in systematic changes of the fitted band characteristics \(A\), \(B\), \(C\), and band width (BW), Figure \ref{fig:year_line}.
@@
 \begin{figure}
-\includegraphics[width=18.0 cm]{p1f6.pdf}
-\caption{Conjugate measurements for two selected events on (a--d) March 14, 2010 and (e-h) April 6, 2010.
+	\centering
+	\includegraphics[width=\textwidth]{p1f6.pdf}
+	\caption[Cojugate patchy day example]{Conjugate measurements for two selected events on (a--d) March 14, 2010 and (e-h) April 6, 2010.
 Shown are the IMF By (blue) and Bz (red) components (Figures \ref{fig:patchy_example}a and \ref{fig:patchy_example}e), MCM line-of-sight velocity versus UT and slant range in beam 14 (Figures \ref{fig:patchy_example}b and \ref{fig:patchy_example}f), RKN velocity in beam 4 (Figure \ref{fig:patchy_example}c and \ref{fig:patchy_example}g), and RISR-N electron density versus UT and altitude (Figure \ref{fig:patchy_example}d and \ref{fig:patchy_example}h).
@@
 Figure \ref{fig:patchy_example}b shows an echo band that fits well with the average pattern of Figure \ref{fig:month_avg_occ}a (note more limited ranges for Figure \ref{fig:patchy_example}).
-This is a clear example of persistent backscatter at MCM where echoes are present throughout the entire day with only range changing in agreement with the model of Equation (\ref{E:Model}).
+This is a clear example of persistent backscatter at MCM where echoes are present throughout the entire day with only range changing in agreement with the model of Equation (\ref{eqn:band}).
 In contrast, Figure \ref{fig:patchy_example}f shows an example of ``patchy backscatter'' at MCM, particularly after 12 UT.
@@
 \begin{figure}
-\includegraphics[width=12.0 cm,angle=90]{p1f7.pdf}
-\caption{Contour plots of the (a) average occurrence within the band, (b) band range, (c) model peak F-region electron density (\(N_mF_2\)) and (d) solar zenith angle (SZA) versus UT and month in 2010.
+	\centering
+	\includegraphics[height=\textwidth,angle=90]{p1f7.pdf}
+	\caption[Summary of annual and diurnal variations]{Contour plots of the (a) average occurrence within the band, (b) band range, (c) model peak F-region electron density (\(N_mF_2\)) and (d) solar zenith angle (SZA) versus UT and month in 2010.
 Color bars are given to the right.
@@
 \section{Summary and Conclusions}
+\label{sec:p1conclusions}
 
@@
 
-1.
-Backscatter location can be described reasonably well by using a simple model involving a cosine fit to diurnal variation in echo band position.
+\begin{enumerate}
+\item Backscatter location can be described reasonably well by using a simple model involving a cosine fit to diurnal variation in echo band position.
 The derived fit characteristics of the echo band display a clear pattern of diurnal, seasonal, and solar cycle changes that is very similar to those of the solar zenith angle and model peak \(F\) region density.
@@
 Conjugate observations of backscatter and density indicate that polar patches transported from the dayside are a viable source for this ionization.
-
-2.
-Backscatter occurrence statistic exhibits evidence of both suppression and enhancement by sunlight, with a clear occurrence peak observed near the solar terminator.
+\item Backscatter occurrence statistic exhibits evidence of both suppression and enhancement by sunlight, with a clear occurrence peak observed near the solar terminator.
 Suppression occurs at the largest SZA values observed and is consistent with diminishing role of gradients in irregularity production and increasing role of conducting \(E\) layer.
 Enhancement by sunlight is mostly associated with more favorable propagation conditions at intermediate SZA values and large-scale gradients near the solar terminator possibly serving as seeds for further structuring processes.
-
-3.
-Indirect control by the solar wind variations is limited to the nightside, with enhanced backscatter occurrence observed during intervals of negative IMF By and Bz.
+\item Indirect control by the solar wind variations is limited to the nightside, with enhanced backscatter occurrence observed during intervals of negative IMF By and Bz.
 The IMF control of small-scale irregularity production can occur through IMF-controlled variations in the high-latitude convection pattern that, in turn, lead to changes in mutual orientations between convection and prevalent gradient vectors, with plasma structuring favoring certain orientations.
+\end{enumerate}
 

### 2017-01-29 paper2.tex
Formatting updated

@@
 when an elongated structure is oriented along the radar boresite.
-The convection direction that results in the strongest asymmetries exhibits a strong dependence on the altitude, with the optimal convection being parallel to the gradient reversal in the E region, perpendicular to it in the F region, and at some angle between these extremes in the transitional region.
+The convection direction that results in the strongest asymmetries exhibits a strong dependence on the altitude, with the optimal convection being parallel to the gradient reversal in the \(E\) region, perpendicular to it in the \(F\) region, and at some angle between these extremes in the transitional region.
 Implications for observations of polar patches by oblique scanning radars within the Super Dual Auroral Radar Network are discussed.
@@
 Plasma density structures in the high-latitude ionosphere range in size from a few centimeters to hundreds of kilometers.
-In the ionospheric F region, they typically move with the large-scale plasma convection, i.e.
+In the ionospheric \(F\) region, they typically move with the large-scale plasma convection, i.e.
 their phase velocity is the \(\vec{E}\times\vec{B}\) drift velocity.
@@
 The GDI growth rate is generally dependent on the relative directions of the background density gradient vector \(\nabla n_0\) and the convection electric field \(\vec{E}\) or, equivalently, the convection velocity \(\vec{V}_E=\vec{E}\times\vec{B}/B^2\).
-It is widely accepted that the maximum GDI growth rate in the F region occurs when the density gradient \(\nabla n_0\) is parallel to the convection drift velocity \(\vec{V}_E\) \citep{Keskinen1982a}, while in the E region the maximum occurs when \(\nabla n_0\parallel \vec{E}\).
-An analytic expression for the GDI growth rate in the F region has been derived in numerous studies, with most of these derivations considering a particular subset of mutual orientations between \(\nabla n_0\), \(\vec{V}_E\), and the wave vector \(\vec{k}\).
+It is widely accepted that the maximum GDI growth rate in the \(F\) region occurs when the density gradient \(\nabla n_0\) is parallel to the convection drift velocity \(\vec{V}_E\) \citep{Keskinen1982a}, while in the \(E\) region the maximum occurs when \(\nabla n_0\parallel \vec{E}\).
+An analytic expression for the GDI growth rate in the \(F\) region has been derived in numerous studies, with most of these derivations considering a particular subset of mutual orientations between \(\nabla n_0\), \(\vec{V}_E\), and the wave vector \(\vec{k}\).
 This typically results in the upper limit of the GDI growth rate of \(\gamma=V_E/L\), where \(L\) is a characteristic gradient scale length defined through \(L^{-1}\equiv |\nabla n_0|/n_0\).
-More general vector configurations and growth rate expressions in the F region have also been considered \citep{Keskinen1982a,Makarevich2014c}, but their implications for studying the asymmetries between leading and trailing edges of large-scale structures have not been investigated.
-Further, it has been demonstrated recently that the GDI growth rate patterns change dramatically between the E and F regions \citep{Makarevich2014c}.
+More general vector configurations and growth rate expressions in the \(F\) region have also been considered \citep{Keskinen1982a,Makarevich2014c}, but their implications for studying the asymmetries between leading and trailing edges of large-scale structures have not been investigated.
+Further, it has been demonstrated recently that the GDI growth rate patterns change dramatically between the \(E\) and \(F\) regions \citep{Makarevich2014c}.
 This means that one expects different signatures of small-scale irregularities and their asymmetries near large-scale structures at different altitudinal regions.
@@
 Oblique-sounding coherent radars within the Super Dual Auroral Radar Network (SuperDARN) routinely detect small-scale irregularities in a broad range of altitudes and directions \citep{Chisham2007}.
-Importantly, both E and F region irregularities are observed by the same radar \citep[e.g.
-recent papers by][]{Lamarche2015,Forsythe2015} and measurements of the background convection velocity \(\vec{V}_E\) are routinely derived from F-region observations.
+Importantly, both \(E\) and \(F\) region irregularities are observed by the same radar \citep[e.g.
+recent papers by][]{Lamarche2015,Forsythe2015} and measurements of the background convection velocity \(\vec{V}_E\) are routinely derived from \(F\)-region observations.
 In addition to their ability to observe large-scale structures in context of background plasma convection, SuperDARN radars provided important observational evidence in support of the GDI mechanism being responsible for the irregularity formation and asymmetries near gradient reversals \citep{Milan2002a,Koustov2012}.
@@
 The aim of this study is to investigate the asymmetries between the GDI growth rates near gradient reversals that often occur in the vicinity of large-scale density structures.
-The specific objectives are to (1) develop a quantitative approach for comparing the growth rates on either side of the structure,  (2) analyse how the growth rates and asymmetries change for different orientations of the background density gradient and convection velocity and determine if there is an optimal radar orientation for which the asymmetry is the strongest, and (3) investigate how the growth rate and asymmetries change between the ionospheric E and F layers.
+The specific objectives are to (1) develop a quantitative approach for comparing the growth rates on either side of the structure,  (2) analyse how the growth rates and asymmetries change for different orientations of the background density gradient and convection velocity and determine if there is an optimal radar orientation for which the asymmetry is the strongest, and (3) investigate how the growth rate and asymmetries change between the ionospheric \(E\) and \(F\) layers.
 
@@
 MSISE-90 is an empirical model of the densities and temperatures of neutral species in the Earth's atmosphere \citep{Hedin1991}.
-For this study, the neutral densities were found for February 15, 2014 at 20 UT at McMurdo station, Antarctica (77.88\(\deg\)S, 166.73\(\deg\)E, geographic), however the parameters considered are expected to be relatively stable so this selection does not constrain the results.
+For this study, the neutral densities were found for February 15, 2014 at 20 UT at McMurdo station, Antarctica (77.88\(\deg\)S, 166.73\(\deg\)\(E\), geographic), however the parameters considered are expected to be relatively stable so this selection does not constrain the results.
 
 \begin{figure}
+	\centering
 	\includegraphics[width=\textwidth,angle=180]{f01.pdf}
-\caption{(a) Altitudinal variations of parameters of interest.
+	\caption[Overview of altitudinal variations and model set-up of MCM]{(a) Altitudinal variations of parameters of interest.
 Shown are the ratios of collision frequencies and gyrofrequencies for electrons \(-r_e=\nu_e/|\Omega_e|\) (blue) and ions \(r_i=\nu_i/\Omega_i\) (green), the anisotropy parameter \(\psi=-r_i r_e\) (pink), and the conductivity ratio \(R=\sigma_H/\sigma_P = \left(r_i+r_e\right)/\left(1+\psi\right)\) (dashed black line on top of the green one).
 (b) Virtual height of radar scattering volume versus the slant range.
-The areas shaded with red, green, and blue color refer to the E region, transitional region, and F region, respectively.
+The areas shaded with red, green, and blue color refer to the \(E\) region, transitional region, and \(F\) region, respectively.
 (c) Geometry of observations and the gradient-drift instability (GDI) growth rate pattern in the radar's field-of-view (FoV).
@@
 The normalized GDI growth rate at every cell in the radar's FoV was calculated and the radar cells are colored accordingly, with the color bar shown in the bottom-right corner.
-The E region, transitional region, and F region are outlined in red, green, and blue, respectively.}
+The \(E\) region, transitional region, and \(F\) region are outlined in red, green, and blue, respectively.}
 	\label{fig:1}
@@
 Figure \ref{fig:1}b shows the standard SuperDARN virtual height model \citep{Chisham2008}.
-The ranges where the beam is within the E region (red), F region (blue), and transitional region (green) are indicated.
+The ranges where the beam is within the \(E\) region (red), \(F\) region (blue), and transitional region (green) are indicated.
 The virtual height model assumes a particular height depending on the slant range.
 The model is based on where the signal is typically coming from for any given range.
-In the standard model the virtual height is constant throughout the E and F regions, but changes rapidly in the transitional region between 600 km and 800 km as shown in Figure \ref{fig:1}b.
+In the standard model the virtual height is constant throughout the \(E\) and \(F\) regions, but changes rapidly in the transitional region between 600 km and 800 km as shown in Figure \ref{fig:1}b.
 For individual observations in each beam, the height variations will be more gradual, but on average it is represented well by the model shown in Figure \ref{fig:1}b.
-A revised virtual height model in which virtual height changes with slant range has also been developed \citep{Chisham2008}, but since in the current study we were interested in the differences between the E and F region behavior (rather than small differences within each region), a standard virtual height model has been used.
+A revised virtual height model in which virtual height changes with slant range has also been developed \citep{Chisham2008}, but since in the current study we were interested in the differences between the \(E\) and \(F\) region behavior (rather than small differences within each region), a standard virtual height model has been used.
 
@@
 This alternate presentation has been shown in \citet{Makarevich2014c}, their figure 1 and figure 2.
-These figures show the GDI growth rate in all wavevector directions for set gradient and drift velocity directions for the E and F regions, respectively.
+These figures show the GDI growth rate in all wavevector directions for set gradient and drift velocity directions for the \(E\) and \(F\) regions, respectively.
 
@@
 
-Figure \ref{fig:2}a shows a gradient reversal through the center of the F-region portion of the FoV that is perpendicular to the boresite (\(\theta=0\deg\)).
+Figure \ref{fig:2}a shows a gradient reversal through the center of the \(F\)-region portion of the FoV that is perpendicular to the boresite (\(\theta=0\deg\)).
 The growth rate on both sides of the gradient reversal is small (close to zero), but it is slightly negative on the leading edge and slightly positive on the trailing edge.
@@
 In the transitional region, the growth rate is again close to zero near the boresite, but increases to \(\sim0.7\) on either edge of the beam.
-The E region shows a completely different pattern, where the growth rate is negative on the left side and smoothly transitions to positive on the right side.
-In addition, the most extreme values for growth rate in the E region are above \(\pm1.0\), i.e.
-much higher than what is observed in the F region.
+The \(E\) region shows a completely different pattern, where the growth rate is negative on the left side and smoothly transitions to positive on the right side.
+In addition, the most extreme values for growth rate in the \(E\) region are above \(\pm1.0\), i.e.
+much higher than what is observed in the \(F\) region.
 
@@
 	\includegraphics[width=\textwidth]{f02.pdf}
-\caption{The GDI growth rate within the radar's FoV for four cases of gradient reversal orientation and position (white line).
+	\caption[GDI growth rate within radar FoV]{The GDI growth rate within the radar's FoV for four cases of gradient reversal orientation and position (white line).
 In all cases, the convection direction \(\vec{V}_E\) (red arrow) is perpendicular to the gradient reversal direction.
@@
 The GDI growth rate at each radar cell is given by the cell color, with the color bar given in the bottom-left corner of panel (a).
-Panels (a) and (b) show a gradient reversal centered in the F region, while panels (c) and (d) show the reversal in the transitional region.
+Panels (a) and (b) show a gradient reversal centered in the \(F\) region, while panels (c) and (d) show the reversal in the transitional region.
 The angles \(\theta\) and \(\phi\) (as defined in Figure \ref{fig:1}c) are also given in the bottom-right corner of each panel.}
@@
 
-Figure \ref{fig:2}b shows an inclined gradient reversal through the FoV in the F region (\(\theta=40\deg\)).
+Figure \ref{fig:2}b shows an inclined gradient reversal through the FoV in the \(F\) region (\(\theta=40\deg\)).
 Again, the growth rate is negative on the leading edge and positive on the trailing edge, but the difference is much more dramatic.
@@
 The transitional region shows a similar pattern with lower positive values on the left side and growth rate above 1.0 on the right side.
-The growth rate is above 1.0 throughout the E region as well.
+The growth rate is above 1.0 throughout the \(E\) region as well.
 
@@
 One limitation of this expression is, however, that it can only be applied when the conductivity ratio \(R\) is approximately constant, i.e.
-in the E region or in the F region, but not in the transitional region.
+in the \(E\) region or in the \(F\) region, but not in the transitional region.
 
@@
 The numerical method results, Section \ref{sec:num}, are shown for the averages (solid lines) and standard deviations (shaded regions).
-The blue and red curves correspond to the E- and F-region cases.
+The blue and red curves correspond to the \(E\)- and \(F\)-region cases.
 The analytic method results, Section \ref{sec:ana}, are also presented for the expected average differences (dashed lines).
@@
 \begin{figure}
+	\centering
 	\includegraphics[width=\textwidth,angle=180]{f03.pdf}
-\caption{The difference in the GDI growth rate across the gradient reversal for four selected angles between the reversal and convection direction.
+	\caption[GDI growth rates for set angle between \(\vec{V}_E\) and \(\vec{G}\)]{The difference in the GDI growth rate across the gradient reversal for four selected angles between the reversal and convection direction.
 The gradient reversal direction \(\theta\) is on the \(x\)-axis and the average growth rate difference across the reversal \(\Delta\gamma\) is on the \(y\)-axis for each panel.
@@
 The dashed lines show the behavior of \(\Delta\gamma\) calculated using an analytic expression.
-The trends for both the E-region (red) and F-region (blue) cases are shown in each panel.}
+The trends for both the \(E\)-region (red) and \(F\)-region (blue) cases are shown in each panel.}
 	\label{fig:3}
@@
 the case when the convection direction is parallel to the gradient reversal.
-This corresponds to \(\vec{V}_E \perp \vec{G}\) or \(\vec{E} \parallel \vec{G}\) on the trailing edge, which is the condition for most favorable gradient orientation in the E region \citep{Makarevich2014c}.
-The growth rate difference in the E region (red) is negative at all gradient reversal directions \(\theta\) and it changes with \(\theta\) in a sinusoidal fashion.
+This corresponds to \(\vec{V}_E \perp \vec{G}\) or \(\vec{E} \parallel \vec{G}\) on the trailing edge, which is the condition for most favorable gradient orientation in the \(E\) region \citep{Makarevich2014c}.
+The growth rate difference in the \(E\) region (red) is negative at all gradient reversal directions \(\theta\) and it changes with \(\theta\) in a sinusoidal fashion.
 It has the lowest \(\Delta\gamma\) minimum values observed out of all 4 cases (\(\sim-10\) at \(90\deg\) and \(270\deg\)).
-The average growth rate difference for the F region (blue) also shows a periodic behavior with \(\theta\); it is positive for \(0\deg\)--\(90\deg\) and \(180\deg\)--\(270\deg\) and negative for all other angles.
-The amplitude of oscillations is much smaller than in the E region.
+The average growth rate difference for the \(F\) region (blue) also shows a periodic behavior with \(\theta\); it is positive for \(0\deg\)--\(90\deg\) and \(180\deg\)--\(270\deg\) and negative for all other angles.
+The amplitude of oscillations is much smaller than in the \(E\) region.
 The segments where \(\Delta\gamma>0\) correspond to the growth rate on the leading edge exceeding that on the trailing edge.
@@
 These show the gradual transition between the extreme cases of the convection direction parallel to the gradient reversal, Figure \ref{fig:3}a, and the convection direction perpendicular to the reversal, Figure \ref{fig:3}d.
-The general E-region pattern starts to move towards more positive values and the F-region pattern moves towards more negative values.
+The general \(E\)-region pattern starts to move towards more positive values and the \(F\)-region pattern moves towards more negative values.
 The location of the peaks in growth rate difference changes as well.
-For the E region, the location of the lowest \(\Delta\gamma\) moves gradually from \(\theta\approx 90\deg, 270\deg\) when \(\phi-\theta=0\deg\) to \(\theta\approx 45\deg, 225\deg\) when \(\phi-\theta=90\deg\).
-Similarly, for the F region, the location of the lowest \(\Delta\gamma\) moves from \(\theta\approx 135\deg, 315\deg\) to \(\theta\approx 90\deg, 270\deg\) between these two situations.
+For the \(E\) region, the location of the lowest \(\Delta\gamma\) moves gradually from \(\theta\approx 90\deg, 270\deg\) when \(\phi-\theta=0\deg\) to \(\theta\approx 45\deg, 225\deg\) when \(\phi-\theta=90\deg\).
+Similarly, for the \(F\) region, the location of the lowest \(\Delta\gamma\) moves from \(\theta\approx 135\deg, 315\deg\) to \(\theta\approx 90\deg, 270\deg\) between these two situations.
 
 Figure \ref{fig:3}d shows the case when the convection direction is perpendicular to the gradient reversal, which corresponds to \(\vec{V}_E \parallel \vec{G}\).
-From previous studies, this is the condition where the growth rate on the trailing edge is the greatest in the F region.
-Similar to Figure \ref{fig:3}a, the growth rate difference in Figure \ref{fig:3}d for the F region (blue) is negative for all directions of the gradient reversal and has the lowest values observed for the F region (\(\sim-2\)) at \(90\deg\) and \(270\deg\).
+From previous studies, this is the condition where the growth rate on the trailing edge is the greatest in the \(F\) region.
+Similar to Figure \ref{fig:3}a, the growth rate difference in Figure \ref{fig:3}d for the \(F\) region (blue) is negative for all directions of the gradient reversal and has the lowest values observed for the \(F\) region (\(\sim-2\)) at \(90\deg\) and \(270\deg\).
 This corresponds to the largest contrast between the leading and trailing edge of the structure.
@@
 For other \(\theta\) values, the difference may be larger or smaller than at other values of \(\phi-\theta\).
-The E-region (red) extreme values (maximum \(\Delta\gamma\approx 4\) and minimum \(\Delta\gamma\approx -6\)) have the lowest magnitude out of all 4 cases shown in panels (a)--(d).
-This indicates that the contrast in the E region is the smallest for this particular orientation, although it is still much larger than that observed in the F region.
+The \(E\)-region (red) extreme values (maximum \(\Delta\gamma\approx 4\) and minimum \(\Delta\gamma\approx -6\)) have the lowest magnitude out of all 4 cases shown in panels (a)--(d).
+This indicates that the contrast in the \(E\) region is the smallest for this particular orientation, although it is still much larger than that observed in the \(F\) region.
 
 Throughout all 4 cases, there are several common features.
-The amplitude of variations of the growth rate difference in the F region is consistently much less than that in the E region.
-Generally, there is very good agreement between the numerical method and the analytic method in both the E and F regions.
+The amplitude of variations of the growth rate difference in the \(F\) region is consistently much less than that in the \(E\) region.
+Generally, there is very good agreement between the numerical method and the analytic method in both the \(E\) and \(F\) regions.
 Typically, the results from the analytic method fall well within one standard deviation of the numerical method.
@@
 The analysis presented in Figure \ref{fig:3} can be extended to include any arbitrary orientation between the gradient reversal and convection direction by allowing \(\phi\) and \(\theta\) to vary independently.
-Both the numerical method and the analytical method can be used to find the average difference in growth rate for any direction of the gradient reversal and plasma convection in the E and F regions.
+Both the numerical method and the analytical method can be used to find the average difference in growth rate for any direction of the gradient reversal and plasma convection in the \(E\) and \(F\) regions.
 Figure \ref{fig:4} presents the results of this analysis in the form of contour plots of the growth rate difference \(\Delta\gamma\) versus angles \(\theta\) and \(\phi\).
-Figures \ref{fig:4}a and \ref{fig:4}b show the results of the analytic method in the E and F regions, respectively.
-The remaining 4 panels show the results from the numerical method in (c) the E region, (d) the F region, (e) transitional region, and (f) all regions combined.
+Figures \ref{fig:4}a and \ref{fig:4}b show the results of the analytic method in the \(E\) and \(F\) regions, respectively.
+The remaining 4 panels show the results from the numerical method in (c) the \(E\) region, (d) the \(F\) region, (e) transitional region, and (f) all regions combined.
 With exception of Figure \ref{fig:4}e, different values of \(\Delta\gamma\) along the gradient reversal were included in the averaging only for the points with specified virtual heights, e.g.
@@
 \begin{figure}
+	\centering
 	\includegraphics[width=\textwidth,angle=180]{f04.pdf}
-\caption{Contour plots of the growth rate difference \(\Delta\gamma\)  versus the two angles of interest.
+	\caption[GDI growth rates for arbitrary angle between \(\vec{V}_E\) and \(\vec{G}\)]{Contour plots of the growth rate difference \(\Delta\gamma\)  versus the two angles of interest.
 The direction of the gradient reversal \(\theta\) is on the \(x\)-axis and the convection direction \(\phi\) is on the \(y\)-axis.
 The color bar is given to the right.
-Panels (a) and (b) show the results using an analytic approach in the E and F regions, respectively.
-Panels (c)--(f) show the results using a numerical approach in the E, F, transitional, and all regions, respectively.
+Panels (a) and (b) show the results using an analytic approach in the \(E\) and \(F\) regions, respectively.
+Panels (c)--(f) show the results using a numerical approach in the \(E\), \(F\), transitional, and all regions, respectively.
 The dashed black and white line in panels (c) and (f) indicates the average value of the conductivity ratio \(R\) along the gradient reversal for different angles \(\theta\) in the transitional region and all regions, respectively.
@@
 
-Figure \ref{fig:4}a shows the average growth rate difference across the gradient reversal found by the analytic method in the E region.
+Figure \ref{fig:4}a shows the average growth rate difference across the gradient reversal found by the analytic method in the \(E\) region.
 The same information but for 4 specific cases of \(\phi-\theta\) were shown in Figure \ref{fig:3}.
@@
 
-Figure \ref{fig:4}b presents the same information but for the F region.
+Figure \ref{fig:4}b presents the same information but for the \(F\) region.
 This refers to the blue dashed lines in Figure \ref{fig:3}.
@@
 
-Figure \ref{fig:4}c and d present the growth rate difference in the E and F regions, respectively, found with the numerical method.
-A comparison of Figures \ref{fig:4}a and b with Figures \ref{fig:4}c and d shows that the results of the analytic and numerical methods are very similar for both the E and F regions.
+Figure \ref{fig:4}c and d present the growth rate difference in the \(E\) and \(F\) regions, respectively, found with the numerical method.
+A comparison of Figures \ref{fig:4}a and b with Figures \ref{fig:4}c and d shows that the results of the analytic and numerical methods are very similar for both the \(E\) and \(F\) regions.
 The results for the numerical method (Figures \ref{fig:4}c and d) are not as smooth as the analytic method results, but this is simply due to the discrete steps that are used in the numerical method.
-Comparison between the E and F region patterns, however, reveals a few interesting features.
-First, the location of the minima moves from \(\phi=90\deg,270\deg\) in the E region to \(\phi=0\deg,180\deg\) in the F region.
+Comparison between the \(E\) and \(F\) region patterns, however, reveals a few interesting features.
+First, the location of the minima moves from \(\phi=90\deg,270\deg\) in the \(E\) region to \(\phi=0\deg,180\deg\) in the \(F\) region.
 These results agree with what was observed in Figure \ref{fig:3}.
-In addition, there are large regions of strongly positive \(\Delta\gamma\) (dark red) in the E region (Figures \ref{fig:4}a and \ref{fig:4}c) that do not appear in the F region.
-There are some areas of positive \(\Delta\gamma\) in the F region, but they are much more limited and have a maximum value of only \(\sim 1\).
-This means that while there are significant areas where the growth rate on the leading edge is greater than that on the trailing edge in the E region, the growth rate on the trailing edge almost always exceeds the leading edge in the F region.
+In addition, there are large regions of strongly positive \(\Delta\gamma\) (dark red) in the \(E\) region (Figures \ref{fig:4}a and \ref{fig:4}c) that do not appear in the \(F\) region.
+There are some areas of positive \(\Delta\gamma\) in the \(F\) region, but they are much more limited and have a maximum value of only \(\sim 1\).
+This means that while there are significant areas where the growth rate on the leading edge is greater than that on the trailing edge in the \(E\) region, the growth rate on the trailing edge almost always exceeds the leading edge in the \(F\) region.
 This is an important generalization of the well-known result that growth rates are more positive on the trailing edge, where \(\vec{G}\cdot \vec{V}_E>0\).
@@
 Figures \ref{fig:4}e shows results of the numerical analysis of the transitional region.
-This panel does not have a comparison with the analytic method because the analytic method requires a constant conductivity ratio \(R\) which changes rapidly from \(\sim\)5.08 in the E region to 0.001 in the F region, Figure \ref{fig:1}a.
+This panel does not have a comparison with the analytic method because the analytic method requires a constant conductivity ratio \(R\) which changes rapidly from \(\sim\)5.08 in the \(E\) region to 0.001 in the \(F\) region, Figure \ref{fig:1}a.
 To demonstrate this, the average \(R\) value along the gradient reversal is plotted using the black-white dashed line and the scale shown on the right.
@@
 Different altitudes correspond to hugely different values of \(R\), which causes the variation in the conductivity ratio shown in Figure \ref{fig:4}e.
-On average, however, the conductivity ratio is somewhere in between its values in the E and F regions, so the pattern shown is between the E region pattern and F region pattern.
-The minimum is between \(\phi=90\deg\) (the E region value) and \(\phi=180\deg\) (the F region value) and has a value of \(\sim -3\).
+On average, however, the conductivity ratio is somewhere in between its values in the \(E\) and \(F\) regions, so the pattern shown is between the \(E\) region pattern and \(F\) region pattern.
+The minimum is between \(\phi=90\deg\) (the \(E\) region value) and \(\phi=180\deg\) (the \(F\) region value) and has a value of \(\sim -3\).
 This will be further examined in Figure \ref{fig:5}.
@@
 with averaging done over all points along the gradient reversal regardless of their virtual height.
-For most angles of the gradient reversal, only the F region will be sampled, Figure \ref{fig:1}c.
-The gradient reversal will only pass through the transitional and E region if it is close to parallel with the boresite, i.e.
+For most angles of the gradient reversal, only the \(F\) region will be sampled, Figure \ref{fig:1}c.
+The gradient reversal will only pass through the transitional and \(E\) region if it is close to parallel with the boresite, i.e.
 when \(\theta\) is close to either \(90\deg\) or \(270\deg\).
-This is why the majority of the pattern in Figure \ref{fig:4}f resembles the F region patterns (Figure \ref{fig:4}d), but changes sharply around \(\theta=90\deg\) and \(\theta=270\deg\) due to contributions from the E region pattern (Figure \ref{fig:4}c).
+This is why the majority of the pattern in Figure \ref{fig:4}f resembles the \(F\) region patterns (Figure \ref{fig:4}d), but changes sharply around \(\theta=90\deg\) and \(\theta=270\deg\) due to contributions from the \(E\) region pattern (Figure \ref{fig:4}c).
 This can also be seen in the conductivity ratio.
-For most angles, average \(R\) only has contribution from the F region, so it is at the F-region value (very close to zero on this scale).
-However, close to \(\theta=90\deg\) and \(\theta=270\deg\) average \(R\) includes E-region values (\(\sim5\)), so it increases sharply.
+For most angles, average \(R\) only has contribution from the \(F\) region, so it is at the \(F\)-region value (very close to zero on this scale).
+However, close to \(\theta=90\deg\) and \(\theta=270\deg\) average \(R\) includes \(E\)-region values (\(\sim5\)), so it increases sharply.
 This causes \(\Delta\gamma\) to be relatively low with values between \(-1.5\) and \(+1.5\) for most angles \(\theta\); it changes around \(\theta=90\deg\) and \(\theta=270\deg\) to have values less than \(-2\).
@@
 Figure \ref{fig:5}a shows the average conductivity ratio (black line) and height (orange line) from 550 km to 850 km in slant range.
-The ranges where the beam is in the E (red), F (blue), and transitional (green) regions are shaded.
-Both the height and the conductivity ratio are constant in the E and F regions since virtual height is constant at these ranges in the standard virtual height model, Figure \ref{fig:1}b.
+The ranges where the beam is in the \(E\) (red), \(F\) (blue), and transitional (green) regions are shaded.
+Both the height and the conductivity ratio are constant in the \(E\) and \(F\) regions since virtual height is constant at these ranges in the standard virtual height model, Figure \ref{fig:1}b.
 The height increases linearly in the transitional region, while the conductivity ratio decreases rapidly between 600 km and 616 km (shaded in yellow) at the lowest ranges within the transitional region.
@@
 \begin{figure}
+	\centering
 	\includegraphics[width=\textwidth,angle=180]{f05.pdf}
-\caption{Analysis of the transitional region.
-(a) The average conductivity ratio \(R\) (black line) and the radar virtual height (orange line) versus the slant range for the entire transitional region and parts of the E and F regions.
+	\caption[GDI growth rate within transitional region]{Analysis of the transitional region.
+(a) The average conductivity ratio \(R\) (black line) and the radar virtual height (orange line) versus the slant range for the entire transitional region and parts of the \(E\) and \(F\) regions.
 The scales for the conductivity ratio and height are given on the left and right of panel (a), respectively.
-The E (red), transitional (yellow and green) and F (blue) regions are highlighted.
+The \(E\) (red), transitional (yellow and green) and \(F\) (blue) regions are highlighted.
 The region highlighted with yellow is examined closer in panels (b) and (c).
@@
 The value of \(R\) for the analytic method was taken from the average \(R\) for each 2 km segment, as shown in Figure \ref{fig:5}b.
-From 600--602 km, \(R\) is close to the E region value (\(\sim5.1\)), so the patterns for \(\Delta\gamma\) are close to those for the E region (Figures \ref{fig:4}a and c).
-However, by 614--616 km, the value of \(R\) has decreased substantially and the patterns resemble the F-region patterns (Figures \ref{fig:4}b and d) much more closely.
+From 600--602 km, \(R\) is close to the \(E\) region value (\(\sim5.1\)), so the patterns for \(\Delta\gamma\) are close to those for the \(E\) region (Figures \ref{fig:4}a and c).
+However, by 614--616 km, the value of \(R\) has decreased substantially and the patterns resemble the \(F\)-region patterns (Figures \ref{fig:4}b and d) much more closely.
 From 600 km to 616 km, the location of lowest value of \(\Delta\gamma\) changes from \(\phi=90\deg,270\deg\) to \(\phi=0\deg,180\deg\).
-This represents the change from where the lowest \(\Delta\gamma\) values are in the E region to their location in the F region.
-In addition, the lowest value of \(\Delta\gamma\) also increases from less than \(-8\) (E region) to \(\sim-3\) (F region).
-The areas of highly positive \(\Delta\gamma\) that appeared in the E region (Figures \ref{fig:4}a and c) are not evident above 606 km.
+This represents the change from where the lowest \(\Delta\gamma\) values are in the \(E\) region to their location in the \(F\) region.
+In addition, the lowest value of \(\Delta\gamma\) also increases from less than \(-8\) (\(E\) region) to \(\sim-3\) (\(F\) region).
+The areas of highly positive \(\Delta\gamma\) that appeared in the \(E\) region (Figures \ref{fig:4}a and c) are not evident above 606 km.
 This indicates that after this point, there is no longer a significant subset of orientations where the growth rate on the leading edge is larger than that on the trailing edge.
@@
 The fluid theory itself also applies directly to long wavelengths, i.e scales much larger than the ion mean free path.
-The observations with which we compare our results refer to small-scale irregularities and, more specifically, decameter-scale waves in the E and F regions as well as in the transitional region, in the case of SuperDARN observations.
-In the E region up to 130 km in altitude, these conditions may be satisfied for decameter scales \citep[e.g.][]{Fejer1984}.
-In the F region, the long-wavelength case refers to significantly larger scales (\(\sim\) 1000 m) below which the fluid theory is not necessarily directly applicable.
-However, observational studies have found the behavior of decameter-scale irregularities in the F region to be consistent with many fluid theory predictions, including their propagation with the \(\vec{E}\times\vec{B}\) drift velocity and stronger waves on the trailing edge of gradient reversals.
-This leads to a common interpretation that decameter-scale waves observed by SuperDARN and  kilometer-scale waves in the F region described by a standard fluid approach are closely related, possibly by a turbulent cascade \citep[e.g.][]{Tsunoda1988}.
-Within the above discussed limits of applicability of the basic equation (\ref{eqn:Mak14gr}) employed in the present modeling, GDI is operational throughout the wide range of ionospheric altitudes spanning both the E and F regions.
-This is because there is always a range of propagation and gradient directions for which growth rates are positive in both the E and F regions.
+The observations with which we compare our results refer to small-scale irregularities and, more specifically, decameter-scale waves in the \(E\) and \(F\) regions as well as in the transitional region, in the case of SuperDARN observations.
+In the \(E\) region up to 130 km in altitude, these conditions may be satisfied for decameter scales \citep[e.g.][]{Fejer1984}.
+In the \(F\) region, the long-wavelength case refers to significantly larger scales (\(\sim\) 1000 m) below which the fluid theory is not necessarily directly applicable.
+However, observational studies have found the behavior of decameter-scale irregularities in the \(F\) region to be consistent with many fluid theory predictions, including their propagation with the \(\vec{E}\times\vec{B}\) drift velocity and stronger waves on the trailing edge of gradient reversals.
+This leads to a common interpretation that decameter-scale waves observed by SuperDARN and  kilometer-scale waves in the \(F\) region described by a standard fluid approach are closely related, possibly by a turbulent cascade \citep[e.g.][]{Tsunoda1988}.
+Within the above discussed limits of applicability of the basic equation (\ref{eqn:Mak14gr}) employed in the present modeling, GDI is operational throughout the wide range of ionospheric altitudes spanning both the \(E\) and \(F\) regions.
+This is because there is always a range of propagation and gradient directions for which growth rates are positive in both the \(E\) and \(F\) regions.
 
@@
 In the polar cap, HF radars rely on beam refraction through a dense ionosphere to observe field-aligned irregularities \citep{Bristow2011,Koustov2014}.
-In addition, high background electron density in both the E and F regions can smooth gradients required for GDI and limit instability growth \citep{Ruohoniemi1997,Koustov2004}.
-Large conductivity in the E region can also short out F region irregularities \citep{Danskin2002,Kane2012,Lamarche2015}.
+In addition, high background electron density in both the \(E\) and \(F\) regions can smooth gradients required for GDI and limit instability growth \citep{Ruohoniemi1997,Koustov2004}.
+Large conductivity in the \(E\) region can also short out \(F\) region irregularities \citep{Danskin2002,Kane2012,Lamarche2015}.
 An enhanced D region may act to attenuate backscatter signals, although this effect is not always obvious \citep{Danskin2002}.
-Nonlinear processes in the F region have been theoretically shown to alter the dependence on the propagation direction while still showing significant anisotropy similar to that predicted by the linear theory \citep{Keskinen1981b,Keskinen1982a,Keskinen1983}.
-The situation in the E region appears to be somewhat different, with some echoes also exhibiting strong anisotropy \citep[e.g.][]{Ierkic1980,Makarevich2008} and other echoes showing isotropic power variation \citep{Ierkic1980,Koustov2001} presumably because of nonlinear processes.
-The results of the present study are thus more likely to apply in the F region.
+Nonlinear processes in the \(F\) region have been theoretically shown to alter the dependence on the propagation direction while still showing significant anisotropy similar to that predicted by the linear theory \citep{Keskinen1981b,Keskinen1982a,Keskinen1983}.
+The situation in the \(E\) region appears to be somewhat different, with some echoes also exhibiting strong anisotropy \citep[e.g.][]{Ierkic1980,Makarevich2008} and other echoes showing isotropic power variation \citep{Ierkic1980,Koustov2001} presumably because of nonlinear processes.
+The results of the present study are thus more likely to apply in the \(F\) region.
 
@@
 
-The results from Figure \ref{fig:4} show that \(\Delta\gamma\) is almost always negative in the F region, which implies that the growth rate on the trailing edge of a large-scale density structure is almost always larger than that on the leading edge.
-This agrees with previous observations of polar patches and sun-aligned arcs in the F region ionosphere that consistently show a greater irregularity intensity on the trailing edge than the leading edge of these structures \citep{Weber1984,Milan2002a,Koustov2012}.
+The results from Figure \ref{fig:4} show that \(\Delta\gamma\) is almost always negative in the \(F\) region, which implies that the growth rate on the trailing edge of a large-scale density structure is almost always larger than that on the leading edge.
+This agrees with previous observations of polar patches and sun-aligned arcs in the \(F\) region ionosphere that consistently show a greater irregularity intensity on the trailing edge than the leading edge of these structures \citep{Weber1984,Milan2002a,Koustov2012}.
 Greater power from radar backscatter is indicative of more decameter-scale structuring, which agrees with a greater growth rate on the trailing edge.
@@
 
-In the E region, large values of \(\Delta\gamma\) that are both positive and negative indicate that the growth rate can be high on both the leading and trailing edge.
-Several previous studies have investigated large-scale density structures in the E region \citep{Milan1999b,Milan2001a,Milan2002a}, but none of these have focused on asymmetries in irregularity characteristics associated with gradient reversals so it is difficult to compare these modeling results with observations.
+In the \(E\) region, large values of \(\Delta\gamma\) that are both positive and negative indicate that the growth rate can be high on both the leading and trailing edge.
+Several previous studies have investigated large-scale density structures in the \(E\) region \citep{Milan1999b,Milan2001a,Milan2002a}, but none of these have focused on asymmetries in irregularity characteristics associated with gradient reversals so it is difficult to compare these modeling results with observations.
 
-\citet{Koustov2008} presented data in their figure 5 that showed smaller echo power on the leading edge of a sun-aligned arc in the F region, but strong backscatter power on both sides of the arc in the short range gates, corresponding to the E-region or transitional region.
+\citet{Koustov2008} presented data in their figure 5 that showed smaller echo power on the leading edge of a sun-aligned arc in the \(F\) region, but strong backscatter power on both sides of the arc in the short range gates, corresponding to the \(E\)-region or transitional region.
 The arc was approximately parallel to the radar boresite and moved towards the left of the radar FoV, corresponding to \(\Delta\gamma\approx 0\) in our model (\(\theta=90\deg,\phi\approx 180\deg\) in figure \ref{fig:4}c).
 Slight changes in the convection direction (\(\phi\)) could create both the condition where \(\Delta\gamma>0\) (greater growth on leading edge) or \(\Delta\gamma<0\) (greater growth on trailing edge).
-Although this proposed mechanism seems reasonable, it is important to note that the sun-aligned arc examined in \citet{Koustov2008} has been mapped to an altitude of 250 km and there is no information about its vertical extent, so it is possible that the density enhancement does not reach the E region at all and the structuring is due to a different process.
+Although this proposed mechanism seems reasonable, it is important to note that the sun-aligned arc examined in \citet{Koustov2008} has been mapped to an altitude of 250 km and there is no information about its vertical extent, so it is possible that the density enhancement does not reach the \(E\) region at all and the structuring is due to a different process.
 
-An important new result of this study was that the difference in growth rate depends strongly on the orientation of the gradient reversal and convection direction in both the E and F regions.
-In general, the maximum \(\Delta\gamma\) occurred at \(\theta=90\deg,270\deg\) in both the E and F regions, regardless of \(\phi\).
+An important new result of this study was that the difference in growth rate depends strongly on the orientation of the gradient reversal and convection direction in both the \(E\) and \(F\) regions.
+In general, the maximum \(\Delta\gamma\) occurred at \(\theta=90\deg,270\deg\) in both the \(E\) and \(F\) regions, regardless of \(\phi\).
 This corresponds to the gradient reversal being aligned along the radar boresite with the gradient direction being perpendicular to the boresite.
@@
 
-In the F region, the peak in the \(\Delta\gamma_{avg}\) magnitude occurs when the convection direction is perpendicular to the radar boresite \(\phi=90\deg\).
-In the E region, the peak is when the convection direction is either parallel or antiparallel to the boresite \(\phi=0\deg, 180\deg\).
+In the \(F\) region, the peak in the \(\Delta\gamma_{avg}\) magnitude occurs when the convection direction is perpendicular to the radar boresite \(\phi=90\deg\).
+In the \(E\) region, the peak is when the convection direction is either parallel or antiparallel to the boresite \(\phi=0\deg, 180\deg\).
 This corresponds to the convection direction being parallel to the wave vector.
@@
 To test it, one would need to follow the track of a large elongated polar patch or sun-aligned arc.
-Although polar patches can retain their form in the F region for several hours \citep{Hosokawa2014}, it is likely that the patch would move a significant lateral distance from the boresite.
+Although polar patches can retain their form in the \(F\) region for several hours \citep{Hosokawa2014}, it is likely that the patch would move a significant lateral distance from the boresite.
 The structure may have to be observed by multiple SuperDARN radars from different directions unless a highly elongated patch (>1500 km) is observed, similar to that described in \citet{Hosokawa2014}.
 
-Another challenge in observing altitudinal transitions is associated with a requirements for a large density structure to have a significant vertical extent, at least from the lower F region through a significant portion of the transitional region.
+Another challenge in observing altitudinal transitions is associated with a requirements for a large density structure to have a significant vertical extent, at least from the lower \(F\) region through a significant portion of the transitional region.
 The vertical extent of these structures has not been investigated systematically, although several examples of density enhancements associated with these structures have been presented.
-Thus \citet{Schlatter2013} examined density enhancements due to high-energy auroral precipitation in the E and F region between \(\sim100\) km and \(\sim150\) km, see their Figure 1a.
+Thus \citet{Schlatter2013} examined density enhancements due to high-energy auroral precipitation in the \(E\) and \(F\) region between \(\sim100\) km and \(\sim150\) km, see their Figure 1a.
 \citet{Jayachandran2012} estimated the maximum vertical thickness of sun-aligned arcs as \(\sim84\) km.
@@
 \item Asymmetry is also strongly dependent on the background convection direction relative to the gradient reversal, but this dependence changes with altitude.
-Maximum asymmetry occurs when the background convection is parallel to a gradient reversal in the E region, but perpendicular to a gradient reversal in the F region.
-There is a relatively smooth transition between configurations associated with maximum asymmetry in the E and F layers.
-In the E region, either the leading or trailing edge can have significant positive (or negative) growth rate, resulting in the asymmetry favoring either edge, depending on the structure orientation.
-This is in contrast with the F region, where growth rate on the trailing edge is almost always larger than the growth rate on the leading edge, but with a much lower absolute maximum than in the E region.
+Maximum asymmetry occurs when the background convection is parallel to a gradient reversal in the \(E\) region, but perpendicular to a gradient reversal in the \(F\) region.
+There is a relatively smooth transition between configurations associated with maximum asymmetry in the \(E\) and \(F\) layers.
+In the \(E\) region, either the leading or trailing edge can have significant positive (or negative) growth rate, resulting in the asymmetry favoring either edge, depending on the structure orientation.
+This is in contrast with the \(F\) region, where growth rate on the trailing edge is almost always larger than the growth rate on the leading edge, but with a much lower absolute maximum than in the \(E\) region.
 This strong altitudinal dependence is due to the rapid decrease of the ratio of Hall and Pedersen conductivities at higher altitudes.
 \item An oblique scanning radar is optimally oriented to observe large-scale and elongated density structures in the high-latitude ionosphere when the radar's boresite is closest in direction to the structure elongation for any given altitude and convection direction since this configuration is associated with the largest contrast between leading and trailing edges.
-The convection direction that leads to the largest contrast depends on the altitude and for HF radars that predominantly receive backscatter from the F region the optimal convection direction is perpendicular to the boresite.
+The convection direction that leads to the largest contrast depends on the altitude and for HF radars that predominantly receive backscatter from the \(F\) region the optimal convection direction is perpendicular to the boresite.
 \end{enumerate}
 
-\appendix
+%\appendix
+\begin{subappendices}
 \section{Analytic Expression for the GDI Growth Rate Difference}\label{app:1}
@@
 To find the average difference in growth rate, we integrate over \(\alpha\) across the radar FoV between \(-\alpha_0\) and \(+\alpha_0\) with respect to the boresite.
-In this integration, we assume that the conductivity ratio \(R\) is a constant for any combination of angles, which is a fair assumption for observations in the E region or in the F region.
+In this integration, we assume that the conductivity ratio \(R\) is a constant for any combination of angles, which is a fair assumption for observations in the \(E\) region or in the \(F\) region.
 \begin{equation}
@@
 where a new constant  has been defined: \(C\equiv\sin(2\alpha_0)/2\alpha_0\). For a typical SuperDARN radar, \(\alpha_0=26\deg\) and \(C\approx 0.87\).
+
+\end{subappendices}
 

### 2017-01-30 paper1.tex
Further formatting adjustments

@@
 
-\chapter{Solar control of \(F\) region radar backscatter: Further insights from observations in the southern polar cap}
+\chapter[Solar control of \(F\) region radar backscatter: Further insights from observations in the southern polar cap]{Solar control of \(F\) region radar backscatter: Further insights from observations in the southern polar cap\footnote{Published as: {Lamarche}, L.~J., and R.~A.
+{Makarevich} (2015), Solar control of {F} region radar backscatter: Further insights from observations in the southern polar cap, \textit{J.
+Geophys.
+Res.
+Space Physics}, \textit{120}(11), 9875--9890, \doi{10.1002/2015JA021663}.}}
 
-%\begin{abstract}
+\label{sec:paper1}
 
+\section*{Abstract}
 The role of solar wind and illumination in production of small-scale \(F\) region plasma irregularities is investigated using a 4-year data set collected by the Super Dual Auroral Radar Network (SuperDARN) facility at the McMurdo station, Antarctica (MCM).
@@
 Indirect solar wind control is also exerted by the IMF-dependent convection pattern, since the gradient-drift instability favors certain orientations between the plasma density gradients and convection velocity.
-
-%\end{abstract}
 
@@
 
-%\acknowledgments This work was supported by NSF grants AGS-1248127 and AGS-1341902.
-SuperDARN data are freely available through the SuperDARN website at Virginia Polytechnic Institute and State University http://vt.superdarn.org/.
-Fortran source code for IRI 2007 is freely available for download from the Community Coordinated Modeling Center (CCMC) at http://ccmc.gsfc.nasa.gov/.
-OMNI data is available from NASA Goddard Space Flight Center Space Physics Data Facility at http://omniweb.gsfc.nasa.gov.
+\section*{Acknowledgments}
+This work was supported by NSF grants AGS-1248127 and AGS-1341902.
+SuperDARN data are freely available through the SuperDARN website at Virginia Polytechnic Institute and State University \url{http://vt.superdarn.org/}.
+Fortran source code for IRI 2012 is freely available for download from the Community Coordinated Modeling Center (CCMC) at \url{http://ccmc.gsfc.nasa.gov/}.
+OMNI data is available from NASA Goddard Space Flight Center Space Physics Data Facility at \url{http://omniweb.gsfc.nasa.gov}.
 The Resolute Bay incoherent scatter radar (North) is operated by SRI International on behalf of the US National Science Foundation under NSF Cooperative Agreement AGS-1133009.
-The data are accessible from the SRI International online database at http://amisr.sri.com/database/.
+The data are accessible from the SRI International online database at \url{http://amisr.sri.com/database/}.
 The authors thank W.
@@
 
-%\newpage
 \bibliographystyle{uafthesis}
 \bibliography{references}
-
-%\end{article}
-%\end{document}
-
-%3 points:
-
-%Solar control of irregularity production is not restricted to dayside
-%Production peak near solar terminator implies breakup of large-scale structures
-%Importance of IMF-controlled convection orientations relative to gradients

### 2017-01-30 paper2.tex
Further formatting adjustments

@@
-%\documentclass[angeo, manuscript]{copernicus}
+% Chapter 3: paper2
 
-%\begin{document}
+\chapter[A modeling study of asymmetries in plasma irregularity characteristics near gradient reversals]{A modeling study of asymmetries in plasma irregularity characteristics near gradient reversals\footnote{Published as: {Lamarche}, L.~J., and R.~A.
+{Makarevich} (2016), A modeling study of asymmetries in plasma irregularity characteristics near gradient reversals, \textit{Ann.
+Geophysicae}, \textit{34}, 709--723, \doi{10.5194/angeo-34-709-2016}.}}
 
-%\linenumbers
+\label{sec:paper2}
 
-\chapter{A modeling study of asymmetries in plasma irregularity characteristics near gradient reversals}
-
-%\Author[1]{Leslie J.}{Lamarche}
-%\Author[1]{Roman A.}{Makarevich}
-
-%\affil[1]{Geophysical Institute and Department of Physics, University of Alaska Fairbanks, Fairbanks, Alaska, USA}
-
-%\runningtitle{Asymmetries in Plasma Irregularities}
-
-%\runningauthor{Lamarche}
-
-%\correspondence{Leslie J. Lamarche (ljlamarche@alaska.edu)}
-
-%\received{}
-%\pubdiscuss{} %% only important for two-stage journals
-%\revised{}
-%\accepted{}
-%\published{}
-
-%\firstpage{1}
-
-%\maketitle
-
-%\begin{abstract}
-
+\section*{Abstract}
 Asymmetries in plasma density irregularity generation between the leading and trailing edges of the large-scale plasma density structures in the high-latitude ionosphere are investigated.
@@
 
-%\end{abstract}
-
 \section{Introduction}
+\label{sec:p2intro}
 Plasma density structures in the high-latitude ionosphere range in size from a few centimeters to hundreds of kilometers.
@@
 
-\section{Modeling of Growth Rates within Radar's Field-of-View}\label{Sec:2}
+\section{Modeling of Growth Rates within Radar's Field-of-View}
+\label{sec:model_fov}
 In this study, we investigate behavior of the GDI growth rate within the field-of-view (FoV) of an oblique scanning radar with a particular focus on asymmetries between the leading and trailing edges of large-scale density structures in the ionosphere.
@@
 In practice, the gradient magnitude will also change depending on the distance from the structure, but the directional and altitudinal effects in asymmetries will be the same.
-This is further examined in Section \ref{Sec:Rev} where a more realistic approach is employed.
+This is further examined in Section \ref{sec:model_reversal} where a more realistic approach is employed.
 The important common features are that the GDI growth rate and its asymmetries are strongly dependent on the altitude and direction of observations.
@@
 
-\section{Modeling of Growth Rate Differences Across the Gradient Reversal}\label{Sec:Rev}
-
+\section{Modeling of Growth Rate Differences Across the Gradient Reversal}
+\label{sec:model_reversal}
 In order to quantify the above-noted differences in the growth rate for different orientations of the gradient reversal, the average change in the growth rate across the reversal \(\Delta\gamma_{avg}\) was found for any general configuration of the gradient reversal and the convection direction.
@@
 \subsection{Numerical Method}
-\label{sec:num}
+\label{sec:numerical}
 In the first method, evenly spaced points were selected along the gradient reversal within the radar FoV.
@@
 
-\subsection{Analytic Method}\label{sec:ana}
+\subsection{Analytic Method}
+\label{sec:analytic}
 
-An alternative approach is to develop an analytic expression for \(\Delta\gamma_{avg}\) for any general direction of the gradient reversal \(\theta\) and convection \(\phi\) as described in Appendix \ref{app:1}.
+An alternative approach is to develop an analytic expression for \(\Delta\gamma_{avg}\) for any general direction of the gradient reversal \(\theta\) and convection \(\phi\) as described in Appendix \ref{sec:p2appendix}.
 The result is
@@
 \end{equation}
-function \(\Theta\) being the Heaviside step function, and \(C=0.87\) for SuperDARN observations, Appendix \ref{app:1}.
+function \(\Theta\) being the Heaviside step function, and \(C=0.87\) for SuperDARN observations, Appendix \ref{sec:p2appendix}.
 This expression can be readily used to compute \(\Delta\gamma_{avg}\) for any given combination of \(\theta\) and \(\phi\).
@@
 The two methods for finding the average difference in the GDI growth rate across the gradient reversal can now be employed to characterize different orientations.
-In this section, we consider a more general case of different relative orientations of the gradient reversal and the convection direction, not necessarily restricted to the case of \(\theta- \phi= 90\deg\) that was considered in Section \ref{Sec:2}.
+In this section, we consider a more general case of different relative orientations of the gradient reversal and the convection direction, not necessarily restricted to the case of \(\theta- \phi= 90\deg\) that was considered in Section \ref{sec:model_fov}.
 In other words, we consider all possible differences \(\theta- \phi\).
@@
 The average growth rate difference is shown on the \(y\) axis in arbitrary units.
-The numerical method results, Section \ref{sec:num}, are shown for the averages (solid lines) and standard deviations (shaded regions).
+The numerical method results, Section \ref{sec:numerical}, are shown for the averages (solid lines) and standard deviations (shaded regions).
 The blue and red curves correspond to the \(E\)- and \(F\)-region cases.
-The analytic method results, Section \ref{sec:ana}, are also presented for the expected average differences (dashed lines).
+The analytic method results, Section \ref{sec:analytic}, are also presented for the expected average differences (dashed lines).
 Because the growth rate difference is calculated by subtracting the trailing edge \(\gamma\) from the leading edge \(\gamma\), negative \(\Delta\gamma\) indicates larger instability growth on the trailing edge than the leading edge.
@@
 \section{Discussion}
+\label{sec:p2discussion}
 In this study, we have applied the GDI linear fluid theory to study the asymmetries in the GDI growth rate on either side of a large-scale density structure, particularly as applied to observations with oblique scanning radars such as SuperDARN.
@@
 \section{Conclusions}
+\label{sec:p2conclusions}
 A model of asymmetries in plasma wave generation due to sharp reversals in the plasma density gradient near elongated density structures in the high-latitude ionosphere revealed the following:
@@
 
-%\appendix
 \begin{subappendices}
-\section{Analytic Expression for the GDI Growth Rate Difference}\label{app:1}
+\section{Analytic Expression for the GDI Growth Rate Difference}
+\label{sec:p2appendix}
 
@@
 Consider a single point that lies on the gradient reversal within the radar's FoV.
-In the geometry illustrated in Figure \ref{fig:1}c, the relevant vectors at this point can be written as \(\vec{\hat k}=-\sin\alpha\vec{\hat{e}}_1-\cos\alpha\vec{\hat{e}}_2\), \(\vec{\hat e}=-\sin\phi\vec{\hat{e}}_1+\cos\phi\vec{\hat{e}}_2\), \(\hat{\vec{g}}=\pm[\sin\theta\vec{\hat{e}}_1-\cos\theta\vec{\hat{e}}_2]\), where angles \(\theta\) and \(\phi\) were previously defined in Section \ref{Sec:2}, angle \(\alpha\) is the angle between the wave vector and the boresite direction, and unit vectors \(\vec{\hat{e}}_1\)  and \(\vec{\hat{e}}_2\) are  perpendicular and parallel to the boresite, respectively.
+In the geometry illustrated in Figure \ref{fig:1}c, the relevant vectors at this point can be written as \(\vec{\hat k}=-\sin\alpha\vec{\hat{e}}_1-\cos\alpha\vec{\hat{e}}_2\), \(\vec{\hat e}=-\sin\phi\vec{\hat{e}}_1+\cos\phi\vec{\hat{e}}_2\), \(\hat{\vec{g}}=\pm[\sin\theta\vec{\hat{e}}_1-\cos\theta\vec{\hat{e}}_2]\), where angles \(\theta\) and \(\phi\) were previously defined in Section \ref{sec:model_fov}, angle \(\alpha\) is the angle between the wave vector and the boresite direction, and unit vectors \(\vec{\hat{e}}_1\)  and \(\vec{\hat{e}}_2\) are  perpendicular and parallel to the boresite, respectively.
 
@@
 
-%\begin{acknowledgements}
-%This work was supported by NSF grants AGS-1248127 and PLR-1443504. The MSISE-90 model is available from the GSFC/SPDF OMNIWeb interface at http://omniweb.gsfc.nasa.gov/.
-%\end{acknowledgements}
+\section*{Acknowledgements}
+This work was supported by NSF grants AGS-1248127 and PLR-1443504. The MSISE-90 model is available from the GSFC/SPDF OMNIWeb interface at \url{http://omniweb.gsfc.nasa.gov/}.
 
@@
 \bibliography{references}
-
-%\newpage
-
-%\addtocounter{figure}{-1}\renewcommand{\thefigure}{\arabic{figure}a}
-
-%\end{document}

### 2017-01-30 introduction.tex
Formatting updated
Further revisions made to Observation and Motivation/Objectives sections in Introduction

@@
 \begin{figure}
+	\centering
 	\includegraphics[width=\textwidth,angle=0]{ste.pdf}
@@
 \begin{figure}
+	\centering
 	\includegraphics[width=\textwidth]{densprofile.pdf}
@@
 \begin{figure}
+	\centering
 	\includegraphics[width=0.5\textwidth]{irregularity.pdf}
@@
 \begin{figure}
+	\centering
 	\includegraphics[width=\textwidth,angle=180]{SuperDARNmap.pdf}
@@
 \begin{figure}
+	\centering
 	\includegraphics[width=\textwidth]{ISRmap.pdf}
@@
 \begin{figure}
-	\includegraphics[width=0.75\textwidth,angle=0]{magnetosphere.pdf}
+	\centering
+	\includegraphics[width=\textwidth,angle=0]{magnetosphere.pdf}
 	\caption[Magnetic reconnection footprint in the ionosphere]{Diagram of the interactions between the IMF, magnetosphere, and polar cap ionosphere.
@@
 \begin{figure}
-	\includegraphics[width=0.75\textwidth,angle=0]{patch_formation.pdf}
+	\centering
+	\includegraphics[width=\textwidth,angle=0]{patch_formation.pdf}
 	\caption[Polar patch formation]{Polar patch formation through transient magnetopause reconnection.
@@
 \begin{split}
-	H_\alpha = S_\alpha F_\alpha + D_\alpha^{-1} F_\parallel \quad\quad\quad 
+	H_\alpha = S_\alpha F_\alpha + D_\alpha^{-1} F_\parallel \quad\quad\quad\quad&\quad\quad\quad\quad 
 	C_\alpha = \frac{T_\alpha}{m_\alpha \Omega_\alpha} \\
-	F_\alpha = i k_\perp^2 D_\alpha + \vec{G}\cdot\vec{k}_\perp D_\alpha + \vec{G}\cdot\vec{k}\times\uvec{b} \quad\quad\quad
+	F_\alpha = i k_\perp^2 D_\alpha + \vec{G}\cdot\vec{k}_\perp D_\alpha + \vec{G}\cdot\vec{k}\times\uvec{b} \quad\quad&\quad\quad\quad\quad
 	F_\parallel = i k_\parallel^2 + \vec{G}\cdot\vec{k}_\parallel \\
-	S_\alpha = \frac{1}{1+D_\alpha^2} \quad\quad\quad
+	S_\alpha = \frac{1}{1+D_\alpha^2} \quad\quad\quad\quad\quad\quad&\quad
 	D_\alpha = -\frac{i}{\Omega_\alpha}(\omega-\vec{k}\cdot{V}_{\alpha 0})+r_\alpha
@@
 \begin{split}
-	b = -\frac{\vec{G}\cdot\vec{k}\times\uvec{b}}{k_\perp^2} \quad\quad\quad
+	b = -\frac{\vec{G}\cdot\vec{k}\times\uvec{b}}{k_\perp^2} \quad\quad&\quad\quad\quad\quad
 	y = \frac{k_\parallel}{k_\perp} \\
-	R = \frac{\sigma_H}{\sigma_P} = \frac{D_i+D_e}{1+\Psi} \quad\quad
-	\psi = -r_ir_e \quad\quad
+	R = \frac{\sigma_H}{\sigma_P} = \frac{D_i+D_e}{1+\Psi} \quad\quad\quad\quad\quad
+	\psi =& -r_ir_e \quad\quad\quad\quad\quad
 	\hat{\psi} = \psi\left(1+\frac{k_\parallel^2}{r_e^2 k_\perp^2}\right) \\
-	\vec{V}_d = \vec{V}_{e0}-\vec{V}_{i0} = (r_i-r_e)\left[s_es_i(1+\psi)\left(R\vec{V}_E-\frac{\vec{E}}{B}\right)\right]
+	\vec{V}_d = \vec{V}_{e0}-\vec{V}_{i0} = (r_i-r_e)&\left[s_es_i(1+\psi)\left(R\vec{V}_E-\frac{\vec{E}}{B}\right)\right]
 \end{split}
@@
 \begin{table}
+	\centering
+	\renewcommand{\arraystretch}{1.5}
+	\setlength{\tabcolsep}{15pt}
 	\begin{tabular}{l l}
@@
 All five signatures were observed as expected, strengthening the idea that transient magnetopause reconnection is responsible for patch formation.
-Furthermore, \citet{Carlson2006} identified and tracked a series of patches directly from the subauroral plasma reservoir and show that the boundary moves equatorwards before relaxing poleward, one of the main predictions of the transient magnetopause reconnection mechanism \citep{Lockwood1992b}.
+Furthermore, \citet{Carlson2006} identified and tracked a series of patches directly from the subauroral plasma reservoir and showed that the boundary moves equatorwards before relaxing poleward, one of the main predictions of the transient magnetopause reconnection mechanism \citep{Lockwood1992b}.
 
@@
 Studies of patches using all-sky airglow imagers (ASI) found that the density gradient on the leading edge of polar patches tends to be 2--3 times steeper than that on the trailing edge \citep{Hosokawa2016}.
-Additionally, large finger-like structures on the trailing edge of patches have been identified that were tens to hundreds of kilometers in size and agreed with predictions made by GDI simulations \citep{Gondarenko2004b}.
+Additionally, large finger-like structures on the trailing edge of patches have been identified that were tens to hundreds of kilometers in size and agreed with predictions made by GDI simulations \citep{Gondarenko2004b,Hosokawa2016}.
 \citet{Moen2012} made direct measurements of plasma density structuring using the ICI-2 sounding rocket.
@@
 
-Recent radar observations suggested that in the polar cap, plasma irregularity production may be much higher than at lower latitudes such that when background plasma conditions are favorable for radar propagation, FAIs are observed nearly continuously \citep{Bristow2011}.
-However, there are still definite diurnal, seasonal, and solar cycle trends as far as radar observations at lower, auroral latitudes are concerned \citep{Kane2012}.
+Recent radar observations suggested that plasma irregularity production in the polar cap may be much higher than at lower latitudes.
+When background plasma conditions are favorable for radar propagation in the polar caps, FAIs are observed nearly continuously \citep{Bristow2011}, however, there are still definite diurnal, seasonal, and solar cycle trends similar to what is observed at lower, auroral \citep{Kane2012}.
 There is an increase in echo occurrence at solar maximum, particularly in the midnight sector \citep{Milan1997,Koustov2004}.
@@
 Properly-oriented density gradients and electric fields are a necessary condition for GDI to be operational, Section \ref{sec:lit_instabilities}, and the stronger these are, the more likely the irregularity production rate will be high.
-Experimental studies focused on the latter factor, with both backscatter power and occurrence showing an increase with increasing electric field strength \citep{Danskin2002,Makarevich2014b}.
+Experimental studies focused on the latter factor, with both backscatter power and occurrence showing an increase with increasing electric field strength \citep{Fukumoto2000,Danskin2002,Makarevich2014b}.
 
@@
 The outline of this thesis is as follows.
-Chapter \ref{sec:introduction} provides an overview of the polar ionosphere and both plasma structuring mechanisms at both large and small scales.
+Chapter \ref{sec:introduction} provides an overview of the polar ionosphere and both plasma structuring mechanisms at both large- and small-scales.
 Chapter \ref{sec:paper1} presents an experimental investigation of decameter-scale plasma irregularities in the polar F-region and their solar control, with a particular focus on statistical analysis of data collected with a recently-deployed SuperDARN radar in the southern polar cap.

### 2017-01-30 abstract.tex
Abstract revised

@@
 
-Factors that contribute to the production and observation of plasma irregularities in the polar \(F\) region ionosphere are investigated.
-The primary focus of this thesis is decameter-scale irregularities observed with radars in the Super Dual Auroral Radar Network (SuperDARN).
+Plasma in the Earth's ionosphere is highly irregular on scales ranging between a few centimeters and hundreds of kilometers.
+Small-scale irregularities or plasma waves can scatter radio waves resulting in a loss of signal for navigation and communication networks.
+The polar region is particularly susceptible to strong disturbances due to its direct connection with the Sun's magnetic field and energetic particles.
+In this thesis, factors that contribute to the production of decameter-scale plasma irregularities in the polar \(F\) region ionosphere are investigated.
+Both global and local control of irregularity production are studied, i.e.
+we consider global solar control through solar illumination and solar wind as well as much more local control by plasma density gradients and convection electric field.
 
-First, solar control on irregularity production is evaluated.
-A band of ranges where echoes are statistically probable is identified.
-The location of this band within a radar's field of view varies diurnally and seasonally, but its motion can be explained through variations in background density with solar illumination effecting radar beam propagation.
-Backscatter occurrence peaks at the terminator.
-High backscatter occurrence at night implies an additional source of nighttime plasma density that is not accounted for by traditional models.
-An example is presented that shows frequent polar patches may be the source of this density.
-Nighttime occurrence is enhanced when a positive IMF By component is present, probably due to changes in the convection pattern creating a more favorable situation for gradient-drift instability growth.
+In the first experimental study, solar control of irregularity production is investigated using the Super Dual Auroral Radar Network (SuperDARN) radar at McMurdo, Antarctica.
+The occurrence trends for irregularities are analyzed statistically and a model is developed that describes the location of radar echoes within the radar's field-of-view.
+The trends are explained through variations in background density with solar illumination affecting radar beam propagation.
+However, it is found that the irregularity occurrence during the night is higher than expected from ray tracing simulations based on a standard ionospheric density model.
+The high occurrence at night implies an additional source of plasma density and it is proposed that large-scale density enhancements called polar patches may and be the source of this density.
 
-The gradient-drift instability is modeled within the field of view of a SuperDARN radar to identify when the growth rate is largest and examine asymmetry around large-scale structures, such as polar patches.
-Dependencies on the relative directions of the gradient, plasma drift, and wavevector are all considered throughout the ionosphere.
-In the \(F\) region, the strongest asymmetry is found when an elongated structure is oriented along the radar's boresight and moving perpendicular to its direction of elongation.
-These results have implications for observations made with the SuperDARN network, but assume the directional dependencies predicted by linear GDI theory are applicable at decameter scales.
+The second study is concerned with modeling irregularity characterics near a large-scale density gradient reversal, such as those expected near polar patches, with a particular focus on the asymmetry of the irregularity growth rate across the gradient reversal.
+Directional dependencies on the plasma density gradient, plasma drift, and wavevector are analyzed in the context of the recently developed general fluid theory of the gradient-drift instability.
+In the ionospheric \(F\) region, the strongest asymmetry is found when an elongated structure is oriented along the radar's boresight and moving perpendicular to its direction of elongation.
+These results have important implications for finding optimal configurations for oblique-scanning ionospheric radars such as SuperDARN to observe gradient reversals.
 
-To test the accuracy of the predictions found from the model, irregularities surrounding a polar patch are analyzed in the context of linear gradient-drift instability theory.
-Backscatter power and occurrence from decameter-scale irregularities are found using measurements from the SuperDARN radar at Rankin Inlet while background density gradients and electric fields are found from the north face of the Resolute Bay Incoherent Scatter Radar.
-Particular emphasis is placed on directional dependence and whether small-scale irregularities show any signs of anisotropy.
-While low powered echoes seem to follow the predictions of linear theory, higher power echoes quickly exhibit much more complicated, nonlinear behavior.
+To test the predictions of the developed model and the general theory of the gradient-drift instability, an experimental investigation is presented focusing on decameter-scale irregularities near a polar patch and the previously uninvestigated directional dependence of irregularity characteristics.
+Backscatter power and occurrence of irregularities are analyzed using measurements from the SuperDARN radar at Rankin Inlet, Canada, while background density gradients and convection electric fields are found from the north face of the Resolute Bay Incoherent Scatter Radar.
+It is shown that irregularity occurrence tends to follow the expected trends better than irregularity power, suggesting that while the gradient-drift instability may be a dominant process in generating small-scale irregularities, other mechanisms such as a shear-driven instability or nonlinear process may exert greater control over their intensity.
+
+It is concluded from this body of work that the production of small-scale plasma irregularities in the polar \(F\)-region ionosphere is controlled both by global factors such as solar illumination as well as local plasma density gradients and electric fields.
+In general, linear gradient-drift instability theory describes small-scale irregularity production well, particularly for low amplitude perturbations.
+The production of irregularities is complex, and while ground-based radars are invaluable tools to study the ionosphere, care must be taken to interpret results correctly.

### 2017-02-01 conclusion.tex
Text corrections made to Conclusion

--- 2017-01-20 conclusion.tex
@@
 Plasma irregularities in the \(F\)-region ionosphere have been observed for decades.
-However, despite extensive data sets collected over these years, supprisingly little is known about these irregularities, especially at small scales.
+However, despite extensive data sets collected over these years, surprisingly little is known about these irregularities, especially at small scales.
 This is in part due to the complex and turbulent nature of the ionosphere, particularly in the polar regions, but also due to some of the inherent limitations in traditional observation techniques which can make it very challenging to identify exactly what processes are operational.
-In this Chapter, we summarize the results that have previously been discussed in Chapters \ref{sec:paper1}--\ref{sec:paper3}, shown in Figure \ref{fig:conclusion}, and consider what overarching conclusions can be derived from them.
-Additionally, Section \ref{sec:futurework} will provide recommendations for future research which will help answer some of the still outstanding questions and further our understanding of small-scale polar \(F\)-region plasma irregularities.
+In Section \ref{sec:summary}, we summarize the results that have previously been discussed in Chapters \ref{sec:paper1}--\ref{sec:paper3}, and consider what overarching conclusions can be derived from them.
+In the following Section \ref{sec:futurework}, we also provide recommendations for future research which will help answer some of the still outstanding questions and further our understanding of small-scale polar \(F\)-region plasma irregularities.
 
@@
 
-In the polar caps, backscatter is often observed at nighttime even when background plasma density is low and propagation conditions are traditionally thought to be too poor for FAIs to be observed with HF radar.
+Figure \ref{fig:conclusion} summarizes the major results from this body of work.
+From the experimental results presented in Chapter \ref{sec:paper1}, radar backscatter from ionospheric irregularities is most likely to be observed at particular slant ranges in the radar FoV, Figure \ref{fig:conclusion}.
+The location of this window of high echo occurrence varies diurnally and seasonally, primarily due to the changing background plasma density in the ionosphere.
+The background density controls how far a radar beam must travel to be perpendicular to the magnetic field and to eventually be scattered back from FAIs.
+This can create an appearance that small-scale plasma irregularities are more common in certain physical locations while in reality, they may exist over a much wider area but it is easier to observe them in particular locations.
+
+It was also demonstraighted in Chapter \ref{sec:paper1} that, in the polar caps, backscatter is often observed at nighttime even when background plasma density is low and propagation conditions are traditionally thought to be too poor for FAIs to be observed with a HF radar.
 This implies an additional source of nighttime plasma density that is not easily accounted for by models.
-Frequent polar patches that propagate from the dayside to the nightside could be the source of this additional density.
-The requirements for confirming this hypotheis will be detailed further in Section \ref{sec:futurework}.
+Frequent polar patches that propagate from the dayside to the nightside could be the source of this additional density, Figure \ref{fig:conclusion}.
+The requirements for confirming this hypothesis will be detailed further in Section \ref{sec:fw_patches}.
 
@@
   \includegraphics[width=\textwidth]{conclusion.pdf}
-  \caption[Irregularity production factors]{Summary of the main conclusions in this thesis as to what factors most effect pasma irregularity production.}
+  \caption[Irregularity production factors]{Summary of the main conclusions in this thesis.
+The radar of interest (MCM) is highlighted.
+Polar patches (orange) drift antisunwards across the polar cap following the background plasma convection (red contours).
+Nighttime irregularity observation is enhanced for a positive IMF \(B_y\) component.
+Production is maximized near the terminator.
+Small-scale irregularity production surrounding a large-scale structure is found to exhibit both asymmetry and anisotropy.
+Radars are most likely to observe this asymmetry if the patch is elongated along the radar boresight.}
   \label{fig:conclusion}
@@
 
-Nighttime observations of FAIs is also higher with a positive IMF By component.
-This could be due to the control that the By component exerts over ionospheric convection patterns.
-In a convectional two-cell convection pattern with predominantly negative IMF Bz, the plasma flows antisunward from dayside to nightside across the terminator.
-In this arrangement, the plasma flow, \(\vec{V}_E\), and the density gradient, \(\vec{G}\), are antiparallel, the least favorable orientation for GDI growth.
-However, with a significant IMF By component, the orientation of the convection cells rotates such that plasma no longer convects directly antisunward, creating a situation where GDI is more favorable and plasma structures can develop more readily.
-Additionally, an IMF By component can contribute to the formation of polar patches by creating convection perturbations that separate high density patches from the daytime reservoir, as described in Section \ref{sec:lit_patches}.
-In this way, the IMF By component may be indirectly responsible for the increased observation of backscatter.
+One other result presented in Chapter \ref{sec:paper1} was that nighttime occurrence of FAIs was also higher for time intervals with a positive IMF \(B_y\) component, Figure \ref{fig:conclusion}.
+This could be due to the control that the \(B_y\) component exerts over ionospheric convection patterns, as explained below.
+In a classical two-cell convection pattern with predominantly negative IMF \(B_z\), the plasma flows antisunward from dayside to nightside across the terminator.
+In this configuration, the plasma flow, \(\vec{V}_E\), and the density gradient, \(\vec{G}\), are antiparallel, the least favorable orientation for GDI growth.
+However, with a significant IMF \(B_y\) component, the orientation of the convection cells rotates such that plasma no longer convects directly antisunward, creating a situation where GDI is more favorable and plasma structures can develop more readily.
+Additionally, a significant IMF \(B_y\) component can contribute to the formation of polar patches by creating convection perturbations that separate high density patches from the daytime reservoir, as described in Section \ref{sec:lit_patches}.
+In this way, the IMF \(B_y\) component may be indirectly responsible for the increased observation of backscatter.
 If patch occurrence increases, it can contribute to the nighttime background density, as described above, which can improve propagation conditions so that FAIs can be observed by
@@
 
-In addition to simply providing additional background density so nighttime propagation conditions are sufficient for the radar to observe FAIs, polar patches can contribute to plasma irregularity formation directly by providing background density gradients for GDI to be operational on.
+In addition to simply providing additional background density so nighttime propagation conditions are sufficient for the radar to observe FAIs, polar patches can contribute to plasma irregularity formation directly by providing background density gradients on which GDI can operate.
 This raises the important question of exactly how and where plasma structures form around polar patches.
-Observations have long suggested there is asymmetry in plasma structuring surrounding polar patches \citep{Weber1984,Milan2002b,Koustov2012}, but the results presented in Chapter \ref{sec:paper2} suggest that production may also be anisotropic, Figure \ref{fig:conclusion}.
-Asymmetry and anisotropy are independent such that the plasma structures can exhibit either of them independently or both or neither.
+Observations have long suggested there is asymmetry in plasma structuring surrounding polar patches \citep{Weber1984,Milan2002b,Koustov2012}, but the results presented in Chapters \ref{sec:paper2} and \ref{sec:paper3} suggest that production may also be anisotropic, Figure \ref{fig:conclusion}.
 
-The plasma drift and wavevector directions influence the asymmetry that is observed surrounding polar patches.
-The direction that the patch is drifting relative to its elongation direction impacts the orientations of \(\vec{G}\) and \(\vec{V}_E\) and controls the GDI growth rate.
-This also creates the asymmetry between the different edges of the patch because as the gradient direction changes, the irregularity growth rate changes.
-Additionally, linear theory predicts that irregularity growth is anisotropic, so the orientation of the patch relative to the radar can also impact the observed irregularities.
-In the \(F\) region, the GDI growth rate is highest for wavevectors perpendicular to the gradient, \(\vec{G}\).
-Although plasma waves can occur in all directions, a SuperDARN radar will only observe waves who's wavevectors are parallel to a radar beam.
-Because of this, irregularities as well as structure asymmetry are most likely to be maximized if the radar's boresight is parallel to the structure's elongation, Figure \ref{fig:conclusion}.
+%An important conclusion that can be derived from this work is, therefore, that asymmetry and anisotropy are independent such that the plasma structures can exhibit either of them independently or both or neither.
 
-Plasma irregularities are observed most often close to the terminator.
-This is likely due to the presence of a very large density gradient between high density solar illuminated plasma on the dayside and low density plasma on the nightside.
-The large background gradient promotes irregularity production through GDI.
+%The other group of results presented in Chapters \ref{sec:paper2} and \ref{sec:paper3} is related to the fact that the plasma drift and wavevector directions influence the asymmetry that is observed near polar patches.
+
+In Chapter \ref{sec:paper2}, it was demonstrated from the modeling point of view that the direction that the patch is drifting relative to its elongation direction impacts the orientations of \(\vec{G}\) and \(\vec{V}_E\) and controls the GDI growth rate.
+This also creates asymmetry between the leading and trailing edges of the patch because when the gradient direction switches, the irregularity growth rate changes from positive to negative (or vise versa) and irregularity damping occurs instead of growth.
+
+Additionally, linear theory predicts that the irregularity growth rate is depends on the wavevector direction, \(\uvec{k}\), so instability growth may anisotropic and the orientation of the patch and convection velocity relative to the radar, not just relative to each other, can also impact the observed irregularities.
+In Chapter \ref{sec:paper2}, it was demonstrated that, in the \(F\) region, the GDI growth rate is highest for wavevectors perpendicular to the gradient, \(\vec{G}\).
+Although plasma waves can occur in all directions, a SuperDARN radar will only observe waves whose wavevectors are antiparallel to a radar beam, due to the Bragg scatter condition.
+Because of this, irregularity growth rate as well as structure asymmetry are most likely to be maximized if the radar's boresight is parallel to the structure's elongation, Figure \ref{fig:conclusion}.
+This is further confirmed by the results of Chapter \ref{sec:paper3}, where the GDI growth rate was calculated from measurements of the density gradient, convection electric field, and wavevector direction.
+When the relative orientations of all three quantities results in a positive GDI growth rate, the occurrence of plasma irregularities is greater.
+An important conclusion that can be derived from this work is, therefore, that while asymmetry and anisotropy are independent processes, the small-scale plasma structuring surrounding polar patches seems to exhibit both.
+
+In addition to large-scale density structures such as polar patches, the solar terminator can also serve as a source of density gradients required for the gradient-drift instability to operate.
+In Chapter \ref{sec:paper1}, experimental evidence was presented that plasma irregularities are observed most often close to the terminator.
 In addition, if a SuperDARN radar itself is positioned close to the terminator, it is observing the density gradient along its boresight, Figure \ref{fig:conclusion}.
-As discussed above, if small-scale irregularity growth is directionally dependent, this is additionally the optimal orientation from which to observe the plasma structures, leading to increased echo occurrence.
+As was demonstrated in Chapter \ref{sec:paper2} and discussed above, if small-scale irregularity growth is directionally dependent, this is also the optimal orientation from which to observe the plasma structures, leading to increased echo occurrence.
+Thus modeling presented in Chapter \ref{sec:paper2} provide an alternative interpretation of earlier experimental results presented in Chapter \ref{sec:paper1}.
+
+%The production and characteristics of small-scale plasma irregularities and the factors that control them are complex.
+Radar studies must differentiate between production and propagation factors, determining if in fact irregularity production is more favorable or if background conditions just make observations of irregularities more likely.
+This thesis helps advance our understanding by identifying a number of situations where small-scale irregularities are observed and establishing plausible reasons why either production is favorable according to plasma theory or radar observation is likely.
 
@@
 
-Although much progress has been made in understanding small-scale plasma structuring and HF radar observations, many of the results presented in this thesis give rise to further questions that must be addressed.
+Although much progress has been made in understanding small-scale plasma structuring using HF radar observations, many of the results presented in this thesis give rise to further questions that must be addressed.
+In this section, four groups of possible follow-up studies are discussed.
+These incude
+\begin{enumerate*}[label={(\arabic*)}] 
+	\item statistical studies of polar patches using SuperDARN, 
+	\item modeling of the gradient-drift instability growth rates at decameter scales, 
+	\item multi-radar observations of irregularity characteristics near large-scale density gradients, and 
+	\item investigation of the role of fractional perturbation density in controlling the irregularity power and occurrence.
+\end{enumerate*}
 
-In Chapter \ref{sec:paper1}, the occurence of radar backscatter durring nighttime in the polar cap was found to be much higher than expected \citep{Bristow2011}, particularly given that raytracing simulations predict that the orthagonality condition is rarely met and therefore there should be little to no backscatter, Figure \ref{fig:month_avg_occ}.
-The proposed hypothesis to explain this discreperency is that polar patches, which are not typically well represented by ionospheric density models, occur frequently in the nighttime \(F\) region and, on average, enhance the background plasma density enough that propagation conditions are sufficient for HF radar backscatter to regularly be observed.
-Experimental surveys of polar patches have shown low occurrence at nighttimes, however the occurrence of patches and their intensity increase at equinox, which is where frequent nighttime backscatter is predominantly observed \citep{Rodger1996}.
-The expected signature in the radar data from this kind of senario would be a night with a series of isolate ``patches'' of backscatter.
+\subsection{Polar Patches in SuperDARN Observations}
+\label{sec:fw_patches}
+In Chapter \ref{sec:paper1}, it was found that the occurrence of radar backscatter during nighttime in the polar cap is much higher than expected, particularly given that raytracing simulations predict that the orthogonality condition is rarely met and therefore there should be little to no backscatter, Figure \ref{fig:month_avg_occ}.
+A hypothesis proposed in Chapter \ref{sec:paper1} was that polar patches, which are not typically well represented by ionospheric density models, occur frequently in the nighttime \(F\) region and, on average, enhance the background plasma density enough that propagation conditions are sufficient for HF radar backscatter to regularly be observed.
+Experimental surveys of polar patches have shown low occurrence at nighttimes, however the occurrence of patches and their intensity increase at equinox \citep{Rodger1996}, which is where frequent nighttime backscatter is predominantly observed according to our observations, Chapter \ref{sec:paper1}.
+The expected signature in the radar data from this kind of scenario would be a night with a series of isolated ``patches'' of backscatter.
 An example of such an event was presented, Figure \ref{fig:patchy_example}, but it is unknown how often events like this occur.
-Additionally, the intensity and size of a patch nessesary to cause the radar beam to refract sufficiently must be established, possibly through further use of raytracing simulations with customizable density profiles.
+Additionally, the intensity and size of a patch necessary to cause the radar beam to refract sufficiently must be established, possibly through further use of raytracing simulations with customizable density profiles.
 
-One of the major drawbacks to linear GDI theory is that it inherently assumes long wavelengths, which means if the \(F\) region, it technically should not be directly applicable to decameter-scale irregularities measured by HF radars.
+\subsection{GDI Theory and Modeling at Decameter Scales}
+\label{sec:fw_gdi}
+One of the major drawbacks to linear GDI theory is that it inherently assumes long wavelengths, which means that in the \(F\) region, it may not be directly applicable to decameter-scale irregularities measured by HF radars.
 Generally, it is assumed that some form of turbulent cascade connects plasma structures of different scale sizes in the ionosphere such that predictions made at large scales can still be applied to decameter-scale observations \citep{Tsunoda1988}.
-This helps explain some of the observations of asymmetry around polar patches, which is predicted at large scales but often observed at decameter scales with radars \citep[e.g.][]{Weber1984,Milan2002b,Koustov2012}.
-This asymmetry was also confirmed in the GDI growth rate from the results shown in Chapter \ref{sec:paper2}.
-However, these results also predicted anisotropy in the growth rate, which is more challenging to confirm from small-scale irregularity obsersavtions, although some evidence of such has been provided in Chapter \ref{sec:paper3}.
-It may be possible to further the theoretical understanding of GDI at small scales by developing a three dimentional picture of it using kinetic theory.
-\citet{Basu1995} developed such a formulation for a specific geometry throughout all altitudes within the ionosphere, but expanding this to general vector directions would not be trivial, mathematically.
-However, modeling irregularity growth with such a formulation may be helpful to understand how plasma structures develop at such small scales.
+This helps explain some of the observations of asymmetry around polar patches, which is predicted at large scales but often observed at decameter scales with radars \citep[e.g.][]{Milan2002b,Koustov2012}.
+This asymmetry was also confirmed from the results presented in Chapter \ref{sec:paper2}.
+However, these results also predicted anisotropy in the growth rate, which is more challenging to confirm from small-scale irregularity observations, although some evidence of such has been provided in Chapter \ref{sec:paper3}.
+It may be possible to further the theoretical understanding of GDI at small scales by developing a 3D picture using kinetic theory.
+\citet{Basu1995} developed such a formulation for a specific vector geometry throughout all altitudes within the ionosphere, but expanding this to general vector directions would not be trivial.
+However, modeling irregularity growth with such a formulation may be necessary in order to further our understanding of how plasma structures develop at such small scales.
 
+\subsection{Multi-Radar Observations of Irregularities and Density Gradients}
+\label{sec:fw_multiradar}
 The results presented in Chapter \ref{sec:paper3} could also be improved with a more extensive data set.
-The event chosen for analysis in that study was in 2011 and was selected on the basis of good data coverage from both RISR-N and RKN and a strong density gradient being apparent in the overlapping FoV.
-Since then, RISR-C has been deployed with a FoV just south of RISR-N (Figure \ref{fig:isrmap}) such that observations between both radars can be coordinated and they can act as a single instrument with a FoV twice as large.
+A single event chosen for analysis in that study was selected on the basis of good data coverage from both RISR-N and RKN and a strong density gradient being apparent in the overlapping FoV.
+A recent deployment of the southern-looking face of the Resolute Bay Incoherent Scatter Radar, also known as RISR-Canada or RISR-C, Figure \ref{fig:isrmap}, created new opportunities in this area.
+The observations between both radars can be coordinated and they can act as a single instrument with a FoV twice as large.
 This allows a polar patch to be tracked over a longer distance and more individual data points to be captured.
-Additionally, specific modes can be utilized that give a much finner grid of RISR beams, allowing measurements to be made closer together and smaller scale gradients to be be considered.
-The downside to these modes is that it takes longer for each set of measurements to be made, so the temporal cadence decreases.
-However, the required integration time for an ISR decreases if the background ionospheric density is higher, so the the temporal resolution can improve for measurements taken durring summer or daytime.
-Another advantage to the RISR-N/RISR-C locations is that the SuperDARN radars at Clyde River (CLY) and Inuvik (INV) (Figure \ref{fig:superdarnmap}), both of which share at least part of the RISR-N/RISR-C FoV, but observe it from different directions.
+Additionally, specific modes can be utilized that give a much finer grid of RISR beams, allowing measurements to be made closer together and smaller scale gradients to be evaluated.
+The downside to these modes is that it takes longer for each set of measurements to be made so that the temporal resolution becomes poorer.
+However, the required integration time for an ISR decreases if the background ionospheric density is higher, so the temporal resolution can improve for measurements taken during summer or daytime.
+Another advantage to the RISR-N/RISR-C setup is that the SuperDARN radars at Clyde River (CLY) and Inuvik (INV) (Figure \ref{fig:superdarnmap}) cover at least part of the RISR-N/RISR-C FoV and observe it from different directions.
 This presents an unique opportunity to further investigate the directional dependence of small-scale irregularities by observing the differences in backscatter observed from the same scattering volume at different directions.
-Although straight forward in theory, this experiment would most likely prove nontrivial in practice.
-Backscatter tends to be limited to the statistical high occurrence band shown in Figure \ref{fig:month_avg_occ}, and at least with INV, this band probably occurs at closer ranges than where the FoV overlaps other radars, so coincident backscatter is probably rare.
-The probability of backscatter being observed is better when the background density is within the range where propagation conditions are most favorable, when the ionosphere is sunlit.
+Although straight-forward in theory, this experiment would most likely prove nontrivial in practice for the following reasons.
+Backscatter tends to be limited spatially to the statistical high occurrence band shown in Figure \ref{fig:month_avg_occ}, and at least with INV, this band probably occurs at closer ranges than needed, so coincident backscatter is probably rare.
+The probability of backscatter being observed is better when the background density is within the range where propagation conditions are most favorable, i.e.
+when the ionosphere is sunlit for this case.
 Additionally, it is advantageous to choose a time when the occurrence of backscatter is statistically high.
-Based on results from Chapter \ref{sec:paper1}, the best time for such an experiment would therefore be daytime durring an equinox month.
+Based on results from Chapter \ref{sec:paper1}, the best time for such an experiment would therefore be daytime near equinox.
 
+\subsection{Irregularity Characteristics and Fractional Perturbation Density}
+\label{sec:fw_fracdens}
 In Chapter \ref{sec:paper3}, both the occurrence and power of radar backscatter are used as proxies to indicate the ``amount'' of irregularities that exist in the ionosphere.
@@
 The factors that control each need to be isolated, but this is nontrivial.
-Echo occurrence generally agrees with the linear GDI theory predictions but echo power only does for relatively low power.
-Irregularities may develop as predicted by linear GDI theory when their amplitudes are small but other nonlinear processes may impact larger density perturbations, such as shear-driven instabilities \citep{Gondarenko2006} or multistep processes \citep{Carlson2012}.
-By the radar equation, the power returned should be proportional to the product of the background density and fractional density, squared, Equation \ref{eqn:fractional_density}.
+Echo occurrence trends generally agree with the linear GDI theory predictions, but this is generally not the case for echo power.
+Irregularities may develop as predicted by linear GDI theory when their amplitudes are small but other processes may impact larger density perturbations, such as shear-driven instabilities \citep{Gondarenko2006} or multistep processes \citep{Carlson2012}.
+\(B_y\) the radar equation, the power returned should be proportional to the product of the background density and fractional density, squared, Equation \ref{eqn:fractional_density}.
 However, there is no general consensus on what controls the fractional density in the \(F\) region, although it has been suggested that it is either constant or a function of convection speed, \(V_E\) \citep{Kustov1988,Haldoupis1990,Makarevich2014b}.
-In Chapter \ref{sec:paper3}, we additionally propose it may be a function of the instability growth rate, \(\gamma\).
-Because ground based radars are so often used to investigate ionospheric irregularities, this is an important issue to resolve to develop an accurate picture of how radar observations relate to plasma irregularities.
+In Chapter \ref{sec:paper3}, we additionally propose that it may be a function of the instability growth rate, \(\gamma\).
+Because ground-based radars are so often used to investigate ionospheric irregularities, this is an important issue to resolve to develop an accurate picture of how radar observations relate to plasma irregularities.
+
+Our understanding of ionospheric plasma irregularities is in no way complete.
+The issues outlined in this section are a small subset of questions to be addressed by future research.
+Ground-based radar systems such as those employed extensively in this thesis will undoubtedly play an important role in this work.
+The widespread nature of the SuperDARN network makes it ideal for statistical studies of small-scale irregularities and mapping large-scale convection.
+Although scarcer, ISRs can provide a great wealth of data products, often in quite high resolution.
+Other instruments, e.g.
+satellites, ASIs, and ASIPs, can further supplement future studies of ionospheric irregularities.
+Equally important are models and theoretical studies of GDI as well as other instability mechanisms that provide insight as to how instabilities develop and explain observed trends.
+Ionospheric irregularities are part of a complex system and there are still many open questions about them, but the work presented here has advanced our understanding of their characteristics, development, and where and when we can expect to observe them.
 

### 2017-02-01 thesis.sh

@@ -20,6 +20,8 @@ bibtex paper1
 bibtex paper2
 bibtex conclusion
 
+# makeindex -s LLamarcheThesis.ist -t LLamarcheThesis.alg -o LLamarcheThesis.acr
+
 pdflatex LLamarcheThesis.tex
 
 pdflatex LLamarcheThesis.tex

### 2017-02-01 acronyms.tex
List of Acronyms added

% Acronyms

\begin{longtable}{ll}
\textbf{AACGM} & Altitude-Adjusted Corrected Geomagnetic coordinate system \\
\textbf{AC} & Alternating Code Pulse \\
\textbf{ACF} & Autocorrelation Function \\
\textbf{AMISR} & Advanced Modular Incoherent Scatter Radar \\
\textbf{ASI} & All-Sky Imager \\
\textbf{ASIP} & All Sky Imaging Photometer \\
\textbf{BSN} & Bow Shock Nose \\
\textbf{CCI} & Current-Conductive Instability \\
\textbf{CCMC} & Community Coordinated Modeling Center \\
\textbf{CLY} & Clyde River SuperDARN Radar \\
\textbf{CME} & Coronal Mass Ejection \\
\textbf{COSPAR} & Committee on Space Research \\
\textbf{CSR} & Coherent Scatter Radar \\
\textbf{EISCAT} & European Incoherent Scatter Association \\
\textbf{ESR} & European Scatter Radar \\
\textbf{FAI} & Field Aligned Irregularity \\
\textbf{FBI} & Farley-Buneman Instability \\
\textbf{FoV} & Field of View \\
\textbf{GDI} & Gradient Drift Instability \\
\textbf{GNSS} & Global Navigation Satellite System \\
\textbf{GPS} & Global Positioning System \\
\textbf{HF} & High Frequency \\
\textbf{IMF} & Interplanetary Magnetic Field \\
\textbf{INV} & Inuvik SuperDARN radar \\
\textbf{IRI} & International Reference Ionosphere \\
\textbf{ISR} & Incoherent Scatter Radar \\
\textbf{KHI} & Kelvin-Helmholtz Instability \\
\textbf{LoS} & Line of Sight \\
\textbf{LP} & Long Pulse \\
\textbf{MCM} & McMurdo Station SuperDARN radar \\
\textbf{MLAT} & Magnetic Latitude \\
\textbf{MLON} & Magnetic Longitude \\
\textbf{MSISE} & Mass Spectrometer - Incoherent Scatter Extended \\
\textbf{PFISR} & Poker Flat Incoherent Scatter Radar \\
\textbf{PCS} & Polar Cap South index \\
\textbf{RISR} & Resolute Bay Incoherent Scatter Radar \\
\textbf{RISR-C} & Resolute Bay Incoherent Scatter Radar Canada \\
\textbf{RISR-N} & Resolute Bay Incoherent Scatter Radar North \\
\textbf{RKN} & Rankin Inlet SuperDARN radar \\
\textbf{SFJ} & Sondrestrom \\
\textbf{STARE} & Scandanavian Twin Auroral Radar Experiment \\
\textbf{SuperDARN} & Super Dual Auroral Radar Network \\
\textbf{SZA} & Solar Zenith Angle \\
\textbf{TEC} & Total Electron Content \\
\textbf{URSI} & International Union of Radio Science \\
\textbf{UT} & Universal Time \\
\textbf{VHF} & Very High Frequency
\end{longtable}

### 2017-02-01 introduction.tex
List of Acronyms added

@@
 The collision frequency depends on the density of a species relative to the density of all other species in the plasma and can be calculated from a series of standard expressions \citep{Schunk1980,Schunk2009}.
-In the E region, \(\Omega_e \gg \nu_e\), so electrons are magnetized.
+In the \(E\) region, \(\Omega_e \gg \nu_e\), so electrons are magnetized.
 However, the motion on ions is dominated by collisions between ions and neutrals (\(\nu_i > \Omega_i\)), so ions in the \(E\) region are considered collisional.
@@
 The model requires the input of a location (latitude, longitude, and altitude) and time (year, date, and time).
-Additionally, the model gives the option to choose between two F peak models and three bottomside thickness models.
+Additionally, the model gives the option to choose between two \(F\) peak models and three bottomside thickness models.
 There are a variety of other optional input parameters such as sunspot number, ionospheric index, and F10.7 radio flux which can help make the output parameters more accurate for a particular situation.
@@
 Although it is fairly accurate at mid-latitudes \citep{Coisson2006,Bilitza2012}, it does not necessarily represent variations within the polar cap well \citep{Themens2014,Makarevich2015b}.
-In particular, the height of the F-region peak tends to be underestimated in daytime and the bottomside thickness does not show realistic seasonal and diurnal variation \citep{Themens2014}.
+In particular, the height of the \(F\)-region peak tends to be underestimated in daytime and the bottomside thickness does not show realistic seasonal and diurnal variation \citep{Themens2014}.
 

### 2017-02-01 paper1.tex
List of Acronyms added

@@
 In addition to direct effects of solar illumination, there is also interest in the indirect ways in which the Sun may control irregularity production in the ionosphere.
-It has been shown that negative IMF Bz generally leads to higher echo occurrence in the auroral \(F\) region \citep{Ballatore2001}.
+It has been shown that negative IMF \(B_z\) generally leads to higher echo occurrence in the auroral \(F\) region \citep{Ballatore2001}.
 An even stronger IMF control has been demonstrated for the auroral \(E\) region irregularities \citep{Makarevich2012}.
@@
 
-The overall aim of this study is to evaluate the extent of solar control on the production of \(F\)-region density irregularities in the polar cap ionosphere.
-Specifically, the objectives are to (1) quantify variations in \(F\) region backscatter location for different time sectors, seasons, and solar cycle phases, (2) analyze direct effects of solar illumination in irregularity production utilizing a comprehensive model of echo location, and (3)  establish the relationship between echo occurrence and IMF in the polar cap \(F\) region.
+The overall aim of this study is to evaluate the extent of solar control on the production of \(F\)-region density irregularities in the polar cap ionosphere. Specifically, the objectives are to 
+\begin{enumerate*}[label={(\arabic*)}]
+	\item quantify variations in \(F\) region backscatter location for different time sectors, seasons, and solar cycle phases,
+	\item analyze direct effects of solar illumination in irregularity production utilizing a comprehensive model of echo location, and
+	\item  establish the relationship between echo occurrence and IMF in the polar cap \(F\) region.
+\end{enumerate*}
 
@@
 The data were further shifted by 9--12 min to account for propagation from the BSN to the high-latitude ionosphere using a method described by \citet{Khan1999} that was also employed by \citet{Makarevich2012}.
-The IMF By and Bz components were averaged over 10-min intervals.
+The IMF \(B_y\) and \(B_z\) components were averaged over 10-min intervals.
 Any periods with IMF information missing were excluded from the analysis.
@@
 	\includegraphics[width=\textwidth]{p1f5.pdf}
-	\caption[Occurrence versus UT and IMF conditions]{The MCM echo occurrence versus UT color coded in the average (a) IMF By, (b) IMF Bz, and (c) PCS values.}
+	\caption[Occurrence versus UT and IMF conditions]{The MCM echo occurrence versus UT color coded in the average (a) IMF \(B_y\), (b) IMF \(B_z\), and (c) PCS values.}
 	\label{fig:ut_occ}
@@
 The data were binned using 1-h UT and 2\% occurrence bins and smoothed using a \(3\times3\) window.
-Figure \ref{fig:ut_occ} presents contour plots of echo occurrence  versus UT, with color representing average (a) IMF By, (b) IMF Bz, and (c) PCS index.
+Figure \ref{fig:ut_occ} presents contour plots of echo occurrence  versus UT, with color representing average (a) IMF \(B_y\), (b) IMF \(B_z\), and (c) PCS index.
 Figures \ref{fig:ut_occ}a and \ref{fig:ut_occ}b show almost no features in the first half of the day (0--12 UT; daytime observations).
 The second half of the day (12--24 UT; nighttime observations) exhibits a definite color structure for both IMF components.
-The echo occurrence was generally higher for more negative By ad Bz.
-Moreover, in Figure \ref{fig:ut_occ}a, the bins with IMF By below \(-2\) nT (blue colors) ar elocated at progressively lower occurrence values.
-This feature is less evident for IMF Bz, which indicates that IMF By component has a stronger association with the F-region echo occurrence than IMF Bz.
+The echo occurrence was generally higher for more negative \(B_y\) ad \(B_z\).
+Moreover, in Figure \ref{fig:ut_occ}a, the bins with IMF \(B_y\) below \(-2\) nT (blue colors) ar elocated at progressively lower occurrence values.
+This feature is less evident for IMF \(B_z\), which indicates that IMF \(B_y\) component has a stronger association with the F-region echo occurrence than IMF \(B_z\).
 In Figure \ref{fig:ut_occ}c, the average PCS index is somewhat enhanced before 12 UT, a result consisten with previous statistical analysis of the PCS index \citep[e.g.][Figure 4]{Troshichev2007}.
@@
 MCM routinely observes backscatter from ionospheric irregularities over extended time periods and large spatial areas within the southern polar cap (``persistent backscatter''), which allowed us to reliably quantify both backscatter location and occurrence characteristics and investigate their control by the solar wind and illumination.
-By considering both location and occurrence we achieved better understanding and differentiation between radio- and plasma-physical factors or, alternatively, between irregularity detection and production.
+\(B_y\) considering both location and occurrence we achieved better understanding and differentiation between radio- and plasma-physical factors or, alternatively, between irregularity detection and production.
 Below we further discuss three groups of issues aligned with three objectives of this study.
@@
 Figure \ref{fig:patchy_example} shows an example of conjugate measurements during two selected events that fall within the equinoctial period of Figure \ref{fig:month_avg_occ}a.
-Shown are IMF By and Bz components, Figure \ref{fig:patchy_example}a, MCM line-of-sight velocity in beam 14 versus UT and range, Figure \ref{fig:patchy_example}b, RKN velocity in nominally-conjugate beam 4, Figure \ref{fig:patchy_example}c, and northern face of the Resolute Bay incoherent scatter radar (RISR-N) measurements of electron density versus UT and altitude, Figure \ref{fig:patchy_example}d, on March 14, 2010.
+Shown are IMF \(B_y\) and \(B_z\) components, Figure \ref{fig:patchy_example}a, MCM line-of-sight velocity in beam 14 versus UT and range, Figure \ref{fig:patchy_example}b, RKN velocity in nominally-conjugate beam 4, Figure \ref{fig:patchy_example}c, and northern face of the Resolute Bay incoherent scatter radar (RISR-N) measurements of electron density versus UT and altitude, Figure \ref{fig:patchy_example}d, on March 14, 2010.
 Figures \ref{fig:patchy_example}e--\ref{fig:patchy_example}h present the same information but for a second event on April 6, 2010.
@@
 	\caption[Cojugate patchy day example]{Conjugate measurements for two selected events on (a--d) March 14, 2010 and (e-h) April 6, 2010.
-Shown are the IMF By (blue) and Bz (red) components (Figures \ref{fig:patchy_example}a and \ref{fig:patchy_example}e), MCM line-of-sight velocity versus UT and slant range in beam 14 (Figures \ref{fig:patchy_example}b and \ref{fig:patchy_example}f), RKN velocity in beam 4 (Figure \ref{fig:patchy_example}c and \ref{fig:patchy_example}g), and RISR-N electron density versus UT and altitude (Figure \ref{fig:patchy_example}d and \ref{fig:patchy_example}h).
+Shown are the IMF \(B_y\) (blue) and \(B_z\) (red) components (Figures \ref{fig:patchy_example}a and \ref{fig:patchy_example}e), MCM line-of-sight velocity versus UT and slant range in beam 14 (Figures \ref{fig:patchy_example}b and \ref{fig:patchy_example}f), RKN velocity in beam 4 (Figure \ref{fig:patchy_example}c and \ref{fig:patchy_example}g), and RISR-N electron density versus UT and altitude (Figure \ref{fig:patchy_example}d and \ref{fig:patchy_example}h).
 The color bar is the same for Figures \ref{fig:patchy_example}b, \ref{fig:patchy_example}f, \ref{fig:patchy_example}c, and \ref{fig:patchy_example}g and is given to the right of Figure \ref{fig:patchy_example}f.}
@@
 When IMF is relatively stable on a time scale of \(\sim\)20 min, as is the case in the first event, backscatter is persistent.
-When IMF exhibits strong perturbations in By with a period of \(\sim\)10 min, as is near 16 UT in the second event, nighttime backscatter is patchy.
+When IMF exhibits strong perturbations in \(B_y\) with a period of \(\sim\)10 min, as is near 16 UT in the second event, nighttime backscatter is patchy.
 Daytime backscatter seen by RKN during the same interval does not seem to be affected, i.e.
@@
 This study found no clear evidence on such a control in the dayside polar cap, Figure \ref{fig:ut_occ}.
-On the nightside (12--24 UT), however, a clear pattern emerged where echo occurrence was enhanced for negative IMF By and Bz, Figure \ref{fig:ut_occ}.
-Interestingly, however, the IMF Bz pattern was somewhat different from that of IMF By, Figure \ref{fig:ut_occ}, and in fact less clearly defined.
-Nevertheless, occurrence was generally higher for more negative IMF Bz, Figure \ref{fig:ut_occ}c.
-This result is consistent with the study by \citet{Ballatore2001} that also found higher occurrence for negative IMF Bz in the auroral zone.
-The new and unexpected result of the current study was that IMF By effects were more clearly defined in the occurrence of backscatter from the polar cap.
+On the nightside (12--24 UT), however, a clear pattern emerged where echo occurrence was enhanced for negative IMF \(B_y\) and \(B_z\), Figure \ref{fig:ut_occ}.
+Interestingly, however, the IMF \(B_z\) pattern was somewhat different from that of IMF \(B_y\), Figure \ref{fig:ut_occ}, and in fact less clearly defined.
+Nevertheless, occurrence was generally higher for more negative IMF \(B_z\), Figure \ref{fig:ut_occ}c.
+This result is consistent with the study by \citet{Ballatore2001} that also found higher occurrence for negative IMF \(B_z\) in the auroral zone.
+The new and unexpected result of the current study was that IMF \(B_y\) effects were more clearly defined in the occurrence of backscatter from the polar cap.
 Another new result of the current study was that the occurrence dependence on IMF was changing in a systematic way with UT on the nightside, Figures \ref{fig:ut_occ}.
@@
 High-latitude convection patterns are directly controlled by solar wind variations.
-Statistical convection patterns are derived from satellites and radars and are routinely parameterized by IMF By and Bz \citep{Haaland2007, Pettigrew2010}.
+Statistical convection patterns are derived from satellites and radars and are routinely parameterized by IMF \(B_y\) and \(B_z\) \citep{Haaland2007, Pettigrew2010}.
 One mechanism through which IMF-controlled variations in convection may be responsible for changes in occurrence of plasma irregularities involves changing orientations between plasma density gradients \(\vec{G}\) and convection velocity \(\vec{V}_E\) vectors.
@@
 Enhancement by sunlight is mostly associated with more favorable propagation conditions at intermediate SZA values and large-scale gradients near the solar terminator possibly serving as seeds for further structuring processes.
-\item Indirect control by the solar wind variations is limited to the nightside, with enhanced backscatter occurrence observed during intervals of negative IMF By and Bz.
+	\item Indirect control by the solar wind variations is limited to the nightside, with enhanced backscatter occurrence observed during intervals of negative IMF \(B_y\) and \(B_z\).
 The IMF control of small-scale irregularity production can occur through IMF-controlled variations in the high-latitude convection pattern that, in turn, lead to changes in mutual orientations between convection and prevalent gradient vectors, with plasma structuring favoring certain orientations.

### 2017-02-01 paper2.tex
List of Acronyms added

@@
 The aim of this study is to investigate the asymmetries between the GDI growth rates near gradient reversals that often occur in the vicinity of large-scale density structures.
-The specific objectives are to (1) develop a quantitative approach for comparing the growth rates on either side of the structure,  (2) analyse how the growth rates and asymmetries change for different orientations of the background density gradient and convection velocity and determine if there is an optimal radar orientation for which the asymmetry is the strongest, and (3) investigate how the growth rate and asymmetries change between the ionospheric \(E\) and \(F\) layers.
+The specific objectives are to
+\begin{enumerate*}[label={(\arabic*)}]
+	\item develop a quantitative approach for comparing the growth rates on either side of the structure,
+	\item analyze how the growth rates and asymmetries change for different orientations of the background density gradient and convection velocity and determine if there is an optimal radar orientation for which the asymmetry is the strongest, and
+	\item investigate how the growth rate and asymmetries change between the ionospheric \(E\) and \(F\) layers.
+\end{enumerate*}
 
@@
 Although polar patches can retain their form in the \(F\) region for several hours \citep{Hosokawa2014}, it is likely that the patch would move a significant lateral distance from the boresite.
-The structure may have to be observed by multiple SuperDARN radars from different directions unless a highly elongated patch (>1500 km) is observed, similar to that described in \citet{Hosokawa2014}.
+The structure may have to be observed by multiple SuperDARN radars from different directions unless a highly elongated patch (\(>1500\) km) is observed, similar to that described in \citet{Hosokawa2014}.
 

### 2017-02-01 abstract.tex
Corrections to Abstract and Intro

@@
 The occurrence trends for irregularities are analyzed statistically and a model is developed that describes the location of radar echoes within the radar's field-of-view.
-The trends are explained through variations in background density with solar illumination affecting radar beam propagation.
+The trends are explained through variations in background plasma density with solar illumination affecting radar beam propagation.
 However, it is found that the irregularity occurrence during the night is higher than expected from ray tracing simulations based on a standard ionospheric density model.
-The high occurrence at night implies an additional source of plasma density and it is proposed that large-scale density enhancements called polar patches may and be the source of this density.
+The high occurrence at night implies an additional source of plasma density and it is proposed that large-scale density enhancements called polar patches may be the source of this density.
 
@@
 It is concluded from this body of work that the production of small-scale plasma irregularities in the polar \(F\)-region ionosphere is controlled both by global factors such as solar illumination as well as local plasma density gradients and electric fields.
-In general, linear gradient-drift instability theory describes small-scale irregularity production well, particularly for low amplitude perturbations.
+In general, linear gradient-drift instability theory describes small-scale irregularity production well, particularly for low-amplitude perturbations.
 The production of irregularities is complex, and while ground-based radars are invaluable tools to study the ionosphere, care must be taken to interpret results correctly.

### 2017-02-01 introduction.tex
Corrections to Abstract and Intro

@@
 Plasmas are fluid masses of charged particles, often formed when neutral atoms and molecules in a gas are ionized by external energy.
-Instances of plasma include everything from florescent lightbulbs and flames to more exotic examples such as nuclear fusion reactor experiments, most stars, nebulas, and the interstellar medium.
+Instances of plasma include everything from fluorescent lightbulbs and flames to more exotic examples such as nuclear fusion reactor experiments, most stars, nebulas, and the interstellar medium.
 In fact, about 99\% of visible matter in the known universe exists in the plasma state.
-Of particular interest to this thesis is the plasma in Earth's ionosphere.
+Of particular interest to this thesis is the plasma in the Earth's ionosphere.
 
-Ionospheric plasma is partially ionized, meaning only a fraction of the particles are ionized and the rest are neutral.
-Considering the dynamics of plasmas therefore requires not only collisions between particles to be considered but also charge-charge interactions and external and internal electric and magnetic fields.
+Ionospheric plasma is partially ionized, meaning that only a fraction of the particles are ionized and the rest are neutral.
+Investigation of the dynamics of plasmas therefore requires not only collisions between particles to be considered but also charge-charge interactions and external and internal electric and magnetic fields.
 These dynamics can generate waves in the plasma, which cause density perturbations.
@@
 A large number of sunspots, high occurrence of fast and dense plasma outbursts, and a highly variable magnetic field indicate solar maximum.
-Conversely, solar minimum is characterized by a low sunspot number and slow and relatively steady outflow.
+Conversely, solar minimum is characterized by a low sunspot number and slow and relatively steady outflow of plasma.
 This thesis examines factors that control irregularity production in the Earth's ionosphere, which can be strongly influenced by solar activity.
@@
 Closed magnetic field lines from the Earth are shown in grey, while open field lines are shown in light blue.
-The ionosphere is shown in yellow above the Earth's surface.}
+The ionosphere is shown in yellow above the Earth's surface (not to scale).}
 	\label{fig:ste}
@@
 \label{sec:ionosphere}
-The focus of this body of work is the Earth's ionosphere, a region of the upper atmosphere that extends approximately between 50--1000 km in altitude where neutral gases have been partially ionized, resulting in free electrons and ions mixed with neutral particles.
+The focus of this body of work is the Earth's ionosphere, a region of the upper atmosphere that extends approximately 50--1000 km in altitude where neutral gases have been partially ionized, resulting in free electrons and ions mixed with neutral particles.
 Two competing processes occur in the ionosphere to create a peak in the electron density.
@@
 	\includegraphics[width=\textwidth]{densprofile.pdf}
-	\caption[Ionosphere density profile]{Altitude profiles of electron density throughout the ionosphere found from the IRI model.  Typical extents of the \(D\), \(E\), and \(F\) regions are identified.}
+	\caption[Ionosphere density profiles]{Altitude profiles of electron density throughout the ionosphere found from the IRI model.
+Typical extents of the \(D\), \(E\), and \(F\) regions are identified.}
 	\label{fig:densprofile}
@@
 This can be determined by comparing the collision frequency, \(\nu_\alpha\), to the gyrofrequency, \(\Omega_\alpha\), of a particular species, where \(\alpha\) refers to either electrons, \(e\), or ions, \(i\).
-If \(\Omega_\alpha \gg \nu_\alpha\), the particle completes many gyrations between each collision, so its motion is determined mostly by the magnetic field and it is magnetized.
-If \(\nu_\alpha \gg \Omega_\alpha\), the particle collides with other particles much more often than it complete a gyration, so it is considered collisional.
-The gyrofrequency is dependent on the mass and charge of the particle and the strength of the magnetic field, \(\Omega_\alpha = q_\alpha B/m_\alpha\), so is roughly constant throughout the ionosphere because the magnetic field strength does not change much throughout these altitudes.
+If \(\Omega_\alpha \gg \nu_\alpha\), the particle completes many gyrations between each collision, so its motion is determined mostly by the magnetic field and it is considered to be magnetized.
+If \(\nu_\alpha \gg \Omega_\alpha\), the particle collides with other particles much more often than it completes a gyration, so it is considered collisional.
+The gyrofrequency is dependent on the mass and charge of the particle and the strength of the magnetic field, \(\Omega_\alpha = q_\alpha B/m_\alpha\), so it is roughly constant throughout the ionosphere because the magnetic field strength does not change much throughout these altitudes.
 The collision frequency depends on the density of a species relative to the density of all other species in the plasma and can be calculated from a series of standard expressions \citep{Schunk1980,Schunk2009}.
 In the \(E\) region, \(\Omega_e \gg \nu_e\), so electrons are magnetized.
-However, the motion on ions is dominated by collisions between ions and neutrals (\(\nu_i > \Omega_i\)), so ions in the \(E\) region are considered collisional.
+However, the motion of ions is dominated by collisions between ions and neutrals (\(\nu_i > \Omega_i\)), so ions in the \(E\) region are considered collisional.
 In the \(F\) region where the neutral density is much lower, both ions and electrons are magnetized, \(\Omega_\alpha \gg \nu_\alpha\).
@@
 Intermediate-scale plasma irregularities are those with sizes of 100 m -- 30 km \citep{Kelley2009}.
-These are also referred to as scintillation-causing structures and are responsible for many of the aderse space weather effects that are observed on Earth.
-One important example of a space weather effect is when the ionosphere changes the phase and amplitude of the original signal, called radio scintillation, which affects the ability of satellites to communicate with ground receivers.
+These are also referred to as scintillation-causing structures and are responsible for many of the adverse space weather effects that are observed on Earth.
+One important example of a space weather effect is when the ionosphere changes the phase and amplitude of the original radio signal, called radio scintillation, which affects the ability of satellites to communicate with ground receivers.
 This can cause communication blackouts or introduce large errors into GPS/GNSS (Global Positioning System/Global Navigation Satellite System) calculations, affecting navigation.
@@
 Small-scale structuring is generally considered to be all irregularities less than 100 m \citep{Kelley2009}.
-Although small-scale irregularities are not responsible for scintillation that directly impact GPS/GNSS performance, they can be the easiest to study because of large datasets collected with a large number of ionospheric radars around the world that have been in operation for many decades (discussed further in Section \ref{sec:superdarn}).
+Although small-scale irregularities are not responsible for scintillation that directly impact GPS/GNSS performance, they can be the easiest to study because of extensive datasets collected with a large number of ionospheric radars around the world that have been in operation for many decades (discussed further in Section \ref{sec:superdarn}).
 Because electrons move far more easily along magnetic field lines than across them, density perturbations tend to be aligned with the magnetic field and are often known as field-aligned irregularities (FAI), Figure \ref{fig:polarirreg}d.
-The irregularity classifications presented here are rough categories only and a high degree of coupling between different scales occurs, which makes plasma structures of all scales (large, intermediate, and small) important for plasma structuring processes.
+The irregularity classifications presented here are rough categories only and a high degree of coupling occurs between different scales, which makes plasma structures of all scales (large, intermediate, and small) important for plasma structuring processes.
 For instance, large-scale structures such as polar patches may very well have internal intermediate- and/or small-scale structuring.
@@
 \label{sec:csr}
-In this thesis, we make extensive use of Coherent Scatter Radars (CSR), which have been used to study the ionosphere for the last half century.
+In this thesis, we make extensive use of coherent scatter radars (CSRs), which have been used to study the ionosphere for the last half century.
 The radar transmits a radio wave, which can scatter off plasma structures in the ionosphere, and then be detected by the receiving antennas.
@@
 This allows the distance between the radar and the target to be found (based on the time difference between when the pulse is transmitted and received).
-The size of the range gate is determined by the length of a radar pulse, Equation \ref{eqn:range_gate}
+The size of the range gate is determined by the length of a radar pulse,
 \begin{equation}
@@
 Conversely, if the radar only transmits a pulse of a certain length of a certain length and then ``listens'' for its return, the time between transmission and return is known and the distance the pulse traveled can be calculated and therefore the location of the backscattering volume can be found \citep{Farley1972,Greenwald1983}.
-However, using this simple single-pulse method another pulse cannot be transmitted until the first pulse returns, limiting the temporal resolution that can be achieved.
-This can be improved by using a multipulse scheme, which will be discussed below \citep{Farley1972,Greenwald1983,Greenwald1985}.
+However, using this simple single-pulse method, another pulse cannot be transmitted until the first pulse returns, limiting the temporal resolution that can be achieved.
+This can be improved by using a multi-pulse scheme, which will be discussed below \citep{Farley1972,Greenwald1983,Greenwald1985}.
 
 A short time between pulse returns is necessary to allow Doppler velocities to be calculated, particularly at long ranges.
-The maximum Doppler shifted frequency that can be observed by a radar is the Nyquist frequency, \(f_n\), which is equivalent to half the sampling frequency, \(f_s\), Equation \ref{eqn:nyquist}
+The maximum Doppler shifted frequency that can be observed by a radar is the Nyquist frequency, \(f_n\), which is equivalent to half the sampling frequency, \(f_s\),
 \begin{equation}	
@@
 \end{equation}
-Radars operating in single-pulse mode must have a very low sampling frequency, as described above, particularly at long ranges where the pulse must travel a long distance.
+Radars operating in a single-pulse mode must have a very low sampling frequency, as described above, particularly at long ranges where the pulse must travel a long distance.
 This severely limits the Doppler shift that can be measured, which places an upper limit on the LoS velocity that can be calculated with this technique.
@@
 The Nyquist frequency is then \(f_n = 37.5\) Hz, which corresponds to a maximum measurable Doppler velocity of \(\sim550\) m/s.
-Flows in the ionosphere have been known to exceed 1000--2000 m/s, so single pulse sampling is insufficient for these purposes.
+Flows in the ionosphere have been known to exceed 1000--2000 m/s, so single-pulse sampling is insufficient for these purposes.
 
-Instead of waiting for each individual signal to be received before transmitting the next, ionospheric radars typically employ a multipulse mode.
-In a multipulse mode, the radar transmits a series of pulses with different time intervals or lags between them.
+Instead of waiting for each individual signal to be received before transmitting the next, ionospheric radars typically employ a multi-pulse mode.
+In a multi-pulse mode, the radar transmits a series of pulses with different time intervals or lags between them.
 This can introduce complications when backscatter from different pulses at different ranges is received by the radar at the same time, referred to as cross-range interference, but these effects can be mitigated using correlation techniques \citep{Farley1972}.
@@
 The ACF is used to find the spectral characteristics of the backscatter, from which the LoS velocity, spectral width, and backscatter power can be obtained.
-Overall, multi-pulse techniques are generally considered far more appropriate for ionospheric studies than single pulse techniques \citep{Farley1972,Greenwald1983,Greenwald1985,Barthes1998,Ponomarenko2006}.
+Overall, multi-pulse techniques are generally considered far more appropriate for ionospheric studies than single-pulse techniques \citep{Farley1972,Greenwald1983,Greenwald1985,Barthes1998,Ponomarenko2006}.
 
@@
 Although the FoVs of other radars cover parts of the polar cap, these parts refer to either \(E\) region ranges or the furthest of the \(F\) region ranges, for which there is rarely much backscatter.
-There are currently three polar radars in the northern polar cap, but the FoV of RKN overlaps the viewing area of an important complimentary instrument, the Resolute Bay Incoherent Scatter Radar FoV, discussed in Section \ref{sec:isr}, which makes it particularly useful for multi-instrument studies of the ionosphere.
+There are currently three polar radars in the northern polar cap, but the FoV of RKN overlaps the viewing area of an important complementary instrument, the Resolute Bay Incoherent Scatter Radar, discussed in Section \ref{sec:isr}, which makes it particularly useful for multi-instrument studies of the ionosphere.
 
@@
 Using a model of chemical composition at different altitudes, the average ion mass can also be found.
-In addition, 3D convection velocities and electric field vectors can be derived from LoS ion velocity measurements in some of the more modern ISRs by assuming both \(\vec{E}\) and \(\vec{V}_E\) are constant across the radar's FoV \citep{Heinselman2008}.
+In addition, 3D convection velocities and electric field vectors can be derived from LoS ion velocity measurements in some of the more modern ISRs by assuming that both \(\vec{E}\) and \(\vec{V}_E\) are constant across the radar's FoV \citep{Heinselman2008}.
 More recently, a new algorithm has been developed that used the Lagrange method of undetermined multipliers to generate 2D convection velocity and electric field vectors without making this limiting assumption \citep{Nicolls2014}.
@@
 A map of the ISR systems currently operational near the northern polar cap is shown in Figure \ref{fig:isrmap}.
-The most recent advancement has been the development of Advanced Modular Incoherent Scatter Radars (AMISR) \citep{Heinselman2008}, of which there are currently three in use, one at the Poker Flat Rocket Range (PFISR), just north of Fairbanks, Alaska and two in Resolute Bay, Canada (RISR-N and RISR-C).
+The most recent advancement has been the development of Advanced Modular Incoherent Scatter Radar (AMISR) systems \citep{Heinselman2008}, of which there are currently three in use, one at the Poker Flat Rocket Range (PFISR), just north of Fairbanks, Alaska and two in Resolute Bay, Canada (RISR-N and RISR-C).
 The advantage of these new systems is that they are electronically steerable, unlike older ISRs which consisted of a large dish that had to be physically moved to change the beam direction.
@@
 
-AMISR systems typically have two types of pulses, a long pulse (LP) with 72-km range gates for \(F\)-region studies and an alternating code pulse (AC) with 4.5-km range gates gates for \(E\)-region studies.
+AMISR systems typically have two types of pulses, a long pulse (LP) with 72-km range gates for \(F\)-region studies and an alternating code pulse (AC) with 4.5-km range gates for \(E\)-region studies.
 The exact number and configuration of beams depends on the radar mode used and can be changed easily due to the system's electronic steering.
@@
 One of the most important advantages of ISR systems over CSRs is the ability to directly measure electron density in the ionosphere.
-However, because ISRs are much scarcer and are only typically run in WorldDay mode a few days out of every month, data are not as widely available.
-In addition, ISRs do not actually observe small-scale coherent plasma structures like the SuperDARN HF network does.
+However, because ISRs are much scarcer and most are only typically run a few days out of every month, data are not as widely available.
+In addition, ISRs do not actually observe small-scale coherent plasma structures like the SuperDARN array does.
 However, they can directly measure large-scale structures and image density variations on scales larger that 100 km (limited by the number of beams and range gate size).
@@
 The primary ISR used in the present volume of work is the north face of the AMISR system at Resolute Bay, Canada, RISR-N, shown in orange in Figure \ref{fig:isrmap}.
-The FoV of RISR-N overlaps that of the RKN SuperDARN radar, red outline in Figure \ref{fig:isrmap}, making it ideal for these types of comparison studies in the polar cap.
+The FoV of RISR-N overlaps that of the SuperDARN RKN radar, red outline in Figure \ref{fig:isrmap}, making it ideal for these types of comparison studies in the polar cap.
 The two experimental studies that employ both the SuperDARN RKN and RISR-N systems are presented in Chapters \ref{sec:paper1} and \ref{sec:paper3}.
@@
 Although it is fairly accurate at mid-latitudes \citep{Coisson2006,Bilitza2012}, it does not necessarily represent variations within the polar cap well \citep{Themens2014,Makarevich2015b}.
-In particular, the height of the \(F\)-region peak tends to be underestimated in daytime and the bottomside thickness does not show realistic seasonal and diurnal variation \citep{Themens2014}.
+In particular, the height of the \(F\)-region peak tends to be underestimated in daytime and the bottomside thickness does not show realistic seasonal and diurnal variations \citep{Themens2014}.
 
@@
 An example of an altitude profile of plasma density provided by the IRI model was shown in Figure \ref{fig:densprofile}.
-The displayed profiles shows how the electron density changes with altitude at local noon (day) and midnight (night) on January 1, 2007 at \(75\deg\) N, \(0\deg\) E, geographic.
+The displayed profiles show how the electron density changes with altitude at local noon (marked Day in Figure \ref{fig:densprofile}) and midnight (Night) on January 1, 2007 at \(75\deg\) N, \(0\deg\) E, geographic.
 
@@
 This can be done through standard raytracing tools based on numerical solutions to the Hamiltonian raypath equations, as shown in Chapter \ref{sec:paper1} \citep{Haselgrove1963,Jones1975}, however these tools require a 2D density profile along the path that the radar beam travels.
-It is possible to assume a Gaussian distribution or Chapman layer as a simple model of the F-region peak, but it is much more instructive to use output from the IRI model, as is done in this work, as it is sensitive to changes in latitude as well as seasonal and diurnal variations.
+It is possible to assume a Gaussian distribution or Chapman layer as a simple model of the \(F\)-region peak, but it is much more instructive to use output from the IRI model, as is done in this work, as it is sensitive to changes in latitude as well as seasonal and diurnal variations.
 
@@
 Large-scale structuring in the polar cap can take a variety of forms, Section \ref{sec:polar_structure}, but here the focus is on large regions of enhanced plasma density known as polar patches  \citep[e.g.][]{Weber1984,Weber1986,Buchau1983,Buchau1985}.
-One of the primary interests of this thesis is the relationships between irregularities of different scales and small-scale structuring.
+One of the primary interests of this thesis is the relationships between large-scale plasma structures ans small-scale irregularity formation.
 This structuring has long been observed around polar patches \citep{Weber1984,Milan2002b,Moen2012}.
@@
 Figure \ref{fig:magnetosphere} is a schematic of how the IMF, magnetosphere, and ionospheric polar cap are all interconnected.
-In general, magnetic field lines in the magnetosphere can either be closed (grey lines in Figure \ref{fig:magnetosphere}), meaning they connect only the north and south pole of the Earth, or open (light blue lines in Figure \ref{fig:magnetosphere}), when one of the ends of the field lines originates from the Sun such that the field line is connected directly to the IMF.
+In general, magnetic field lines in the magnetosphere can either be closed (grey lines in Figure \ref{fig:magnetosphere}), meaning they connect only the regions near the north and south pole of the Earth, or open (light blue lines in Figure \ref{fig:magnetosphere}), when one of the ends of the field lines originates from the Sun such that the field line is connected directly to the IMF.
 IMF lines that are not connected to the Earth are shown in dark blue in Figure \ref{fig:magnetosphere}.
@@
 In the polar cap at ionospheric altitudes, the magnetic field strength is relatively constant at around \(5\times 10^{-5}\) T, so any changes in magnetic flux must be due to changes in the polar cap area, \(A_{pc}\) due to the polar cap boundary expanding or contracting \citep{Lockwood1992a}.
-The polar cap boundary changes in response to magnetic reconnection between the magnetosphere and the IMF  on both the day- and night-sides changing the amount of open magnetic flux.
+The polar cap boundary changes in response to magnetic reconnection between the magnetosphere and the IMF  on both the day and night sides changing the amount of open magnetic flux.
 
@@
 In all panels, the dayside is the top of the figure and the green line shows the open-closed polar cap boundary, similar to Figure \ref{fig:magnetosphere}.
-Plasma flow contours are shown in black and the light orange region represents high density from the dayside reservoir.
+Plasma flow contours are shown in black and the orange region represents high density from the dayside reservoir.
 The reconnection burst starts at Figure \ref{fig:patch_formation}a and the increasing amount of open magnetic flux moves the dayside open-closed boundary equatorwards, as described above \citep{Cowley1991}.
@@
 Also operational in the \(E\) region but more dominant at higher altitudes is the gradient-drift instability (GDI) \citep{Simon1963,Hoh1963,Linson1970}.
-If a wave has a component along the magnetic field line, the current-conductive instability (CCI) can be operational even if GDI is not \citep{Hoh1960,Ossakow1979,Chaturvedi1981}.
+If a wave propagation vector has a component along the magnetic field line, the current-convective instability (CCI) can be operational even if GDI is not \citep{Hoh1960,Ossakow1979,Chaturvedi1981}.
 Additional instability mechanisms emerge if shears are considered, when \(\nabla E \neq 0\), such as the Kelvin-Helmholtz instability (KHI) \citep{Kintner1977,DAngelo1965}.
 
-In the polar F-region ionosphere, GDI is typically considered to be the dominant structuring process \citep{Weber1984,Cerisier1985,Basu1988,Tsunoda1988,Fukumoto2000}.
+In the polar \(F\)-region ionosphere, GDI is typically considered to be the dominant structuring process \citep{Weber1984,Cerisier1985,Basu1988,Tsunoda1988,Fukumoto2000}.
 GDI is operational when high-density perturbations in the plasma move to regions of lower background density and low-density perturbations in the plasma move to regions of higher background density, such that the wave amplitude grows relative to background conditions.
@@
 If the drift causes high- and low-density perturbations to move as described above, the instability is operational and the perturbations grow.
-If instead high density perturbations drift into regions of higher background density and low density perturbations drift into regions of lower background density, the wave is damped.
+If instead high-density perturbations drift into regions of higher background density and low-density perturbations drift into regions of lower background density, the wave is damped.
 
@@
 An even more general consideration involves arbitrary gradient and plasma drift directions \citep{Keskinen1982a,Makarevich2014c}.
-In this more general case, the directional dependence of GDI has been shown to be significant, such that the growth rate can actually change signs (determining whether wave growth or damping occurs) with different wavevector and gradient directions.
+In this more general case, the directional dependence of GDI has been shown to be significant, such that the growth rate can actually
+change signs (determining whether wave growth or damping occurs) with different wavevector and gradient directions.
 In addition, these expressions have historically been developed for a particular altitudinal regime, usually in the \(F\) region.
@@
 \begin{split}
-	b = -\frac{\vec{G}\cdot\vec{k}\times\uvec{b}}{k_\perp^2} \quad\quad&\quad\quad\quad\quad
-	y = \frac{k_\parallel}{k_\perp} \\
-	R = \frac{\sigma_H}{\sigma_P} = \frac{D_i+D_e}{1+\Psi} \quad\quad\quad\quad\quad
-	\psi =& -r_ir_e \quad\quad\quad\quad\quad
+	\Psi = -D_iD_e \quad\quad\quad\quad\quad\quad
+	\psi =& -r_ir_e \quad\quad\quad\quad
 	\hat{\psi} = \psi\left(1+\frac{k_\parallel^2}{r_e^2 k_\perp^2}\right) \\
+	b = -\frac{\vec{G}\cdot\vec{k}\times\uvec{b}}{k_\perp^2} \quad\quad\quad\quad\quad\quad
+	y =& \frac{k_\parallel}{k_\perp} \quad\quad\quad\quad\quad
+	R = \frac{\sigma_H}{\sigma_P} = \frac{D_i+D_e}{1+\Psi} \\
 	\vec{V}_d = \vec{V}_{e0}-\vec{V}_{i0} = (r_i-r_e)&\left[s_es_i(1+\psi)\left(R\vec{V}_E-\frac{\vec{E}}{B}\right)\right]
@@
 
+From Equation \ref{eqn:gdi}, the growth rate exhibits a clear dependence on the propagation direction in the field-perpendicular plane \(\uvec{k}\) and from heretofore, we will refer to this dependence as an \textit{anisotropy}  or the \textit{directional dependence}.
 Although linear GDI theory as described above predicts clear dependence on wavevector direction \citep{Makarevich2014c}, this theory is only valid in the long-wavelength limit, \(\vec{k}\cdot\vec{V}_E \ll \nu_\alpha\).
@@
 Later, 2D simulations were expanded to include magnetosphere-ionosphere coupling, but it was found that this only had a substantial effect for very large-scale structures \citep{Keskinen1990}.
-3D simulations were introduced of GDI surrounding a large plasma density enhancement \citep{Guzdar1998} and were later improved by including plasma dynamics parallel to the magnetic field and inertial effects so that secondary KHI and tertiary shear-driven instability processes could be modeled \citep{Gondarenko1999}.
+3D simulations of GDI were introduced surrounding a large plasma density enhancement \citep{Guzdar1998} and were later improved by including plasma dynamics parallel to the magnetic field and inertial effects so that secondary KHI and tertiary shear-driven instability processes could be modeled \citep{Gondarenko1999}.
 These simulations consistently showed asymmetry between the leading and trailing edge of large-scale density structures and fluctuations in plasma density that were as much as 10\%--20\% of the background plasma \citep{Gondarenko2004a,Gondarenko2004b}.
@@
 
-%/  Additionally, 2D power spectrum could be calculated and initially considered plasma waves along a large density structure but were later expanded to include magnetosphere-ionosphere coupling \citep{Keskinen1981,Keskinen1982,Keskinen1990}.
-
-%\citet{Keskinen1982} found plasma waves on the trailing edge of at large density enhancement are unstable.
-\citet{Keskinen1990} expanded these results by considering coupling between the magnetosphere and ionosphere, but found that this coupling only has a substantial effect for very large-scale structuring.
-In addition, this simulation predicted instability evolution that was far faster than what observations had shown.
-3D simulations of GDI surrounding a plasma density enhancement were introduced in \citet{Guzdar1998}.
-\citet{Gondarenko1999} improved this 3D simulation further by including plasma dynamics parallel to the magnetic field and inertial effects.
-These factors are necessary for the simulation to model secondary KHI and tertiary shear-driven instability processes.
-These simulations provided insight in how exactly GDI structuring develops.
-
-%Both the density and velocity spectra were found to be anisotropic, however increasing the impact of ion-inertial factors in simulations caused both spectra to become more isotropic \citep{Gondarenko2001}.
-
-%Fluctuations in density were found that were as much at 10--20\% of the background plasma density \citep{Gondarenko2004a}.
-Although asymmetry between structuring on the leading and trailing edges of a large-scale density structure was observed in all simulations, rapidly changing convection velocity could complicate this \citep{Gondarenko2004b}.
-For instance, if a large scale structure was drifting one direction and convection patters change suddenly so that it is now drifting in the opposite direction, the leading edge becomes the trailing edge and vice versa and both edges may exhibit some structuring.
-Additionally, both edges may become structured if large velocity shears are initially present surrounding the density structure \citep{Gondarenko2006}.
-Large shears additionally cause KHI to be operational as a primary structuring mechanism, but over time GDI will still be the dominant structuring process.
-
 \subsection{Observations of Polar Cap Structuring}
 \label{sec:lit_observations}
+A comprehensive review of polar patch observations has been provided by \citet{Crowley1996}.
+In the below overview, we focus on the observations that are most relevant to this thesis.
 Polar patches were first observed in the 1960s \citep{Hill1963}, but much of the work of characterizing them using optical and radio techniques was not accomplished until two decades later \citep{Weber1981,Weber1984,Weber1986,Buchau1983,Buchau1985}.
@@
 Patches tend to drift antisunwards at speeds of 500--1000 m/s, following background convection \citep{Buchau1983,Weber1984}.
-
-%\citet{Coley1998} examined the occurrence of polar patches in both the northern and southern hemispheres with DE-2 and found that
 
@@
 Patches convect from the dayside through the cusp/throat region towards the polar cap \citep{Kelly1984,Foster1984,Foster1985,Foster1993,Sojka1982,delaBeaujardiere1985}.
-A comprehensive review of polar patch observations has been provided by \citet{Crowley1996}.
-
-%Several studies had previously reported density enhanced patches convecting from the dayside through the cusp/throat region toward the polar cap \citep{Kelly1984,Foster1984,Foster1985,Foster1993}.
-This also agrees with observations made by \citet{Sojka1982,delaBeaujardiere1985}
-
-%\citet{Rodger1996} also considered occurrence of patches but with a HF radar in Halley, Antarctica.
-The diurnal variation in patch occurrence was found to peak at magnetic noon and the seasonal variation in equinox months.
-The only correlation between hourly averaged solar wind parameters that was found was a bias towards patches occurring with a negative IMF \(B_z\) component.
-
-%Over time, the trailing edge of a patch became steeper than the leading edge and irregularities developed on the trailing edge.
-
-%\citet{Weber1981} identified sun-aligned arcs that extend  along the sun-earth line through large parts of the polar cap using all-sky imaging photometers (ASIP).
-These arcs are over 1000 km long and \(\sim100\) km wide and are produced by soft electron precipitation.
-
-%Additionally, \citet{Buchau1983} found that polar patches were more likely to be observed for southward IMF \(B_z\) while sun-aligned arcs were more common for northwards \(B_z\).
-
-%\citet{Buchau1983} used a combinations of Digisonde and ASIP measurements to identify large luminous patches in the polar cap in addition to sun-aligned arcs.
-These patches could have densities up to \(10^6\) cm\(^{-3}\) and tended to drift antisunward.
-
-%\citet{Weber1984} found patches of enhanced ionization between 800--1000 km during moderately disturbed geomagnetic conditions, again using a combination of ASIP images, ionosonde measurements, and data from Dynamic Explorer 2 (DE-2).
-These findings confirmed that density enhancements tended to drift antisunward at speeds of 500-1000 m/s and patches were not colocated with enhanced particle within the polar cap, indicating that this is not the source of the density enhancement.
-Over time, the trailing edge of a patch became steeper than the leading edge and irregularities developed on the trailing edge.
-
-%Patches are not colocated with enhanced particle precipitation within the polar cap, indicating that this is not the source of the density enhancement \citep{Weber1984}.
-Conversely, \citet{Buchau1985} found that plasma densities within patches were comparable with those measured at dayside sub-auroral latitudes, suggesting that dayside solar-illuminated plasma is the source of dense plasma that form polar patches.
 
@@
 One of the key observations in this regard was that, over time, the trailing edge of a patch became steeper than the leading edge and irregularities developed on the trailing edge \citep{Weber1984}.
-Backscatter power from HF radars also tended to be greater on the trailing edges of both moving polar patches \citep{Milan2002b} and sun-aligned arcs identified from ASIP data \citep{Koustov2012}, which is generally attributed to GDI being unstable on the trailing edge but stable on the leading edge of such large, moving density structures.
+Backscatter power from HF radars also tended to be greater on the trailing edges of both moving polar patches \citep{Milan2002b} and sun-aligned arcs identified from ASIP data \citep{Koustov2012}, which is generally attributed to plasma being unstable to GDI on the trailing edge but stable on the leading edge of such large and moving density structures.
 Studies of patches using all-sky airglow imagers (ASI) found that the density gradient on the leading edge of polar patches tends to be 2--3 times steeper than that on the trailing edge \citep{Hosokawa2016}.
-Additionally, large finger-like structures on the trailing edge of patches have been identified that were tens to hundreds of kilometers in size and agreed with predictions made by GDI simulations \citep{Gondarenko2004b,Hosokawa2016}.
+Additionally, large finger-like structures on the trailing edge of patches have been identified by \citet{Hosokawa2016} that were tens to hundreds of kilometers in size and agreed with predictions made by GDI simulations \citep{Gondarenko2004b}.
 \citet{Moen2012} made direct measurements of plasma density structuring using the ICI-2 sounding rocket.
@@
 Recent radar observations suggested that plasma irregularity production in the polar cap may be much higher than at lower latitudes.
-When background plasma conditions are favorable for radar propagation in the polar caps, FAIs are observed nearly continuously \citep{Bristow2011}, however, there are still definite diurnal, seasonal, and solar cycle trends similar to what is observed at lower, auroral \citep{Kane2012}.
+When background plasma conditions are favorable for radar propagation in the polar caps, FAIs are observed nearly continuously \citep{Bristow2011}, however, there are still definite diurnal, seasonal, and solar cycle trends similar to what is observed at lower, auroral latitudes \citep{Kane2012}.
 There is an increase in echo occurrence at solar maximum, particularly in the midnight sector \citep{Milan1997,Koustov2004}.
@@
 Furthermore, models and simulations of ionospheric dynamics in the polar caps often incorporate plasma structuring details, such as instability growth mechanism.
-Although GDI is often considered the dominant instability mechanism in the \(F\) region, it is helpful to understand which other mechanisms may be operational and how they are interconnected.
+Although GDI is often considered the dominant instability mechanism in the \(F\) region, it is helpful to understand what other mechanisms may be operational and how they are interconnected.
 
@@
 Both global and local control of irregularity production is studied, i.e.
-we consider global solar control through solar illumination and solar wind ans well as much more local control by plasma density gradients and convection electric field.
-The majority of observational data have been measured by ground-based ionospheric radars, both CSR systems (Section \ref{sec:csr}) and ISR systems (Section \ref{sec:isr}), with particular emphasis on radars within the SuperDARN network, Section \ref{sec:superdarn}.
-These datasets have also been complimented with information from models as well as satellite data when necessary.
+we consider global solar control through solar illumination and solar wind as well as much more local control by plasma density gradients and convection electric field.
+The majority of observational data used in this thesis have been obtained by ground-based ionospheric radars, both CSR systems (Section \ref{sec:csr}) and ISR systems (Section \ref{sec:isr}), with particular emphasis on radars within the SuperDARN array, Section \ref{sec:superdarn}.
+These datasets have also been complemented with information from models as well as satellite data when necessary.
 Experimental results are considered in the context of a linear fluid theory of the gradient-drift instability.
@@
 The first objective of this body of work is to evaluate the extent of solar control over irregularity production in the polar \(F\) region.
-The occurrence of radar backscatter is considered in the contex both solar illumination (considered direct solar control) and IMF factors (considered indirect solar control).
+The occurrence of radar backscatter is considered in the context of both solar illumination (considered direct solar control) and IMF factors (considered indirect solar control).
 This gives a global or macrophysical view of where and when plasma irregularities are most likely to occur.
@@
 The second objective is to investigate asymmetry in the GDI growth rate surrounding large-scale density structures.
-A model is used to identifying where plasma irregularities occur surrounding a polar patch and how they can be observed by a ground-based HF radar.
-The model considers a large, elongated polar patch drifting in an arbitrary direction relative to a radar and assumes irregularity production is directly dependent to the linear GDI growth rate.
+A model is used to identify where plasma irregularities occur surrounding a polar patch and how they can be observed by a ground-based HF radar.
+The model considers a large, elongated polar patch drifting in an arbitrary direction relative to a radar and assumes irregularity production is directly dependent on the linear GDI growth rate.
 The model results are expected to contribute to more accurate interpretation of observations of irregularities near polar patches.
 
-The third objective is to investigate directional dependencies of small-scale plasma irregularities produced through the gradient-drift instability operating in the linear regiem.
+The third objective is to investigate directional dependencies of small-scale plasma irregularities produced through the gradient-drift instability operating in the linear regime.
 Two-dimensional measurements of density gradients and electric fields are made such that the directionally-dependent GDI growth rate can be calculated, which is then compared with plasma irregularity power and occurrence in a particular location.
@@
 The outline of this thesis is as follows.
-Chapter \ref{sec:introduction} provides an overview of the polar ionosphere and both plasma structuring mechanisms at both large- and small-scales.
-Chapter \ref{sec:paper1} presents an experimental investigation of decameter-scale plasma irregularities in the polar F-region and their solar control, with a particular focus on statistical analysis of data collected with a recently-deployed SuperDARN radar in the southern polar cap.
+Chapter \ref{sec:introduction} provides an overview of the polar ionosphere and plasma structuring mechanisms at both large and small scales.
+Chapter \ref{sec:paper1} presents an experimental investigation of decameter-scale plasma irregularities in the polar \(F\)-region and their solar control, with a particular focus on statistical analysis of data collected with a recently-deployed SuperDARN radar in the southern polar cap.
 The same results were also presented in a recent journal article \citep{Lamarche2015}.
@@
 These results have also been recently published \citep{Lamarche2016}.
-Chapter \ref{sec:paper3} presents an experimental investigation of the directional dependence of small-scale plasma structures by considering simultaneous measurements of electric field and density gradients from an AMISR  system and FAIs observed by SuperDARN in the context of predictions made by linear GDI theory and the model results presented in Chapter \ref{sec:paper3}.
+Chapter \ref{sec:paper3} presents an experimental investigation of the directional dependence of small-scale plasma structures by considering simultaneous measurements of electric field and density gradients from an AMISR  system and FAIs observed by SuperDARN in the context of predictions made by linear GDI theory and the model results presented in Chapter \ref{sec:paper2}.
 The same results have been submitted for publication \citep{Lamarche2017}.

### 2017-02-01 thesis.sh

@@ -20,8 +20,6 @@ bibtex paper1
 bibtex paper2
 bibtex conclusion
 
-# makeindex -s LLamarcheThesis.ist -t LLamarcheThesis.alg -o LLamarcheThesis.acr
-
 pdflatex LLamarcheThesis.tex
 
 pdflatex LLamarcheThesis.tex

### 2017-02-02 acknowledgements.tex

% Acknowlegements

First and foremost, I would like to thank my advisor, Dr. Roman Makarevich for his guidance throughout this project, which would not have been possible without his support, feedback, and constructive criticism.  I would also like to acknowledge the efforts of the other members of my committee, Dr. William Bristow, Dr. Mark Conde, and Dr. Hui Zhang for always providing new ideas and perspectives.  Additionally, I truely appreciate the support and encouragement I received from the entire physics department at UAF, particularly Dr. David Neuman, Dr. Renate Wackerbauer, Dr. Curt Szuberla, Dr. Chung-Sang Ng, Dr. Peter Delamere, Ellen Craig and Jeanie Talbot.  I would like to extend a particularly special thank you to Flora G**** and all the staff of the Keith B. Mather Libray for providing invaluable assistance in retrieving sources for my research, no matter how old and obscure my requests were.

To my family, thank you all for the supporting me for years, particularly my Mom, for making me a much better writer than I have any right to be as a scientist, my Dad, for always being there to answer my programming questions and helping me fix my code when I had now idea how I had broken it, and my brother, Henry, generally for just existing and providing a friendly voice of reason when I really needed it.  Additionally, thank you to my friends Katie, Ari, Bonnie, Heather, Jayne, and Stephanie for being there over these last few years despite being 4000 miles away, to Laura for giving me perspective and a laugh regardless of the time difference.  Finally, I would like to extend a very special thank you to Allison, Rosie, Emma, and Kinley on the homefront for their tireless effort in keeping me sane, healthy, and happy throughout this whole process.  I couldn't have done it without you guys.  Thank you.

### 2017-02-02 paper3.tex
Paper3 added to thesis

%% 11/23/2015
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% AGUJournalTemplate.tex: this template file is for articles formatted with LaTeX
%
% This file includes commands and instructions
% given in the order necessary to produce a final output that will
% satisfy AGU requirements. 
%
% You may copy this file and give it your
% article name, and enter your text.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% PLEASE DO NOT USE YOUR OWN MACROS
% DO NOT USE \newcommand, \renewcommand, or \def, etc.
%
% FOR FIGURES, DO NOT USE \psfrag or \subfigure.
% DO NOT USE \psfrag or \subfigure commands.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% All questions should be e-mailed to latex@agu.org.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% Step 1: Set the \documentclass
%
% There are two options for article format:
%
% 1) PLEASE USE THE DRAFT OPTION TO SUBMIT YOUR PAPERS.
% The draft option produces double spaced output.
% 
% 2) numberline will give you line numbers.

%% To submit your paper:
%\documentclass[draft,linenumbers]{agujournal}
%\draftfalse

%\usepackage{graphicx}
%\usepackage{amsmath}
%\usepackage{lineno}
%
%% For final version.
% \documentclass{agujournal}

% Now, type in the journal name: \journalname{<Journal Name>}

% ie, \journalname{Journal of Geophysical Research}
%% Choose from this list of Journals:
%
% JGR-Atmospheres
% JGR-Biogeosciences
% JGR-Earth Surface
% JGR-Oceans
% JGR-Planets
% JGR-Solid Earth
% JGR-Space Physics
% Global Biochemical Cycles
% Geophysical Research Letters
% Paleoceanography
% Radio Science
% Reviews of Geophysics
% Tectonics
% Space Weather
% Water Resource Research
% Geochemistry, Geophysics, Geosystems
% Journal of Advances in Modeling Earth Systems (JAMES)
% Earth's Future
% Earth and Space Science
%
%
%
%\journalname{JGR-Space Physics}
%
%
%\begin{document}
%
%% ------------------------------------------------------------------------ %%
%  Title
% 
% (A title should be specific, informative, and brief. Use
% abbreviations only if they are defined in the abstract. Titles that
% start with general keywords then specific terms are optimized in
% searches)
%
%% ------------------------------------------------------------------------ %%

% Example: \title{This is a test title}

%\title{Radar observations of density gradients, electric fields, and plasma irregularities near polar cap patches in the context of the gradient-drift instability}
\chapter{Radar observations of density gradients, electric fields, and plasma irregularities near polar cap patches in the context of the gradient-drift instability}

%% ------------------------------------------------------------------------ %%
%
%  AUTHORS AND AFFILIATIONS
%
%% ------------------------------------------------------------------------ %%

% Authors are individuals who have significantly contributed to the
% research and preparation of the article. Group authors are allowed, if
% each author in the group is separately identified in an appendix.)

% List authors by first name or initial followed by last name and
% separated by commas. Use \affil{} to number affiliations, and
% \thanks{} for author notes.  
% Additional author notes should be indicated with \thanks{} (for
% example, for current addresses).

% Example: \authors{A. B. Author\affil{1}\thanks{Current address, Antartica}, B. C. Author\affil{2,3}, and D. E.
% Author\affil{3,4}\thanks{Also funded by Monsanto.}}

%\authors{Leslie J. Lamarche\affil{1} and Roman A. Makarevich\affil{1}}

% \affiliation{1}{First Affiliation}
% \affiliation{2}{Second Affiliation}
% \affiliation{3}{Third Affiliation}
% \affiliation{4}{Fourth Affiliation}

%\affiliation{1}{Geophysical Institute, University of Alaska Fairbanks, 903 Koyukuk Drive, PO Box 757320, Fairbanks, AK, 99775-7320, USA}
%(repeat as many times as is necessary)

%% Corresponding Author:
% Corresponding author mailing address and e-mail address:

% (include name and email addresses of the corresponding author.  More
% than one corresponding author is allowed in this LaTeX file and for
% publication; but only one corresponding author is allowed in our
% editorial system.)

% Example: \correspondingauthor{First and Last Name}{email@address.edu}

%\correspondingauthor{Leslie Lamarche}{ljlamarche@alaska.edu}

%% Keypoints, final entry on title page.

% Example: 
% \begin{keypoints}
% \item	List up to three key points (at least one is required)
% \item	Key Points summarize the main points and conclusions of the article
% \item	Each must be 100 characters or less with no special characters or punctuation 
% \end{keypoints}

%  List up to three key points (at least one is required)
%  Key Points summarize the main points and conclusions of the article
%  Each must be 100 characters or less with no special characters or punctuation

%\begin{keypoints}
%\item Radar echo power of small-scale structures increases as large-scale gradient magnitude increases, implying a connection between scale sizes
%\item Directional dependence of small-scale structures is observed on all sides of a polar cap patch and not restricted to the leading/trailing edge 
%\item Gradient-drift instability theory describes echo occurrence behavior including its directional dependence better than echo strength 
%\end{keypoints}

%% ------------------------------------------------------------------------ %%
%
%  ABSTRACT
%
% A good abstract will begin with a short description of the problem
% being addressed, briefly describe the new data or analyses, then
% briefly states the main conclusion(s) and how they are supported and
% uncertainties. 
%% ------------------------------------------------------------------------ %%

%% \begin{abstract} starts the second page 
%\begin{abstract}
\section*{Abstract}
We present observations of plasma density gradients, electric fields, and small-scale plasma irregularities near a polar cap patch made by the Super Dual Auroral Radar Network radar at Rankin Inlet (RKN) and the northern face of Resolute Bay Incoherent Scatter Radar (RISN-N). RKN echo power and occurrence are analyzed in the context of gradient-drift instability (GDI) theory, with a particular focus on the previously uninvestigated 2D dependencies on wave propagation, electric field, and gradient vectors, with the latter two quantities evaluated directly from RISR-N measurements. It is shown that higher gradient and electric field components along the wavevector generally lead to the higher observed echo occurrence, which is consistent with the expected higher GDI growth rate, but the relationship with echo power is far less straight forward. The RKN echo power increases monotonically as the predicted linear growth rate approaches zero from negative values, but does not continue this trend into positive growth rate values, in contrast with GDI predictions. The observed greater consistency of echo occurrence with GDI predictions suggests that GDI operating in the linear regime can control basic plasma structuring, but measured echo strength may be affected by other processes and factors, such as multi-step or nonlinear processes or a shear-driven instability.
%\end{abstract}

%% ------------------------------------------------------------------------ %%
%
%  TEXT
%
%% ------------------------------------------------------------------------ %%

%%% Suggested section heads:
% \section{Introduction}
% 
% The main text should start with an introduction. Except for short
% manuscripts (such as comments and replies), the text should be divided
% into sections, each with its own heading.

% Headings should be sentence fragments and do not begin with a
% lowercase letter or number. Examples of good headings are:

% \section{Materials and Methods}
% Here is text on Materials and Methods.
%
% \subsection{A descriptive heading about methods}
% More about Methods.
% 
% \section{Data} (Or section title might be a descriptive heading about data)
% 
% \section{Results} (Or section title might be a descriptive heading about the
% results)
% 
% \section{Conclusions}

\section{Introduction}
\label{sec:p3intro}

Plasma irregularities in the ionosphere have been investigated for several decades using both in situ (sounding rockets and satellites) and remote (radars and optical instruments) techniques \citep[e.g. a comprehensive review by][]{Fejer1980}.  These observations revealed that ionospheric plasma is highly structured on scales ranging from a few centimeters to hundreds of kilometers.  The gradient-drift instability (GDI) has historically been considered one of the dominant drivers of this structuring at high latitudes \citep{Tsunoda1988}.  Many factors influence the strength of this instability, including plasma density, \(n\), density gradient, \(\nabla n\), electric field, \(\vec{E}\), and direction of the wavevector, \(\uvec{k}=\vec{k}/k\).  These factors can roughly be divided into two groups: density-related and electric-field-related which are further discussed below.

Plasma density can affect the observation of ionospheric structuring in a number of ways, including controlling both plasma physical factors, which regulate generation of plasma waves in the ionosphere, and propagation factors, which affects how well irregularities can be observed using radars \citep[e.g.][]{Danskin2002,Koustov2004,Lamarche2015}.  Enhanced echo power can be the result of both improved GDI conditions and a particular plasma density creating more favorable radar propagation conditions in the ionosphere  \citep[e.g.][]{Koustov2012}.  Enhanced background plasma density can smooth out existing gradients and reduce the effectiveness of GDI \citep{Ruohoniemi1997}.  However, enhanced plasma density can also increase radar backscatter in that the radar cross section is proportional to the amplitude of density perturbations squared, \(\delta n^2\), so if the fractional density \(\delta n/n\) is approximately constant, the cross section and backscatter power are effectively proportional to \(n^2\) \citep{Makarevich2014b}.  The direct effect of gradients in the plasma density has rarely been studied either in terms of the gradient magnitude \(\lvert\nabla n\rvert\), strength \(\lvert\nabla n\rvert/n\), or direction, in part because these are difficult to evaluate experimentally.

Observations of greater structuring on the trailing edge than the leading edge of polar cap patches provide one important example of how the direction of density gradients can be important for structuring \citep[e.g.][]{Weber1984,Milan2002b,Koustov2012,Moen2012}. Typically, this observation is interpreted in terms of GDI linear theory, according to which the growth rate is greatest when the gradient \(\nabla n\) is parallel to the drift velocity direction \(\vec{V}_E = \vec{E}\times\vec{B}/B^2\) and least when they are antiparallel.  Although previous studies have been successful in replicating this result in GDI models \citep{Gondarenko2004b,Gondarenko2006}, there have not been direct comparisons of the predicted and measured GDI growth rate dependencies on \(\nabla n\) and \(\vec{V}_E\) when both quantities are considered to be two-dimensional and independent.

Implicit in the above relationship  between gradient strength vector \(\vec{G} = \nabla n/n\) and the drift velocity \(\vec{V}_E\) (parallel or antiparallel) is an assumption about their typical mutual orientation as well as the associated assumption about directions of the peak growth rate in terms of both the wavevector and the gradient. This leads to significant simplifications in the GDI growth rate expression \citep{Linson1970,Tsunoda1988}. In context of large-scale density structures at high-latitudes drifting perpendicular to their direction of elongation \citep{Makarevitch2004c,Makarevich2015b} such as polar cap patches and sun-aligned arcs, \(\vec{G}\) and \(\vec{V}_E\) are parallel on the trailing edge and antiparallel on the leading edge. However, there can be large variations in this \citep{Makarevitch2004c} and, in general, the growth rate changes with the wavevector direction \citep{Keskinen1982a,Keskinen1983,Makarevich2014c}. More insight can therefore be gained by considering the directions of vector quantities instead of considering a vector configuration that results in the peak growth.

In this study, a general expression for the GDI growth rate is used that considers arbitrary directions of the density gradient \(\uvec{g}\), electric field \(\uvec{e}\), and wavevector \(\uvec{k}\), and is applicable throughout the\(F\) region \citep[Equation 23]{Makarevich2014c}.

%\begin{linenomath*}
\begin{equation}
  \label{eqn:gdi}
  \gamma = \frac{1}{B}\left(\uvec{k}\cdot\vec{E}\right)\left(\uvec{k}\cdot\uvec{b}\times\vec{G}\right) = GV_E\left(\uvec{k}\cdot\uvec{e}\right)\left(\uvec{k}\cdot\uvec{b}\times\uvec{g}\right)
\end{equation}
%\end{linenomath*}

Equation (\ref{eqn:gdi}) describes the dependence on the magnitudes of the electric field $E$ or, equivalently, convection speed \(V_E=E/B\) and the density gradient strength \(G\). In addition, it describes a dependence on the purely directional factors \(\uvec{k}\cdot\uvec{e}\) and \(\uvec{k}\cdot\uvec{b}\times\uvec{g}\). Past experimental investigations focused on a very limited subset of these dependencies.

The 2D electric field \(\vec{E}\)  and plasma drift velocity \(\vec{V}_E\) can be found from incoherent scatter radar (ISR) measurements, using the standard ISR method utilizing line-of-sight ion velocity data \citep{Hei08}.  Additionally, ISRs have been previously used to determine density patterns in the ionosphere \citep{Semeter2009,Dahlgren2012a,Dahlgren2012b}. \citet{Dahlgren2012b} qualitatively compared the location of irregularities detected by a coherent HF radar to high-density regions identified by an ISR.  The observations were described in the context of GDI predictions, but the growth rate has not been explicitly evaluated. The current study aims to fill this gap by calculating 2D density gradient fields based on ISR plasma density patterns. Previous studies that evaluated plasma density gradients have found them primarily from in situ measurements from sounding rockets \citep{Moen2012,Lynch2015}, but that method inherently only allows one component of the gradient to be resolved and only allows for measurements along the rocket's flight path.  The new method employed here allows a field of 2D gradient vectors to be calculated, which, together with the 2D electric fields also provided by ISR, allows for experimental investigations of plasma structuring in context of factors \(\uvec{k}\cdot\vec{E}\) and \(\uvec{k}\cdot\uvec{b}\times\vec{G}\) as well their directional equivalents \(\uvec{k}\cdot\uvec{e}\) and \(\uvec{k}\cdot\uvec{b}\times\uvec{g}\), Equation (\ref{eqn:gdi}).

In order to investigate the relationship between GDI growth rate and the degree of plasma structuring in the ionosphere, structuring must be quantified and measured.  Information on plasma structuring at small scales (\(\lambda \approx 10\) m) is widely available from the Super Dual Auroral Radar Network (SuperDARN), so this study will focus on small-scale irregularities.  Specifically, this study uses the SuperDARN radar at Rankin Inlet (RKN) to measure ionospheric echo strength and occurrence as an indication of the degree of plasma structuring in the ionosphere. The ISR at Resolute Bay, which will be used to determine electric fields and plasma density gradients, observes a portion of the RKN field-of-view (FoV) in the \(F\) region.  Although both echo occurrence and strength have been used as proxies for irregularity structuring \citep[e.g.][]{Milan2002b,Koustov2012}, any differences between these two indicators have not been analyzed, particularly in the context of the above-discussed factors that are expected to control the wave characteristics.

The aim of this study is to investigate the relationship between factors that control the directionally-dependent GDI growth rate and plasma structuring in the ionosphere by analyzing coincident measurements of echo occurrence, power, density gradients, and the electric field made by RKN and RISR-N.  The specific objectives are to 
\begin{enumerate*}[label={(\arabic*)}]
	\item evaluate how plasma density and density gradients affect the strength and occurrence of echoes measured by RKN,
	\item investigate the directional dependency of plasma structuring in the ionosphere in the context of predictions made by linear GDI theory, and
	\item determine if there is a difference between echo occurrence and power as observed by RKN and consider implications for plasma structuring.
\end{enumerate*}

\section{Experiment Configuration and Data Postprocessing}
\label{sec:experiment}
SuperDARN is a network of HF coherent radars that are designed to study plasma irregularities and large-scale plasma motion in the ionosphere \citep{Chisham2007}.  It consists of over 30 individual radars in both the northern and southern hemispheres at \mbox{mid-,} auroral, and polar latitudes.  Each radar has 16--24 beams, spanning \(\sim3.25\deg\) horizontally and \(\sim30\deg\) vertically, and 75--100 range bins, typically 45-km long.  A complete scan of all beams can occur in 60 s or 120 s, depending on the mode of operation.  Each radar measures Doppler velocity, signal-to-noise ratio (SNR) or power, and spectral width of ionospheric echoes using a 17-lag autocorrelation function.  A more detailed description of the SuperDARN network can be found in \citet{Chisham2007}.

In the current study, the ground scatter was eliminated using the standard criteria, \(|V-\Delta V| < 30\) m/s and \(|W-\Delta W| < 35\) m/s, where \(V\) is the Doppler velocity, \(W\) is spectral width, and \(\Delta V\) and \(\Delta W\) represent the uncertainty in these quantities, respectively.  To limit the contribution of interference and other spurious echoes, measurements were also eliminated if the SNR was less than 3 dB or the spectral width was greater than 500 m/s.  Further filtering of ground scatter was achieved by considering only echoes where the velocity measured by RKN was reasonably close to that measured by RISR-N, as described in Section \ref{sec:velocity}.

This study utilized the northern face of the Incoherent Scatter Radar (ISR) at Resolute Bay, Canada (RISR-N).  This is one of the Advanced Modular Incoherent Scatter Radar (AMISR) systems that consists of an electronically steerable phased array, allowing the radar to make measurements in different directions close to simultaneously by quickly cycling through all beam directions \citep{Bahcivan2010}.  Typically, AMISR systems have two types of pulses, a long pulse for \(F\)-region observations with 72-km range resolution and an alternating code pulse for \(E\)-region observations with 4.5-km range resolution.  The radar measures a power spectrum and the electron density, ion and electron temperatures, and ion drift velocity are found using a standard ISR technique \citep{Evans1969,Risbeth1985,Nicolls2007}.  Calibration accounts for the noise and system constants \citep{Nicolls2007} and is done using either plasma line measurements (for summer daytime) or ionosonde measurements (for all other times) \citep{Bahcivan2010,Themens2014a}.  3D convection velocity and electric field vectors in the $F$ region are also found by combining all ion drift components measured at a particular magnetic latitude (MLAT) and assuming that both \(\vec{E}\) and \(\vec{V}_E\) are constant in magnetic longitude (MLON) \citep{Heinselman2008}.

The common mode for RISR-N is the WorldDay mode, which is typically scheduled for 5--6 days per month.  The technical specification, such as number and orientation of beams, pulse sequences, and cadence, of the WorldDay mode changes slightly depending on the rendition of the mode that is currently being used.  For the event investigated in this study, RISR-N was run in WorldDay64m, in which 11 beams are used to collect data at approximately 70-s intervals.

Figure \ref{fig:f1} shows the FoV of both RKN, located at 62.82\(\deg\)N, 93.11\(\deg\)W, geographic (72.6\(\deg\)N, 26.4\(\deg\)W, geomagnetic), and RISR-N, located at 74.73\(\deg\)N, 94.91\(\deg\)W, geographic (82.8\(\deg\)N, 36.8\(\deg\)W geomagnetic).  Geomagnetic coordinates were determined using the Altitude Adjusted Corrected Geomagnetic Coordinates (AACGM) model \citep{Shepard2014}.  The RKN footprint consists of individual radar cells, outlined in dashed lines.  Cells where SNR \(>\) 3 dB are color-coded in SNR measured at 18:16--18:17 UT on November 30, 2011. The locations of the 11 RISR-N beams at 300 km are shown by circles, with the color of each circle representing the electron density, \(n\), measured by RISR-N at 300 km during the post-integration interval closest to that of RKN.

\begin{figure}
	\includegraphics[width=\textwidth]{fig1.pdf}
  \caption{Experimental setup showing the FoVs of RKN and RISR-N. Lines of constant magnetic latitude are shown at \(75\deg\)N, \(80\deg\)N, and \(85\deg\)N. The pink outline shows the footprint of the full FoV of RISR-N at 400 km.  Raw RISR-N electron density measurements (\(n\)) on November 30, 2011 at 18:16 UT at 300 km are shown with colored circles, with the color scale given in the top-right corner.  Filled contours show interpolated electron density at 300 km, using the same color scale.  The interpolated area has been restricted to the 10 southernmost beams. In addition, RKN cells are outlined.  Cells with a signal-to-noise ratio (SNR) greater than 3 dB are color coded in SNR, according to the left color bar in top-right corner. RKN beam 7, which lies along \(335\deg\)E MLON, is outlined in black.}
  \label{fig:f1}
\end{figure}

To calculate gradients, RISR-N density measurements at 300 km were first converted to a local Cartesian coordinate system and then interpolated to a grid with 10-km resolution in both the \(x\) and \(y\) directions.  The result of the interpolation is shown as filled contours in Figure \ref{fig:f1}, also colored in electron density according to the color bar in the top-right corner.  2D density gradient vectors were found by considering the average density change over two adjacent grid points in both the \(x\) and \(y\) directions.

To find coincident points between RKN and RISR-N data, simultaneous times were first found in both data sets.  RISR-N and RKN data points were considered coincident if the start time of the RISR-N integration window occurred within the 60-s interval of the RKN FoV scan.  Because RISR-N data had a longer time cadence than the RKN data, a limited number of RKN measurements did not have a coincident RISR-N measurement within 60 s; these were not considered. To find spatially coincident points, the point in the interpolated density grid (or constant MLON electric field measurement) that was closest to the center of each RKN radar cell was found.  Two points were only considered coincident if they were within 20 km of each other.

\section{Event Overview and Velocity Analysis}
The snapshot shown in Figure \ref{fig:f1} is part of the event on November 30, 2011.  This event was selected because it included a high-density structure that passed through the common FoV area of RKN and RISR-N.  The structure has a density several times higher than the surrounding plasma and seems to move coherently with the convection flow.  It is observed in the local noon sector moving approximately poleward (antisunwards), so it can be categorized as a polar cap patch. The event is summarized in Figure \ref{fig:f2}.  Figure \ref{fig:f2}a shows a north-south keogram of electron density as measured along \(335\deg\) MLON by RISR-N at 17:00--19:00 UT.  The high-density structure that is of particular interest occurred at \(\sim\)18:15 UT and is well defined by the two oblique black lines.

\begin{figure}
	\includegraphics[width=\textwidth,angle=180]{fig2.pdf}
  \caption{Left column is a keogram representation of (a and b) RISR-N electron density, (c) RKN line-of-sight velocity, and (d) RKN SNR for an event on November 30, 2011, 17:00--19:00 UT.  Color scales for all quantities are given to the right of the respective row. Keograms are taken at 335\(\deg\)E MLON and all RKN data are from beam 7.  Panel (b) also shows 2D plasma drift velocity derived from RISR-N measurements superimposed as vectors every 3 minutes.  Panels (c) and (d) also show contours of electron density plotted over RKN data. The high-density structure examined in this study is highlighted by the two oblique black lines between \(\sim\)18:00 and 18:30 UT. Panel (e) shows 16 snapshot 2D views of electron density, SNR, and plasma drift velocity within the time period highlighted with the pink bar at the top of panels (a)--(d).  Each RKN cell is color-coded in SNR. Interpolated electron density is shown as colored contour lines. Raw electron density measurements are the ten colored circles.  2D plasma drift velocity vectors are shown at \(335\deg\)E MLON.}
  \label{fig:f2}
\end{figure}

Figure \ref{fig:f2}b shows the same keogram, but with 2D velocity vectors calculated from RISR-N measurements superimposed over the density data every 3 min.  Velocity data is available at \(\sim\)1-min intervals and these higher-resolution data were used for later analysis, but a subset is shown here to keep the plot readable. In the structure region, plasma velocity is primarily northwards with a small eastwards component, indicating that the structure is moving north-east.  Figure \ref{fig:f2}c shows the line-of-sight velocity measured in RKN beam 7, which is aligned with the \(335\deg\) MLON meridian, Figure \ref{fig:f1}. Contours of electron density from RISR-N (same as in Figure \ref{fig:f2}a) are shown over the RKN data. Figure \ref{fig:f2}d has the same format as Figure \ref{fig:f2}c, but shows the SNR measured by RKN. Most of the RKN echoes occur along the steep transition between the high- and low-density regions.  The velocity data indicates that the structure was moving consistently away from the radar with a speed of \(\sim\)500 m/s.  If the structure is moving north-east, as indicated in Figure \ref{fig:f2}b, the greatest power is observed south of it, or on the trailing edge.

Figure \ref{fig:f2}e contains sixteen 2D snapshot views of the event at 18:09--18:27 UT (period shown by the pink bars at the top of Figures \ref{fig:f2}a--\ref{fig:f2}d). Each snapshot is similar to that shown in Figure \ref{fig:f1} with SNR measured by RKN shown by color-filled radar cells, RISR-N electron density measurements shown as colored circles, interpolated electron density as colored contours, and plasma velocity as a line of 2D vectors at \(335\deg\) MLON. In this view, the interpolated density contours are left unfilled so as to not obscure the RKN SNR data.

From velocity vectors in Figure \ref{fig:f2}e, the plasma drifts mostly northwards at the beginning of the time interval and more eastwards later during the event, similar to what is seen in Figure \ref{fig:f2}b. This results in the structure moving in a generally north-east direction, which can be also seen by looking at the interpolated density contours.  Thus, at 18:16 UT, the density peak of the patch is approximately centered in the panel, but it moves towards the top-right corner by 18:20 UT and has exited the panel to the top-right by 18:22 UT. In addition, the region where backscatter is observed by RKN (SNR \(>\) 3 dB) seems to follow the patch of high densities through the panels.  Between 18:09 and 18:12 UT, most of the observed backscatter occurs in the lower part of the panel, south of \(85\deg\) MLAT.  At 18:13--18:22 UT, the backscatter region moves northwards as the patch moves north-east.  This again indicates structuring on the trailing edge of the polar cap patch, while there is little evidence for a similar degree of structuring on the leading edge.  This effect is most obvious in panels for 18:19--18:21 UT, where the red circles and dark red contour are northward of the RKN cells filled with the red color.  That is, SNR is 40--50 dB on the trailing (south-west) edge, while SNR is closer to 20--30 dB on the leading (north-east) edge.

In finding coincident points between RISR-N and RKN, all possible spatial conjunctions are considered.  This generally includes the entire FoV of RISR-N with coincident RKN echoes.  From Figure \ref{fig:f1}, this includes RKN beams 4--10 and up to 9 range gates.  This allows large portions of the density structures to be considered.  In some cases, e.g. Figure \ref{fig:f2}e at 18:20 UT, the entire patch exists within the observable area and all edges can be considered.  In other cases, radars sample only parts of density structures.  To minimize possible biases in the orientation of the density structures and their edges with respect to RKN, this study also considers the data collected over the entire 24-h period of November 30, 2011.  However, certain biases may still be present because this study is dominated by a single strong density enhancement and the patch moved through the FoV with a relatively constant velocity.  These biases are discussed further in Section \ref{sec:discussion2}.

\label{sec:velocity}
In addition to the standard SuperDARN method for eliminating ground scatter, this study further required that the velocities measured by RKN and RISR-N be comparable, since low SuperDARN velocities are one of the characteristics of ground scatter.  Figure \ref{fig:f3} is a scatter plot of coincident RKN and RISR-N velocities.  In this figure and all subsequent ones in this paper, black points refer to all observations throughout the event where there were coincident RISR-N and RKN echoes and red points represent only those echoes observed within the structure identified by the two oblique black lines in Figure \ref{fig:f2}.  Because RKN only measures velocity along the radar beam, RKN velocities are compared with RISR-N velocities projected onto the appropriate RKN beam direction.

\begin{figure}
	\includegraphics[width=\textwidth,angle=180]{fig3.pdf}
  \caption{RKN line-of-sight velocity versus the projection of 2D RISR-N velocity onto the closest RKN beam. Black dots represent all coincident echoes that were observed by RKN and RISR-N simultaneously over the day considered, whereas red dots represent only echoes that occurred within the high-density structure.  The blue dashed line is the ideal coincidence line. The solid blue lines outline a region where the RKN velocity equals the RISR-N velocity within \(\pm\)500 m/s.}
  \label{fig:f3}
\end{figure}

In Figure \ref{fig:f3}, there is a clustering of all echoes (black) around (0 m/s, 0 m/s) and a clustering of structure echoes (red) around (\(-700\) m/s, \(-700\) m/s).  The dashed blue line shows where RKN velocity equals the projected RISR-N velocity.  There is a secondary population composed of both all echoes and structure echoes that is centered around RKN velocity 0 m/s and is spread between RISR-N velocity \(-1500\) m/s and \(-500\) m/s.  Although it is difficult to distinguish in this presentation, echo occurrence analysis (not shown) does show a depression in occurrence around (\(-500\) m/s, 0 m/s), so this value is used as a distinction between populations where RKN velocities approximately equal RISR-N velocity and populations where this condition is not met, probably more representative of ground scatter or other spurious echoes.  The two solid blue lines are the projected RISR-N velocity plus or minus 500 m/s, so all observations that fall between these two lines represent RKN velocities that are within 500 m/s of the corresponding RISR-N velocity.  Points with greater deviation (all points not within the two solid blue lines) are discarded from the following analysis, which effectively removed the most spurious measurements, while retaining a reasonable number of points.

\section{Plasma Irregularities and Density-Related Factors}
In order to investigate how density-related factors may affect echo power and occurrence, Figure \ref{fig:f4} shows scatter plots of echo power versus (a) electron density, \(n\), (b), electron density gradient magnitude, \(|\nabla n|\), and (c), and gradient strength, \(G = |\nabla n/n|\). Points are binned and averages (crosses) and standard deviations (shaded rectangles) are shown for both all echoes and only echoes within the structure.  Averages and standard deviations are only calculated if there are at least 3 points in a bin.  The thick black and red lines show count histograms for all echoes and only echoes within the structure, respectively.

\begin{figure}
\includegraphics[width=\textwidth,angle=180]{fig4.pdf}
  \caption{Scatter plots of RKN echo power versus RISR-N measurements of (a) electron density, \(n\), (b) density gradient magnitude, \(|\nabla n|\), and (c) gradient strength, \(G=|\nabla n/n|\).  Black dots represent all echoes observed over the day, whereas red dots represent echoes that occur within the high-density structure.  Points have been binned and crosses indicate the average power of each bin, while shaded rectangles show one standard deviation. Thick black and red lines are point count histograms; scale is on the right axis.}
  \label{fig:f4}
\end{figure}

In Figure \ref{fig:f4}a, most points are clustered at \(2\times10^{11}\)--\(5\times10^{11}\) \# m\(^{-3}\), corresponding to a peak in the all-counts histogram (black).  The structure-count histogram (red) has a secondary peak at this location, but its highest peak occurs at a much higher density, around \(1.1\times10^{12}\) \# m\(^{-3}\).  The latter feature is expected as the structure was specifically chosen to contain high density. The average power of all echoes increases approximately monotonically, particularly for densities above \(4\times10^{11}\) \# m\(^{-3}\). Average powers observed from structure echoes are typically higher than that for all echoes.  This is again expected because the structure was selected by a particular region of RKN data with high backscatter power.  The exception to this occurs at very high densities, above \(1\times10^{12}\) \# m\(^{-3}\), where there are few points with high densities that are not within the structure, so the two sets become very similar.  The average power of echoes within the structure does not increase monotonically, but instead appears to peak at around \(6\times10^{11}\) \# m\(^{-3}\).

Figure \ref{fig:f4}b shows the dependence on density gradient magnitude, \(|\nabla n|\).  For all echoes, counts peak at low gradients and decrease.  For structure echoes, counts peak for gradients at \(2\times10^5\) \# m\(^{-4}\), although the count peak is very broad.  Average power increases nearly monotonically with an increasing gradient magnitude for both all echoes and structure echoes, except for very large densities. The structure echo power is again higher.

Finally, Figure \ref{fig:f4}c shows the dependence on gradient strength, defined as \(G = |\nabla n/n|\).  In this case, both all-echo and structure-echo counts peak at about \(2\times10^{-6}\) m\(^{-1}\).  Similar to the density gradient, average power for both all echoes and structure echoes increases as gradient strength increases.  Interestingly, this trend is actually clearer and stronger in the gradient case (Figure \ref{fig:f4}b) than the gradient strength case (Figure \ref{fig:f4}c), which is unexpected as the GDI growth rate expression depends directly on the gradient strength, \(\vec{G}\), not the density gradient vector, \(\nabla n\), equation \ref{eqn:gdi}.  This feature is further discussed in Section \ref{sec:discussion1}.

\section{Plasma Irregularities and Directional Factors}
Figure \ref{fig:f5} contains similar scatter plots to Figure \ref{fig:f4}, but these demonstrate the dependence of echo power and counts on the two factors that directly impact the GDI linear growth rate, the rotated gradient component along the wavevector, \(\uvec{k}\cdot\uvec{b}\times\vec{G}\), and the electric field component along the wavevector, \(\uvec{k}\cdot\vec{E}\). An increase in either of these factors should theoretically be reflected in higher GDI growth rates, Equation \ref{eqn:gdi}, and this could occur through increase in the entire factor or its purely directional part, e.g. \(\uvec{k}\cdot\uvec{e}\), with of both of these possibilities investigated below. The wavevector direction \(\uvec{k}\) is determined from the RKN beam that viewed a particular scattering volume.  From the Bragg scatter condition, the received wave will have a wavevector pointing towards the radar and matching in direction the irregularity wavevector.

\begin{figure}
	\includegraphics[width=\textwidth,angle=180]{fig5.pdf}
  \caption{Scatter plots of RKN echo power versus RISR-N measurements of (a) the gradient strength component, \(\uvec{k}\cdot\uvec{b}\times\vec{G}\), (b) the electric field component \(\uvec{k}\cdot\vec{E}\), (c) the normalized gradient component, \(\uvec{k}\cdot\uvec{b}\times\uvec{g}\), and (d) the normalized electric field component, \(\uvec{k}\cdot\uvec{e}\). The meaning of dots, crosses, vertical bars, and lines is the same as in Figure \ref{fig:f4}.}
  \label{fig:f5}
\end{figure}

Power and counts are found from RKN, while the gradients, \(\vec{G}\), and electric fields, \(\vec{E}\), are 2D vectors evaluated by RISR-N, Section \ref{sec:experiment}. For brevity, the quantity \(\vec{G}\), previously described as the gradient strength, will be referred to simply as the gradient for the remainder of this paper.  In this experiment, the wavevector direction \(\uvec{k}\) is defined by the orientation of the RKN radar beam examining a particular scattering volume because the radar will only observe plasma waves propagating in this particular direction. Therefore, the first factor in equation \ref{eqn:gdi} is the component of the 2D electric field measured by RISR-N along the coincident RKN beam and the second factor is the component of the rotated 2D plasma density gradient along the coincident RKN beam.  Figure \ref{fig:f5}a shows the gradient component, \(\uvec{k}\cdot\uvec{b}\times\vec{G}\), and Figure \ref{fig:f5}b shows the electric field component, \(\uvec{k}\cdot\vec{E}\).  The purely directional parts of these factors or the normalized component of the gradient and electric field are shown in Figures \ref{fig:f5}c and \ref{fig:f5}d, respectively.  It is important to consider the normalized components independently to differentiate trends due to the directions of the electric field and gradients relative to the wavevector from those due to the magnitude of the electric field or gradient.

In Figure \ref{fig:f5}a, the average power of all echoes tends to increase as the gradient component increases from zero. At negative values, the power is smaller than at positive values and is approximately constant below 20 dB. The trend in the structure echoes is less obvious and it seems to increase as  \(\uvec{k}\cdot\uvec{b}\times\vec{G}\) increases up to \(6\times10^{-6}\) m\(^{-1}\), but then levels off or even decreases slightly.  Counts for both all echoes and structure echoes peak at 0, but the distribution of all echoes is roughly symmetric, whereas the count histogram of structure echoes is biased towards positive values of  \(\uvec{k}\cdot\uvec{b}\times\vec{G}\).

Figure \ref{fig:f5}b shows that, while the average power of all echoes is approximately constant with changing \(\uvec{k}\cdot\vec{E}\), the structure echoes exhibit a peak in power at \(\sim\)0.01 mV/m.
Both all echoes and structure echoes are biased in occurrence towards positive values of \(\uvec{k}\cdot\vec{E}\). The all-echo distribution peaks at \(\sim\)0 mV/m and falls off very quickly for negative values of \(\uvec{k}\cdot\vec{E}\). The structure-echo distribution grows steadily between $-$0.015 mV/m and $+$0.025 mV/m and then drops off suddenly due to low occurrence of high-electric-field measurements.

The purely directional part of the gradient, Figure \ref{fig:f5}c, does not appear to follow any trend in the average power of either all or structure echoes.  Structure average power is consistently higher, as expected, but both sets are approximately constant in power for all values of the normalized gradient component.  Importantly though, the occurrence of both all echoes and structure echoes peaks strongly at \(\uvec{k}\cdot\uvec{b}\times\uvec{g} = 1\) and exhibits a secondary peak at \(\uvec{k}\cdot\uvec{b}\times\uvec{g} = -1\).

The purely directional part of the electric field, Figure \ref{fig:f5}d, again has average power approximately constant for all echoes, but average power for structure echoes peaks at \(\sim\)0.4.  The occurrence distribution for all echoes again peaks strongly at \(\uvec{k}\cdot\uvec{e} = 1\) with a secondary peak at \(\uvec{k}\cdot\uvec{e} = -1\). Counts grow steadily between \(\uvec{k}\cdot\uvec{e} = -0.3\) and \(\uvec{k}\cdot\uvec{e} = 0.9\), leading up to the primary peak.  The occurrence distribution for structure echoes is biased towards positive values of \(\uvec{k}\cdot\uvec{e}\) (there are no structure echoes observed below $-$0.3) and peaks at \(\sim\)0.5.

Figure \ref{fig:f6} further examines the RKN echo power and occurrence versus the (a) dot product of the gradient and plasma drift velocity \(\vec{G}\cdot\vec{V}_E\) and (b) the expected GDI growth rate, with both evaluated from RISR-N measurements.  Similar to Figure \ref{fig:f5}, purely directional factors are shown for (c) the dot product of the gradient and the plasma drift velocity, \(\uvec{g}\cdot\mathbf{\hat{v}}_E\)  and (d) the GDI growth rate, given by \(\gamma^\prime \equiv \gamma/\left(GV_E\right) = (\uvec{k}\cdot\uvec{e})(\uvec{k}\cdot\uvec{b}\times\uvec{g})\), from equation (\ref{eqn:gdi}).  The normalized components are again important to determine if the relative directions of \(\vec{G}\) and \(\vec{V}_E\) impact the power and occurrence of backscatter independent of their magnitude.

\begin{figure}
	\includegraphics[width=\textwidth,angle=180]{fig6.pdf}
  \caption{Same as Figure \ref{fig:f5}, but for (a) the dot product of the gradient and plasma drift velocity, \(\vec{G}\cdot\vec{V}_E\), (b) the calculated GDI growth rate, \(\gamma\), (c) the normalized dot product of the  gradient and plasma drift velocity, \(\uvec{g}\cdot\mathbf{\hat{v}}_E\), and (d) the normalized GDI growth rate, \(\gamma'\).}
  \label{fig:f6}
\end{figure}

In Figure \ref{fig:f6}a, the trend in the average power of either all echoes or structure echoes is not monotonic, but structure echoes seem to show a small increase in average power as \(\vec{G}\cdot\vec{V}_E\) increases, and all echoes tend to have higher power for positive values of \(\vec{G}\cdot\vec{V}_E\) than for negative values. The occurrence distribution for all echoes is centered around 0, but counts are slightly higher for positive \(\vec{G}\cdot\vec{V}_E\).  Occurrence for structure echoes is biased towards positive values of \(\vec{G}\cdot\vec{V}_E\).  In Figure \ref{fig:f6}b, average power for all echoes is roughly constant and less than 20 dB when \(\gamma < 0\) s\(^{-1}\), but increases for positive values of \(\gamma\).  The occurrence distribution of all echoes is again centered around zero, but counts tend to be slightly higher for positive \(\gamma\).  The average power of structure echoes increases monotonically as \(\gamma\) goes from negative values to zero.  The behavior for positive \(\gamma\) is not monotonic, although the average power is always relatively high.

Turning to purely directional dependencies, in Figure \ref{fig:f6}c, average power for all echoes is relatively constant at \(\sim\)20 dB, while average power for structure echoes is higher, but also variable.  Both occurrence histograms for all echoes and structure echoes have a strong peak where \(\uvec{g}\cdot\mathbf{\hat{v}}_E = 1\), i.e. where the gradient is parallel to the drift velocity. The all-echo histogram also has a secondary peak at \(\uvec{g}\cdot\mathbf{\hat{v}}_E = -1\), i.e. where the gradient is antiparallel to the drift velocity.  Figure \ref{fig:f6}d shows that the average power is relatively insensitive to changes in the normalized growth rate, although power for both all echoes and structure echoes appears to have a peak around \(\gamma'=0.4\).  The occurrence histogram for all echoes shows that echoes are observed for all values of \(\gamma'\) and peak at zero, but the distribution is asymmetric such that more counts are observed for \(\gamma'\) positive than negative.  Structure echoes are only observed for \(\gamma'>-0.3\), and are biased towards positive \(\gamma'\).

Figure \ref{fig:f7} investigates the same factors as Figure \ref{fig:f5}, but here, scatter plots show the echo power (given as the color) versus the gradient component \(\uvec{k}\cdot\uvec{b}\times\vec{G}\) and electric field component \(\uvec{k}\cdot\vec{E}\), with the color bar shown to the right of the figure.  Figure \ref{fig:f7}a shows all echoes.  The solid black line in this and all other panels in this figure shows a linear fit to all points; equation is given at the bottom of the panel.  This linear fit includes all echoes considered in this study (SNR $>$ 3 dB).  The linear Pearson correlation coefficient of all points is given in the bottom-right corner of the panel.  Figure \ref{fig:f7}b shows power for all echoes versus purely directional factors \(\uvec{k}\cdot\uvec{b}\times\uvec{g}\) and \(\uvec{k}\cdot\uvec{e}\).  Figures \ref{fig:f7}c and \ref{fig:f7}d present the same information, but only structure echoes are shown.  Additionally in these two panels, the pink line shows the linear fit for points where SNR \(>\) 30 dB.  The equation of the line and linear correlation of this subset of points are again given at the bottom of each panel in pink.  In Figure \ref{fig:f7}c, the two linear fits are close to identical, so the pink line lies almost directly on top of the black line.

\begin{figure}
\includegraphics[width=\textwidth,angle=180]{fig7.pdf}
  \caption{Scatter plots of RISR-N measurements of the electric field component, \(\uvec{k}\cdot\vec{E}\), versus the gradient component, \(\uvec{k}\cdot\uvec{b}\times\vec{G}\).  Points are color coded in RKN SNR, according to the color bar to the right.  Panels (a) and (b) show all coincident echoes, whereas panels (c) and (d) show only echoes within the structure.  Panels (a) and (c) show the electric field and gradient component, whereas panels (b) and (d) show the normalized components, representing directional dependence.  The black line in each panel is the linear fit of all data points in the panel (SNR \(>\) 3 dB).  The pink line in panels (c) and (d) shows the same for a subset of points with high power (SNR \(>\) 30 dB).  Equations for each line are given in the bottom-left corner of each panel and the correlation values are shown in the bottom-right corner.}
  \label{fig:f7}
\end{figure}

In Figure \ref{fig:f7}a, points are clustered around \(\uvec{k}\cdot\uvec{b}\times\vec{G} = 0\) m\(^{-1}\) and \(\uvec{k}\cdot\vec{E} = 0.01\) mV/m.  Figure \ref{fig:f7}b has less localized clustering of points.  The bias of points toward positive values of \(\uvec{k}\cdot\vec{E}\) agrees with the results seen in Figures \ref{fig:f5}a and \ref{fig:f5}b.  As seen in Figure \ref{fig:f7}c as well as previous figures, structure echoes tend to have higher power.  The spread of the points is also substantially less for structure echoes than all echoes.  This can be seen in the higher linear correlation in Figure \ref{fig:f7}c than Figure \ref{fig:f7}a.  The same is true comparing Figures \ref{fig:f7}b and \ref{fig:f7}d, i.e. linear correlation is substantially higher in Figure \ref{fig:f7}d than Figure \ref{fig:f7}b. When only high power echoes are considered (SRN \(>\) 30 dB), linear correlations increase even more.  This can also be seen from a simple visual examination of Figures \ref{fig:f7}c and \ref{fig:f7}d, where the electric field component shows a clear increase with the gradient component, particularly for high-power echoes with SNR \(>\) 30 dB shown by red dots. Note that, in all cases, the relationship between the two factors is approximately linear and slope of the linear fit is positive within uncertainty. This finding is further discussed in Section \ref{sec:discussion2}.

\section{Discussion}
In this study, small-scale plasma irregularity characteristics were analyzed in the context of factors that affect the linear growth rate of the gradient-drift instability.  Previous experimental studies of GDI and its controlling factors focused on the convection electric field, electron density, fractional perturbation density, and density gradients derived from 1D measurements \citep[e.g.][]{Villian1986,Fukumoto1999,Fukumoto2000,Danskin2002,Oksavik2010,Moen2012,Makarevich2014b,Burston2016}.  In contrast, the current study focused on the previously uninvestigated directional dependence of radar echo occurrence and intensity and comparison with theoretical predictions for arbitrary gradient, electric field, and wave propagation vectors. To accomplish this, 2D density gradient vectors were estimated from multi-point and direct measurements of plasma density by an incoherent scatter radar and, in combination with electric field measurements by an ISR, used to evaluate the GDI growth rate.  In addition, differences between echo occurrence and backscatter power were examined in context of the GDI theory predictions. Below, we discuss three groups of issues aligned with the objectives of this study.

\subsection{Effects of Density-Related Factors}
\label{sec:discussion1}
The fluid linear theory of GDI predicts that an increase in plasma density gradient strength \(G\) will increase irregularity production through GDI \citep{Keskinen1983,Makarevich2014c}. The current experimental results were generally consistent with this expectation. In particular, the echo power showed an expected increase with \(G\), Figure \ref{fig:f4}c. The point scatter was large but the increase was monotonic for bins with significant statistics. One of  most interesting aspects of these observations arguably concerns small scales of irregularities considered (\(\lambda \approx 10\) m), as explained below.

The GDI growth rate applies to long wavelengths \(\lambda >1\) km in the $F$-region due to the fact that the fluid theory assumes the low-frequency (\(\omega \ll \nu_{in}\)) and long-wavelength (\(\lambda \gg 2\pi V_E\nu_{in}^{-1}\)) limits, where \(\nu_{in}\) is the ion-neutral collision frequency \citep[e.g.][]{Makarevich2014c}. In addition, density gradients estimated in this study refer to relative large scales \(L=G^{-1} =10\) km. Despite the differences in scales, strong similarities were observed between GDI predictions and measured irregularity characteristics at small scales. Because of these similarities, large- and small-scale irregularities are typically assumed to be closely related, possibly by a turbulent cascade \citep[e.g.][]{Tsunoda1988}. Our observations, including that of an increase in echo power with gradient strength \(G\), provide additional support for this idea.

Another factor that can affect both the irregularity occurrence and strength is the background plasma density \(n\) itself. To achieve perpendicularity to the magnetic field in the polar cap, HF radars rely on beam refraction, which is dependent on the plasma density \citep{Bristow2011,Koustov2014,Lamarche2015}.  High plasma density can also reduce the occurrence of backscatter observed by radar by smoothing the gradients required for GDI \citep{Ruohoniemi1997,Koustov2004}, through \(E\)-region shorting \citep{Dan02,Kan12,Lam15}, or through attenuation of the signal through a dense \(D\) region \citep{Danskin2002}.  The peak in occurrence of all echoes in Figure \ref{fig:f4}a agrees with previous results that there is an optimal density for HF radar observations of \(F\)-region irregularities \citep{Danskin2002,Makarevich2014b,Koustov2014,Lamarche2015}.

In the current study, however, a nearly monotonic increase in average power of all echoes with density was also found, Figure \ref{fig:f4}a, which can potentially be explained through radar backscatter theory. The backscatter power is known to be proportional to the perturbation intensity squared, \((\delta n)^2\), which can be represented as the product of the fractional perturbation density, \(\delta n/n\), and the background density, \(n\), squared.  If the fractional density is approximately constant, backscatter power is proportional to the background density squared, \(n^2\), which would result in a monotonic increase in backscatter power as \(n\) increases \citep{Kustov1988,Haldoupis1990,Makarevich2014b}. This explanation assumes a particular behavior of the fractional perturbation density \(\delta n/n\), with little knowledge currently being available either from experiment or theory and with other behaviors being possible, as elaborated on below.

The current study has also considered the effects of the magnitude of the density gradient, \(|\nabla n|\), and the gradient strength, \(G = |\nabla n/n|\), Figures \ref{fig:f4}b and \ref{fig:f4}c.  The average power of both structure and all echoes showed an increase with both of these factors.  However, the relationship between power and \(|\nabla n|\) was somewhat stronger than that between power and \(G\), particularly in the case of structure echoes. This is puzzling as the expression for the GDI growth rate, equation \ref{eqn:gdi}, is dependent on \(G\), not \(|\nabla n|\).  This may be due to the previously mentioned density effect. When background density, \(n\), is low, \(G\) is high, but propagation conditions may be insufficient for the radar to observe irregularities.  In other words, even if instability growth is high in the \(F\) region and the plasma is highly structured, the radar beam does not observe it because it has not been refracted enough to be perpendicular to the magnetic field, resulting in an artificial undersampling of points where \(G\) is large.  An alternative explanation can be offered using a suitably modified argument involving the fractional density. Thus if the fractional density is proportional to the growth rate, \(\gamma\), instead of being approximately constant, backscatter power should be proportional to \((\gamma n)^2\). Because the growth rate is directly proportional to the magnitude of the gradient strength, equation (\ref{eqn:gdi}), the power will be proportional to the magnitude of the density gradient squared:

\begin{equation}
	\label{eqn:densgrad}
	P(V_E,n) \propto \left(\frac{\delta n}{n}\right)^2 n^2 \propto \gamma^2 n^2 \propto G^2 n^2 = |\nabla n|^2
\end{equation}

From the current observations, it is not possible to conclude which of the three density-related parameters is most important and why, but they do suggest that both background density \(n\) and gradient strength \(G\) are important and that additional knowledge about fractional density behavior may be key to our understanding of complex interplay of the plasma- and radio-physical factors.

\subsection{Directional Effects of Electric Field and Gradient Components}
\label{sec:discussion2}
One important new aspect of the current study is the experimental investigation of the directional dependence of the GDI growth rate, as represented by equation  (\ref{eqn:gdi}). This directional dependence has been previously predicted theoretically \citep{Keskinen1982a,Makarevich2014c,Lamarche2016}, but no experimental evidence is currently available, apart from asymmetry between leading and trailing edges of large-scale density structures. Because the growth rate \(\gamma\) is directly dependent on both the gradient component, \(\uvec{k}\cdot\uvec{b}\times\vec{G}\), and the electric field component, \(\uvec{k}\cdot\vec{E}\), the relative directions of the gradient, electric field, and wavevector are predicted to strongly influence plasma structuring in the ionosphere, equation (\ref{eqn:gdi}).  A new method was used to find 2D plasma density gradients based on RISR-N measurements and these, along with electric field measurements, are used here to examine directional dependencies.

A summary of the results that agree with these predictions is presented in Table \ref{tab:results}.  The relationship between \(\vec{G}\) and \(\vec{V}_E\) described in entries 9 and 10 in Table \ref{tab:results} matches previously-reported observations of an asymmetry between the trailing and leading edge \citep[][]{Weber1984,Milan2002b,Koustov2012,Moe2012}. Most of these previous studies observed asymmetry only qualitatively and compared only the leading and trailing edge of patches instead of considering all directions around it.  The new approach employed here allowed a quantitative assessment using direct estimates from measurements. Apart from results 9 and 10, all other features in Table  \ref{tab:results} have not been reported before.

\begin{table}
\caption{A summary of the experimentally observed features that are consistent with the gradient-drift instability theory for arbitrary directions of the wavevector $\uvec{k}$, electric field $\uvec{e}$, and density gradient $\uvec{g}$.}
\label{tab:results}
\centering
\begin{tabular}{|r|l|c|}
	\hline
	\textbf{\#}& \textbf{Experimental Feature} & \textbf{Figure} \\
	\hline
	1&Increase in average power of all echoes with \(\uvec{k}\cdot\uvec{b}\times\vec{G}\) & \ref{fig:f5}a \\
	2&Greater occurrence of structure echoes when \(\uvec{k}\cdot\uvec{b}\times\vec{G}\) is positive & \ref{fig:f5}a, \ref{fig:f7}c \\
	3&Increase in average power of structure echoes as \(\uvec{k}\cdot\vec{E}\) crosses zero & \ref{fig:f5}b \\
	4&Greater occurrence of all echoes when \(\uvec{k}\cdot\vec{E}\) is positive & \ref{fig:f5}b, \ref{fig:f7}a \\
	5&Monotonic increase in occurrence of structure echoes with \(\uvec{k}\cdot\vec{E}\) & \ref{fig:f5}b \\
	6&Strong peak in occurrence when \(\uvec{k}\cdot\uvec{b}\times\uvec{g} = 1\) & \ref{fig:f5}c \\
	7&Increase in all echo occurrence leading to a peak at \(\uvec{k}\cdot\uvec{e}=1\) & \ref{fig:f5}d \\
	8&Greater occurrence of structure echoes when \(\uvec{k}\cdot\uvec{e}\) is positive & \ref{fig:f5}d \\
	9&Higher average power when \(\vec{G}\cdot\vec{V}_E\) is positive & \ref{fig:f6}a \\
	10&Strong peak in occurrence when \(\uvec{g}\cdot\mathbf{\hat{v}}_E = 1\) & \ref{fig:f6}c \\
	11&More high power echoes when \(\uvec{k}\cdot\uvec{b}\times\vec{G}\) and \(\uvec{k}\cdot\vec{E}\) are positive & \ref{fig:f7}a \\
	12&Greater occurrence of structure echoes when \(\uvec{k}\cdot\uvec{b}\times\vec{G}\) and \(\uvec{k}\cdot\vec{E}\) are positive & \ref{fig:f7}c \\
	\hline
\end{tabular}
\end{table}

Many of the results in Table \ref{tab:results} are based on Figure \ref{fig:f5}, which considers the dependence on the electric field term, \(\uvec{k}\cdot\vec{E}\), and the gradient term, \(\uvec{k}\cdot\uvec{b}\times\vec{G}\), of the growth rate independently.  Because the growth rate is the product of these two quantities, it can be challenging to interpret these plots, but they still provide useful insight into directional dependencies.  The two strong peaks at \(\uvec{k}\cdot\uvec{b}\times\uvec{g}=1\) and \(\uvec{k}\cdot\uvec{b}\times\uvec{g}=-1\) in Figure \ref{fig:f5}c indicate that many measurements were made when the gradient was perpendicular to the radar's boresight, \(\uvec{k}\perp\vec{G}\).  As discussed previously, the peaks in power in Figures \ref{fig:f5}b and \ref{fig:f5}d possibly indicate a preferred electric field direction for the event.  However, considering all echoes, a full range of electric field directions is observed (points exist for all values of \(\uvec{k}\cdot\uvec{e}\)), but there is a systematic increase in occurrence when \(\uvec{k}\cdot\uvec{e}>1\).  This indicates that irregularity growth is more favorable in the direction of the electric field, agreeing with previous results regarding the directional dependence of plasma irregularities \citep{Lamarche2016}.

In relation to result 10, the strong peak in echo occurrence at \(\uvec{g}\cdot\mathbf{\hat{v}}_E = 1\) corresponds to \(\vec{G} \parallel \vec{V}_E\), which is considered the most favorable orientation for GDI growth.  What was unexpected in the observations presented in Figure \ref{fig:f6}c is the secondary peak at \(\uvec{g}\cdot\mathbf{\hat{v}}_E = -1\), corresponding to \(\vec{G}\) antiparallel to \(\vec{V}_E\), which is the least favorable configuration for GDI growth. It is possible that this secondary peak is due to a different instability mechanism, but the directional dependencies of other mechanisms are also rarely considered \citep{Burston2016}.  \(\vec{G}\) antiparallel to \(\vec{V}_E\) corresponds physically to the leading edge of a drifting plasma patch.  Conventionally there is thought to be little to no structuring on the leading edge of drifting large-scale density structures, but some simulations have shown if there is a significant velocity shear along the edge of the patch, the leading edge can become structured, although this will happen  only after a long period of time \citep{Gondarenko2006}. Thus a secondary peak in echo occurrence on the leading edge of the currently examined density structure may be due to this, non-GDI mechanism.  Additionally, it is possible that the patch rotated before entering the FoV such that the leading edge became the trailing edge and vice versa, which would lead to both sides exhibiting structuring \citep{Oksavik2010}.  The magnitude of these two peaks may be due to some directional bias within the event.  For some portions of the event, the patch is completely within the radar's FoV so measurements are taken from all edges.  However, because the patch does move almost directly northwards through the FoV, predominantly the leading and trailing edges are seen by the radar at times.  This means that the leading and trailing edges may be oversampled, which could explain the sharp peaks in occurrence when \(\vec{G}\) parallel or antiparallel to \(\vec{V}_E\) relative to other directions.

One of the interesting features in Figure \ref{fig:f7} is the linear nature of the trend, particularly for the normalized components of high power echoes, Figure \ref{fig:f7}d.  GDI linear theory predicts a clustering of points where both \(\uvec{k}\cdot\uvec{b}\times\vec{G}\) and \(\uvec{k}\cdot\vec{E}\) are either positive or negative, or two clusters, one in quadrant I and one in quadrant III.  Figure \ref{fig:f7}c does show a relatively strong cluster of points in quadrant I but very few points in quadrant III.  Additionally, in Figure \ref{fig:f7}d, points appear to form a linear pattern, which is confirmed by the relatively high correlation coefficients.  A linear relationship between the electric field \(\vec{E}\) and density gradient \(\vec{G}\) is not necessarily expected, but if the majority of echoes are observed when \(\vec{G}\parallel\vec{V}_E\), the linear relationship between the normalized components does make some sense.  In this case \(\vec{E}\parallel\uvec{b}\times\vec{G}\), so a linear relationship between their normalized components is to be expected.

In this study, the plasma drift is assumed to be driven primarily by convection electric fields, with the neutral wind contribution assumed to be negligible.  In general, this is reasonable because the data collected refers to the \(F\)-region, where neutral wind speed is much smaller than that of the \(\vec{E}\times\vec{B}\) convection, particularly for the present dataset which included many convection velocity measurements exceeding 500 m/s, Figures \ref{fig:f2}c and \ref{fig:f3}.  If, however, neutral winds were to contribute significantly at these altitudes, the effective convection electric field would be reduced and a corresponding reduction in the calculated growth rates would occur, so the growth rates presented here would be overestimated.

\subsection{Differences Between Echo Power and Occurrence}
\label{sec:discussion3}
This study considered both echo power and occurrence, and it has been found that they do not necessarily follow the same trends.  This can be seen, for example, in Figure \ref{fig:f5}a, where the average power for all echoes gradually increases but occurrence peaks at zero, or in Figure \ref{fig:f5}b, where the average power for structure echoes peaks around 0.01 mV/m but occurrence increases monotonically to 0.025 mV/m.  In general, the echo occurrence seems to agree with GDI linear theory somewhat better than average power, Table \ref{tab:results}.  This is evident in Figures \ref{fig:f6}b and \ref{fig:f6}d, in which there is slightly greater occurrence for positive growth rates than negative growth rates, and structure echoes in particular are biased towards positive growth rates.  On the other hand, average power seems to either exhibit very little trend for positive growth rates or even decreases in the normalized case.

One possible conclusion that can be drawn from this is that GDI is responsible for a basic level of plasma structuring in the ionosphere, but as waves increase in amplitude, other processes may take effect.  This could be a nonlinear combination of the Kelvin-Helmholtz instability (KHI) and GDI \citep{Gondarenko2006} or some kind of multi-step plasma structuring process \citep{Carlson2007,Carlson2008}.  Turbulent processes have also been proposed as contributing as much if not more than GDI to plasma structuring in the polar ionosphere under certain conditions \citep{Burston2010,Spicher2015,Burston2016}.

Structure echoes do show a steady increase in average power while the growth rate is negative, Figure \ref{fig:f6}b, for both the non-normalized (between \(-0.001\) s\(^{-1}\) and 0  s\(^{-1}\)) and normalized cases (between \(-0.3\) and 0.1).  While this agrees well with the notion that increasing the growth rate should increase the strength of irregularities in the ionosphere, conventional understanding of a linear growth is that if the growth rate is negative, waves are damped, so there should be no structuring and no backscatter observed at all.

To further consider this, it is important to recall that the growth rate predicts only how fast instabilities will grow in the linear regime, not their current amplitude.  This means that points with a positive growth rate could have structures in the process of developing while points with a negative growth rate could have preexisting structures that are being damped, which could explain how some backscatter is observed even when the growth rate is negative.  The electric field and density gradients were measured spatially and temporally coincidently with the backscatter measurements, so the calculated growth rate represents the local growth rate at a particular moment in time and does not give any information about the history of growth of the irregularity.  The shortest rise times (\(1/\gamma\)) were still on the order of several minutes while the cadence of measurement was \(\sim1\) per minute, so it is reasonable that measurements of \(\gamma\) and echoes were made faster than the local response time.

With respect to the estimates of the growth rates, the highest rates were around 0.002 s\(^{-1}\), Figure \ref{fig:f6}c, which corresponds to a rise time of \(\sim8\) min.  However, the peak in echo occurrence occurred at \(\gamma\approx0.0001\) s\(^{-1}\), corresponding to a rise time of over 2.5 hours. In comparison with previous studies, \citet{Moe12} found through sounding rocket measurements \(\gamma = 0.021\) s\(^{-1}\) and a rise time of 47.6 s.  \citet{Carlson2007} estimated the rise time when GDI is the only instability mechanism operational as 5--10 min, but they also proposed a 2-step process in which KHI seeds GDI, in which case the rise time would be \(\sim50\) s.  \citet{Burston2016} found that on the edge of polar cap patches, the GDI growth rate is greater than 0.0017 s\(^{-1}\) (10-min rise time) approximately 81\% of the time and greater than 0.016 s\(^{-1}\) (1-min rise time) approximately 11\% of the time. The fastest growth rates estimated in the current study thus agree well with GDI growth rates estimated from previous studies as well as observations of how quickly structuring develops around polar cap patches \citep{Carlson2007}.  Overall, it appears that structuring can be directionally dependent and echo power and occurrence are not necessarily related in a simple way.

\section{Conclusions}
Signatures of plasma structuring in the polar \(F\)-region ionosphere have been investigated in the context of factors that control directionally-dependent growth rate of the gradient-drift instability.  The GDI growth rate was evaluated using 2D density gradient and electric field vectors measured by the northern face of the Resolute Bay Incoherent Scatter Radar, while information about structuring was inferred from echo occurrence and power measured by the SuperDARN radar at Rankin Inlet.  The findings are as follows:
\begin{enumerate}
	\item Echo power in general increases as plasma density, \(n\), density gradient magnitude, \(|\nabla n|\), and gradient strength, \(G = |\nabla n/n|\), all increase.  Small-scale structuring increases with large-scale gradients, implying strong coupling between different scale sizes.  The relationship between echo power and density gradient magnitude is stronger than that with the gradient strength, in contrast with GDI linear theory prediction, possibly suggesting that the fractional density is directly related to the growth rate.
	\item Small-scale plasma structuring is observed on all edges of a polar cap patch, with plasma irregularity characteristics exhibiting a dependence on the edge orientation with respect to the propagation direction. Stronger structuring is not restricted to the trailing edge and generally observed when the wavevector is oriented in a direction relative to the density gradient or electric field where irregularity growth rate is more favorable according to a general GDI theory.  Echo occurrence peaks when the relative orientations of the gradient and plasma drift velocity is most favorable (parallel), however, there is also a strong secondary peak when the orientations are least favorable (antiparallel) and the gradient-drift instability should be linearly damped.  This could be explained by a strong velocity shear causing structuring on the leading edge of the patch.
	\item Echo occurrence tends to follow the trends predicted by the linear theory of the gradient-drift instability better than echo power, suggesting that while GDI may be a dominant process in generating small-scale irregularities throughout the ionosphere, other mechanisms such as shear-driven KHI or nonlinear processes may exert greater control over the strength of those irregularities.
\end{enumerate}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%  ACKNOWLEDGMENTS
%
% The acknowledgments must list:
%
% •	All funding sources related to this work from all authors
%
% •	Any real or perceived financial conflicts of interests for any
%	author
%
% •	Other affiliations for any author that may be perceived as
% 	having a conflict of interest with respect to the results of this
% 	paper.
%
% •	A statement that indicates to the reader where the data
% 	supporting the conclusions can be obtained (for example, in the
% 	references, tables, supporting information, and other databases).
%
% It is also the appropriate place to thank colleagues and other contributors. 
% AGU does not normally allow dedications.

\section*{Acknowledgments}
This work was supported by NSF grants AGS-1248127 and PLR-1443504.  SuperDARN data are freely available through the SuperDARN website at Virginia Tech \url{http://vt.superdarn.org/}.  The Resolute Bay Incoherent Scatter Radar (North) is operated by SRI International on behalf of the U.S. National Science Foundation under NSF Cooperative Agreement AGS-1133009.  The data are accessible from the SRI International online database at \url{http://amisr.sri.com/database/}.

%%  REFERENCE LIST AND TEXT CITATIONS
%
% Either type in your references using
%
% \begin{thebibliography}{}
% \bibitem[{\textit{Kobayashi et~al.}}(2003)]{R2013} Kobayashi, T.,
% Tran, A.~H., Nishijo, H., Ono, T., and Matsumoto, G.  (2003).
% Contribution of hippocampal place cell activity to learning and
% formation of goal-directed navigation in rats. \textit{Neuroscience}
% 117, 1025--1035.
%
% \bibitem{}
% Text
% \end{thebibliography}
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Or, to use BibTeX:
%
% Follow these steps
%
% 1. Type in \bibliography{<name of your .bib file>} 
%    Run LaTeX on your LaTeX file.
%
% 2. Run BiBTeX on your LaTeX file.
%
% 3. Open the new .bbl file containing the reference list and
%   copy all the contents into your LaTeX file here.
%
% 4. Run LaTeX on your new file which will produce the citations.
%
% AGU does not want a .bib or a .bbl file. Please copy in the contents of your .bbl file here.

%\bibliography{myrefll}
%
%\providecommand{\ddegrees}{\ensuremath{\deg}}
%\begin{thebibliography}{46}
%\providecommand{\natexlab}[1]{#1}
%\expandafter\ifx\csname urlstyle\endcsname\relax
%  \providecommand{\doi}[1]{doi:\discretionary{}{}{}#1}\else
%  \providecommand{\doi}{doi:\discretionary{}{}{}\begingroup
%  \urlstyle{rm}\Url}\fi
%
%\bibitem[{\textit{Bahcivan et~al.}(2010)\textit{Bahcivan, Tsunoda, Nicolls, and
%  Heinselman}}]{Bah10}
%Bahcivan, H., R.~Tsunoda, M.~Nicolls, and C.~Heinselman (2010), Initial
%  ionospheric observations made by the new {Resolute} incoherent scatter radar
%  and comparison to solar wind {IMF}, \textit{Geophys. Res. Lett.},
%  \textit{37}, L15103, \doi{10.1029/2010GL043632}.
%
%\bibitem[{\textit{Bristow et~al.}(2011)\textit{Bristow, Spaleta, and
%  Parris}}]{Bri11}
%Bristow, W.~A., J.~Spaleta, and R.~T. Parris (2011), {First observations of
%  ionospheric irregularities and flows over the south geomagnetic pole from the
%  Super Dual Auroral Radar Network (SuperDARN) HF radar at McMurdo Station,
%  Antarctica}, \textit{J. Geophys. Res.}, \textit{116}, A12325,
%  \doi{10.1029/2011JA016834}.
%
%\bibitem[{\textit{Burston et~al.}(2010)\textit{Burston, Astin, Mitchell,
%  Alfonsi, Pedersen, and Skone}}]{Bur10}
%Burston, R., I.~Astin, C.~Mitchell, L.~Alfonsi, T.~Pedersen, and S.~Skone
%  (2010), Turbulent times in the northern polar ionosphere?, \textit{J.
%  Geophys. Res.}, \textit{115}, A04310, \doi{10.1029/2009JA014813}.
%
%\bibitem[{\textit{Burston et~al.}(2016)\textit{Burston, Mitchell, and
%  Astin}}]{Bur16}
%Burston, R., C.~Mitchell, and I.~Astin (2016), Polar cap plasma patch primary
%  linear instability growth rates compared, \textit{J. Geophys. Res. Space
%  Physics}, \textit{121}, 3439--3451, \doi{10.1002/2015JA021895}.
%
%\bibitem[{\textit{Carlson et~al.}(2007)\textit{Carlson, Pedersen, Basu,
%  Keskinen, and Moen}}]{Car07}
%Carlson, H.~C., T.~Pedersen, S.~Basu, M.~Keskinen, and J.~Moen (2007), Case for
%  a new process, not mechanism, for cusp irregularity production, \textit{J.
%  Geophys. Res.}, \textit{112}, A11304, \doi{10.1029/2007JA012384}.
%
%\bibitem[{\textit{Carlson et~al.}(2008)\textit{Carlson, Oksavik, and
%  Moen}}]{Car08}
%Carlson, H.~C., K.~Oksavik, and J.~Moen (2008), On a new process for cusp
%  irregularity production, \textit{Ann. Geophysicae}, \textit{26}, 2871--2885.
%
%\bibitem[{\textit{Chisham et~al.}(2007)\textit{Chisham, Lester, Milan, Freeman,
%  Bristow, Grocott, McWilliams, Ruohoniemi, Yeoman, Dyson, Greenwald, Kikuchi,
%  Pinnock, Rash, Sato, Sofko, Villain, and Walker}}]{Chi07}
%Chisham, G., M.~Lester, S.~E. Milan, M.~P. Freeman, W.~A. Bristow, A.~Grocott,
%  K.~A. McWilliams, J.~M. Ruohoniemi, T.~Yeoman, P.~L. Dyson, R.~A. Greenwald,
%  T.~Kikuchi, M.~Pinnock, J.~P.~S. Rash, N.~Sato, G.~J. Sofko, J.-P. Villain,
%  and A.~D.~M. Walker (2007), A decade of the {Super Dual Auroral Radar Network
%  (SuperDARN)}: scientific achievements, new techniques and future directions,
%  \textit{Surveys in Geophysics}, \textit{28}, 33--109,
%  \doi{10.1007/s10712-007-9017-8}.
%
%\bibitem[{\textit{Dahlgren et~al.}(2012{\natexlab{a}})\textit{Dahlgren, Perry,
%  Semeter, St.-Maurice, Hosokawa, Nicolls, Greffen, Shiokawa, and
%  Heinselman}}]{Dah12b}
%Dahlgren, H., G.~W. Perry, J.~L. Semeter, J.-P. St.-Maurice, K.~Hosokawa, M.~J.
%  Nicolls, M.~Greffen, K.~Shiokawa, and C.~Heinselman (2012{\natexlab{a}}),
%  {Space-time variability of polar cap patches: Direct evidence for internal
%  plasma structuring}, \textit{J. Geophys. Res.}, \textit{117}, A09312,
%  \doi{10.1029/2012JA017961}.
%
%\bibitem[{\textit{Dahlgren et~al.}(2012{\natexlab{b}})\textit{Dahlgren,
%  Semeter, Hosokawa, Nicolls, Butler, Johnsen, Shiokawa, and
%  Heinselman}}]{Dah12a}
%Dahlgren, H., J.~L. Semeter, K.~Hosokawa, M.~J. Nicolls, T.~W. Butler, M.~G.
%  Johnsen, K.~Shiokawa, and Heinselman (2012{\natexlab{b}}), Direct
%  three-dimensional imaging of polar ionospheric strucutes with the {Resolute
%  Bay Incoherent Scatter Radar}, \textit{Geophys. Res. Lett.}, \textit{39},
%  L05104, \doi{10.1029/2012GL050895}.
%
%\bibitem[{\textit{Danskin et~al.}(2002)\textit{Danskin, Koustov, Ogawa,
%  Nishitani, Nozawa, Milan, Lester, and Andre}}]{Dan02}
%Danskin, D.~W., A.~V. Koustov, T.~Ogawa, N.~Nishitani, S.~Nozawa, S.~E. Milan,
%  M.~Lester, and D.~Andre (2002), On the factors controlling occurrence of
%  {F}-region coherent echoes, \textit{Ann. Geophysicae}, \textit{20}(9),
%  1385--1397.
%
%\bibitem[{\textit{Evans}(1969)}]{Eva69}
%Evans, J.~V. (1969), Theory and practice of ionosphere study by {Thomson
%  Scatter Radar}, \textit{Proc. {IEEE}}, \textit{57}, 496--530.
%
%\bibitem[{\textit{Fejer and Kelley}(1980)}]{Fej80}
%Fejer, B.~G., and M.~C. Kelley (1980), Ionospheric irregularities,
%  \textit{Geophys. Rev.}, \textit{18}, 401--454.
%
%\bibitem[{\textit{{Fukumoto} et~al.}(1999)\textit{{Fukumoto}, {Nishitani},
%  {Ogawa}, {Sato}, {Yamagishi}, and {Yukimatu}}}]{Fuk99}
%{Fukumoto}, M., N.~{Nishitani}, T.~{Ogawa}, N.~{Sato}, H.~{Yamagishi}, and
%  A.~{Yukimatu} (1999), Statistical analysis of echo power, {Doppler} velocity
%  and spectral width obtained with the {Syowa South HF radar}, \textit{Adv.
%  Polar Upper Atmos. Res.}, \textit{13}, 37--47.
%
%\bibitem[{\textit{{Fukumoto} et~al.}(2000)\textit{{Fukumoto}, {Nishitani},
%  {Ogawa}, {Sato}, {Yamagishi}, and {Yukimatu}}}]{Fuk00}
%{Fukumoto}, M., N.~{Nishitani}, T.~{Ogawa}, N.~{Sato}, H.~{Yamagishi}, and
%  A.~S. {Yukimatu} (2000), Statistical study of {Doppler} velocity and echo
%  power around 75\(\deg\) magnetic latitude using data obtained with the
%  {Syowa} {East} {HF} radar in 1997, \textit{Adv. Polar Upper Atmos. Res.},
%  \textit{14}, 93--102.
%
%\bibitem[{\textit{Gondarenko and Guzdar}(2004)}]{Gon04}
%Gondarenko, N.~A., and P.~N. Guzdar (2004), Plasma patch structuring by the
%  nonlinear evolution of the gradient drift instability in the high-latitude
%  ionosphere, \textit{J. Geophys. Res.}, \textit{109}, A09301,
%  \doi{10.1029/2004JA010504}.
%
%\bibitem[{\textit{Gondarenko and Guzdar}(2006)}]{Gon06}
%Gondarenko, N.~A., and P.~N. Guzdar (2006), Nonlinear three-dimensional
%  simulations of mesoscale structuring by multiple drives in high-latitude
%  plasma patches, \textit{J. Geophys. Res.}, \textit{111}, A08302,
%  \doi{10.1029/2006JA011701}.
%
%\bibitem[{\textit{Haldoupis et~al.}(1990)\textit{Haldoupis, Nielsen, and
%  Schlegel}}]{Hal90}
%Haldoupis, C., E.~Nielsen, and K.~Schlegel (1990), Dependence of radar auroral
%  scattering cross section on the ambient electron density and the
%  destabilizing electric field, \textit{Ann. Geophysicae}, \textit{8},
%  195--212.
%
%\bibitem[{\textit{{Heinselman} and {Nicolls}}(2008)}]{Hei08}
%{Heinselman}, C.~J., and M.~J. {Nicolls} (2008), {A Bayesian approach to
%  electric field and E-region neutral wind estimation with the Poker Flat
%  Advanced Modular Incoherent Scatter Radar}, \textit{Radio Sci.}, \textit{43},
%  RS5013, \doi{10.1029/2007RS003805}.
%
%\bibitem[{\textit{Kane et~al.}(2012)\textit{Kane, Makarevich, and
%  Devlin}}]{Kan12}
%Kane, T.~A., R.~A. Makarevich, and J.~C. Devlin (2012), {{HF} radar
%  observations of ionospheric backscatter during geomagnetically quiet
%  periods}, \textit{Ann. Geophysicae}, \textit{30}, 221--233,
%  \doi{10.5194/angeo-30-221-2012}.
%
%\bibitem[{\textit{{Keskinen} and {Ossakow}}(1982)}]{Kes82a}
%{Keskinen}, M.~J., and S.~L. {Ossakow} (1982), {Nonlinear evolution of plasma
%  enhancements in the auroral ionosphere. I - Long wavelength irregularities},
%  \textit{J. Geophys. Res.}, \textit{87}, 144--150.
%
%\bibitem[{\textit{{Keskinen} and {Ossakow}}(1983)}]{Kes83}
%{Keskinen}, M.~J., and S.~L. {Ossakow} (1983), Theories of high-latitude
%  ionospheric irregularities: A review, \textit{Radio Sci.}, \textit{18}(6),
%  1077--1091.
%
%\bibitem[{\textit{{Koustov} et~al.}(2004)\textit{{Koustov}, {Sofko},
%  {Andr{\'e}}, {Danskin}, and {Benkevitch}}}]{Kou04}
%{Koustov}, A.~V., G.~J. {Sofko}, D.~{Andr{\'e}}, D.~W. {Danskin}, and L.~V.
%  {Benkevitch} (2004), {Seasonal variation of HF radar F region echo occurrence
%  in the midnight sector}, \textit{J. Geophys. Res.}, \textit{109}, A06305,
%  \doi{10.1029/2003JA010337}.
%
%\bibitem[{\textit{{Koustov} et~al.}(2012)\textit{{Koustov}, {Hosokawa},
%  {Nishitani}, {Shiokawa}, and {Liu}}}]{Kou12}
%{Koustov}, A.~V., K.~{Hosokawa}, N.~{Nishitani}, K.~{Shiokawa}, and H.~{Liu}
%  (2012), {Signatures of moving polar cap arcs in the F-region PolarDARN
%  echoes}, \textit{Ann. Geophysicae}, \textit{30}, 441--455.
%
%\bibitem[{\textit{{Koustov} et~al.}(2014)\textit{{Koustov}, {Ponomarenko},
%  {Ghezelbash}, {Themens}, and {Jayachandran}}}]{Kou14}
%{Koustov}, A.~V., P.~V. {Ponomarenko}, M.~{Ghezelbash}, D.~R. {Themens}, and
%  P.~T. {Jayachandran} (2014), {Electron density and electric field over
%  Resolute Bay and F region ionospheric echo detection with the Rankin Inlet
%  and Inuvik SuperDARN radars}, \textit{Radio Sci.}, pp. 1194--1205,
%  \doi{10.1002/2014RS005579}.
%
%\bibitem[{\textit{Kustov et~al.}(1988)\textit{Kustov, Uspenskii, and
%  Williams}}]{Kus88}
%Kustov, A.~V., M.~V. Uspenskii, and F.~J.~S. Williams (1988), Saturation
%  effects of auroral-electrojet turbulence, \textit{Geom. Aeron.}, \textit{28},
%  923--927.
%
%\bibitem[{\textit{Lamarche and Makarevich}(2015)}]{Lam15}
%Lamarche, L.~J., and R.~A. Makarevich (2015), {Solar control of F region radar
%  backscatter: Further insights from observations in the southern polar cap},
%  \textit{J. Geophys. Res. Space Physics}, \textit{120}(11), 9875--9890,
%  \doi{10.1002/2015JA021663}.
%
%\bibitem[{\textit{Lamarche and Makarevich}(2016)}]{Lam16}
%Lamarche, L.~J., and R.~A. Makarevich (2016), A modeling study of asymmetries
%  in plasma irregularity characteristics near gradient reversals, \textit{Ann.
%  Geophysicae}, \textit{34}, 709--723, \doi{10.5194/angeo-34-709-2016}.
%
%\bibitem[{\textit{{Linson} and {Workman}}(1970)}]{Lin70}
%{Linson}, L.~M., and J.~B. {Workman} (1970), Formation of striations in
%  ionospheric plasma clouds, \textit{J. Geophys. Res.}, \textit{75}, 3211.
%
%\bibitem[{\textit{Lynch et~al.}(2015)\textit{Lynch, Hampton, Zettergren,
%  Bekkeng, Conde, Fernandes, Horak, Lessard, Miceli, Michell, Moen, Nicolls,
%  Powell, and Samara}}]{Lyn15}
%Lynch, K.~A., D.~K. Hampton, M.~Zettergren, T.~A. Bekkeng, M.~Conde, P.~A.
%  Fernandes, P.~Horak, M.~Lessard, R.~Miceli, R.~Michell, J.~Moen, M.~Nicolls,
%  S.~P. Powell, and M.~Samara (2015), {MICA} sounding rocket observations of
%  the conductivity-gradient-generated auroral ionospheric responses:
%  Small-scale structures with large-scale drivers, \textit{J. Geophys. Res.
%  Space Physics}, \textit{120}, 9661--9682, \doi{10.1002/2014JA020860}.
%
%\bibitem[{\textit{Makarevich}(2014)}]{Mak14c}
%Makarevich, R.~A. (2014), {Symmetry considerations in the two-fluid theory of
%  the gradient-drift instability in the lower ionosphere}, \textit{J. Geophys.
%  Res.}, \textit{119}, \doi{10.1002/2014JA020292}.
%
%\bibitem[{\textit{Makarevich and Bristow}(2014)}]{Mak14b}
%Makarevich, R.~A., and W.~A. Bristow (2014), Coordinated radar observations of
%  plasma wave characteristics in the auroral {F} region, \textit{Ann.
%  Geophysicae}, \textit{32}, 875--888.
%
%\bibitem[{\textit{Makarevich et~al.}(2015)\textit{Makarevich, Lamarche, and
%  Nicolls}}]{Mak15b}
%Makarevich, R.~A., L.~J. Lamarche, and M.~J. Nicolls (2015), {Resolute Bay
%  Incoherent Scatter Radar} observations of plasma structures in the vicinity
%  of polar holes, \textit{J. Geophys. Res. Space Physics}, \textit{120},
%  \doi{10.1002/2015JA021443}.
%
%\bibitem[{\textit{Makarevitch et~al.}(2004)\textit{Makarevitch, Honary, McCrea,
%  and Howells}}]{Mak04c}
%Makarevitch, R.~A., F.~Honary, I.~W. McCrea, and V.~S.~C. Howells (2004),
%  Imaging riometer observations of drifting absorption patches in the morning
%  sector, \textit{Ann. Geophysicae}, \textit{22}, 3461--3478.
%
%\bibitem[{\textit{Milan et~al.}(2002)\textit{Milan, Lester, and
%  Yeoman}}]{Mil02b}
%Milan, S.~E., M.~Lester, and T.~K. Yeoman (2002), {HF} radar polar patch
%  formation revisited: summer and winter variations in dayside plasma
%  structuring, \textit{Ann. Geophysicae}, \textit{20}, 487--499.
%
%\bibitem[{\textit{Moen et~al.}(2012)\textit{Moen, Oksavik, Abe, Lester, Saito,
%  Bekkeng, and Jacobsen}}]{Moe12}
%Moen, J., K.~Oksavik, T.~Abe, M.~Lester, Y.~Saito, T.~A. Bekkeng, and K.~S.
%  Jacobsen (2012), {First in situ measurements of HF radar echoing targets},
%  \textit{Geophys. Res. Lett.}, \textit{39}, L07104,
%  \doi{10.1029/2012GL051407}.
%
%\bibitem[{\textit{Nicolls et~al.}(2007)\textit{Nicolls, Heinselman, Hope,
%  Ranjan, Kelley, and Kelly}}]{Nic07}
%Nicolls, M.~J., C.~J. Heinselman, E.~A. Hope, S.~Ranjan, M.~C. Kelley, and
%  J.~D. Kelly (2007), Imaging of polar mesosphere summer echoes with the 450
%  {MHz} {Poker Flat Advanced Modular Incoherent Scatter Radar},
%  \textit{Geophys. Res. Lett.}, \textit{34}, L20102,
%  \doi{10.1029/2007GL031476}.
%
%\bibitem[{\textit{Oksavik et~al.}(2010)\textit{Oksavik, Barth, Moen, and
%  Lester}}]{Oks10}
%Oksavik, K., V.~L. Barth, J.~Moen, and M.~Lester (2010), On the entry and
%  transit of high-density plasma across the polar cap, \textit{J. Geophys.
%  Res.}, \textit{115}, \doi{10.1029/2010JA015817}.
%
%\bibitem[{\textit{Rishbeth and Williams}(1985)}]{Ris85}
%Rishbeth, H., and P.~J.~S. Williams (1985), The {EISCAT} ionospheric radar: The
%  system and its early results, \textit{Q. J. R. Astron. Soc.}, \textit{26},
%  478--512.
%
%\bibitem[{\textit{Ruohoniemi and Greenwald}(1997)}]{Ruo97}
%Ruohoniemi, J.~M., and R.~A. Greenwald (1997), {Rates of scattering occurrence
%  in routine HF radar observations during solar cycle maximum}, \textit{Radio
%  Sci.}, \textit{32}, 1051--1070.
%
%\bibitem[{\textit{{Semeter} et~al.}(2009)\textit{{Semeter}, {Butler},
%  {Heinselman}, {Nicolls}, {Kelly}, and {Hampton}}}]{Sem09}
%{Semeter}, J., T.~{Butler}, C.~{Heinselman}, M.~{Nicolls}, J.~{Kelly}, and
%  D.~{Hampton} (2009), {Volumetric imaging of the auroral ionosphere: Initial
%  results from PFISR}, \textit{J. Atmos. Sol. Terr. Phys.}, \textit{71},
%  738--743.
%
%\bibitem[{\textit{Shepherd}(2014)}]{She14}
%Shepherd, S.~G. (2014), Altitude-adjusted corrected geomagnetic coordinates:
%  Definition and functional approximations, \textit{J. Geophys. Res.},
%  \textit{119}(9), 7501--7521, \doi{10.1002/2014JA020264}.
%
%\bibitem[{\textit{Spicher et~al.}(2015)\textit{Spicher, Miloch, Clausen, and
%  Moen}}]{Spi15}
%Spicher, A., W.~J. Miloch, L.~B.~N. Clausen, and J.~I. Moen (2015), Plasma
%  turbulence and coherent structures in the polar cap observed by the {ICI}-2
%  sounding rocket, \textit{J. Geophys. Res. Space Physics}, \textit{120},
%  10,959--10,978, \doi{10.1002/2015JA021634}.
%
%\bibitem[{\textit{Themens et~al.}(2014)\textit{Themens, Jayachandran, Nicolls,
%  and MacDougall}}]{The14a}
%Themens, D., P.~Jayachandran, M.~J. Nicolls, and J.~W. MacDougall (2014), {A
%  Top to bottom evaluation of IRI-2007 within the polar cap}, \textit{J.
%  Geophys. Res.}, \textit{119}, \doi{10.1002/2014JA020052}.
%
%\bibitem[{\textit{{Tsunoda}}(1988)}]{Tsu88}
%{Tsunoda}, R.~T. (1988), {High-latitude F region irregularities - A review and
%  synthesis}, \textit{Geophys. Rev.}, \textit{26}, 719--760.
%
%\bibitem[{\textit{{Villain} et~al.}(1986)\textit{{Villain}, {Beghin}, and
%  {Hanuise}}}]{Vil86}
%{Villain}, J.~P., C.~{Beghin}, and C.~{Hanuise} (1986), {ARCAD3-SAFARI}
%  coordinated study of auroral and polar {F}-region ionospheric irregularities,
%  \textit{Ann. Geophysicae}, \textit{4}, 61--68.
%
%\bibitem[{\textit{{Weber} et~al.}(1984)\textit{{Weber}, {Buchau}, {Moore},
%  {Sharber}, {Livingston}, {Winningham}, and {Reinisch}}}]{Web84}
%{Weber}, E.~J., J.~{Buchau}, J.~G. {Moore}, J.~R. {Sharber}, R.~C.
%  {Livingston}, J.~D. {Winningham}, and B.~W. {Reinisch} (1984), {F layer
%  ionization patches in the polar cap}, \textit{J. Geophys. Res.}, \textit{89},
%  1683--1694.
%
%\end{thebibliography}

%% After you run BibTeX, Copy in the contents of the .bbl file here:

%Text here ===>>>

%%

%  Numbered lines in equations:
%  To add line numbers to lines in equations,
%  \begin{linenomath*}
%  \begin{equation}
%  \end{equation}
%  \end{linenomath*}

%
% ---------------
% EXAMPLE TABLE
%
% \begin{table}
% \caption{Time of the Transition Between Phase 1 and Phase 2$^{a}$}
% \centering
% \begin{tabular}{l c}
% \hline
%  Run  & Time (min)  \\
% \hline
%   $l1$  & 260   \\
%   $l2$  & 300   \\
%   $l3$  & 340   \\
%   $h1$  & 270   \\
%   $h2$  & 250   \\
%   $h3$  & 380   \\
%   $r1$  & 370   \\
%   $r2$  & 390   \\
% \hline
% \multicolumn{2}{l}{$^{a}$Footnote text here.}
% \end{tabular}
% \end{table}

%% Enter Figures and Tables near as possible to where they are first mentioned:
%
% DO NOT USE \psfrag or \subfigure commands.
%
% Figure captions go below the figure.
% Table titles go above tables;  other caption information
%  should be placed in last line of the table, using
% \multicolumn2l{$^a$ This is a table note.}
%
%----------------
% EXAMPLE FIGURE
%
% \begin{figure}[h]
% \centering
% when using pdflatex, use pdf file:
% \includegraphics[width=20pc]{figsamp.pdf}
%
% when using dvips, use .eps file:
% \includegraphics[width=20pc]{figsamp.eps}
%
% \caption{Short caption}
% \label{figone}
%  \end{figure}

%% SIDEWAYS FIGURE and TABLE 
% AGU prefers the use of {sidewaystable} over {landscapetable} as it causes fewer problems.
%
% \begin{sidewaysfigure}
% \includegraphics[width=20pc]{figsamp}
% \caption{caption here}
% \label{newfig}
% \end{sidewaysfigure}
% 
%  \begin{sidewaystable}
%  \caption{Caption here}
% \label{tab:signif_gap_clos}
%  \begin{tabular}{ccc}
% one&two&three\\
% four&five&six
%  \end{tabular}
%  \end{sidewaystable}

%% If using numbered lines, please surround equations with \begin{linenomath*}...\end{linenomath*}
%\begin{linenomath*}
%\begin{equation}
%y|{f} \sim g(m, \sigma),
%\end{equation}
%\end{linenomath*}

%%% End of body of article

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Optional Appendix goes here
%
% The \appendix command resets counters and redefines section heads
%
% After typing \appendix
%
%\section{Here Is Appendix Title}
% will show
% A: Here Is Appendix Title
%
%\appendix
%\section{Here is a sample appendix}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% Optional Glossary, Notation or Acronym section goes here:
%
%%%%%%%%%%%%%%  
% Glossary is only allowed in Reviews of Geophysics
%  \begin{glossary}
%  \term{Term}
%   Term Definition here
%  \term{Term}
%   Term Definition here
%  \term{Term}
%   Term Definition here
%  \end{glossary}

%
%%%%%%%%%%%%%%
% Acronyms
%   \begin{acronyms}
%   \acro{Acronym}
%   Definition here
%   \acro{EMOS}
%   Ensemble model output statistics 
%   \acro{ECMWF}
%   Centre for Medium-Range Weather Forecasts
%   \end{acronyms}

%
%%%%%%%%%%%%%%
% Notation 
%   \begin{notation}
%   \notation{$a+b$} Notation Definition here
%   \notation{$e=mc^2$} 
%   Equation in German-born physicist Albert Einstein's theory of special
%  relativity that showed that the increased relativistic mass ($m$) of a
%  body comes from the energy of motion of the body—that is, its kinetic
%  energy ($E$)—divided by the speed of light squared ($c^2$).
%   \end{notation}

%% ------------------------------------------------------------------------ %%
%% Citations

% Please use ONLY \citet and \citep for reference citations.
% DO NOT use other cite commands (e.g., \cite, \citeyear, \nocite, \citealp, etc.).

%% Example \citet and \citep:
%  ...as shown by \citet{Boug10}, \citet{Buiz07}, \citet{Fra10},
%  \citet{Ghel00}, and \citet{Leit74}.

%  ...as shown by \citep{Boug10}, \citep{Buiz07}, \citep{Fra10},
%  \citep{Ghel00, Leit74}.

%  ...has been shown \citep [e.g.,][]{Boug10,Buiz07,Fra10}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Track Changes:
% To add words, \added{<word added>}
% To delete words, \deleted{<word deleted>}
% To replace words, \replace{<word to be replaced>}{<replacement word>}
% To explain why change was made: \explain{<explanation>} This will put
% a comment into the right margin.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% At the end of the document, use \listofchanges, which will list the
% changes and the page and line number where the change was made.

% When final version, \listofchanges will not produce anything,
% \added{<word or words>} word will be printed, \deleted{<word or words} will take away the word,
% \replaced{<delete this word>}{<replace with this word>} will print only the replacement word.
%  In the final version, \explain will not print anything.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%
%\listofchanges
%%%

%\end{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Supporting Information
%% (Optional) See AGUSuppInfoSamp.tex/pdf for requirements 
%% for Supporting Information.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%More Information and Advice:
%
%------------------------------------------------------------------------ %%
%
% SECTION HEADS
%
%------------------------------------------------------------------------ %%
%
%Capitalize the first letter of each word (except for
%prepositions, conjunctions, and articles that are
%three or fewer letters).
%
%AGU follows standard outline style; therefore, there cannot be a section 1 without
%a section 2, or a section 2.3.1 without a section 2.3.2.
%Please make sure your section numbers are balanced.
%---------------
%Level 1 head
%
%Use the \section{} command to identify level 1 heads;
%type the appropriate head wording between the curly
%brackets, as shown below.
%
%An example:
%\section{Level 1 Head: Introduction}
%
%---------------
%Level 2 head
%
%Use the \subsection{} command to identify level 2 heads.
%An example:
%\subsection{Level 2 Head}
%
%---------------
%Level 3 head
%
%Use the \subsubsection{} command to identify level 3 heads
%An example:
%\subsubsection{Level 3 Head}
%
%---------------
%Level 4 head
%
%Use the \subsubsubsection{} command to identify level 3 heads
%An example:
%\subsubsubsection{Level 4 Head} An example.
%
%------------------------------------------------------------------------ %%
%
% IN-TEXT LISTS
%
%------------------------------------------------------------------------ %%
%
%Do not use bulleted lists; enumerated lists are okay.
%\begin{enumerate}
%\item
%\item
%\item
%\end{enumerate}
%
%------------------------------------------------------------------------ %%
%
% EQUATIONS
%
%------------------------------------------------------------------------ %%
%
%Single-line equations are centered.
%Equation arrays will appear left-aligned.
%
%Math coded inside display math mode \[ ...\]
% will not be numbered, e.g.,:
% \[ x^2=y^2 + z^2\]
%
% Math coded inside \begin{equation} and \end{equation} will
% be automatically numbered, e.g.,:
% \begin{equation}
% x^2=y^2 + z^2
% \end{equation}
%
%
%To create multiline equations, use the
%\begin{eqnarray} and \end{eqnarray} environment
%as demonstrated below.
%\begin{eqnarray}
%  x_{1} & = & (x - x_{0}) \cos \Theta \nonumber \\
%        && + (y - y_{0}) \sin \Theta  \nonumber \\
%  y_{1} & = & -(x - x_{0}) \sin \Theta \nonumber \\
%        && + (y - y_{0}) \cos \Theta.
%\end{eqnarray}
%
%If you don't want an equation number, use the star form:
%\begin{eqnarray*}...\end{eqnarray*}
%
%Break each line at a sign of operation
%(+, -, etc.) if possible, with the sign of operation
%on the new line.
%
%Indent second and subsequent lines to align with
%the first character following the equal sign on the
%first line.
%
%Use an \hspace{} command to insert horizontal space
%into your equation if necessary. Place an appropriate
%unit of measure between the curly braces, e.g.
%\hspace{1in}; you may have to experiment to achieve
%the correct amount of space.
%
%
%------------------------------------------------------------------------ %%
%
% EQUATION NUMBERING: COUNTER
%
%------------------------------------------------------------------------ %%
%
%You may change equation numbering by resetting
%the equation counter or by explicitly numbering
%an equation.
%
%To explicitly number an equation, type \eqnum{}
%(with the desired number between the brackets)
%after the \begin{equation} or \begin{eqnarray}
%command.  The \eqnum{} command will affect only
%the equation it appears with; LaTeX will number
%any equations appearing later in the manuscript
%according to the equation counter.
%
%
%If you have a multiline equation that needs only
%one equation number, use a \nonumber command in
%front of the double backslashes (\\) as shown in
%the multiline equation above.
%
%If you are using line numbers, remember to surround
%equations with \begin{linenomath*}...\end{linenomath*}
%
% To add line numbers to lines in equations:
% \begin{linenomath*}
% \begin{equation}
% \end{equation}
% \end{linenomath*}
%
%
%

### 2017-02-02 thesis.sh

@@ -18,6 +18,7 @@ bibtex LLamarcheThesis
 bibtex introduction
 bibtex paper1
 bibtex paper2
+bibtex paper3
 bibtex conclusion
 
 pdflatex LLamarcheThesis.tex

### 2017-02-02 conclusion.tex
Formatting corrections made

@@
   \caption[Irregularity production factors]{Summary of the main conclusions in this thesis.
-The radar of interest (MCM) is highlighted.
+The radar of interest (MCM) is highlighted in pink.
 Polar patches (orange) drift antisunwards across the polar cap following the background plasma convection (red contours).
@@
 Small-scale irregularity production surrounding a large-scale structure is found to exhibit both asymmetry and anisotropy.
-Radars are most likely to observe this asymmetry if the patch is elongated along the radar boresight.}
+Radars are most likely to observe this asymmetry if the patch is elongated along the radar boresight.
+Backscatter is most likely to be observed in a particular window of slant ranges within the radar FoV.}
   \label{fig:conclusion}

### 2017-02-02 introduction.tex
Paper3 added to thesis
Formatting corrections made

@@
 
-Plasmas are fluid masses of charged particles, often formed when neutral atoms and molecules in a gas are ionized by external energy.
+Plasmas are fluid masses of charged particles, often formed when neutral atoms and mol\-e\-cules in a gas are ionized by external energy.
 Instances of plasma include everything from fluorescent lightbulbs and flames to more exotic examples such as nuclear fusion reactor experiments, most stars, nebulas, and the interstellar medium.
@@
 	\includegraphics[width=\textwidth,angle=180]{SuperDARNmap.pdf}
-	\caption[SuperDARN map]{Fields-of-view of all operational SuperDARN radars in both the northern (left) and southern (right) hemispheres.
-Polar-latitude radars are shown with the red outlines.
-The radars at Rankin Inlet (northern hemisphere) and McMurdo Station (southern hemisphere) are shaded in red because these two instruments are particularly important to this body of work.
+	\caption[SuperDARN map]{Fields-of-view of all operational SuperDARN radars in both the northern (a) and southern (b) hemispheres.
+Polar-latitude radars are shown shaded in blue.
+The radars at Rankin Inlet (northern hemisphere) and McMurdo Station (southern hemisphere) are outlined in pink because these two instruments are particularly important to this body of work.
 Lines of constant magnetic latitude are shown in yellow at \(\Lambda = 40\deg, 60\deg, 80\deg\) in the northern hemisphere and at \(\Lambda = -40\deg,-60\deg,-80\deg\) in the southern hemisphere.}
@@
 	\caption[ISR map]{FoVs of all ISRs in the northern polar region.
-The individual radars shown are identified by their commonly used acronyms: Poker Flat (PFISR), Resolute Bay North (RISR-N), Resolute Bay Canada (RISR-C), Sondrestrom (SFJ).
+The individual radars shown are identified by their commonly used acronyms: Poker Flat (PFISR), Resolute Bay North (RISR-N), Resolute Bay Canada (RISR-C), Sondrestrom (SFJ), European Incoherent Scatter Scientific Association (EISCAT), and EISCAT Svalbard (ESR).
 RISR-N, which is primarily used throughout this thesis, is shown in orange.
-The FoV of the SuperDARN RKN radar is shown by the red outline.
+The FoV of the SuperDARN RKN radar is shown by the pink outline.
 Lines of constant MLAT are shown at \(\Lambda=80\deg,60\deg\) in yellow.}

### 2017-02-02 paper2.tex
Corrections to Abstract and Intro
Formatting corrections made

@@
 
-Figure \ref{fig:1}a presents altitudinal dependencies of several parameters of interest.
+Figure \ref{fig:model_setup}a presents altitudinal dependencies of several parameters of interest.
 Shown are the ratios of collision frequencies and gyrofrequencies for electrons \(-r_e=\nu_e/|\Omega_e|\) (blue) and ions \(r_i=\nu_i/\Omega_i\) (green), the anisotropy parameter \(\psi=-r_i r_e\) (pink), and the conductivity ratio \(R=\sigma_H/\sigma_P = \left(r_i+r_e\right)/\left(1+\psi\right)\) (dashed black line on top of the green one).
 The above notations are the same as in the recent study by \citet{Makarevich2014c}, where the electrons gyrofrequency \(\Omega_e=-|e|B/m_e\) and the ratio \(r_e=\nu_e/\Omega_e\) are negative, while other quantities of interest are positive.
-The \(x\) axis in Figure \ref{fig:1}a does not have any units since all presented parameters are dimensionless quantities.
+The \(x\) axis in Figure \ref{fig:model_setup}a does not have any units since all presented parameters are dimensionless quantities.
 
@@
 	\centering
-	\includegraphics[width=\textwidth,angle=180]{f01.pdf}
-	\caption[Overview of altitudinal variations and model set-up of MCM]{(a) Altitudinal variations of parameters of interest.
+	\includegraphics[width=\textwidth,angle=180]{p2f1.pdf}
+	\caption[Overview of altitudinal variations and model setup of MCM]{(a) Altitudinal variations of parameters of interest.
 Shown are the ratios of collision frequencies and gyrofrequencies for electrons \(-r_e=\nu_e/|\Omega_e|\) (blue) and ions \(r_i=\nu_i/\Omega_i\) (green), the anisotropy parameter \(\psi=-r_i r_e\) (pink), and the conductivity ratio \(R=\sigma_H/\sigma_P = \left(r_i+r_e\right)/\left(1+\psi\right)\) (dashed black line on top of the green one).
@@
 The \(E\) region, transitional region, and \(F\) region are outlined in red, green, and blue, respectively.}
-	\label{fig:1}
+	\label{fig:model_setup}
 \end{figure}
@@
 For an oblique-sounding radar like SuperDARN, different range gates refer to different altitudes, with a virtual height model normally used to characterize this altitudinal dependence of the radar scattering volume as a function of the slant range.
-Figure \ref{fig:1}b shows the standard SuperDARN virtual height model \citep{Chisham2008}.
+Figure \ref{fig:model_setup}b shows the standard SuperDARN virtual height model \citep{Chisham2008}.
 The ranges where the beam is within the \(E\) region (red), \(F\) region (blue), and transitional region (green) are indicated.
@@
 The model is based on where the signal is typically coming from for any given range.
-In the standard model the virtual height is constant throughout the \(E\) and \(F\) regions, but changes rapidly in the transitional region between 600 km and 800 km as shown in Figure \ref{fig:1}b.
-For individual observations in each beam, the height variations will be more gradual, but on average it is represented well by the model shown in Figure \ref{fig:1}b.
+In the standard model the virtual height is constant throughout the \(E\) and \(F\) regions, but changes rapidly in the transitional region between 600 km and 800 km as shown in Figure \ref{fig:model_setup}b.
+For individual observations in each beam, the height variations will be more gradual, but on average it is represented well by the model shown in Figure \ref{fig:model_setup}b.
 A revised virtual height model in which virtual height changes with slant range has also been developed \citep{Chisham2008}, but since in the current study we were interested in the differences between the \(E\) and \(F\) region behavior (rather than small differences within each region), a standard virtual height model has been used.
 
-Figure \ref{fig:1}c presents a typical observational geometry in the horizontal plane.
+Figure \ref{fig:model_setup}c presents a typical observational geometry in the horizontal plane.
 Here we show a FoV of the radar at McMurdo Antarctic station with the geographic South Pole located to the left and down of panel (c) (in the center of concentric circles that show geographic parallels).
@@
 
-Figure \ref{fig:1}c presents a fan plot of the McMurdo radar's FoV with a large-scale structure across the FoV (thick white line).
+Figure \ref{fig:model_setup}c presents a fan plot of the McMurdo radar's FoV with a large-scale structure across the FoV (thick white line).
 It it assumed that the field of gradient vectors \(\vec{G}=\nabla n_0/n_0\) is uniform in magnitude, with directions always pointing towards the structure throughout the FoV.
@@
 The first angle describes the structure orientation relative to the radar.
-This is the angle \(\theta\) between the gradient reversal direction and the direction perpendicular to the boresite (FoV center direction); this perpendicular direction is shown by the dashed horizontal line in Figure \ref{fig:1}c.
-The second angle \(\phi\) is the angle between the structure drift velocity (red vector in Figure \ref{fig:1}c) and the dashed horizontal line.
+This is the angle \(\theta\) between the gradient reversal direction and the direction perpendicular to the boresite (FoV center direction); this perpendicular direction is shown by the dashed horizontal line in Figure \ref{fig:model_setup}c.
+The second angle \(\phi\) is the angle between the structure drift velocity (red vector in Figure \ref{fig:model_setup}c) and the dashed horizontal line.
 Since these structures tend to move with the background convection velocity \citep{Makarevich2015b}, this drift velocity is simply \(\vec{V}_E= \vec{E}\times\vec{B}/B^2\).
@@
 Finally, the GDI growth rates were found for each radar beam-range cell as follows.
-The location and height of the center of each radar cell was found with a standard positioning method utilizing virtual heights shown in Figure \ref{fig:1}b.
-The height was used to find the value of the conductivity ratio \(R\) using information presented in Figure \ref{fig:1}a.
+The location and height of the center of each radar cell was found with a standard positioning method utilizing virtual heights shown in Figure \ref{fig:model_setup}b.
+The height was used to find the value of the conductivity ratio \(R\) using information presented in Figure \ref{fig:model_setup}a.
 Using these quantities, the GDI growth rate at any point within the radar FoV can be calculated \citep[eqn.
@@
 The quantity \(GV_E\) will be the same for all vector orientations and heights.
-The anisotropy parameter \(\psi\) will change at different heights, but it is always small relative to 1, Figure \ref{fig:1}a, so its effect is minimal.
+The anisotropy parameter \(\psi\) will change at different heights, but it is always small relative to 1, Figure \ref{fig:model_setup}a, so its effect is minimal.
 Since this study focuses primarily on different relative orientations of the gradient and convection directions and altitude regimes, it is convenient to use a normalized growth rate \(\gamma^\prime\).
@@
 
-Each radar cell in Figure \ref{fig:1}c is color-coded in the GDI growth rate at its center, with the color bar given in the bottom-right corner of Figure \ref{fig:1}c.
+Each radar cell in Figure \ref{fig:model_setup}c is color-coded in the GDI growth rate at its center, with the color bar given in the bottom-right corner of Figure \ref{fig:model_setup}c.
 On the left side of the radar FoV, the growth rate is positive on the leading edge of the gradient reversal and negative on the trailing edge, but both are very close to zero.
@@
 
-Figure \ref{fig:1}c presented the growth rate pattern for one particular orientation of the gradient reversal and plasma convection.
-Figure \ref{fig:2} presents the same analysis for four different orientations and locations of the gradient reversal.
+Figure \ref{fig:model_setup}c presented the growth rate pattern for one particular orientation of the gradient reversal and plasma convection.
+Figure \ref{fig:fov_grs} presents the same analysis for four different orientations and locations of the gradient reversal.
 In all cases, the convection direction is assumed to be perpendicular to the gradient reversal direction (\(\phi-\theta=90\deg\)).
@@
 
-Figure \ref{fig:2}a shows a gradient reversal through the center of the \(F\)-region portion of the FoV that is perpendicular to the boresite (\(\theta=0\deg\)).
+Figure \ref{fig:fov_grs}a shows a gradient reversal through the center of the \(F\)-region portion of the FoV that is perpendicular to the boresite (\(\theta=0\deg\)).
 The growth rate on both sides of the gradient reversal is small (close to zero), but it is slightly negative on the leading edge and slightly positive on the trailing edge.
@@
 \begin{figure}
-	\includegraphics[width=\textwidth]{f02.pdf}
+	\includegraphics[width=\textwidth]{p2f2.pdf}
 	\caption[GDI growth rate within radar FoV]{The GDI growth rate within the radar's FoV for four cases of gradient reversal orientation and position (white line).
 In all cases, the convection direction \(\vec{V}_E\) (red arrow) is perpendicular to the gradient reversal direction.
-The orientation of the vectors \(\vec{G}\), \(\vec{V}_E\), \(\vec{E}\), and \(\vec{k}\) are shown in each panel, using the same colors as in Figure \ref{fig:1}c.
+The orientation of the vectors \(\vec{G}\), \(\vec{V}_E\), \(\vec{E}\), and \(\vec{k}\) are shown in each panel, using the same colors as in Figure \ref{fig:model_setup}c.
 The GDI growth rate at each radar cell is given by the cell color, with the color bar given in the bottom-left corner of panel (a).
 Panels (a) and (b) show a gradient reversal centered in the \(F\) region, while panels (c) and (d) show the reversal in the transitional region.
-The angles \(\theta\) and \(\phi\) (as defined in Figure \ref{fig:1}c) are also given in the bottom-right corner of each panel.}
-	\label{fig:2}
+The angles \(\theta\) and \(\phi\) (as defined in Figure \ref{fig:model_setup}c) are also given in the bottom-right corner of each panel.}
+	\label{fig:fov_grs}
 \end{figure}
 
-Figure \ref{fig:2}b shows an inclined gradient reversal through the FoV in the \(F\) region (\(\theta=40\deg\)).
+Figure \ref{fig:fov_grs}b shows an inclined gradient reversal through the FoV in the \(F\) region (\(\theta=40\deg\)).
 Again, the growth rate is negative on the leading edge and positive on the trailing edge, but the difference is much more dramatic.
@@
 
-Figure \ref{fig:2}c presents the same configuration as Figure \ref{fig:2}a, except that a gradient reversal is centered in the transitional region.
-Similar to Figure \ref{fig:2}a, the growth rate is close to zero on both sizes of the reversal in the center of the radar FoV.
+Figure \ref{fig:fov_grs}c presents the same configuration as Figure \ref{fig:fov_grs}a, except that a gradient reversal is centered in the transitional region.
+Similar to Figure \ref{fig:fov_grs}a, the growth rate is close to zero on both sizes of the reversal in the center of the radar FoV.
 However, there are more extreme values (\(\sim\mp0.5\)) near the edge beams.
-Figure \ref{fig:2}d shows an inclined gradient reversal in the transitional region.
+Figure \ref{fig:fov_grs}d shows an inclined gradient reversal in the transitional region.
 The growth rate difference across the gradient is again more extreme further away from the boresite.
@@
 In order to quantify the above-noted differences in the growth rate for different orientations of the gradient reversal, the average change in the growth rate across the reversal \(\Delta\gamma_{avg}\) was found for any general configuration of the gradient reversal and the convection direction.
-The approach presented in Figures \ref{fig:1}c and \ref{fig:2} was useful for presenting and emphasizing general spatial patterns in the growth rate.
+The approach presented in Figures \ref{fig:model_setup}c and \ref{fig:fov_grs} was useful for presenting and emphasizing general spatial patterns in the growth rate.
 In this section, we employ two new methods that are more robust, independent of the way the the radar FoV is divided into beam-range cells, and more suitable for evaluating the growth rate differences in the vicinity of large-scale structures.
@@
 
-Figure \ref{fig:3} examines the growth rate difference across the gradient reversal for 4 selected angle differences \(\theta- \phi\) as indicated in the top-right corner of each panel.
+Figure \ref{fig:set_angle} examines the growth rate difference across the gradient reversal for 4 selected angle differences \(\theta- \phi\) as indicated in the top-right corner of each panel.
 In each case, the average growth rate difference, notated simply by \(\Delta\gamma\) for the remainder of this paper, is calculated for different \(\theta\) values from \(0\deg\) to \(360\deg\).
@@
 	\centering
-	\includegraphics[width=\textwidth,angle=180]{f03.pdf}
+	\includegraphics[width=\textwidth,angle=180]{p2f3.pdf}
 	\caption[GDI growth rates for set angle between \(\vec{V}_E\) and \(\vec{G}\)]{The difference in the GDI growth rate across the gradient reversal for four selected angles between the reversal and convection direction.
@@
 The trends for both the \(E\)-region (red) and \(F\)-region (blue) cases are shown in each panel.}
-	\label{fig:3}
+	\label{fig:set_angle}
 \end{figure}
 
-Figure \ref{fig:3}a shows  a case of \(\phi-\theta=0\deg\), i.e.
+Figure \ref{fig:set_angle}a shows  a case of \(\phi-\theta=0\deg\), i.e.
 the case when the convection direction is parallel to the gradient reversal.
@@
 However, \(\Delta\gamma\approx 1\) and \(\Delta\gamma\approx -1\) are the maximum and minimum values, respectively, so the contrast across the gradient reversal is relatively small regardless of whether the leading or trailing edge has the greater growth rate.
-The important conclusion from this analysis is that the contrast or the asymmetry magnitude depends strongly on the structure orientation angle \(\theta\), which confirms the previous results based on two angles \(\theta\) only, Figures \ref{fig:2}a and b.
+The important conclusion from this analysis is that the contrast or the asymmetry magnitude depends strongly on the structure orientation angle \(\theta\), which confirms the previous results based on two angles \(\theta\) only, Figures \ref{fig:fov_grs}a and b.
 
-Figures \ref{fig:3}b and \ref{fig:3}c present the same analysis, but with a set angle of \(30\deg\) and \(60\deg\), respectively, between the gradient reversal and the convection direction.
-These show the gradual transition between the extreme cases of the convection direction parallel to the gradient reversal, Figure \ref{fig:3}a, and the convection direction perpendicular to the reversal, Figure \ref{fig:3}d.
+Figures \ref{fig:set_angle}b and \ref{fig:set_angle}c present the same analysis, but with a set angle of \(30\deg\) and \(60\deg\), respectively, between the gradient reversal and the convection direction.
+These show the gradual transition between the extreme cases of the convection direction parallel to the gradient reversal, Figure \ref{fig:set_angle}a, and the convection direction perpendicular to the reversal, Figure \ref{fig:set_angle}d.
 The general \(E\)-region pattern starts to move towards more positive values and the \(F\)-region pattern moves towards more negative values.
@@
 
-Figure \ref{fig:3}d shows the case when the convection direction is perpendicular to the gradient reversal, which corresponds to \(\vec{V}_E \parallel \vec{G}\).
+Figure \ref{fig:set_angle}d shows the case when the convection direction is perpendicular to the gradient reversal, which corresponds to \(\vec{V}_E \parallel \vec{G}\).
 From previous studies, this is the condition where the growth rate on the trailing edge is the greatest in the \(F\) region.
-Similar to Figure \ref{fig:3}a, the growth rate difference in Figure \ref{fig:3}d for the \(F\) region (blue) is negative for all directions of the gradient reversal and has the lowest values observed for the \(F\) region (\(\sim-2\)) at \(90\deg\) and \(270\deg\).
+Similar to Figure \ref{fig:set_angle}a, the growth rate difference in Figure \ref{fig:set_angle}d for the \(F\) region (blue) is negative for all directions of the gradient reversal and has the lowest values observed for the \(F\) region (\(\sim-2\)) at \(90\deg\) and \(270\deg\).
 This corresponds to the largest contrast between the leading and trailing edge of the structure.
@@
 
-The analysis presented in Figure \ref{fig:3} can be extended to include any arbitrary orientation between the gradient reversal and convection direction by allowing \(\phi\) and \(\theta\) to vary independently.
+The analysis presented in Figure \ref{fig:set_angle} can be extended to include any arbitrary orientation between the gradient reversal and convection direction by allowing \(\phi\) and \(\theta\) to vary independently.
 Both the numerical method and the analytical method can be used to find the average difference in growth rate for any direction of the gradient reversal and plasma convection in the \(E\) and \(F\) regions.
-Figure \ref{fig:4} presents the results of this analysis in the form of contour plots of the growth rate difference \(\Delta\gamma\) versus angles \(\theta\) and \(\phi\).
-Figures \ref{fig:4}a and \ref{fig:4}b show the results of the analytic method in the \(E\) and \(F\) regions, respectively.
+Figure \ref{fig:any_angle} presents the results of this analysis in the form of contour plots of the growth rate difference \(\Delta\gamma\) versus angles \(\theta\) and \(\phi\).
+Figures \ref{fig:any_angle}a and \ref{fig:any_angle}b show the results of the analytic method in the \(E\) and \(F\) regions, respectively.
 The remaining 4 panels show the results from the numerical method in (c) the \(E\) region, (d) the \(F\) region, (e) transitional region, and (f) all regions combined.
-With exception of Figure \ref{fig:4}e, different values of \(\Delta\gamma\) along the gradient reversal were included in the averaging only for the points with specified virtual heights, e.g.
-115 km for Figure \ref{fig:4}c.
-In Figure \ref{fig:4}f, all points along the gradient reversal within the FoV were considered.
+With exception of Figure \ref{fig:any_angle}e, different values of \(\Delta\gamma\) along the gradient reversal were included in the averaging only for the points with specified virtual heights, e.g.
+115 km for Figure \ref{fig:any_angle}c.
+In Figure \ref{fig:any_angle}f, all points along the gradient reversal within the FoV were considered.
 
 The discontinuities in \(\Delta\gamma\) that occur on diagonal lines \(\phi=\theta\), \(\phi=\theta+180\deg\), \(\phi=\theta-180\deg\) are due to the fact that the convection velocity vector crosses the gradient reversal at these points and the leading and trailing edges of the gradient reversal switch.
-This feature is common in all panels in Figure \ref{fig:4}.
+This feature is common in all panels in Figure \ref{fig:any_angle}.
 On these discontinuities, the \(\Delta\gamma\) values always switch their sign, but their magnitude is the same on both sides of the discontinuity.
@@
 	\centering
-	\includegraphics[width=\textwidth,angle=180]{f04.pdf}
+	\includegraphics[width=\textwidth,angle=180]{p2f4.pdf}
 	\caption[GDI growth rates for arbitrary angle between \(\vec{V}_E\) and \(\vec{G}\)]{Contour plots of the growth rate difference \(\Delta\gamma\)  versus the two angles of interest.
@@
 The scale for \(R\) is shown on the right of panels (c) and (f).}
-	\label{fig:4}
+	\label{fig:any_angle}
 \end{figure}
 
-Figure \ref{fig:4}a shows the average growth rate difference across the gradient reversal found by the analytic method in the \(E\) region.
-The same information but for 4 specific cases of \(\phi-\theta\) were shown in Figure \ref{fig:3}.
-Specifically, each dashed red line in Figure \ref{fig:3} represents one diagonal cross section in Figure \ref{fig:4}a.
-For example, the red dashed line in Figure \ref{fig:3}a refers to the main diagonal line \(\phi=\theta\) and the red dashed line in Figure \ref{fig:3}d refers to the line \(\phi=\theta+90\deg\).
+Figure \ref{fig:any_angle}a shows the average growth rate difference across the gradient reversal found by the analytic method in the \(E\) region.
+The same information but for 4 specific cases of \(\phi-\theta\) were shown in Figure \ref{fig:set_angle}.
+Specifically, each dashed red line in Figure \ref{fig:set_angle} represents one diagonal cross section in Figure \ref{fig:any_angle}a.
+For example, the red dashed line in Figure \ref{fig:set_angle}a refers to the main diagonal line \(\phi=\theta\) and the red dashed line in Figure \ref{fig:set_angle}d refers to the line \(\phi=\theta+90\deg\).
 The peak magnitudes in \(\Delta\gamma\) occur at \(\theta=90\deg,270\deg\) and \(\phi=90\deg,270\deg\).
-This agrees with what is shown in Figure \ref{fig:3}a.
+This agrees with what is shown in Figure \ref{fig:set_angle}a.
 Physically, this means that the largest contrast in growth rate across the gradient reversal occurs when the reversal is passing through the radar FoV along the boresite and convection is parallel or antiparallel to the reversal.
@@
 
-Figure \ref{fig:4}b presents the same information but for the \(F\) region.
-This refers to the blue dashed lines in Figure \ref{fig:3}.
+Figure \ref{fig:any_angle}b presents the same information but for the \(F\) region.
+This refers to the blue dashed lines in Figure \ref{fig:set_angle}.
 In this case, the peak magnitudes in \(\Delta\gamma\) occur at \(\theta=90\deg,270\deg\) and \(\phi=0\deg,180\deg\).
@@
 
-Figure \ref{fig:4}c and d present the growth rate difference in the \(E\) and \(F\) regions, respectively, found with the numerical method.
-A comparison of Figures \ref{fig:4}a and b with Figures \ref{fig:4}c and d shows that the results of the analytic and numerical methods are very similar for both the \(E\) and \(F\) regions.
-The results for the numerical method (Figures \ref{fig:4}c and d) are not as smooth as the analytic method results, but this is simply due to the discrete steps that are used in the numerical method.
+Figure \ref{fig:any_angle}c and d present the growth rate difference in the \(E\) and \(F\) regions, respectively, found with the numerical method.
+A comparison of Figures \ref{fig:any_angle}a and b with Figures \ref{fig:any_angle}c and d shows that the results of the analytic and numerical methods are very similar for both the \(E\) and \(F\) regions.
+The results for the numerical method (Figures \ref{fig:any_angle}c and d) are not as smooth as the analytic method results, but this is simply due to the discrete steps that are used in the numerical method.
 Comparison between the \(E\) and \(F\) region patterns, however, reveals a few interesting features.
 First, the location of the minima moves from \(\phi=90\deg,270\deg\) in the \(E\) region to \(\phi=0\deg,180\deg\) in the \(F\) region.
-These results agree with what was observed in Figure \ref{fig:3}.
-In addition, there are large regions of strongly positive \(\Delta\gamma\) (dark red) in the \(E\) region (Figures \ref{fig:4}a and \ref{fig:4}c) that do not appear in the \(F\) region.
+These results agree with what was observed in Figure \ref{fig:set_angle}.
+In addition, there are large regions of strongly positive \(\Delta\gamma\) (dark red) in the \(E\) region (Figures \ref{fig:any_angle}a and \ref{fig:any_angle}c) that do not appear in the \(F\) region.
 There are some areas of positive \(\Delta\gamma\) in the \(F\) region, but they are much more limited and have a maximum value of only \(\sim 1\).
@@
 
-Figures \ref{fig:4}e shows results of the numerical analysis of the transitional region.
-This panel does not have a comparison with the analytic method because the analytic method requires a constant conductivity ratio \(R\) which changes rapidly from \(\sim\)5.08 in the \(E\) region to 0.001 in the \(F\) region, Figure \ref{fig:1}a.
+Figures \ref{fig:any_angle}e shows results of the numerical analysis of the transitional region.
+This panel does not have a comparison with the analytic method because the analytic method requires a constant conductivity ratio \(R\) which changes rapidly from \(\sim\)5.08 in the \(E\) region to 0.001 in the \(F\) region, Figure \ref{fig:model_setup}a.
 To demonstrate this, the average \(R\) value along the gradient reversal is plotted using the black-white dashed line and the scale shown on the right.
@@
 As angle \(\theta\) changes, the gradient reversal passes through different portions of the FoV with different heights.
-Different altitudes correspond to hugely different values of \(R\), which causes the variation in the conductivity ratio shown in Figure \ref{fig:4}e.
+Different altitudes correspond to hugely different values of \(R\), which causes the variation in the conductivity ratio shown in Figure \ref{fig:any_angle}e.
 On average, however, the conductivity ratio is somewhere in between its values in the \(E\) and \(F\) regions, so the pattern shown is between the \(E\) region pattern and \(F\) region pattern.
 The minimum is between \(\phi=90\deg\) (the \(E\) region value) and \(\phi=180\deg\) (the \(F\) region value) and has a value of \(\sim -3\).
-This will be further examined in Figure \ref{fig:5}.
+This will be further examined in Figure \ref{fig:transitional}.
 
-Finally, Figure \ref{fig:4}f presents the same analysis for all regions considered together, i.e.
+Finally, Figure \ref{fig:any_angle}f presents the same analysis for all regions considered together, i.e.
 with averaging done over all points along the gradient reversal regardless of their virtual height.
-For most angles of the gradient reversal, only the \(F\) region will be sampled, Figure \ref{fig:1}c.
+For most angles of the gradient reversal, only the \(F\) region will be sampled, Figure \ref{fig:model_setup}c.
 The gradient reversal will only pass through the transitional and \(E\) region if it is close to parallel with the boresite, i.e.
 when \(\theta\) is close to either \(90\deg\) or \(270\deg\).
-This is why the majority of the pattern in Figure \ref{fig:4}f resembles the \(F\) region patterns (Figure \ref{fig:4}d), but changes sharply around \(\theta=90\deg\) and \(\theta=270\deg\) due to contributions from the \(E\) region pattern (Figure \ref{fig:4}c).
+This is why the majority of the pattern in Figure \ref{fig:any_angle}f resembles the \(F\) region patterns (Figure \ref{fig:any_angle}d), but changes sharply around \(\theta=90\deg\) and \(\theta=270\deg\) due to contributions from the \(E\) region pattern (Figure \ref{fig:any_angle}c).
 This can also be seen in the conductivity ratio.
@@
 \subsection{Behavior in the Transitional Region}
-Figure \ref{fig:5} shows a more detailed analysis of the transitional region.
+Figure \ref{fig:transitional} shows a more detailed analysis of the transitional region.
 The transitional region was divided into segments that were 2 km in slant range.
-In each of these segments, the analysis presented in Figure \ref{fig:4} was performed and the average height and average conductivity ratio were found.
-Figure \ref{fig:5}a shows the average conductivity ratio (black line) and height (orange line) from 550 km to 850 km in slant range.
+In each of these segments, the analysis presented in Figure \ref{fig:any_angle} was performed and the average height and average conductivity ratio were found.
+Figure \ref{fig:transitional}a shows the average conductivity ratio (black line) and height (orange line) from 550 km to 850 km in slant range.
 The ranges where the beam is in the \(E\) (red), \(F\) (blue), and transitional (green) regions are shaded.
-Both the height and the conductivity ratio are constant in the \(E\) and \(F\) regions since virtual height is constant at these ranges in the standard virtual height model, Figure \ref{fig:1}b.
+Both the height and the conductivity ratio are constant in the \(E\) and \(F\) regions since virtual height is constant at these ranges in the standard virtual height model, Figure \ref{fig:model_setup}b.
 The height increases linearly in the transitional region, while the conductivity ratio decreases rapidly between 600 km and 616 km (shaded in yellow) at the lowest ranges within the transitional region.
 The \(R\) value continues to decrease throughout the transitional region, but above \(\sim620\) km in the slant range, it is so close to zero that further changes do not affect the results appreciably.
-The region highlighted in yellow, where most of the change in \(R\) occurs, is examined more closely in Figures \ref{fig:5}b and \ref{fig:5}c.
+The region highlighted in yellow, where most of the change in \(R\) occurs, is examined more closely in Figures \ref{fig:transitional}b and \ref{fig:transitional}c.
 
@@
 	\centering
-	\includegraphics[width=\textwidth,angle=180]{f05.pdf}
+	\includegraphics[width=\textwidth,angle=180]{p2f5.pdf}
 	\caption[GDI growth rate within transitional region]{Analysis of the transitional region.
@@
 In addition to the \(R\) and height continuous variations, it also gives the average value of \(R\) for each 2-km virtual range gate.
-The same 2-km virtual gates are considered in panel (c) that shows a similar analysis to Figure \ref{fig:4}.
+The same 2-km virtual gates are considered in panel (c) that shows a similar analysis to Figure \ref{fig:any_angle}.
 The top row shows the results of the numerical approach (N) and the bottom row shows the analytic approach results (A).}
-	\label{fig:5}
+	\label{fig:transitional}
 \end{figure}
 
-Figure \ref{fig:5}b has the same format as Figure \ref{fig:5}a, but it focuses on the 16-km-wide region in the slant range highlighted with yellow.
+Figure \ref{fig:transitional}b has the same format as Figure \ref{fig:transitional}a, but it focuses on the 16-km-wide region in the slant range highlighted with yellow.
 In these 16 km, the conductivity ratio (black line) decreases by approximately one order of magnitude, from \(\sim\)5 to 0.5.
@@
 
-Each small panel in Figure \ref{fig:5}c has the same format as Figure \ref{fig:4}a.
+Each small panel in Figure \ref{fig:transitional}c has the same format as Figure \ref{fig:any_angle}a.
 The top row shows the numerical analysis and the bottom show the analytic analysis of each 2 km segment.
-The value of \(R\) for the analytic method was taken from the average \(R\) for each 2 km segment, as shown in Figure \ref{fig:5}b.
-From 600--602 km, \(R\) is close to the \(E\) region value (\(\sim5.1\)), so the patterns for \(\Delta\gamma\) are close to those for the \(E\) region (Figures \ref{fig:4}a and c).
-However, by 614--616 km, the value of \(R\) has decreased substantially and the patterns resemble the \(F\)-region patterns (Figures \ref{fig:4}b and d) much more closely.
+The value of \(R\) for the analytic method was taken from the average \(R\) for each 2 km segment, as shown in Figure \ref{fig:transitional}b.
+From 600--602 km, \(R\) is close to the \(E\) region value (\(\sim5.1\)), so the patterns for \(\Delta\gamma\) are close to those for the \(E\) region (Figures \ref{fig:any_angle}a and c).
+However, by 614--616 km, the value of \(R\) has decreased substantially and the patterns resemble the \(F\)-region patterns (Figures \ref{fig:any_angle}b and d) much more closely.
 From 600 km to 616 km, the location of lowest value of \(\Delta\gamma\) changes from \(\phi=90\deg,270\deg\) to \(\phi=0\deg,180\deg\).
@@
 In addition, the lowest value of \(\Delta\gamma\) also increases from less than \(-8\) (\(E\) region) to \(\sim-3\) (\(F\) region).
-The areas of highly positive \(\Delta\gamma\) that appeared in the \(E\) region (Figures \ref{fig:4}a and c) are not evident above 606 km.
+The areas of highly positive \(\Delta\gamma\) that appeared in the \(E\) region (Figures \ref{fig:any_angle}a and c) are not evident above 606 km.
 This indicates that after this point, there is no longer a significant subset of orientations where the growth rate on the leading edge is larger than that on the trailing edge.
@@
 
-The results from Figure \ref{fig:4} show that \(\Delta\gamma\) is almost always negative in the \(F\) region, which implies that the growth rate on the trailing edge of a large-scale density structure is almost always larger than that on the leading edge.
+The results from Figure \ref{fig:any_angle} show that \(\Delta\gamma\) is almost always negative in the \(F\) region, which implies that the growth rate on the trailing edge of a large-scale density structure is almost always larger than that on the leading edge.
 This agrees with previous observations of polar patches and sun-aligned arcs in the \(F\) region ionosphere that consistently show a greater irregularity intensity on the trailing edge than the leading edge of these structures \citep{Weber1984,Milan2002a,Koustov2012}.
@@
 \citet{Koustov2008} presented data in their figure 5 that showed smaller echo power on the leading edge of a sun-aligned arc in the \(F\) region, but strong backscatter power on both sides of the arc in the short range gates, corresponding to the \(E\)-region or transitional region.
-The arc was approximately parallel to the radar boresite and moved towards the left of the radar FoV, corresponding to \(\Delta\gamma\approx 0\) in our model (\(\theta=90\deg,\phi\approx 180\deg\) in figure \ref{fig:4}c).
+The arc was approximately parallel to the radar boresite and moved towards the left of the radar FoV, corresponding to \(\Delta\gamma\approx 0\) in our model (\(\theta=90\deg,\phi\approx 180\deg\) in figure \ref{fig:any_angle}c).
 Slight changes in the convection direction (\(\phi\)) could create both the condition where \(\Delta\gamma>0\) (greater growth on leading edge) or \(\Delta\gamma<0\) (greater growth on trailing edge).
@@
 
-In this section, we derive an analytic expression for the average difference in the GDI growth rate across the gradient reversal passing through the radar's viewing area, Figure \ref{fig:1}c.
+In this section, we derive an analytic expression for the average difference in the GDI growth rate across the gradient reversal passing through the radar's viewing area, Figure \ref{fig:model_setup}c.
 The starting point is Equation (\ref{eqn:Mak14gr}).
 Consider a single point that lies on the gradient reversal within the radar's FoV.
-In the geometry illustrated in Figure \ref{fig:1}c, the relevant vectors at this point can be written as \(\vec{\hat k}=-\sin\alpha\vec{\hat{e}}_1-\cos\alpha\vec{\hat{e}}_2\), \(\vec{\hat e}=-\sin\phi\vec{\hat{e}}_1+\cos\phi\vec{\hat{e}}_2\), \(\hat{\vec{g}}=\pm[\sin\theta\vec{\hat{e}}_1-\cos\theta\vec{\hat{e}}_2]\), where angles \(\theta\) and \(\phi\) were previously defined in Section \ref{sec:model_fov}, angle \(\alpha\) is the angle between the wave vector and the boresite direction, and unit vectors \(\vec{\hat{e}}_1\)  and \(\vec{\hat{e}}_2\) are  perpendicular and parallel to the boresite, respectively.
+In the geometry illustrated in Figure \ref{fig:model_setup}c, the relevant vectors at this point can be written as \(\vec{\hat k}=-\sin\alpha\vec{\hat{e}}_1-\cos\alpha\vec{\hat{e}}_2\), \(\vec{\hat e}=-\sin\phi\vec{\hat{e}}_1+\cos\phi\vec{\hat{e}}_2\), \(\hat{\vec{g}}=\pm[\sin\theta\vec{\hat{e}}_1-\cos\theta\vec{\hat{e}}_2]\), where angles \(\theta\) and \(\phi\) were previously defined in Section \ref{sec:model_fov}, angle \(\alpha\) is the angle between the wave vector and the boresite direction, and unit vectors \(\vec{\hat{e}}_1\)  and \(\vec{\hat{e}}_2\) are  perpendicular and parallel to the boresite, respectively.
 

### 2017-02-02 paper3.tex
Formatting corrections made

@@
-%% 11/23/2015
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% AGUJournalTemplate.tex: this template file is for articles formatted with LaTeX
-%
-% This file includes commands and instructions
-% given in the order necessary to produce a final output that will
-% satisfy AGU requirements. 
-%
-% You may copy this file and give it your
-% article name, and enter your text.
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% PLEASE DO NOT USE YOUR OWN MACROS
-% DO NOT USE \newcommand, \renewcommand, or \def, etc.
-%
-% FOR FIGURES, DO NOT USE \psfrag or \subfigure.
-% DO NOT USE \psfrag or \subfigure commands.
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-% All questions should be e-mailed to latex@agu.org.
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-% Step 1: Set the \documentclass
-%
-% There are two options for article format:
-%
-% 1) PLEASE USE THE DRAFT OPTION TO SUBMIT YOUR PAPERS.
-% The draft option produces double spaced output.
-% 
-% 2) numberline will give you line numbers.
+\chapter[Radar observations of density gradients, electric fields, and plasma irregularities near polar cap patches in the context of the gradient-drift instability]{Radar observations of density gradients, electric fields, and plasma irregularities near polar cap patches in the context of the gradient-drift instability\footnote{Submitted as: {Lamarche}, L.~J., and R.~A.
+{Makarevich} (2017), Radar observations of density gradients, electric fields, and plasma irregularities near polar cap patches in the context of the gradient-drift instability, \textit{J.
+Geophys.
+Res.
+Space Physics}.}}
 
-%% To submit your paper:
-%\documentclass[draft,linenumbers]{agujournal}
-%\draftfalse
+\label{sec:paper3}
 
-%\usepackage{graphicx}
-%\usepackage{amsmath}
-%\usepackage{lineno}
-%
-%% For final version.
-% \documentclass{agujournal}
-
-% Now, type in the journal name: \journalname{<Journal Name>}
-
-% ie, \journalname{Journal of Geophysical Research}
-%% Choose from this list of Journals:
-%
-% JGR-Atmospheres
-% JGR-Biogeosciences
-% JGR-Earth Surface
-% JGR-Oceans
-% JGR-Planets
-% JGR-Solid Earth
-% JGR-Space Physics
-% Global Biochemical Cycles
-% Geophysical Research Letters
-% Paleoceanography
-% Radio Science
-% Reviews of Geophysics
-% Tectonics
-% Space Weather
-% Water Resource Research
-% Geochemistry, Geophysics, Geosystems
-% Journal of Advances in Modeling Earth Systems (JAMES)
-% Earth's Future
-% Earth and Space Science
-%
-%
-%
-%\journalname{JGR-Space Physics}
-%
-%
-%\begin{document}
-%
-%% ------------------------------------------------------------------------ %%
-%  Title
-% 
-% (A title should be specific, informative, and brief. Use
-% abbreviations only if they are defined in the abstract. Titles that
-% start with general keywords then specific terms are optimized in
-% searches)
-%
-%% ------------------------------------------------------------------------ %%
-
-% Example: \title{This is a test title}
-
-%\title{Radar observations of density gradients, electric fields, and plasma irregularities near polar cap patches in the context of the gradient-drift instability}
-\chapter{Radar observations of density gradients, electric fields, and plasma irregularities near polar cap patches in the context of the gradient-drift instability}
-
-%% ------------------------------------------------------------------------ %%
-%
-%  AUTHORS AND AFFILIATIONS
-%
-%% ------------------------------------------------------------------------ %%
-
-% Authors are individuals who have significantly contributed to the
-% research and preparation of the article. Group authors are allowed, if
-% each author in the group is separately identified in an appendix.)
-
-% List authors by first name or initial followed by last name and
-% separated by commas. Use \affil{} to number affiliations, and
-% \thanks{} for author notes.  
-% Additional author notes should be indicated with \thanks{} (for
-% example, for current addresses).
-
-% Example: \authors{A. B. Author\affil{1}\thanks{Current address, Antartica}, B. C. Author\affil{2,3}, and D. E.
-% Author\affil{3,4}\thanks{Also funded by Monsanto.}}
-
-%\authors{Leslie J. Lamarche\affil{1} and Roman A. Makarevich\affil{1}}
-
-% \affiliation{1}{First Affiliation}
-% \affiliation{2}{Second Affiliation}
-% \affiliation{3}{Third Affiliation}
-% \affiliation{4}{Fourth Affiliation}
-
-%\affiliation{1}{Geophysical Institute, University of Alaska Fairbanks, 903 Koyukuk Drive, PO Box 757320, Fairbanks, AK, 99775-7320, USA}
-%(repeat as many times as is necessary)
-
-%% Corresponding Author:
-% Corresponding author mailing address and e-mail address:
-
-% (include name and email addresses of the corresponding author.  More
-% than one corresponding author is allowed in this LaTeX file and for
-% publication; but only one corresponding author is allowed in our
-% editorial system.)
-
-% Example: \correspondingauthor{First and Last Name}{email@address.edu}
-
-%\correspondingauthor{Leslie Lamarche}{ljlamarche@alaska.edu}
-
-%% Keypoints, final entry on title page.
-
-% Example: 
-% \begin{keypoints}
-% \item	List up to three key points (at least one is required)
-% \item	Key Points summarize the main points and conclusions of the article
-% \item	Each must be 100 characters or less with no special characters or punctuation 
-% \end{keypoints}
-
-%  List up to three key points (at least one is required)
-%  Key Points summarize the main points and conclusions of the article
-%  Each must be 100 characters or less with no special characters or punctuation
-
-%\begin{keypoints}
-%\item Radar echo power of small-scale structures increases as large-scale gradient magnitude increases, implying a connection between scale sizes
-%\item Directional dependence of small-scale structures is observed on all sides of a polar cap patch and not restricted to the leading/trailing edge 
-%\item Gradient-drift instability theory describes echo occurrence behavior including its directional dependence better than echo strength 
-%\end{keypoints}
-
-%% ------------------------------------------------------------------------ %%
-%
-%  ABSTRACT
-%
-% A good abstract will begin with a short description of the problem
-% being addressed, briefly describe the new data or analyses, then
-% briefly states the main conclusion(s) and how they are supported and
-% uncertainties. 
-%% ------------------------------------------------------------------------ %%
-
-%% \begin{abstract} starts the second page 
-%\begin{abstract}
 \section*{Abstract}
@@
 The observed greater consistency of echo occurrence with GDI predictions suggests that GDI operating in the linear regime can control basic plasma structuring, but measured echo strength may be affected by other processes and factors, such as multi-step or nonlinear processes or a shear-driven instability.
-%\end{abstract}
-
-%% ------------------------------------------------------------------------ %%
-%
-%  TEXT
-%
-%% ------------------------------------------------------------------------ %%
-
-%%% Suggested section heads:
-% \section{Introduction}
-% 
-% The main text should start with an introduction. Except for short
-% manuscripts (such as comments and replies), the text should be divided
-% into sections, each with its own heading.
-
-% Headings should be sentence fragments and do not begin with a
-% lowercase letter or number. Examples of good headings are:
-
-% \section{Materials and Methods}
-% Here is text on Materials and Methods.
-%
-% \subsection{A descriptive heading about methods}
-% More about Methods.
-% 
-% \section{Data} (Or section title might be a descriptive heading about data)
-% 
-% \section{Results} (Or section title might be a descriptive heading about the
-% results)
-% 
-% \section{Conclusions}
 
@@
 
-%\begin{linenomath*}
 \begin{equation}
-  \label{eqn:gdi}
+  \label{eqn:p3gdi}
   \gamma = \frac{1}{B}\left(\uvec{k}\cdot\vec{E}\right)\left(\uvec{k}\cdot\uvec{b}\times\vec{G}\right) = GV_E\left(\uvec{k}\cdot\uvec{e}\right)\left(\uvec{k}\cdot\uvec{b}\times\uvec{g}\right)
 \end{equation}
-%\end{linenomath*}
 
-Equation (\ref{eqn:gdi}) describes the dependence on the magnitudes of the electric field $E$ or, equivalently, convection speed \(V_E=E/B\) and the density gradient strength \(G\).
+Equation (\ref{eqn:p3gdi}) describes the dependence on the magnitudes of the electric field $E$ or, equivalently, convection speed \(V_E=E/B\) and the density gradient strength \(G\).
 In addition, it describes a dependence on the purely directional factors \(\uvec{k}\cdot\uvec{e}\) and \(\uvec{k}\cdot\uvec{b}\times\uvec{g}\).
@@
 Previous studies that evaluated plasma density gradients have found them primarily from in situ measurements from sounding rockets \citep{Moen2012,Lynch2015}, but that method inherently only allows one component of the gradient to be resolved and only allows for measurements along the rocket's flight path.
-The new method employed here allows a field of 2D gradient vectors to be calculated, which, together with the 2D electric fields also provided by ISR, allows for experimental investigations of plasma structuring in context of factors \(\uvec{k}\cdot\vec{E}\) and \(\uvec{k}\cdot\uvec{b}\times\vec{G}\) as well their directional equivalents \(\uvec{k}\cdot\uvec{e}\) and \(\uvec{k}\cdot\uvec{b}\times\uvec{g}\), Equation (\ref{eqn:gdi}).
+The new method employed here allows a field of 2D gradient vectors to be calculated, which, together with the 2D electric fields also provided by ISR, allows for experimental investigations of plasma structuring in context of factors \(\uvec{k}\cdot\vec{E}\) and \(\uvec{k}\cdot\uvec{b}\times\vec{G}\) as well their directional equivalents \(\uvec{k}\cdot\uvec{e}\) and \(\uvec{k}\cdot\uvec{b}\times\uvec{g}\), Equation (\ref{eqn:p3gdi}).
 
@@
 
-Figure \ref{fig:f1} shows the FoV of both RKN, located at 62.82\(\deg\)N, 93.11\(\deg\)W, geographic (72.6\(\deg\)N, 26.4\(\deg\)W, geomagnetic), and RISR-N, located at 74.73\(\deg\)N, 94.91\(\deg\)W, geographic (82.8\(\deg\)N, 36.8\(\deg\)W geomagnetic).
+Figure \ref{fig:radar_map} shows the FoV of both RKN, located at 62.82\(\deg\)N, 93.11\(\deg\)W, geographic (72.6\(\deg\)N, 26.4\(\deg\)W, geomagnetic), and RISR-N, located at 74.73\(\deg\)N, 94.91\(\deg\)W, geographic (82.8\(\deg\)N, 36.8\(\deg\)W geomagnetic).
 Geomagnetic coordinates were determined using the Altitude Adjusted Corrected Geomagnetic Coordinates (AACGM) model \citep{Shepard2014}.
@@
 \begin{figure}
-	\includegraphics[width=\textwidth]{fig1.pdf}
-  \caption{Experimental setup showing the FoVs of RKN and RISR-N.
+	\includegraphics[width=\textwidth]{p3f1.pdf}
+  \caption[RKN/RISR-N Experimental Setup]{Experimental setup showing the FoVs of RKN and RISR-N.
 Lines of constant magnetic latitude are shown at \(75\deg\)N, \(80\deg\)N, and \(85\deg\)N.
@@
 RKN beam 7, which lies along \(335\deg\)E MLON, is outlined in black.}
-  \label{fig:f1}
+  \label{fig:radar_map}
 \end{figure}
@@
 To calculate gradients, RISR-N density measurements at 300 km were first converted to a local Cartesian coordinate system and then interpolated to a grid with 10-km resolution in both the \(x\) and \(y\) directions.
-The result of the interpolation is shown as filled contours in Figure \ref{fig:f1}, also colored in electron density according to the color bar in the top-right corner.
+The result of the interpolation is shown as filled contours in Figure \ref{fig:radar_map}, also colored in electron density according to the color bar in the top-right corner.
 2D density gradient vectors were found by considering the average density change over two adjacent grid points in both the \(x\) and \(y\) directions.
@@
 \section{Event Overview and Velocity Analysis}
-The snapshot shown in Figure \ref{fig:f1} is part of the event on November 30, 2011.
+The snapshot shown in Figure \ref{fig:radar_map} is part of the event on November 30, 2011.
 This event was selected because it included a high-density structure that passed through the common FoV area of RKN and RISR-N.
@@
 It is observed in the local noon sector moving approximately poleward (antisunwards), so it can be categorized as a polar cap patch.
-The event is summarized in Figure \ref{fig:f2}.
-Figure \ref{fig:f2}a shows a north-south keogram of electron density as measured along \(335\deg\) MLON by RISR-N at 17:00--19:00 UT.
+The event is summarized in Figure \ref{fig:event_overview}.
+Figure \ref{fig:event_overview}a shows a north-south keogram of electron density as measured along \(335\deg\) MLON by RISR-N at 17:00--19:00 UT.
 The high-density structure that is of particular interest occurred at \(\sim\)18:15 UT and is well defined by the two oblique black lines.
@@
 \begin{figure}
-	\includegraphics[width=\textwidth,angle=180]{fig2.pdf}
-  \caption{Left column is a keogram representation of (a and b) RISR-N electron density, (c) RKN line-of-sight velocity, and (d) RKN SNR for an event on November 30, 2011, 17:00--19:00 UT.
+	\includegraphics[width=\textwidth,angle=180]{p3f2.pdf}
+  \caption[Polar Patch Event Overview]{Left column is a keogram representation of (a and b) RISR-N electron density, (c) RKN line-of-sight velocity, and (d) RKN SNR for an event on November 30, 2011, 17:00--19:00 UT.
 Color scales for all quantities are given to the right of the respective row.
@@
 2D plasma drift velocity vectors are shown at \(335\deg\)E MLON.}
-  \label{fig:f2}
+  \label{fig:event_overview}
 \end{figure}
 
-Figure \ref{fig:f2}b shows the same keogram, but with 2D velocity vectors calculated from RISR-N measurements superimposed over the density data every 3 min.
+Figure \ref{fig:event_overview}b shows the same keogram, but with 2D velocity vectors calculated from RISR-N measurements superimposed over the density data every 3 min.
 Velocity data is available at \(\sim\)1-min intervals and these higher-resolution data were used for later analysis, but a subset is shown here to keep the plot readable.
 In the structure region, plasma velocity is primarily northwards with a small eastwards component, indicating that the structure is moving north-east.
-Figure \ref{fig:f2}c shows the line-of-sight velocity measured in RKN beam 7, which is aligned with the \(335\deg\) MLON meridian, Figure \ref{fig:f1}.
-Contours of electron density from RISR-N (same as in Figure \ref{fig:f2}a) are shown over the RKN data.
-Figure \ref{fig:f2}d has the same format as Figure \ref{fig:f2}c, but shows the SNR measured by RKN.
+Figure \ref{fig:event_overview}c shows the line-of-sight velocity measured in RKN beam 7, which is aligned with the \(335\deg\) MLON meridian, Figure \ref{fig:radar_map}.
+Contours of electron density from RISR-N (same as in Figure \ref{fig:event_overview}a) are shown over the RKN data.
+Figure \ref{fig:event_overview}d has the same format as Figure \ref{fig:event_overview}c, but shows the SNR measured by RKN.
 Most of the RKN echoes occur along the steep transition between the high- and low-density regions.
 The velocity data indicates that the structure was moving consistently away from the radar with a speed of \(\sim\)500 m/s.
-If the structure is moving north-east, as indicated in Figure \ref{fig:f2}b, the greatest power is observed south of it, or on the trailing edge.
+If the structure is moving north-east, as indicated in Figure \ref{fig:event_overview}b, the greatest power is observed south of it, or on the trailing edge.
 
-Figure \ref{fig:f2}e contains sixteen 2D snapshot views of the event at 18:09--18:27 UT (period shown by the pink bars at the top of Figures \ref{fig:f2}a--\ref{fig:f2}d).
-Each snapshot is similar to that shown in Figure \ref{fig:f1} with SNR measured by RKN shown by color-filled radar cells, RISR-N electron density measurements shown as colored circles, interpolated electron density as colored contours, and plasma velocity as a line of 2D vectors at \(335\deg\) MLON.
+Figure \ref{fig:event_overview}e contains sixteen 2D snapshot views of the event at 18:09--18:27 UT (period shown by the pink bars at the top of Figures \ref{fig:event_overview}a--\ref{fig:event_overview}d).
+Each snapshot is similar to that shown in Figure \ref{fig:radar_map} with SNR measured by RKN shown by color-filled radar cells, RISR-N electron density measurements shown as colored circles, interpolated electron density as colored contours, and plasma velocity as a line of 2D vectors at \(335\deg\) MLON.
 In this view, the interpolated density contours are left unfilled so as to not obscure the RKN SNR data.
 
-From velocity vectors in Figure \ref{fig:f2}e, the plasma drifts mostly northwards at the beginning of the time interval and more eastwards later during the event, similar to what is seen in Figure \ref{fig:f2}b.
+From velocity vectors in Figure \ref{fig:event_overview}e, the plasma drifts mostly northwards at the beginning of the time interval and more eastwards later during the event, similar to what is seen in Figure \ref{fig:event_overview}b.
 This results in the structure moving in a generally north-east direction, which can be also seen by looking at the interpolated density contours.
@@
 This generally includes the entire FoV of RISR-N with coincident RKN echoes.
-From Figure \ref{fig:f1}, this includes RKN beams 4--10 and up to 9 range gates.
+From Figure \ref{fig:radar_map}, this includes RKN beams 4--10 and up to 9 range gates.
 This allows large portions of the density structures to be considered.
 In some cases, e.g.
-Figure \ref{fig:f2}e at 18:20 UT, the entire patch exists within the observable area and all edges can be considered.
+Figure \ref{fig:event_overview}e at 18:20 UT, the entire patch exists within the observable area and all edges can be considered.
 In other cases, radars sample only parts of density structures.
@@
 However, certain biases may still be present because this study is dominated by a single strong density enhancement and the patch moved through the FoV with a relatively constant velocity.
-These biases are discussed further in Section \ref{sec:discussion2}.
+These biases are discussed further in Section \ref{sec:p3discussion2}.
 
@@
 In addition to the standard SuperDARN method for eliminating ground scatter, this study further required that the velocities measured by RKN and RISR-N be comparable, since low SuperDARN velocities are one of the characteristics of ground scatter.
-Figure \ref{fig:f3} is a scatter plot of coincident RKN and RISR-N velocities.
-In this figure and all subsequent ones in this paper, black points refer to all observations throughout the event where there were coincident RISR-N and RKN echoes and red points represent only those echoes observed within the structure identified by the two oblique black lines in Figure \ref{fig:f2}.
+Figure \ref{fig:velocity} is a scatter plot of coincident RKN and RISR-N velocities.
+In this figure and all subsequent ones in this paper, black points refer to all observations throughout the event where there were coincident RISR-N and RKN echoes and red points represent only those echoes observed within the structure identified by the two oblique black lines in Figure \ref{fig:event_overview}.
 Because RKN only measures velocity along the radar beam, RKN velocities are compared with RISR-N velocities projected onto the appropriate RKN beam direction.
@@
 \begin{figure}
-	\includegraphics[width=\textwidth,angle=180]{fig3.pdf}
-  \caption{RKN line-of-sight velocity versus the projection of 2D RISR-N velocity onto the closest RKN beam.
+	\includegraphics[width=\textwidth,angle=180]{p3f3.pdf}
+  \caption[RKN/RISR-N Velocity Analysis]{RKN line-of-sight velocity versus the projection of 2D RISR-N velocity onto the closest RKN beam.
 Black dots represent all coincident echoes that were observed by RKN and RISR-N simultaneously over the day considered, whereas red dots represent only echoes that occurred within the high-density structure.
@@
 The solid blue lines outline a region where the RKN velocity equals the RISR-N velocity within \(\pm\)500 m/s.}
-  \label{fig:f3}
+  \label{fig:velocity}
 \end{figure}
 
-In Figure \ref{fig:f3}, there is a clustering of all echoes (black) around (0 m/s, 0 m/s) and a clustering of structure echoes (red) around (\(-700\) m/s, \(-700\) m/s).
+In Figure \ref{fig:velocity}, there is a clustering of all echoes (black) around (0 m/s, 0 m/s) and a clustering of structure echoes (red) around (\(-700\) m/s, \(-700\) m/s).
 The dashed blue line shows where RKN velocity equals the projected RISR-N velocity.
@@
 \section{Plasma Irregularities and Density-Related Factors}
-In order to investigate how density-related factors may affect echo power and occurrence, Figure \ref{fig:f4} shows scatter plots of echo power versus (a) electron density, \(n\), (b), electron density gradient magnitude, \(|\nabla n|\), and (c), and gradient strength, \(G = |\nabla n/n|\).
+In order to investigate how density-related factors may affect echo power and occurrence, Figure \ref{fig:density} shows scatter plots of echo power versus (a) electron density, \(n\), (b), electron density gradient magnitude, \(|\nabla n|\), and (c), and gradient strength, \(G = |\nabla n/n|\).
 Points are binned and averages (crosses) and standard deviations (shaded rectangles) are shown for both all echoes and only echoes within the structure.
@@
 \begin{figure}
-\includegraphics[width=\textwidth,angle=180]{fig4.pdf}
-  \caption{Scatter plots of RKN echo power versus RISR-N measurements of (a) electron density, \(n\), (b) density gradient magnitude, \(|\nabla n|\), and (c) gradient strength, \(G=|\nabla n/n|\).
+\includegraphics[width=\textwidth,angle=180]{p3f4.pdf}
+  \caption[Irregularity Dependence on Density Factors]{Scatter plots of RKN echo power versus RISR-N measurements of (a) electron density, \(n\), (b) density gradient magnitude, \(|\nabla n|\), and (c) gradient strength, \(G=|\nabla n/n|\).
 Black dots represent all echoes observed over the day, whereas red dots represent echoes that occur within the high-density structure.
@@
 Thick black and red lines are point count histograms; scale is on the right axis.}
-  \label{fig:f4}
+  \label{fig:density}
 \end{figure}
 
-In Figure \ref{fig:f4}a, most points are clustered at \(2\times10^{11}\)--\(5\times10^{11}\) \# m\(^{-3}\), corresponding to a peak in the all-counts histogram (black).
+In Figure \ref{fig:density}a, most points are clustered at \(2\times10^{11}\)--\(5\times10^{11}\) \# m\(^{-3}\), corresponding to a peak in the all-counts histogram (black).
 The structure-count histogram (red) has a secondary peak at this location, but its highest peak occurs at a much higher density, around \(1.1\times10^{12}\) \# m\(^{-3}\).
@@
 
-Figure \ref{fig:f4}b shows the dependence on density gradient magnitude, \(|\nabla n|\).
+Figure \ref{fig:density}b shows the dependence on density gradient magnitude, \(|\nabla n|\).
 For all echoes, counts peak at low gradients and decrease.
@@
 
-Finally, Figure \ref{fig:f4}c shows the dependence on gradient strength, defined as \(G = |\nabla n/n|\).
+Finally, Figure \ref{fig:density}c shows the dependence on gradient strength, defined as \(G = |\nabla n/n|\).
 In this case, both all-echo and structure-echo counts peak at about \(2\times10^{-6}\) m\(^{-1}\).
 Similar to the density gradient, average power for both all echoes and structure echoes increases as gradient strength increases.
-Interestingly, this trend is actually clearer and stronger in the gradient case (Figure \ref{fig:f4}b) than the gradient strength case (Figure \ref{fig:f4}c), which is unexpected as the GDI growth rate expression depends directly on the gradient strength, \(\vec{G}\), not the density gradient vector, \(\nabla n\), equation \ref{eqn:gdi}.
-This feature is further discussed in Section \ref{sec:discussion1}.
+Interestingly, this trend is actually clearer and stronger in the gradient case (Figure \ref{fig:density}b) than the gradient strength case (Figure \ref{fig:density}c), which is unexpected as the GDI growth rate expression depends directly on the gradient strength, \(\vec{G}\), not the density gradient vector, \(\nabla n\), equation \ref{eqn:p3gdi}.
+This feature is further discussed in Section \ref{sec:p3discussion1}.
 
 \section{Plasma Irregularities and Directional Factors}
-Figure \ref{fig:f5} contains similar scatter plots to Figure \ref{fig:f4}, but these demonstrate the dependence of echo power and counts on the two factors that directly impact the GDI linear growth rate, the rotated gradient component along the wavevector, \(\uvec{k}\cdot\uvec{b}\times\vec{G}\), and the electric field component along the wavevector, \(\uvec{k}\cdot\vec{E}\).
-An increase in either of these factors should theoretically be reflected in higher GDI growth rates, Equation \ref{eqn:gdi}, and this could occur through increase in the entire factor or its purely directional part, e.g.
+Figure \ref{fig:directional} contains similar scatter plots to Figure \ref{fig:density}, but these demonstrate the dependence of echo power and counts on the two factors that directly impact the GDI linear growth rate, the rotated gradient component along the wavevector, \(\uvec{k}\cdot\uvec{b}\times\vec{G}\), and the electric field component along the wavevector, \(\uvec{k}\cdot\vec{E}\).
+An increase in either of these factors should theoretically be reflected in higher GDI growth rates, Equation \ref{eqn:p3gdi}, and this could occur through increase in the entire factor or its purely directional part, e.g.
 \(\uvec{k}\cdot\uvec{e}\), with of both of these possibilities investigated below.
@@
 \begin{figure}
-	\includegraphics[width=\textwidth,angle=180]{fig5.pdf}
-  \caption{Scatter plots of RKN echo power versus RISR-N measurements of (a) the gradient strength component, \(\uvec{k}\cdot\uvec{b}\times\vec{G}\), (b) the electric field component \(\uvec{k}\cdot\vec{E}\), (c) the normalized gradient component, \(\uvec{k}\cdot\uvec{b}\times\uvec{g}\), and (d) the normalized electric field component, \(\uvec{k}\cdot\uvec{e}\).
-The meaning of dots, crosses, vertical bars, and lines is the same as in Figure \ref{fig:f4}.}
-  \label{fig:f5}
+	\includegraphics[width=\textwidth,angle=180]{p3f5.pdf}
+  \caption[Irregularity Dependence on Directional Factors]{Scatter plots of RKN echo power versus RISR-N measurements of (a) the gradient strength component, \(\uvec{k}\cdot\uvec{b}\times\vec{G}\), (b) the electric field component \(\uvec{k}\cdot\vec{E}\), (c) the normalized gradient component, \(\uvec{k}\cdot\uvec{b}\times\uvec{g}\), and (d) the normalized electric field component, \(\uvec{k}\cdot\uvec{e}\).
+The meaning of dots, crosses, vertical bars, and lines is the same as in Figure \ref{fig:density}.}
+  \label{fig:directional}
 \end{figure}
@@
 In this experiment, the wavevector direction \(\uvec{k}\) is defined by the orientation of the RKN radar beam examining a particular scattering volume because the radar will only observe plasma waves propagating in this particular direction.
-Therefore, the first factor in equation \ref{eqn:gdi} is the component of the 2D electric field measured by RISR-N along the coincident RKN beam and the second factor is the component of the rotated 2D plasma density gradient along the coincident RKN beam.
-Figure \ref{fig:f5}a shows the gradient component, \(\uvec{k}\cdot\uvec{b}\times\vec{G}\), and Figure \ref{fig:f5}b shows the electric field component, \(\uvec{k}\cdot\vec{E}\).
-The purely directional parts of these factors or the normalized component of the gradient and electric field are shown in Figures \ref{fig:f5}c and \ref{fig:f5}d, respectively.
+Therefore, the first factor in equation \ref{eqn:p3gdi} is the component of the 2D electric field measured by RISR-N along the coincident RKN beam and the second factor is the component of the rotated 2D plasma density gradient along the coincident RKN beam.
+Figure \ref{fig:directional}a shows the gradient component, \(\uvec{k}\cdot\uvec{b}\times\vec{G}\), and Figure \ref{fig:directional}b shows the electric field component, \(\uvec{k}\cdot\vec{E}\).
+The purely directional parts of these factors or the normalized component of the gradient and electric field are shown in Figures \ref{fig:directional}c and \ref{fig:directional}d, respectively.
 It is important to consider the normalized components independently to differentiate trends due to the directions of the electric field and gradients relative to the wavevector from those due to the magnitude of the electric field or gradient.
 
-In Figure \ref{fig:f5}a, the average power of all echoes tends to increase as the gradient component increases from zero.
+In Figure \ref{fig:directional}a, the average power of all echoes tends to increase as the gradient component increases from zero.
 At negative values, the power is smaller than at positive values and is approximately constant below 20 dB.
@@
 
-Figure \ref{fig:f5}b shows that, while the average power of all echoes is approximately constant with changing \(\uvec{k}\cdot\vec{E}\), the structure echoes exhibit a peak in power at \(\sim\)0.01 mV/m.
+Figure \ref{fig:directional}b shows that, while the average power of all echoes is approximately constant with changing \(\uvec{k}\cdot\vec{E}\), the structure echoes exhibit a peak in power at \(\sim\)0.01 mV/m.
 Both all echoes and structure echoes are biased in occurrence towards positive values of \(\uvec{k}\cdot\vec{E}\).
@@
 
-The purely directional part of the gradient, Figure \ref{fig:f5}c, does not appear to follow any trend in the average power of either all or structure echoes.
+The purely directional part of the gradient, Figure \ref{fig:directional}c, does not appear to follow any trend in the average power of either all or structure echoes.
 Structure average power is consistently higher, as expected, but both sets are approximately constant in power for all values of the normalized gradient component.
@@
 
-The purely directional part of the electric field, Figure \ref{fig:f5}d, again has average power approximately constant for all echoes, but average power for structure echoes peaks at \(\sim\)0.4.
+The purely directional part of the electric field, Figure \ref{fig:directional}d, again has average power approximately constant for all echoes, but average power for structure echoes peaks at \(\sim\)0.4.
 The occurrence distribution for all echoes again peaks strongly at \(\uvec{k}\cdot\uvec{e} = 1\) with a secondary peak at \(\uvec{k}\cdot\uvec{e} = -1\).
@@
 
-Figure \ref{fig:f6} further examines the RKN echo power and occurrence versus the (a) dot product of the gradient and plasma drift velocity \(\vec{G}\cdot\vec{V}_E\) and (b) the expected GDI growth rate, with both evaluated from RISR-N measurements.
-Similar to Figure \ref{fig:f5}, purely directional factors are shown for (c) the dot product of the gradient and the plasma drift velocity, \(\uvec{g}\cdot\mathbf{\hat{v}}_E\)  and (d) the GDI growth rate, given by \(\gamma^\prime \equiv \gamma/\left(GV_E\right) = (\uvec{k}\cdot\uvec{e})(\uvec{k}\cdot\uvec{b}\times\uvec{g})\), from equation (\ref{eqn:gdi}).
+Figure \ref{fig:gamma} further examines the RKN echo power and occurrence versus the (a) dot product of the gradient and plasma drift velocity \(\vec{G}\cdot\vec{V}_E\) and (b) the expected GDI growth rate, with both evaluated from RISR-N measurements.
+Similar to Figure \ref{fig:directional}, purely directional factors are shown for (c) the dot product of the gradient and the plasma drift velocity, \(\uvec{g}\cdot\mathbf{\hat{v}}_E\)  and (d) the GDI growth rate, given by \(\gamma^\prime \equiv \gamma/\left(GV_E\right) = (\uvec{k}\cdot\uvec{e})(\uvec{k}\cdot\uvec{b}\times\uvec{g})\), from equation (\ref{eqn:p3gdi}).
 The normalized components are again important to determine if the relative directions of \(\vec{G}\) and \(\vec{V}_E\) impact the power and occurrence of backscatter independent of their magnitude.
@@
 \begin{figure}
-	\includegraphics[width=\textwidth,angle=180]{fig6.pdf}
-  \caption{Same as Figure \ref{fig:f5}, but for (a) the dot product of the gradient and plasma drift velocity, \(\vec{G}\cdot\vec{V}_E\), (b) the calculated GDI growth rate, \(\gamma\), (c) the normalized dot product of the  gradient and plasma drift velocity, \(\uvec{g}\cdot\mathbf{\hat{v}}_E\), and (d) the normalized GDI growth rate, \(\gamma'\).}
-  \label{fig:f6}
+	\includegraphics[width=\textwidth,angle=180]{p3f6.pdf}
+  \caption[Irregularity dependence on GDI growth rate]{Same as Figure \ref{fig:directional}, but for (a) the dot product of the gradient and plasma drift velocity, \(\vec{G}\cdot\vec{V}_E\), (b) the calculated GDI growth rate, \(\gamma\), (c) the normalized dot product of the  gradient and plasma drift velocity, \(\uvec{g}\cdot\mathbf{\hat{v}}_E\), and (d) the normalized GDI growth rate, \(\gamma'\).}
+  \label{fig:gamma}
 \end{figure}
 
-In Figure \ref{fig:f6}a, the trend in the average power of either all echoes or structure echoes is not monotonic, but structure echoes seem to show a small increase in average power as \(\vec{G}\cdot\vec{V}_E\) increases, and all echoes tend to have higher power for positive values of \(\vec{G}\cdot\vec{V}_E\) than for negative values.
+In Figure \ref{fig:gamma}a, the trend in the average power of either all echoes or structure echoes is not monotonic, but structure echoes seem to show a small increase in average power as \(\vec{G}\cdot\vec{V}_E\) increases, and all echoes tend to have higher power for positive values of \(\vec{G}\cdot\vec{V}_E\) than for negative values.
 The occurrence distribution for all echoes is centered around 0, but counts are slightly higher for positive \(\vec{G}\cdot\vec{V}_E\).
 Occurrence for structure echoes is biased towards positive values of \(\vec{G}\cdot\vec{V}_E\).
-In Figure \ref{fig:f6}b, average power for all echoes is roughly constant and less than 20 dB when \(\gamma < 0\) s\(^{-1}\), but increases for positive values of \(\gamma\).
+In Figure \ref{fig:gamma}b, average power for all echoes is roughly constant and less than 20 dB when \(\gamma < 0\) s\(^{-1}\), but increases for positive values of \(\gamma\).
 The occurrence distribution of all echoes is again centered around zero, but counts tend to be slightly higher for positive \(\gamma\).
@@
 
-Turning to purely directional dependencies, in Figure \ref{fig:f6}c, average power for all echoes is relatively constant at \(\sim\)20 dB, while average power for structure echoes is higher, but also variable.
+Turning to purely directional dependencies, in Figure \ref{fig:gamma}c, average power for all echoes is relatively constant at \(\sim\)20 dB, while average power for structure echoes is higher, but also variable.
 Both occurrence histograms for all echoes and structure echoes have a strong peak where \(\uvec{g}\cdot\mathbf{\hat{v}}_E = 1\), i.e.
@@
 where the gradient is antiparallel to the drift velocity.
-Figure \ref{fig:f6}d shows that the average power is relatively insensitive to changes in the normalized growth rate, although power for both all echoes and structure echoes appears to have a peak around \(\gamma'=0.4\).
+Figure \ref{fig:gamma}d shows that the average power is relatively insensitive to changes in the normalized growth rate, although power for both all echoes and structure echoes appears to have a peak around \(\gamma'=0.4\).
 The occurrence histogram for all echoes shows that echoes are observed for all values of \(\gamma'\) and peak at zero, but the distribution is asymmetric such that more counts are observed for \(\gamma'\) positive than negative.
@@
 
-Figure \ref{fig:f7} investigates the same factors as Figure \ref{fig:f5}, but here, scatter plots show the echo power (given as the color) versus the gradient component \(\uvec{k}\cdot\uvec{b}\times\vec{G}\) and electric field component \(\uvec{k}\cdot\vec{E}\), with the color bar shown to the right of the figure.
-Figure \ref{fig:f7}a shows all echoes.
+Figure \ref{fig:eg_scatter} investigates the same factors as Figure \ref{fig:directional}, but here, scatter plots show the echo power (given as the color) versus the gradient component \(\uvec{k}\cdot\uvec{b}\times\vec{G}\) and electric field component \(\uvec{k}\cdot\vec{E}\), with the color bar shown to the right of the figure.
+Figure \ref{fig:eg_scatter}a shows all echoes.
 The solid black line in this and all other panels in this figure shows a linear fit to all points; equation is given at the bottom of the panel.
@@
 The linear Pearson correlation coefficient of all points is given in the bottom-right corner of the panel.
-Figure \ref{fig:f7}b shows power for all echoes versus purely directional factors \(\uvec{k}\cdot\uvec{b}\times\uvec{g}\) and \(\uvec{k}\cdot\uvec{e}\).
-Figures \ref{fig:f7}c and \ref{fig:f7}d present the same information, but only structure echoes are shown.
+Figure \ref{fig:eg_scatter}b shows power for all echoes versus purely directional factors \(\uvec{k}\cdot\uvec{b}\times\uvec{g}\) and \(\uvec{k}\cdot\uvec{e}\).
+Figures \ref{fig:eg_scatter}c and \ref{fig:eg_scatter}d present the same information, but only structure echoes are shown.
 Additionally in these two panels, the pink line shows the linear fit for points where SNR \(>\) 30 dB.
 The equation of the line and linear correlation of this subset of points are again given at the bottom of each panel in pink.
-In Figure \ref{fig:f7}c, the two linear fits are close to identical, so the pink line lies almost directly on top of the black line.
+In Figure \ref{fig:eg_scatter}c, the two linear fits are close to identical, so the pink line lies almost directly on top of the black line.
 
 \begin{figure}
-\includegraphics[width=\textwidth,angle=180]{fig7.pdf}
-  \caption{Scatter plots of RISR-N measurements of the electric field component, \(\uvec{k}\cdot\vec{E}\), versus the gradient component, \(\uvec{k}\cdot\uvec{b}\times\vec{G}\).
+\includegraphics[width=\textwidth,angle=180]{p3f7.pdf}
+  \caption[Relationship between the electric field and gradient components]{Scatter plots of RISR-N measurements of the electric field component, \(\uvec{k}\cdot\vec{E}\), versus the gradient component, \(\uvec{k}\cdot\uvec{b}\times\vec{G}\).
 Points are color coded in RKN SNR, according to the color bar to the right.
@@
 Equations for each line are given in the bottom-left corner of each panel and the correlation values are shown in the bottom-right corner.}
-  \label{fig:f7}
+  \label{fig:eg_scatter}
 \end{figure}
 
-In Figure \ref{fig:f7}a, points are clustered around \(\uvec{k}\cdot\uvec{b}\times\vec{G} = 0\) m\(^{-1}\) and \(\uvec{k}\cdot\vec{E} = 0.01\) mV/m.
-Figure \ref{fig:f7}b has less localized clustering of points.
-The bias of points toward positive values of \(\uvec{k}\cdot\vec{E}\) agrees with the results seen in Figures \ref{fig:f5}a and \ref{fig:f5}b.
-As seen in Figure \ref{fig:f7}c as well as previous figures, structure echoes tend to have higher power.
+In Figure \ref{fig:eg_scatter}a, points are clustered around \(\uvec{k}\cdot\uvec{b}\times\vec{G} = 0\) m\(^{-1}\) and \(\uvec{k}\cdot\vec{E} = 0.01\) mV/m.
+Figure \ref{fig:eg_scatter}b has less localized clustering of points.
+The bias of points toward positive values of \(\uvec{k}\cdot\vec{E}\) agrees with the results seen in Figures \ref{fig:directional}a and \ref{fig:directional}b.
+As seen in Figure \ref{fig:eg_scatter}c as well as previous figures, structure echoes tend to have higher power.
 The spread of the points is also substantially less for structure echoes than all echoes.
-This can be seen in the higher linear correlation in Figure \ref{fig:f7}c than Figure \ref{fig:f7}a.
-The same is true comparing Figures \ref{fig:f7}b and \ref{fig:f7}d, i.e.
-linear correlation is substantially higher in Figure \ref{fig:f7}d than Figure \ref{fig:f7}b.
+This can be seen in the higher linear correlation in Figure \ref{fig:eg_scatter}c than Figure \ref{fig:eg_scatter}a.
+The same is true comparing Figures \ref{fig:eg_scatter}b and \ref{fig:eg_scatter}d, i.e.
+linear correlation is substantially higher in Figure \ref{fig:eg_scatter}d than Figure \ref{fig:eg_scatter}b.
 When only high power echoes are considered (SRN \(>\) 30 dB), linear correlations increase even more.
-This can also be seen from a simple visual examination of Figures \ref{fig:f7}c and \ref{fig:f7}d, where the electric field component shows a clear increase with the gradient component, particularly for high-power echoes with SNR \(>\) 30 dB shown by red dots.
+This can also be seen from a simple visual examination of Figures \ref{fig:eg_scatter}c and \ref{fig:eg_scatter}d, where the electric field component shows a clear increase with the gradient component, particularly for high-power echoes with SNR \(>\) 30 dB shown by red dots.
 Note that, in all cases, the relationship between the two factors is approximately linear and slope of the linear fit is positive within uncertainty.
-This finding is further discussed in Section \ref{sec:discussion2}.
+This finding is further discussed in Section \ref{sec:p3discussion2}.
 
 \section{Discussion}
+\label{sec:p3discussion}
 In this study, small-scale plasma irregularity characteristics were analyzed in the context of factors that affect the linear growth rate of the gradient-drift instability.
@@
 \subsection{Effects of Density-Related Factors}
-\label{sec:discussion1}
+\label{sec:p3discussion1}
 The fluid linear theory of GDI predicts that an increase in plasma density gradient strength \(G\) will increase irregularity production through GDI \citep{Keskinen1983,Makarevich2014c}.
 The current experimental results were generally consistent with this expectation.
-In particular, the echo power showed an expected increase with \(G\), Figure \ref{fig:f4}c.
+In particular, the echo power showed an expected increase with \(G\), Figure \ref{fig:density}c.
 The point scatter was large but the increase was monotonic for bins with significant statistics.
@@
 High plasma density can also reduce the occurrence of backscatter observed by radar by smoothing the gradients required for GDI \citep{Ruohoniemi1997,Koustov2004}, through \(E\)-region shorting \citep{Dan02,Kan12,Lam15}, or through attenuation of the signal through a dense \(D\) region \citep{Danskin2002}.
-The peak in occurrence of all echoes in Figure \ref{fig:f4}a agrees with previous results that there is an optimal density for HF radar observations of \(F\)-region irregularities \citep{Danskin2002,Makarevich2014b,Koustov2014,Lamarche2015}.
+The peak in occurrence of all echoes in Figure \ref{fig:density}a agrees with previous results that there is an optimal density for HF radar observations of \(F\)-region irregularities \citep{Danskin2002,Makarevich2014b,Koustov2014,Lamarche2015}.
 
-In the current study, however, a nearly monotonic increase in average power of all echoes with density was also found, Figure \ref{fig:f4}a, which can potentially be explained through radar backscatter theory.
+In the current study, however, a nearly monotonic increase in average power of all echoes with density was also found, Figure \ref{fig:density}a, which can potentially be explained through radar backscatter theory.
 The backscatter power is known to be proportional to the perturbation intensity squared, \((\delta n)^2\), which can be represented as the product of the fractional perturbation density, \(\delta n/n\), and the background density, \(n\), squared.
@@
 
-The current study has also considered the effects of the magnitude of the density gradient, \(|\nabla n|\), and the gradient strength, \(G = |\nabla n/n|\), Figures \ref{fig:f4}b and \ref{fig:f4}c.
+The current study has also considered the effects of the magnitude of the density gradient, \(|\nabla n|\), and the gradient strength, \(G = |\nabla n/n|\), Figures \ref{fig:density}b and \ref{fig:density}c.
 The average power of both structure and all echoes showed an increase with both of these factors.
 However, the relationship between power and \(|\nabla n|\) was somewhat stronger than that between power and \(G\), particularly in the case of structure echoes.
-This is puzzling as the expression for the GDI growth rate, equation \ref{eqn:gdi}, is dependent on \(G\), not \(|\nabla n|\).
+This is puzzling as the expression for the GDI growth rate, equation \ref{eqn:p3gdi}, is dependent on \(G\), not \(|\nabla n|\).
 This may be due to the previously mentioned density effect.
@@
 Thus if the fractional density is proportional to the growth rate, \(\gamma\), instead of being approximately constant, backscatter power should be proportional to \((\gamma n)^2\).
-Because the growth rate is directly proportional to the magnitude of the gradient strength, equation (\ref{eqn:gdi}), the power will be proportional to the magnitude of the density gradient squared:
+Because the growth rate is directly proportional to the magnitude of the gradient strength, equation (\ref{eqn:p3gdi}), the power will be proportional to the magnitude of the density gradient squared:
 
@@
 \subsection{Directional Effects of Electric Field and Gradient Components}
-\label{sec:discussion2}
-One important new aspect of the current study is the experimental investigation of the directional dependence of the GDI growth rate, as represented by equation  (\ref{eqn:gdi}).
+\label{sec:p3discussion2}
+One important new aspect of the current study is the experimental investigation of the directional dependence of the GDI growth rate, as represented by equation  (\ref{eqn:p3gdi}).
 This directional dependence has been previously predicted theoretically \citep{Keskinen1982a,Makarevich2014c,Lamarche2016}, but no experimental evidence is currently available, apart from asymmetry between leading and trailing edges of large-scale density structures.
-Because the growth rate \(\gamma\) is directly dependent on both the gradient component, \(\uvec{k}\cdot\uvec{b}\times\vec{G}\), and the electric field component, \(\uvec{k}\cdot\vec{E}\), the relative directions of the gradient, electric field, and wavevector are predicted to strongly influence plasma structuring in the ionosphere, equation (\ref{eqn:gdi}).
+Because the growth rate \(\gamma\) is directly dependent on both the gradient component, \(\uvec{k}\cdot\uvec{b}\times\vec{G}\), and the electric field component, \(\uvec{k}\cdot\vec{E}\), the relative directions of the gradient, electric field, and wavevector are predicted to strongly influence plasma structuring in the ionosphere, equation (\ref{eqn:p3gdi}).
 A new method was used to find 2D plasma density gradients based on RISR-N measurements and these, along with electric field measurements, are used here to examine directional dependencies.
@@
 \begin{table}
-\caption{A summary of the experimentally observed features that are consistent with the gradient-drift instability theory for arbitrary directions of the wavevector $\uvec{k}$, electric field $\uvec{e}$, and density gradient $\uvec{g}$.}
-\label{tab:results}
 \centering
-\begin{tabular}{|r|l|c|}
-	\hline
+\begin{tabular}{rlc}
+	%\hline
 	\textbf{\#}& \textbf{Experimental Feature} & \textbf{Figure} \\
 	\hline
-	1&Increase in average power of all echoes with \(\uvec{k}\cdot\uvec{b}\times\vec{G}\) & \ref{fig:f5}a \\
-	2&Greater occurrence of structure echoes when \(\uvec{k}\cdot\uvec{b}\times\vec{G}\) is positive & \ref{fig:f5}a, \ref{fig:f7}c \\
-	3&Increase in average power of structure echoes as \(\uvec{k}\cdot\vec{E}\) crosses zero & \ref{fig:f5}b \\
-	4&Greater occurrence of all echoes when \(\uvec{k}\cdot\vec{E}\) is positive & \ref{fig:f5}b, \ref{fig:f7}a \\
-	5&Monotonic increase in occurrence of structure echoes with \(\uvec{k}\cdot\vec{E}\) & \ref{fig:f5}b \\
-	6&Strong peak in occurrence when \(\uvec{k}\cdot\uvec{b}\times\uvec{g} = 1\) & \ref{fig:f5}c \\
-	7&Increase in all echo occurrence leading to a peak at \(\uvec{k}\cdot\uvec{e}=1\) & \ref{fig:f5}d \\
-	8&Greater occurrence of structure echoes when \(\uvec{k}\cdot\uvec{e}\) is positive & \ref{fig:f5}d \\
-	9&Higher average power when \(\vec{G}\cdot\vec{V}_E\) is positive & \ref{fig:f6}a \\
-	10&Strong peak in occurrence when \(\uvec{g}\cdot\mathbf{\hat{v}}_E = 1\) & \ref{fig:f6}c \\
-	11&More high power echoes when \(\uvec{k}\cdot\uvec{b}\times\vec{G}\) and \(\uvec{k}\cdot\vec{E}\) are positive & \ref{fig:f7}a \\
-	12&Greater occurrence of structure echoes when \(\uvec{k}\cdot\uvec{b}\times\vec{G}\) and \(\uvec{k}\cdot\vec{E}\) are positive & \ref{fig:f7}c \\
-	\hline
+	1&Increase in average power of all echoes with \(\uvec{k}\cdot\uvec{b}\times\vec{G}\) & \ref{fig:directional}a \\
+	2&Greater occurrence of structure echoes when \(\uvec{k}\cdot\uvec{b}\times\vec{G}\) is positive & \ref{fig:directional}a, \ref{fig:eg_scatter}c \\
+	3&Increase in average power of structure echoes as \(\uvec{k}\cdot\vec{E}\) crosses zero & \ref{fig:directional}b \\
+	4&Greater occurrence of all echoes when \(\uvec{k}\cdot\vec{E}\) is positive & \ref{fig:directional}b, \ref{fig:eg_scatter}a \\
+	5&Monotonic increase in occurrence of structure echoes with \(\uvec{k}\cdot\vec{E}\) & \ref{fig:directional}b \\
+	6&Strong peak in occurrence when \(\uvec{k}\cdot\uvec{b}\times\uvec{g} = 1\) & \ref{fig:directional}c \\
+	7&Increase in all echo occurrence leading to a peak at \(\uvec{k}\cdot\uvec{e}=1\) & \ref{fig:directional}d \\
+	8&Greater occurrence of structure echoes when \(\uvec{k}\cdot\uvec{e}\) is positive & \ref{fig:directional}d \\
+	9&Higher average power when \(\vec{G}\cdot\vec{V}_E\) is positive & \ref{fig:gamma}a \\
+	10&Strong peak in occurrence when \(\uvec{g}\cdot\mathbf{\hat{v}}_E = 1\) & \ref{fig:gamma}c \\
+	11&More high power echoes when \(\uvec{k}\cdot\uvec{b}\times\vec{G}\) and \(\uvec{k}\cdot\vec{E}\) are positive & \ref{fig:eg_scatter}a \\
+	12&Greater occurrence of structure echoes when \(\uvec{k}\cdot\uvec{b}\times\vec{G}\) and \(\uvec{k}\cdot\vec{E}\) are positive & \ref{fig:eg_scatter}c \\
+	%\hline
 \end{tabular}
+\caption[Results consistent with linear GDI theory]{A summary of the experimentally observed features that are consistent with the gradient-drift instability theory for arbitrary directions of the wavevector \(\uvec{k}\), electric field \(\uvec{e}\), and density gradient \(\uvec{g}\).}
+\label{tab:results}
 \end{table}
 
-Many of the results in Table \ref{tab:results} are based on Figure \ref{fig:f5}, which considers the dependence on the electric field term, \(\uvec{k}\cdot\vec{E}\), and the gradient term, \(\uvec{k}\cdot\uvec{b}\times\vec{G}\), of the growth rate independently.
+Many of the results in Table \ref{tab:results} are based on Figure \ref{fig:directional}, which considers the dependence on the electric field term, \(\uvec{k}\cdot\vec{E}\), and the gradient term, \(\uvec{k}\cdot\uvec{b}\times\vec{G}\), of the growth rate independently.
 Because the growth rate is the product of these two quantities, it can be challenging to interpret these plots, but they still provide useful insight into directional dependencies.
-The two strong peaks at \(\uvec{k}\cdot\uvec{b}\times\uvec{g}=1\) and \(\uvec{k}\cdot\uvec{b}\times\uvec{g}=-1\) in Figure \ref{fig:f5}c indicate that many measurements were made when the gradient was perpendicular to the radar's boresight, \(\uvec{k}\perp\vec{G}\).
-As discussed previously, the peaks in power in Figures \ref{fig:f5}b and \ref{fig:f5}d possibly indicate a preferred electric field direction for the event.
+The two strong peaks at \(\uvec{k}\cdot\uvec{b}\times\uvec{g}=1\) and \(\uvec{k}\cdot\uvec{b}\times\uvec{g}=-1\) in Figure \ref{fig:directional}c indicate that many measurements were made when the gradient was perpendicular to the radar's boresight, \(\uvec{k}\perp\vec{G}\).
+As discussed previously, the peaks in power in Figures \ref{fig:directional}b and \ref{fig:directional}d possibly indicate a preferred electric field direction for the event.
 However, considering all echoes, a full range of electric field directions is observed (points exist for all values of \(\uvec{k}\cdot\uvec{e}\)), but there is a systematic increase in occurrence when \(\uvec{k}\cdot\uvec{e}>1\).
@@
 In relation to result 10, the strong peak in echo occurrence at \(\uvec{g}\cdot\mathbf{\hat{v}}_E = 1\) corresponds to \(\vec{G} \parallel \vec{V}_E\), which is considered the most favorable orientation for GDI growth.
-What was unexpected in the observations presented in Figure \ref{fig:f6}c is the secondary peak at \(\uvec{g}\cdot\mathbf{\hat{v}}_E = -1\), corresponding to \(\vec{G}\) antiparallel to \(\vec{V}_E\), which is the least favorable configuration for GDI growth.
+What was unexpected in the observations presented in Figure \ref{fig:gamma}c is the secondary peak at \(\uvec{g}\cdot\mathbf{\hat{v}}_E = -1\), corresponding to \(\vec{G}\) antiparallel to \(\vec{V}_E\), which is the least favorable configuration for GDI growth.
 It is possible that this secondary peak is due to a different instability mechanism, but the directional dependencies of other mechanisms are also rarely considered \citep{Burston2016}.
@@
 
-One of the interesting features in Figure \ref{fig:f7} is the linear nature of the trend, particularly for the normalized components of high power echoes, Figure \ref{fig:f7}d.
+One of the interesting features in Figure \ref{fig:eg_scatter} is the linear nature of the trend, particularly for the normalized components of high power echoes, Figure \ref{fig:eg_scatter}d.
 GDI linear theory predicts a clustering of points where both \(\uvec{k}\cdot\uvec{b}\times\vec{G}\) and \(\uvec{k}\cdot\vec{E}\) are either positive or negative, or two clusters, one in quadrant I and one in quadrant III.
-Figure \ref{fig:f7}c does show a relatively strong cluster of points in quadrant I but very few points in quadrant III.
-Additionally, in Figure \ref{fig:f7}d, points appear to form a linear pattern, which is confirmed by the relatively high correlation coefficients.
+Figure \ref{fig:eg_scatter}c does show a relatively strong cluster of points in quadrant I but very few points in quadrant III.
+Additionally, in Figure \ref{fig:eg_scatter}d, points appear to form a linear pattern, which is confirmed by the relatively high correlation coefficients.
 A linear relationship between the electric field \(\vec{E}\) and density gradient \(\vec{G}\) is not necessarily expected, but if the majority of echoes are observed when \(\vec{G}\parallel\vec{V}_E\), the linear relationship between the normalized components does make some sense.
@@
 In this study, the plasma drift is assumed to be driven primarily by convection electric fields, with the neutral wind contribution assumed to be negligible.
-In general, this is reasonable because the data collected refers to the \(F\)-region, where neutral wind speed is much smaller than that of the \(\vec{E}\times\vec{B}\) convection, particularly for the present dataset which included many convection velocity measurements exceeding 500 m/s, Figures \ref{fig:f2}c and \ref{fig:f3}.
+In general, this is reasonable because the data collected refers to the \(F\)-region, where neutral wind speed is much smaller than that of the \(\vec{E}\times\vec{B}\) convection, particularly for the present dataset which included many convection velocity measurements exceeding 500 m/s, Figures \ref{fig:event_overview}c and \ref{fig:velocity}.
 If, however, neutral winds were to contribute significantly at these altitudes, the effective convection electric field would be reduced and a corresponding reduction in the calculated growth rates would occur, so the growth rates presented here would be overestimated.
@@
 \subsection{Differences Between Echo Power and Occurrence}
-\label{sec:discussion3}
+\label{sec:p3discussion3}
 This study considered both echo power and occurrence, and it has been found that they do not necessarily follow the same trends.
-This can be seen, for example, in Figure \ref{fig:f5}a, where the average power for all echoes gradually increases but occurrence peaks at zero, or in Figure \ref{fig:f5}b, where the average power for structure echoes peaks around 0.01 mV/m but occurrence increases monotonically to 0.025 mV/m.
+This can be seen, for example, in Figure \ref{fig:directional}a, where the average power for all echoes gradually increases but occurrence peaks at zero, or in Figure \ref{fig:directional}b, where the average power for structure echoes peaks around 0.01 mV/m but occurrence increases monotonically to 0.025 mV/m.
 In general, the echo occurrence seems to agree with GDI linear theory somewhat better than average power, Table \ref{tab:results}.
-This is evident in Figures \ref{fig:f6}b and \ref{fig:f6}d, in which there is slightly greater occurrence for positive growth rates than negative growth rates, and structure echoes in particular are biased towards positive growth rates.
+This is evident in Figures \ref{fig:gamma}b and \ref{fig:gamma}d, in which there is slightly greater occurrence for positive growth rates than negative growth rates, and structure echoes in particular are biased towards positive growth rates.
 On the other hand, average power seems to either exhibit very little trend for positive growth rates or even decreases in the normalized case.
@@
 
-Structure echoes do show a steady increase in average power while the growth rate is negative, Figure \ref{fig:f6}b, for both the non-normalized (between \(-0.001\) s\(^{-1}\) and 0  s\(^{-1}\)) and normalized cases (between \(-0.3\) and 0.1).
+Structure echoes do show a steady increase in average power while the growth rate is negative, Figure \ref{fig:gamma}b, for both the non-normalized (between \(-0.001\) s\(^{-1}\) and 0  s\(^{-1}\)) and normalized cases (between \(-0.3\) and 0.1).
 While this agrees well with the notion that increasing the growth rate should increase the strength of irregularities in the ionosphere, conventional understanding of a linear growth is that if the growth rate is negative, waves are damped, so there should be no structuring and no backscatter observed at all.
@@
 
-With respect to the estimates of the growth rates, the highest rates were around 0.002 s\(^{-1}\), Figure \ref{fig:f6}c, which corresponds to a rise time of \(\sim8\) min.
+With respect to the estimates of the growth rates, the highest rates were around 0.002 s\(^{-1}\), Figure \ref{fig:gamma}c, which corresponds to a rise time of \(\sim8\) min.
 However, the peak in echo occurrence occurred at \(\gamma\approx0.0001\) s\(^{-1}\), corresponding to a rise time of over 2.5 hours.
@@
 \section{Conclusions}
+\label{sec:p3conclusions}
 Signatures of plasma structuring in the polar \(F\)-region ionosphere have been investigated in the context of factors that control directionally-dependent growth rate of the gradient-drift instability.
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-%  ACKNOWLEDGMENTS
-%
-% The acknowledgments must list:
-%
-% •	All funding sources related to this work from all authors
-%
-% •	Any real or perceived financial conflicts of interests for any
-%	author
-%
-% •	Other affiliations for any author that may be perceived as
-% 	having a conflict of interest with respect to the results of this
-% 	paper.
-%
-% •	A statement that indicates to the reader where the data
-% 	supporting the conclusions can be obtained (for example, in the
-% 	references, tables, supporting information, and other databases).
-%
-% It is also the appropriate place to thank colleagues and other contributors. 
-% AGU does not normally allow dedications.
-
 \section*{Acknowledgments}
@@
 
-%%  REFERENCE LIST AND TEXT CITATIONS
-%
-% Either type in your references using
-%
-% \begin{thebibliography}{}
-% \bibitem[{\textit{Kobayashi et~al.}}(2003)]{R2013} Kobayashi, T.,
-% Tran, A.~H., Nishijo, H., Ono, T., and Matsumoto, G.  (2003).
-% Contribution of hippocampal place cell activity to learning and
-% formation of goal-directed navigation in rats. \textit{Neuroscience}
-% 117, 1025--1035.
-%
-% \bibitem{}
-% Text
-% \end{thebibliography}
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Or, to use BibTeX:
-%
-% Follow these steps
-%
-% 1. Type in \bibliography{<name of your .bib file>} 
-%    Run LaTeX on your LaTeX file.
-%
-% 2. Run BiBTeX on your LaTeX file.
-%
-% 3. Open the new .bbl file containing the reference list and
-%   copy all the contents into your LaTeX file here.
-%
-% 4. Run LaTeX on your new file which will produce the citations.
-%
-% AGU does not want a .bib or a .bbl file. Please copy in the contents of your .bbl file here.
-
-%\bibliography{myrefll}
-%
-%\providecommand{\ddegrees}{\ensuremath{\deg}}
-%\begin{thebibliography}{46}
-%\providecommand{\natexlab}[1]{#1}
-%\expandafter\ifx\csname urlstyle\endcsname\relax
-%  \providecommand{\doi}[1]{doi:\discretionary{}{}{}#1}\else
-%  \providecommand{\doi}{doi:\discretionary{}{}{}\begingroup
-%  \urlstyle{rm}\Url}\fi
-%
-%\bibitem[{\textit{Bahcivan et~al.}(2010)\textit{Bahcivan, Tsunoda, Nicolls, and
-%  Heinselman}}]{Bah10}
-%Bahcivan, H., R.~Tsunoda, M.~Nicolls, and C.~Heinselman (2010), Initial
-%  ionospheric observations made by the new {Resolute} incoherent scatter radar
-%  and comparison to solar wind {IMF}, \textit{Geophys. Res. Lett.},
-%  \textit{37}, L15103, \doi{10.1029/2010GL043632}.
-%
-%\bibitem[{\textit{Bristow et~al.}(2011)\textit{Bristow, Spaleta, and
-%  Parris}}]{Bri11}
-%Bristow, W.~A., J.~Spaleta, and R.~T. Parris (2011), {First observations of
-%  ionospheric irregularities and flows over the south geomagnetic pole from the
-%  Super Dual Auroral Radar Network (SuperDARN) HF radar at McMurdo Station,
-%  Antarctica}, \textit{J. Geophys. Res.}, \textit{116}, A12325,
-%  \doi{10.1029/2011JA016834}.
-%
-%\bibitem[{\textit{Burston et~al.}(2010)\textit{Burston, Astin, Mitchell,
-%  Alfonsi, Pedersen, and Skone}}]{Bur10}
-%Burston, R., I.~Astin, C.~Mitchell, L.~Alfonsi, T.~Pedersen, and S.~Skone
-%  (2010), Turbulent times in the northern polar ionosphere?, \textit{J.
-%  Geophys. Res.}, \textit{115}, A04310, \doi{10.1029/2009JA014813}.
-%
-%\bibitem[{\textit{Burston et~al.}(2016)\textit{Burston, Mitchell, and
-%  Astin}}]{Bur16}
-%Burston, R., C.~Mitchell, and I.~Astin (2016), Polar cap plasma patch primary
-%  linear instability growth rates compared, \textit{J. Geophys. Res. Space
-%  Physics}, \textit{121}, 3439--3451, \doi{10.1002/2015JA021895}.
-%
-%\bibitem[{\textit{Carlson et~al.}(2007)\textit{Carlson, Pedersen, Basu,
-%  Keskinen, and Moen}}]{Car07}
-%Carlson, H.~C., T.~Pedersen, S.~Basu, M.~Keskinen, and J.~Moen (2007), Case for
-%  a new process, not mechanism, for cusp irregularity production, \textit{J.
-%  Geophys. Res.}, \textit{112}, A11304, \doi{10.1029/2007JA012384}.
-%
-%\bibitem[{\textit{Carlson et~al.}(2008)\textit{Carlson, Oksavik, and
-%  Moen}}]{Car08}
-%Carlson, H.~C., K.~Oksavik, and J.~Moen (2008), On a new process for cusp
-%  irregularity production, \textit{Ann. Geophysicae}, \textit{26}, 2871--2885.
-%
-%\bibitem[{\textit{Chisham et~al.}(2007)\textit{Chisham, Lester, Milan, Freeman,
-%  Bristow, Grocott, McWilliams, Ruohoniemi, Yeoman, Dyson, Greenwald, Kikuchi,
-%  Pinnock, Rash, Sato, Sofko, Villain, and Walker}}]{Chi07}
-%Chisham, G., M.~Lester, S.~E. Milan, M.~P. Freeman, W.~A. Bristow, A.~Grocott,
-%  K.~A. McWilliams, J.~M. Ruohoniemi, T.~Yeoman, P.~L. Dyson, R.~A. Greenwald,
-%  T.~Kikuchi, M.~Pinnock, J.~P.~S. Rash, N.~Sato, G.~J. Sofko, J.-P. Villain,
-%  and A.~D.~M. Walker (2007), A decade of the {Super Dual Auroral Radar Network
-%  (SuperDARN)}: scientific achievements, new techniques and future directions,
-%  \textit{Surveys in Geophysics}, \textit{28}, 33--109,
-%  \doi{10.1007/s10712-007-9017-8}.
-%
-%\bibitem[{\textit{Dahlgren et~al.}(2012{\natexlab{a}})\textit{Dahlgren, Perry,
-%  Semeter, St.-Maurice, Hosokawa, Nicolls, Greffen, Shiokawa, and
-%  Heinselman}}]{Dah12b}
-%Dahlgren, H., G.~W. Perry, J.~L. Semeter, J.-P. St.-Maurice, K.~Hosokawa, M.~J.
-%  Nicolls, M.~Greffen, K.~Shiokawa, and C.~Heinselman (2012{\natexlab{a}}),
-%  {Space-time variability of polar cap patches: Direct evidence for internal
-%  plasma structuring}, \textit{J. Geophys. Res.}, \textit{117}, A09312,
-%  \doi{10.1029/2012JA017961}.
-%
-%\bibitem[{\textit{Dahlgren et~al.}(2012{\natexlab{b}})\textit{Dahlgren,
-%  Semeter, Hosokawa, Nicolls, Butler, Johnsen, Shiokawa, and
-%  Heinselman}}]{Dah12a}
-%Dahlgren, H., J.~L. Semeter, K.~Hosokawa, M.~J. Nicolls, T.~W. Butler, M.~G.
-%  Johnsen, K.~Shiokawa, and Heinselman (2012{\natexlab{b}}), Direct
-%  three-dimensional imaging of polar ionospheric strucutes with the {Resolute
-%  Bay Incoherent Scatter Radar}, \textit{Geophys. Res. Lett.}, \textit{39},
-%  L05104, \doi{10.1029/2012GL050895}.
-%
-%\bibitem[{\textit{Danskin et~al.}(2002)\textit{Danskin, Koustov, Ogawa,
-%  Nishitani, Nozawa, Milan, Lester, and Andre}}]{Dan02}
-%Danskin, D.~W., A.~V. Koustov, T.~Ogawa, N.~Nishitani, S.~Nozawa, S.~E. Milan,
-%  M.~Lester, and D.~Andre (2002), On the factors controlling occurrence of
-%  {F}-region coherent echoes, \textit{Ann. Geophysicae}, \textit{20}(9),
-%  1385--1397.
-%
-%\bibitem[{\textit{Evans}(1969)}]{Eva69}
-%Evans, J.~V. (1969), Theory and practice of ionosphere study by {Thomson
-%  Scatter Radar}, \textit{Proc. {IEEE}}, \textit{57}, 496--530.
-%
-%\bibitem[{\textit{Fejer and Kelley}(1980)}]{Fej80}
-%Fejer, B.~G., and M.~C. Kelley (1980), Ionospheric irregularities,
-%  \textit{Geophys. Rev.}, \textit{18}, 401--454.
-%
-%\bibitem[{\textit{{Fukumoto} et~al.}(1999)\textit{{Fukumoto}, {Nishitani},
-%  {Ogawa}, {Sato}, {Yamagishi}, and {Yukimatu}}}]{Fuk99}
-%{Fukumoto}, M., N.~{Nishitani}, T.~{Ogawa}, N.~{Sato}, H.~{Yamagishi}, and
-%  A.~{Yukimatu} (1999), Statistical analysis of echo power, {Doppler} velocity
-%  and spectral width obtained with the {Syowa South HF radar}, \textit{Adv.
-%  Polar Upper Atmos. Res.}, \textit{13}, 37--47.
-%
-%\bibitem[{\textit{{Fukumoto} et~al.}(2000)\textit{{Fukumoto}, {Nishitani},
-%  {Ogawa}, {Sato}, {Yamagishi}, and {Yukimatu}}}]{Fuk00}
-%{Fukumoto}, M., N.~{Nishitani}, T.~{Ogawa}, N.~{Sato}, H.~{Yamagishi}, and
-%  A.~S. {Yukimatu} (2000), Statistical study of {Doppler} velocity and echo
-%  power around 75\(\deg\) magnetic latitude using data obtained with the
-%  {Syowa} {East} {HF} radar in 1997, \textit{Adv. Polar Upper Atmos. Res.},
-%  \textit{14}, 93--102.
-%
-%\bibitem[{\textit{Gondarenko and Guzdar}(2004)}]{Gon04}
-%Gondarenko, N.~A., and P.~N. Guzdar (2004), Plasma patch structuring by the
-%  nonlinear evolution of the gradient drift instability in the high-latitude
-%  ionosphere, \textit{J. Geophys. Res.}, \textit{109}, A09301,
-%  \doi{10.1029/2004JA010504}.
-%
-%\bibitem[{\textit{Gondarenko and Guzdar}(2006)}]{Gon06}
-%Gondarenko, N.~A., and P.~N. Guzdar (2006), Nonlinear three-dimensional
-%  simulations of mesoscale structuring by multiple drives in high-latitude
-%  plasma patches, \textit{J. Geophys. Res.}, \textit{111}, A08302,
-%  \doi{10.1029/2006JA011701}.
-%
-%\bibitem[{\textit{Haldoupis et~al.}(1990)\textit{Haldoupis, Nielsen, and
-%  Schlegel}}]{Hal90}
-%Haldoupis, C., E.~Nielsen, and K.~Schlegel (1990), Dependence of radar auroral
-%  scattering cross section on the ambient electron density and the
-%  destabilizing electric field, \textit{Ann. Geophysicae}, \textit{8},
-%  195--212.
-%
-%\bibitem[{\textit{{Heinselman} and {Nicolls}}(2008)}]{Hei08}
-%{Heinselman}, C.~J., and M.~J. {Nicolls} (2008), {A Bayesian approach to
-%  electric field and E-region neutral wind estimation with the Poker Flat
-%  Advanced Modular Incoherent Scatter Radar}, \textit{Radio Sci.}, \textit{43},
-%  RS5013, \doi{10.1029/2007RS003805}.
-%
-%\bibitem[{\textit{Kane et~al.}(2012)\textit{Kane, Makarevich, and
-%  Devlin}}]{Kan12}
-%Kane, T.~A., R.~A. Makarevich, and J.~C. Devlin (2012), {{HF} radar
-%  observations of ionospheric backscatter during geomagnetically quiet
-%  periods}, \textit{Ann. Geophysicae}, \textit{30}, 221--233,
-%  \doi{10.5194/angeo-30-221-2012}.
-%
-%\bibitem[{\textit{{Keskinen} and {Ossakow}}(1982)}]{Kes82a}
-%{Keskinen}, M.~J., and S.~L. {Ossakow} (1982), {Nonlinear evolution of plasma
-%  enhancements in the auroral ionosphere. I - Long wavelength irregularities},
-%  \textit{J. Geophys. Res.}, \textit{87}, 144--150.
-%
-%\bibitem[{\textit{{Keskinen} and {Ossakow}}(1983)}]{Kes83}
-%{Keskinen}, M.~J., and S.~L. {Ossakow} (1983), Theories of high-latitude
-%  ionospheric irregularities: A review, \textit{Radio Sci.}, \textit{18}(6),
-%  1077--1091.
-%
-%\bibitem[{\textit{{Koustov} et~al.}(2004)\textit{{Koustov}, {Sofko},
-%  {Andr{\'e}}, {Danskin}, and {Benkevitch}}}]{Kou04}
-%{Koustov}, A.~V., G.~J. {Sofko}, D.~{Andr{\'e}}, D.~W. {Danskin}, and L.~V.
-%  {Benkevitch} (2004), {Seasonal variation of HF radar F region echo occurrence
-%  in the midnight sector}, \textit{J. Geophys. Res.}, \textit{109}, A06305,
-%  \doi{10.1029/2003JA010337}.
-%
-%\bibitem[{\textit{{Koustov} et~al.}(2012)\textit{{Koustov}, {Hosokawa},
-%  {Nishitani}, {Shiokawa}, and {Liu}}}]{Kou12}
-%{Koustov}, A.~V., K.~{Hosokawa}, N.~{Nishitani}, K.~{Shiokawa}, and H.~{Liu}
-%  (2012), {Signatures of moving polar cap arcs in the F-region PolarDARN
-%  echoes}, \textit{Ann. Geophysicae}, \textit{30}, 441--455.
-%
-%\bibitem[{\textit{{Koustov} et~al.}(2014)\textit{{Koustov}, {Ponomarenko},
-%  {Ghezelbash}, {Themens}, and {Jayachandran}}}]{Kou14}
-%{Koustov}, A.~V., P.~V. {Ponomarenko}, M.~{Ghezelbash}, D.~R. {Themens}, and
-%  P.~T. {Jayachandran} (2014), {Electron density and electric field over
-%  Resolute Bay and F region ionospheric echo detection with the Rankin Inlet
-%  and Inuvik SuperDARN radars}, \textit{Radio Sci.}, pp. 1194--1205,
-%  \doi{10.1002/2014RS005579}.
-%
-%\bibitem[{\textit{Kustov et~al.}(1988)\textit{Kustov, Uspenskii, and
-%  Williams}}]{Kus88}
-%Kustov, A.~V., M.~V. Uspenskii, and F.~J.~S. Williams (1988), Saturation
-%  effects of auroral-electrojet turbulence, \textit{Geom. Aeron.}, \textit{28},
-%  923--927.
-%
-%\bibitem[{\textit{Lamarche and Makarevich}(2015)}]{Lam15}
-%Lamarche, L.~J., and R.~A. Makarevich (2015), {Solar control of F region radar
-%  backscatter: Further insights from observations in the southern polar cap},
-%  \textit{J. Geophys. Res. Space Physics}, \textit{120}(11), 9875--9890,
-%  \doi{10.1002/2015JA021663}.
-%
-%\bibitem[{\textit{Lamarche and Makarevich}(2016)}]{Lam16}
-%Lamarche, L.~J., and R.~A. Makarevich (2016), A modeling study of asymmetries
-%  in plasma irregularity characteristics near gradient reversals, \textit{Ann.
-%  Geophysicae}, \textit{34}, 709--723, \doi{10.5194/angeo-34-709-2016}.
-%
-%\bibitem[{\textit{{Linson} and {Workman}}(1970)}]{Lin70}
-%{Linson}, L.~M., and J.~B. {Workman} (1970), Formation of striations in
-%  ionospheric plasma clouds, \textit{J. Geophys. Res.}, \textit{75}, 3211.
-%
-%\bibitem[{\textit{Lynch et~al.}(2015)\textit{Lynch, Hampton, Zettergren,
-%  Bekkeng, Conde, Fernandes, Horak, Lessard, Miceli, Michell, Moen, Nicolls,
-%  Powell, and Samara}}]{Lyn15}
-%Lynch, K.~A., D.~K. Hampton, M.~Zettergren, T.~A. Bekkeng, M.~Conde, P.~A.
-%  Fernandes, P.~Horak, M.~Lessard, R.~Miceli, R.~Michell, J.~Moen, M.~Nicolls,
-%  S.~P. Powell, and M.~Samara (2015), {MICA} sounding rocket observations of
-%  the conductivity-gradient-generated auroral ionospheric responses:
-%  Small-scale structures with large-scale drivers, \textit{J. Geophys. Res.
-%  Space Physics}, \textit{120}, 9661--9682, \doi{10.1002/2014JA020860}.
-%
-%\bibitem[{\textit{Makarevich}(2014)}]{Mak14c}
-%Makarevich, R.~A. (2014), {Symmetry considerations in the two-fluid theory of
-%  the gradient-drift instability in the lower ionosphere}, \textit{J. Geophys.
-%  Res.}, \textit{119}, \doi{10.1002/2014JA020292}.
-%
-%\bibitem[{\textit{Makarevich and Bristow}(2014)}]{Mak14b}
-%Makarevich, R.~A., and W.~A. Bristow (2014), Coordinated radar observations of
-%  plasma wave characteristics in the auroral {F} region, \textit{Ann.
-%  Geophysicae}, \textit{32}, 875--888.
-%
-%\bibitem[{\textit{Makarevich et~al.}(2015)\textit{Makarevich, Lamarche, and
-%  Nicolls}}]{Mak15b}
-%Makarevich, R.~A., L.~J. Lamarche, and M.~J. Nicolls (2015), {Resolute Bay
-%  Incoherent Scatter Radar} observations of plasma structures in the vicinity
-%  of polar holes, \textit{J. Geophys. Res. Space Physics}, \textit{120},
-%  \doi{10.1002/2015JA021443}.
-%
-%\bibitem[{\textit{Makarevitch et~al.}(2004)\textit{Makarevitch, Honary, McCrea,
-%  and Howells}}]{Mak04c}
-%Makarevitch, R.~A., F.~Honary, I.~W. McCrea, and V.~S.~C. Howells (2004),
-%  Imaging riometer observations of drifting absorption patches in the morning
-%  sector, \textit{Ann. Geophysicae}, \textit{22}, 3461--3478.
-%
-%\bibitem[{\textit{Milan et~al.}(2002)\textit{Milan, Lester, and
-%  Yeoman}}]{Mil02b}
-%Milan, S.~E., M.~Lester, and T.~K. Yeoman (2002), {HF} radar polar patch
-%  formation revisited: summer and winter variations in dayside plasma
-%  structuring, \textit{Ann. Geophysicae}, \textit{20}, 487--499.
-%
-%\bibitem[{\textit{Moen et~al.}(2012)\textit{Moen, Oksavik, Abe, Lester, Saito,
-%  Bekkeng, and Jacobsen}}]{Moe12}
-%Moen, J., K.~Oksavik, T.~Abe, M.~Lester, Y.~Saito, T.~A. Bekkeng, and K.~S.
-%  Jacobsen (2012), {First in situ measurements of HF radar echoing targets},
-%  \textit{Geophys. Res. Lett.}, \textit{39}, L07104,
-%  \doi{10.1029/2012GL051407}.
-%
-%\bibitem[{\textit{Nicolls et~al.}(2007)\textit{Nicolls, Heinselman, Hope,
-%  Ranjan, Kelley, and Kelly}}]{Nic07}
-%Nicolls, M.~J., C.~J. Heinselman, E.~A. Hope, S.~Ranjan, M.~C. Kelley, and
-%  J.~D. Kelly (2007), Imaging of polar mesosphere summer echoes with the 450
-%  {MHz} {Poker Flat Advanced Modular Incoherent Scatter Radar},
-%  \textit{Geophys. Res. Lett.}, \textit{34}, L20102,
-%  \doi{10.1029/2007GL031476}.
-%
-%\bibitem[{\textit{Oksavik et~al.}(2010)\textit{Oksavik, Barth, Moen, and
-%  Lester}}]{Oks10}
-%Oksavik, K., V.~L. Barth, J.~Moen, and M.~Lester (2010), On the entry and
-%  transit of high-density plasma across the polar cap, \textit{J. Geophys.
-%  Res.}, \textit{115}, \doi{10.1029/2010JA015817}.
-%
-%\bibitem[{\textit{Rishbeth and Williams}(1985)}]{Ris85}
-%Rishbeth, H., and P.~J.~S. Williams (1985), The {EISCAT} ionospheric radar: The
-%  system and its early results, \textit{Q. J. R. Astron. Soc.}, \textit{26},
-%  478--512.
-%
-%\bibitem[{\textit{Ruohoniemi and Greenwald}(1997)}]{Ruo97}
-%Ruohoniemi, J.~M., and R.~A. Greenwald (1997), {Rates of scattering occurrence
-%  in routine HF radar observations during solar cycle maximum}, \textit{Radio
-%  Sci.}, \textit{32}, 1051--1070.
-%
-%\bibitem[{\textit{{Semeter} et~al.}(2009)\textit{{Semeter}, {Butler},
-%  {Heinselman}, {Nicolls}, {Kelly}, and {Hampton}}}]{Sem09}
-%{Semeter}, J., T.~{Butler}, C.~{Heinselman}, M.~{Nicolls}, J.~{Kelly}, and
-%  D.~{Hampton} (2009), {Volumetric imaging of the auroral ionosphere: Initial
-%  results from PFISR}, \textit{J. Atmos. Sol. Terr. Phys.}, \textit{71},
-%  738--743.
-%
-%\bibitem[{\textit{Shepherd}(2014)}]{She14}
-%Shepherd, S.~G. (2014), Altitude-adjusted corrected geomagnetic coordinates:
-%  Definition and functional approximations, \textit{J. Geophys. Res.},
-%  \textit{119}(9), 7501--7521, \doi{10.1002/2014JA020264}.
-%
-%\bibitem[{\textit{Spicher et~al.}(2015)\textit{Spicher, Miloch, Clausen, and
-%  Moen}}]{Spi15}
-%Spicher, A., W.~J. Miloch, L.~B.~N. Clausen, and J.~I. Moen (2015), Plasma
-%  turbulence and coherent structures in the polar cap observed by the {ICI}-2
-%  sounding rocket, \textit{J. Geophys. Res. Space Physics}, \textit{120},
-%  10,959--10,978, \doi{10.1002/2015JA021634}.
-%
-%\bibitem[{\textit{Themens et~al.}(2014)\textit{Themens, Jayachandran, Nicolls,
-%  and MacDougall}}]{The14a}
-%Themens, D., P.~Jayachandran, M.~J. Nicolls, and J.~W. MacDougall (2014), {A
-%  Top to bottom evaluation of IRI-2007 within the polar cap}, \textit{J.
-%  Geophys. Res.}, \textit{119}, \doi{10.1002/2014JA020052}.
-%
-%\bibitem[{\textit{{Tsunoda}}(1988)}]{Tsu88}
-%{Tsunoda}, R.~T. (1988), {High-latitude F region irregularities - A review and
-%  synthesis}, \textit{Geophys. Rev.}, \textit{26}, 719--760.
-%
-%\bibitem[{\textit{{Villain} et~al.}(1986)\textit{{Villain}, {Beghin}, and
-%  {Hanuise}}}]{Vil86}
-%{Villain}, J.~P., C.~{Beghin}, and C.~{Hanuise} (1986), {ARCAD3-SAFARI}
-%  coordinated study of auroral and polar {F}-region ionospheric irregularities,
-%  \textit{Ann. Geophysicae}, \textit{4}, 61--68.
-%
-%\bibitem[{\textit{{Weber} et~al.}(1984)\textit{{Weber}, {Buchau}, {Moore},
-%  {Sharber}, {Livingston}, {Winningham}, and {Reinisch}}}]{Web84}
-%{Weber}, E.~J., J.~{Buchau}, J.~G. {Moore}, J.~R. {Sharber}, R.~C.
-%  {Livingston}, J.~D. {Winningham}, and B.~W. {Reinisch} (1984), {F layer
-%  ionization patches in the polar cap}, \textit{J. Geophys. Res.}, \textit{89},
-%  1683--1694.
-%
-%\end{thebibliography}
-
-%% After you run BibTeX, Copy in the contents of the .bbl file here:
-
-%Text here ===>>>
-
-%%
-
-%  Numbered lines in equations:
-%  To add line numbers to lines in equations,
-%  \begin{linenomath*}
-%  \begin{equation}
-%  \end{equation}
-%  \end{linenomath*}
-
-%
-% ---------------
-% EXAMPLE TABLE
-%
-% \begin{table}
-% \caption{Time of the Transition Between Phase 1 and Phase 2$^{a}$}
-% \centering
-% \begin{tabular}{l c}
-% \hline
-%  Run  & Time (min)  \\
-% \hline
-%   $l1$  & 260   \\
-%   $l2$  & 300   \\
-%   $l3$  & 340   \\
-%   $h1$  & 270   \\
-%   $h2$  & 250   \\
-%   $h3$  & 380   \\
-%   $r1$  & 370   \\
-%   $r2$  & 390   \\
-% \hline
-% \multicolumn{2}{l}{$^{a}$Footnote text here.}
-% \end{tabular}
-% \end{table}
-
-%% Enter Figures and Tables near as possible to where they are first mentioned:
-%
-% DO NOT USE \psfrag or \subfigure commands.
-%
-% Figure captions go below the figure.
-% Table titles go above tables;  other caption information
-%  should be placed in last line of the table, using
-% \multicolumn2l{$^a$ This is a table note.}
-%
-%----------------
-% EXAMPLE FIGURE
-%
-% \begin{figure}[h]
-% \centering
-% when using pdflatex, use pdf file:
-% \includegraphics[width=20pc]{figsamp.pdf}
-%
-% when using dvips, use .eps file:
-% \includegraphics[width=20pc]{figsamp.eps}
-%
-% \caption{Short caption}
-% \label{figone}
-%  \end{figure}
-
-%% SIDEWAYS FIGURE and TABLE 
-% AGU prefers the use of {sidewaystable} over {landscapetable} as it causes fewer problems.
-%
-% \begin{sidewaysfigure}
-% \includegraphics[width=20pc]{figsamp}
-% \caption{caption here}
-% \label{newfig}
-% \end{sidewaysfigure}
-% 
-%  \begin{sidewaystable}
-%  \caption{Caption here}
-% \label{tab:signif_gap_clos}
-%  \begin{tabular}{ccc}
-% one&two&three\\
-% four&five&six
-%  \end{tabular}
-%  \end{sidewaystable}
-
-%% If using numbered lines, please surround equations with \begin{linenomath*}...\end{linenomath*}
-%\begin{linenomath*}
-%\begin{equation}
-%y|{f} \sim g(m, \sigma),
-%\end{equation}
-%\end{linenomath*}
-
-%%% End of body of article
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Optional Appendix goes here
-%
-% The \appendix command resets counters and redefines section heads
-%
-% After typing \appendix
-%
-%\section{Here Is Appendix Title}
-% will show
-% A: Here Is Appendix Title
-%
-%\appendix
-%\section{Here is a sample appendix}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-% Optional Glossary, Notation or Acronym section goes here:
-%
-%%%%%%%%%%%%%%  
-% Glossary is only allowed in Reviews of Geophysics
-%  \begin{glossary}
-%  \term{Term}
-%   Term Definition here
-%  \term{Term}
-%   Term Definition here
-%  \term{Term}
-%   Term Definition here
-%  \end{glossary}
-
-%
-%%%%%%%%%%%%%%
-% Acronyms
-%   \begin{acronyms}
-%   \acro{Acronym}
-%   Definition here
-%   \acro{EMOS}
-%   Ensemble model output statistics 
-%   \acro{ECMWF}
-%   Centre for Medium-Range Weather Forecasts
-%   \end{acronyms}
-
-%
-%%%%%%%%%%%%%%
-% Notation 
-%   \begin{notation}
-%   \notation{$a+b$} Notation Definition here
-%   \notation{$e=mc^2$} 
-%   Equation in German-born physicist Albert Einstein's theory of special
-%  relativity that showed that the increased relativistic mass ($m$) of a
-%  body comes from the energy of motion of the body—that is, its kinetic
-%  energy ($E$)—divided by the speed of light squared ($c^2$).
-%   \end{notation}
-
-%% ------------------------------------------------------------------------ %%
-%% Citations
-
-% Please use ONLY \citet and \citep for reference citations.
-% DO NOT use other cite commands (e.g., \cite, \citeyear, \nocite, \citealp, etc.).
-
-%% Example \citet and \citep:
-%  ...as shown by \citet{Boug10}, \citet{Buiz07}, \citet{Fra10},
-%  \citet{Ghel00}, and \citet{Leit74}.
-
-%  ...as shown by \citep{Boug10}, \citep{Buiz07}, \citep{Fra10},
-%  \citep{Ghel00, Leit74}.
-
-%  ...has been shown \citep [e.g.,][]{Boug10,Buiz07,Fra10}.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Track Changes:
-% To add words, \added{<word added>}
-% To delete words, \deleted{<word deleted>}
-% To replace words, \replace{<word to be replaced>}{<replacement word>}
-% To explain why change was made: \explain{<explanation>} This will put
-% a comment into the right margin.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% At the end of the document, use \listofchanges, which will list the
-% changes and the page and line number where the change was made.
-
-% When final version, \listofchanges will not produce anything,
-% \added{<word or words>} word will be printed, \deleted{<word or words} will take away the word,
-% \replaced{<delete this word>}{<replace with this word>} will print only the replacement word.
-%  In the final version, \explain will not print anything.
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%
-%\listofchanges
-%%%
-
-%\end{document}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Supporting Information
-%% (Optional) See AGUSuppInfoSamp.tex/pdf for requirements 
-%% for Supporting Information.
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-%More Information and Advice:
-%
-%------------------------------------------------------------------------ %%
-%
-% SECTION HEADS
-%
-%------------------------------------------------------------------------ %%
-%
-%Capitalize the first letter of each word (except for
-%prepositions, conjunctions, and articles that are
-%three or fewer letters).
-%
-%AGU follows standard outline style; therefore, there cannot be a section 1 without
-%a section 2, or a section 2.3.1 without a section 2.3.2.
-%Please make sure your section numbers are balanced.
-%---------------
-%Level 1 head
-%
-%Use the \section{} command to identify level 1 heads;
-%type the appropriate head wording between the curly
-%brackets, as shown below.
-%
-%An example:
-%\section{Level 1 Head: Introduction}
-%
-%---------------
-%Level 2 head
-%
-%Use the \subsection{} command to identify level 2 heads.
-%An example:
-%\subsection{Level 2 Head}
-%
-%---------------
-%Level 3 head
-%
-%Use the \subsubsection{} command to identify level 3 heads
-%An example:
-%\subsubsection{Level 3 Head}
-%
-%---------------
-%Level 4 head
-%
-%Use the \subsubsubsection{} command to identify level 3 heads
-%An example:
-%\subsubsubsection{Level 4 Head} An example.
-%
-%------------------------------------------------------------------------ %%
-%
-% IN-TEXT LISTS
-%
-%------------------------------------------------------------------------ %%
-%
-%Do not use bulleted lists; enumerated lists are okay.
-%\begin{enumerate}
-%\item
-%\item
-%\item
-%\end{enumerate}
-%
-%------------------------------------------------------------------------ %%
-%
-% EQUATIONS
-%
-%------------------------------------------------------------------------ %%
-%
-%Single-line equations are centered.
-%Equation arrays will appear left-aligned.
-%
-%Math coded inside display math mode \[ ...\]
-% will not be numbered, e.g.,:
-% \[ x^2=y^2 + z^2\]
-%
-% Math coded inside \begin{equation} and \end{equation} will
-% be automatically numbered, e.g.,:
-% \begin{equation}
-% x^2=y^2 + z^2
-% \end{equation}
-%
-%
-%To create multiline equations, use the
-%\begin{eqnarray} and \end{eqnarray} environment
-%as demonstrated below.
-%\begin{eqnarray}
-%  x_{1} & = & (x - x_{0}) \cos \Theta \nonumber \\
-%        && + (y - y_{0}) \sin \Theta  \nonumber \\
-%  y_{1} & = & -(x - x_{0}) \sin \Theta \nonumber \\
-%        && + (y - y_{0}) \cos \Theta.
-%\end{eqnarray}
-%
-%If you don't want an equation number, use the star form:
-%\begin{eqnarray*}...\end{eqnarray*}
-%
-%Break each line at a sign of operation
-%(+, -, etc.) if possible, with the sign of operation
-%on the new line.
-%
-%Indent second and subsequent lines to align with
-%the first character following the equal sign on the
-%first line.
-%
-%Use an \hspace{} command to insert horizontal space
-%into your equation if necessary. Place an appropriate
-%unit of measure between the curly braces, e.g.
-%\hspace{1in}; you may have to experiment to achieve
-%the correct amount of space.
-%
-%
-%------------------------------------------------------------------------ %%
-%
-% EQUATION NUMBERING: COUNTER
-%
-%------------------------------------------------------------------------ %%
-%
-%You may change equation numbering by resetting
-%the equation counter or by explicitly numbering
-%an equation.
-%
-%To explicitly number an equation, type \eqnum{}
-%(with the desired number between the brackets)
-%after the \begin{equation} or \begin{eqnarray}
-%command.  The \eqnum{} command will affect only
-%the equation it appears with; LaTeX will number
-%any equations appearing later in the manuscript
-%according to the equation counter.
-%
-%
-%If you have a multiline equation that needs only
-%one equation number, use a \nonumber command in
-%front of the double backslashes (\\) as shown in
-%the multiline equation above.
-%
-%If you are using line numbers, remember to surround
-%equations with \begin{linenomath*}...\end{linenomath*}
-%
-% To add line numbers to lines in equations:
-% \begin{linenomath*}
-% \begin{equation}
-% \end{equation}
-% \end{linenomath*}
-%
-%
-%
+\bibliographystyle{uafthesis}
+\bibliography{references}

### 2017-02-03 acknowledgements.tex
Acknowledgements revised

@@
 First and foremost, I would like to thank my advisor, Dr.
-Roman Makarevich for his guidance throughout this project, which would not have been possible without his support, feedback, and constructive criticism.
+Roman Makarevich for his guidance throughout this project, which would not have been possible without his continuous support and feedback.
 I would also like to acknowledge the efforts of the other members of my committee, Dr.
@@
 Mark Conde, and Dr.
-Hui Zhang for always providing new ideas and perspectives.
+Hui Zhang for offering new ideas and perspectives.
 Additionally, I truely appreciate the support and encouragement I received from the entire physics department at UAF, particularly Dr.
@@
 Chung-Sang Ng, Dr.
-Peter Delamere, Ellen Craig and Jeanie Talbot.
-I would like to extend a particularly special thank you to Flora G**** and all the staff of the Keith B.
-Mather Libray for providing invaluable assistance in retrieving sources for my research, no matter how old and obscure my requests were.
+Peter Delamere, Ellen Craig and Jeanie Talbot, as well as my fellow grad students, Victoriya, Christina, Jason, Robin, Joe, Nathan, Soma, Bishwa, Vitaliy, and John for both the homework and coding assistance and companionship over the last several years.
+I would like to extend a particularly special thank you to Flora Grabowska and all the staff of the Keith B.
+Mather Libray for providing invaluable assistance in retrieving a great number of old, obscure, and difficult to locate sources for my research.
 
-To my family, thank you all for the supporting me for years, particularly my Mom, for making me a much better writer than I have any right to be as a scientist, my Dad, for always being there to answer my programming questions and helping me fix my code when I had now idea how I had broken it, and my brother, Henry, generally for just existing and providing a friendly voice of reason when I really needed it.
-Additionally, thank you to my friends Katie, Ari, Bonnie, Heather, Jayne, and Stephanie for being there over these last few years despite being 4000 miles away, to Laura for giving me perspective and a laugh regardless of the time difference.
-Finally, I would like to extend a very special thank you to Allison, Rosie, Emma, and Kinley on the homefront for their tireless effort in keeping me sane, healthy, and happy throughout this whole process.
+To my family, thank you all for supporting me for years, particularly my Mom, for helping me become a much better writer than I would have been otherwise, my Dad, for always picking up the phone for my programming emergencies, and my brother, Henry, for giving me a friendly voice of reason when I really needed it.
+Additionally, I would like to thank my friends back East, Katie, Ari, Bonnie, Heather, Jayne, and Stephanie for supporting me over these last few years despite being 4000 miles away, and to Laura for always brightening my spirits and making my life feel a little less hectic.
+Finally, a huge thank you to Allison, Rosie, Emma, and Kinley on the homefront for their tireless effort in keeping me sane, healthy, and happy throughout this whole process.
 I couldn't have done it without you guys.

### 2017-02-03 introduction.tex
Acknowledgements revised

@@
 Additionally, polar patches occur relatively frequently, providing a good opportunity to study irregularity coupling at different scales \citep{Rodger1996}.
-In theory, many of the results presented in this work pertaining to the strong gradients on the edges of patches could also easily be applied to the edges of polar holes, but to date, far less attention has been paid to polar holes in the literature, making it challenging to accurately put observations in context \citep{Makarevich2015}.
+In theory, many of the results presented in this work pertaining to the strong gradients on the edges of patches could also easily be applied to the edges of polar holes, but to date, far less attention has been paid to polar holes in the literature, making it challenging to accurately put observations in context \citep{Makarevich2015b}.
 

### 2017-02-03 paper1.tex
Corrections to Abstract and Intro
Formatting corrections made
Acknowledgements revised

@@
 Following the same idea, a reasonable assumption is that additional density implied by our observations was transported from somewhere else.
-Several phenomena may be responsible for this including polar patches \citep{Wever1984}, Sun-aligned arcs \citep{Valladares1994}, and storm-enhanced densities \citep{Foster1993}.
+Several phenomena may be responsible for this including polar patches \citep{Weber1984}, Sun-aligned arcs \citep{Valladares1994}, and storm-enhanced densities \citep{Foster1993}.
 The polar patches in particular are known to travel long distances across the polar cap and even well into the nightside auroral zone \citep{Moen2007,Oksavik2010}.
@@
 	\includegraphics[width=\textwidth]{p1f6.pdf}
-	\caption[Cojugate patchy day example]{Conjugate measurements for two selected events on (a--d) March 14, 2010 and (e-h) April 6, 2010.
+	\caption[An example of conjugate observations of patchy backscatter]{Conjugate measurements for two selected events on (a--d) March 14, 2010 and (e-h) April 6, 2010.
 Shown are the IMF \(B_y\) (blue) and \(B_z\) (red) components (Figures \ref{fig:patchy_example}a and \ref{fig:patchy_example}e), MCM line-of-sight velocity versus UT and slant range in beam 14 (Figures \ref{fig:patchy_example}b and \ref{fig:patchy_example}f), RKN velocity in beam 4 (Figure \ref{fig:patchy_example}c and \ref{fig:patchy_example}g), and RISR-N electron density versus UT and altitude (Figure \ref{fig:patchy_example}d and \ref{fig:patchy_example}h).

### 2017-02-03 paper3.tex
Acknowledgements revised

@@
 
-The 2D electric field \(\vec{E}\)  and plasma drift velocity \(\vec{V}_E\) can be found from incoherent scatter radar (ISR) measurements, using the standard ISR method utilizing line-of-sight ion velocity data \citep{Hei08}.
+The 2D electric field \(\vec{E}\)  and plasma drift velocity \(\vec{V}_E\) can be found from incoherent scatter radar (ISR) measurements, using the standard ISR method utilizing line-of-sight ion velocity data \citep{Heinselman2008}.
 Additionally, ISRs have been previously used to determine density patterns in the ionosphere \citep{Semeter2009,Dahlgren2012a,Dahlgren2012b}.
@@
 Typically, AMISR systems have two types of pulses, a long pulse for \(F\)-region observations with 72-km range resolution and an alternating code pulse for \(E\)-region observations with 4.5-km range resolution.
-The radar measures a power spectrum and the electron density, ion and electron temperatures, and ion drift velocity are found using a standard ISR technique \citep{Evans1969,Risbeth1985,Nicolls2007}.
-Calibration accounts for the noise and system constants \citep{Nicolls2007} and is done using either plasma line measurements (for summer daytime) or ionosonde measurements (for all other times) \citep{Bahcivan2010,Themens2014a}.
+The radar measures a power spectrum and the electron density, ion and electron temperatures, and ion drift velocity are found using a standard ISR technique \citep{Evans1969,Rishbeth1985,Nicolls2007a}.
+Calibration accounts for the noise and system constants \citep{Nicolls2007a} and is done using either plasma line measurements (for summer daytime) or ionosonde measurements (for all other times) \citep{Bahcivan2010,Themens2014}.
 3D convection velocity and electric field vectors in the $F$ region are also found by combining all ion drift components measured at a particular magnetic latitude (MLAT) and assuming that both \(\vec{E}\) and \(\vec{V}_E\) are constant in magnetic longitude (MLON) \citep{Heinselman2008}.
@@
 Figure \ref{fig:radar_map} shows the FoV of both RKN, located at 62.82\(\deg\)N, 93.11\(\deg\)W, geographic (72.6\(\deg\)N, 26.4\(\deg\)W, geomagnetic), and RISR-N, located at 74.73\(\deg\)N, 94.91\(\deg\)W, geographic (82.8\(\deg\)N, 36.8\(\deg\)W geomagnetic).
-Geomagnetic coordinates were determined using the Altitude Adjusted Corrected Geomagnetic Coordinates (AACGM) model \citep{Shepard2014}.
+Geomagnetic coordinates were determined using the Altitude Adjusted Corrected Geomagnetic Coordinates (AACGM) model \citep{Shepherd2014}.
 The RKN footprint consists of individual radar cells, outlined in dashed lines.
@@
 In this study, small-scale plasma irregularity characteristics were analyzed in the context of factors that affect the linear growth rate of the gradient-drift instability.
-Previous experimental studies of GDI and its controlling factors focused on the convection electric field, electron density, fractional perturbation density, and density gradients derived from 1D measurements \citep[e.g.][]{Villian1986,Fukumoto1999,Fukumoto2000,Danskin2002,Oksavik2010,Moen2012,Makarevich2014b,Burston2016}.
+Previous experimental studies of GDI and its controlling factors focused on the convection electric field, electron density, fractional perturbation density, and density gradients derived from 1D measurements \citep[e.g.][]{Villain1986,Fukumoto1999,Fukumoto2000,Danskin2002,Oksavik2010,Moen2012,Makarevich2014b,Burston2016}.
 In contrast, the current study focused on the previously uninvestigated directional dependence of radar echo occurrence and intensity and comparison with theoretical predictions for arbitrary gradient, electric field, and wave propagation vectors.
@@
 To achieve perpendicularity to the magnetic field in the polar cap, HF radars rely on beam refraction, which is dependent on the plasma density \citep{Bristow2011,Koustov2014,Lamarche2015}.
-High plasma density can also reduce the occurrence of backscatter observed by radar by smoothing the gradients required for GDI \citep{Ruohoniemi1997,Koustov2004}, through \(E\)-region shorting \citep{Dan02,Kan12,Lam15}, or through attenuation of the signal through a dense \(D\) region \citep{Danskin2002}.
+High plasma density can also reduce the occurrence of backscatter observed by radar by smoothing the gradients required for GDI \citep{Ruohoniemi1997,Koustov2004}, through \(E\)-region shorting \citep{Danskin2002,Kane2012,Lamarche2015}, or through attenuation of the signal through a dense \(D\) region \citep{Danskin2002}.
 The peak in occurrence of all echoes in Figure \ref{fig:density}a agrees with previous results that there is an optimal density for HF radar observations of \(F\)-region irregularities \citep{Danskin2002,Makarevich2014b,Koustov2014,Lamarche2015}.
@@
 A summary of the results that agree with these predictions is presented in Table \ref{tab:results}.
-The relationship between \(\vec{G}\) and \(\vec{V}_E\) described in entries 9 and 10 in Table \ref{tab:results} matches previously-reported observations of an asymmetry between the trailing and leading edge \citep[][]{Weber1984,Milan2002b,Koustov2012,Moe2012}.
+The relationship between \(\vec{G}\) and \(\vec{V}_E\) described in entries 9 and 10 in Table \ref{tab:results} matches previously-reported observations of an asymmetry between the trailing and leading edge \citep[][]{Weber1984,Milan2002b,Koustov2012,Moen2012}.
 Most of these previous studies observed asymmetry only qualitatively and compared only the leading and trailing edge of patches instead of considering all directions around it.
@@
 However, the peak in echo occurrence occurred at \(\gamma\approx0.0001\) s\(^{-1}\), corresponding to a rise time of over 2.5 hours.
-In comparison with previous studies, \citet{Moe12} found through sounding rocket measurements \(\gamma = 0.021\) s\(^{-1}\) and a rise time of 47.6 s.
+In comparison with previous studies, \citet{Moen2012} found through sounding rocket measurements \(\gamma = 0.021\) s\(^{-1}\) and a rise time of 47.6 s.
 \citet{Carlson2007} estimated the rise time when GDI is the only instability mechanism operational as 5--10 min, but they also proposed a 2-step process in which KHI seeds GDI, in which case the rise time would be \(\sim50\) s.

### 2017-02-03 acknowledgements.tex
Minor revisions to Introduction and Conclusion

@@
 Hui Zhang for offering new ideas and perspectives.
-Additionally, I truely appreciate the support and encouragement I received from the entire physics department at UAF, particularly Dr.
+Additionally, I truly appreciate the support and encouragement I received from the entire physics department at UAF, particularly Dr.
 David Neuman, Dr.
@@
 I would like to extend a particularly special thank you to Flora Grabowska and all the staff of the Keith B.
-Mather Libray for providing invaluable assistance in retrieving a great number of old, obscure, and difficult to locate sources for my research.
+Mather Library for providing invaluable assistance in retrieving a great number of old, obscure, and difficult to locate sources for my research.
 

### 2017-02-03 conclusion.tex
Minor revisions to Introduction and Conclusion

@@
 This is in part due to the complex and turbulent nature of the ionosphere, particularly in the polar regions, but also due to some of the inherent limitations in traditional observation techniques which can make it very challenging to identify exactly what processes are operational.
-In Section \ref{sec:summary}, we summarize the results that have previously been discussed in Chapters \ref{sec:paper1}--\ref{sec:paper3}, and consider what overarching conclusions can be derived from them.
+In Section \ref{sec:summary}, we summarize the results that have previously been presented in Chapters \ref{sec:paper1}--\ref{sec:paper3}, and consider what overarching conclusions can be derived from them.
 In the following Section \ref{sec:futurework}, we also provide recommendations for future research which will help answer some of the still outstanding questions and further our understanding of small-scale polar \(F\)-region plasma irregularities.
@@
 
-It was also demonstraighted in Chapter \ref{sec:paper1} that, in the polar caps, backscatter is often observed at nighttime even when background plasma density is low and propagation conditions are traditionally thought to be too poor for FAIs to be observed with a HF radar.
+It was also demonstrated in Chapter \ref{sec:paper1} that, in the polar caps, backscatter is often observed at nighttime even when background plasma density is low and propagation conditions are traditionally thought to be too poor for FAIs to be observed with a HF radar.
 This implies an additional source of nighttime plasma density that is not easily accounted for by models.
@@
 
-In addition to simply providing additional background density so nighttime propagation conditions are sufficient for the radar to observe FAIs, polar patches can contribute to plasma irregularity formation directly by providing background density gradients on which GDI can operate.
+In addition to simply enhancing background density so nighttime propagation conditions are sufficient for the radar to observe FAIs, polar patches can contribute to plasma irregularity formation directly by providing background density gradients on which GDI can operate.
 This raises the important question of exactly how and where plasma structures form around polar patches.
@@
 
-Additionally, linear theory predicts that the irregularity growth rate is depends on the wavevector direction, \(\uvec{k}\), so instability growth may anisotropic and the orientation of the patch and convection velocity relative to the radar, not just relative to each other, can also impact the observed irregularities.
+Additionally, linear theory predicts that the irregularity growth rate is depends on the wavevector direction, \(\uvec{k}\), so instability growth may be anisotropic and the orientation of the patch and convection velocity relative to the radar, not just relative to each other, can also impact the observed irregularities.
 In Chapter \ref{sec:paper2}, it was demonstrated that, in the \(F\) region, the GDI growth rate is highest for wavevectors perpendicular to the gradient, \(\vec{G}\).
@@
 When the relative orientations of all three quantities results in a positive GDI growth rate, the occurrence of plasma irregularities is greater.
-An important conclusion that can be derived from this work is, therefore, that while asymmetry and anisotropy are independent processes, the small-scale plasma structuring surrounding polar patches seems to exhibit both.
+An important conclusion that can be derived from this work is, therefore, that asymmetry and anisotropy should not be used interchangeably and that the small-scale plasma structuring surrounding polar patches seems to exhibit both.
 
@@
 As was demonstrated in Chapter \ref{sec:paper2} and discussed above, if small-scale irregularity growth is directionally dependent, this is also the optimal orientation from which to observe the plasma structures, leading to increased echo occurrence.
-Thus modeling presented in Chapter \ref{sec:paper2} provide an alternative interpretation of earlier experimental results presented in Chapter \ref{sec:paper1}.
+Thus modeling results presented in Chapter \ref{sec:paper2} provide an alternative interpretation of earlier experimental results presented in Chapter \ref{sec:paper1}.
 
@@
 A hypothesis proposed in Chapter \ref{sec:paper1} was that polar patches, which are not typically well represented by ionospheric density models, occur frequently in the nighttime \(F\) region and, on average, enhance the background plasma density enough that propagation conditions are sufficient for HF radar backscatter to regularly be observed.
-Experimental surveys of polar patches have shown low occurrence at nighttimes, however the occurrence of patches and their intensity increase at equinox \citep{Rodger1996}, which is where frequent nighttime backscatter is predominantly observed according to our observations, Chapter \ref{sec:paper1}.
+Experimental surveys of polar patches have shown low occurrence at nighttimes, however the occurrence of patches and their intensity increase at equinox \citep{Rodger1996}, which is when frequent nighttime backscatter is predominantly observed according to our observations, Chapter \ref{sec:paper1}.
 The expected signature in the radar data from this kind of scenario would be a night with a series of isolated ``patches'' of backscatter.
@@
 However, these results also predicted anisotropy in the growth rate, which is more challenging to confirm from small-scale irregularity observations, although some evidence of such has been provided in Chapter \ref{sec:paper3}.
-It may be possible to further the theoretical understanding of GDI at small scales by developing a 3D picture using kinetic theory.
+It may be possible to further the theoretical understanding of GDI at small scales by developing a 3D formulism using kinetic theory.
 \citet{Basu1995} developed such a formulation for a specific vector geometry throughout all altitudes within the ionosphere, but expanding this to general vector directions would not be trivial.
@@
 \label{sec:fw_fracdens}
-In Chapter \ref{sec:paper3}, both the occurrence and power of radar backscatter are used as proxies to indicate the ``amount'' of irregularities that exist in the ionosphere.
-However, the trends from these quantities are rarely in good agreement, giving rise to the question what, if anything, does each of them indicate about the plasma structuring in the ionosphere.
+In Chapter \ref{sec:paper3}, both the occurrence and power of radar backscatter were used as proxies to indicate the ``amount'' of irregularities that exist in the ionosphere.
+However, the trends from these quantities were rarely in good agreement, giving rise to the question what, if anything, does each of them indicate about the plasma structuring in the ionosphere.
 The factors that control each need to be isolated, but this is nontrivial.
@@
 Irregularities may develop as predicted by linear GDI theory when their amplitudes are small but other processes may impact larger density perturbations, such as shear-driven instabilities \citep{Gondarenko2006} or multistep processes \citep{Carlson2012}.
-\(B_y\) the radar equation, the power returned should be proportional to the product of the background density and fractional density, squared, Equation \ref{eqn:fractional_density}.
+By the radar equation, the power returned should be proportional to the product of the background density and fractional density, squared, Equation \ref{eqn:densgrad}.
 However, there is no general consensus on what controls the fractional density in the \(F\) region, although it has been suggested that it is either constant or a function of convection speed, \(V_E\) \citep{Kustov1988,Haldoupis1990,Makarevich2014b}.
@@
 
-Our understanding of ionospheric plasma irregularities is in no way complete.
+Overall, this body of work has demonstrated that small-scale plasma irregularities in the polar \(F\) region develop in accordance with directionally-dependent linear GDI theory, although large amplitude perturbations quickly become controlled by more complicated processes.
+Additionally, global factors can cause large-scale changes in background density and plasma convection patterns, which in turn effect the opperation of GDI on the microphysical level.
+
+%Our understanding of ionospheric plasma irregularities is in no way complete.
 The issues outlined in this section are a small subset of questions to be addressed by future research.

### 2017-02-03 introduction.tex
Minor revisions to Introduction and Conclusion

@@
 At the magnetopause, magnetic reconnection can occur between the IMF and the Earth's magnetic field creating open field lines, light blue in Figure \ref{fig:ste}.
-This reconnection occurs on last closed field line intersecting the magnetopause (highlighted in yellow in Figure \ref{fig:ste}).
+This reconnection occurs on last closed field line intersecting the magnetopause (highlighted in orange in Figure \ref{fig:ste}).
 Magnetic field lines originating from the Earth are then directly connected to the Sun through the solar wind, which consequently allows highly energized solar particles to enter the magnetosphere.
@@
 If \(\nu_\alpha \gg \Omega_\alpha\), the particle collides with other particles much more often than it completes a gyration, so it is considered collisional.
-The gyrofrequency is dependent on the mass and charge of the particle and the strength of the magnetic field, \(\Omega_\alpha = q_\alpha B/m_\alpha\), so it is roughly constant throughout the ionosphere because the magnetic field strength does not change much throughout these altitudes.
+The gyrofrequency is dependent on the mass (\(m_\alpha\)) and charge of the particle (\(q_\alpha\)) and the strength of the magnetic field (\(B\)), \(\Omega_\alpha = q_\alpha B/m_\alpha\), so it is roughly constant throughout the ionosphere because the magnetic field strength does not change much throughout these altitudes.
 The collision frequency depends on the density of a species relative to the density of all other species in the plasma and can be calculated from a series of standard expressions \citep{Schunk1980,Schunk2009}.
@@
 Small-scale structuring is generally considered to be all irregularities less than 100 m \citep{Kelley2009}.
-Although small-scale irregularities are not responsible for scintillation that directly impact GPS/GNSS performance, they can be the easiest to study because of extensive datasets collected with a large number of ionospheric radars around the world that have been in operation for many decades (discussed further in Section \ref{sec:superdarn}).
+Although small-scale irregularities are not responsible for scintillation that directly impacts GPS/GNSS performance, they can be the easiest to study because of extensive datasets collected with a large number of ionospheric radars around the world that have been in operation for many decades (discussed further in Section \ref{sec:superdarn}).
 Because electrons move far more easily along magnetic field lines than across them, density perturbations tend to be aligned with the magnetic field and are often known as field-aligned irregularities (FAI), Figure \ref{fig:polarirreg}d.
@@
 However, using this simple single-pulse method, another pulse cannot be transmitted until the first pulse returns, limiting the temporal resolution that can be achieved.
-This can be improved by using a multi-pulse scheme, which will be discussed below \citep{Farley1972,Greenwald1983,Greenwald1985}.
+This can be improved by using a multi-pulse scheme, which is further discussed below \citep{Farley1972,Greenwald1983,Greenwald1985}.
 
@@
 The Nyquist frequency is then \(f_n = 37.5\) Hz, which corresponds to a maximum measurable Doppler velocity of \(\sim550\) m/s.
-Flows in the ionosphere have been known to exceed 1000--2000 m/s, so single-pulse sampling is insufficient for these purposes.
+Flows in the ionosphere are known to exceed 1000--2000 m/s, so single-pulse sampling is insufficient for these purposes.
 
@@
 The first observations of the ionosphere by CSRs were made in the 1930s \citep{Eckersley1937,Harang1938}.
-Over the next several decades, numerous ground-based radars have been used to advance our understanding of the ionosphere and radio aurora \citep{Hultqvist1964,Leadabrand1965,Unwin1972,Sahr1996}.
+Over the next several decades, numerous ground-based radars have been used to advance our understanding of the ionosphere  \citep{Hultqvist1964,Leadabrand1965,Unwin1972,Sahr1996}.
 In the 1970s and 1980s, the Scandanavian Twin Auroral Radar Experiment (STARE) produced the first continuous large data-set, similar to how modern CSR systems are run.
-STARE consisted of two very high frequency (VHF) radars with overlapping fields-of-view (FoV) that were designed to measure FAIs in the \(E\) region of the ionosphere \citep{Greenwald1997}.
+STARE consisted of two very high frequency (VHF) radars with overlapping FoVs that were designed to measure FAIs in the \(E\) region of the ionosphere \citep{Greenwald1997}.
 The advantage of two radars with overlapping FoVs was the ability to observe the same structures from two different directions.
@@
 Although the IRI model is very useful for providing background plasma density conditions where no instruments are available to make measurements, it is important to bear in mind that it is a model and therefore only represents the average conditions.
-Although it is fairly accurate at mid-latitudes \citep{Coisson2006,Bilitza2012}, it does not necessarily represent variations within the polar cap well \citep{Themens2014,Makarevich2015b}.
+Although it is fairly accurate at mid-latitudes \citep{Coisson2006,Bilitza2012}, it does not necessarily represent well variations within the polar cap \citep{Themens2014,Makarevich2015b}.
 In particular, the height of the \(F\)-region peak tends to be underestimated in daytime and the bottomside thickness does not show realistic seasonal and diurnal variations \citep{Themens2014}.
@@
 Large-scale structuring in the polar cap can take a variety of forms, Section \ref{sec:polar_structure}, but here the focus is on large regions of enhanced plasma density known as polar patches  \citep[e.g.][]{Weber1984,Weber1986,Buchau1983,Buchau1985}.
-One of the primary interests of this thesis is the relationships between large-scale plasma structures ans small-scale irregularity formation.
+One of the primary interests of this thesis is the relationships between large-scale plasma structures and small-scale irregularity formation.
 This structuring has long been observed around polar patches \citep{Weber1984,Milan2002b,Moen2012}.
@@
 The first observations of polar patches were made in the 1960s by \citet{Hill1963}.
-They can range in size from 100 km up to 1500 km and can have peak densities as much as eight times that of the background ionospheric plasma \citep{Weber1986,Hosokawa2014}.
+Polar patches can range in size from 100 km up to 1500 km and can have peak densities as much as eight times that of the background ionospheric plasma \citep{Weber1986,Hosokawa2014}.
 Polar patches are known to emerge on the dayside and then propagate with the background plasma convection across the polar cap to the nightside, where they either disintegrate or recombine with enhanced-density plasma in the auroral oval \citep{Weber1985,Weber1986}.
@@
 The Farley-Buneman instability (FBI), or the modified two-stream instability tends to be a major factor in the \(E\) region where ion inertial effects are strong \citep{Farley1963,Buneman1963}.
-Also operational in the \(E\) region but more dominant at higher altitudes is the gradient-drift instability (GDI) \citep{Simon1963,Hoh1963,Linson1970}.
+Also operational in the \(E\) region but more dominant at higher, \(F\)-region altitudes is the gradient-drift instability (GDI) \citep{Simon1963,Hoh1963,Linson1970}.
 If a wave propagation vector has a component along the magnetic field line, the current-convective instability (CCI) can be operational even if GDI is not \citep{Hoh1960,Ossakow1979,Chaturvedi1981}.
@@
 	S_\alpha = \frac{1}{1+D_\alpha^2} \quad\quad\quad\quad\quad\quad&\quad
-	D_\alpha = -\frac{i}{\Omega_\alpha}(\omega-\vec{k}\cdot{V}_{\alpha 0})+r_\alpha
+	D_\alpha = -\frac{i}{\Omega_\alpha}(\omega-\vec{k}\cdot\vec{V}_{\alpha 0})+r_\alpha
 \end{split}
 \end{equation}
-Throughout this thesis, \(T_\alpha\) and \(m_\alpha\) refer to the temperature and mass of species \(\alpha\), respectively.
+Throughout this thesis, \(T_\alpha\) refer to the temperature of species \(\alpha\).
 The unit vector \(\uvec{b}\) is in the direction of the magnetic field \(\vec{B}\) and \(\vec{k}_\perp\) is the component of the wavevector \(\vec{k}\) perpendicular to the magnetic field while \(\vec{k}_\parallel\) is the component parallel to it.

### 2017-02-03 conclusion.tex
THESIS TO COMMITTEE

@@
   \includegraphics[width=\textwidth]{conclusion.pdf}
-  \caption[Irregularity production factors]{Summary of the main conclusions in this thesis.
+  \caption[Irregularity production factors]{{\:}Irregularity production factors\\ Summary of the main conclusions in this thesis.
 The radar of interest (MCM) is highlighted in pink.
@@
 
-%An important conclusion that can be derived from this work is, therefore, that asymmetry and anisotropy are independent such that the plasma structures can exhibit either of them independently or both or neither.
-
-%The other group of results presented in Chapters \ref{sec:paper2} and \ref{sec:paper3} is related to the fact that the plasma drift and wavevector directions influence the asymmetry that is observed near polar patches.
-
 In Chapter \ref{sec:paper2}, it was demonstrated from the modeling point of view that the direction that the patch is drifting relative to its elongation direction impacts the orientations of \(\vec{G}\) and \(\vec{V}_E\) and controls the GDI growth rate.
@@
 
-Additionally, linear theory predicts that the irregularity growth rate is depends on the wavevector direction, \(\uvec{k}\), so instability growth may be anisotropic and the orientation of the patch and convection velocity relative to the radar, not just relative to each other, can also impact the observed irregularities.
+Additionally, linear theory predicts that the irregularity growth rate depends on the wavevector direction, \(\uvec{k}\), so instability growth may be anisotropic and the orientation of the patch and convection velocity relative to the radar, not just relative to each other, can also impact the observed irregularities.
 In Chapter \ref{sec:paper2}, it was demonstrated that, in the \(F\) region, the GDI growth rate is highest for wavevectors perpendicular to the gradient, \(\vec{G}\).
@@
 In addition, if a SuperDARN radar itself is positioned close to the terminator, it is observing the density gradient along its boresight, Figure \ref{fig:conclusion}.
-As was demonstrated in Chapter \ref{sec:paper2} and discussed above, if small-scale irregularity growth is directionally dependent, this is also the optimal orientation from which to observe the plasma structures, leading to increased echo occurrence.
+As was demonstrated in Chapter \ref{sec:paper2} and discussed above that if small-scale irregularity growth is directionally dependent, this is also the optimal orientation from which to observe the plasma structures, leading to increased echo occurrence.
 Thus modeling results presented in Chapter \ref{sec:paper2} provide an alternative interpretation of earlier experimental results presented in Chapter \ref{sec:paper1}.
-
-%The production and characteristics of small-scale plasma irregularities and the factors that control them are complex.
-Radar studies must differentiate between production and propagation factors, determining if in fact irregularity production is more favorable or if background conditions just make observations of irregularities more likely.
-This thesis helps advance our understanding by identifying a number of situations where small-scale irregularities are observed and establishing plausible reasons why either production is favorable according to plasma theory or radar observation is likely.
 
@@
 An example of such an event was presented, Figure \ref{fig:patchy_example}, but it is unknown how often events like this occur.
-Additionally, the intensity and size of a patch necessary to cause the radar beam to refract sufficiently must be established, possibly through further use of raytracing simulations with customizable density profiles.
+Additionally, the intensity and size of a patch necessary to cause the radar beam to refract sufficiently must be established, possibly through further use of ray tracing simulations with customizable density profiles.
 
@@
 This presents an unique opportunity to further investigate the directional dependence of small-scale irregularities by observing the differences in backscatter observed from the same scattering volume at different directions.
-Although straight-forward in theory, this experiment would most likely prove nontrivial in practice for the following reasons.
+Although straight forward in theory, this experiment would most likely prove nontrivial in practice for the following reasons.
 Backscatter tends to be limited spatially to the statistical high occurrence band shown in Figure \ref{fig:month_avg_occ}, and at least with INV, this band probably occurs at closer ranges than needed, so coincident backscatter is probably rare.
@@
 However, there is no general consensus on what controls the fractional density in the \(F\) region, although it has been suggested that it is either constant or a function of convection speed, \(V_E\) \citep{Kustov1988,Haldoupis1990,Makarevich2014b}.
-In Chapter \ref{sec:paper3}, we additionally propose that it may be a function of the instability growth rate, \(\gamma\).
+In Chapter \ref{sec:paper3}, we additionally propose that it may be a function of the instability growth rate, \(\gamma\), and through that, the gradient strength, \(G\).
 Because ground-based radars are so often used to investigate ionospheric irregularities, this is an important issue to resolve to develop an accurate picture of how radar observations relate to plasma irregularities.
 
-Overall, this body of work has demonstrated that small-scale plasma irregularities in the polar \(F\) region develop in accordance with directionally-dependent linear GDI theory, although large amplitude perturbations quickly become controlled by more complicated processes.
-Additionally, global factors can cause large-scale changes in background density and plasma convection patterns, which in turn effect the opperation of GDI on the microphysical level.
-
-%Our understanding of ionospheric plasma irregularities is in no way complete.
-The issues outlined in this section are a small subset of questions to be addressed by future research.
-Ground-based radar systems such as those employed extensively in this thesis will undoubtedly play an important role in this work.
-The widespread nature of the SuperDARN network makes it ideal for statistical studies of small-scale irregularities and mapping large-scale convection.
-Although scarcer, ISRs can provide a great wealth of data products, often in quite high resolution.
-Other instruments, e.g.
-satellites, ASIs, and ASIPs, can further supplement future studies of ionospheric irregularities.
-Equally important are models and theoretical studies of GDI as well as other instability mechanisms that provide insight as to how instabilities develop and explain observed trends.
-Ionospheric irregularities are part of a complex system and there are still many open questions about them, but the work presented here has advanced our understanding of their characteristics, development, and where and when we can expect to observe them.
+Overall, this body of work has demonstrated that small-scale plasma irregularities in the polar \(F\) region exhibit a previously unreported dependence on the propagation direction in accordance with directionally-dependent linear GDI theory, although large-amplitude perturbations quickly become controlled by more complicated processes.
+Additionally, global factors can cause large-scale changes in background density and plasma convection patterns, which in turn affect the operation of GDI on the microphysical level.
 

### 2017-02-03 introduction.tex
THESIS TO COMMITTEE

@@
 	\includegraphics[width=\textwidth,angle=0]{ste.pdf}
-	\caption[Solar-terrestrial environment]{Diagram of the solar-terrestrial environment.
+	\caption[Solar-terrestrial environment]{{\:}Solar-terrestrial environment\\ Diagram of the solar-terrestrial environment.
 The solar wind, green arrows, contains both a steady stream of charged particles and the interplanetary magnetic field, dark blue.
@@
 	\includegraphics[width=\textwidth]{densprofile.pdf}
-	\caption[Ionosphere density profiles]{Altitude profiles of electron density throughout the ionosphere found from the IRI model.
+	\caption[Ionosphere density profiles]{{\:}Ionosphere density profiles\\ Altitude profiles of electron density throughout the ionosphere found from the IRI model.
 Typical extents of the \(D\), \(E\), and \(F\) regions are identified.}
@@
 The \(E\) region typically exists between 90 and 130 km with a density peak between 105 and 110 km, depending on factors such as time of day and season.
-Above 130 km is considered to be the F region, which has a density peak around 250 km, Figure \ref{fig:densprofile} \citep{Kivelson1995}.
+Above 130 km is considered to be the \(F\) region, which has a density peak around 250 km, Figure \ref{fig:densprofile} \citep{Kivelson1995}.
 The \(F\) region can actually have two peaks under some daytime conditions, notated as the \(F1\) peak and the \(F2\) peak.
@@
 If \(\nu_\alpha \gg \Omega_\alpha\), the particle collides with other particles much more often than it completes a gyration, so it is considered collisional.
-The gyrofrequency is dependent on the mass (\(m_\alpha\)) and charge of the particle (\(q_\alpha\)) and the strength of the magnetic field (\(B\)), \(\Omega_\alpha = q_\alpha B/m_\alpha\), so it is roughly constant throughout the ionosphere because the magnetic field strength does not change much throughout these altitudes.
+The gyrofrequency is dependent on the mass (\(m_\alpha\)) and charge (\(q_\alpha\)) of the particle, and the strength of the magnetic field (\(B\)), \(\Omega_\alpha = q_\alpha B/m_\alpha\), so it is roughly constant throughout the ionosphere because the magnetic field strength does not change much throughout these altitudes.
 The collision frequency depends on the density of a species relative to the density of all other species in the plasma and can be calculated from a series of standard expressions \citep{Schunk1980,Schunk2009}.
@@
 	\includegraphics[width=0.5\textwidth]{irregularity.pdf}
-	\caption[Plasma irregularities in the polar ionosphere]{Plasma density irregularities occur in the polar cap at different scale sizes.
+	\caption[Plasma irregularities in the polar ionosphere]{{\:}Plasma irregularities in the polar ionosphere\\ Plasma density irregularities occur in the polar cap at different scale sizes.
 Polar patches, orange in Figures \ref{fig:polarirreg}a and \ref{fig:polarirreg}b, move with the background ionospheric convection, indicated by red arrows.
@@
 	\includegraphics[width=\textwidth,angle=180]{SuperDARNmap.pdf}
-	\caption[SuperDARN map]{Fields-of-view of all operational SuperDARN radars in both the northern (a) and southern (b) hemispheres.
+	\caption[SuperDARN map]{{\:}SuperDARN map\\ Fields-of-view of all operational SuperDARN radars in both the northern (a) and southern (b) hemispheres.
 Polar-latitude radars are shown shaded in blue.
@@
 In the studies presented here, data were primarily used from the SuperDARN radars located at Rankin Inlet, Canada (RKN) and McMurdo Station, Antarctica (MCM).
-These two radars have been shaded in Figure \ref{fig:superdarnmap}.
+These two radars have been outlined in pink in Figure \ref{fig:superdarnmap}.
 MCM is the only polar radar in the southern hemisphere that has been operational for more than 5 years.
@@
 	\includegraphics[width=\textwidth]{ISRmap.pdf}
-	\caption[ISR map]{FoVs of all ISRs in the northern polar region.
+	\caption[ISR map]{{\:}ISR map\\ FoVs of all ISRs in the northern polar region.
 The individual radars shown are identified by their commonly used acronyms: Poker Flat (PFISR), Resolute Bay North (RISR-N), Resolute Bay Canada (RISR-C), Sondrestrom (SFJ), European Incoherent Scatter Scientific Association (EISCAT), and EISCAT Svalbard (ESR).
@@
 High speed flow channels can ``cut'' segments off of a tongue of high-density plasma that has been pulled into the polar cap by convection \citep{Valladares1994,Valladares1998}.
-Models have also shown that fast-changing large-scale convection patterns does produce density structures similar to patches \citep{Anderson1988}.
+Models have also shown that fast-changing large-scale convection patterns do produce density structures similar to patches \citep{Anderson1988}.
 However, the mechanism that seems to be dominant for the majority of polar patches observed is transient magnetic reconnection \citep{Carlson2012}.
@@
 	\includegraphics[width=\textwidth,angle=0]{magnetosphere.pdf}
-	\caption[Magnetic reconnection footprint in the ionosphere]{Diagram of the interactions between the IMF, magnetosphere, and polar cap ionosphere.
+	\caption[Magnetic reconnection footprint in the ionosphere]{{\:}Magnetic reconnection footprint in the ionosphere\\ Diagram of the interactions between the IMF, magnetosphere, and polar cap ionosphere.
 Closed field lines in the Earth's magnetosphere are shown in grey, while open field lines are light blue.
@@
 Before the convection flow stops completely, Figure \ref{fig:patch_formation}e, the blob of dense plasma ``pinches off'' from the dayside plasma to become an independent patch \citep{Lockwood1992b}.
-The exact mechanism by which this happens is not well understood, but it is thought to be related to small variations in the IMF By component shifting the convection pattern slightly \citep{Cowley1980,Lockwood1992b}.
+The exact mechanism by which this happens is not well understood, but it is thought to be related to small variations in the IMF \(B_y\) component shifting the convection pattern slightly \citep{Cowley1980,Lockwood1992b}.
 By the time shown in Figure \ref{fig:patch_formation}f, the open-closed boundary has returned completely to its original position and there is a newly formed patch within the polar cap.
@@
 	\includegraphics[width=\textwidth,angle=0]{patch_formation.pdf}
-	\caption[Polar patch formation]{Polar patch formation through transient magnetopause reconnection.
+	\caption[Polar patch formation]{{\:}Polar patch formation\\ Polar patch formation through transient magnetopause reconnection.
 The open-closed boundary is the green line.
@@
 	\begin{tabular}{l l}
-	Limiting Case & Growth Rate, \(\gamma\) \\
+	\textbf{Limiting Case} & \textbf{Growth Rate, \(\gamma\)} \\
 	\hline
@@
 	\end{tabular}
-	\caption[Irregularity growth rates]{Irregularity growth rates in the ionosphere for three different limiting cases \citep{Makarevich2016a}.}
+	\caption[Irregularity growth rates]{{\:}Irregularity growth rates\\ Irregularity growth rates in the ionosphere for three different limiting cases \citep{Makarevich2016a}.}
 	\label{tab:gdi_exps}

### 2017-02-03 paper1.tex
THESIS TO COMMITTEE

@@
 
-\chapter[Solar control of \(F\) region radar backscatter: Further insights from observations in the southern polar cap]{Solar control of \(F\) region radar backscatter: Further insights from observations in the southern polar cap\footnote{Published as: {Lamarche}, L.~J., and R.~A.
+\chapter[Solar Control of \(F\) Region Radar Backscatter: Further Insights from Observations in the Southern Polar Cap]{Solar Control of \(F\) Region Radar Backscatter: Further Insights from Observations in the Southern Polar Cap\footnote{Published as: {Lamarche}, L.~J., and R.~A.
 {Makarevich} (2015), Solar control of {F} region radar backscatter: Further insights from observations in the southern polar cap, \textit{J.
@@
 	\includegraphics[height=\textwidth,angle=90]{p1f1.pdf}
-	\caption[Monthly average echo occurrence and electron density]{Monthly average plots of (a--d) echo occurrence and (e--h) peak electron density for 2010.
+	\caption[Monthly average echo occurrence and electron density]{{\:}Monthly average echo occurrence and electron density\\ Monthly average plots of (a--d) echo occurrence and (e--h) peak electron density for 2010.
 Four months are selected (one for each season) to show seasonal variations.
@@
 The final observation about Figure \ref{fig:month_avg_occ} concerns the results of raytracing simulations that were conducted for each hour of day 15 in each of the four months shown in Figure \ref{fig:month_avg_occ} using a version of standard ray tracing tools based on numerical solutions of the Hamiltonian raypath equations \citep{Haselgrove1963,Jones1975} that allow the consideration of arbitrary density distributions.
-In these simulations, IRI densities in a 2-D simulation domain aligned with beam 14 of MCM were used.
+In these simulations, IRI densities in a 2D simulation domain aligned with beam 14 of MCM were used.
 That is, plasma densities employed were varying both horizontally and vertically.
@@
 
+\begin{figure}
+	\centering
+	\includegraphics[height=\textwidth,angle=90]{p1f2.pdf}
+	\caption[Trends in high-occurrence band]{{\:}Trends in high-occurrence band\\ The variations in the echo band (a) amplitude of the fitted cosine function \(A\), (b) time shift or time of day of the peak \(B\), (c) average position of the band or the vertical shift of the cosine function from zero \(C\), and (d) width (BW).
+Different lines are for different years with the legend given in Figure \ref{fig:year_line}a.
+The fit errors are given by the vertical bars.}
+	\label{fig:year_line}
+\end{figure}
+
 The same analysis was performed for the entire period of interest 2010--2013 and the band characteristics are presented in Figure \ref{fig:year_line}.
@@
 when band position was located furthest away from the radar.
-
-\begin{figure}
-	\centering
-	\includegraphics[height=\textwidth,angle=90]{p1f2.pdf}
-	\caption[Trends in high-occurrence band]{The variations in the echo band (a) amplitude of the fitted cosine function \(A\), (b) time shift or time of day of the peak \(B\), (c) average position of the band or the vertical shift of the cosine function from zero \(C\), and (d) width (BW).
-Different lines are for different years with the legend given in Figure \ref{fig:year_line}a.
-The fit errors are given by the vertical bars.}
-	\label{fig:year_line}
-\end{figure}
 
@@
 	\includegraphics[height=\textwidth,angle=90]{p1f3.pdf}
-	\caption[Solar cycle trends of high-occurrence band]{Contour plots of the (a--d) echo band position \(r(t)\) and (e)--(h) occurrence versus UT and season.
+	\caption[Solar cycle trends of high-occurrence band]{{\:}Solar cycle trends of high-occurrence band\\ Contour plots of the (a--d) echo band position \(r(t)\) and (e)--(h) occurrence versus UT and season.
 Each row shows the data for a different year as given in the top right corner.
@@
 	\includegraphics[width=\textwidth]{p1f4.pdf}
-	\caption[Band position and occurrence versus SZA and \(N_mF_2\)]{The echo band range versus the (a) solar zenith angle (SZA) and (b) model peak F2 electron density (\(N_mF_2\)).
+	\caption[Band position and occurrence versus SZA and \(N_mF_2\)]{{\:}Band position and occurrence versus SZA and \(N_mF_2\)\\ The echo band range versus the (a) solar zenith angle (SZA) and (b) model peak F2 electron density (\(N_mF_2\)).
 Each point is color-coded in model \(E\) region density.
@@
 	\includegraphics[width=\textwidth]{p1f5.pdf}
-	\caption[Occurrence versus UT and IMF conditions]{The MCM echo occurrence versus UT color coded in the average (a) IMF \(B_y\), (b) IMF \(B_z\), and (c) PCS values.}
+	\caption[Occurrence versus UT and IMF conditions]{{\:}Occurrence versus UT and IMF conditions\\ The MCM echo occurrence versus UT color coded in the average (a) IMF \(B_y\), (b) IMF \(B_z\), and (c) PCS values.}
 	\label{fig:ut_occ}
@@
 	\includegraphics[width=\textwidth]{p1f6.pdf}
-	\caption[An example of conjugate observations of patchy backscatter]{Conjugate measurements for two selected events on (a--d) March 14, 2010 and (e-h) April 6, 2010.
+	\caption[An example of conjugate observations of patchy backscatter]{{\:}An example of conjugate observations of patchy backscatter\\ Conjugate measurements for two selected events on (a--d) March 14, 2010 and (e-h) April 6, 2010.
 Shown are the IMF \(B_y\) (blue) and \(B_z\) (red) components (Figures \ref{fig:patchy_example}a and \ref{fig:patchy_example}e), MCM line-of-sight velocity versus UT and slant range in beam 14 (Figures \ref{fig:patchy_example}b and \ref{fig:patchy_example}f), RKN velocity in beam 4 (Figure \ref{fig:patchy_example}c and \ref{fig:patchy_example}g), and RISR-N electron density versus UT and altitude (Figure \ref{fig:patchy_example}d and \ref{fig:patchy_example}h).
@@
 This is consistent with the Doppler velocity measured by MCM within the patch (dark blue colors).
-Similar consistency between the patch drift and plasma convection velocities was observed by \citet{Bristow2011} using MCM observations and, more recently, by \citet{Makarevich2015} using direct observations of 2-D convection velocity and density enhancements by RISR-N in the northern polar cap.
+Similar consistency between the patch drift and plasma convection velocities was observed by \citet{Bristow2011} using MCM observations and, more recently, by \citet{Makarevich2015} using direct observations of 2D convection velocity and density enhancements by RISR-N in the northern polar cap.
 This impolies that polar patches tend to move with the background convection, at least deep within the polar cap.
@@
 	\includegraphics[height=\textwidth,angle=90]{p1f7.pdf}
-	\caption[Summary of annual and diurnal variations]{Contour plots of the (a) average occurrence within the band, (b) band range, (c) model peak F-region electron density (\(N_mF_2\)) and (d) solar zenith angle (SZA) versus UT and month in 2010.
+	\caption[Summary of annual and diurnal variations]{{\:}Summary of annual and diurnal variations\\ Contour plots of the (a) average occurrence within the band, (b) band range, (c) model peak F-region electron density (\(N_mF_2\)) and (d) solar zenith angle (SZA) versus UT and month in 2010.
 Color bars are given to the right.

### 2017-02-03 paper2.tex
THESIS TO COMMITTEE

@@
 
-\chapter[A modeling study of asymmetries in plasma irregularity characteristics near gradient reversals]{A modeling study of asymmetries in plasma irregularity characteristics near gradient reversals\footnote{Published as: {Lamarche}, L.~J., and R.~A.
+\chapter[A Modeling Study of Asymmetries in Plasma Irregularity Characteristics Near Gradient Reversals]{A Modeling Study of Asymmetries in Plasma Irregularity Characteristics Near Gradient Reversals\footnote{Published as: {Lamarche}, L.~J., and R.~A.
 {Makarevich} (2016), A modeling study of asymmetries in plasma irregularity characteristics near gradient reversals, \textit{Ann.
@@
 	\includegraphics[width=\textwidth,angle=180]{p2f1.pdf}
-	\caption[Overview of altitudinal variations and model setup of MCM]{(a) Altitudinal variations of parameters of interest.
+	\caption[Overview of altitudinal variations and model setup of MCM]{{\:}Overview of altitudinal variations and model setup of MCM\\ (a) Altitudinal variations of parameters of interest.
 Shown are the ratios of collision frequencies and gyrofrequencies for electrons \(-r_e=\nu_e/|\Omega_e|\) (blue) and ions \(r_i=\nu_i/\Omega_i\) (green), the anisotropy parameter \(\psi=-r_i r_e\) (pink), and the conductivity ratio \(R=\sigma_H/\sigma_P = \left(r_i+r_e\right)/\left(1+\psi\right)\) (dashed black line on top of the green one).
@@
 	\includegraphics[width=\textwidth]{p2f2.pdf}
-	\caption[GDI growth rate within radar FoV]{The GDI growth rate within the radar's FoV for four cases of gradient reversal orientation and position (white line).
+	\caption[GDI growth rate within radar FoV]{{\:}GDI growth rate within radar FoV\\ The GDI growth rate within the radar's FoV for four cases of gradient reversal orientation and position (white line).
 In all cases, the convection direction \(\vec{V}_E\) (red arrow) is perpendicular to the gradient reversal direction.
@@
 	\includegraphics[width=\textwidth,angle=180]{p2f3.pdf}
-	\caption[GDI growth rates for set angle between \(\vec{V}_E\) and \(\vec{G}\)]{The difference in the GDI growth rate across the gradient reversal for four selected angles between the reversal and convection direction.
+	\caption[GDI growth rates for set angle between \(\vec{V}_E\) and \(\vec{G}\)]{{\:}GDI growth rates for set angle between \(\vec{V}_E\) and \(\vec{G}\)\\ The difference in the GDI growth rate across the gradient reversal for four selected angles between the reversal and convection direction.
 The gradient reversal direction \(\theta\) is on the \(x\)-axis and the average growth rate difference across the reversal \(\Delta\gamma\) is on the \(y\)-axis for each panel.
@@
 	\includegraphics[width=\textwidth,angle=180]{p2f4.pdf}
-	\caption[GDI growth rates for arbitrary angle between \(\vec{V}_E\) and \(\vec{G}\)]{Contour plots of the growth rate difference \(\Delta\gamma\)  versus the two angles of interest.
+	\caption[GDI growth rates for arbitrary angle between \(\vec{V}_E\) and \(\vec{G}\)]{{\:}GDI growth rates for arbitrary angle between \(\vec{V}_E\) and \(\vec{G}\)\\ Contour plots of the growth rate difference \(\Delta\gamma\)  versus the two angles of interest.
 The direction of the gradient reversal \(\theta\) is on the \(x\)-axis and the convection direction \(\phi\) is on the \(y\)-axis.
@@
 	\includegraphics[width=\textwidth,angle=180]{p2f5.pdf}
-	\caption[GDI growth rate within transitional region]{Analysis of the transitional region.
+	\caption[GDI growth rate within transitional region]{{\:}GDI growth rate within transitional region\\ Analysis of the transitional region.
 (a) The average conductivity ratio \(R\) (black line) and the radar virtual height (orange line) versus the slant range for the entire transitional region and parts of the \(E\) and \(F\) regions.

### 2017-02-03 paper3.tex
THESIS TO COMMITTEE

@@
-\chapter[Radar observations of density gradients, electric fields, and plasma irregularities near polar cap patches in the context of the gradient-drift instability]{Radar observations of density gradients, electric fields, and plasma irregularities near polar cap patches in the context of the gradient-drift instability\footnote{Submitted as: {Lamarche}, L.~J., and R.~A.
+\chapter[Radar Observations of Density Gradients, Electric Fields, and Plasma Irregularities Near Polar Cap Patches in the Context of the Gradient-Drift Instability]{Radar Observations of Density Gradients, Electric Fields, and Plasma Irregularities Near Polar Cap Patches in the Context of the Gradient-Drift Instability\footnote{Submitted as: {Lamarche}, L.~J., and R.~A.
 {Makarevich} (2017), Radar observations of density gradients, electric fields, and plasma irregularities near polar cap patches in the context of the gradient-drift instability, \textit{J.
@@
 Res.
-Space Physics}.}}
+Space Physics}, \textit{122}, \doi{10.1002/2016JA023702}.}}
 
@@
 
-In this study, a general expression for the GDI growth rate is used that considers arbitrary directions of the density gradient \(\uvec{g}\), electric field \(\uvec{e}\), and wavevector \(\uvec{k}\), and is applicable throughout the\(F\) region \citep[Equation 23]{Makarevich2014c}.
+In this study, a general expression for the GDI growth rate is used that considers arbitrary directions of the density gradient \(\uvec{g}\), electric field \(\uvec{e}\), and wavevector \(\uvec{k}\), and is applicable throughout the \(F\) region \citep[Equation 23]{Makarevich2014c}.
 
@@
 
-Figure \ref{fig:radar_map} shows the FoV of both RKN, located at 62.82\(\deg\)N, 93.11\(\deg\)W, geographic (72.6\(\deg\)N, 26.4\(\deg\)W, geomagnetic), and RISR-N, located at 74.73\(\deg\)N, 94.91\(\deg\)W, geographic (82.8\(\deg\)N, 36.8\(\deg\)W geomagnetic).
-Geomagnetic coordinates were determined using the Altitude Adjusted Corrected Geomagnetic Coordinates (AACGM) model \citep{Shepherd2014}.
-The RKN footprint consists of individual radar cells, outlined in dashed lines.
-Cells where SNR \(>\) 3 dB are color-coded in SNR measured at 18:16--18:17 UT on November 30, 2011.
-The locations of the 11 RISR-N beams at 300 km are shown by circles, with the color of each circle representing the electron density, \(n\), measured by RISR-N at 300 km during the post-integration interval closest to that of RKN.
-
 \begin{figure}
 	\includegraphics[width=\textwidth]{p3f1.pdf}
-  \caption[RKN/RISR-N Experimental Setup]{Experimental setup showing the FoVs of RKN and RISR-N.
+  \caption[RKN/RISR-N experimental setup]{{\:}RKN/RISR-N experimental setup\\ Experimental setup showing the FoVs of RKN and RISR-N.
 Lines of constant magnetic latitude are shown at \(75\deg\)N, \(80\deg\)N, and \(85\deg\)N.
@@
 \end{figure}
+
+Figure \ref{fig:radar_map} shows the FoV of both RKN, located at 62.82\(\deg\)N, 93.11\(\deg\)W, geographic (72.6\(\deg\)N, 26.4\(\deg\)W, geomagnetic), and RISR-N, located at 74.73\(\deg\)N, 94.91\(\deg\)W, geographic (82.8\(\deg\)N, 36.8\(\deg\)W geomagnetic).
+Geomagnetic coordinates were determined using the Altitude Adjusted Corrected Geomagnetic Coordinates (AACGM) model \citep{Shepherd2014}.
+The RKN footprint consists of individual radar cells, outlined in dashed lines.
+Cells where SNR \(>\) 3 dB are color-coded in SNR measured at 18:16--18:17 UT on November 30, 2011.
+The locations of the 11 RISR-N beams at 300 km are shown by circles, with the color of each circle representing the electron density, \(n\), measured by RISR-N at 300 km during the post-integration interval closest to that of RKN.
 
@@
 	\includegraphics[width=\textwidth,angle=180]{p3f2.pdf}
-  \caption[Polar Patch Event Overview]{Left column is a keogram representation of (a and b) RISR-N electron density, (c) RKN line-of-sight velocity, and (d) RKN SNR for an event on November 30, 2011, 17:00--19:00 UT.
+  \caption[Polar patch event overview]{{\:}Polar patch event overview\\ Left column is a keogram representation of (a and b) RISR-N electron density, (c) RKN line-of-sight velocity, and (d) RKN SNR for an event on November 30, 2011, 17:00--19:00 UT.
 Color scales for all quantities are given to the right of the respective row.
@@
 	\includegraphics[width=\textwidth,angle=180]{p3f3.pdf}
-  \caption[RKN/RISR-N Velocity Analysis]{RKN line-of-sight velocity versus the projection of 2D RISR-N velocity onto the closest RKN beam.
+  \caption[RKN/RISR-N velocity analysis]{{\:}RKN/RISR-N velocity analysis\\ RKN line-of-sight velocity versus the projection of 2D RISR-N velocity onto the closest RKN beam.
 Black dots represent all coincident echoes that were observed by RKN and RISR-N simultaneously over the day considered, whereas red dots represent only echoes that occurred within the high-density structure.
@@
 \includegraphics[width=\textwidth,angle=180]{p3f4.pdf}
-  \caption[Irregularity Dependence on Density Factors]{Scatter plots of RKN echo power versus RISR-N measurements of (a) electron density, \(n\), (b) density gradient magnitude, \(|\nabla n|\), and (c) gradient strength, \(G=|\nabla n/n|\).
+  \caption[Irregularity dependence on density factors]{{\:}Irregularity dependence on density factors\\ Scatter plots of RKN echo power versus RISR-N measurements of (a) electron density, \(n\), (b) density gradient magnitude, \(|\nabla n|\), and (c) gradient strength, \(G=|\nabla n/n|\).
 Black dots represent all echoes observed over the day, whereas red dots represent echoes that occur within the high-density structure.
@@
 	\includegraphics[width=\textwidth,angle=180]{p3f5.pdf}
-  \caption[Irregularity Dependence on Directional Factors]{Scatter plots of RKN echo power versus RISR-N measurements of (a) the gradient strength component, \(\uvec{k}\cdot\uvec{b}\times\vec{G}\), (b) the electric field component \(\uvec{k}\cdot\vec{E}\), (c) the normalized gradient component, \(\uvec{k}\cdot\uvec{b}\times\uvec{g}\), and (d) the normalized electric field component, \(\uvec{k}\cdot\uvec{e}\).
+  \caption[Irregularity dependence on directional factors]{{\:}Irregularity dependence on directional factors\\ Scatter plots of RKN echo power versus RISR-N measurements of (a) the gradient strength component, \(\uvec{k}\cdot\uvec{b}\times\vec{G}\), (b) the electric field component \(\uvec{k}\cdot\vec{E}\), (c) the normalized gradient component, \(\uvec{k}\cdot\uvec{b}\times\uvec{g}\), and (d) the normalized electric field component, \(\uvec{k}\cdot\uvec{e}\).
 The meaning of dots, crosses, vertical bars, and lines is the same as in Figure \ref{fig:density}.}
@@
 	\includegraphics[width=\textwidth,angle=180]{p3f6.pdf}
-  \caption[Irregularity dependence on GDI growth rate]{Same as Figure \ref{fig:directional}, but for (a) the dot product of the gradient and plasma drift velocity, \(\vec{G}\cdot\vec{V}_E\), (b) the calculated GDI growth rate, \(\gamma\), (c) the normalized dot product of the  gradient and plasma drift velocity, \(\uvec{g}\cdot\mathbf{\hat{v}}_E\), and (d) the normalized GDI growth rate, \(\gamma'\).}
+  \caption[Irregularity dependence on GDI growth rate]{{\:}Irregularity dependence on GDI growth rate\\ Same as Figure \ref{fig:directional}, but for (a) the dot product of the gradient and plasma drift velocity, \(\vec{G}\cdot\vec{V}_E\), (b) the calculated GDI growth rate, \(\gamma\), (c) the normalized dot product of the  gradient and plasma drift velocity, \(\uvec{g}\cdot\mathbf{\hat{v}}_E\), and (d) the normalized GDI growth rate, \(\gamma'\).}
   \label{fig:gamma}
@@
 \includegraphics[width=\textwidth,angle=180]{p3f7.pdf}
-  \caption[Relationship between the electric field and gradient components]{Scatter plots of RISR-N measurements of the electric field component, \(\uvec{k}\cdot\vec{E}\), versus the gradient component, \(\uvec{k}\cdot\uvec{b}\times\vec{G}\).
+  \caption[Relationship between the electric field and gradient components]{{\:}Relationship between the electric field and gradient components\\ Scatter plots of RISR-N measurements of the electric field component, \(\uvec{k}\cdot\vec{E}\), versus the gradient component, \(\uvec{k}\cdot\uvec{b}\times\vec{G}\).
 Points are color coded in RKN SNR, according to the color bar to the right.
@@
 
-The GDI growth rate applies to long wavelengths \(\lambda >1\) km in the $F$-region due to the fact that the fluid theory assumes the low-frequency (\(\omega \ll \nu_{in}\)) and long-wavelength (\(\lambda \gg 2\pi V_E\nu_{in}^{-1}\)) limits, where \(\nu_{in}\) is the ion-neutral collision frequency \citep[e.g.][]{Makarevich2014c}.
+The GDI growth rate applies to long wavelengths \(\lambda >1\) km in the \(F\)-region due to the fact that the fluid theory assumes the low-frequency (\(\omega \ll \nu_{in}\)) and long-wavelength (\(\lambda \gg 2\pi V_E\nu_{in}^{-1}\)) limits, where \(\nu_{in}\) is the ion-neutral collision frequency \citep[e.g.][]{Makarevich2014c}.
 In addition, density gradients estimated in this study refer to relative large scales \(L=G^{-1} =10\) km.
@@
 \begin{tabular}{rlc}
-	%\hline
 	\textbf{\#}& \textbf{Experimental Feature} & \textbf{Figure} \\
@@
 	12&Greater occurrence of structure echoes when \(\uvec{k}\cdot\uvec{b}\times\vec{G}\) and \(\uvec{k}\cdot\vec{E}\) are positive & \ref{fig:eg_scatter}c \\
-	%\hline
 \end{tabular}
-\caption[Results consistent with linear GDI theory]{A summary of the experimentally observed features that are consistent with the gradient-drift instability theory for arbitrary directions of the wavevector \(\uvec{k}\), electric field \(\uvec{e}\), and density gradient \(\uvec{g}\).}
+\caption[Results consistent with linear GDI theory]{{\:}Results consistent with linear GDI theory\\ A summary of the experimentally observed features that are consistent with the gradient-drift instability theory for arbitrary directions of the wavevector \(\uvec{k}\), electric field \(\uvec{e}\), and density gradient \(\uvec{g}\).}
 \label{tab:results}

### 2017-02-06 introduction.tex
Formatting corrections made

@@
 	\centering
-	\includegraphics[width=0.5\textwidth]{irregularity.pdf}
+	\includegraphics[height=0.75\textheight]{irregularity.pdf}
 	\caption[Plasma irregularities in the polar ionosphere]{{\:}Plasma irregularities in the polar ionosphere\\ Plasma density irregularities occur in the polar cap at different scale sizes.
@@
 \end{equation}
-An even more general consideration involves arbitrary gradient and plasma drift directions \citep{Keskinen1982a,Makarevich2014c}.
+An even more general consideration involves arbitrary gradient and plasma drift directions \citep{Keskinen1982,Makarevich2014c}.
 In this more general case, the directional dependence of GDI has been shown to be significant, such that the growth rate can actually
@@
 For the most part, the results of nonlinear numerical simulations very closely match expectations from linear GDI theory.
-In particular, it was found that the spatial power spectra are anisotropic with most of the power concentrated in the direction where linear GDI growth is most favorable \citep{Keskinen1981a,Keskinen1981b,Keskinen1982a,Gondarenko2001,Gondarenko2004b}.
+In particular, it was found that the spatial power spectra are anisotropic with most of the power concentrated in the direction where linear GDI growth is most favorable \citep{Keskinen1981a,Keskinen1981b,Keskinen1982,Gondarenko2001,Gondarenko2004b}.
 However, very large shears or stronger ion-inertial factors can cause isotropy \citep{Gondarenko2001,Gondarenko2006}.

### 2017-02-06 paper1.tex
Formatting corrections made

@@
 
-\chapter[Solar Control of \(F\) Region Radar Backscatter: Further Insights from Observations in the Southern Polar Cap]{Solar Control of \(F\) Region Radar Backscatter: Further Insights from Observations in the Southern Polar Cap\footnote{Published as: {Lamarche}, L.~J., and R.~A.
+\chapter[Solar Control of \textit{F} Region Radar Backscatter: Further Insights from Observations in the Southern Polar Cap]{Solar Control of \textit{F} Region Radar Backscatter: Further Insights from Observations in the Southern Polar Cap\footnote{Published as: {Lamarche}, L.~J., and R.~A.
 {Makarevich} (2015), Solar control of {F} region radar backscatter: Further insights from observations in the southern polar cap, \textit{J.
@@
 
-Production of these irregularities is reasonably well understood on a microphysical level, with GDI promoting irregularity growth for certain orientations of the density gradient vector relative to the convection electric field \citep{Keskinen1982a,Keskinen1982b,Makarevich2014c}.
+Production of these irregularities is reasonably well understood on a microphysical level, with GDI promoting irregularity growth for certain orientations of the density gradient vector relative to the convection electric field \citep{Keskinen1982,Keskinen1983a,Makarevich2014c}.
 The role of global factors, on the other hand, such as geomagnetic activity or solar wind conditions is much less investigated.
@@
 
-\subsection{\(\mathbf{F}\) Region Echo Band: A Comprehensive Model and Density Implications}
+\subsection{\textit{F} Region Echo Band: A Comprehensive Model and Density Implications}
 
@@
 This is consistent with the Doppler velocity measured by MCM within the patch (dark blue colors).
-Similar consistency between the patch drift and plasma convection velocities was observed by \citet{Bristow2011} using MCM observations and, more recently, by \citet{Makarevich2015} using direct observations of 2D convection velocity and density enhancements by RISR-N in the northern polar cap.
+Similar consistency between the patch drift and plasma convection velocities was observed by \citet{Bristow2011} using MCM observations and, more recently, by \citet{Makarevich2015b} using direct observations of 2D convection velocity and density enhancements by RISR-N in the northern polar cap.
 This impolies that polar patches tend to move with the background convection, at least deep within the polar cap.
@@
 
-\subsection{\(\mathbf{F}\) Region Irregularities: Suppression and Enhancement by Sunlight}
+\subsection{\textit{F} Region Irregularities: Suppression and Enhancement by Sunlight}
 
@@
 One mechanism through which IMF-controlled variations in convection may be responsible for changes in occurrence of plasma irregularities involves changing orientations between plasma density gradients \(\vec{G}\) and convection velocity \(\vec{V}_E\) vectors.
-GDI can more easily excite waves when a component \(\vec{G}\cdot\vec{V}_E\) is more positive, with \(\vec{G}\parallel\vec{V}_E\) being most favorable orientation \citep{Keskinen1982a,Makarevich2014c}.
+GDI can more easily excite waves when a component \(\vec{G}\cdot\vec{V}_E\) is more positive, with \(\vec{G}\parallel\vec{V}_E\) being most favorable orientation \citep{Keskinen1982,Makarevich2014c}.
 IMF-controlled changes in \(\vec{V}_E\) can thus potentially lead to changes in \(\vec{G}\cdot\vec{V}_E\), which, in turn, will change GDI growth rate and occurrence of GDI-produced irregularities.

### 2017-02-06 paper2.tex
Formatting corrections made

@@
 The GDI growth rate is generally dependent on the relative directions of the background density gradient vector \(\nabla n_0\) and the convection electric field \(\vec{E}\) or, equivalently, the convection velocity \(\vec{V}_E=\vec{E}\times\vec{B}/B^2\).
-It is widely accepted that the maximum GDI growth rate in the \(F\) region occurs when the density gradient \(\nabla n_0\) is parallel to the convection drift velocity \(\vec{V}_E\) \citep{Keskinen1982a}, while in the \(E\) region the maximum occurs when \(\nabla n_0\parallel \vec{E}\).
+It is widely accepted that the maximum GDI growth rate in the \(F\) region occurs when the density gradient \(\nabla n_0\) is parallel to the convection drift velocity \(\vec{V}_E\) \citep{Keskinen1982}, while in the \(E\) region the maximum occurs when \(\nabla n_0\parallel \vec{E}\).
 An analytic expression for the GDI growth rate in the \(F\) region has been derived in numerous studies, with most of these derivations considering a particular subset of mutual orientations between \(\nabla n_0\), \(\vec{V}_E\), and the wave vector \(\vec{k}\).
 This typically results in the upper limit of the GDI growth rate of \(\gamma=V_E/L\), where \(L\) is a characteristic gradient scale length defined through \(L^{-1}\equiv |\nabla n_0|/n_0\).
-More general vector configurations and growth rate expressions in the \(F\) region have also been considered \citep{Keskinen1982a,Makarevich2014c}, but their implications for studying the asymmetries between leading and trailing edges of large-scale structures have not been investigated.
+More general vector configurations and growth rate expressions in the \(F\) region have also been considered \citep{Keskinen1982,Makarevich2014c}, but their implications for studying the asymmetries between leading and trailing edges of large-scale structures have not been investigated.
 Further, it has been demonstrated recently that the GDI growth rate patterns change dramatically between the \(E\) and \(F\) regions \citep{Makarevich2014c}.
@@
 An enhanced D region may act to attenuate backscatter signals, although this effect is not always obvious \citep{Danskin2002}.
-Nonlinear processes in the \(F\) region have been theoretically shown to alter the dependence on the propagation direction while still showing significant anisotropy similar to that predicted by the linear theory \citep{Keskinen1981b,Keskinen1982a,Keskinen1983}.
+Nonlinear processes in the \(F\) region have been theoretically shown to alter the dependence on the propagation direction while still showing significant anisotropy similar to that predicted by the linear theory \citep{Keskinen1981b,Keskinen1982,Keskinen1983b}.
 The situation in the \(E\) region appears to be somewhat different, with some echoes also exhibiting strong anisotropy \citep[e.g.][]{Ierkic1980,Makarevich2008} and other echoes showing isotropic power variation \citep{Ierkic1980,Koustov2001} presumably because of nonlinear processes.

### 2017-02-06 paper3.tex
Formatting corrections made

@@
 In context of large-scale density structures at high-latitudes drifting perpendicular to their direction of elongation \citep{Makarevitch2004c,Makarevich2015b} such as polar cap patches and sun-aligned arcs, \(\vec{G}\) and \(\vec{V}_E\) are parallel on the trailing edge and antiparallel on the leading edge.
-However, there can be large variations in this \citep{Makarevitch2004c} and, in general, the growth rate changes with the wavevector direction \citep{Keskinen1982a,Keskinen1983,Makarevich2014c}.
+However, there can be large variations in this \citep{Makarevitch2004c} and, in general, the growth rate changes with the wavevector direction \citep{Keskinen1982,Keskinen1983b,Makarevich2014c}.
 More insight can therefore be gained by considering the directions of vector quantities instead of considering a vector configuration that results in the peak growth.
@@
 \label{sec:p3discussion1}
-The fluid linear theory of GDI predicts that an increase in plasma density gradient strength \(G\) will increase irregularity production through GDI \citep{Keskinen1983,Makarevich2014c}.
+The fluid linear theory of GDI predicts that an increase in plasma density gradient strength \(G\) will increase irregularity production through GDI \citep{Keskinen1983b,Makarevich2014c}.
 The current experimental results were generally consistent with this expectation.
@@
 One important new aspect of the current study is the experimental investigation of the directional dependence of the GDI growth rate, as represented by equation  (\ref{eqn:p3gdi}).
-This directional dependence has been previously predicted theoretically \citep{Keskinen1982a,Makarevich2014c,Lamarche2016}, but no experimental evidence is currently available, apart from asymmetry between leading and trailing edges of large-scale density structures.
+This directional dependence has been previously predicted theoretically \citep{Keskinen1982,Makarevich2014c,Lamarche2016}, but no experimental evidence is currently available, apart from asymmetry between leading and trailing edges of large-scale density structures.
 Because the growth rate \(\gamma\) is directly dependent on both the gradient component, \(\uvec{k}\cdot\uvec{b}\times\vec{G}\), and the electric field component, \(\uvec{k}\cdot\vec{E}\), the relative directions of the gradient, electric field, and wavevector are predicted to strongly influence plasma structuring in the ionosphere, equation (\ref{eqn:p3gdi}).

### 2017-02-10 conclusion.tex
Minor corrections

@@
 Polar patches (orange) drift antisunwards across the polar cap following the background plasma convection (red contours).
-Nighttime irregularity observation is enhanced for a positive IMF \(B_y\) component.
+Nighttime irregularity observation is enhanced for a negative IMF \(B_y\) component.
 Production is maximized near the terminator.
@@
 
-One other result presented in Chapter \ref{sec:paper1} was that nighttime occurrence of FAIs was also higher for time intervals with a positive IMF \(B_y\) component, Figure \ref{fig:conclusion}.
+One other result presented in Chapter \ref{sec:paper1} was that nighttime occurrence of FAIs was also higher for time intervals with a negative IMF \(B_y\) component, Figure \ref{fig:conclusion}.
 This could be due to the control that the \(B_y\) component exerts over ionospheric convection patterns, as explained below.

### 2017-04-29 abstract50.tex
FINAL VERSION POST GRAD SCHOOL CORRECTIONS

%% 50 word abstract

Plasma in the Earth's ionosphere is highly irregular on scales ranging
between a few centimeters and hundreds of kilometers. Small-scale
irregularities or plasma waves in the Earth's ionosphere can scatter radio waves resulting in a loss of
signal for navigation and communication networks. The polar region is
particularly susceptible to strong disturbances due to its direct connection with
the Sun's magnetic field and energetic particles.

In this work, factors that
contribute to the production of small-scale plasma irregularities in the polar
F-region ionosphere are investigated.

We present results from two
experimental studies using the Super Dual Auroral Radar Network and the
Resolute Bay Incoherent Scatter Radar facilities and one modeling study
utilizing recently developed general theory of the gradient-drift instability
(GDI). It is demonstrated that small-scale plasma irregularities in the polar
F-region exhibit a previously unreported dependence on the propagation
direction in accordance with directionally-dependent GDI theory, although
large-amplitude perturbations quickly become controlled by more complicated
processes. The production of irregularities is complex, and while ground-based
radars are invaluable tools to study the ionosphere, care must be taken to
interpret results correctly.

Plasma irregularities in the Earth's polar ionosphere can scatter radio waves resulting in disruptions to navigation/communication systems. It is demonstrated that small-scale irregularities depend on the propagation direction in accordance with gradient-drift instability theory, although large-amplitude perturbations are controlled by more complicated processes.

### 2017-04-29 conclusion.tex
FINAL VERSION POST GRAD SCHOOL CORRECTIONS

@@
 Frequent polar patches that propagate from the dayside to the nightside could be the source of this additional density, Figure \ref{fig:conclusion}.
+This is supported by the fact that the annual maximums in nighttime backscatter occurrence occur durring equinox months, which is also when the occurrence of polar patches is the highest \citep{Rodger1996}.
 The requirements for confirming this hypothesis will be detailed further in Section \ref{sec:fw_patches}.
@@
   \includegraphics[width=\textwidth]{conclusion.pdf}
-  \caption[Irregularity production factors]{{\:}Irregularity production factors\\ Summary of the main conclusions in this thesis.
+  \caption[Irregularity production factors]{{\:}Irregularity production factors\vspace{1mm}
+
+  Summary of the main conclusions in this thesis.
 The radar of interest (MCM) is highlighted in pink.
@@
 However, with a significant IMF \(B_y\) component, the orientation of the convection cells rotates such that plasma no longer convects directly antisunward, creating a situation where GDI is more favorable and plasma structures can develop more readily.
+The bias towards greater nighttime occurrence when the \(B_y\) component was negative rather than positive could be due to variations in geomagnetic activity with the orientation of the \(B_y\) component, particularly durring equinox months and under extreme solar wind conditions \citep{Zhao2012}.
 Additionally, a significant IMF \(B_y\) component can contribute to the formation of polar patches by creating convection perturbations that separate high density patches from the daytime reservoir, as described in Section \ref{sec:lit_patches}.
 In this way, the IMF \(B_y\) component may be indirectly responsible for the increased observation of backscatter.
-If patch occurrence increases, it can contribute to the nighttime background density, as described above, which can improve propagation conditions so that FAIs can be observed by
-HF radars.
+If patch occurrence increases, it can contribute to the nighttime background density, as described above, which can improve propagation conditions so that FAIs can be observed by HF radars.
 
@@
 In Chapter \ref{sec:paper1}, experimental evidence was presented that plasma irregularities are observed most often close to the terminator.
-In addition, if a SuperDARN radar itself is positioned close to the terminator, it is observing the density gradient along its boresight, Figure \ref{fig:conclusion}.
+The diurnal and seasonal variations demonstrated in Figure \ref{fig:ut_occ} that follow the terminator also could be due to the Russel-McPherron effect, which predicts annual variations in geomagnetic activity due to changing orientations between the solar equatorial coordinate system and solar magnetospheric coordinate system \citep{Russell1973,Zhao2012}.
+In addition, if a SuperDARN radar itself is positioned close to the terminator, it is observing the density gradient along its boresight, with the gradient in the same direction as the plasma drift, Figure \ref{fig:conclusion}.
 As was demonstrated in Chapter \ref{sec:paper2} and discussed above that if small-scale irregularity growth is directionally dependent, this is also the optimal orientation from which to observe the plasma structures, leading to increased echo occurrence.
@@
 
-\section{Recommendations for Future Research on \(F\) Region Plasma Irregularities}
+\section{Recommendations for Future Research on \textit{F} Region Plasma Irregularities}
 \label{sec:futurework}

### 2017-04-29 introduction.tex
FINAL VERSION POST GRAD SCHOOL CORRECTIONS

@@
 The frequency of sunspots and large outbursts of plasma changes with the 11-year-long solar cycle.
-A large number of sunspots, high occurrence of fast and dense plasma outbursts, and a highly variable magnetic field indicate solar maximum.
+A large number of sunspots, high occurrence of fast and dense plasma outbursts, strong x-ray or extreme ultraviolet (EUV) flux, and a highly variable magnetic field indicate solar maximum.
 Conversely, solar minimum is characterized by a low sunspot number and slow and relatively steady outflow of plasma.
@@
 
-The Sun interacts with the rest of the solar system through the solar wind, a continuous outflow of plasma that carries the Sun's magnetic field with it.
+The Sun interacts with the rest of the solar system through solar radiation and the solar wind, a continuous outflow of plasma that carries the Sun's magnetic field with it.
 This is shown in Figure \ref{fig:ste}.
-The solar wind, represented by the green arrows in Figure \ref{fig:ste}, typically has a density around 10 cm\(^{-3}\) and is traveling away from the Sun at approximately 400 km/s at the Earth's orbit.
+The solar wind, represented by the blue arrows in Figure \ref{fig:ste}, typically has a density around 10 cm\(^{-3}\) and is traveling away from the Sun on the order of 400 km/s at the Earth's orbit.
 However, like the Sun, the solar wind is highly dynamic and these parameters can change drastically over a period of minutes.
 CMEs, for instance, create a substantial increase in density and are often preceded by a shock, a region of hot, compressed plasma caused by the CME moving at supersonic speeds.
-In addition to the outflow of plasma, the solar wind also contains the Sun's magnetic field extended into the outer limits of the solar system, called the interplanetary magnetic field (IMF), dark blue lines in Figure \ref{fig:ste}.
-Because the Sun has a short rotation period (about 33 days), the IMF gets twisted into a spiral, commonly called the Parker Spiral \citep{Parker1958}.
+In addition to the outflow of plasma, the solar wind also contains the Sun's magnetic field extended into the outer limits of the solar system, called the interplanetary magnetic field (IMF), orange lines in Figure \ref{fig:ste}.
+Because the Sun has a short rotation period (about 27 days at the equator), the IMF gets twisted into a spiral, commonly called the Parker Spiral \citep{Parker1958,Kivelson1995}.
 When the IMF reaches the Earth, the magnetic field is typically oriented at a \(45\deg\) inclination relative to the ecliptic, but this is highly variable depending on solar wind conditions.
@@
 	\includegraphics[width=\textwidth,angle=0]{ste.pdf}
-	\caption[Solar-terrestrial environment]{{\:}Solar-terrestrial environment\\ Diagram of the solar-terrestrial environment.
-The solar wind, green arrows, contains both a steady stream of charged particles and the interplanetary magnetic field, dark blue.
-The red region represents the magnetosheath, the shock layer formed due to the supersonic plasma outflow from the Sun.
+	\caption[Solar-terrestrial environment]{{\:}Solar-terrestrial environment\vspace{1mm}
+
+	Diagram of the solar-terrestrial environment.
+The solar wind, blue arrows, contains both a steady stream of charged particles and the interplanetary magnetic field, orange.
+The purple region represents the magnetosheath, the shock layer formed due to the supersonic plasma outflow from the Sun.
+The bow shock is indicated by the green line.
 The magnetopause encloses the magnetosphere dominated by the geomagnetic field and is shown by the black dashed line.
@@
 The IMF interacts with the Earth's own magnetic field, and this interaction creates the magnetosphere.
-Without the IMF, the magnetosphere would be approximately in a dipole shape, but dynamic pressure from the solar wind causes the Sun-facing side to be compressed to 6--10 Earth radii and the opposite side to be stretched into an extended tail (hundreds of Earth radii).
-Because the solar wind is typically a supersonic flow, a bow shock forms on the dayside of the magnetosphere.
-The heated and compressed solar wind plasma created by the bow shock is known as the magnetosheath, the red region in Figure \ref{fig:ste}.
+Without the solar wind, the Earth's magnetosphere would be approximately in a dipole shape, but dynamic pressure from the solar wind causes the Sun-facing side to be compressed to 6--10 Earth radii and the opposite side to be stretched into an extended tail (hundreds of Earth radii).
+Because the solar wind is typically a supersonic flow, a bow shock forms on the dayside of the magnetosphere, green line in Figure \ref{fig:ste}.
+The heated and compressed solar wind plasma created by the bow shock is known as the magnetosheath, the purple region in Figure \ref{fig:ste}.
 The magnetopause, black dashed line, is the actual boundary between solar wind plasma in the magnetosheath and magnetospheric plasma.
 Magnetic field lines that are part of the closed magnetosphere are shown in grey in Figure \ref{fig:ste}.
-However, the magnetosphere is not fully closed.
-At the magnetopause, magnetic reconnection can occur between the IMF and the Earth's magnetic field creating open field lines, light blue in Figure \ref{fig:ste}.
+However, usually the magnetosphere is not fully closed.
+At the magnetopause, magnetic reconnection can occur between the IMF and the Earth's magnetic field creating magnetic field lines that directly connect the Sun to the Earth via the solar wind, commonly referred to as open field lines, light blue in Figure \ref{fig:ste}.
 This reconnection occurs on last closed field line intersecting the magnetopause (highlighted in orange in Figure \ref{fig:ste}).
-Magnetic field lines originating from the Earth are then directly connected to the Sun through the solar wind, which consequently allows highly energized solar particles to enter the magnetosphere.
-Open field lines tend to occur around the Earth's polar caps, making these regions particularly interesting to study due to very complex behavior and coupled interactions between the Earth and the Sun.
+Open field lines allow highly energized solar particles to enter the magnetosphere and tend to occur around the Earth's polar caps, making these regions particularly interesting to study due to very complex behavior and coupled interactions between the Earth and the Sun.
 
@@
 \label{sec:ionosphere}
-The focus of this body of work is the Earth's ionosphere, a region of the upper atmosphere that extends approximately 50--1000 km in altitude where neutral gases have been partially ionized, resulting in free electrons and ions mixed with neutral particles.
+The focus of this work is the Earth's ionosphere, a region of the upper atmosphere that extends approximately 50--1000 km in altitude where neutral gases have been partially ionized, resulting in free electrons and ions mixed with neutral particles.
 Two competing processes occur in the ionosphere to create a peak in the electron density.
 Solar illumination and particle precipitation ionize neutrals, and the density of ions increases as altitude decreases because there are more neutral molecules available to ionize.
-However, below a certain point, the concentration of ions becomes large enough that recombination becomes a substantial factor, reducing the ion density.
-The exact altitude where this density peak occurs varries with time of day, location, season, and solar cycle.
+However, ionizing radiation does not penetrate the atmosphere easily below a certain altitude, and ion recombination becomes a substantial factor, reducing the ion density.
+The exact altitude where this density peak occurs varies with time of day, location, season, and solar cycle.
 Representative model examples of electron density changes with altitude are shown in Figure \ref{fig:densprofile}, found from the International Reference Ionosphere (IRI) model run on January 1, 2007 at \(75\deg\) N, \(0\deg\) E, geographic (the IRI model is discussed further in Section \ref{sec:iri}).
@@
 	\includegraphics[width=\textwidth]{densprofile.pdf}
-	\caption[Ionosphere density profiles]{{\:}Ionosphere density profiles\\ Altitude profiles of electron density throughout the ionosphere found from the IRI model.
-Typical extents of the \(D\), \(E\), and \(F\) regions are identified.}
+	\caption[Ionosphere density profiles]{{\:}Ionosphere density profiles\vspace{1mm}
+
+	Altitude profiles of electron density throughout the ionosphere found from the IRI model.  Typical extents of the \(D\), \(E\), and \(F\) regions are identified.}
 	\label{fig:densprofile}
@@
 
-Above 50 km, the primary neutral particles are atomic oxygen, O, molecular oxygen, O\(_2\), and molecular nitrogen, N\(_2\).
-Common ions at these altitudes are therefore mostly composed of oxygen and nitrogen, O\(^+\), NO\(^+\), O\(_2^+\), N\(_2^+\), but hydrogen, H\(^+\), can contribute as well, particularly at higher altitudes \citep{Kelley2009}.
+Above 50 km, the primary neutral particles are atomic oxygen, O, molecular oxygen, O\(_2\), and molecular nitrogen, N\(_2\), with N\(_2\) dominating at low altitudes but O becoming predominate above 250 km \citep{Kelley2009}.
+Common ions at these altitudes are therefore mostly composed of oxygen and nitrogen, O\(^+\), NO\(^+\), and O\(_2^+\), but hydrogen, H\(^+\), can contribute as well, particularly at higher altitudes \citep{Kelley2009}.
 In the ionosphere, neutral densities are typically several orders of magnitude higher than ion densities, however even a small number of charged particles changes the behavior of the region dramatically.
@@
 In Figure \ref{fig:densprofile}, the daytime profile is shown with the solid line next to a nighttime profile shown with a dashed line.
-Note that night electron densities are lower at all altitudes due to the lack of photoionization during the night time.
+Note that night electron densities are generally lower at all altitudes due to the lack of photoionization during the night time, however densities can be high in the cases of aurora or a sporadic \(E\) layer.
 In addition, the \(F\)-region peak moves to higher altitudes at night.
@@
 The main issue is whether the large-scale motion of a particle is controlled more by the magnetic field (the particle is magnetized) or by collisions with other particles (the particle is collisional).
-This can be determined by comparing the collision frequency, \(\nu_\alpha\), to the gyrofrequency, \(\Omega_\alpha\), of a particular species, where \(\alpha\) refers to either electrons, \(e\), or ions, \(i\).
+This can be determined by comparing the collision frequency, \(\nu_\alpha\), to the gyrofrequency, \(\Omega_\alpha\), of a particular species, where \(\alpha\) refers to either electrons, \(e\), or the various ion species, \(i\).
 If \(\Omega_\alpha \gg \nu_\alpha\), the particle completes many gyrations between each collision, so its motion is determined mostly by the magnetic field and it is considered to be magnetized.
@@
 The gyrofrequency is dependent on the mass (\(m_\alpha\)) and charge (\(q_\alpha\)) of the particle, and the strength of the magnetic field (\(B\)), \(\Omega_\alpha = q_\alpha B/m_\alpha\), so it is roughly constant throughout the ionosphere because the magnetic field strength does not change much throughout these altitudes.
-The collision frequency depends on the density of a species relative to the density of all other species in the plasma and can be calculated from a series of standard expressions \citep{Schunk1980,Schunk2009}.
+Collision frequencies depend on the relative densities of different species within the plasma and can be calculated from a series of standard expressions \citep{Schunk1980,Schunk2009}.
 In the \(E\) region, \(\Omega_e \gg \nu_e\), so electrons are magnetized.
@@
 Polar holes are plasma density depletions that typically occur in the \(F\) region slightly poleward of the auroral oval \citep{Brinton1978}.
-Sun-aligned arcs are large density forms that stretch across the polar cap along the Sun-Earth line.
+Sun-aligned arcs are large auroral forms that stretch across the polar cap along the Sun-Earth line.
 They are very narrow and often characterized by large and complex velocity shears on either side of the arc \citep{Valladares1991}.
@@
 	\includegraphics[height=0.75\textheight]{irregularity.pdf}
-	\caption[Plasma irregularities in the polar ionosphere]{{\:}Plasma irregularities in the polar ionosphere\\ Plasma density irregularities occur in the polar cap at different scale sizes.
+	\caption[Plasma irregularities in the polar ionosphere]{{\:}Plasma irregularities in the polar ionosphere\vspace{1mm}
+
+	Plasma density irregularities occur in the polar cap at different scale sizes.
 Polar patches, orange in Figures \ref{fig:polarirreg}a and \ref{fig:polarirreg}b, move with the background ionospheric convection, indicated by red arrows.
@@
 These are also referred to as scintillation-causing structures and are responsible for many of the adverse space weather effects that are observed on Earth.
-One important example of a space weather effect is when the ionosphere changes the phase and amplitude of the original radio signal, called radio scintillation, which affects the ability of satellites to communicate with ground receivers.
+One important example of a space weather effect is when ionospheric irregularities cause fluctuations of the phase and amplitude of the original radio signal, called radio scintillation, which affects the ability of satellites to communicate with ground receivers.
 This can cause communication blackouts or introduce large errors into GPS/GNSS (Global Positioning System/Global Navigation Satellite System) calculations, affecting navigation.
@@
 The radar transmits a radio wave, which can scatter off plasma structures in the ionosphere, and then be detected by the receiving antennas.
-The time difference between when the signal was transmitted and when it was detected and the power and phase of the returned signal can then be used to determine characteristics of plasma irregularities.
+The time difference between when the signal was transmitted and when it was detected determines the location of plasma irregularities in the ionosphere and the power, phase, Doppler shift, and Doppler width of the returned signal determines their characteristics.
 
-There are three measurements typically made by CSR systems: backscatter power, line-of-sight (LoS) velocity, and spectral width.
+There are three measurements typically made by CSR systems: backscatter power, line-of-sight (LoS) velocity (derived from the Doppler shift), and spectral width.
 Backscatter power is the power of the returned signal that the radar's receivers detect.
-Typically, a threshold is selected for the minimum power acceptable for a return to be considered an actual signal instead of background noise.
+Typically, a threshold is selected for the minimum signal-to-noise ratio (SNR) acceptable for a return to be considered an actual signal instead of background noise.
 LoS velocity is the component of the total plasma irregularity velocity that is measured along the transmitted signal direction.
-Because radars measure velocity through the Doppler shift of backscatter, only a component of the velocity vector can be measured.
-Spectral width represents the width of the Doppler power spectrum and can be used to infer other characteristics of irregularities, such as their lifetime \citep{Greenwald1985}.
+Because radars measure velocity through the Doppler shift of backscatter, only a component of the velocity vector can be measured \citep{Greenwald1985}.
+Spectral width represents the broadening of the returned Doppler power spectrum and can be used to infer other characteristics of irregularities.
+For instance, a large spectral width can indicate irregularities that have a short lifetime, possibly due to Bohm diffusion, or fluctuations in the plasma drift velocity due to turbulence \citep{Ponomarenko2006}.
 
-CSR systems can operate by transmitting pulses instead of continuously.
+Radars can operate by transmitting pulses instead of a continuous signal.
 This allows the distance between the radar and the target to be found (based on the time difference between when the pulse is transmitted and received).
-The size of the range gate is determined by the length of a radar pulse,
+The lower limit of the range gate size is determined by the length of a radar pulse,
 \begin{equation}
@@
 When the radar transmits continuously, backscatter will be received, but because it is impossible to know when that signal was transmitted, the difference in time between transmission and receiving cannot be used to calculate how far away the backscatter volume is.
-Conversely, if the radar only transmits a pulse of a certain length of a certain length and then ``listens'' for its return, the time between transmission and return is known and the distance the pulse traveled can be calculated and therefore the location of the backscattering volume can be found \citep{Farley1972,Greenwald1983}.
+Conversely, if the radar only transmits a pulse of a certain length and then ``listens'' for its return, the time between transmission and return is known and the distance the pulse traveled can be calculated and therefore the location of the backscattering volume can be found \citep{Farley1972,Greenwald1983}.
 However, using this simple single-pulse method, another pulse cannot be transmitted until the first pulse returns, limiting the temporal resolution that can be achieved.
@@
 
-A short time between pulse returns is necessary to allow Doppler velocities to be calculated, particularly at long ranges.
-The maximum Doppler shifted frequency that can be observed by a radar is the Nyquist frequency, \(f_n\), which is equivalent to half the sampling frequency, \(f_s\),
+A short time between pulse returns is necessary to allow Doppler velocities to be calculated.
+The maximum Doppler shifted frequency, \(\delta f\), that can be observed by a radar is the Nyquist frequency, \(f_n\), which is half the sampling frequency, \(f_s\),
 \begin{equation}	
 	\label{eqn:nyquist}
-	f_n = \frac{f_s}{2}.
+	\delta f < f_n = \frac{f_s}{2}.
 \end{equation}
-Radars operating in a single-pulse mode must have a very low sampling frequency, as described above, particularly at long ranges where the pulse must travel a long distance.
-This severely limits the Doppler shift that can be measured, which places an upper limit on the LoS velocity that can be calculated with this technique.
-For an ionospheric radar observing a structure 2000 km away, it would take about 13 ms for a signal traveling at the speed of light to travel to the structure and back to the radar, which corresponds to a sampling frequency of \(f_s = 75\) Hz.
+Radars operating in a single-pulse mode must have a very low sampling frequency, as described above, particularly when the extent over which echoes are expected is large and the pulse must travel a long distance.
+This results in a low Nyquist frequency which severely limits the Doppler shift that can be measured and places an upper limit on the LoS velocity that can be calculated with this technique.
+For an ionospheric radar observing targets a maximum of 2000 km away, it would take about 13 ms for a signal traveling at the speed of light to travel the extent of the target area and back to the radar, which corresponds to a sampling frequency of \(f_s = 75\) Hz.
 The Nyquist frequency is then \(f_n = 37.5\) Hz, which corresponds to a maximum measurable Doppler velocity of \(\sim550\) m/s.
@@
 The ACF is used to find the spectral characteristics of the backscatter, from which the LoS velocity, spectral width, and backscatter power can be obtained.
-Overall, multi-pulse techniques are generally considered far more appropriate for ionospheric studies than single-pulse techniques \citep{Farley1972,Greenwald1983,Greenwald1985,Barthes1998,Ponomarenko2006}.
+Overall, multi-pulse techniques are generally considered far more appropriate for ionospheric studies than single-pulse techniques \citep{Farley1972,Greenwald1983,Greenwald1985,Sulzer1993,Barthes1998,Ponomarenko2006}.
 
 The first observations of the ionosphere by CSRs were made in the 1930s \citep{Eckersley1937,Harang1938}.
-Over the next several decades, numerous ground-based radars have been used to advance our understanding of the ionosphere  \citep{Hultqvist1964,Leadabrand1965,Unwin1972,Sahr1996}.
+Over the next several decades, numerous ground-based radars were used to advance our understanding of the ionosphere  \citep{Hultqvist1964,Leadabrand1965,Unwin1972,Sahr1996}.
 In the 1970s and 1980s, the Scandanavian Twin Auroral Radar Experiment (STARE) produced the first continuous large data-set, similar to how modern CSR systems are run.
@@
 	\includegraphics[width=\textwidth,angle=180]{SuperDARNmap.pdf}
-	\caption[SuperDARN map]{{\:}SuperDARN map\\ Fields-of-view of all operational SuperDARN radars in both the northern (a) and southern (b) hemispheres.
+	\caption[SuperDARN map]{{\:}SuperDARN map\vspace{1mm}
+
+	Fields-of-view of all operational SuperDARN radars in both the northern (a) and southern (b) hemispheres.
 Polar-latitude radars are shown shaded in blue.
@@
 
-The most important condition for observation of FAIs with CSRs is the so-called perpendicularity condition, according to which the radar beam has to reach the scattering volume perpendicular to the local magnetic field direction.
-For this reason, it is necessary to use HF radars to study ionospheric structures in the polar cap because the magnetic field is close to vertical and the beam must be refracted through a dense ionosphere to reach perpendicularity with the magnetic field.
-VHF radar beams are not refracted enough in the polar cap and the beam will pass through the ionosphere without FAIs producing backscatter.
+The most important condition for the observation of FAIs with CSR systems with colocated transmitters and receivers is the so-called perpendicularity condition, according to which the radar beam has to reach the scattering volume perpendicular to the local magnetic field direction.
+At mid-latitudes, this is plausable through straight line propigation because the magnetic field is angled relative to the Earth's surface, so a radar with a suitably chosen elevation angle can produce a beam perpendicular to the local magnetic field.
+However, at the polar cap, the magnetic field is close to vertical so it is necessary to use HF radars to study ionospheric structures because the beam must be refracted through a dense ionosphere to reach perpendicularity with the magnetic field.
+VHF radar beams are not refracted enough in the polar cap and backscatter produced by FAIs will not return to the radar.
 SuperDARN radars operate nominally between 8--20 MHz, which, according to the Bragg scatter condition, refers to decameter-scale plasma waves.
 Each radar consists of 16 independent antennas that transmit a radio wave and receive backscatter from the ionosphere.
-Beams are electronically steerable and most radars have between 16 and 24 beams in normal operation mode, each beam being \(3.25\deg\) in azimuth.
+Beams are electronically steerable and most radars have between 16 and 24 beams in normal operation mode, each beam spanning \(3.25\deg\) in azimuth.
 Each beam consists of 75--100 range gates, which are generally either 15 km, 30 km, or 45 km in length \citep{Chisham2007}.
-SuperDARN radars were originally designed to used a 7-pulse ACF \citep{Farley1972,Greenwald1983,Greenwald1985}, however since 2011, most radars have begun using an 8-pulse sequence.
+SuperDARN radars were originally designed to use a 7-pulse ACF \citep{Farley1972,Greenwald1983,Greenwald1985}, however since 2011, most radars have begun using an 8-pulse sequence.
 The multi-pulse increment, \(\tau\), is 2400 \(\mu\)s, which corresponds to a Nyquist frequency of about 200 Hz, small enough to measure plasma drift velocities up to 3000 m/s.
@@
 This creates a vast database of high-quality measurements of small-scale FAIs, which is very useful for other studies on irregularity occurrence and plasma structuring.
-In the studies presented here, data were primarily used from the SuperDARN radars located at Rankin Inlet, Canada (RKN) and McMurdo Station, Antarctica (MCM).
+Data used in the studies presented here came primarily from the SuperDARN radars located at Rankin Inlet, Canada (RKN) and McMurdo Station, Antarctica (MCM).
 These two radars have been outlined in pink in Figure \ref{fig:superdarnmap}.
@@
 This can be attributed to the effect that the much slower ions have on free electrons in the plasma through ion acoustic waves \citep{Bowles1958}.
-The practical results of this is that ISRs can use radio waves of a much longer wavelength than originally anticipated, on the order of \(\sim\)1 m.
+The practical results of this is that ISRs of much lower sensitivity can be used to study the ionosphere.
+Very high sensitivity is only necessary for wavelengths less than the Debye length when true Thompson scatter occurs.
 
@@
 This is done using the plasma frequency measured from either the plasma line of the ISR spectra during summer daytime period or ionosonde measurements for all other times \citep{Bahcivan2010,Themens2014}.
+The plasma line refers to a discrete line in an ISR spectra that occurs at a Doppler shift approximately equal to the plasma frequency, from which the plasma density can be derived \citep[e.g.][Figure 1]{Evans1969}.
 
@@
 	\includegraphics[width=\textwidth]{ISRmap.pdf}
-	\caption[ISR map]{{\:}ISR map\\ FoVs of all ISRs in the northern polar region.
+	\caption[ISR map]{{\:}ISR map\vspace{1mm}
+
+	FoVs of all ISRs in the northern polar region.
 The individual radars shown are identified by their commonly used acronyms: Poker Flat (PFISR), Resolute Bay North (RISR-N), Resolute Bay Canada (RISR-C), Sondrestrom (SFJ), European Incoherent Scatter Scientific Association (EISCAT), and EISCAT Svalbard (ESR).
@@
 One of the most important advantages of ISR systems over CSRs is the ability to directly measure electron density in the ionosphere.
-However, because ISRs are much scarcer and most are only typically run a few days out of every month, data are not as widely available.
+However, because ISRs are much scarcer and typically only run a few days each month, data are not as widely available.
 In addition, ISRs do not actually observe small-scale coherent plasma structures like the SuperDARN array does.
@@
 	\item Plasma production by cusp particle precipitation \citep{Rodger1994,Millward1999}
-	\item Plasma flow jet channels that cut continuous streaches of plasma into segments 	\citep{Valladares1998}
-	\item Alfven wave coupling \citep{Prikryl1999}.
+	\item Plasma flow jet channels that cut continuous stretches of plasma into segments 	\citep{Valladares1998}
+	\item Alfv\'{e}n wave coupling \citep{Prikryl1999}.
 \end{enumerate}
@@
 In general, magnetic field lines in the magnetosphere can either be closed (grey lines in Figure \ref{fig:magnetosphere}), meaning they connect only the regions near the north and south pole of the Earth, or open (light blue lines in Figure \ref{fig:magnetosphere}), when one of the ends of the field lines originates from the Sun such that the field line is connected directly to the IMF.
-IMF lines that are not connected to the Earth are shown in dark blue in Figure \ref{fig:magnetosphere}.
+IMF lines that are not connected to the Earth are shown in orange in Figure \ref{fig:magnetosphere}.
 The polar cap boundary is generally defined as the border between open and closed magnetic field lines, shown by the green line in Figure \ref{fig:magnetosphere}.
+
+\begin{figure}
+	\centering
+	\includegraphics[width=\textwidth,angle=0]{magnetosphere.pdf}
+	\caption[Magnetic reconnection footprint in the ionosphere]{{\:}Magnetic reconnection footprint in the ionosphere\vspace{1mm}
+
+	Diagram of the interactions between the IMF, magnetosphere, and polar cap ionosphere.
+Closed field lines in the Earth's magnetosphere are shown in grey, while open field lines are light blue.
+The polar cap open-closed boundary is in green.
+Magnetic field lines that are part of the IMF are shown in orange.
+Magnetic reconnection occurs along the X-line (red) on the dayside of the magnetopause (dashed line).
+The X-line can be mapped back along magnetic field lines (yellow) to the ionosphere, where its projection is the merging gap (pink dotted line).
+During reconnection events, the open-closed boundary expands equatorwards as more open field lines are formed and there is plasma flow across the merging gap.
+Figure adapted from \citet{Cowley1991}.}
+	\label{fig:magnetosphere}
+\end{figure}
 
@@
 \end{equation}
-In the polar cap at ionospheric altitudes, the magnetic field strength is relatively constant at around \(5\times 10^{-5}\) T, so any changes in magnetic flux must be due to changes in the polar cap area, \(A_{pc}\) due to the polar cap boundary expanding or contracting \citep{Lockwood1992a}.
+In the polar cap at ionospheric altitudes, the plasma is incompressible so the magnetic field line density is relatively constant and any changes in magnetic flux must be due to changes in the polar cap area, \(A_{pc}\) due to the polar cap boundary expanding or contracting \citep{Lockwood1992a}.
 The polar cap boundary changes in response to magnetic reconnection between the magnetosphere and the IMF on both the day and night sides changing the amount of open magnetic flux.
-
-\begin{figure}
-	\centering
-	\includegraphics[width=\textwidth,angle=0]{magnetosphere.pdf}
-	\caption[Magnetic reconnection footprint in the ionosphere]{{\:}Magnetic reconnection footprint in the ionosphere\\ Diagram of the interactions between the IMF, magnetosphere, and polar cap ionosphere.
-Closed field lines in the Earth's magnetosphere are shown in grey, while open field lines are light blue.
-The polar cap open-closed boundary is in green.
-Magnetic field lines that are part of the IMF are shown in dark blue.
-Magnetic reconnection occurs along the X-line (red) on the dayside of the magnetopause (dashed line).
-The X-line can be mapped back along magnetic field lines (orange) to the ionosphere, where its projection is the merging gap (pink dotted line).
-During reconnection events, the open-closed boundary expands equatorwards as more open field lines are formed and there is plasma flow across the merging gap.
-Figure adapted from \citet{Cowley1991}.}
-	\label{fig:magnetosphere}
-\end{figure}
 
@@
 In all panels, the dayside is the top of the figure and the green line shows the open-closed polar cap boundary, similar to Figure \ref{fig:magnetosphere}.
-Plasma flow contours are shown in black and the orange region represents high density from the dayside reservoir.
+Plasma flow contours are shown in black and the orange region represents high density plasma from the dayside reservoir.
 The reconnection burst starts at Figure \ref{fig:patch_formation}a and the increasing amount of open magnetic flux moves the dayside open-closed boundary equatorwards, as described above \citep{Cowley1991}.
@@
 
+\begin{figure}
+	\centering
+	\includegraphics[width=\textwidth,angle=0]{patch_formation.pdf}
+	\caption[Polar patch formation]{{\:}Polar patch formation\vspace{1mm}
+
+	Polar patch formation through transient magnetopause reconnection.
+The open-closed boundary is the green line.
+Black contours represent plasma flow.
+Dense daytime plasma is shown by the orange regions.
+The burst of reconnection is occurring in Figures \ref{fig:patch_formation}a--\ref{fig:patch_formation}c, but stops in Figures \ref{fig:patch_formation}d--\ref{fig:patch_formation}f.
+Figure adapted from \citet{Cowley1991}.}
+	\label{fig:patch_formation}
+\end{figure}
+
 At Figure \ref{fig:patch_formation}d, the burst of reconnection stops and the open-closed boundary begins to relax polewards \citep{Cowley1991}.
-As it moves, it pulls the blob of dense plasma within the polar cap with it, separating it from the reservoir on the dayside \citep{Lockwood1992b}.
+As it moves, continuing convection begins to pull from lower-density plasma instead of the high-density dayside reservoir, separating the dense blob in the polar cap \citep{Lockwood1992b}.
 Before the convection flow stops completely, Figure \ref{fig:patch_formation}e, the blob of dense plasma ``pinches off'' from the dayside plasma to become an independent patch \citep{Lockwood1992b}.
@@
 In this way, a reconnection burst can excite plasma flow before the convection from the previous burst has stopped, and a series of overlapping reconnection events like this can create continuous plasma flow through the polar cap \citep{Cowley1991}.
-
-\begin{figure}
-	\centering
-	\includegraphics[width=\textwidth,angle=0]{patch_formation.pdf}
-	\caption[Polar patch formation]{{\:}Polar patch formation\\ Polar patch formation through transient magnetopause reconnection.
-The open-closed boundary is the green line.
-Black contours represent plasma flow.
-Dense daytime plasma is shown by the orange regions.
-The burst of reconnection is occurring in Figures \ref{fig:patch_formation}a--\ref{fig:patch_formation}c, but stops in Figures \ref{fig:patch_formation}d--\ref{fig:patch_formation}f.
-Figure adapted from \citet{Cowley1991}.}
-	\label{fig:patch_formation}
-\end{figure}
 
@@
 GDI is operational when high-density perturbations in the plasma move to regions of lower background density and low-density perturbations in the plasma move to regions of higher background density, such that the wave amplitude grows relative to background conditions.
-As plasma drifts and carries irregularities with it, electrons and ions have slightly different drifts and create a perturbed electric field.
-This perturbed electric field combined with the background magnetic field causes the density perturbations to drift in the \(\vec{E}\times\vec{B}\) direction.
+As plasma drifts and carries irregularities with it, electrons and ions have slightly different velocities and create a perturbed electric field.
+This perturbed electric field \(\vec{E}'\) combined with the background magnetic field causes the density perturbations to drift in the \(\vec{E}'\times\vec{B}\) direction.
 If the drift causes high- and low-density perturbations to move as described above, the instability is operational and the perturbations grow.
 If instead high-density perturbations drift into regions of higher background density and low-density perturbations drift into regions of lower background density, the wave is damped.
+Because GDI is considered the primary structuring mechanism at \(F\)-region altitudes, it will be the focus of this work.
+Other instabilities, notably FBI and CCI, have been briefly described previously, and details can be found in the accompanying references \citep{Farley1963,Buneman1963,Ossakow1979}.
 
@@
 	F_\parallel = i k_\parallel^2 + \vec{G}\cdot\vec{k}_\parallel \\
-	S_\alpha = \frac{1}{1+D_\alpha^2} \quad\quad\quad\quad\quad\quad&\quad
+	S_\alpha = \frac{1}{1+r_\alpha^2} \quad\quad\quad\quad\quad\quad&\quad
 	D_\alpha = -\frac{i}{\Omega_\alpha}(\omega-\vec{k}\cdot\vec{V}_{\alpha 0})+r_\alpha
@@
 
-For the purposes of this thesis, the cold plasma limit expression in Table \ref{tab:gdi_exps} is most relevant and can be rewritten using the definition of \(b\) \citep{Makarevich2014c,Makarevich2016a}
+Most of the quantities defined above are simply for mathematic convenience, however \(\psi\) is also known as the anisotropy parameter and \(R\) is the ratio of the Hall conductivity, \(\sigma_H\), to the Pedersen conductivity, \(\sigma_P\), both of which are discussed further in Chapter \ref{sec:paper2}.
+Additionally, \(\vec{V}_d\) is known as the differential plasma velocity and is the difference between the background electron and ion drift.
+Furthermore, the standard isothermal ion-acoustic speed, \(C_s\), is used in the expressions in Table \ref{tab:gdi_exps}.
+For the purposes of this thesis, the cold plasma limit is most relevant and can be rewritten using the definition of \(b\) \citep{Makarevich2014c,Makarevich2016a}
 \begin{equation}
@@
 	\centering
+	\caption[Irregularity growth rates]{{\:}Irregularity growth rates\vspace{1mm}
+
+	Irregularity growth rates in the ionosphere for three different limiting cases \citep{Makarevich2016a}.}
+	\vspace{5pt}
 	\renewcommand{\arraystretch}{1.5}
@@
 	GDI/FBI in \(E\) Region & \(\gamma = \frac{1}{1+\hat{\psi}}\left[\frac{\hat{\psi}}{\nu_i}\left(\left(\frac{\vec{V}_d\cdot \vec{k}}{1+\hat{\psi}}\right)^2-C_s^2 k_\perp^2\right)+\frac{br_i\vec{V}_d\cdot \vec{k}}{1+\hat{\psi}}\right]\) \\
-	GDI/CCI in \(F\) Region & \(\gamma = -\frac{b}{\psi+y^2}\left[\frac{\psi\vec{k}\cdot\vec{E}_{0\perp}+E_{0\parallel}k_\parallel}{B}\right]+C\left[r_ek_\perp^2-\frac{k_\parallel^2}{r_i}\left(1+\frac{r_i^2}{\psi+y^2}\right)\right]\) \\
+	GDI/CCI in \(F\) Region & \(\gamma = -\frac{b}{\psi+y^2}\left[\frac{\psi\vec{k}\cdot\vec{E}_{0\perp}+E_{0\parallel}k_\parallel}{B}\right]+\frac{C_s^2}{\Omega_i}\left[r_ek_\perp^2-\frac{k_\parallel^2}{r_i}\left(1+\frac{r_i^2}{\psi+y^2}\right)\right]\) \\
 	Cold Plasma in \(F\) Region & \(\gamma = \frac{b}{1+\psi}\left(R\vec{V}_E-\frac{\vec{E}_{0\perp}}{B}\right)\cdot\vec{k}\) \\
 	\end{tabular}
-	\caption[Irregularity growth rates]{{\:}Irregularity growth rates\\ Irregularity growth rates in the ionosphere for three different limiting cases \citep{Makarevich2016a}.}
 	\label{tab:gdi_exps}
@@
 These plasma flows are pulsed in the region of the polar cap boundary \citep{Lockwood1993a,Lockwood1993b}, as predicted by \citet{Cowley1991}.
-Observations made by an all sky imaging photometer (ASIP) in Svalbard in 1984 were fully consistent with patch formation through transient magnetopause reconnection \citep{Carlson1996,Carlson2002}.
+Observations made by an all sky imaging photometer (ASIP) in Svalbard in 1984 were fully consistent with patch formation through transient magnetopause reconnection \citep{Carlson1996,Carlson2002} and periods where polar patches developed on the dayside have been found to correspond to flux transfer events (FTE), or periods of ``bursty'' magnetic reconnection \citep{Rodger1996,Walsh2014}.
 \citet{Carlson2004} identified five signatures of transient magnetic reconnection and then examined a patch formation event for these signatures using the EISCAT Svalbard radar.
@@
 Furthermore, \citet{Carlson2006} identified and tracked a series of patches directly from the subauroral plasma reservoir and showed that the boundary moves equatorwards before relaxing poleward, one of the main predictions of the transient magnetopause reconnection mechanism \citep{Lockwood1992b}.
+Similar efforts using TEC measurements from GPS receivers have confirmed that patches form from dense dayside plasma convecting into the polar cap and ``pinch off'' due to bursty reconnection events allowing low density plasma to isolate them \citep{Zhang2013,Walsh2014}.
 
@@
 Backscatter power from HF radars also tended to be greater on the trailing edges of both moving polar patches \citep{Milan2002b} and sun-aligned arcs identified from ASIP data \citep{Koustov2012}, which is generally attributed to plasma being unstable to GDI on the trailing edge but stable on the leading edge of such large and moving density structures.
-Studies of patches using all-sky airglow imagers (ASI) found that the density gradient on the leading edge of polar patches tends to be 2--3 times steeper than that on the trailing edge \citep{Hosokawa2016}.
+Contradicting the results of \citet{Weber1984}, studies of patches using all-sky airglow imagers (ASI) found that the density gradient on the leading edge of polar patches tends to be 2--3 times steeper than that on the trailing edge \citep{Hosokawa2016}.
 Additionally, large finger-like structures on the trailing edge of patches have been identified by \citet{Hosokawa2016} that were tens to hundreds of kilometers in size and agreed with predictions made by GDI simulations \citep{Gondarenko2004b}.
@@
 
-The first objective of this body of work is to evaluate the extent of solar control over irregularity production in the polar \(F\) region.
-The occurrence of radar backscatter is considered in the context of both solar illumination (considered direct solar control) and IMF factors (considered indirect solar control).
+The first objective of this body of work is to evaluate the extent of solar control over irregularity occurrence in the polar \(F\) region.
+Although radar backscatter is the lower limit to plasma irregularity occurrence, examining the occurrence of radar backscatter in the context of both solar illumination (considered direct solar control) and IMF factors (considered indirect solar control) reveals physically relevant trends.
 This gives a global or macrophysical view of where and when plasma irregularities are most likely to occur.
@@
 The second objective is to investigate asymmetry in the GDI growth rate surrounding large-scale density structures.
-A model is used to identify where plasma irregularities occur surrounding a polar patch and how they can be observed by a ground-based HF radar.
+A model is used to identify where the growth rate is high and hense where plasma irregularities are likely to occur surrounding a polar patch and how they can be observed by a ground-based HF radar.
 The model considers a large, elongated polar patch drifting in an arbitrary direction relative to a radar and assumes irregularity production is directly dependent on the linear GDI growth rate.
@@
 Chapter \ref{sec:paper3} presents an experimental investigation of the directional dependence of small-scale plasma structures by considering simultaneous measurements of electric field and density gradients from an AMISR  system and FAIs observed by SuperDARN in the context of predictions made by linear GDI theory and the model results presented in Chapter \ref{sec:paper2}.
-The same results have been submitted for publication \citep{Lamarche2017}.
+The same results have been recently published \citep{Lamarche2017}.
 Chapter \ref{sec:conclusion} presents a summary of the most significant results of this body of work, their implications for interpretation of radar observations of plasma structures in the polar cap and suggestions for future research involving coherent and incoherent scatter radar facilities in the polar region.

### 2017-04-29 paper1.tex
FINAL VERSION POST GRAD SCHOOL CORRECTIONS

@@
 
-\section*{Abstract}
+\section{Abstract}
 The role of solar wind and illumination in production of small-scale \(F\) region plasma irregularities is investigated using a 4-year data set collected by the Super Dual Auroral Radar Network (SuperDARN) facility at the McMurdo station, Antarctica (MCM).
@@
 The nighttime occurrence, taken separately, was demonstrated to be closely related to SZA-controlled \(F\) region density  variations \citep{Kane2012}, but no similarly clear relationship has been demonstrated for datasets including daytime points.
-One possible reason is the competing effects of the electric filed, particularly in the noon sector near the cusp region \citep{Ghezelbash2014a,Ghezelbash2014b}.
+One possible reason is the competing effects of the electric field, particularly in the noon sector near the cusp region \citep{Ghezelbash2014a,Ghezelbash2014b}.
 
@@
 	\includegraphics[height=\textwidth,angle=90]{p1f1.pdf}
-	\caption[Monthly average echo occurrence and electron density]{{\:}Monthly average echo occurrence and electron density\\ Monthly average plots of (a--d) echo occurrence and (e--h) peak electron density for 2010.
+	\caption[Monthly average echo occurrence and electron density]{{\:}Monthly average echo occurrence and electron density\vspace{1mm}
+
+	Monthly average plots of (a--d) echo occurrence and (e--h) peak electron density for 2010.
 Four months are selected (one for each season) to show seasonal variations.
@@
 These observations refer to the region just equatorward of the polar cap near the noon sector (looking through the polar cap) and thus may represent a radar signature of the dayside cusp \citep{Yeoman2002}.
-In the following analysis, we focus on the maid \(F\)-region band between 500 km and 1500 km which refers to observations in the polar cap.
+In the following analysis, we focus on the made \(F\)-region band between 500 km and 1500 km which refers to observations in the polar cap.
 
@@
 Further, similar variations were also observed for many individual days.
-IN summer months, the variation was largely flat, Figure \ref{fig:month_avg_occ}d, which can be also represented by a cosine variation shifted from zero with a small amplitude.
+In summer months, the variation was largely flat, Figure \ref{fig:month_avg_occ}d, which can be also represented by a cosine variation shifted from zero with a small amplitude.
 To further characterize this diurnal variation, a cosine function of the form
@@
 Rays separated by \(1\deg\) in elevation were found under the assumption of ordinary mode propagation.
-Locations were founds where the ray was nearly perpendicular (within \(1\deg\)) to the local magnetic field based on the International Geomagnetic Reference Field 12 model \citep{Thebault2015}.
+Locations were found where the ray was nearly perpendicular (within \(1\deg\)) to the local magnetic field based on the International Geomagnetic Reference Field 12 model \citep{Thebault2015}.
 
 The range locations where orthogonality condition was expected are plotted in Figures \ref{fig:month_avg_occ}e--\ref{fig:month_avg_occ}h by six lines corresponding to six different options of the IRI model described above.
-The red (grey) kines refer to three URSI (CCIR) options.
+The red (grey) lines refer to three URSI (CCIR) options.
 These lines are seen between the solid and bottom dashed white lines, indicating a reasonable agreement between the expected and observed backscatter locations.
@@
 	\includegraphics[height=\textwidth,angle=90]{p1f2.pdf}
-	\caption[Trends in high-occurrence band]{{\:}Trends in high-occurrence band\\ The variations in the echo band (a) amplitude of the fitted cosine function \(A\), (b) time shift or time of day of the peak \(B\), (c) average position of the band or the vertical shift of the cosine function from zero \(C\), and (d) width (BW).
+	\caption[Trends in high-occurrence band]{{\:}Trends in high-occurrence band\vspace{1mm}
+
+	The variations in the echo band (a) amplitude of the fitted cosine function \(A\), (b) time shift or time of day of the peak \(B\), (c) average position of the band or the vertical shift of the cosine function from zero \(C\), and (d) width (BW).
 Different lines are for different years with the legend given in Figure \ref{fig:year_line}a.
@@
 Similarly, the average echo occurrence is shown in Figures \ref{fig:year_color}e--\ref{fig:year_color}h.
-Figures \ref{fig:year_color}a--\ref{fig:year_color}d also present the fitted band position R(t) as line contours using the same colour scheme.
+Figures \ref{fig:year_color}a--\ref{fig:year_color}d also present the fitted band position R(t) as line contours using the same color scheme.
 
@@
 Outside the gaps, the \(R(t)\) contours match well the underlying color pattern.
-INside the gaps, they represent extrapolation of the same pattern.
-This information is alos useful ot analyze, particularly considering that the fitted errors for the months that contibute the most to the gaps were less tan 20\%, Figures \ref{fig:month_avg_occ}b and \ref{fig:year_line}.
+Inside the gaps, they represent extrapolation of the same pattern.
+This information is also useful to analyze, particularly considering that the fitted errors for the months that contribute the most to the gaps were less than 20\%, Figures \ref{fig:month_avg_occ}b and \ref{fig:year_line}.
 
@@
 	\includegraphics[height=\textwidth,angle=90]{p1f3.pdf}
-	\caption[Solar cycle trends of high-occurrence band]{{\:}Solar cycle trends of high-occurrence band\\ Contour plots of the (a--d) echo band position \(r(t)\) and (e)--(h) occurrence versus UT and season.
+	\caption[Solar cycle trends of high-occurrence band]{{\:}Solar cycle trends of high-occurrence band\vspace{1mm}
+
+	Contour plots of the (a--d) echo band position \(r(t)\) and (e)--(h) occurrence versus UT and season.
 Each row shows the data for a different year as given in the top right corner.
@@
 	\includegraphics[width=\textwidth]{p1f4.pdf}
-	\caption[Band position and occurrence versus SZA and \(N_mF_2\)]{{\:}Band position and occurrence versus SZA and \(N_mF_2\)\\ The echo band range versus the (a) solar zenith angle (SZA) and (b) model peak F2 electron density (\(N_mF_2\)).
+	\caption[Band position and occurrence versus SZA and \(N_mF_2\)]{{\:}Band position and occurrence versus SZA and \(N_mF_2\)\vspace{1mm}
+
+	The echo band range versus the (a) solar zenith angle (SZA) and (b) model peak F2 electron density (\(N_mF_2\)).
 Each point is color-coded in model \(E\) region density.
@@
 	\includegraphics[width=\textwidth]{p1f5.pdf}
-	\caption[Occurrence versus UT and IMF conditions]{{\:}Occurrence versus UT and IMF conditions\\ The MCM echo occurrence versus UT color coded in the average (a) IMF \(B_y\), (b) IMF \(B_z\), and (c) PCS values.}
+	\caption[Occurrence versus UT and IMF conditions]{{\:}Occurrence versus UT and IMF conditions\vspace{1mm}
+
+	The MCM echo occurrence versus UT color coded in the average (a) IMF \(B_y\), (b) IMF \(B_z\), and (c) PCS values.}
 	\label{fig:ut_occ}
@@
 This feature is less evident for IMF \(B_z\), which indicates that IMF \(B_y\) component has a stronger association with the F-region echo occurrence than IMF \(B_z\).
-In Figure \ref{fig:ut_occ}c, the average PCS index is somewhat enhanced before 12 UT, a result consisten with previous statistical analysis of the PCS index \citep[e.g.][Figure 4]{Troshichev2007}.
+In Figure \ref{fig:ut_occ}c, the average PCS index is somewhat enhanced before 12 UT, a result consistent with previous statistical analysis of the PCS index \citep[e.g.][Figure 4]{Troshichev2007}.
 There is no clear relationship between the echo occurrence and the PCS index except for the nighttime occurrence to be enhanced at large PCS values and the daytime occurrence to be either high or low at large PCS values.
@@
 MCM routinely observes backscatter from ionospheric irregularities over extended time periods and large spatial areas within the southern polar cap (``persistent backscatter''), which allowed us to reliably quantify both backscatter location and occurrence characteristics and investigate their control by the solar wind and illumination.
-\(B_y\) considering both location and occurrence we achieved better understanding and differentiation between radio- and plasma-physical factors or, alternatively, between irregularity detection and production.
+By considering both location and occurrence we achieved better understanding and differentiation between radio- and plasma-physical factors or, alternatively, between irregularity detection and production.
 Below we further discuss three groups of issues aligned with three objectives of this study.
@@
 	\includegraphics[width=\textwidth]{p1f6.pdf}
-	\caption[An example of conjugate observations of patchy backscatter]{{\:}An example of conjugate observations of patchy backscatter\\ Conjugate measurements for two selected events on (a--d) March 14, 2010 and (e-h) April 6, 2010.
+	\caption[An example of conjugate observations of patchy backscatter]{{\:}An example of conjugate observations of patchy backscatter\vspace{1mm}
+
+	Conjugate measurements for two selected events on (a--d) March 14, 2010 and (e-h) April 6, 2010.
 Shown are the IMF \(B_y\) (blue) and \(B_z\) (red) components (Figures \ref{fig:patchy_example}a and \ref{fig:patchy_example}e), MCM line-of-sight velocity versus UT and slant range in beam 14 (Figures \ref{fig:patchy_example}b and \ref{fig:patchy_example}f), RKN velocity in beam 4 (Figure \ref{fig:patchy_example}c and \ref{fig:patchy_example}g), and RISR-N electron density versus UT and altitude (Figure \ref{fig:patchy_example}d and \ref{fig:patchy_example}h).
@@
 Similar consistency between the patch drift and plasma convection velocities was observed by \citet{Bristow2011} using MCM observations and, more recently, by \citet{Makarevich2015b} using direct observations of 2D convection velocity and density enhancements by RISR-N in the northern polar cap.
-This impolies that polar patches tend to move with the background convection, at least deep within the polar cap.
+This implies that polar patches tend to move with the background convection, at least deep within the polar cap.
 
@@
 	\includegraphics[height=\textwidth,angle=90]{p1f7.pdf}
-	\caption[Summary of annual and diurnal variations]{{\:}Summary of annual and diurnal variations\\ Contour plots of the (a) average occurrence within the band, (b) band range, (c) model peak F-region electron density (\(N_mF_2\)) and (d) solar zenith angle (SZA) versus UT and month in 2010.
+	\caption[Summary of annual and diurnal variations]{{\:}Summary of annual and diurnal variations\vspace{1mm}
+
+	Contour plots of the (a) average occurrence within the band, (b) band range, (c) model peak F-region electron density (\(N_mF_2\)) and (d) solar zenith angle (SZA) versus UT and month in 2010.
 Color bars are given to the right.
@@
 
-One other mechanism through which IMF can exert influence over echo occurrence rates is through its indirect control of geomagnetic substorms.
+One other mechanism through which the IMF can exert influence over echo occurrence rates is through its indirect control of geomagnetic substorms.
 It is well known that enhanced particle precipitation and convection electric fields during substorms can be a significant source for plasma structures in the auroral region \citep[e.g.][]{Tsunoda1988}.
 At decameter scales, however, propagation conditions appear ot be a major limiting factor, with broad reductions in the echo occurrence rates observed in the nightside auroral zone \citep{Wild2008}.
-On the other hand, this is not the case in the polar cap, where no noticeable change in the cho occurrence is observed \citep[Figures 1 and 2]{Wild2008}.
+On the other hand, this is not the case in the polar cap, where no noticeable change in the echo occurrence is observed \citep[Figures 1 and 2]{Wild2008}.
 The absence of noticeable occurrence dependence on geomagnetic activity represented by PCS, Figure \ref{fig:ut_occ}c, is possibly related with that.
@@
 Fortran source code for IRI 2012 is freely available for download from the Community Coordinated Modeling Center (CCMC) at \url{http://ccmc.gsfc.nasa.gov/}.
-OMNI data is available from NASA Goddard Space Flight Center Space Physics Data Facility at \url{http://omniweb.gsfc.nasa.gov}.
+OMNI data is available from the NASA Goddard Space Flight Center Space Physics Data Facility at \url{http://omniweb.gsfc.nasa.gov}.
 The Resolute Bay incoherent scatter radar (North) is operated by SRI International on behalf of the US National Science Foundation under NSF Cooperative Agreement AGS-1133009.

### 2017-04-29 paper2.tex
FINAL VERSION POST GRAD SCHOOL CORRECTIONS

@@
 
-\section*{Abstract}
+\section{Abstract}
 Asymmetries in plasma density irregularity generation between the leading and trailing edges of the large-scale plasma density structures in the high-latitude ionosphere are investigated.
@@
 	\includegraphics[width=\textwidth,angle=180]{p2f1.pdf}
-	\caption[Overview of altitudinal variations and model setup of MCM]{{\:}Overview of altitudinal variations and model setup of MCM\\ (a) Altitudinal variations of parameters of interest.
+	\caption[Overview of altitudinal variations and model setup of MCM]{{\:}Overview of altitudinal variations and model setup of MCM\vspace{1mm}
+
+	(a) Altitudinal variations of parameters of interest.
 Shown are the ratios of collision frequencies and gyrofrequencies for electrons \(-r_e=\nu_e/|\Omega_e|\) (blue) and ions \(r_i=\nu_i/\Omega_i\) (green), the anisotropy parameter \(\psi=-r_i r_e\) (pink), and the conductivity ratio \(R=\sigma_H/\sigma_P = \left(r_i+r_e\right)/\left(1+\psi\right)\) (dashed black line on top of the green one).
@@
 	\includegraphics[width=\textwidth]{p2f2.pdf}
-	\caption[GDI growth rate within radar FoV]{{\:}GDI growth rate within radar FoV\\ The GDI growth rate within the radar's FoV for four cases of gradient reversal orientation and position (white line).
+	\caption[GDI growth rate within radar FoV]{{\:}GDI growth rate within radar FoV\vspace{1mm}
+
+	The GDI growth rate within the radar's FoV for four cases of gradient reversal orientation and position (white line).
 In all cases, the convection direction \(\vec{V}_E\) (red arrow) is perpendicular to the gradient reversal direction.
@@
 	\includegraphics[width=\textwidth,angle=180]{p2f3.pdf}
-	\caption[GDI growth rates for set angle between \(\vec{V}_E\) and \(\vec{G}\)]{{\:}GDI growth rates for set angle between \(\vec{V}_E\) and \(\vec{G}\)\\ The difference in the GDI growth rate across the gradient reversal for four selected angles between the reversal and convection direction.
+	\caption[GDI growth rates for set angle between \(\vec{V}_E\) and \(\vec{G}\)]{{\:}GDI growth rates for set angle between \(\vec{V}_E\) and \(\vec{G}\)\vspace{1mm}
+
+	The difference in the GDI growth rate across the gradient reversal for four selected angles between the reversal and convection direction.
 The gradient reversal direction \(\theta\) is on the \(x\)-axis and the average growth rate difference across the reversal \(\Delta\gamma\) is on the \(y\)-axis for each panel.
@@
 	\includegraphics[width=\textwidth,angle=180]{p2f4.pdf}
-	\caption[GDI growth rates for arbitrary angle between \(\vec{V}_E\) and \(\vec{G}\)]{{\:}GDI growth rates for arbitrary angle between \(\vec{V}_E\) and \(\vec{G}\)\\ Contour plots of the growth rate difference \(\Delta\gamma\)  versus the two angles of interest.
+	\caption[GDI growth rates for arbitrary angle between \(\vec{V}_E\) and \(\vec{G}\)]{{\:}GDI growth rates for arbitrary angle between \(\vec{V}_E\) and \(\vec{G}\)\vspace{1mm}
+
+	Contour plots of the growth rate difference \(\Delta\gamma\)  versus the two angles of interest.
 The direction of the gradient reversal \(\theta\) is on the \(x\)-axis and the convection direction \(\phi\) is on the \(y\)-axis.
@@
 	\includegraphics[width=\textwidth,angle=180]{p2f5.pdf}
-	\caption[GDI growth rate within transitional region]{{\:}GDI growth rate within transitional region\\ Analysis of the transitional region.
+	\caption[GDI growth rate within transitional region]{{\:}GDI growth rate within transitional region\vspace{1mm}
+
+	Analysis of the transitional region.
 (a) The average conductivity ratio \(R\) (black line) and the radar virtual height (orange line) versus the slant range for the entire transitional region and parts of the \(E\) and \(F\) regions.
@@
 In this study we used the SuperDARN radar at McMurdo Station, Antarctica to model spatial variations of the GDI growth rate within this radar's field-of-view and the resulting asymmetries near gradient reversals.
-Despite this, the results should be also applicable to any of the high-latitude SuperDARN radars in either the northern or southern hemisphere which observe large-scale density structures with gradient reversals.
+Despite this, the results should also be applicable to any of the high-latitude SuperDARN radars in either the northern or southern hemisphere which observe large-scale density structures with gradient reversals.
 To be more accurate, the parameters of the neutral density model would have to be changed to match a particular location, but these corrections would be very minor.

### 2017-04-29 paper3.tex
FINAL VERSION POST GRAD SCHOOL CORRECTIONS

@@
-\chapter[Radar Observations of Density Gradients, Electric Fields, and Plasma Irregularities Near Polar Cap Patches in the Context of the Gradient-Drift Instability]{Radar Observations of Density Gradients, Electric Fields, and Plasma Irregularities Near Polar Cap Patches in the Context of the Gradient-Drift Instability\footnote{Submitted as: {Lamarche}, L.~J., and R.~A.
+\chapter[Radar Observations of Density Gradients, Electric Fields, and Plasma Irregularities Near Polar Cap Patches in the Context of the Gradient-Drift Instability]{Radar Observations of Density Gradients, Electric Fields, and Plasma Irregularities Near Polar Cap Patches in the Context of the Gradient-Drift Instability\footnote{Published as: {Lamarche}, L.~J., and R.~A.
 {Makarevich} (2017), Radar observations of density gradients, electric fields, and plasma irregularities near polar cap patches in the context of the gradient-drift instability, \textit{J.
@@
 
-\section*{Abstract}
+\section{Abstract}
 We present observations of plasma density gradients, electric fields, and small-scale plasma irregularities near a polar cap patch made by the Super Dual Auroral Radar Network radar at Rankin Inlet (RKN) and the northern face of Resolute Bay Incoherent Scatter Radar (RISN-N).
@@
 	\includegraphics[width=\textwidth]{p3f1.pdf}
-  \caption[RKN/RISR-N experimental setup]{{\:}RKN/RISR-N experimental setup\\ Experimental setup showing the FoVs of RKN and RISR-N.
+  \caption[RKN/RISR-N experimental setup]{{\:}RKN/RISR-N experimental setup\vspace{1mm}
+
+  Experimental setup showing the FoVs of RKN and RISR-N.
 Lines of constant magnetic latitude are shown at \(75\deg\)N, \(80\deg\)N, and \(85\deg\)N.
@@
 	\includegraphics[width=\textwidth,angle=180]{p3f2.pdf}
-  \caption[Polar patch event overview]{{\:}Polar patch event overview\\ Left column is a keogram representation of (a and b) RISR-N electron density, (c) RKN line-of-sight velocity, and (d) RKN SNR for an event on November 30, 2011, 17:00--19:00 UT.
+  \caption[Polar patch event overview]{{\:}Polar patch event overview\vspace{1mm}
+
+  Left column is a keogram representation of (a and b) RISR-N electron density, (c) RKN line-of-sight velocity, and (d) RKN SNR for an event on November 30, 2011, 17:00--19:00 UT.
 Color scales for all quantities are given to the right of the respective row.
@@
 	\includegraphics[width=\textwidth,angle=180]{p3f3.pdf}
-  \caption[RKN/RISR-N velocity analysis]{{\:}RKN/RISR-N velocity analysis\\ RKN line-of-sight velocity versus the projection of 2D RISR-N velocity onto the closest RKN beam.
+  \caption[RKN/RISR-N velocity analysis]{{\:}RKN/RISR-N velocity analysis\vspace{1mm}
+
+  RKN line-of-sight velocity versus the projection of 2D RISR-N velocity onto the closest RKN beam.
 Black dots represent all coincident echoes that were observed by RKN and RISR-N simultaneously over the day considered, whereas red dots represent only echoes that occurred within the high-density structure.
@@
 \includegraphics[width=\textwidth,angle=180]{p3f4.pdf}
-  \caption[Irregularity dependence on density factors]{{\:}Irregularity dependence on density factors\\ Scatter plots of RKN echo power versus RISR-N measurements of (a) electron density, \(n\), (b) density gradient magnitude, \(|\nabla n|\), and (c) gradient strength, \(G=|\nabla n/n|\).
+  \caption[Irregularity dependence on density factors]{{\:}Irregularity dependence on density factors\vspace{1mm}
+
+  Scatter plots of RKN echo power versus RISR-N measurements of (a) electron density, \(n\), (b) density gradient magnitude, \(|\nabla n|\), and (c) gradient strength, \(G=|\nabla n/n|\).
 Black dots represent all echoes observed over the day, whereas red dots represent echoes that occur within the high-density structure.
@@
 	\includegraphics[width=\textwidth,angle=180]{p3f5.pdf}
-  \caption[Irregularity dependence on directional factors]{{\:}Irregularity dependence on directional factors\\ Scatter plots of RKN echo power versus RISR-N measurements of (a) the gradient strength component, \(\uvec{k}\cdot\uvec{b}\times\vec{G}\), (b) the electric field component \(\uvec{k}\cdot\vec{E}\), (c) the normalized gradient component, \(\uvec{k}\cdot\uvec{b}\times\uvec{g}\), and (d) the normalized electric field component, \(\uvec{k}\cdot\uvec{e}\).
+  \caption[Irregularity dependence on directional factors]{{\:}Irregularity dependence on directional factors\vspace{1mm}
+
+  Scatter plots of RKN echo power versus RISR-N measurements of (a) the gradient strength component, \(\uvec{k}\cdot\uvec{b}\times\vec{G}\), (b) the electric field component \(\uvec{k}\cdot\vec{E}\), (c) the normalized gradient component, \(\uvec{k}\cdot\uvec{b}\times\uvec{g}\), and (d) the normalized electric field component, \(\uvec{k}\cdot\uvec{e}\).
 The meaning of dots, crosses, vertical bars, and lines is the same as in Figure \ref{fig:density}.}
@@
 	\includegraphics[width=\textwidth,angle=180]{p3f6.pdf}
-  \caption[Irregularity dependence on GDI growth rate]{{\:}Irregularity dependence on GDI growth rate\\ Same as Figure \ref{fig:directional}, but for (a) the dot product of the gradient and plasma drift velocity, \(\vec{G}\cdot\vec{V}_E\), (b) the calculated GDI growth rate, \(\gamma\), (c) the normalized dot product of the  gradient and plasma drift velocity, \(\uvec{g}\cdot\mathbf{\hat{v}}_E\), and (d) the normalized GDI growth rate, \(\gamma'\).}
+  \caption[Irregularity dependence on GDI growth rate]{{\:}Irregularity dependence on GDI growth rate\vspace{1mm}
+
+  Same as Figure \ref{fig:directional}, but for (a) the dot product of the gradient and plasma drift velocity, \(\vec{G}\cdot\vec{V}_E\), (b) the calculated GDI growth rate, \(\gamma\), (c) the normalized dot product of the  gradient and plasma drift velocity, \(\uvec{g}\cdot\mathbf{\hat{v}}_E\), and (d) the normalized GDI growth rate, \(\gamma'\).}
   \label{fig:gamma}
@@
 \includegraphics[width=\textwidth,angle=180]{p3f7.pdf}
-  \caption[Relationship between the electric field and gradient components]{{\:}Relationship between the electric field and gradient components\\ Scatter plots of RISR-N measurements of the electric field component, \(\uvec{k}\cdot\vec{E}\), versus the gradient component, \(\uvec{k}\cdot\uvec{b}\times\vec{G}\).
+  \caption[Relationship between the electric field and gradient components]{{\:}Relationship between the electric field and gradient components\vspace{1mm}
+
+  Scatter plots of RISR-N measurements of the electric field component, \(\uvec{k}\cdot\vec{E}\), versus the gradient component, \(\uvec{k}\cdot\uvec{b}\times\vec{G}\).
 Points are color coded in RKN SNR, according to the color bar to the right.
@@
 \centering
+\caption[Results consistent with linear GDI theory]{{\:}Results consistent with linear GDI theory\vspace{1mm}
+
+A summary of the experimentally observed features that are consistent with the gradient-drift instability theory for arbitrary directions of the wavevector \(\uvec{k}\), electric field \(\uvec{e}\), and density gradient \(\uvec{g}\).}
+\vspace{5pt}
 \begin{tabular}{rlc}
@@
 \end{tabular}
-\caption[Results consistent with linear GDI theory]{{\:}Results consistent with linear GDI theory\\ A summary of the experimentally observed features that are consistent with the gradient-drift instability theory for arbitrary directions of the wavevector \(\uvec{k}\), electric field \(\uvec{e}\), and density gradient \(\uvec{g}\).}
 \label{tab:results}
@@
 However, because the patch does move almost directly northwards through the FoV, predominantly the leading and trailing edges are seen by the radar at times.
-This means that the leading and trailing edges may be oversampled, which could explain the sharp peaks in occurrence when \(\vec{G}\) parallel or antiparallel to \(\vec{V}_E\) relative to other directions.
+This means that the leading and trailing edges may be oversampled, which could explain the sharp peaks in occurrence when \(\vec{G}\) is parallel or antiparallel to \(\vec{V}_E\) relative to other directions.
 

### 2017-04-29 thesis.sh

@@ -25,6 +25,8 @@ pdflatex LLamarcheThesis.tex
 
 pdflatex LLamarcheThesis.tex
 
+pdflatex LLamarcheThesis.tex
+
 rm *.aux
 rm *.bbl
 rm *.blg
